### 2025-04-23 appendix.tex
Initial Overleaf Import

% appendix.tex
\appendix

\section{Appendix A}

\label{appendix:gradient_fields}
\subsection{Interpreting The Gradient Fields}
\label{sec:interpret_gradient_fields}
The gradient descent field contains information about the loss landscape and the update path from the initial point. We chose 2 parameters for each program, corresponding to the 2 dimensions in the gradient fields. For \BPNoise, the x-axis is the HP cutoff and the y axis is the LP cut off. The values for the parameters are normalized from -1 to 1. In \BPNoise{} for example, the possible values for LP cut are defined in the \texttt{hslider} function in the range of 50 Hz-1000 Hz. The initial point is 100 Hz, which is normalized to around -0.95, as shown in the graph. The initial HP cutoff is set at 20, which normalizes to around -0.75 in the range of 1 Hz - 120 Hz.

The initial and target parameters are shown with green circle \greencircle~and star \greenstar~respectively. Each circle in the connected red circles \redcircles~ corresponds to an update to the initial values. The ideal successful experiment would entail a series of red dots connecting the initial values to the target values, meaning that each update step took the initial parameter values closer to the target. The background of the gradient field is a heatmap based on the value of the loss. The highest loss corresponds to a darker color, and the lowest loss---which should be around the target value---should be the white.

The arrow in the middle of each heatmap square corresponds to the direction and magnitude of the gradients. The gradients here can be extremely large; signal processing functions can be unpredictable, and the gradients are calculated backwards sequentially through the signal processing chain, resembling similar calculations---associated with exploding gradients---in recurrent neural networks~\cite{gers2000learning}. For this reason gradient clipping is applied~\cite{goodfellow2016deep}.

\subsection{\BPNoise}
In \BPNoise, a noise signal is fed through a band-pass (BP) filter, which removes the frequencies between the low-pass (LP) and high-pass (HP) cutoff. As shown in Listing~\ref{lst:program0}, the target (or \texttt{true\_params}) is the HP of 100, and LP of 900, and the initial parameters are 20 and 100 respectively.

Figure~\ref{fig:p0_losses} depicts the learning trajectory for all loss functions. Changes in filter cut-offs is mainly reflected in the frequency content of the sound.  All loss functions except for DTW-Onset are successful in finding the target. These results make intuitive sense, as the DTW-Onset loss primarily works by matching amplitude envelopes, yet the other loss functions primarily measure changes in frequency content.

The heatmaps of L1-Spec, SIMSE-Spec, and JTFS also show low loss values closer to the target, unlike the DTW-Onset function.

\subsection{\AddSineSaw}
\AddSineSaw{} is an additive program that combines a saw and a sine wave function (each with a different frequency) to create the output. The initial values for the frequencies being 800 Hz and 300 Hz, and the target of 200 Hz and 100 Hz.

As show in Figure~\ref{sec:program1}, the Spectrogram losses quickly fail this test. The loss heatmaps here look reasonable, yet most gradients point to the wrong direction, and the iterative update steps quickly fall outside the parameter range as defined by the code, which makes recovery impossible. The DTW-Onset loss also fails to find the target, but gets close to the correct Saw frequency. What is notable here is that for the 3 failed loss functions, their gradient field indicates that the Sine Frequency (the vertical y-axis features) is not considered an important parameter, as most of the gradients do not have a strong vertical direction. We can speculate that this is due to the strong harmonic content of the saw wave relative to the sine waves.

A notable characteristic of the heatmaps is the diagonal line where the sine and saw frequencies are equal. All points on this line are outliers relative to their neighbouring points. This is because of the harmonic overlap where the sine frequency falls directly on top of the fundamental frequency of the saw wave.

\subsection{\AmpMod}
Matching the dynamic amplitude change in the sound caused by the Sine oscillator is the main challange of this program, and we see that DTW-Onset loss is the only function successful at this task. JTFS and L1-Spectrogram loss find a value close to the true \textit{amp}, yet do not move towards the correct \textit{carrier} value. The DTW-Onset heatmap also shows reasonable gradients and loss-values corresponding to distance from the target.

\subsection{\FMMod}
All loss functions except DTW-Onset perform poorly and fall outside the acceptable range within a few updates. DTW-Onset appears as the only promising function here. This loss crosses the correct amplitude modulation frequency, but fails to converge to the correct carrier frequency.

In the DTW-Onset heatmap we see little change in loss values in the vertical plane corresponding to the carrier, however, the amplitude dimension has the minimal loss value around the target parameter. This again shows that the DTW-Onset loss excels at matching sound envelopes (or changes in sound volume over time) but is insensitive to frequency content. 
\begin{figure*}[htbp]
    \centering
    % Top row
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \adjustbox{valign=t}{\includegraphics[width=\textwidth]{images/experiment_plots/p0_L1_Spec.png}}
        \caption{L1 spectrogram loss}
        \label{fig:p0_spec}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \adjustbox{valign=t}{\includegraphics[width=\textwidth]{images/experiment_plots/p0_SIMSE_Spec.png}}
        \caption{SIMSE spectrogram loss}
        \label{fig:p0_simse}
    \end{subfigure}
    \vspace{0.5cm} 
    % Bottom row
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \adjustbox{valign=t}{\includegraphics[width=\textwidth]{images/experiment_plots/p0_DTW_Onset.png}}
        \caption{DTW onset loss.}
        \label{fig:p0_dtw}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \adjustbox{valign=t}{\includegraphics[width=\textwidth]{images/experiment_plots/p0_JTFS.png}}
        \caption{JTFS loss}
        \label{fig:p0_jtfs}
    \end{subfigure}

    \caption{Gradient descent fields for each loss with respect to \BPNoise. The initial and target parameters are shown with green circle \greencircle~and star \greenstar~respectively. High loss corresponds to dark-blue and low corresponds to white.}
    \label{fig:p0_losses}
\end{figure*}

\begin{figure*}[htbp]
    \centering
    % Top row
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \adjustbox{valign=t}{\includegraphics[width=\textwidth]{images/experiment_plots/p1_L1_Spec.png}}
        \caption{L1 spectrogram loss}
        \label{fig:p1_spec}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \adjustbox{valign=t}{\includegraphics[width=\textwidth]{images/experiment_plots/p1_SIMSE_Spec.png}}
        \caption{SIMSE spectrogram loss}
        \label{fig:p1_simse}
    \end{subfigure}
    \vspace{0.5cm} 
    % Bottom row
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \adjustbox{valign=t}{\includegraphics[width=\textwidth]{images/experiment_plots/p1_DTW_Onset.png}}
        \caption{DTW onset loss.}
        \label{fig:p1_dtw}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \adjustbox{valign=t}{\includegraphics[width=\textwidth]{images/experiment_plots/p1_JTFS.png}}
        \caption{JTFS loss}
        \label{fig:p1_jtfs}
    \end{subfigure}

    \caption{Gradient descent fields for each loss with respect to \AddSineSaw{}. The initial and target parameters are shown with green circle \greencircle~and star \greenstar~respectively. High loss corresponds to dark-blue and low corresponds to white.}
    \label{fig:p1_losses}
\end{figure*}

\begin{figure*}[htbp]
    \centering
    % Top row
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \adjustbox{valign=t}{\includegraphics[width=\textwidth]{images/experiment_plots/p2_L1_Spec.png}}
        \caption{L1 spectrogram loss}
        \label{fig:p2_spec}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \adjustbox{valign=t}{\includegraphics[width=\textwidth]{images/experiment_plots/p2_SIMSE_Spec.png}}
        \caption{SIMSE spectrogram loss}
        \label{fig:p2_simse}
    \end{subfigure}
    \vspace{0.5cm} 
    % Bottom row
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \adjustbox{valign=t}{\includegraphics[width=\textwidth]{images/experiment_plots/p2_DTW_Onset.png}}
        \caption{DTW onset loss.}
        \label{fig:p2_dtw}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \adjustbox{valign=t}{\includegraphics[width=\textwidth]{images/experiment_plots/p2_JTFS.png}}
        \caption{JTFS loss}
        \label{fig:p2_jtfs}
    \end{subfigure}

    \caption{Gradient descent fields for each loss with respect to \AmpMod. The initial and target parameters are shown with green circle \greencircle~and star \greenstar~respectively. High loss corresponds to dark-blue and low corresponds to white.}
    \label{fig:p2_losses}
\end{figure*}

\begin{figure*}[htbp]
    \centering
    % Top row
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \adjustbox{valign=t}{\includegraphics[width=\textwidth]{images/experiment_plots/p3_L1_Spec.png}}
        \caption{L1 spectrogram loss}
        \label{fig:p3_spec}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \adjustbox{valign=t}{\includegraphics[width=\textwidth]{images/experiment_plots/p3_SIMSE_Spec.png}}
        \caption{SIMSE spectrogram loss}
        \label{fig:p3_simse}
    \end{subfigure}
    \vspace{0.5cm} 
    % Bottom row
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \adjustbox{valign=t}{\includegraphics[width=\textwidth]{images/experiment_plots/p3_DTW_Onset.png}}
        \caption{DTW onset loss.}
        \label{fig:p3_dtw}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \centering
        \adjustbox{valign=t}{\includegraphics[width=\textwidth]{images/experiment_plots/p3_JTFS.png}}
        \caption{JTFS loss}
        \label{fig:p3_jtfs}
    \end{subfigure}

    \caption{Gradient descent fields for each loss with respect to \FMMod. The initial and target parameters are shown with green circle \greencircle~and star \greenstar~respectively. High loss corresponds to dark-blue and low corresponds to white.}
    \label{fig:p3_losses}
\end{figure*}

% \section{Appendix B}
% \label{appendix:parameters}
% List of arbitrary parameters used in this work.\\
% % name,  value
% % learning rate, 0.45
% % iterations, 200
% % number of experiments, 200
% % bootstrapping k, 1000
% % bootstrapping n, 50\% or 100
% % sample length, 1 sec
% % sample rate, 44100 (cite)
% \begin{verbatim}
% | Name                  | Value           |
% |-----------------------|-----------------|
% | Learning Rate         | 0.45            |
% | Iterations            | 200             |
% | Number of Experiments | 200             |
% | Bootstrapping k       | 1000            |
% | Bootstrapping n       | 50% or 100      |
% | Sample Length         | 1 sec           |
% | Sample Rate           | 44100 (cite)    |
% | DTW gamma             | 1               |
% \end{verbatim}

### 2025-04-23 extra_figures/evaluation_graph.tex
Initial Overleaf Import

\begin{figure}
\resizebox{\textwidth}{!}{ % Automatically scales diagram to fit page width
\begin{tikzpicture}[
    function/.style={rectangle, draw, rounded corners, minimum width=2cm, minimum height=1cm, text centered, node distance=1.5cm},
    distribution/.style={ellipse, draw, minimum width=2cm, minimum height=1cm, text centered, node distance=1.2cm},
    gaussian/.style={draw, smooth, samples=100, domain=-2.5:2.5},
    snpk/.style={rectangle, draw, rounded corners, minimum width=2.5cm, minimum height=1cm, text centered, node distance=1.5cm, fill=blue!20},
    level 1/.style={sibling distance=6cm, level distance=2cm},
    level 2/.style={sibling distance=2cm, level distance=3cm},
    mss/.style={fill=red!50, draw=black},
    ploss/.style={fill=blue!50, draw=black},
    likert/.style={fill=green!40, draw=black}
]

% Existing functions and distributions diagram
\node[function] {Synthesizer Program}
    child { node[function] {DTW\_Onset}
        child { node[distribution, mss] {MSS}
            child[grow=down, level distance=1.5cm] {
                node[draw=none] {
                    \begin{tikzpicture}[scale=0.7]
                        \draw[red] plot (\x/3, {exp(-\x*\x)});  % Red Gaussian
                    \end{tikzpicture}
                }
            }
        }
        child { node[distribution, ploss] {P-Loss}
            child[grow=down, level distance=1.5cm] {
                node[draw=none] {
                    \begin{tikzpicture}[scale=0.7]
                        \draw[blue] plot (\x/3, {exp(-\x*\x)});  % Blue Gaussian
                    \end{tikzpicture}
                }
            }
        }
        child { node[distribution, likert] {Likert}
            child[grow=down, level distance=1.5cm] {
                node[draw=none] {
                    \begin{tikzpicture}[scale=0.7]
                        \draw[green] plot (\x/3, {exp(-\x*\x)});  % Green Gaussian
                    \end{tikzpicture}
                }
            }
        }
    }
    child { node[function] {L1\_Spec}
        child { node[distribution, mss] {MSS} }
        child { node[distribution, ploss] {P-Loss} }
        child { node[distribution, likert] {Likert} }
    }
    child { node[function] {SIMSE\_Spec}
        child { node[distribution, mss] {MSS} }
        child { node[distribution, ploss] {P-Loss} }
        child { node[distribution, likert] {Likert} }
    }
    child { node[function] {JTFS}
        child { node[distribution, mss] {MSS}
            child[grow=down, level distance=1.5cm] {
                node[draw=none] {
                    \begin{tikzpicture}[scale=0.7]
                        \draw[red] plot (\x/3, {exp(-\x*\x)});  % Red Gaussian
                    \end{tikzpicture}
                }
            }
        }
        child { node[distribution, ploss] {P-Loss}
            child[grow=down, level distance=1.5cm] {
                node[draw=none] {
                    \begin{tikzpicture}[scale=0.7]
                        \draw[blue] plot (\x/3, {exp(-\x*\x)});  % Blue Gaussian
                    \end{tikzpicture}
                }
            }
        }
        child { node[distribution, likert] {Likert}
            child[grow=down, level distance=1.5cm] {
                node[draw=none] {
                    \begin{tikzpicture}[scale=0.7]
                        \draw[green] plot (\x/3, {exp(-\x*\x)});  % Green Gaussian
                    \end{tikzpicture}
                }
            }
        }
    };

% Label for the second row (Distributions)
\node at (0,-6.5) {\textbf{$\longleftarrow$\ \ \large Bootstrapped Distributions\ \ $\longrightarrow$}};

% Left side: SNPK with MSS
\node[snpk, text height=5ex, align=center, fill=red!30, draw=black] at (-6,-9) {SNPK with 4 \\ MSS distributions};
\draw[->] (-6,-9.5) -- (-6,-10.5) node[below, below] {\textbf{Ranks 1-4}};
\draw[red!40, thick] (-4.5, -8.4) -- (-2.75, -7.75);  % Line 1
\draw[red!40, thick] (-5.25, -8.4) -- (-3.5, -7.75);  % Line 2
\draw[red!40, thick] (-5.75, -8.4) -- (-5.5, -7.75);  % Line 3
\draw[red!40, thick] (-6.25, -8.4) -- (-9, -7.75);  % Line 4

% Middle: SNPK with P-Loss
\node[snpk, text height=5ex, align=center, fill=blue!40, draw=black] at (0,-9) {SNPK with 4 \\ P-Loss distributions};
\draw[->] (0,-9.5) -- (0,-10.5) node[below, below] {\textbf{Ranks 1-4}};
\draw[blue!40, thick] (1.5, -8.4) -- (3.25, -7.75);  % Line 1
\draw[blue!40, thick] (0.75, -8.4) -- (1.5, -7.75);  % Line 2
\draw[blue!40, thick] (-0.25, -8.4) -- (-1, -7.75);  % Line 3
\draw[blue!40, thick] (-0.75, -8.4) -- (-3, -7.75);  % Line 4

% Right side: SNPK with Likert
\node[snpk, text height=5ex, align=center, fill=green!30, draw=black] at (6,-9) {SNPK with 4 \\ Likert distributions};
\draw[->] (6,-9.5) -- (6,-10.5) node[below, below] {\textbf{Ranks 1-4}};
\draw[green!50, thick] (7.5, -8.4) -- (8.75, -7.75);  % Line 1
\draw[green!50, thick] (6.75, -8.4) -- (6, -7.75);    % Line 2
\draw[green!50, thick] (6.25, -8.4) -- (4, -7.75);   % Line 3
\draw[green!50, thick] (5.75, -8.4) -- (3, -7.75);     % Line 4

\end{tikzpicture}
}
\caption{For each synthesizer program, we assign ranks to the four loss functions  (\DTWEnv, \LoneSpec, \SIMSESpec, \JTFS) in three different ways, using MSS, P-Loss, or manually assigned Likert scores.}
\label{fig:posthoc_evaluation}
\end{figure}

### 2025-04-23 glossary.tex
Initial Overleaf Import

% glossary.tex
\newglossaryentry{FFT}{
    name={FFT},
    description={Fast Fourier Transforms}
}
\newglossaryentry{STFT}{
    name={STFT},
    description={Short Time Fourier Transform, a Fourier-based technique used in audio signal processing}
}

\newglossaryentry{DSP}{
    name={DSP},
    description={Digital Signal Processing}
}

\newglossaryentry{MSS}{
name = {MSS},
description={Multi-Scale Spectrogram}
}

\newglossaryentry{HC}{
name = {HC},
description={Hill-Climber algorithm}
}
\newglossaryentry{JTFS}{
name = {JTFS},
description={Joint-time Frequency Scattering}
}

\newglossaryentry{SIMSE}{
name = {SIMSE},
description = {Structured Similarity Metric}
}

\newglossaryentry{DTW}{
name = {DTW},
description={Dynamic Time Warping}
}

\newglossaryentry{DDSP}{
    name={DDSP},
    description={Differentiable Digital Signal Processing, overlap of methods from signal processing with machine learning}
}

\newglossaryentry{FM}{
    name={FM},
    description={A method of sound synthesis where the frequency of one oscillator modulates the frequency of another}
}

\newglossaryentry{Gradient descent}{
    name={Gradient descent},
    description={An optimization algorithm used to minimize a function by iteratively moving towards the lowest point}
}

\newglossaryentry{GA}{
name = {GA},
description={genetic algorithm}
}

% \newglossary{HC}{
% name = {HC},
% description={Hill climber}

% }

\newglossaryentry{NN}{
name = {NN},
description = {Neural network}
}

\newglossaryentry{VST}{
name = {VST},
description = {Virtual Studio Technology. Virtual software synthesizers which can run standalone or in a DAW}
}

\newglossaryentry{DAW}{
name = {DAW},
description = {Digital Audio Workstation}
}

\newglossaryentry{CI}{
name = {CI},
description = {Confidence interval}
}

\newglossaryentry{LFO}{
name = {LFO},
description = {low frequency oscillator}
}

\newglossaryentry{MFCC}{
name = {MFCC},
description = {Mel-frequency cepstral coefficients}
}

\newglossaryentry{AM}{
name = {AM},
description = {Amplitude modulation synthesis. A method of sound synthesis where one oscillator modulates the amplitude of another
}
}

\newglossaryentry{ADSR}{
    name = {ADSR},
    description= {Attack, delay, sustain, release. An envelope that shapes a sound. }
}

\newglossaryentry{OSR}{
name = {OSR},
description = {Open set recognition}
}

\newglossaryentry{RL}{
name = {RL},
description = {Reinforcement learning}
}

\newglossaryentry{SGD}{
name = {SGD},
description = {Stochastic gradient descent}
}

\newglossaryentry{DL}{
name = {DL},
description = {Deep learning}
}

\newglossaryentry{SOTA}{
name = {SOTA},
description = {State of the art}
}

\newglossaryentry{ML}{
name = {ML},
description = {Machine learning}
}

\newglossaryentry{SNPK}{
name = {SNPK},
description = {Non-Parametric Scott-Knott}
}

### 2025-04-23 main.tex
Initial Overleaf Import

\documentclass[lettersize,journal]{IEEEtran}% \documentclass[twocolumn,10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{stfloats} % for positioning of figure* on the same page
\usepackage{caption}
\usepackage{tikz}
\usepackage[inline]{enumitem}
\usepackage{amsmath}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage[breaklinks=true,colorlinks=true, allcolors=blue]{hyperref}
\usepackage{breakcites}
\usepackage{microtype}
\usepackage{lipsum}
\usepackage{xcolor}
\usepackage{array}
\usepackage{float}
\usepackage{adjustbox}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{csquotes}
\usepackage{makecell}
\usepackage{pdfpages}

\usetikzlibrary{positioning,shapes.geometric}
\usepackage[acronym,toc]{glossaries}

% \input{glossary.tex}

% Initialize the glossary
\makeglossaries

\definecolor{highlight}{RGB}{00,150,00}
\definecolor{todo}{RGB}{200,50,00}
\definecolor{rank1}{HTML}{70FF70}
\definecolor{rank2}{HTML}{858585}
\definecolor{rank3}{HTML}{454545}
\definecolor{rank4}{HTML}{000000}

\newcommand{\staronsquare}{
    \tikz[baseline=0ex]{
        % Blue square with increased size
        \fill[blue] (0,0) rectangle (2ex,2ex);
        % Yellow star character centered on the square, larger font size
        \node at (1ex,0.8ex) {\textcolor{yellow}{\large *}};
    }
}

\newcommand{\vertdashred}[1][0.3cm]{% Default height is 2cm
    \tikz[baseline]{\draw[red, dashed,line width=0.5mm] (0,0) -- (0,#1);}
}

\newcommand{\SIMSESpec}{\texttt{SIMSE\_Spec}}
\newcommand{\LoneSpec}{\texttt{L1\_Spec}}
\newcommand{\JTFS}{\texttt{JTFS}}
\newcommand{\DTWEnv}{\texttt{DTW\_Envelope}}

\newcommand{\LossSelect}{\textbf{Loss Selection}}
\newcommand{\SynthSelect}{\textbf{Synthesis Selection}}
\newcommand{\PeriodicLoss}{\textbf{Periodic Loss}}
\newcommand{\OutDomain}{\textbf{Out-Domain Generation}}

\newcommand{\BPNoise}{\textbf{BP-Noise}}  
\newcommand{\AddSineSaw}{\textbf{Additive-SineSaw}}  
\newcommand{\AmpMod}{\textbf{Noise-AM}}  
\newcommand{\FMMod}{\textbf{SineSaw-AM}}

\input{code_style}

\title{Sound Similarity, Sound Matching, and Differentiable Audio Synthesizer Control}
\author{ }

\begin{document}

\printglossary
\clearpage

\maketitle
\begin{abstract}
Manual sound design with a synthesizer is an iterative process. When an artist has a target sound in mind, they listen to the output of their chosen synthesizer, determine whether the output is closer or farther from the target, and adjust the synthesizer's parameters until satisfied. There are practically an infinite number of parameter combinations, and their effect on the output sound is hard to predict. Iterative sound-matching is the automation of this process, more formally, the automatic programming of an audio synthesizer to generate outputs similar to a target sound while guided by a loss function that quantifies sound similarity.
Past works in sound-matching often relied on a narrow range of synthesizers and overlooked the possible interactions between the loss function and synthesis parameters.
% There are critical choices which need to be made in sound-matching: these are the choice of synthesizer, algorithms for search and optimization, the source of sounds, and similarity (or loss) functions.  We categorize sound-matching works based on these choices, and point out a number of unresolved issues and interesting yet unexplored areas. 
We analyzed the interplay between four simple subtractive, additive, and FM synthesizers when guided by four different novel or established similarity functions.
For each loss function and synthesizer pair, we ran 300 experiments with randomized initial values and targets, and used the final parameter configurations and spectrogram distances as key performance evaluation metrics of the performance. A post-hoc analysis of these performance measures showed that the best‐performing loss function varies based on the method of synthesis at hand, which highlights the value of novel similarity measures and de-prioritizes the search for the elusive universally optimal sound similarity measure.
\end{abstract}

%  Like other problems in computing science, there appears to be an ever present drive towards finding optimal solutions to the problem, and outperforming state-of-the-art.

\section{Introduction}
\subsection{Digital Synthesis With Sound-Matching}
Since the 1970s, artists such as Suzanne Ciani~\cite{ciani_life_in_waves} have explored the use of audio synthesizers to replicate natural sounds---such as a bottle cap popping off---in a way that preserves the essential qualities of the original while showcasing the characteristics of the synthesizer. This approach goes beyond mere replication and uses synthesizers for artistic re-interpretation of sounds. Inspired by this, we aim to emulate and automate this process using digital audio synthesizers.

A digital audio synthesizer is any software used for the creation and manipulation of audio. Since the advent of digital signal processing (\gls{DSP}) in the 1960s~\cite{stranneby2004digital}, a wide variety of parametric audio generation functions have been proposed~\cite{lyons1997understanding,russ1999sound,shier2020spiegelib}. Manual sound design with a synthesizer is an iterative process of trial and error; when the artist has a desired (or target) sound in mind, sound design involves listening to the output, determining whether the output is closer or farther from the target, and adjusting the parameters accordingly until the target is reached~\cite{russ1999sound}. This paper explores the automation of this process, which we call \textit{iterative sound-matching}.

Automatic sound design has been a subject of research for decades, with reduction of tinkering time and creation of ``interesting" sounds as the main motivators \cite{krekovic2019insights,turian2020sorry,horner1993machine,salimi2020make,esling2019flow,engel2020ddsp,mitchell2007evolutionary,shier2020spiegelib}. \textbf{Sound-matching} is a common form of automatic sound design explored in literature, where the goal is to find parameters for a synthesizer that replicate \textit{all} or \textit{some} of the characteristics of a target sound as best as possible, given some kind of similarity metric~\cite{horner1993machine,mitchell2007evolutionary}. Not all sound-matching works follow the iterative process of manual sound design. \textit{Iterative} sound-matching refers to works which mimic the back and forth process of listening to a synthesizer's output and updating the parameters until the termination conditions are met.

A brief description of the iterative sound-matching workflow may be useful here: The two necessary components for sound-matching are the \textit{loss function}, which determines how close two sounds are, and the \textit{parametric synthesizer} that outputs sounds based on input parameters. In the iterative workflow, a target sound is selected, the synthesizer outputs sounds based on its parameters, and the loss function determines the proximity of the output to the target. From there, the synthesizer parameters are updated, the output and target are compared, and so on.

With the goal of implementing effective iterative sound-matching algorithms for sound designers, we identify a number of weaknesses in current relevant literature (see Section~\ref{sec:lacking}). The experiments in this work pertain to two core issues, which we named \LossSelect~and \SynthSelect. The former refers to the search for an optimal loss function, lack of novel algorithms, and the focus on outperforming state of the art (\gls{SOTA}), while the latter refers to the lack of diversity in synthesis methods, and their interaction with the loss function.

In the case of \LossSelect, there have been many works and experiments comparing the accuracy of loss functions and choosing the best performers~\cite{vahidi2023mesostructures,turian2020sorry,engel2020ddsp,uzrad2024diffmoog,han2023perceptual,masuda2021soundmatch,turian2020sorry,bruford2024synthesizer}.  We note that these claims regarding the effectiveness of one function versus another have consistently been made in limited contexts that may not generalize to other settings. One such limitation is the synthesizers used, and ties into the problem of \SynthSelect, as typically only a single (and often, simple) synthesizer has been used for testing the effectiveness of loss functions. Furthermore, there is a lack of diversity in the underlying functions used by the synthesizers, with simple Frequency Modulation (\gls{FM}) being the most common method of synthesis~\cite{horner1993machine,vahidi2023mesostructures,mitchell2007evolutionary,caspe2023envelope}.

Considering the interplay between the two problems, we design experiments that address them simultaneously. We hypothesize that the performance of a similarity measure is influenced by other factors in the environment, particularly the method of synthesis. We combine 4 loss functions with 4 differentiable synthesis methods, and compare their performance (as measured by the similarity of the final output or parameters to the target or target parameters). Our experiments show that when accounting for different synthesizers, there is no consistent choice of sound similarity measure for optimal performance in sound-matching. In some experiments, \DTWEnv, a novel loss function that only looks at envelope similarity, matches or outperforms commonly used methods of audio comparison. These results imply that rather than aiming to outperform the SOTA and finding a universal best performer, more effort should be placed towards diversifying the existing similarity measures of sound as well as conducting experiments using a variety of synthesis methods. Sound design is an inherently creative endeavor, and would likely benefit from a diversified range of solutions for different approaches.

\subsection{Differentiable Sound-Matching}
An important caveat is that this work takes a \textit{differentiable}, machine learning (\gls{ML}) inspired approach to iterative sound-matching. A differentiable approach requires differentiable loss and synthesizer functions, allowing the synthesizer parameters to be updated using gradient descent~\cite{goodfellow2016deep,boyd2004convex}. Ideally, these updates are in a direction that improves audio similarity by reducing loss. Differentiable iterative sound-matching has only recently become viable thanks to continued advances in both computation hardware~\cite{moore2012introduction,schaller1997moore}, and machine learning software~\cite{deepmind2020jax,paszke2019pytorch}.

 As we discuss in Section~\ref{sec:related_works}, nearly all past works pertaining to iterative sound-matching have used genetic algorithms (\gls{GA})~\cite{holland1992genetic} where parameters are randomly changed with hopes of increasing similarity. Differentiable settings have the benefit of making more directed updates to the synthesizer parameters, but have major drawbacks. Other than requiring more computation power, differentiable functions require careful implementation of operations that are continuous, numerically stable, and of course, differentiable.

\section{Background}
In this Section, we give a formal definition of sound-matching and the necessary components. The \textit{type} of sound-matching problem is dependent on the implementation of these components and the source of target sounds. We give a summary of different possible implementations and goals for sound-matching, and provide brief examples of previous works where applicable. We also give a typology of sound-matching types and major perennial issues that need to be addressed in future works. Finally, we introduce our formal definition of \textit{iterative sound-matching} in Section~\ref{sec:iterative_sound_matching}.

% Due to their pivotal role in our experiments, a discussion of similarity measures, or \textit{loss functions} is provided in Section~\ref{sec:loss_funcs}. A formal definition of loss functions and specific implementations of four loss functions are given.

\label{sec:background}

\subsection{Digital Audio Synthesis}
\label{sec:diffSynth}
A digital audio synthesizer is any software used for the creation and manipulation of digital audio. Digital synthesizers use signal processing chains with a variety of digital signal processing (\gls{DSP}) functions to create sounds. These functions are often parametric, and the set of parameters given to the synthesizer's chain of DSP functions is called a synthesizer program.

The simplest form of DSP function is a sinusoidal tone. For example, consider:
\[ x(n) = sin( 2\pi n T)\]

where $T$ represents time in seconds, and $n$ is the frequency parameter. If $n$ is 1, then $x(n)$ would be a waveform with a frequency of 1 hertz as the value of $x(n)$ would oscillate once per second. Waveform generators are called oscillators for this reason. If $n$ is in the range of 30 to 14,000 (the upper limit can be higher for some), $x(n)$ would be an audible tone~\cite{smith2007mathematics,jilek2014reference}.

Since the advent of digital signal processing in the 1960s~\cite{stranneby2004digital}, a wide variety of parametric audio generation functions have been proposed~\cite{lyons1997understanding,russ1999sound,shier2020spiegelib}. This includes oscillators, filters, equalizers, and envelopes, which can be used sequentially or in parallel~\cite{lyons1997understanding,russ1999sound}.  Sound design with a synthesizer is done by modifying the parameters of the functions until reaching a desired output~\cite{roads1996computer,pinch2004analog}.

A relatively small subset of these synthesis methods have been used for automatic sound-matching, notably additive and/or subtractive synthesizers~\cite{esling2019flow,yee2018automatic,mitchell2007evolutionary,salimi2020make}, physical modeling~\cite{riionheimo2003parameter}, and frequency/amplitude modulation (\gls{FM}/\gls{AM})~\cite{horner1993machine,vahidi2023mesostructures}.

Differential DSP (\gls{DDSP}) is a general term for applications that combine machine learning techniques such as gradient descent~\cite{goodfellow2016deep,boyd2004convex} with DSP~\cite{engel2020ddsp}. Here, we take a differential approach to sound-matching using DDSP, i.e., the use of differentiable loss functions and gradient descent to update synthesizer parameters to reduce the difference between the output and the target sound.

Implementing complex DSP functions in a differentiable manner can be challenging, and effective differentiable audio similarity measures require careful mathematical expression of the desired attributes of sound. These issues have likely contributed to the limited exploration of this domain. Perhaps the most relevant recent work is the comparison of two differentiable loss functions with a simple FM synthesizer by Vahidi \textit{et al.}~\cite{vahidi2023mesostructures}.

\subsection{Sound Representation and Loss Functions}
\label{sec:loss_funcs}
A digital sound (or an audio signal) is a series of numbers~\cite{smith1991viewpoints,smith2007mathematics}. To compare two digital sounds, the two corresponding series are passed to a function that measures their similarity. Two signals can sound identical to our ears, without having any values in common~\cite{moore2012introduction}. This necessitates the use of proxy representations (or feature extractors) when comparing sounds automatically. Similarity between the target sound and the synthesizer output is then measured by some form of subtraction and summation of the proxy representations.

In sound-matching, particularly in a Deep Learning (\gls{DL}) context~\cite{goodfellow2016deep}, the similarity function can also be called a \textit{loss} function, where the emphasis is on measurement and reduction of the distance between target and output. It is important to note that there is a close relationship between the loss function $L$ and the sound representation function $\phi$. A distance measure $d$ is applied to the features extracted by $\phi$.

\[
L(\theta, t) =  d\langle\phi(t),\phi(x)\rangle
\]

\noindent

A proxy representation is the output of the function \( \phi \), which can be thought of as a feature extraction function that maps the sounds \( t \) and \( x \) to their respective representations. 
The proportionality or distance metric $d$ has typically been the L1 distance~\cite{turian2020sorry}, calculated as the mean of the absolute difference between every point in the proxy representation~\cite{engel2020ddsp,vahidi2023mesostructures}:
\[
L(\theta, t) = \left\| \phi(t) - \phi(x) \right\|_1
\]

Here we discuss four methods of audio representation and the corresponding loss functions. Three of which have been used in previous works, and a novel method which we will make use of in our own experiments.

\subsubsection{Parameter Loss}
A common measure of similarity in sound-matching is the distance between synthesizer parameter sets, referred to as ``P-Loss"~\cite{han2023perceptual}. Typically, for the implementation of P-Loss the parameter sets are treated as vectors in space, and L1 or L2 distance is applied. There are two major limitations to this approach: First, the target and output sound must be made by the same synthesizer; otherwise the parameter sets cannot be compared (see Section~\ref{sec:matching_types}). Second, the relationship between synthesizer parameters and the audio output is not linear~\cite{shier2020spiegelib,han2023perceptual,esling2019flow}.

\subsubsection{Fourier Spectrograms}
\label{sec:fourier_specs}
Fourier-based transformations such as short-time Fourier transforms (\gls{STFT}), Mel-spectrograms, and Mel-frequency cepstral coefficients have been viewed as the de facto and state-of-the-art representation of audio~\cite{beauchamp2003error,mitchell2007evolutionary,yee2018automatic}, however, there are many issues associated with their use in sound-matching~\cite{turian2020sorry,vahidi2023mesostructures,han2023perceptual,uzrad2024diffmoog}. Fourier transformations allow for the conversion of a signal from the time-domain to the frequency domain. Audio spectrograms can be generated by segmentation of a piece of audio into overlapping windows followed by the application of Fourier transforms to each window. They are costly to compute, but provide a better temporal view of changes in frequency content~\cite{muller2007dynamic,smith2007mathematics}. There are different types of spectrograms that have a basis in Fourier transformations, but the most notable and commonly used is the STFT.

What we call \textit{Fourier-based Spectrograms} are variations on the STFT approach. For example, Mel-Spectrograms \textit{bin} frequencies on a near-logarithmic scale to better match human perception of frequencies~\cite{muller2007dynamic}. Multi-scale spectrograms (\gls{MSS}) used in recent works are a simple weighted average of multiple spectrograms with different parameters such as window size, number of frequency bins, and hop size~\cite{engel2020ddsp,vahidi2023mesostructures}; this may provide some improvements at a higher computational cost~\cite{turian2020sorry,engel2020ddsp}.

\subsubsection{Joint-Time Frequency Spectrum}
Recent works have focused on the limitations of parameter and spectral loss functions in sound-matching~\cite{vahidi2023mesostructures,uzrad2024diffmoog}, seeking to create more effective general solutions for the comparison of audio. 
Noting the aforementioned weaknesses of comparing STFT spectrograms, Vahidi \textit{et al.} proposed differentiable Joint-Time Frequency Scattering (\gls{JTFS})~\cite{anden2015joint} as an alternative to spectrogram loss in sound-matching, and showed improved performance in sound-matching with differentiable chirplet synthesizers~\cite{vahidi2023mesostructures}.  JTFS is the result of the application of a 2D wavelet transformation to the time-frequency representation of a signal~\cite{anden2015joint}.

\subsubsection{Dynamic Envelope Warping}
Dynamic Time Warping (DTW) is a method for measuring similarity between multi-dimensional time-series~\cite{rabiner1993fundamentals,muller2007dynamic,giorgino2009computing}. Given any two time-series $X = \{x_1,x_2,...,x_m\}$ and $Y = \{y_1,y_2,...,y_n\}$, we have indices $i\in\{1...m\}$ and $j\in\{1...n\}$ defining $X$ and $Y$. When the series are \textit{warped}, these indices change to expand or contract different portions of the series. To borrow the notation given by Muller~\cite{muller2007dynamic}, warped indices are a sequence $p=(p_1,...,p_L)$, where \(p_\ell = (m_\ell, n_\ell) \in [1 : m] \times [1 : n] \text{ for } \ell \in [1 : L]\), meaning that the indices for $X$ and $Y$ are reorganized under special conditions. In classical DTW, these conditions are \textit{monotonicity}, \textit{boundary matching}, and \textit{single step-size}.

DTW measures the distance between the time-series \textit{after} alignments, typically using Euclidean distance. The distance between a time-series and shifted versions of itself would be 0, regardless of shift amount~\cite{tavenard.blog.dtw}. Additional rules can be imposed to keep alignments locally constrained~\cite{itakura1975minimum,sakoe1978dynamic}.
% \textcolor{red}{DTW was originally used for signal comparison but using it the way we do is novel}

\subsection{Automatic Sound Matching}
\label{sec:sound_matching_definition}
In past works the sound-matching problem has been tackled from a variety of perspectives as it involves a number of modular parts that need to be considered: the choice of synthesizer $g$, the target sounds of interest, the representation function $\phi$, and the heuristic for finding the optimal $\theta$. We discuss these components in the upcoming sections, and give a coherent typology of the sound-matching tasks in Section~\ref{sec:matching_types}.

To give a formal definition of sound-matching, we expand the definition given by recent works~\cite{vahidi2023mesostructures,han2023perceptual}. The major components are as follows: 
\begin{itemize}
    \item $g(\theta)$: Parametric audio synthesizer $g$ which takes on parameters $\theta$ 
    \item $x$: The output of $g$, given a set of parameters $\theta$ or $g(\theta) = x$ 
    \item $t$: The target sound which we want to replicate or imitate. 
    \item $\phi$: Representation function or feature extractor. $\phi$ is applied to $x$ and $t$ to facilitate their comparison using the loss function.
    \item $L$: Loss or error function. $L(\theta,t)$ is a measure of distance between $x$ and $t$. Often proportional to the subtraction of their representations, or $ \phi(x) - \phi(t)$
\end{itemize}

We view the implementation of these components not as an optimization problem with a correct answer, but a \textit{creative} endeavor depending on the artist's needs. Consider the case where the goal is to create an 8-bit version of a high quality, organic snare drum~\cite{collins2007loop}: in such a case, there is no optimal or ``correct" version of an 8-bit snare sound, and multiple correct answers exist.

\subsection{Sound Matching Types}
\label{sec:matching_types}

Sound-matching tasks can be categorized based on the choices of target sound $t$, the loss $L$ and representation function $\phi$, and the heuristic used for minimizing $\theta$. We will discuss the subtypes of sound-matching tasks in the upcoming sub-sections.

\subsubsection{In-Domain vs Out-of-Domain}
\label{sec:in-domain}
The choice of domain depends on whether we want to use the same synthesizer for the target and output sounds, which we call \textit{in-domain}, or have target sounds that came from sources other than the synthesizer, or \textit{out-of-domain}. To paraphrase the description given by Masuda \textit{et al.}~\cite{masuda2021soundmatch}, if $g$, the synthesizer of choice, can accurately replicate the target sound $t$, or put differently, if $t$ itself is an output of $g$, then the sound-matching task is \textit{in-domain}. If $t$ is not an output of $g$, then the sound-matching task is \textit{out-of-domain}. In-domain tasks in general are simpler, and often there is a guarantee that there is a correct answer to the sound-matching problem, particularly if the goal is accurate replication of the sound. If the target sound is out-of-domain, replication is not guaranteed, and the goal becomes the \textit{imitation} of some aspect of sound.

\subsubsection{Replication vs Imitation}
Regardless of the domain, the generation goal can be \textit{replication} or \textit{imitation} of the target sound. In replication, the goal is to make an identical copy of the target sound. With imitation, the goal is to make new sounds that only retain a subset of the target's sonic features. While closely related to the in-domain versus out-of-domain problem, the choice of replication versus imitation is more dependent on how the loss and representation functions are defined.

From a signal perspective, replication is a scenario where the goal is to find a $\theta$ such that $t-x=0$ where $x=g(\theta)$. From a sound perspective, this definition of replication is unnecessarily restrictive, as $t$ and $x$ can sound identical but contain very different values. For example, if $t$ and $x$ were periodic waveforms at an audible frequency $f$ with different phase values, their subtraction would not be 0, yet they would sound identical to our ears, as our ears are insensitive to phase values~\cite{moore2012introduction}. Representation functions such as STFT spectrograms or Mel-frequency cepstral coefficients (\gls{MFCC}) become viable in these cases, as they allow for separation and removal of the phase information within an audio signal. The \textit{sound replication} goal can be defined, in layman's terms, as ``$x$ sounds exactly like $t$". More formally, the goal is to find parameters such that $\phi(t)-\phi(x)=0$, where $\phi$ transforms an audio into a spectrogram that removes the phase information from sounds, but otherwise represents the sonic information as accurately as possible.

With \textit{sound imitation}, the goal is the replication of some aspect of the target sound. Instrument conversion is a common example of imitation. As an example, a note played by a violin is converted to the same note played by a piano. Conversion of even longer pieces has been made possible with large enough datasets~\cite{engel2020ddsp}. Imitation of sound envelopes (how the loudness changes over time) is another interesting idea~\cite{caspe2023envelope}. Transfer of other sound properties such as reverb and delay has; also been a subject of research~\cite{engel2020ddsp}.

A target sound can have many properties of interest, and the imitation task is the transfer of a subset of those properties to other sounds. Relative to replication, imitation is a very open ended goal. From a research point of view, success in the replication task can be measured more reliably, often without the need for manual evaluation. With imitation, listening tests are needed to ensure quality of outputs, particularly in a supervised learning setting where learning is done by examples~\cite{salimi2020make,masuda2021soundmatch}. The need for subjective evaluations has perhaps contributed to the lack of research in imitation and out-of-domain sound-matching.

\subsubsection{Supervised Versus Direct Optimization}
\label{sec:optimization}

The goal of sound-matching is to find the optimal parameters $\theta^*$ that minimize the loss between the synthesizer output and the target sound. 
\[
\theta^* = \arg\min_{\theta} L(\theta,t)
\]

The heuristics in past works can be broadly split into two categories:
\begin{enumerate}
    \item Supervised methods which use datasets of sound and parameter pairs to describe the generation objective. Often with a large number of input and output examples~\cite{engel2020ddsp,salimi2020make,yee2018automatic,esling2019flow}
    \item Unsupervised or \textit{direct sound optimization} methods~\cite{uzrad2024diffmoog}. Usually require at most one target example, and a loss function that measures the difference between the target and synthesizer outputs. 
\end{enumerate}

Genetic algorithms (\gls{GA})~\cite{holland1992genetic} can be used for direct sound optimization, and have been the earliest and most common heuristic in sound-matching~\cite{horner1993machine,mitchell2007evolutionary,yee2018automatic,uzrad2024diffmoog}. These algorithms start with arbitrary parameter sets that can be treated as an evolving population where the genomes are the parameter values. The most fit members of the group are the parameters which perform the best in the loss function. The most fit parameters create a new generation of parameters via mutation (random change in subset of parameters) and crossovers (combination of parameter sets), and the process repeats until the goal or a maximum number of generations is reached.

One important aspect of the heuristics is whether they imitate the iterative nature of manual sound design---that is, listening to the output and adjusting the synthesizer parameters recursively until the desired goal is reached. Supervised methods generate their final output in a single step, while direct sound optimization methods refine their outputs iteratively until termination, which resembles manual sound design more closely.

\subsection{Iterative Sound Matching}
\label{sec:iterative_sound_matching}
The diagram in Figure~\ref{fig:sound_design_loop_iterative} shows a general model for iterative sound-matching that covers previous works in this domain. To begin the process, the parameters are arbitrarily initialized (usually random generation), the similarity of the target and output is measured, and the parameter is \textit{optimized} with the goal of increasing the similarity, or alternatively, reducing the loss. This process repeats until termination.

\begin{figure}[ht]
    \centering
\begin{tikzpicture}[node distance=2cm, auto]

    % Nodes
    \node (start) [text centered] {\( g (\hat{\theta}) = x \)};
    \node (L) [above of=start,right of=start, text centered] {\( L(\hat{\theta},t) {\ \propto \ } \phi(x) - \phi(t) \)};
    \node (optimize) [below of=L,right of=L, text centered] {Optimize $\hat{\theta}$};
    \node (new_theta) [below of=optimize,right of = start, text centered] {New \( \hat{\theta} \)};

    % Highlight and arrow for the start node
    \draw[->, very thick, red] (start) ++(-2,1.5) -- (start)
        node[midway, below, align=center, sloped, color=red] {Start}
        node[midway, above, align=center, sloped, color=red] {Random $\hat{\theta}$};

    % Arrows with multi-line labels
    \draw[->, bend left] (start) to node[midway, right, align=center] {} (L);
    \draw[->, bend left] (L) to node[midway, right, align=center] {} (optimize);
    \draw[->, bend left] (optimize) to node[midway, below, align=center] {} (new_theta);
    \draw[->, bend left] (new_theta) to node[midway, left, align=center] {} (start);

\end{tikzpicture}
    \caption{ Iterative approach to sound design. Synthesizer $g$ with arbitrary initialized parameters $\hat{\theta}$, creates sound $x$. The target sound $t$ is used as the goal. Parameters $\hat{\theta}$ are adjusted to minimize error (or loss) $L(\hat{\theta},t)$, where $L$ is proportional to the difference between the representations of $x$ and $t$. $\phi$ is the audio feature extractor, or representation function.}
    \label{fig:sound_design_loop_iterative}
\end{figure}

Genetic mutation has predominantly been the most common direct optimization method in previous works, particularly in non-differentiable settings. 
In a differentiable setting, gradient descent can be used to directly optimize the parameters of a synthesizer~\cite{vahidi2023mesostructures}. In the upcoming ``Past Works'' section, we will discuss the evolution of prior sound-matching research, with \textit{differentiable iterative sound-matching} being the latest extension of the field.

\begin{table*}[t]
\centering
\setlength{\tabcolsep}{3pt} % Reduce horizontal spacing
\renewcommand{\arraystretch}{1.05} % Reduce vertical spacing
\scriptsize % Reduce font size
\resizebox{\textwidth}{!}{
\begin{tabular}{|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{1.2cm}|>{\raggedright\arraybackslash}m{1.2cm}|>{\raggedright\arraybackslash}m{1cm}|>{\raggedright\arraybackslash}m{1cm}|}
\hline
\textbf{Work} & \textbf{Synthesis} & \textbf{Loss} & \textbf{Heuristics} & \textbf{Goal} & \textbf{Domain} & \textbf{Year} & \textbf{Gen. } \\
\hline
Horner \textit{\textit{et al.}}~\cite{horner1993machine} & Non-diff FM & STFT (McAulay-Quatieri) & Direct & Replicate & In & 1991 & Iter \\
\hline
Mitchel \textit{\textit{et al.}}~\cite{mitchell2007evolutionary} & Non-diff FM & Spectral Relative & Evolutionary & Replicate & In (Contrived) & 2007 & Iter \\
\hline
Yee-King \textit{\textit{et al.}}~\cite{yee2018automatic} & Non-diff VST & P-Loss (for Supervised), Spectral (MFCC) & Supervised \& Evolutionary & Replicate & In & 2018 & Both \\
\hline
Esling \textit{\textit{et al.}}~\cite{esling2019flow} & Non-diff VST & Spec MSS/SC P-loss & Supervised \& Modeling & Replicate & In* \& Out & 2019 & 1-shot \\
\hline
Shier \textit{\textit{et al.}}~\cite{shier2020spiegelib} & Custom, non-diff & FFT \& Spectral & Supervised \& Direct & Replicate & In & 2020 & Both \\
\hline
Salimi \textit{\textit{et al.}}~\cite{salimi2020make} & Non-diff & STFT and Envelope & Supervised & Imitate & Out & 2020 & 1-shot \\
\hline
Masuda \textit{\textit{et al.}}~\cite{masuda2021soundmatch} & DDSP & P-Loss initially, fine-tune with Spec. & Supervised & Replicate? & In \& Out & 2021 & 1-shot \\
\hline
Han \textit{\textit{et al.}}~\cite{han2023perceptual} & NN-encoder $\rightarrow$ physical model & PNP (approx. of STFT) & Supervised & Replicate & In & 2023 & 1-shot \\
\hline
Vahidi \textit{\textit{et al.}}~\cite{vahidi2023mesostructures} & Differentiable FM & JTFS/MSS & Direct (grad desc.) & Replicate & In & 2023 & Iter \\
\hline
Uzrad \textit{\textit{et al.}}~\cite{uzrad2024diffmoog} & DDSP chains & Synth. Chain and P-loss & Supervised & Replicate & In \& Out & 2024 & 1-shot \\
\hline
\end{tabular}
}
\caption{Summary of select works in sound-matching. Gen. refers to whether the generations are done in 1-shot (generally, output by a neural network) or a search is conducted with the goal of iteratively getting closer to the target sound.}
\label{tab:summary}
\end{table*}

\section{Past Works}
\label{sec:related_works}
Table~\ref{tab:summary} contains an overview of relevant literature, covering important works in sound-matching.  Earlier works in this domain focused on FM synthesis and genetic algorithms (or direct optimization methods)\cite{justice1979analytic,horner1993machine,mitchell2007evolutionary}. Along with the exponential increase in computation power, later works utilized a wider range of approaches to parameter search, estimation, and audio synthesis; particularly with supervised learning and neural networks (\gls{NN})~\cite{yee2018automatic,engel2020ddsp,esling2019flow}.

\subsection{Earlier Works}
FM synthesizers are simple to implement, yet very expressive~\cite{chowning1973synthesis}. Perhaps the earliest foundational work in sound-matching is the analytical approach by Justice~\cite{justice1979analytic} toward the decomposition and recreation of sounds with FM synthesis using one carrier, one modulator, and the corresponding envelopes:
\[ f(t) = I_c(t) \cos(\omega_c t + I_m(t) \cos(\omega_m t))
\]
Where $f(t)$ is the output signal, $I_c(t)$ is the carrier amplitude, $\omega_c$ is the carrier frequency, and $I_m(t)$ and $\omega_m$ are the envelope and frequency for the modulator~\cite{justice1979analytic}.

 The sound-matching works that followed used other heuristics for finding the synthesizer parameters, yet the simple FM synthesizer structure remained unchanged. Horner \textit{et al.}~\cite{horner1993machine} used GAs for re-synthesis of sounds with an FM synthesizer using one modulator and one carrier oscillator, and the McAulay-Quatieri (MQ) method~\cite{mcaulay1986speech} for measuring loss. MQ calculates the STFT and marks the start and end of \textit{trends} in a sound with the assumption that the harmonic content belongs to trends that have a beginning and an end.

Introduction of more complex models of synthesis such as wavetables~\cite{horner2003auto} and physical modeling~\cite{riionheimo2003parameter} rendered interesting results, but led to further questions about the nature and goals of sound-matching. Mitchell and Creasy noted the difficulty in distinguishing between \textit{inefficiency of the optimization engine} and \textit{synthesizer limitations} as the cause of failure or success in sound-matching~\cite{mitchell2007evolutionary}. They offered a \textit{contrived methodology} for finding the best evolutionary method for sound-matching using their FM synthesizer. The methodology postulates that the best search heuristic for in-domain search (where an exact target exists, and a wide range of targets can be produced by sampling from different points in the parameter space) would also be the best-performing heuristic for out-of-domain search on a dataset of muted trumpet tones~\cite{opolko1989mcgill}; however, testing was not extensive and yielded some contradictory results.

Testing sound-matching experiments can be difficult, in part due to the large number of parametric components and navigation of complex non-linear systems. The issue with finding the best approaches to sound-matching is exacerbated when slight changes to the pipeline render previous results moot. Perhaps due to these reasons, the contrived methodology was not extensively tested by Mitchell and Creasy, who only provided a single experiment where the best evolutionary strategy found by in-domain search would also perform well in out-of-domain synthesis. As they noted, modifications to the synthesizer, loss function, and sound domain would lead to a problem with an entirely new search space~\cite{mitchell2007evolutionary}.

\subsection{Introduction of Deep-Learning}
Recent years have seen more works in sound-matching using machine learning techniques. In 2018, Yee-King \textit{et al.} rendered 60,000 audio-parameter pairs from the \textit{Dexed} \gls{VST} synthesizer~\footnote{https://asb2m10.github.io/dexed/}, and showed that NNs trained on this dataset can outperform GA and hill-climber (\gls{HC})~\cite{hoffmann2000heuristic} methods in rendering speed, with slight improvements in MFCC error (used as an objective performance test). The speed improvement appears trivial, considering the iterative nature of GAs when compared to offline training of supervised models. For GAs and HC optimizer, MFCCs were used as a measure of performance as well as a loss function. For training the networks, P-loss was used, since differentiable MFCCs were not possible in their pipeline. Importantly, informal hearing tests revealed that the performance of even the best NN model was unsatisfactory, possibly due to the complex nature of the synthesizer, which features 155 parameters.

% Programming synthesizers and querying sounds by imitating human vocals has also been a subject of interest~\cite{Cartwright2014SynthAssistQA,siamesevocalimitation2018}. For example, Cartwright and Pardo created a framework for searching pre-made patches for a synthesizer to best imitate vocals using audio features such as pitch, spectral centroid, and loudness~\cite{Cartwright2014SynthAssistQA}.

Esling \textit{et al.} also noted the issue of using P-Loss in sound-matching~\cite{esling2019flow}, considering the non-linear mapping between parameters of a synth and the audio output. In their approach, dubbed as \textquote{flowsynth}, they used a large parameter-to-audio dataset for a commercial \gls{VST} synthesizer. In flowsynth, audio is encoded to an auditory space, using a NN autoencoding approach~\cite{hinton2006reducing}. The encodings in the auditory space can then be converted to the parameter space, but only for that particular synthesizer. Using P-Loss and Spectral Convergence (Frobenius norm of the spectrograms) as metrics, flowsynth outperformed other NN approaches that did not use an auditory to parameter space conversion, but formal hearing tests were not conducted.

Masuda \textit{et al.} also applied supervised learning to synthesizer parameter estimation~\cite{masuda2021soundmatch}. Their work highlights the issue of non-linearity in parameter-to-synthesizer outputs and out-of-domain search. This work uses a differentiable subtractive synthesizer (two oscillators and an LP filter). A NN model was pre-trained using an in-domain dataset of randomly selected parameters and P-Loss. After training, the model was fine-tuned using 20,000 out-of-domain sounds from the NSynth dataset~\cite{engel2017neural} and multi-scale spectrogram loss~\cite{engel2020ddsp}. This approach proved more effective---that is, lower multi-level spectral difference in out-of-domain tests---than baseline models, which were either not exposed to out-of-domain sounds or trained exclusively with P-Loss. Subjective hearing tests were conducted, showing a preference for the fine-tuned model~\cite{masuda2021soundmatch}.

Rather than focusing on a particular implementation, Shier \textit{et al.} presented Spieglib, a library for implementation of sound-matching pipelines (which Shier \textit{et al.} referred to as ``automatic parameter search'')~\cite{shier2020spiegelib}. This library provides extensible choices of deep-learning estimators with \textit{tensorflow}~\cite{abadi2016tensorflow}, genetic estimators with \textit{DEAP}~\cite{DEAP_JMLR2012}, feature extractors with \textit{librosa}~\cite{mcfee2015librosa}, and different choices for synthesizers, evaluators, dataloaders. Shier \textit{et al.} provide an example experiment using this library. This example has a setup similar to Yee-King \textit{et al.}~\cite{yee2018automatic}, where the \textit{Dexed} VST was used and different genetic and NN estimators were compared by using MFCC difference as the minimization objective. Depending on the NN architecture, MFCC or STFT spectrograms were used as the input. Unlike previous comparisons of NN and genetic algorithms, the best estimator was the multi-objective non-dominated sorting genetic algorithm III (NSGA-III). This algorithm had multiple minimization objectives: 13-band MFCC, STFT, and five spectral features. Shier \textit{et al.} report that all models performed relatively well at imitation of harmonic distributions, but the multi-objective function was the only optimizer which did \textit{not} struggle with matching temporal envelopes of the audio spectrum~\cite{shier2020spiegelib}.

Differentiable loss functions that use spectrogram differences can be computationally expensive. To mitigate this, Han \textit{et al.}~\cite{han2023perceptual} introduced ``perceptual-neural-physical loss'' (PNP). PNP is an approximation of spectrogram loss functions; specifically, a loss function that uses the L2 norm of the difference between the features of two sounds, or $||\phi(t) - \phi(x)||^2_2$. PNP loss functions are fast and differentiable, but require training and parameter estimation. A Riemannian metric M needs to be calculated for the minimization of locally linear approximation of the ``true" spectral loss function~\cite{han2023perceptual}.

\[
\|\phi(t) - \phi(x)\|_2^2 = \langle \tilde{\theta} - \theta | \mathbf{M}(\theta) | \tilde{\theta} - \theta \rangle + O(\|\tilde{\theta} - \theta\|_2^3). \tag{4}
\]

During the training phase, this metric is calculated alongside the training of the neural parameter estimator. After training, PNP can be used as a fast approximation of the loss function, $\|\phi(t) - \phi(x)\|_2^2 $. Han \textit{et al.} observed slight improvements to run-time (for testing) and minor improvements to the loss values under certain conditions. These experiments were based on supervised learning and physical modeling (of a drum head) for synthesis~\cite{smith2010physical}, and compared PNP, P-Loss, MSS, and JTFS functions~\cite{han2023perceptual}. This approach was more recently applied to other supervised settings using physical and FM synthesizers~\cite{han2024learning}.

Uzrad \textit{et al.}~\cite{uzrad2024diffmoog} took another unique approach to sound-matching: using a differentiable \textit{synthesis chain} of DSP generators and effects and a loss function that combines P-Loss with a \textit{signal-chain loss}. Like other supervised learning replication approaches, during training a target sound $t$ is generated by random  $\theta$ parameter selection, and a model attempts to an output $\tilde{\theta}$ that generates a sound similar to $t$. The synthesizer is a customizable chain of effects, which feed one output as input to the next step of the chain; signal-chain loss compares the parameter and output difference at every output step in the chain~\cite{uzrad2024diffmoog}. Possible chain functionalities are FM/AM, Low Frequency Oscillators (\gls{LFO}), filters, and envelopes. Like the results shown by Masuda \textit{et al.}~\cite{masuda2021soundmatch}, better out-of-domain results were achieved when pre-trained on in-domain data and fine-tuned using out-of-domain NSynth data~\cite{engel2017neural}.

Uzrad \textit{et al.}\cite{uzrad2024diffmoog} used log-spectral distance (LSD) as their loss function, where   
$\text{LSD} = \| \log(S(t)) - \log(S(x)) \|_F$, S is the mel-spectrogram of the sounds and F is the Frobenius norm~\cite{golub2013matrix}. The Frobenius norm of a matrix $A$ is defined as
$ \|A\|_F = \sqrt{\sum_{i,j} |a_{ij}|^2}$. Here, $A$ would be the 2D matrix resulting from subtracting two log-spectrograms. Along with other works, \textbf{non-convexity of loss surfaces when applying gradient optimization} is a noted issue~\cite{turian2020sorry,vahidi2023mesostructures,uzrad2024diffmoog}.

\section{What is Lacking In the Field}
\label{sec:lacking}
  % First, we need to clearly define the characteristics we are looking for with non-generic loss functions, otherwise the chances of finding the sound(s) of interest is slim. Second, we need to move beyond direct replication, and focus on imitating target sounds that are outside the synthesizers capabilities. Third, more complex methods of synthesis need to be explored for iterative sound-matching. Fourth, we need to find optimization methods that can effectively navigate the the sinusoidal nature of loss function landscapes.

Based on the literature reviewed, we find several gaps in past research that makes the application of sound-matching for sound-designers difficult. Thus far, replication of in-domain sounds has been the main subject of research. With in-domain replication, we already have the target sound \textit{and} the program that makes it. This simplifies the problem of measuring performance and is the reason why this type of sound-matching research is by far the most common. However, practical application of sound-matching requires answers to much harder problems. Sound-matching would be an interesting creative tool if we could use it to either:
\begin{enumerate}
 \item \textbf{Replicate out-of-domain sounds}: this would give us the  the parameters that could closely approximate a target sound, and we can modify the sound further by modulating the synthesizer. 
\item \textbf{Imitate in-domain or out-of-domain sounds}: This would allow for applying the characteristics of one sound to another. 
\end{enumerate}

We need to be aware of the ``open set recognition'' (\gls{OSR}) problem here \cite{mundt2019open,gers2000learning}.  OSR is a well documented issue in DL that manifests when the set or category that the network is supposed to reject or accept cannot be explained via examples. With imitation or replication of out-of-domain sounds, we can expect issues related to OSR to occur, since target and output sounds would be from different domains~\cite{salimi2021percussive}.

Perhaps one method of addressing the OSR problem is the use of representations (or feature extraction) methods that reduce sounds to a feature vector in the same domain. For example, a representation that only looks at the envelope of sounds is far less likely to be subject to the OSR problem. \textcolor{highlight}{We highlight the lack of non-generic loss functions as a major weakness in previous works and an interesting area to explore}.

Another weakness \textcolor{highlight}{is that the synthesis methods in past works are not representative of the common methods used by sound designers and modern synthesizers}. Simple FM/AM synthesizers have generally been the go-to methods of sound-matching. We believe that general statements regarding the effectiveness of an approach to iterative sound-matching (particularly in regards to a loss function) can only be reasonably made in the context of the subset of the tested methods of synthesis.
\textcolor{highlight}{The simplicity of parameter optimization approaches} is another area of weakness in past works. Classic genetic algorithms update the parameters randomly, while in recent differentiable settings, simple gradient descent methods are used. Past works have provided issues which arise in differentiable parameter estimation due to periodic loss functions. Perhaps the use of reinforcement learning (\gls{RL}), or other algorithms that continually learn from their environment, could help with this problem.

\subsection{Defining The Major Issues}
\label{sec:addressing_problems}
Based on the weaknesses noted above, we propose four problems that need to be addressed to improve the sound-matching field. 
\begin{enumerate}
    \item \LossSelect: Is there a \textit{best performing} loss function, or is the selection of the loss function dependent on factors such as the sound domain, synthesizer capability and the desired characteristics of the output? 
    \item \SynthSelect: There is a lack of diversity in the synthesis methods used in sound-matching. The use of different DSP functions for iterative sound-matching and their effects on the outcome or the interactions with the loss function have not been tested.
    \item \PeriodicLoss: The sinusoidal nature of loss function landscapes is a problem that appears frequently. In differentiable settings, this problem causes gradient descent updates to local minima. 
    \item \OutDomain: Sound-matching with out-of-domain sounds is an unexplored area, yet a necessary one for practical applications for sound designers.
\end{enumerate}

In Section~\ref{sec:what_is_done}, we discuss our work in simultaneously bridging the knowledge gap with regards to the \LossSelect~and \SynthSelect~problems. The latter two issues are left for future work; possible approaches are discussed in Sections~\ref{sec:summary_conclusion} and~\ref{sec:future}.

\section{Programs, Methodology, and Results}
\label{sec:what_is_done}
In this section we report the results of our efforts in addressing the first two problems. \textbf{We compare the performance of iterative sound-matching experiments, where four differentiable loss functions are applied to four differentiable synthesizers.} What we mean by performance is the similarity of the final output to the target, as measured by P-Loss, MSS, and manual listening.

\textbf{The hypothesis here is that the best performing loss function for sound-matching is influenced by the method of synthesis}. The results of these experiments show that there is no universally optimal loss function, strengthening the idea that the selection of the loss function and synthesis method is not an optimization problem, but a subjective choice depending on the needs of the sound designer.

In Section~\ref{sec:methodology} we lay out the experiment methodology. Section~\ref{sec:loss_implementation} discusses the details of implementation for the loss functions. Section~\ref{sec:programs} covers the characteristics of the programs, and finally Section~\ref{sec:results_posthoc} shows that the effect of loss function choice on the success of iterative sound-matching is dependent on the signal process chain of interest.

\subsection{Experiment Methodology}
\label{sec:methodology}
We have highlighted the lack of attention given to the selection of loss functions in sound-matching, and noted that the few existing works addressing this issue evaluate their proposed losses using only simplistic sound synthesis methods. \textcolor{highlight}{We hypothesize that the best performing loss function is influenced by the synthesis techniques utilized in the synthesizer}. This hypothesis targets the issues of \LossSelect~and \SynthSelect. For our experiments, we test four different loss functions and compare their success in sound-matching using four different synthesizers. We verify the success of the sound-matching process using two automatic measures of sound similarity (MSS and P-Loss) and a manual blinded hearing test conducted by two of the authors where likert scores~\cite{jebb2021review} of 1-5 are assigned to a subset of targets and outputs.

An \textit{experiment} is one iteration of sound-matching, given a loss function, target sound,
and synthesizer program. When the final iteration is complete, a score is given (automatically with an algorithm or with manual listening) to an experiment that measures its success in replicating the target sound. Consider the following example: we have two loss functions, $L_1$ and $L_2$, and a single synthesizer program $P$ with a target parameter set $\theta^*$ where the sound output is $t$, or $P(\theta^*) = t$. We run a large number (in this case, 300 for each loss function) of experiments where we randomly set the initial ($\theta_0$) and target parameters ($\theta^*$) of $P$ and run an iterative sound-matching loop for each loss function, with max number of iterations set to 200. When the loop is terminated, we measure the similarity of the final output sound $x$ (where $P(\theta_{199}) = x$) to $t$, or if using P-Loss we compare $\theta_{199}$ to $\theta^*$. Since we ran 300 experiments for each loss, we can compare the bootstrapped distribution of losses or likert scores~\cite{jebb2021review} assigned by manual review to determine whether $L_1$ or $L_2$ led to better performance. This simplified example leads us to our methodolgy which has four loss functions, four synthesizer programs, and three different methods of scoring sound similarity.

\subsubsection{Determining Best Performers}
Determining the best performing loss function(s) might seem complicated since we have different synthesizer programs, loss functions, and evaluation metrics. Figure~\ref{fig:posthoc_evaluation} is a visualized summary of how the loss functions are ranked in three ways (MSS, P-Loss, Likert scores).

We have four loss functions and four synthesizer programs. The maximum number of iterations set to 200, and 300 experiments for each loss/synthesizer combination. Each synthesizer program  is paired with four loss functions.  Under each of these loss functions, there are 300 experiments. Note that P-Loss and MSS are performance evaluation methods that measure similarity of final output and target sound; They are not one of the four loss functions. P-Loss and MSS are assigned automatically to each of the 300 pairs of target and output sounds. Blinded manual scoring using a Likert scale~\cite{jebb2021review} is assigned to 40 randomly selected sounds, where integers from 1 and 5 are assigned, with 1 meaning no similarity and 5 for near identical.

For each program and loss function pair, we have three distributions that can be used to rank the loss functions. Two with 300 automatically assigned similarities and one with 80 likert scores. For consistency and statistical robustness, the distributions are upsampled to 1000 values using bootstrapping~\cite{tibshirani1993introduction}. Bootstrapping gives an estimation of the distribution for the mean performance of each experiment. $k$ (set to 1000) samples of $n$ values (set to 100\% of the values) are taken with replacement from the empirical distribution (list of 300 performance values for MSS and P-Loss or 80 for manual rankings). The mean performance is calculated for each of the $k$ samples. The $k$ estimates of the mean performance give us a bootstrapped distribution for each loss function.

For each program, the loss functions are ranked using the non-parametric Scott-Knott (\gls{SNPK}) test~\cite{tantithamthavorn2017mvt,tantithamthavorn2018optimization}, which ranks the loss functions from rank 1 (best, or highest distribution values) to maximum rank of 4 (worst, with lowest distribution values). SNPK may determine that multiple functions belong to the same rank, in which case the maximum rank would be less than 4. Lower P-Loss and MSS values indicate higher similarity, therefore (for consistency with manually assigned ranks) the inverse of their output is used, where higher values indicate greater similarity.

\input{extra_figures/evaluation_graph.tex}

\subsubsection{Training Parameters and Gradient Calculations}
 Previous work utilized gradient fields for analysis of the algorithm performance and gradient update~\cite{vahidi2023mesostructures}. Gradient fields are not the center of our analysis, but we include them in Appendix~\ref{appendix:gradient_fields}. During the experiments, the updates applied to the initial parameters are recorded for gradient field generation. For updating the weights, we use RMSProp, which operates similar to stochastic gradient descent (SGD)~\cite{goodfellow2016deep}, with the caveat that the gradients of each weight are scaled by the root-mean-square of past gradients of that weight. We used a fixed learning rate of 0.045 for all experiments. The maximum number of iterations is set to 200, where, based on our observations, the parameters have either irrecoverably diverged outside the acceptable ranges, or are stuck at a local minimum. At the maximum number of iterations the search is terminated and the P-Loss and MSS (as described in Section~\ref{sec:fourier_specs}) are calculated using the iteratively updated parameters, or the final sound output.

Additionally, the gradients are large, and calculated backwards sequentially through the signal processing chain. This backward calculation resembles the \textit{exploding gradients} problem in recurrent neural networks~\cite{gers2000learning}. For this reason, gradient clipping~\cite{goodfellow2016deep} is used as a preemptive measure to prevent large divergences in the parameter updates, but it is unclear if it facilitates learning in these experiments. The gradient clipping function~\cite{goodfellow2016deep} takes a set of gradients $\gamma_i$ and ensures that the $\ell_2$ norm of all gradients, or $||\gamma||_2:=\sqrt{\sum\limits_{i}(\gamma_{i}\cdot\gamma_{i})}$ does not exceed a clipping threshold of c (here set to 1). Clipped gradients $\gamma_i^{\text{clipped}}$ are calculated by multiplying them with a scale factor $\mathcal{S}$
\[
\mathbf{\gamma}_i^{\text{clipped}} = \mathbf{\gamma}_i \cdot \mathcal{S}
\]
Where $\mathcal{S}$ is a value less than or equal to 1, depending on whether the norm of the gradients exceeds the threshold $c$.  
\[
\mathcal{S} = \frac{c}{\max(\|\mathbf{\gamma}\|_2, c)}
\]

\subsection{Loss Function Implementation Details}
\label{sec:loss_implementation}
\subsubsection{STFT Losses}
Due to their ubiquity and lower cost of gradient calculation, we use STFT as the basis of two of the loss functions. As discussed in Section~\ref{sec:matching_types}, the L1 or L2 distances are the most common methods of comparing different spectrograms, however, there are situations where this might lead to shortcomings. An example given by Vahidi et  al.~\cite{vahidi2023mesostructures} is two chirplets (tones that increase in pitch) starting and ending at the same frequencies. If one of these tones is slightly shifted in time, then the spectrograms will no longer overlap at any point, despite their sonic similarity. In this example, we do not know if this is an issue of the spectrograms, or the L1 measure used for their comparison. In the computer vision field, Structural Similarity Index (SSIM)~\cite{wang2004imagesssim,wang2009mean} and Scale-Invariant Mean Squared Error (SIMSE)~\cite{barron2014shapessimse} have been used as alternatives that improved performance in various tasks relative to L1.

We define the \LoneSpec~and \SIMSESpec~functions as the application of L1 and SIMSE to the STFT spectrograms. The STFT spectrogram function uses 512 FFT bins, window size of 600 samples, and hop length (how many samples the window shifts) of 100 samples. The L1 and SIMSE implementations are differentiable.

\subsubsection{JTFS Loss}
The \JTFS~loss function is the application of L1 difference to the JTFS representations of two sounds~\cite{vahidi2023mesostructures}. The code used for the JTFS transformation is the differentiable implementation of a 1-dimensional JTFS function provided by Andreux \textit{et al.}~\cite{kymatio}.

\subsubsection{Soft-DTW Loss}
We use the soft-DTW function, which is differentiable and --- depending on its parameters --- not shift-invariant~\cite{cuturi2017soft,janati2020spatio,tavenard.blog.softdtw}. The loss function \DTWEnv~is the application of the soft-DTW function to the envelope of the two sounds being compared~\cite{lyons1997understanding}. The envelope is calculated by creating the STFT spectrogram of a sound (the same process used in Section~\ref{sec:fourier_specs}) and summing the values at each timestep.

\subsection{The Synthesizers}
\label{sec:programs}
The synthesizer programs are meant to be simple examples that test the building blocks of digital sound synthesis. Subtractive, additive, and FM/AM synthesis are three of the most common techniques in sound design~\cite{smith1991viewpoints}. In subtractive synthesis, frequencies are removed from a sound via digital filters. In additive synthesis, complex sounds are created via the linear combination of simpler sounds~\cite{lyons1997understanding,smith2007introduction}. As discussed previously, FM/AM synthesis refers to the general technique of modulating the frequency or amplitude of a waveform (the carrier) by another waveform (the modulator).

For each program, we provide the Faust code~\cite{orlarey2009faust}. Faust is a functional language for audio synthesis that can succinctly define signal processing chains. The publicly available Faust IDE\footnote{\url{https://faustide.grame.fr/}} is the simplest method for running the programs. A brief overview of each program is given in Table~\ref{tab:programs}. The programs are first defined in Faust, then converted (or \textit{transpiled}) to differentiable Jax functions~\cite{deepmind2020jax} using the DawDreamer library\footnote{\url{https://github.com/DBraun/DawDreamer}}.

 Once the programs are transpiled to differentiable synthesizers, the iterative sound-matching procedure is:
 \begin{enumerate}
    \item Initialize $\theta^*$ and $\hat{\theta}$, i.e., random generation of target and initial parameters uniformly over a predefined range
    \item Generating the output of the synthesizer with $\hat{\theta}$ (the length of the output is set to 1 second with sample-rate of 44100 Hz) 
    \item Calculating the loss between the target and output
    \item Applying gradient updates to the synthesizer parameters
    \item Repeating the second step with the updated parameters $\hat{\theta}$, until maximum number of iterations has been reached
 \end{enumerate}

A general overview of the programs and the predefined parameter ranges is given in Table~\ref{tab:programs}. In the following subsections, we describe the programs and their parameters in more detail, followed by the methodology in Section~\ref{sec:methodology} and results in Section~\ref{sec:results_posthoc}.
\begin{table*}[htbp]
\centering
\begin{adjustbox}{max width=\textwidth}
\begin{tabular}{|c|c|c|c|l|}
\hline
\textbf{ID} & \textbf{Parameters} & \makecell{\textbf{Param. 1}\\\textbf{Range}} & \makecell{\textbf{Param. 2}\\\textbf{Range}} & \textbf{Description} \\
\hline
\BPNoise & hp\_cut, lp\_cut & 50-1000 & 1-120 & Band-pass of Noise signal \\
\hline
\AddSineSaw & saw\_freq, sine\_freq & 20-1000 & 20-1000 & Addition of sine and saw waves \\
\hline
\AmpMod & amp, carrier & 1-20 & 20-1000 & LFO on amplitude of noise osc \\
\hline
\FMMod & amp, carrier & 0-5 & 0-4 & LFO on amplitude of saw osc \\
\hline
\end{tabular}
\end{adjustbox}
\caption{Overview of Program Parameters and the range of values each parameter can take.}
\label{tab:programs}
\end{table*}

\subsubsection{\BPNoise}
\label{sec:program0}
\begin{lstlisting}[caption={\BPNoise}, label={lst:program0},language=Faust,float,floatplacement=htbp,xleftmargin=2em,xrightmargin=1em,firstnumber=0]
import("stdfaust.lib");
lp_cut = hslider("lp_cut",900,50,1000,1);
hp_cut = hslider("hp_cut",100,1,120,1);
process = no.noise:fi.lowpass(3,lp_cut):fi.highpass(10,hp_cut);
\end{lstlisting}

\BPNoise{} is a bare bones example of subtractive synthesis using digital filters~\cite{smith2007introduction}. Despite the ubiquity of subtractive synthesis in practical sound design, it has rarely been tested in differentiable sound-matching. In this program, a noise signal is fed through a band-pass (BP) filter. This removes the frequencies outside the low-pass (LP) and high-pass (HP) cutoffs. The noise generator produces all frequencies with random variation over time. The LP filter removes frequencies over its threshold frequency, and the HP removes frequencies lower than its threshold~\cite{smith2007introduction}. The search parameters are the cutoff thresholds for the LP and HP filters. Listing~\ref{lst:program0} shows the Faust code for \BPNoise.

\subsubsection{\AddSineSaw}
\label{sec:program1}

\begin{lstlisting}[caption={\AddSineSaw}, label={lst:program1},language=Faust,float,floatplacement=TBH,xleftmargin=2em,xrightmargin=1em,firstnumber=0]
import("stdfaust.lib");
saw_freq = hslider("saw_freq",800,20,1000,1);
sine_freq = hslider("sine_freq",300,20,1000,1);
sineOsc(f) = +(f/ma.SR) ~ ma.frac:*(2*ma.PI) : sin;
sawOsc(f) = +(f/ma.SR) ~ ma.frac;
process = sineOsc(sine_freq)+sawOsc(saw_freq);
\end{lstlisting}

\AddSineSaw{} is an additive program that combines a saw and a sine wave function. Like subtractive synthesis, additive synthesis is a common approach to sound design that has not been extensively tested as a benchmark in sound-matching. The search parameters here are the frequency of the sine and saw oscillators. Listing~\ref{lst:program1} is the Faust code for \AddSineSaw.

\subsubsection{\AmpMod}
\label{sec:program2}
\begin{lstlisting}[caption={\AmpMod}, label={lst:program2},language=Faust,float,floatplacement=TBH,xleftmargin=2em,xrightmargin=1em,firstnumber=0]
import("stdfaust.lib");
amp = hslider("amp",0.5,0,5,0.01);
modulator = hslider("modulator",0.5,0,4,0.01);
sineOsc(f) = +(f/ma.SR) ~ ma.frac:*(2*ma.PI) : sin;
process = no.noise*sineOsc(modulator)*amp;
\end{lstlisting}

\AmpMod{} involves amplitude modulation of a noise generator by an LFO with \textit{modulator} as its frequency and a global \textit{amp} value that does not change over time and applies to the entire signal. The search parameters for this program are the LFO frequency and the amp value. The amp value acts as a global volume control, and since the sounds are normalized before analysis, it is likely inconsequential to the performance. This program is meant to be a stepping stone to \FMMod, which utilizes proper AM synthesis. Listing~\ref{lst:program2} is the Faust code for \AmpMod.

\subsubsection{\FMMod}

\label{sec:program3}
\begin{lstlisting}[caption={\FMMod}, label={lst:program3},language=Faust,float,floatplacement=TBH,xleftmargin=2em,xrightmargin=1em,firstnumber=0]
import("stdfaust.lib");
carrier = hslider("carrier",100,20,1000,1);
amp = hslider("amp",6,1,20,1);
sineOsc(f) = +(f/ma.SR) ~ ma.frac:*(2*ma.PI) : sin;
sawOsc(f) = +(f/ma.SR) ~ ma.frac;
process = sineOsc(amp)*sawOsc(carrier);
\end{lstlisting}
Relative to additive and subtractive synthesis, AM/FM synthesis are less common methods of sound design. Due to their frequent use in previous works, we also tested an AM synthesizer. The synthesizer program is the multiplication of a low frequency sine oscillator with frequency parameter \textit{amp}, and a saw oscillator with frequency parameter \textit{carrier}. Listing~\ref{lst:program3} is the Faust code for \FMMod.

\subsection{Post hoc Analysis of Loss Performance}
\label{sec:results_posthoc}
As summarized in Figure~\ref{fig:posthoc_evaluation}, for each program and scoring method, there are four sets of empirical results (a set of 1000 bootstrapped values for each loss function). A post-hoc test is conducted in two stages in order to determine whether there is a difference in performance between loss functions for each program, and if so, which are the best performers. The first stage determines whether there is a difference in group means, with the null hypothesis being that all groups have similar mean ranks. If the null hypothesis is rejected, the second stage ranks the loss functions from best to worst using the non-parametric Scott-Knott algorithm (\gls{SNPK})~\cite{tantithamthavorn2017mvt,tantithamthavorn2018optimization}.

The Kruskal-Wallis test is used for the first stage~\cite{kruskal1952use}. This test works by pooling and ranking all performance evaluation measures for a program (either P-loss or MSS), then testing whether the differences in mean rank for all loss function groups is zero. The Kruskal-Wallis test calculates an H statistic that is compared to a chi-square distribution with k-1 degrees of freedom (k is the number of groups, or 4, and degrees of freedom is 3). Using the significance level of 0.05, if the H statistic is greater than the critical value of the chi-square distribution then the null hypothesis is rejected, meaning that at least one group has a mean rank significantly above others~\cite{kruskal1952use}. Table~\ref{tab:kruskal_auto} shows the Kruskal-Wallis results for every program and automatic evaluation method, and shows that significant differences exist in all cases except MSS for the \AmpMod{} program. The manual ranking results showed significant differences in all cases, as shown in table~\ref{tab:kruskal_manual}.

\begin{table}[ht]
\centering
\begin{tabular}{|c|c|c|c|c|}
\hline
\textbf{Program} & \textbf{Eval. Method} & \textbf{H-Stat.} & \textbf{P-value} & \textbf{Reject} \\
\hline
\BPNoise & MSS      & 81.19  & $1.70 \times 10^{-17}$ & Yes \\
\BPNoise & P-Loss  & 163.60 & $3.07 \times 10^{-35}$ & Yes \\
\AddSineSaw & MSS      & 308.97 & $1.14 \times 10^{-66}$ & Yes \\
\AddSineSaw & P-Loss  & 348.42 & $3.28 \times 10^{-75}$ & Yes \\
\AmpMod & MSS      & 1.35   & $0.7171$               & No \\
\AmpMod & P-Loss  & 366.76 & $3.50 \times 10^{-79}$ & Yes \\
\FMMod & MSS      & 564.65 & $4.65 \times 10^{-122}$ & Yes \\
\FMMod & P-Loss  & 229.19 & $2.07 \times 10^{-49}$ & Yes \\
\hline
\end{tabular}
\caption{Kruskal-Wallis results by program and evaluation method. The null hypothesis is that no difference in performance is observed between loss functions; it is rejected for all programs and evaluation methods with the exception of MSS combined with \AmpMod.}
\label{tab:kruskal_auto}
\end{table}

\begin{table}[ht]
\centering
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Program} & \textbf{H-Stat.} & \textbf{P-value} & \textbf{Reject} \\
\hline
\BPNoise & 9.45  & $2.39 \times 10^{-2}$ & Yes \\
\AddSineSaw & 9.45  & $2.39 \times 10^{-2}$ & Yes \\
\AmpMod & 32.71 & $3.70 \times 10^{-7}$ & Yes \\
\FMMod & 207.58 & $9.69 \times 10^{-45}$ & Yes \\
\hline
\end{tabular}
\caption{Manual ranking Kruskal-Wallis results. Significant differences are observed for all programs.}
\label{tab:kruskal_manual}
\end{table}

\subsubsection{Best Performers For Each Program}
Based on the bootstrapped distribution of scores, the SNPK algorithm ranks the loss functions from 1 (best) to a maximum of 4 (worst). In cases where the distributions are similar, multiple loss functions can be clustered into the same rank.

We use color-coded violin plots to visualize the best performers for each program. In the violin plots, the corresponding colors for each rank are
\colorbox{rank1}{\textcolor{black}{\textbf{1}}} \colorbox{rank2}{\textcolor{white}{\textbf{2}}} \colorbox{rank3}{\textcolor{white}{\textbf{3}}} \colorbox{rank4}{\textcolor{black}{\textcolor{white}{\textbf{4}}}}

\subsubsection{\BPNoise}
For this synthesizer program, the spectrogram-based models performed the best. This makes intuitive sense, as the visual effects of a band-pass filter on white noise are readily apparent in a spectrogram. As shown in Figure~\ref{fig:npsk_p0}, manual hearing test and MSS selected \SIMSESpec~as the best performer and \LoneSpec~as the second-best performer, while P-loss gave the reverse order. 
\begin{figure}[htbp]
    % \centering
 \begin{minipage}{0.045\textwidth}
    \footnotesize
    \raggedleft
    \vspace{-0.4 cm} SIMSE\\[0.25 cm]
    L1\\[0.26 cm]
    JTFS\\[0.26 cm]
    DTW
\end{minipage}%
\begin{minipage}{\textwidth}
    \centering
    \subfloat[Likert Scores]{
        \includegraphics[width=0.3\textwidth]{images/npsk_likert_0.png}
        \label{fig:p0_hearing}
    }\hspace{-0.425cm}  % Reduce space between figures
    \subfloat[Inverse P-Loss]{
        \includegraphics[width=0.3\textwidth]{images/npsk_P-Loss_0.png}
        \label{fig:p0_P-Loss}
    }\hspace{-0.5cm}  % Reduce space between figures
    \subfloat[Inverse MSS]{
        \includegraphics[width=0.3\textwidth]{images/npsk_MSS_0.png}
        \label{fig:p0_MSS}
    }
\end{minipage}
    \caption{\BPNoise~bootstrapped distributions and ranks given by NPSK. Ranks and their corresponding colors from best to worst are \colorbox{rank1}{\textcolor{black}{\textbf{1}}} \colorbox{rank2}{\textcolor{white}{\textbf{2}}} \colorbox{rank3}{\textcolor{white}{\textbf{3}}} \colorbox{rank4}{\textcolor{black}{\textcolor{white}{\textbf{4}}}}
}

    \label{fig:npsk_p0}
\end{figure}

\subsubsection{\AddSineSaw}
As shown in Figure~\ref{fig:npsk_p1}, JTFS was the best performing loss function in all evaluation methods. Moreover, all evaluation methods produced identical results. 
\begin{figure}[htbp]
    % \centering
 \begin{minipage}{0.045\textwidth}
    \footnotesize
    \raggedleft
    \vspace{-0.4 cm} SIMSE\\[0.25 cm]
    L1\\[0.26 cm]
    JTFS\\[0.26 cm]
    DTW
\end{minipage}%
\begin{minipage}{\textwidth}
    \centering
    \subfloat[Likert Scores]{
        \includegraphics[width=0.3\textwidth]{images/npsk_likert_1.png}
        \label{fig:p1_hearing}
    }\hspace{-0.425cm}  % Reduce space between figures
    \subfloat[Inverse P-Loss]{
        \includegraphics[width=0.3\textwidth]{images/npsk_P-Loss_1.png}
        \label{fig:p1_P-Loss}
    }\hspace{-0.5cm}  % Reduce space between figures
    \subfloat[Inverse MSS]{
        \includegraphics[width=0.3\textwidth]{images/npsk_MSS_1.png}
        \label{fig:p1_MSS}
    }
\end{minipage}
    \caption{\AddSineSaw~bootstrapped distributions and ranks given by NPSK. 
}
    \label{fig:npsk_p1}
\end{figure}

\subsubsection{\AmpMod}
IP-Loss and manual hearing tests yield identical rankings, and MSS results vary for for ranks 2 to 4.  It is not surprising that MSS results differ from the hearing tests, as we saw in Table~\ref{tab:kruskal_auto}, MSS showed no significant differences between groups. As shown in Figure~\ref{fig:npsk_p2}, all models selected \DTWEnv~as the best performer. This may be due to \DTWEnv's focus on periodic changes in loudness.
\begin{figure}[H]
    % \centering
 \begin{minipage}{0.045\textwidth}
    \footnotesize
    \raggedleft
    \vspace{-0.4 cm} SIMSE\\[0.25 cm]
    L1\\[0.26 cm]
    JTFS\\[0.26 cm]
    DTW
\end{minipage}%
\begin{minipage}{\textwidth}
    \centering
    \subfloat[Likert Scores]{
        \includegraphics[width=0.3\textwidth]{images/npsk_likert_2.png}
        \label{fig:p2_hearing}
    }\hspace{-0.425cm}  % Reduce space between figures
    \subfloat[Inverse P-Loss]{
        \includegraphics[width=0.3\textwidth]{images/npsk_P-Loss_2.png}
        \label{fig:p2_P-Loss}
    }\hspace{-0.5cm}  % Reduce space between figures
    \subfloat[Inverse MSS]{
        \includegraphics[width=0.3\textwidth]{images/npsk_MSS_2.png}
        \label{fig:p2_MSS}
    }
\end{minipage}
    \caption{\AmpMod~bootstrapped distributions and ranks given by NPSK. 
}
    \label{fig:npsk_p2}
\end{figure}

\subsubsection{\FMMod}
\DTWEnv~is again the best performer here. As shown in Figure~\ref{fig:npsk_p3}, all methods of analysis gave identical rankings to all programs.

\begin{figure}[H]
    % \centering
 \begin{minipage}{0.045\textwidth}
    \footnotesize
    \raggedleft
    \vspace{-0.4 cm} SIMSE\\[0.25 cm]
    L1\\[0.26 cm]
    JTFS\\[0.26 cm]
    DTW
\end{minipage}%
\begin{minipage}{\textwidth}
    \centering
    \subfloat[Likert Scores]{
        \includegraphics[width=0.3\textwidth]{images/npsk_likert_3.png}
        \label{fig:p3_hearing}
    }\hspace{-0.425cm}  % Reduce space between figures
    \subfloat[Inverse P-Loss]{
        \includegraphics[width=0.3\textwidth]{images/npsk_P-Loss_3.png}
        \label{fig:p3_P-Loss}
    }\hspace{-0.5cm}  % Reduce space between figures
    \subfloat[Inverse MSS]{
        \includegraphics[width=0.3\textwidth]{images/npsk_MSS_3.png}
        \label{fig:p3_MSS}
    }
\end{minipage}
    \caption{\FMMod~bootstrapped distributions and ranks given by NPSK. 
}
    \label{fig:npsk_p3}
\end{figure}

\subsection{Consistency In Rankings}

As we've discussed, automatic sound evaluation methods are unreliable, as we do not know how they would align with the often subjective opinion of human listeners. Here, the manual blinded rankings given by the reviewers could be treated as a ground truth, particularly since the two reviewers gave very similar rankings to the experiment results.

To measure the agreement between the two listeners, Spearman's rank correlation is used. The Spearman test returns a p-value for the null hypothesis of no correlation, along with a $\rho$ value between -1 to 1; -1 for perfect negative correlation, 1 for perfect correlation, and 0 for no correlation~\cite{spearman1987proof,rebekic2015pearson}. For each program and loss function, the blind manual hearing test gives us 80 ranks, which are given by two of the authors to a set of 40 randomly selected experiment results. This gives 160 data points per program, and 640 data points for all four programs. We computed Spearman correlations between the two reviewers. Across all programs, the correlation was very strong ($\rho = 0.86$, $p < 10^{-180}$). Per-program correlations were similarly strong and statistically significant: $\rho = 0.71$ for \BPNoise{} ($p < 10^{-25}$), $\rho = 0.64$ for \AddSineSaw{} ($p < 10^{-19}$), $\rho = 0.84$ for \AmpMod{} ($p < 10^{-43}$), and $\rho = 0.85$ for \FMMod{} ($p < 10^{-44}$).

Tables~\ref{tab:p0},~\ref{tab:p1},~\ref{tab:p2}, and~\ref{tab:p3} show the rankings for each program, with ranks different from the corresponding hearing test results marked with an asterisk. Both MSS and P-Loss gave a different rank from the hearing results in 3 out of 16 cases, which---somewhat surprisingly---shows consistency between automatic and manual hearing tests, at least for the simple programs used in this work. We also observe that identical top ranks were consistent across performance evaluation methodologies, with the exception of P-Loss in \BPNoise, which narrowly picks a different spectrogram-based loss function.
\input{extra_figures/ranking_tables}
\subsection{Takeaways}
Perhaps the most important takeaway from our analysis is that the loss function that yields the best sound-matching outcomes varies depending on the synthesizer program. This is perhaps due to the interaction between how the parameters of the synthesizer influence the sound, and the core sonic features of the loss function.

We also see that the novel \SIMSESpec{} and \DTWEnv{} loss functions (to our knowledge, not previously used in iterative sound-matching), lead to the best results in three of the four programs. This shows room for further creative approaches to differentiable sound-similarity measures, particularly since we observed that SOTA loss functions can be ineffective depending on the design of a synthesizer.

Finally, we surprisingly see somewhat consistent results between the rankings assigned by manual hearing tests (which we take as the ground truth), and automatic measures of P-Loss and MSS. Due to the simplicity of our programs and the occasional deviations, we cannot conclude that these automatic measures are substitutes for actual hearing.

\section{Summary and Conclusion}
\label{sec:summary_conclusion}
 Sound-matching is an umbrella term for the algorithmic programming of audio synthesizers, often with the goal of assisting sound designers. Here we provided a history of sound-matching, and proposed ``differentiable iterative sound-matching'' as a natural extension of previous works and recent advances.

We exposed several areas of weakness in the field, for which we presented methods for bridging the knowledge gaps. These issues included loss and synthesizer selection, periodic loss landscapes, and verification of performance measures with in-domain and out-of-domain sounds. The first two issues being the main target of the experiments in this work.

Using simple synthesizers, small sound domains, and a limited number of loss functions, previous works have been able to claim some loss functions or feature extractors for audio as ``better performing" for sound-matching (or other audio-related tasks)\cite{vahidi2023mesostructures,turian2020sorry,masuda2021soundmatch,engel2020ddsp,han2023perceptual}. Is the search for a universally best-performing loss function a worthy endeavor, or perhaps the choice of loss function is domain dependent?

The main hypothesis tested here is whether the performance of differentiable loss functions is influenced by the synthesis techniques used for sound-matching. We created iterative sound-matching experiments by combining four different loss functions with four different sound synthesis programs. We ranked the performance of the iterative sound-matching pipelines for every loss function and program, and observed that the success of the pipeline (that is, how closely the output sound matches the target sound) is program dependent. In other words, different synthesizer programs work best with different loss functions.

 We considered P-Loss and MSS as performance measures, commonly used in other sound-matching works, and compared their rankings to those given by blind hearing tests. We observed that automatic performance measures and manual listening tests were often (though not always) in agreement. Manual verification of sound-matching results remains a necessity for future works.

\section{Weaknesses}
% synthesizer simplicity
% arbitrary parameter selection
Compared to previous work, we presented a relatively more cohesive approach to iterative sound-matching which utilizes and compares four loss functions and four synthesis methods. However, there are many other untested methods of synthesis that can be combined in practically infinite ways.

To limit the scope of our experiments, we had no choice but to use arbitrary parameters for the various signal processing functions. We used bare-bones versions of STFT and JTFS with fixed parameters. We did not test complex synthesizers, which may use different signal generation functions in parallel or sequentially. Arbitrary hyperparameters such as learning rate and max number of iterations were selected for the DL pipeline, and only the RMSProp optimizer was tested.

In Section~\ref{sec:in-domain}, we discussed the importance of sound domains. This work, much like previous work, only utilizes in-domain sounds. That is, the target sound is made by the synthesizer, and the parameters are already known. This simplifies the issues of measuring sound-similarity, but it is not a realistic scenario for practical sound-matching. We plan to address this issue in future works, as we will discuss in the next section.

\section{Future Work}
\label{sec:future}
The analysis of different loss functions sound-matching presented here shows that problem of periodic loss-landscapes is not limited to spectral representations using Fourier transforms~\cite{turian2020sorry,vahidi2023mesostructures,uzrad2024diffmoog,bruford2024synthesizer}. We can postulate that this problem emerges due to the periodic nature of sound. The current trend in contemporary work appears to be the application of domain specific and computationally expensive loss functions~\cite{han2023perceptual,uzrad2024diffmoog}, use of large neural networks~\cite{hershey2017cnn,cramer2019look}, or complex ensemble methods~\cite{turian2022hear}. Such models are useful, but often intractable; furthermore, training such models still requires the definition of simpler loss functions, which emphasizes the need for further work on differentiable and expressive loss function implementations.

With regard to navigation of loss function landscapes, an optimizer that is more aware of fluctuations in gradients would perhaps lead to better solutions than simple gradient descent. Viewing the sound-matching problem as ``the navigation of an agent from an arbitrary point in a gradient field to a target'' closely resembles many classical problems in the field of reinforcement learning (RL)~\cite{sutton2018reinforcement}. Naturally, we can extend our work by applying RL and other heuristic search techniques to the problem of sound-matching, and measuring the performance changes (both quantitative and qualitative) relative to simple gradient descent updates.

Like nearly all previous work in sound-matching, the main measure of success here was the accurate \textit{replication} of sounds (or synth parameters) rather than imitation of sounds outside of the synthesizer's domain. Replication of arbitrary features of sounds is an important component of sound-design, though it is much harder to define or measure. Future works could explore loss functions which only measure certain characteristics of sound (such as \DTWEnv), and whether they can pave the way for better imitation sound-matching.

\clearpage
\bibliographystyle{alpha}
\bibliography{references}

% Include the appendix from the separate file
\include{appendix}
\end{document}

### 2025-04-24 main.tex
Updates from Overleaf

@@
-\documentclass[lettersize,journal]{IEEEtran}% \documentclass[twocolumn,10pt]{article}
-\usepackage[utf8]{inputenc}
-\usepackage{natbib}
+\documentclass[lettersize,journal]{IEEEtran}
+\usepackage{amsmath,amsfonts}
+\usepackage{algorithmic}
+\usepackage{algorithm}
+\usepackage{array}
+\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
+\usepackage{textcomp}
+\usepackage{stfloats}
+\usepackage{url}
+\usepackage{verbatim}
 \usepackage{graphicx}
-\usepackage{stfloats} % for positioning of figure* on the same page
-\usepackage{caption}
-\usepackage{tikz}
-\usepackage[inline]{enumitem}
-\usepackage{amsmath}
-\usepackage{subcaption}
-\usepackage{graphicx}
-\usepackage[breaklinks=true,colorlinks=true, allcolors=blue]{hyperref}
-\usepackage{breakcites}
-\usepackage{microtype}
-\usepackage{lipsum}
-\usepackage{xcolor}
-\usepackage{array}
-\usepackage{float}
-\usepackage{adjustbox}
-\usepackage{listings}
 \usepackage{xcolor}
@@
 \usepackage{makecell}
-\usepackage{pdfpages}
+% my added packages and commands =======================
+\usepackage[numbers]{natbib} %removed cite package
+\usepackage{tikz}
+\usepackage{adjustbox}
+\usepackage{listings}
+
+\usepackage{caption}               % re-enable caption functionality
+\captionsetup[lstlisting]{justification=centering, singlelinecheck=false}
 
 \usetikzlibrary{positioning,shapes.geometric}
-\usepackage[acronym,toc]{glossaries}
-
-% \input{glossary.tex}
-
-% Initialize the glossary
-\makeglossaries
-
+
+\providecommand{\gls}[1]{#1}
 \definecolor{highlight}{RGB}{00,150,00}
@@
 \newcommand{\BPNoise}{\textbf{BP-Noise}}  
-\newcommand{\AddSineSaw}{\textbf{Additive-SineSaw}}  
+\newcommand{\AddSineSaw}{\textbf{Add-SineSaw}}  
 \newcommand{\AmpMod}{\textbf{Noise-AM}}  
@@
 \input{code_style}
+%  ================================
+
+\begin{document}
 
 \title{Sound Similarity, Sound Matching, and Differentiable Audio Synthesizer Control}
-\author{ }
-
-\begin{document}
-
-\printglossary
-\clearpage
-
+
+\author{-}
 \maketitle
+
 \begin{abstract}
@@
 Past works in sound-matching often relied on a narrow range of synthesizers and overlooked the possible interactions between the loss function and synthesis parameters.
-% There are critical choices which need to be made in sound-matching: these are the choice of synthesizer, algorithms for search and optimization, the source of sounds, and similarity (or loss) functions.
-We categorize sound-matching works based on these choices, and point out a number of unresolved issues and interesting yet unexplored areas.
 We analyzed the interplay between four simple subtractive, additive, and FM synthesizers when guided by four different novel or established similarity functions.
@@
 
-%  Like other problems in computing science, there appears to be an ever present drive towards finding optimal solutions to the problem, and outperforming state-of-the-art.
+\begin{IEEEkeywords}
+Machine Learning, Signal Processing, Sound-Matching
+\end{IEEEkeywords}
 
@@
 \subsection{Digital Synthesis With Sound-Matching}
-Since the 1970s, artists such as Suzanne Ciani~\cite{ciani_life_in_waves} have explored the use of audio synthesizers to replicate natural sounds---such as a bottle cap popping off---in a way that preserves the essential qualities of the original while showcasing the characteristics of the synthesizer.
+Since the 1970s, artists such as Suzanne Ciani~\cite{ciani_life_in_waves} have explored the use of audio synthesizers to replicate natural sounds—such as a bottle cap popping off—in a way that preserves the essential qualities of the original while showcasing the characteristics of the synthesizer.
 This approach goes beyond mere replication and uses synthesizers for artistic re-interpretation of sounds.
@@
 Esling \textit{et al.} also noted the issue of using P-Loss in sound-matching~\cite{esling2019flow}, considering the non-linear mapping between parameters of a synth and the audio output.
-In their approach, dubbed as \textquote{flowsynth}, they used a large parameter-to-audio dataset for a commercial \gls{VST} synthesizer.
+In their approach, dubbed as ``flowsynth'', they used a large parameter-to-audio dataset for a commercial \gls{VST} synthesizer.
 In flowsynth, audio is encoded to an auditory space, using a NN autoencoding approach~\cite{hinton2006reducing}.
@@
 In the following subsections, we describe the programs and their parameters in more detail, followed by the methodology in Section~\ref{sec:methodology} and results in Section~\ref{sec:results_posthoc}.
+
 \begin{table*}[htbp]
@@
 \label{sec:program0}
-\begin{lstlisting}[caption={\BPNoise}, label={lst:program0},language=Faust,float,floatplacement=htbp,xleftmargin=2em,xrightmargin=1em,firstnumber=0]
+
+\begin{lstlisting}[caption={\BPNoise}, label={lst:program0}, language=Faust,
+                  float, floatplacement=htbp, xleftmargin=1em, xrightmargin=0.5em, firstnumber=0]
 import("stdfaust.lib");
@@
 Listing~\ref{lst:program0} shows the Faust code for \BPNoise.
-
 \subsubsection{\AddSineSaw}
@@
 
-\begin{lstlisting}[caption={\AddSineSaw}, label={lst:program1},language=Faust,float,floatplacement=TBH,xleftmargin=2em,xrightmargin=1em,firstnumber=0]
+\begin{lstlisting}[caption={\AddSineSaw}, label={lst:program1},language=Faust,float,floatplacement=TBH,xleftmargin=1em,xrightmargin=0.5em,firstnumber=0]
 import("stdfaust.lib");
@@
 \label{sec:program2}
-\begin{lstlisting}[caption={\AmpMod}, label={lst:program2},language=Faust,float,floatplacement=TBH,xleftmargin=2em,xrightmargin=1em,firstnumber=0]
+\begin{lstlisting}[caption={\AmpMod}, label={lst:program2},language=Faust,float,floatplacement=TBH,xleftmargin=1em,xrightmargin=0.5em,firstnumber=0]
 import("stdfaust.lib");
@@
 \label{sec:program3}
-\begin{lstlisting}[caption={\FMMod}, label={lst:program3},language=Faust,float,floatplacement=TBH,xleftmargin=2em,xrightmargin=1em,firstnumber=0]
+\begin{lstlisting}[caption={\FMMod}, label={lst:program3},language=Faust,float,floatplacement=TBH,xleftmargin=1em,xrightmargin=0.5em,firstnumber=0]
 import("stdfaust.lib");
@@
 As shown in Figure~\ref{fig:npsk_p0}, manual hearing test and MSS selected \SIMSESpec~as the best performer and \LoneSpec~as the second-best performer, while P-loss gave the reverse order.
-\begin{figure}[htbp]
-    % \centering
- \begin{minipage}{0.045\textwidth}
+
+\begin{figure*}[htbp]
+    \centering
+    \begin{minipage}{0.03\textwidth}
         \footnotesize
         \raggedleft
-    \vspace{-0.4 cm} SIMSE\\[0.25 cm]
-    L1\\[0.26 cm]
-    JTFS\\[0.26 cm]
+        \vspace{-.2cm}
+        SIMSE\\[0.6cm]
+        L1\\[0.65cm]
+        JTFS\\[0.65cm]
         DTW
     \end{minipage}%
-\begin{minipage}{\textwidth}
+    \begin{minipage}{0.95\textwidth}
         \centering
@@
             \label{fig:p0_hearing}
-    }\hspace{-0.425cm}  % Reduce space between figures
+        }\hspace{0.005\textwidth} % Reduced space between figures
         \subfloat[Inverse P-Loss]{
@@
             \label{fig:p0_P-Loss}
-    }\hspace{-0.5cm}  % Reduce space between figures
+        }\hspace{0.005\textwidth} % Reduced space between figures
         \subfloat[Inverse MSS]{
@@
     \end{minipage}
+    \vspace{-0.5em}
     \caption{\BPNoise~bootstrapped distributions and ranks given by NPSK.
-Ranks and their corresponding colors from best to worst are \colorbox{rank1}{\textcolor{black}{\textbf{1}}} \colorbox{rank2}{\textcolor{white}{\textbf{2}}} \colorbox{rank3}{\textcolor{white}{\textbf{3}}} \colorbox{rank4}{\textcolor{black}{\textcolor{white}{\textbf{4}}}}
-}
-
+Ranks and their corresponding colors from best to worst are \colorbox{rank1}{\textcolor{black}{\textbf{1}}}, \colorbox{rank2}{\textcolor{white}{\textbf{2}}}, \colorbox{rank3}{\textcolor{white}{\textbf{3}}}, \colorbox{rank4}{\textcolor{white}{\textbf{4}}}.}
     \label{fig:npsk_p0}
-\end{figure}
-
-\subsubsection{\AddSineSaw}
-As shown in Figure~\ref{fig:npsk_p1}, JTFS was the best performing loss function in all evaluation methods. Moreover, all evaluation methods produced identical results. 
-\begin{figure}[htbp]
-    % \centering
- \begin{minipage}{0.045\textwidth}
+\end{figure*}
+\begin{figure*}[htbp]
+    \centering
+    \begin{minipage}{0.03\textwidth}
         \footnotesize
         \raggedleft
-    \vspace{-0.4 cm} SIMSE\\[0.25 cm]
-    L1\\[0.26 cm]
-    JTFS\\[0.26 cm]
+        \vspace{-.2cm}
+        SIMSE\\[0.6cm]
+        L1\\[0.65cm]
+        JTFS\\[0.65cm]
         DTW
     \end{minipage}%
-\begin{minipage}{\textwidth}
+    \begin{minipage}{0.95\textwidth}
         \centering
@@
             \label{fig:p1_hearing}
-    }\hspace{-0.425cm}  % Reduce space between figures
+        }\hspace{0.005\textwidth} % Reduced space between figures
         \subfloat[Inverse P-Loss]{
@@
             \label{fig:p1_P-Loss}
-    }\hspace{-0.5cm}  % Reduce space between figures
+        }\hspace{0.005\textwidth} % Reduced space between figures
         \subfloat[Inverse MSS]{
@@
     \end{minipage}
-    \caption{\AddSineSaw~bootstrapped distributions and ranks given by NPSK. 
+    \vspace{-0.5em}
+    \caption{\AddSineSaw~bootstrapped distributions and ranks given by NPSK.}
+    \label{fig:npsk_p1}
+\end{figure*}
+\begin{figure*}[htbp]
+    \centering
+    \begin{minipage}{0.03\textwidth}
+        \footnotesize
+        \raggedleft
+        \vspace{-.2cm}
+        SIMSE\\[0.6cm]
+        L1\\[0.65cm]
+        JTFS\\[0.65cm]
+        DTW
+    \end{minipage}%
+    \begin{minipage}{0.95\textwidth}
+        \centering
+        \subfloat[Likert Scores]{
+            \includegraphics[width=0.3\textwidth]{images/npsk_likert_2.png}
+            \label{fig:p2_hearing}
+        }\hspace{0.005\textwidth} % Reduced space between figures
+        \subfloat[Inverse P-Loss]{
+            \includegraphics[width=0.3\textwidth]{images/npsk_P-Loss_2.png}
+            \label{fig:p2_P-Loss}
+        }\hspace{0.005\textwidth} % Reduced space between figures
+        \subfloat[Inverse MSS]{
+            \includegraphics[width=0.3\textwidth]{images/npsk_MSS_2.png}
+            \label{fig:p2_MSS}
         }
-    \label{fig:npsk_p1}
-\end{figure}
+    \end{minipage}
+    \vspace{-0.5em}
+    \caption{\AmpMod~bootstrapped distributions and ranks given by NPSK.}
+    \label{fig:npsk_p2}
+\end{figure*}
+\begin{figure*}[htbp]
+    \centering
+    \begin{minipage}{0.03\textwidth}
+        \footnotesize
+        \raggedleft
+        \vspace{-.2cm}
+        SIMSE\\[0.6cm]
+        L1\\[0.65cm]
+        JTFS\\[0.65cm]
+        DTW
+    \end{minipage}%
+    \begin{minipage}{0.95\textwidth}
+        \centering
+        \subfloat[Likert Scores]{
+            \includegraphics[width=0.3\textwidth]{images/npsk_likert_3.png}
+            \label{fig:p3_hearing}
+        }\hspace{0.005\textwidth} % Reduced space between figures
+        \subfloat[Inverse P-Loss]{
+            \includegraphics[width=0.3\textwidth]{images/npsk_P-Loss_3.png}
+            \label{fig:p3_P-Loss}
+        }\hspace{0.005\textwidth} % Reduced space between figures
+        \subfloat[Inverse MSS]{
+            \includegraphics[width=0.3\textwidth]{images/npsk_MSS_3.png}
+            \label{fig:p3_MSS}
+        }
+    \end{minipage}
+    \vspace{-0.5em}
+    \caption{\FMMod~bootstrapped distributions and ranks given by NPSK.}
+    \label{fig:npsk_p3}
+\end{figure*}
+
+\subsubsection{\AddSineSaw}
+As shown in Figure~\ref{fig:npsk_p1}, JTFS was the best performing loss function in all evaluation methods. Moreover, all evaluation methods produced identical results.
 
@@
 This may be due to \DTWEnv's focus on periodic changes in loudness.
-\begin{figure}[H]
-    % \centering
- \begin{minipage}{0.045\textwidth}
-    \footnotesize
-    \raggedleft
-    \vspace{-0.4 cm} SIMSE\\[0.25 cm]
-    L1\\[0.26 cm]
-    JTFS\\[0.26 cm]
-    DTW
-\end{minipage}%
-\begin{minipage}{\textwidth}
-    \centering
-    \subfloat[Likert Scores]{
-        \includegraphics[width=0.3\textwidth]{images/npsk_likert_2.png}
-        \label{fig:p2_hearing}
-    }\hspace{-0.425cm}  % Reduce space between figures
-    \subfloat[Inverse P-Loss]{
-        \includegraphics[width=0.3\textwidth]{images/npsk_P-Loss_2.png}
-        \label{fig:p2_P-Loss}
-    }\hspace{-0.5cm}  % Reduce space between figures
-    \subfloat[Inverse MSS]{
-        \includegraphics[width=0.3\textwidth]{images/npsk_MSS_2.png}
-        \label{fig:p2_MSS}
-    }
-\end{minipage}
-    \caption{\AmpMod~bootstrapped distributions and ranks given by NPSK. 
-}
-    \label{fig:npsk_p2}
-\end{figure}
 
@@
 \DTWEnv~is again the best performer here. As shown in Figure~\ref{fig:npsk_p3}, all methods of analysis gave identical rankings to all programs.
-
-\begin{figure}[H]
-    % \centering
- \begin{minipage}{0.045\textwidth}
-    \footnotesize
-    \raggedleft
-    \vspace{-0.4 cm} SIMSE\\[0.25 cm]
-    L1\\[0.26 cm]
-    JTFS\\[0.26 cm]
-    DTW
-\end{minipage}%
-\begin{minipage}{\textwidth}
-    \centering
-    \subfloat[Likert Scores]{
-        \includegraphics[width=0.3\textwidth]{images/npsk_likert_3.png}
-        \label{fig:p3_hearing}
-    }\hspace{-0.425cm}  % Reduce space between figures
-    \subfloat[Inverse P-Loss]{
-        \includegraphics[width=0.3\textwidth]{images/npsk_P-Loss_3.png}
-        \label{fig:p3_P-Loss}
-    }\hspace{-0.5cm}  % Reduce space between figures
-    \subfloat[Inverse MSS]{
-        \includegraphics[width=0.3\textwidth]{images/npsk_MSS_3.png}
-        \label{fig:p3_MSS}
-    }
-\end{minipage}
-    \caption{\FMMod~bootstrapped distributions and ranks given by NPSK. 
-}
-    \label{fig:npsk_p3}
-\end{figure}
 
@@
 
-\clearpage
-\bibliographystyle{alpha}
+\bibliographystyle{IEEEtran}
 \bibliography{references}
-
-% Include the appendix from the separate file
-\include{appendix}
 \end{document}

### 2025-04-24 main.tex
Updates from Overleaf

@@
 \section{Weaknesses}
-% synthesizer simplicity
-% arbitrary parameter selection
 Compared to previous work, we presented a relatively more cohesive approach to iterative sound-matching which utilizes and compares four loss functions and four synthesis methods.

### 2025-04-25 main.tex
Updates from Overleaf

@@
 \input{code_style}
+
 %  ================================
-
 \begin{document}
-
 \title{Sound Similarity, Sound Matching, and Differentiable Audio Synthesizer Control}
@@
 Iterative sound-matching is the automation of this process, more formally, the automatic programming of an audio synthesizer to generate outputs similar to a target sound while guided by a loss function that quantifies sound similarity.
-Past works in sound-matching often relied on a narrow range of synthesizers and overlooked the possible interactions between the loss function and synthesis parameters.
+Past works in sound-matching often relied on a narrow range of synthesizers and overlooked the possible interactions between the loss function and synthesizer parameters.
 We analyzed the interplay between four simple subtractive, additive, and FM synthesizers when guided by four different novel or established similarity functions.
@@
 This approach goes beyond mere replication and uses synthesizers for artistic re-interpretation of sounds.
-Inspired by this, we aim to emulate and automate this process using digital audio synthesizers.
+Pursuing this idea, we aim to emulate and automate this process using digital audio synthesizers.
 
@@
 where $T$ represents time in seconds, and $n$ is the frequency parameter.
-If $n$ is 1, then $x(n)$ would be a waveform with a frequency of 1 hertz as the value of $x(n)$ would oscillate once per second.
+If $n$ is 1, then $x(n)$ would be a waveform with a frequency of 1 hertz, meaning that the value of $x(n)$ oscillates to the same point once every second.
 Waveform generators are called oscillators for this reason.
@@
 After training, PNP can be used as a fast approximation of the loss function, $\|\phi(t) - \phi(x)\|_2^2 $.
-Han \textit{et al.} observed slight improvements to run-time (for testing) and minor improvements to the loss values under certain conditions.
+Han \textit{et al.} observed improvements to run-time (for testing) and slight reduction of the loss values under certain conditions.
 These experiments were based on supervised learning and physical modeling (of a drum head) for synthesis~\cite{smith2010physical}, and compared PNP, P-Loss, MSS, and JTFS functions~\cite{han2023perceptual}.
-This approach was more recently applied to other supervised settings using physical and FM synthesizers~\cite{han2024learning}.
+This approach was more recently applied to other supervised settings using physical and FM synthesizers~\cite{han2024learning}\textcolor{red}{elaborate}.
 

### 2025-04-29 main.tex
Updates from Overleaf

@@
 Manual sound design with a synthesizer is an iterative process.
-When an artist has a target sound in mind, they listen to the output of their chosen synthesizer, determine whether the output is closer or farther from the target, and adjust the synthesizer's parameters until satisfied.
+When an artist has a target sound in mind, they listen to the output of their chosen synthesizer, determine whether the output is closer or farther from the target, and adjust the parameters until satisfied.
 There are practically an infinite number of parameter combinations, and their effect on the output sound is hard to predict.
@@
 This approach goes beyond mere replication and uses synthesizers for artistic re-interpretation of sounds.
-Pursuing this idea, we aim to emulate and automate this process using digital audio synthesizers.
+Pursuing this idea, we aim to automate this process using digital audio synthesizers.
 
@@
 Automatic sound design has been a subject of research for decades, with reduction of tinkering time and creation of ``interesting" sounds as the main motivators \cite{krekovic2019insights,turian2020sorry,horner1993machine,salimi2020make,esling2019flow,engel2020ddsp,mitchell2007evolutionary,shier2020spiegelib}.
-\textbf{Sound-matching} is a common form of automatic sound design explored in literature, where the goal is to find parameters for a synthesizer that replicate \textit{all} or \textit{some} of the characteristics of a target sound as best as possible, given some kind of similarity metric~\cite{horner1993machine,mitchell2007evolutionary}.
-Not all sound-matching works follow the iterative process of manual sound design.
-\textit{Iterative} sound-matching refers to works which mimic the back and forth process of listening to a synthesizer's output and updating the parameters until the termination conditions are met.
-
-A brief description of the iterative sound-matching workflow may be useful here: The two necessary components for sound-matching are the \textit{loss function}, which determines how close two sounds are, and the \textit{parametric synthesizer} that outputs sounds based on input parameters.
+\textbf{Sound-matching} is a common form of automatic sound design explored in literature, where the goal is to find parameters for a synthesizer that replicate \textit{all} or \textit{some} of the characteristics of a target sound as best as possible, given some kind of similarity metric (or loss function)~\cite{horner1993machine,mitchell2007evolutionary}.
+
+\textit{Iterative} sound-matching refers to works that mimic the back-and-forth process of listening to a synthesizer's output and updating the parameters until the termination conditions are met.
+The two necessary components are the \textit{loss function}, which determines how close two sounds are, and the \textit{parametric synthesizer} that outputs sounds based on input parameters.
 In the iterative workflow, a target sound is selected, the synthesizer outputs sounds based on its parameters, and the loss function determines the proximity of the output to the target.
@@
 
-With the goal of implementing effective iterative sound-matching algorithms for sound designers, we identify a number of weaknesses in current relevant literature (see Section~\ref{sec:lacking}).
+Despite the early popularity of the iterative approach~\cite{justice1979analytic,horner1993machine,mitchell2007evolutionary}, recent works in sound-matching have favored inference models, deep neural-networks (\gls{DNN}), and large datasets.
+Rather than continually interacting with the environment, DNN-based models learn from a dataset of inputs (synthesizer parameters) and outputs (corresponding sound), and predict the synthesizer parameters for new sounds~\cite{yee2018automatic,han2023perceptual,han2024learning}.
+
+We believe that iterative sound-matching is interesting for multiple reasons.
+First, it better resembles the human approach to sound-design, which requires continually interacting with the environment and making decisions according to a feedback signal~\cite{sutton2018reinforcement}.
+Second, the trial-and-error approach leaves a trail of parameter sets from the starting point to the end goal, which could be of use to sound designers.
+Third, relative to the DNN-based inference models, it provides a more tractable environment for exploring the complex issue of sound-similarity functions, and their interactions with different methods of synthesis.
+
+With the goal of improving iterative sound-matching algorithms for sound designers, we identify a number of weaknesses in current literature (see Section~\ref{sec:lacking}).
 The experiments in this work pertain to two core issues, which we named \LossSelect~and \SynthSelect.
-The former refers to the search for an optimal loss function, lack of novel algorithms, and the focus on outperforming state of the art (\gls{SOTA}), while the latter refers to the lack of diversity in synthesis methods, and their interaction with the loss function.
+The former refers to the search for an optimal loss function, lack of novel algorithms, and the---perhaps needless---focus on outperforming state of the art (\gls{SOTA}), while the latter refers to the lack of diversity in synthesis methods, and their interaction with the loss function.
 
@@
 Waveform generators are called oscillators for this reason.
-If $n$ is in the range of 30 to 14,000 (the upper limit can be higher for some), $x(n)$ would be an audible tone~\cite{smith2007mathematics,jilek2014reference}.
+If $n$ is in the range of 30 to 20,000 (the upper limit is often lower for some), $x(n)$ could be played as an audible tone~\cite{smith2007mathematics,jilek2014reference}.
 
@@
 Differential DSP (\gls{DDSP}) is a general term for applications that combine machine learning techniques such as gradient descent~\cite{goodfellow2016deep,boyd2004convex} with DSP~\cite{engel2020ddsp}.
-Here, we take a differential approach to sound-matching using DDSP, i.e., the use of differentiable loss functions and gradient descent to update synthesizer parameters to reduce the difference between the output and the target sound.
-
 Implementing complex DSP functions in a differentiable manner can be challenging, and effective differentiable audio similarity measures require careful mathematical expression of the desired attributes of sound.
 These issues have likely contributed to the limited exploration of this domain.
-Perhaps the most relevant recent work is the comparison of two differentiable loss functions with a simple FM synthesizer by Vahidi \textit{et al.}~\cite{vahidi2023mesostructures}.
+Perhaps to most relevant recent work is by Vahidi \textit{et al.}, featuring the comparison of two differentiable loss functions with a simple FM synthesizer~\cite{vahidi2023mesostructures}.
 
@@
 It is important to note that there is a close relationship between the loss function $L$ and the sound representation function $\phi$.
-A distance measure $d$ is applied to the features extracted by $\phi$.
+$L$ is the result of a distance measure $d$ applied to the features extracted by $\phi$.
 
@@
 \label{sec:sound_matching_definition}
-In past works the sound-matching problem has been tackled from a variety of perspectives as it involves a number of modular parts that need to be considered: the choice of synthesizer $g$, the target sounds of interest, the representation function $\phi$, and the heuristic for finding the optimal $\theta$.
+In past works the sound-matching problem has been tackled from a variety of perspectives as it involves a number of modular parts: the choice of synthesizer $g$, the target sounds of interest, the representation function $\phi$, and the heuristic for finding the optimal $\theta$.
 We discuss these components in the upcoming sections, and give a coherent typology of the sound-matching tasks in Section~\ref{sec:matching_types}.

### 2025-04-29 extra_figures/evaluation_graph.tex
Updates from Overleaf

@@
-\begin{figure}
-\resizebox{\textwidth}{!}{ % Automatically scales diagram to fit page width
+\begin{figure*}
+\resizebox{\linewidth}{!}{ % Automatically scales diagram to fit page width
 \begin{tikzpicture}[
-    function/.style={rectangle, draw, rounded corners, minimum width=2cm, minimum height=1cm, text centered, node distance=1.5cm},
-    distribution/.style={ellipse, draw, minimum width=2cm, minimum height=1cm, text centered, node distance=1.2cm},
+    function/.style={rectangle, draw, rounded corners, minimum width=2cm, minimum height=1cm, text centered, node distance=1cm},
+    distribution/.style={ellipse, draw, minimum width=1cm, minimum height=0cm, text centered, node distance=0.3cm},
     gaussian/.style={draw, smooth, samples=100, domain=-2.5:2.5},
     snpk/.style={rectangle, draw, rounded corners, minimum width=2.5cm, minimum height=1cm, text centered, node distance=1.5cm, fill=blue!20},
-    level 1/.style={sibling distance=6cm, level distance=2cm},
-    level 2/.style={sibling distance=2cm, level distance=3cm},
+    level 1/.style={sibling distance=6cm, level distance=1cm},
+    level 2/.style={sibling distance=2cm, level distance=1.2cm},
     mss/.style={fill=red!50, draw=black},
@@
 % Label for the second row (Distributions)
-\node at (0,-6.5) {\textbf{$\longleftarrow$\ \ \large Bootstrapped Distributions\ \ $\longrightarrow$}};
+\node at (0,-4) {\textbf{$\longleftarrow$\ \ \large Bootstrapped Distributions\ \ $\longrightarrow$}};
+% Left side: SNPK with MSS
 
 % Left side: SNPK with MSS
-\node[snpk, text height=5ex, align=center, fill=red!30, draw=black] at (-6,-9) {SNPK with 4 \\ MSS distributions};
-\draw[->] (-6,-9.5) -- (-6,-10.5) node[below, below] {\textbf{Ranks 1-4}};
-\draw[red!40, thick] (-4.5, -8.4) -- (-2.75, -7.75);  % Line 1
-\draw[red!40, thick] (-5.25, -8.4) -- (-3.5, -7.75);  % Line 2
-\draw[red!40, thick] (-5.75, -8.4) -- (-5.5, -7.75);  % Line 3
-\draw[red!40, thick] (-6.25, -8.4) -- (-9, -7.75);  % Line 4
+\node[snpk, text height=5ex, align=center, fill=red!30, draw=black] at (-6,-6) {SNPK with 4 \\ MSS distributions};
+\draw[->] (-6,-6.5) -- (-6,-7.5) node[below, below] {\textbf{Ranks 1-4}};
+\draw[red!40, thick] (-4.5, -5.4) -- (-2.75, -4.75);  % Line 1
+\draw[red!40, thick] (-5.25, -5.4) -- (-3.5, -4.75);  % Line 2
+\draw[red!40, thick] (-5.75, -5.4) -- (-5.5, -4.75);  % Line 3
+\draw[red!40, thick] (-6.25, -5.4) -- (-9, -4.75);    % Line 4
 
 % Middle: SNPK with P-Loss
-\node[snpk, text height=5ex, align=center, fill=blue!40, draw=black] at (0,-9) {SNPK with 4 \\ P-Loss distributions};
-\draw[->] (0,-9.5) -- (0,-10.5) node[below, below] {\textbf{Ranks 1-4}};
-\draw[blue!40, thick] (1.5, -8.4) -- (3.25, -7.75);  % Line 1
-\draw[blue!40, thick] (0.75, -8.4) -- (1.5, -7.75);  % Line 2
-\draw[blue!40, thick] (-0.25, -8.4) -- (-1, -7.75);  % Line 3
-\draw[blue!40, thick] (-0.75, -8.4) -- (-3, -7.75);  % Line 4
+\node[snpk, text height=5ex, align=center, fill=blue!40, draw=black] at (0,-6) {SNPK with 4 \\ P-Loss distributions};
+\draw[->] (0,-6.5) -- (0,-7.5) node[below, below] {\textbf{Ranks 1-4}};
+\draw[blue!40, thick] (1.5, -5.4) -- (3.25, -4.75);   % Line 1
+\draw[blue!40, thick] (0.75, -5.4) -- (1.5, -4.75);   % Line 2
+\draw[blue!40, thick] (-0.25, -5.4) -- (-1, -4.75);   % Line 3
+\draw[blue!40, thick] (-0.75, -5.4) -- (-3, -4.75);   % Line 4
 
 % Right side: SNPK with Likert
-\node[snpk, text height=5ex, align=center, fill=green!30, draw=black] at (6,-9) {SNPK with 4 \\ Likert distributions};
-\draw[->] (6,-9.5) -- (6,-10.5) node[below, below] {\textbf{Ranks 1-4}};
-\draw[green!50, thick] (7.5, -8.4) -- (8.75, -7.75);  % Line 1
-\draw[green!50, thick] (6.75, -8.4) -- (6, -7.75);    % Line 2
-\draw[green!50, thick] (6.25, -8.4) -- (4, -7.75);   % Line 3
-\draw[green!50, thick] (5.75, -8.4) -- (3, -7.75);     % Line 4
+\node[snpk, text height=5ex, align=center, fill=green!30, draw=black] at (6,-6) {SNPK with 4 \\ Likert distributions};
+\draw[->] (6,-6.5) -- (6,-7.5) node[below, below] {\textbf{Ranks 1-4}};
+\draw[green!50, thick] (7.5, -5.4) -- (8.75, -4.75);  % Line 1
+\draw[green!50, thick] (6.75, -5.4) -- (6, -4.75);    % Line 2
+\draw[green!50, thick] (6.25, -5.4) -- (4, -4.75);    % Line 3
+\draw[green!50, thick] (5.75, -5.4) -- (3, -4.75);    % Line 4
 
@@
 \label{fig:posthoc_evaluation}
-\end{figure}
+\end{figure*}

### 2025-04-29 main.tex
Updates from Overleaf

@@
 
-We view the implementation of these components not as an optimization problem with a correct answer, but a \textit{creative} endeavor depending on the artist's needs.
+It could be argued that the implementation of these components is not an optimization problem with a correct answer, but a \textit{creative} endeavor depending on the artist's needs.
 Consider the case where the goal is to create an 8-bit version of a high quality, organic snare drum~\cite{collins2007loop}: in such a case, there is no optimal or ``correct" version of an 8-bit snare sound, and multiple correct answers exist.
@@
 
-From a signal perspective, replication is a scenario where the goal is to find a $\theta$ such that $t-x=0$ where $x=g(\theta)$.
+% From a signal perspective, replication is a scenario where the goal is to find a $\theta$ such that $t-x=0$ where $x=g(\theta)$.
 From a sound perspective, this definition of replication is unnecessarily restrictive, as $t$ and $x$ can sound identical but contain very different values.
@@
 
-With \textit{sound imitation}, the goal is the replication of some aspect of the target sound.
+% With \textit{sound imitation}, the goal is the replication of some aspect of the target sound.
 Instrument conversion is a common example of imitation.
@@
 Imitation of sound envelopes (how the loudness changes over time) is another interesting idea~\cite{caspe2023envelope}.
-Transfer of other sound properties such as reverb and delay has; also been a subject of research~\cite{engel2020ddsp}.
-
-A target sound can have many properties of interest, and the imitation task is the transfer of a subset of those properties to other sounds.
+Transfer of other sound properties such as reverb and delay has also been a subject of research~\cite{engel2020ddsp}.
+
+% A target sound can have many properties of interest, and the imitation task is the transfer of a subset of those properties to other sounds.
 Relative to replication, imitation is a very open ended goal.
@@
 \begin{enumerate}
-    \item Supervised methods which use datasets of sound and parameter pairs to describe the generation objective.
+    \item Supervised (or inference) methods which use datasets of sound and parameter pairs to describe the generation objective.
 Often with a large number of input and output examples~\cite{engel2020ddsp,salimi2020make,yee2018automatic,esling2019flow}
-    \item Unsupervised or \textit{direct sound optimization} methods~\cite{uzrad2024diffmoog}.
+    \item Unsupervised or \textit{direct sound optimization} methods~\cite{horner1993machine,mitchell2007evolutionary,yee2018automatic,vahidi2023mesostructures}.
 Usually require at most one target example, and a loss function that measures the difference between the target and synthesizer outputs.
@@
 
-Genetic algorithms (\gls{GA})~\cite{holland1992genetic} can be used for direct sound optimization, and have been the earliest and most common heuristic in sound-matching~\cite{horner1993machine,mitchell2007evolutionary,yee2018automatic,uzrad2024diffmoog}.
+Genetic algorithms (\gls{GA})~\cite{holland1992genetic} can be used for direct sound optimization, and have been the earliest and most common heuristic in sound-matching~\cite{horner1993machine,mitchell2007evolutionary,yee2018automatic}.
 These algorithms start with arbitrary parameter sets that can be treated as an evolving population where the genomes are the parameter values.
-The most fit members of the group are the parameters which perform the best in the loss function.
-The most fit parameters create a new generation of parameters via mutation (random change in subset of parameters) and crossovers (combination of parameter sets), and the process repeats until the goal or a maximum number of generations is reached.
-
-One important aspect of the heuristics is whether they imitate the iterative nature of manual sound design---that is, listening to the output and adjusting the synthesizer parameters recursively until the desired goal is reached.
+The most fit members of the group are the parameters which perform the best in the loss function, and create a new generation of parameters via mutation (random change in subset of parameters) and crossovers (combination of parameter sets); this process repeats until the goal or a maximum number of generations is reached.
+
+% One important aspect of the heuristics is whether they imitate the iterative nature of manual sound design---that is, listening to the output and adjusting the synthesizer parameters recursively until the desired goal is reached.
 Supervised methods generate their final output in a single step, while direct sound optimization methods refine their outputs iteratively until termination, which resembles manual sound design more closely.
@@
 \label{sec:iterative_sound_matching}
-The diagram in Figure~\ref{fig:sound_design_loop_iterative} shows a general model for iterative sound-matching that covers previous works in this domain.
+ Figure~\ref{fig:sound_design_loop_iterative} shows a general model for iterative sound-matching.
 To begin the process, the parameters are arbitrarily initialized (usually random generation), the similarity of the target and output is measured, and the parameter is \textit{optimized} with the goal of increasing the similarity, or alternatively, reducing the loss.
@@
 \end{tikzpicture}
-    \caption{ Iterative approach to sound design.
+    % \caption{ Iterative approach to sound design.
 Synthesizer $g$ with arbitrary initialized parameters $\hat{\theta}$, creates sound $x$.
@@
 $\phi$ is the audio feature extractor, or representation function.}
+    \caption{ Iterative approach to sound design.}
     \label{fig:sound_design_loop_iterative}
@@
 
-Genetic mutation has predominantly been the most common direct optimization method in previous works, particularly in non-differentiable settings. 
-In a differentiable setting, gradient descent can be used to directly optimize the parameters of a synthesizer~\cite{vahidi2023mesostructures}.
 In the upcoming ``Past Works'' section, we will discuss the evolution of prior sound-matching research, with \textit{differentiable iterative sound-matching} being the latest extension of the field.
@@
  The sound-matching works that followed used other heuristics for finding the synthesizer parameters, yet the simple FM synthesizer structure remained unchanged.
-Horner \textit{et al.}~\cite{horner1993machine} used GAs for re-synthesis of sounds with an FM synthesizer using one modulator and one carrier oscillator, and the McAulay-Quatieri (MQ) method~\cite{mcaulay1986speech} for measuring loss.
-MQ calculates the STFT and marks the start and end of \textit{trends} in a sound with the assumption that the harmonic content belongs to trends that have a beginning and an end.
+Horner \textit{et al.}~\cite{horner1993machine} used GAs for re-synthesis of sounds with an FM synthesizer using one modulator and one carrier oscillator, and the McAulay-Quatieri method~\cite{mcaulay1986speech} for measuring loss.
+
+ % MQ calculates the STFT and marks the start and end of \textit{trends} in a sound with the assumption that the harmonic content belongs to trends that have a beginning and an end.
 
 Introduction of more complex models of synthesis such as wavetables~\cite{horner2003auto} and physical modeling~\cite{riionheimo2003parameter} rendered interesting results, but led to further questions about the nature and goals of sound-matching.
-Mitchell and Creasy noted the difficulty in distinguishing between \textit{inefficiency of the optimization engine} and \textit{synthesizer limitations} as the cause of failure or success in sound-matching~\cite{mitchell2007evolutionary}.
+Mitchell and Creasy noted the difficulty in distinguishing between ``inefficiency of the optimization engine'' and ``synthesizer limitations'' as the cause of failure or success in sound-matching~\cite{mitchell2007evolutionary}.
 They offered a \textit{contrived methodology} for finding the best evolutionary method for sound-matching using their FM synthesizer.
 The methodology postulates that the best search heuristic for in-domain search (where an exact target exists, and a wide range of targets can be produced by sampling from different points in the parameter space) would also be the best-performing heuristic for out-of-domain search on a dataset of muted trumpet tones~\cite{opolko1989mcgill}; however, testing was not extensive and yielded some contradictory results.
-
-Testing sound-matching experiments can be difficult, in part due to the large number of parametric components and navigation of complex non-linear systems.
+As they noted, modifications to the synthesizer, loss function, and sound domain would lead to a problem with an entirely new search space~\cite{mitchell2007evolutionary}.
+
+% Testing sound-matching experiments can be difficult, in part due to the large number of parametric components and navigation of complex non-linear systems.
 The issue with finding the best approaches to sound-matching is exacerbated when slight changes to the pipeline render previous results moot.
 Perhaps due to these reasons, the contrived methodology was not extensively tested by Mitchell and Creasy, who only provided a single experiment where the best evolutionary strategy found by in-domain search would also perform well in out-of-domain synthesis.
-As they noted, modifications to the synthesizer, loss function, and sound domain would lead to a problem with an entirely new search space~\cite{mitchell2007evolutionary}.
-
 \subsection{Introduction of Deep-Learning}
@@
 
-Esling \textit{et al.} also noted the issue of using P-Loss in sound-matching~\cite{esling2019flow}, considering the non-linear mapping between parameters of a synth and the audio output.
+Esling \textit{et al.} noted the issue of using P-Loss in sound-matching~\cite{esling2019flow}, considering the non-linear mapping between parameters of a synth and the audio output.
 In their approach, dubbed as ``flowsynth'', they used a large parameter-to-audio dataset for a commercial \gls{VST} synthesizer.
@@
 
-Rather than focusing on a particular implementation, Shier \textit{et al.} presented Spieglib, a library for implementation of sound-matching pipelines (which Shier \textit{et al.} referred to as ``automatic parameter search'')~\cite{shier2020spiegelib}.
+Rather than focusing on a particular implementation, Shier \textit{et al.} presented Spieglib, a library for implementation of sound-matching pipelines~\cite{shier2020spiegelib}.
+This library provides different choices for DNNs, GAs, synthesizers, and feature extractors.
+Shier \textit{et al.} presented an experiment with a similar setup to Yee-King \textit{et al.}~\cite{yee2018automatic}, however, they found a genetic algorithm to be the best performing.
+
+% Rather than focusing on a particular implementation, Shier \textit{et al.} presented Spieglib, a library for implementation of sound-matching pipelines (which Shier \textit{et al.} referred to as ``automatic parameter search'')~\cite{shier2020spiegelib}.
 This library provides extensible choices of deep-learning estimators with \textit{tensorflow}~\cite{abadi2016tensorflow}, genetic estimators with \textit{DEAP}~\cite{DEAP_JMLR2012}, feature extractors with \textit{librosa}~\cite{mcfee2015librosa}, and different choices for synthesizers, evaluators, dataloaders.
@@
 Uzrad \textit{et al.}~\cite{uzrad2024diffmoog} took another unique approach to sound-matching: using a differentiable \textit{synthesis chain} of DSP generators and effects and a loss function that combines P-Loss with a \textit{signal-chain loss}.
-Like other supervised learning replication approaches, during training a target sound $t$ is generated by random  $\theta$ parameter selection, and a model attempts to an output $\tilde{\theta}$ that generates a sound similar to $t$.
 The synthesizer is a customizable chain of effects, which feed one output as input to the next step of the chain; signal-chain loss compares the parameter and output difference at every output step in the chain~\cite{uzrad2024diffmoog}.
@@
 
-Uzrad \textit{et al.}\cite{uzrad2024diffmoog} used log-spectral distance (LSD) as their loss function, where   
-$\text{LSD} = \| \log(S(t)) - \log(S(x)) \|_F$, S is the mel-spectrogram of the sounds and F is the Frobenius norm~\cite{golub2013matrix}. The Frobenius norm of a matrix $A$ is defined as
-$ \|A\|_F = \sqrt{\sum_{i,j} |a_{ij}|^2}$.
+% Uzrad \textit{et al.}\cite{uzrad2024diffmoog} used log-spectral distance (LSD) as their loss function, where   
+% $\text{LSD} = \| \log(S(t)) - \log(S(x)) \|_F$, S is the mel-spectrogram of the sounds and F is the Frobenius norm~\cite{golub2013matrix}. The Frobenius norm of a matrix $A$ is defined as
+% $ \|A\|_F = \sqrt{\sum_{i,j} |a_{ij}|^2}$.
 Here, $A$ would be the 2D matrix resulting from subtracting two log-spectrograms.
@@
 
-\section{What is Lacking In the Field}
+\subsection{What is Lacking In the Field}
 \label{sec:lacking}
-  % First, we need to clearly define the characteristics we are looking for with non-generic loss functions, otherwise the chances of finding the sound(s) of interest is slim.
+%   % First, we need to clearly define the characteristics we are looking for with non-generic loss functions, otherwise the chances of finding the sound(s) of interest is slim.
 Second, we need to move beyond direct replication, and focus on imitating target sounds that are outside the synthesizers capabilities.
@@
 
-Based on the literature reviewed, we find several gaps in past research that makes the application of sound-matching for sound-designers difficult.
+% Based on the literature reviewed, we find several gaps in past research that makes the application of sound-matching for sound-designers difficult.
 Thus far, replication of in-domain sounds has been the main subject of research.
@@
 Sound-matching would be an interesting creative tool if we could use it to either:
-\begin{enumerate}
- \item \textbf{Replicate out-of-domain sounds}: this would give us the  the parameters that could closely approximate a target sound, and we can modify the sound further by modulating the synthesizer.
-\item \textbf{Imitate in-domain or out-of-domain sounds}: This would allow for applying the characteristics of one sound to another. 
-\end{enumerate}
-
-We need to be aware of the ``open set recognition'' (\gls{OSR}) problem here \cite{mundt2019open,gers2000learning}.
+% \begin{enumerate}
+%  \item \textbf{Replicate out-of-domain sounds}: this would give us the  the parameters that could closely approximate a target sound, and we can modify the sound further by modulating the synthesizer.
+% \item \textbf{Imitate in-domain or out-of-domain sounds}: This would allow for applying the characteristics of one sound to another. 
+% \end{enumerate}
+
+% We need to be aware of the ``open set recognition'' (\gls{OSR}) problem here \cite{mundt2019open,gers2000learning}.
 OSR is a well documented issue in DL that manifests when the set or category that the network is supposed to reject or accept cannot be explained via examples.
@@
 
-Perhaps one method of addressing the OSR problem is the use of representations (or feature extraction) methods that reduce sounds to a feature vector in the same domain.
+% Perhaps one method of addressing the OSR problem is the use of representations (or feature extraction) methods that reduce sounds to a feature vector in the same domain.
 For example, a representation that only looks at the envelope of sounds is far less likely to be subject to the OSR problem.
@@
 
-Another weakness \textcolor{highlight}{is that the synthesis methods in past works are not representative of the common methods used by sound designers and modern synthesizers}.
+% Another weakness \textcolor{highlight}{is that the synthesis methods in past works are not representative of the common methods used by sound designers and modern synthesizers}.
 Simple FM/AM synthesizers have generally been the go-to methods of sound-matching.
 We believe that general statements regarding the effectiveness of an approach to iterative sound-matching (particularly in regards to a loss function) can only be reasonably made in the context of the subset of the tested methods of synthesis.
-\textcolor{highlight}{The simplicity of parameter optimization approaches} is another area of weakness in past works.
+% \textcolor{highlight}{The simplicity of parameter optimization approaches} is another area of weakness in past works.
 Classic genetic algorithms update the parameters randomly, while in recent differentiable settings, simple gradient descent methods are used.
@@
 
-\subsection{Defining The Major Issues}
-\label{sec:addressing_problems}
-Based on the weaknesses noted above, we propose four problems that need to be addressed to improve the sound-matching field. 
+% \subsection{Defining The Major Issues}
+% \label{sec:addressing_problems}
+
+Having looked at the current available literature, we note four major areas of weakness in sound-matching:
 \begin{enumerate}
@@
 
-In Section~\ref{sec:what_is_done}, we discuss our work in simultaneously bridging the knowledge gap with regards to the \LossSelect~and \SynthSelect~problems.
+In Section~\ref{sec:progs_methodology_results}, we discuss our work in simultaneously bridging the knowledge gap with regards to the \LossSelect~and \SynthSelect~problems.
 The latter two issues are left for future work; possible approaches are discussed in Sections~\ref{sec:summary_conclusion} and~\ref{sec:future}.
@@
 \section{Programs, Methodology, and Results}
-\label{sec:what_is_done}
-In this section we report the results of our efforts in addressing the first two problems.
-\textbf{We compare the performance of iterative sound-matching experiments, where four differentiable loss functions are applied to four differentiable synthesizers.} What we mean by performance is the similarity of the final output to the target, as measured by P-Loss, MSS, and manual listening.
+\label{sec:progs_methodology_results}
+Towards addressing the first two problems, we compare the performance iterative sound-matching experiments, where four differentiable loss functions are applied to four differentiable synthesizers.
+What we mean by performance is the similarity of the final output to the target, as measured by P-Loss, MSS, and manual listening.
 
@@
 \label{sec:methodology}
-We have highlighted the lack of attention given to the selection of loss functions in sound-matching, and noted that the few existing works addressing this issue evaluate their proposed losses using only simplistic sound synthesis methods.
+% We have highlighted the lack of attention given to the selection of loss functions in sound-matching, and noted that the few existing works addressing this issue evaluate their proposed losses using only simplistic sound synthesis methods.
 \textcolor{highlight}{We hypothesize that the best performing loss function is influenced by the synthesis techniques utilized in the synthesizer}.
 This hypothesis targets the issues of \LossSelect~and \SynthSelect.
-For our experiments, we test four different loss functions and compare their success in sound-matching using four different synthesizers.
+
+We test four different loss functions and compare their success in sound-matching using four different synthesizers.
 We verify the success of the sound-matching process using two automatic measures of sound similarity (MSS and P-Loss) and a manual blinded hearing test conducted by two of the authors where likert scores~\cite{jebb2021review} of 1-5 are assigned to a subset of targets and outputs.
 
-An \textit{experiment} is one iteration of sound-matching, given a loss function, target sound,
+An \textit{experiment} is one iteration of sound-matching given a loss function, target sound,
 and synthesizer program.
@@
 When the loop is terminated, we measure the similarity of the final output sound $x$ (where $P(\theta_{199}) = x$) to $t$, or if using P-Loss we compare $\theta_{199}$ to $\theta^*$.
-Since we ran 300 experiments for each loss, we can compare the bootstrapped distribution of losses or likert scores~\cite{jebb2021review} assigned by manual review to determine whether $L_1$ or $L_2$ led to better performance.
-This simplified example leads us to our methodolgy which has four loss functions, four synthesizer programs, and three different methods of scoring sound similarity.
+Having 300 experiments gives us a distribution of losses to determine whether $L_1$ or $L_2$ led to better performance.
+This simplified example leads us to our methodology which has four loss functions, four synthesizer programs, and three different methods of scoring sound similarity.
 
@@
 
-\subsubsection{Training Parameters and Gradient Calculations}
- Previous work utilized gradient fields for analysis of the algorithm performance and gradient update~\cite{vahidi2023mesostructures}.
+% \subsubsection{Training Parameters and Gradient Calculations}
+%  Previous work utilized gradient fields for analysis of the algorithm performance and gradient update~\cite{vahidi2023mesostructures}.
 Gradient fields are not the center of our analysis, but we include them in Appendix~\ref{appendix:gradient_fields}.
@@
 
-Additionally, the gradients are large, and calculated backwards sequentially through the signal processing chain.
+% Additionally, the gradients are large, and calculated backwards sequentially through the signal processing chain.
 This backward calculation resembles the \textit{exploding gradients} problem in recurrent neural networks~\cite{gers2000learning}.
@@
 Clipped gradients $\gamma_i^{\text{clipped}}$ are calculated by multiplying them with a scale factor $\mathcal{S}$
-\[
-\mathbf{\gamma}_i^{\text{clipped}} = \mathbf{\gamma}_i \cdot \mathcal{S}
-\]
-Where $\mathcal{S}$ is a value less than or equal to 1, depending on whether the norm of the gradients exceeds the threshold $c$.  
-\[
-\mathcal{S} = \frac{c}{\max(\|\mathbf{\gamma}\|_2, c)}
-\]
+% \[
+% \mathbf{\gamma}_i^{\text{clipped}} = \mathbf{\gamma}_i \cdot \mathcal{S}
+% \]
+% Where $\mathcal{S}$ is a value less than or equal to 1, depending on whether the norm of the gradients exceeds the threshold $c$.  
+% \[
+% \mathcal{S} = \frac{c}{\max(\|\mathbf{\gamma}\|_2, c)}
+% \]
 
@@
 
-For each program, we provide the Faust code~\cite{orlarey2009faust}.
+For each program, we provide the Faust code~\cite{orlarey2009faust}, which can be run in the online IDE~\footnote{\url{https://faustide.grame.fr/}}.
 Faust is a functional language for audio synthesis that can succinctly define signal processing chains.
-The publicly available Faust IDE\footnote{\url{https://faustide.grame.fr/}} is the simplest method for running the programs.
-A brief overview of each program is given in Table~\ref{tab:programs}.
 The programs are first defined in Faust, then converted (or \textit{transpiled}) to differentiable Jax functions~\cite{deepmind2020jax} using the DawDreamer library\footnote{\url{https://github.com/DBraun/DawDreamer}}.
@@
 
-A general overview of the programs and the predefined parameter ranges is given in Table~\ref{tab:programs}.
+% A general overview of the programs and the predefined parameter ranges is given in Table~\ref{tab:programs}.
+
 In the following subsections, we describe the programs and their parameters in more detail, followed by the methodology in Section~\ref{sec:methodology} and results in Section~\ref{sec:results_posthoc}.
 
-\begin{table*}[htbp]
-\centering
-\begin{adjustbox}{max width=\textwidth}
-\begin{tabular}{|c|c|c|c|l|}
-\hline
-\textbf{ID} & \textbf{Parameters} & \makecell{\textbf{Param. 1}\\\textbf{Range}} & \makecell{\textbf{Param. 2}\\\textbf{Range}} & \textbf{Description} \\
-\hline
-\BPNoise & hp\_cut, lp\_cut & 50-1000 & 1-120 & Band-pass of Noise signal \\
-\hline
-\AddSineSaw & saw\_freq, sine\_freq & 20-1000 & 20-1000 & Addition of sine and saw waves \\
-\hline
-\AmpMod & amp, carrier & 1-20 & 20-1000 & LFO on amplitude of noise osc \\
-\hline
-\FMMod & amp, carrier & 0-5 & 0-4 & LFO on amplitude of saw osc \\
-\hline
-\end{tabular}
-\end{adjustbox}
-\caption{Overview of Program Parameters and the range of values each parameter can take.}
-\label{tab:programs}
-\end{table*}
+% \begin{table*}[htbp]
+% \centering
+% \begin{adjustbox}{max width=\textwidth}
+% \begin{tabular}{|c|c|c|c|l|}
+% \hline
+% \textbf{ID} & \textbf{Parameters} & \makecell{\textbf{Param. 1}\\\textbf{Range}} & \makecell{\textbf{Param. 2}\\\textbf{Range}} & \textbf{Description} \\
+% \hline
+% \BPNoise & hp\_cut, lp\_cut & 50-1000 & 1-120 & Band-pass of Noise signal \\
+% \hline
+% \AddSineSaw & saw\_freq, sine\_freq & 20-1000 & 20-1000 & Addition of sine and saw waves \\
+% \hline
+% \AmpMod & amp, carrier & 1-20 & 20-1000 & LFO on amplitude of noise osc \\
+% \hline
+% \FMMod & amp, carrier & 0-5 & 0-4 & LFO on amplitude of saw osc \\
+% \hline
+% \end{tabular}
+% \end{adjustbox}
+% \caption{Overview of Program Parameters and the range of values each parameter can take.}
+% \label{tab:programs}
+% \end{table*}
 
@@
 If the null hypothesis is rejected, the second stage ranks the loss functions from best to worst using the non-parametric Scott-Knott algorithm (\gls{SNPK})~\cite{tantithamthavorn2017mvt,tantithamthavorn2018optimization}.
-
 The Kruskal-Wallis test is used for the first stage~\cite{kruskal1952use}.
-This test works by pooling and ranking all performance evaluation measures for a program (either P-loss or MSS), then testing whether the differences in mean rank for all loss function groups is zero.
+This test works by pooling and ranking all performance evaluation measures for a program, then testing whether the differences in mean rank for all loss function groups is zero.
 The Kruskal-Wallis test calculates an H statistic that is compared to a chi-square distribution with k-1 degrees of freedom (k is the number of groups, or 4, and degrees of freedom is 3).
 Using the significance level of 0.05, if the H statistic is greater than the critical value of the chi-square distribution then the null hypothesis is rejected, meaning that at least one group has a mean rank significantly above others~\cite{kruskal1952use}.
-Table~\ref{tab:kruskal_auto} shows the Kruskal-Wallis results for every program and automatic evaluation method, and shows that significant differences exist in all cases except MSS for the \AmpMod{} program.
-The manual ranking results showed significant differences in all cases, as shown in table~\ref{tab:kruskal_manual}.
+Table~\ref{tab:kruskal_auto} shows the Kruskal-Wallis results for every program and evaluation method, and shows that significant differences exist in all cases except MSS for the \AmpMod{} program.
 
@@
 \BPNoise & P-Loss  & 163.60 & $3.07 \times 10^{-35}$ & Yes \\
+\BPNoise & Manual & 9.45  & $2.39 \times 10^{-2}$ & Yes \\
 \AddSineSaw & MSS      & 308.97 & $1.14 \times 10^{-66}$ & Yes \\
 \AddSineSaw & P-Loss  & 348.42 & $3.28 \times 10^{-75}$ & Yes \\
+\AddSineSaw & Manual & 9.45  & $2.39 \times 10^{-2}$ & Yes \\
 \AmpMod & MSS      & 1.35   & $0.7171$               & No \\
 \AmpMod & P-Loss  & 366.76 & $3.50 \times 10^{-79}$ & Yes \\
+\AmpMod & Manual & 32.71 & $3.70 \times 10^{-7}$ & Yes \\
 \FMMod & MSS      & 564.65 & $4.65 \times 10^{-122}$ & Yes \\
 \FMMod & P-Loss  & 229.19 & $2.07 \times 10^{-49}$ & Yes \\
+\FMMod & Manual &  207.58 & $9.69 \times 10^{-45}$ & Yes \\
 \hline
 \end{tabular}
-\caption{Kruskal-Wallis results by program and evaluation method.
-The null hypothesis is that no difference in performance is observed between loss functions; it is rejected for all programs and evaluation methods with the exception of MSS combined with \AmpMod.}
+\caption{Kruskal-Wallis results by program and evaluation method..}
 \label{tab:kruskal_auto}
@@
 
-\begin{table}[ht]
-\centering
-\begin{tabular}{|c|c|c|c|}
-\hline
-\textbf{Program} & \textbf{H-Stat.} & \textbf{P-value} & \textbf{Reject} \\
-\hline
-\BPNoise & 9.45  & $2.39 \times 10^{-2}$ & Yes \\
-\AddSineSaw & 9.45  & $2.39 \times 10^{-2}$ & Yes \\
-\AmpMod & 32.71 & $3.70 \times 10^{-7}$ & Yes \\
-\FMMod & 207.58 & $9.69 \times 10^{-45}$ & Yes \\
-\hline
-\end{tabular}
-\caption{Manual ranking Kruskal-Wallis results. Significant differences are observed for all programs.}
-\label{tab:kruskal_manual}
-\end{table}
+% \begin{table}[ht]
+% \centering
+% \begin{tabular}{|c|c|c|c|}
+% \hline
+% \textbf{Program} & \textbf{H-Stat.} & \textbf{P-value} & \textbf{Reject} \\
+% \hline
+% \BPNoise & 9.45  & $2.39 \times 10^{-2}$ & Yes \\
+% \AddSineSaw & 9.45  & $2.39 \times 10^{-2}$ & Yes \\
+% \AmpMod & 32.71 & $3.70 \times 10^{-7}$ & Yes \\
+% \FMMod & 207.58 & $9.69 \times 10^{-45}$ & Yes \\
+% \hline
+% \end{tabular}
+% \caption{Manual ranking Kruskal-Wallis results. Significant differences are observed for all programs.}
+% \label{tab:kruskal_manual}
+% \end{table}
 
@@
     \vspace{-0.5em}
-    \caption{\BPNoise~bootstrapped distributions and ranks given by NPSK.
-Ranks and their corresponding colors from best to worst are \colorbox{rank1}{\textcolor{black}{\textbf{1}}}, \colorbox{rank2}{\textcolor{white}{\textbf{2}}}, \colorbox{rank3}{\textcolor{white}{\textbf{3}}}, \colorbox{rank4}{\textcolor{white}{\textbf{4}}}.}
+    \caption{\BPNoise~bootstrapped distributions and ranks given by NPSK.}
     \label{fig:npsk_p0}
@@
 
-Tables~\ref{tab:p0},~\ref{tab:p1},~\ref{tab:p2}, and~\ref{tab:p3} show the rankings for each program, with ranks different from the corresponding hearing test results marked with an asterisk.
+Table~\ref{tab:combined_ranks} shows the rankings for each program, with ranks different from the corresponding hearing test results marked with an asterisk.
 Both MSS and P-Loss gave a different rank from the hearing results in 3 out of 16 cases, which---somewhat surprisingly---shows consistency between automatic and manual hearing tests, at least for the simple programs used in this work.
 We also observe that identical top ranks were consistent across performance evaluation methodologies, with the exception of P-Loss in \BPNoise, which narrowly picks a different spectrogram-based loss function.
-\input{extra_figures/ranking_tables}
+% \input{extra_figures/ranking_tables}
+\begin{table*}[htbp]
+\centering
+\footnotesize
+\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
+\hline
+\textbf{Function} 
+  & \multicolumn{3}{c|}{\BPNoise}
+  & \multicolumn{3}{c|}{\AddSineSaw}
+  & \multicolumn{3}{c|}{\AmpMod}
+  & \multicolumn{3}{c|}{\FMMod} \\
+\cline{2-13}
+  & MSS & P-LOSS & Hearing 
+  & MSS & P-LOSS & Hearing 
+  & MSS & P-LOSS & Hearing 
+  & MSS & P-LOSS & Hearing \\
+\hline
+\textbf{SIMSE} 
+  & 1    & 2*   & 1    
+  & 4    & 4    & 4    
+  & 3*   & 4    & 4    
+  & 2    & 2    & 2    \\
+\textbf{L1}    
+  & 2    & 1*   & 2    
+  & 2    & 2    & 2    
+  & 2    & 2    & 2    
+  & 3    & 3    & 3    \\
+\textbf{JTFS}  
+  & 3    & 4*   & 3    
+  & 1*   & 1    & 1    
+  & 4*   & 3    & 3    
+  & 4    & 4    & 4    \\
+\textbf{DTW}   
+  & 4*   & 3    & 3    
+  & 3    & 3    & 3    
+  & 1    & 1    & 1    
+  & 1    & 1    & 1    \\
+\hline
+\end{tabular}
+\vspace{0.25em}
+\caption{Ranks for each synthesis method (rows) under the three evaluation metrics (columns), across four targets (\BPNoise, \AddSineSaw, \AmpMod, \FMMod).
+Asterisks denote the best (lowest) rank in each column.}
+\label{tab:combined_ranks}
+\end{table*}
+
 \subsection{Takeaways}
-Perhaps the most important takeaway from our analysis is that the loss function that yields the best sound-matching outcomes varies depending on the synthesizer program.
+An important takeaway is that the loss function that yields the best sound-matching outcomes varies depending on the synthesizer program.
 This is perhaps due to the interaction between how the parameters of the synthesizer influence the sound, and the core sonic features of the loss function.
@@
 
-Using simple synthesizers, small sound domains, and a limited number of loss functions, previous works have been able to claim some loss functions or feature extractors for audio as ``better performing" for sound-matching (or other audio-related tasks)\cite{vahidi2023mesostructures,turian2020sorry,masuda2021soundmatch,engel2020ddsp,han2023perceptual}.
+Using simple synthesizers, small sound domains, and a limited number of loss functions, previous works have been able to claim some loss functions or feature extractors for audio as ``better performing" for sound-matching (or other audio-related tasks)~\cite{vahidi2023mesostructures,turian2020sorry,masuda2021soundmatch,engel2020ddsp,han2023perceptual}.
 Is the search for a universally best-performing loss function a worthy endeavor, or perhaps the choice of loss function is domain dependent?
@@
  We considered P-Loss and MSS as performance measures, commonly used in other sound-matching works, and compared their rankings to those given by blind hearing tests.
-We observed that automatic performance measures and manual listening tests were often (though not always) in agreement.
-Manual verification of sound-matching results remains a necessity for future works.
+We observed that automatic performance measures and manual listening tests were often (though not always) in agreement, despite this, we believe that manual verification of sound-matching results remains a necessity for future works.
 

### 2025-04-30 main.tex
Updates from Overleaf

@@
 There are practically an infinite number of parameter combinations, and their effect on the output sound is hard to predict.
-Iterative sound-matching is the automation of this process, more formally, the automatic programming of an audio synthesizer to generate outputs similar to a target sound while guided by a loss function that quantifies sound similarity.
-Past works in sound-matching often relied on a narrow range of synthesizers and overlooked the possible interactions between the loss function and synthesizer parameters.
+Iterative sound-matching is the automation of this process, more formally, the continual programming of an audio synthesizer to generate outputs similar to a target sound while guided by a loss function that quantifies sound similarity.
+We propose differentiable iterative sound-matching as the natural extension of the current available literature.
+Past works in sound-matching often relied on a narrow range of synthesizers in their analysis of ``best'' loss functions and overlooked the possible interactions between the loss and synthesizer parameters.
+Additionally, the results have rarely been verified with manual hearing tests.
 We analyzed the interplay between four simple subtractive, additive, and FM synthesizers when guided by four different novel or established similarity functions.
-For each loss function and synthesizer pair, we ran 300 experiments with randomized initial values and targets, and used the final parameter configurations and spectrogram distances as key performance evaluation metrics of the performance.
+For each loss function and synthesizer pair, we ran 300 experiments with randomized initial values and targets, and used the final parameter configurations, spectrogram distances, and manually assigned scores as key performance evaluation metrics of the performance.
 A post-hoc analysis of these performance measures showed that the best‐performing loss function varies based on the method of synthesis at hand, which highlights the value of novel similarity measures and de-prioritizes the search for the elusive universally optimal sound similarity measure.
@@
 
+% the automatic performance measures have not been compared to human hearing
+% lack of creativity in loss function definitions 
 \begin{IEEEkeywords}
@@
 Ideally, these updates are in a direction that improves audio similarity by reducing loss.
-Differentiable iterative sound-matching has only recently become viable thanks to continued advances in both computation hardware~\cite{moore2012introduction,schaller1997moore}, and machine learning software~\cite{deepmind2020jax,paszke2019pytorch}.
+Differentiable iterative sound-matching has only recently become viable thanks to continued advances in both computation hardware~\cite{moore2012introduction,schaller1997moore}, and machine learning software~\cite{braun2024dac,paszke2019pytorch}.
 
@@
 Waveform generators are called oscillators for this reason.
-If $n$ is in the range of 30 to 20,000 (the upper limit is often lower for some), $x(n)$ could be played as an audible tone~\cite{smith2007mathematics,jilek2014reference}.
+
+% If $n$ is in the range of 30 to 20,000 (the upper limit is often lower for some), $x(n)$ could be played as an audible tone~\cite{smith2007mathematics,jilek2014reference}.
 
@@
 \centering
+\caption{Summary of select works in sound-matching.
+Gen.
+refers to whether the generations are done in 1-shot (generally, output by a neural network) or a search is conducted with the goal of iteratively getting closer to the target sound.}
 \setlength{\tabcolsep}{3pt} % Reduce horizontal spacing
@@
 }
-\caption{Summary of select works in sound-matching.
-Gen.
-refers to whether the generations are done in 1-shot (generally, output by a neural network) or a search is conducted with the goal of iteratively getting closer to the target sound.}
+
 \label{tab:summary}
@@
 When the final iteration is complete, a score is given (automatically with an algorithm or with manual listening) to an experiment that measures its success in replicating the target sound.
-Consider the following example: we have two loss functions, $L_1$ and $L_2$, and a single synthesizer program $P$ with a target parameter set $\theta^*$ where the sound output is $t$, or $P(\theta^*) = t$.
+Consider the following simplified scenario: we have two loss functions, $L_1$ and $L_2$, and a single synthesizer program $P$ with a target parameter set $\theta^*$ where the target sound is $t$, or $P(\theta^*) = t$.
 We run a large number (in this case, 300 for each loss function) of experiments where we randomly set the initial ($\theta_0$) and target parameters ($\theta^*$) of $P$ and run an iterative sound-matching loop for each loss function, with max number of iterations set to 200.
@@
 \subsubsection{Determining Best Performers}
-Determining the best performing loss function(s) might seem complicated since we have different synthesizer programs, loss functions, and evaluation metrics.
-Figure~\ref{fig:posthoc_evaluation} is a visualized summary of how the loss functions are ranked in three ways (MSS, P-Loss, Likert scores).
-
+ Figure~\ref{fig:posthoc_evaluation} is a visualized summary of how the loss functions are ranked.
 We have four loss functions and four synthesizer programs.
 The maximum number of iterations set to 200, and 300 experiments for each loss/synthesizer combination.
-Each synthesizer program  is paired with four loss functions.
-Under each of these loss functions, there are 300 experiments.
-Note that P-Loss and MSS are performance evaluation methods that measure similarity of final output and target sound; They are not one of the four loss functions.
-P-Loss and MSS are assigned automatically to each of the 300 pairs of target and output sounds.
-Blinded manual scoring using a Likert scale~\cite{jebb2021review} is assigned to 40 randomly selected sounds, where integers from 1 and 5 are assigned, with 1 meaning no similarity and 5 for near identical.
+Each synthesizer program  is paired with four loss functions and 300 experiments are conducted and evaluated automatically with P-Loss and MSS.
+Two of the authors assigned blinded Likert scores to 40 randomly selected sounds, where integers from 1 to 5 are assigned, with 1 meaning no similarity and 5 for near identical.
 
@@
 
-% \subsubsection{Training Parameters and Gradient Calculations}
-%  Previous work utilized gradient fields for analysis of the algorithm performance and gradient update~\cite{vahidi2023mesostructures}.
+\subsubsection{Training Parameters and Gradient Calculations}
+ % Previous work utilized gradient fields for analysis of the algorithm performance~\cite{vahidi2023mesostructures}.
 Gradient fields are not the center of our analysis, but we include them in Appendix~\ref{appendix:gradient_fields}.
-During the experiments, the updates applied to the initial parameters are recorded for gradient field generation.
-For updating the weights, we use RMSProp, which operates similar to stochastic gradient descent (SGD)~\cite{goodfellow2016deep}, with the caveat that the gradients of each weight are scaled by the root-mean-square of past gradients of that weight.
+ For updating the synthesizer parameters (or weights), we use RMSProp, which operates similar to stochastic gradient descent (SGD)~\cite{goodfellow2016deep}, with the caveat that the gradients of each weight are scaled by the root-mean-square of past gradients of that weight.
 We used a fixed learning rate of 0.045 for all experiments.
 The maximum number of iterations is set to 200, where, based on our observations, the parameters have either irrecoverably diverged outside the acceptable ranges, or are stuck at a local minimum.
-At the maximum number of iterations the search is terminated and the P-Loss and MSS (as described in Section~\ref{sec:fourier_specs}) are calculated using the iteratively updated parameters, or the final sound output.
-
-% Additionally, the gradients are large, and calculated backwards sequentially through the signal processing chain.
+Additionally, the gradients are large, and calculated backwards sequentially through the signal processing chain.
 This backward calculation resembles the \textit{exploding gradients} problem in recurrent neural networks~\cite{gers2000learning}.
-For this reason, gradient clipping~\cite{goodfellow2016deep} is used as a preemptive measure to prevent large divergences in the parameter updates, but it is unclear if it facilitates learning in these experiments.
-The gradient clipping function~\cite{goodfellow2016deep} takes a set of gradients $\gamma_i$ and ensures that the $\ell_2$ norm of all gradients, or $||\gamma||_2:=\sqrt{\sum\limits_{i}(\gamma_{i}\cdot\gamma_{i})}$ does not exceed a clipping threshold of c (here set to 1).
-Clipped gradients $\gamma_i^{\text{clipped}}$ are calculated by multiplying them with a scale factor $\mathcal{S}$
-% \[
+For this reason, gradient clipping~\cite{goodfellow2016deep} is preemptively used to ensure that the  $\ell_2$ norm of all gradientsdoes not exceed a clipping threshold 1.
 % \mathbf{\gamma}_i^{\text{clipped}} = \mathbf{\gamma}_i \cdot \mathcal{S}
@@
 Faust is a functional language for audio synthesis that can succinctly define signal processing chains.
-The programs are first defined in Faust, then converted (or \textit{transpiled}) to differentiable Jax functions~\cite{deepmind2020jax} using the DawDreamer library\footnote{\url{https://github.com/DBraun/DawDreamer}}.
+The programs are first defined in Faust, then converted (or \textit{transpiled}) to differentiable Jax functions~\cite{braun2024dac} using the DawDreamer library\footnote{\url{https://github.com/DBraun/DawDreamer}}.
 
@@
 \centering
+\caption{Kruskal-Wallis results by program and evaluation method.}
 \begin{tabular}{|c|c|c|c|c|}
@@
 \end{tabular}
-\caption{Kruskal-Wallis results by program and evaluation method..}
 \label{tab:kruskal_auto}
@@
 
-As we've discussed, automatic sound evaluation methods are unreliable, as we do not know how they would align with the often subjective opinion of human listeners.
+Automatic sound evaluation methods are unreliable, as we do not know how they would align with the often subjective opinion of human listeners.
 Here, the manual blinded rankings given by the reviewers could be treated as a ground truth, particularly since the two reviewers gave very similar rankings to the experiment results.
-
-To measure the agreement between the two listeners, Spearman's rank correlation is used.
-The Spearman test returns a p-value for the null hypothesis of no correlation, along with a $\rho$ value between -1 to 1; -1 for perfect negative correlation, 1 for perfect correlation, and 0 for no correlation~\cite{spearman1987proof,rebekic2015pearson}.
+To measure the agreement, Spearman's rank correlation is used.
+The Spearman test returns a p-value for the null hypothesis of no correlation, along with a $\rho$ value between -1 (perfect negative correlation) to 1 (perfect correlation), and middle value of 0 for no correlation~\cite{spearman1987proof,rebekic2015pearson}.
 For each program and loss function, the blind manual hearing test gives us 80 ranks, which are given by two of the authors to a set of 40 randomly selected experiment results.
-This gives 160 data points per program, and 640 data points for all four programs.
+This gives 160 pairs of ranks per program, and 640 for all programs.
 We computed Spearman correlations between the two reviewers.
@@
 \centering
+\caption{Ranks for each synthesis method (rows) under the three evaluation metrics (columns), across four targets (\BPNoise, \AddSineSaw, \AmpMod, \FMMod).
+Asterisks denote difference from Likert score ranks.}
 \footnotesize
@@
   & 3    & 4*   & 3    
-  & 1*   & 1    & 1    
+  & 1   & 1    & 1    
   & 4*   & 3    & 3    
@@
 \end{tabular}
-\vspace{0.25em}
-\caption{Ranks for each synthesis method (rows) under the three evaluation metrics (columns), across four targets (\BPNoise, \AddSineSaw, \AmpMod, \FMMod).
-Asterisks denote the best (lowest) rank in each column.}
+% \vspace{0.25em}
+
 \label{tab:combined_ranks}
@@
 
-We exposed several areas of weakness in the field, for which we presented methods for bridging the knowledge gaps.
-These issues included loss and synthesizer selection, periodic loss landscapes, and verification of performance measures with in-domain and out-of-domain sounds.
+%  We discussed areas of weakness in the field, including loss and synthesizer selection, periodic loss landscapes, and verification of performance measures with in-domain and out-of-domain sounds.
 The first two issues being the main target of the experiments in this work.
 
-Using simple synthesizers, small sound domains, and a limited number of loss functions, previous works have been able to claim some loss functions or feature extractors for audio as ``better performing" for sound-matching (or other audio-related tasks)~\cite{vahidi2023mesostructures,turian2020sorry,masuda2021soundmatch,engel2020ddsp,han2023perceptual}.
+% Using simple synthesizers, small sound domains, and a limited number of loss functions, previous works have been able to claim some loss functions or feature extractors for audio as ``better performing" for sound-matching (or other audio-related tasks)~\cite{vahidi2023mesostructures,turian2020sorry,masuda2021soundmatch,engel2020ddsp,han2023perceptual}.
 Is the search for a universally best-performing loss function a worthy endeavor, or perhaps the choice of loss function is domain dependent?
@@
 
- We considered P-Loss and MSS as performance measures, commonly used in other sound-matching works, and compared their rankings to those given by blind hearing tests.
+P-Loss and MSS have frequently been used as automatic performance measures, yet their ``preference'' has rarely been compared to human rankings.
 We observed that automatic performance measures and manual listening tests were often (though not always) in agreement, despite this, we believe that manual verification of sound-matching results remains a necessity for future works.
 
-\section{Weaknesses}
-Compared to previous work, we presented a relatively more cohesive approach to iterative sound-matching which utilizes and compares four loss functions and four synthesis methods.
-However, there are many other untested methods of synthesis that can be combined in practically infinite ways.
+\textbf{Weaknesses:} Compared to previous work, we presented a more cohesive approach to iterative sound-matching which utilizes and compares four loss functions and four synthesis methods.
+However, there are many other methods of synthesis and sound-similarity that can be combined in practically infinite ways.
 
@@
 We used bare-bones versions of STFT and JTFS with fixed parameters.
-We did not test complex synthesizers, which may use different signal generation functions in parallel or sequentially.
+We did not test complex synthesizers.
 Arbitrary hyperparameters such as learning rate and max number of iterations were selected for the DL pipeline, and only the RMSProp optimizer was tested.
 
-In Section~\ref{sec:in-domain}, we discussed the importance of sound domains.
  This work, much like previous work, only utilizes in-domain sounds.
@@
 
-\section{Future Work}
 \label{sec:future}
-The analysis of different loss functions sound-matching presented here shows that problem of periodic loss-landscapes is not limited to spectral representations using Fourier transforms~\cite{turian2020sorry,vahidi2023mesostructures,uzrad2024diffmoog,bruford2024synthesizer}.
+\textbf{Future Work: }The analysis of different loss functions sound-matching presented here shows that problem of periodic loss-landscapes is not limited to spectral representations using Fourier transforms~\cite{turian2020sorry,vahidi2023mesostructures,uzrad2024diffmoog,bruford2024synthesizer}.
 We can postulate that this problem emerges due to the periodic nature of sound.

### 2025-05-01 main.tex
abstract update

@@
 \begin{abstract}
-Manual sound design with a synthesizer is an iterative process.
-When an artist has a target sound in mind, they listen to the output of their chosen synthesizer, determine whether the output is closer or farther from the target, and adjust the parameters until satisfied.
+Manual sound design with a synthesizer is inherently iterative: an artist compares the synthesized output to a mental target, adjusts parameters, and repeats until satisfied.
 There are practically an infinite number of parameter combinations, and their effect on the output sound is hard to predict.
-Iterative sound-matching is the automation of this process, more formally, the continual programming of an audio synthesizer to generate outputs similar to a target sound while guided by a loss function that quantifies sound similarity.
-We propose differentiable iterative sound-matching as the natural extension of the current available literature.
-Past works in sound-matching often relied on a narrow range of synthesizers in their analysis of ``best'' loss functions and overlooked the possible interactions between the loss and synthesizer parameters.
-Additionally, the results have rarely been verified with manual hearing tests.
-We analyzed the interplay between four simple subtractive, additive, and FM synthesizers when guided by four different novel or established similarity functions.
-For each loss function and synthesizer pair, we ran 300 experiments with randomized initial values and targets, and used the final parameter configurations, spectrogram distances, and manually assigned scores as key performance evaluation metrics of the performance.
-A post-hoc analysis of these performance measures showed that the best‐performing loss function varies based on the method of synthesis at hand, which highlights the value of novel similarity measures and de-prioritizes the search for the elusive universally optimal sound similarity measure.
+Iterative sound-matching automates this workflow by continually programming a synthesizer under the guidance of a loss function that quantifies similarity to the target sound.
+Prior comparisons of loss functions have typically favored one metric over another, but only within narrow settings: limited synthesis methods, few loss types, often without blind listening tests.
+This leaves open whether a universally optimal loss exists, or the choice of loss remains a creative decision conditioned on the synthesis method and the sound designer's preferences.
+We propose differentiable iterative sound-matching as the natural extension of the available literature, since it combines the manual approach to sound design with modern advances in machine learning.
+To analyze the variability of loss function performance across synthesizers, we implemented differentiable subtractive, additive, and AM synthesizers, paired with four novel or established differentiable loss functions.
+For each of the sixteen synthesizer–loss combinations, we ran 300 randomized sound-matching trials.
+We calculated parameter differences, spectrogram-distance metrics, and gathered blind listening scores.
+A post-hoc analysis shows that no single loss function dominates across all synthesis methods; instead, the optimal measure depends on the architecture at hand.
+These findings underscore the value of developing new similarity metrics tailored to specific synthesis techniques, rather than pursuing a one-size-fits-all solution.
 \end{abstract}

### 2025-05-01 main.tex
Updates from Overleaf

@@
 \usepackage{listings}
-
-\usepackage{caption}               % re-enable caption functionality
+\usepackage{caption}         
+% re-enable caption functionality
 \captionsetup[lstlisting]{justification=centering, singlelinecheck=false}
@@
 
-\author{-}
+\author{Amir Salimi, Abram Hindle, Osmar R Zaiane}
+
+% The paper headers
+\markboth{IEEE TRANSACTIONS ON AUDIO, SPEECH AND LANGUAGE PROCESSING, VOL -, 2025}%
+{Shell \MakeLowercase{\textit{et al.}}: A Sample Article Using IEEEtran.cls for IEEE Journals}
+
 \maketitle
@@
 \textbf{Sound-matching} is a common form of automatic sound design explored in literature, where the goal is to find parameters for a synthesizer that replicate \textit{all} or \textit{some} of the characteristics of a target sound as best as possible, given some kind of similarity metric (or loss function)~\cite{horner1993machine,mitchell2007evolutionary}.
-
-\textit{Iterative} sound-matching refers to works that mimic the back-and-forth process of listening to a synthesizer's output and updating the parameters until the termination conditions are met.
+Our analysis of past works shows that this general goal of ``sound-matching'' can be broken down in to very different subtypes, many of them rarely or never explored.
+This leads us to a typology of sound-matching tasks, and the importance of \textit{differentiable, iterative sound-matching} as the extension of current work.
+
+What we call ``iterative sound-matching'' refers to works that mimic the back-and-forth process of listening to a synthesizer's output and updating the parameters until the termination conditions are met.
 The two necessary components are the \textit{loss function}, which determines how close two sounds are, and the \textit{parametric synthesizer} that outputs sounds based on input parameters.
@@
 
-Despite the early popularity of the iterative approach~\cite{justice1979analytic,horner1993machine,mitchell2007evolutionary}, recent works in sound-matching have favored inference models, deep neural-networks (\gls{DNN}), and large datasets.
-Rather than continually interacting with the environment, DNN-based models learn from a dataset of inputs (synthesizer parameters) and outputs (corresponding sound), and predict the synthesizer parameters for new sounds~\cite{yee2018automatic,han2023perceptual,han2024learning}.
-
-We believe that iterative sound-matching is interesting for multiple reasons.
+Despite the early popularity of the iterative approach~\cite{justice1979analytic,horner1993machine,mitchell2007evolutionary}, recent works in sound-matching have favored the supervised approach using inference models, deep neural-networks (\gls{DNN}), and large datasets.
+Rather than continually interacting with the environment, these models learn from a dataset of inputs (synthesizer parameters) and outputs (corresponding sound), and predict the synthesizer parameters for new sounds~\cite{yee2018automatic,han2023perceptual,han2024learning}.
+
+Iterative sound-matching is interesting for multiple reasons.
 First, it better resembles the human approach to sound-design, which requires continually interacting with the environment and making decisions according to a feedback signal~\cite{sutton2018reinforcement}.
 Second, the trial-and-error approach leaves a trail of parameter sets from the starting point to the end goal, which could be of use to sound designers.
-Third, relative to the DNN-based inference models, it provides a more tractable environment for exploring the complex issue of sound-similarity functions, and their interactions with different methods of synthesis.
+Third, relative to the DNN-based inference models, it provides a more tractable environment with room for creative approaches to sound-similarity definitions.
 
@@
 The experiments in this work pertain to two core issues, which we named \LossSelect~and \SynthSelect.
-The former refers to the search for an optimal loss function, lack of novel algorithms, and the---perhaps needless---focus on outperforming state of the art (\gls{SOTA}), while the latter refers to the lack of diversity in synthesis methods, and their interaction with the loss function.
+The former refers to the search for an optimal loss function, lack of novel algorithms, and the---perhaps needless---focus on outperforming state of the art (\gls{SOTA}), while the latter refers to the lack of diversity in synthesis methods.
 
@@
 We hypothesize that the performance of a similarity measure is influenced by other factors in the environment, particularly the method of synthesis.
-We combine 4 loss functions with 4 differentiable synthesis methods, and compare their performance (as measured by the similarity of the final output or parameters to the target or target parameters).
+We combine 4 loss functions with 4 differentiable synthesizers, and compare their performance (as measured by the similarity of the final output or parameters to the target or target parameters).
 Our experiments show that when accounting for different synthesizers, there is no consistent choice of sound similarity measure for optimal performance in sound-matching.
 In some experiments, \DTWEnv, a novel loss function that only looks at envelope similarity, matches or outperforms commonly used methods of audio comparison.
-These results imply that rather than aiming to outperform the SOTA and finding a universal best performer, more effort should be placed towards diversifying the existing similarity measures of sound as well as conducting experiments using a variety of synthesis methods.
+These results imply that rather than aiming to outperform the SOTA and finding a universal best performer, more effort could be placed towards diversifying the existing similarity measures of sound as well as conducting experiments using a variety of synthesis methods.
 Sound design is an inherently creative endeavor, and would likely benefit from a diversified range of solutions for different approaches.
@@
 \subsection{Differentiable Sound-Matching}
-An important caveat is that this work takes a \textit{differentiable}, machine learning (\gls{ML}) inspired approach to iterative sound-matching.
+This work takes a \textit{differentiable} approach to iterative sound-matching, a natural extension of early iterative sound-matching combined with recent advances in Machine Learning (\gls{ML}) software~\cite{braun2024dac,paszke2019pytorch} and increase computational power~\cite{moore2012introduction,schaller1997moore}.
 A differentiable approach requires differentiable loss and synthesizer functions, allowing the synthesizer parameters to be updated using gradient descent~\cite{goodfellow2016deep,boyd2004convex}.
 Ideally, these updates are in a direction that improves audio similarity by reducing loss.
-Differentiable iterative sound-matching has only recently become viable thanks to continued advances in both computation hardware~\cite{moore2012introduction,schaller1997moore}, and machine learning software~\cite{braun2024dac,paszke2019pytorch}.
 
@@
 Differentiable settings have the benefit of making more directed updates to the synthesizer parameters, but have major drawbacks.
-Other than requiring more computation power, differentiable functions require careful implementation of operations that are continuous, numerically stable, and of course, differentiable.
+Other than requiring more computation power, differentiable functions require careful implementation of operations that are continuous and numerically stable; this makes the implementation of signal processing functions quite difficult, which explains the scarcity of works in this domain.
 
@@
 \hline
-Masuda \textit{\textit{et al.}}~\cite{masuda2021soundmatch} & DDSP & P-Loss initially, fine-tune with Spec. & Supervised & Replicate? & In \& Out & 2021 & 1-shot \\
+Masuda \textit{\textit{et al.}}~\cite{masuda2021soundmatch} & DDSP & P-Loss initially, fine-tune with Spec. & Supervised & Both & In \& Out & 2021 & 1-shot \\
 \hline
@@
 To mitigate this, Han \textit{et al.}~\cite{han2023perceptual} introduced ``perceptual-neural-physical loss'' (PNP).
-PNP is an approximation of spectrogram loss functions; specifically, a loss function that uses the L2 norm of the difference between the features of two sounds, or $||\phi(t) - \phi(x)||^2_2$.
+PNP is an approximation of loss functions; specifically, a loss function that uses the L2 norm of the difference between the features of two sounds, or $||\phi(t) - \phi(x)||^2_2$, where $\phi$ could be a spectrogram or JTFS function.
 PNP loss functions are fast and differentiable, but require training and parameter estimation.
@@
 
-During the training phase, this metric is calculated alongside the training of the neural parameter estimator.
-After training, PNP can be used as a fast approximation of the loss function, $\|\phi(t) - \phi(x)\|_2^2 $.
-Han \textit{et al.} observed improvements to run-time (for testing) and slight reduction of the loss values under certain conditions.
-These experiments were based on supervised learning and physical modeling (of a drum head) for synthesis~\cite{smith2010physical}, and compared PNP, P-Loss, MSS, and JTFS functions~\cite{han2023perceptual}.
-This approach was more recently applied to other supervised settings using physical and FM synthesizers~\cite{han2024learning}\textcolor{red}{elaborate}.
+During the training phase, this metric is calculated alongside the neural parameter estimator.
+After training, PNP can be used as a fast approximation of computationally demanding loss functions such as JTFS~\cite{han2023perceptual,han2024learning}, paired with FM synthesizers and a differentiable physical model of a drum head~\cite{smith2010physical}.
 
@@
 
-\section{Summary and Conclusion}
+\section{Summary, Weaknesses, and Conclusion}
 \label{sec:summary_conclusion}
- Sound-matching is an umbrella term for the algorithmic programming of audio synthesizers, often with the goal of assisting sound designers.
-Here we provided a history of sound-matching, and proposed ``differentiable iterative sound-matching'' as a natural extension of previous works and recent advances.
-
-%  We discussed areas of weakness in the field, including loss and synthesizer selection, periodic loss landscapes, and verification of performance measures with in-domain and out-of-domain sounds.
-The first two issues being the main target of the experiments in this work.
-
-% Using simple synthesizers, small sound domains, and a limited number of loss functions, previous works have been able to claim some loss functions or feature extractors for audio as ``better performing" for sound-matching (or other audio-related tasks)~\cite{vahidi2023mesostructures,turian2020sorry,masuda2021soundmatch,engel2020ddsp,han2023perceptual}.
-Is the search for a universally best-performing loss function a worthy endeavor, or perhaps the choice of loss function is domain dependent?
+\textbf{Conclusion:} Sound-matching is an umbrella term for the algorithmic programming of audio synthesizers, often with the goal of assisting sound designers.
+Here we provided a history of sound-matching, and discussed the importance of ``differentiable iterative sound-matching'' as a natural extension current literature.
 
@@
 We used bare-bones versions of STFT and JTFS with fixed parameters.
-We did not test complex synthesizers.
+We did not test complex synthesizers using parallel and sequential DSP functions.
 Arbitrary hyperparameters such as learning rate and max number of iterations were selected for the DL pipeline, and only the RMSProp optimizer was tested.
@@
 This simplifies the issues of measuring sound-similarity, but it is not a realistic scenario for practical sound-matching.
-We plan to address this issue in future works, as we will discuss in the next section.
+This problem is left for future work, as we will discuss in the next section.
 
 \label{sec:future}
-\textbf{Future Work: }The analysis of different loss functions sound-matching presented here shows that problem of periodic loss-landscapes is not limited to spectral representations using Fourier transforms~\cite{turian2020sorry,vahidi2023mesostructures,uzrad2024diffmoog,bruford2024synthesizer}.
+\textbf{Future Work: }The analysis of different loss functions sound-matching shows that problem of periodic loss-landscapes is not limited to spectral representations using Fourier transforms~\cite{turian2020sorry,vahidi2023mesostructures,uzrad2024diffmoog,bruford2024synthesizer}.
 We can postulate that this problem emerges due to the periodic nature of sound.
-The current trend in contemporary work appears to be the application of domain specific and computationally expensive loss functions~\cite{han2023perceptual,uzrad2024diffmoog}, use of large neural networks~\cite{hershey2017cnn,cramer2019look}, or complex ensemble methods~\cite{turian2022hear}.
-Such models are useful, but often intractable; furthermore, training such models still requires the definition of simpler loss functions, which emphasizes the need for further work on differentiable and expressive loss function implementations.
-
 With regard to navigation of loss function landscapes, an optimizer that is more aware of fluctuations in gradients would perhaps lead to better solutions than simple gradient descent.
@@
 
+The current trend in contemporary work appears to be the application of domain specific and computationally expensive loss functions~\cite{han2023perceptual,uzrad2024diffmoog}, use of large neural networks~\cite{hershey2017cnn,cramer2019look}, or complex ensemble methods~\cite{turian2022hear}.
+Such models are useful, but often intractable; furthermore, training such models still requires the definition of simpler loss functions, which emphasizes the need for further work on differentiable and expressive loss function implementations.
+
 Like nearly all previous work in sound-matching, the main measure of success here was the accurate \textit{replication} of sounds (or synth parameters) rather than imitation of sounds outside of the synthesizer's domain.

### 2025-05-02 main.tex
Updates from Overleaf

@@
 \usepackage{tikz}
+\usetikzlibrary{positioning,shapes.geometric}
 \usepackage{adjustbox}
@@
 \usepackage{caption}         
-% re-enable caption functionality
 \captionsetup[lstlisting]{justification=centering, singlelinecheck=false}
-
-\usetikzlibrary{positioning,shapes.geometric}
 
@@
 \definecolor{rank4}{HTML}{000000}
-
-\newcommand{\staronsquare}{
-    \tikz[baseline=0ex]{
-        % Blue square with increased size
-        \fill[blue] (0,0) rectangle (2ex,2ex);
-        % Yellow star character centered on the square, larger font size
-        \node at (1ex,0.8ex) {\textcolor{yellow}{\large *}};
-    }
-}
-
-\newcommand{\vertdashred}[1][0.3cm]{% Default height is 2cm
-    \tikz[baseline]{\draw[red, dashed,line width=0.5mm] (0,0) -- (0,#1);}
-}
-
 \newcommand{\SIMSESpec}{\texttt{SIMSE\_Spec}}
@@
 \newcommand{\DTWEnv}{\texttt{DTW\_Envelope}}
-
 \newcommand{\LossSelect}{\textbf{Loss Selection}}
@@
 \newcommand{\OutDomain}{\textbf{Out-Domain Generation}}
-
 \newcommand{\BPNoise}{\textbf{BP-Noise}}  
@@
 \begin{document}
-\title{Sound Similarity, Sound Matching, and Differentiable Audio Synthesizer Control}
+\title{Measuring Sound Similarity for Iterative Sound-Matching}
 
@@
 Manual sound design with a synthesizer is inherently iterative: an artist compares the synthesized output to a mental target, adjusts parameters, and repeats until satisfied.
-There are practically an infinite number of parameter combinations, and their effect on the output sound is hard to predict.
 Iterative sound-matching automates this workflow by continually programming a synthesizer under the guidance of a loss function that quantifies similarity to the target sound.
 Prior comparisons of loss functions have typically favored one metric over another, but only within narrow settings: limited synthesis methods, few loss types, often without blind listening tests.
-This leaves open whether a universally optimal loss exists, or the choice of loss remains a creative decision conditioned on the synthesis method and the sound designer's preferences.
+This leaves open whether a universally optimal loss exists, or the choice of loss remains a creative decision conditioned on the synthesis method and the sound designer's preference.
 We propose differentiable iterative sound-matching as the natural extension of the available literature, since it combines the manual approach to sound design with modern advances in machine learning.
@@
 Automatic sound design has been a subject of research for decades, with reduction of tinkering time and creation of ``interesting" sounds as the main motivators \cite{krekovic2019insights,turian2020sorry,horner1993machine,salimi2020make,esling2019flow,engel2020ddsp,mitchell2007evolutionary,shier2020spiegelib}.
-\textbf{Sound-matching} is a common form of automatic sound design explored in literature, where the goal is to find parameters for a synthesizer that replicate \textit{all} or \textit{some} of the characteristics of a target sound as best as possible, given some kind of similarity metric (or loss function)~\cite{horner1993machine,mitchell2007evolutionary}.
+\textit{Sound-matching} is a common form of automatic sound design explored in literature, where the goal is to find parameters for a synthesizer that replicate \textit{all} or \textit{some} of the characteristics of a target sound as best as possible, given some kind of similarity metric (or loss function)~\cite{horner1993machine,mitchell2007evolutionary}.
 Our analysis of past works shows that this general goal of ``sound-matching'' can be broken down in to very different subtypes, many of them rarely or never explored.
@@
 
-What we call ``iterative sound-matching'' refers to works that mimic the back-and-forth process of listening to a synthesizer's output and updating the parameters until the termination conditions are met.
+``iterative sound-matching'' refers to works that mimic the back-and-forth process of listening to a synthesizer's output and updating the parameters until the termination conditions are met.
 The two necessary components are the \textit{loss function}, which determines how close two sounds are, and the \textit{parametric synthesizer} that outputs sounds based on input parameters.
@@
 
-Despite the early popularity of the iterative approach~\cite{justice1979analytic,horner1993machine,mitchell2007evolutionary}, recent works in sound-matching have favored the supervised approach using inference models, deep neural-networks (\gls{DNN}), and large datasets.
+Despite the early popularity of the iterative approach~\cite{justice1979analytic,horner1993machine,mitchell2007evolutionary}, recent works in sound-matching have favored supervised learning using inference models, deep neural-networks (\gls{DNNs}), and large datasets.
 Rather than continually interacting with the environment, these models learn from a dataset of inputs (synthesizer parameters) and outputs (corresponding sound), and predict the synthesizer parameters for new sounds~\cite{yee2018automatic,han2023perceptual,han2024learning}.
@@
 
-With the goal of improving iterative sound-matching algorithms for sound designers, we identify a number of weaknesses in current literature (see Section~\ref{sec:lacking}).
-The experiments in this work pertain to two core issues, which we named \LossSelect~and \SynthSelect.
+We discuss the history of sound-matching, and provide a nomenclature for different subtypes of sound-matching tasks.
+A number of weaknesses in current literature (see Section~\ref{sec:lacking}) are identified.
+The experiments in this work focus on two of these core issues, referred to as  \LossSelect~and \SynthSelect.
 The former refers to the search for an optimal loss function, lack of novel algorithms, and the---perhaps needless---focus on outperforming state of the art (\gls{SOTA}), while the latter refers to the lack of diversity in synthesis methods.
@@
 Considering the interplay between the two problems, we design experiments that address them simultaneously.
-We hypothesize that the performance of a similarity measure is influenced by other factors in the environment, particularly the method of synthesis.
+Testing the hypothesis that the performance of a similarity measure is influenced by other factors in the environment, particularly the method of synthesis.
 We combine 4 loss functions with 4 differentiable synthesizers, and compare their performance (as measured by the similarity of the final output or parameters to the target or target parameters).
@@
 \subsection{Differentiable Sound-Matching}
-This work takes a \textit{differentiable} approach to iterative sound-matching, a natural extension of early iterative sound-matching combined with recent advances in Machine Learning (\gls{ML}) software~\cite{braun2024dac,paszke2019pytorch} and increase computational power~\cite{moore2012introduction,schaller1997moore}.
+It is important to note that this work takes a \textit{differentiable} approach to iterative sound-matching, a natural extension of early iterative sound-matching combined with recent advances in Machine Learning (\gls{ML}) software~\cite{braun2024dac,paszke2019pytorch} and increase in computational power~\cite{moore2012introduction,schaller1997moore}.
 A differentiable approach requires differentiable loss and synthesizer functions, allowing the synthesizer parameters to be updated using gradient descent~\cite{goodfellow2016deep,boyd2004convex}.
@@
 
- As we discuss in Section~\ref{sec:related_works}, nearly all past works pertaining to iterative sound-matching have used genetic algorithms (\gls{GA})~\cite{holland1992genetic} where parameters are randomly changed with hopes of increasing similarity.
-Differentiable settings have the benefit of making more directed updates to the synthesizer parameters, but have major drawbacks.
-Other than requiring more computation power, differentiable functions require careful implementation of operations that are continuous and numerically stable; this makes the implementation of signal processing functions quite difficult, which explains the scarcity of works in this domain.
+ Genetic algorithms (\gls{GA})~\cite{holland1992genetic} have been the predominant approach to iterative sound-matching, where parameters are randomly changed with hopes of increasing similarity.
+Differentiable settings have the benefit of making goal-oriented updates to the synthesizer parameters (the goal being the minimization of loss).
+There are also drawbacks: other than requiring more computation power, differentiable functions require careful implementation of operations that are continuous and numerically stable; this makes the implementation of signal processing functions quite difficult, contributing to the scarcity of works in this domain.
 
 \section{Background}
+\label{sec:background}
 In this Section, we give a formal definition of sound-matching and the necessary components.
 The \textit{type} of sound-matching problem is dependent on the implementation of these components and the source of target sounds.
-We give a summary of different possible implementations and goals for sound-matching, and provide brief examples of previous works where applicable.
-We also give a typology of sound-matching types and major perennial issues that need to be addressed in future works.
-Finally, we introduce our formal definition of \textit{iterative sound-matching} in Section~\ref{sec:iterative_sound_matching}.
-
-% Due to their pivotal role in our experiments, a discussion of similarity measures, or \textit{loss functions} is provided in Section~\ref{sec:loss_funcs}.
-A formal definition of loss functions and specific implementations of four loss functions are given.
-
-\label{sec:background}
+We give a summary of different possible component implementations and goals for sound-matching, and provide brief examples of previous works where applicable.
+These discussions lead to a nomenclature for types of sound-matching, as well as major perennial issues that need to be addressed in future works.
+Finally, we introduce our formal definition of \textit{iterative sound-matching}.
 
@@
 In other words, different synthesizer programs work best with different loss functions.
+Notably, we see that our DTW and SIMSE based differentiable loss functions (\DTWEnv{} and \SIMSESpec) can outperform what are regarded as the SOTA loss functions, although this is highly synthesizer dependent.
 

### 2025-05-05 main.tex
Updates from Overleaf

@@
 
-\author{Amir Salimi, Abram Hindle, Osmar R Zaiane}
+\author{Amir Salimi, Abram Hindle, Osmar R. Za{\"i}ane}
 
@@
 A proxy representation is the output of the function \( \phi \), which can be thought of as a feature extraction function that maps the sounds \( t \) and \( x \) to their respective representations. 
-The proportionality or distance metric $d$ has typically been the L1 distance~\cite{turian2020sorry}, calculated as the mean of the absolute difference between every point in the proxy representation~\cite{engel2020ddsp,vahidi2023mesostructures}:
+The proportionality or distance metric $d$ has typically been the L1 distance~\cite{turian2020sorry,richard2025model}, calculated as the mean of the absolute difference between every point in the proxy representation~\cite{engel2020ddsp,vahidi2023mesostructures}:
 \[

### 2025-05-07 main.tex
last edits

@@
 For each of the sixteen synthesizer–loss combinations, we ran 300 randomized sound-matching trials.
-We calculated parameter differences, spectrogram-distance metrics, and gathered blind listening scores.
-A post-hoc analysis shows that no single loss function dominates across all synthesis methods; instead, the optimal measure depends on the architecture at hand.
-These findings underscore the value of developing new similarity metrics tailored to specific synthesis techniques, rather than pursuing a one-size-fits-all solution.
+Performance is measured with parameter differences, spectrogram-distance metrics, and blind listening scores; a fair amount of consistency between the three performance measures is observed.
+A post-hoc analysis shows that each of these loss functions has the potential to be the most appropriate.
+These findings underscore the value of expanding the scope of sound-matching experiments, and developing new similarity metrics tailored to specific synthesis techniques, rather than pursuing one-size-fits-all solutions.
 \end{abstract}
@@
 \textit{Sound-matching} is a common form of automatic sound design explored in literature, where the goal is to find parameters for a synthesizer that replicate \textit{all} or \textit{some} of the characteristics of a target sound as best as possible, given some kind of similarity metric (or loss function)~\cite{horner1993machine,mitchell2007evolutionary}.
-Our analysis of past works shows that this general goal of ``sound-matching'' can be broken down in to very different subtypes, many of them rarely or never explored.
+Our analysis of past works shows that this general goal of ``sound-matching'' can be broken down in to very different subtypes, many of them rarely or never explored, despite their relevance to practical applications.
 This leads us to a typology of sound-matching tasks, and the importance of \textit{differentiable, iterative sound-matching} as the extension of current work.
@@
 Second, the trial-and-error approach leaves a trail of parameter sets from the starting point to the end goal, which could be of use to sound designers.
-Third, relative to the DNN-based inference models, it provides a more tractable environment with room for creative approaches to sound-similarity definitions.
+Third, relative to the DNN-based inference models, it provides a more tractable environment with room for creative approaches to sound-similarity definitions, and more amenable to deployment on different synthesizers as it does not require pre-training on large datasets.
 
@@
 We give a summary of different possible component implementations and goals for sound-matching, and provide brief examples of previous works where applicable.
-These discussions lead to a nomenclature for types of sound-matching, as well as major perennial issues that need to be addressed in future works.
-Finally, we introduce our formal definition of \textit{iterative sound-matching}.
+These discussions lead to a nomenclature for types of sound-matching, as well as major perennial issues in the field.
 
@@
 In past works the sound-matching problem has been tackled from a variety of perspectives as it involves a number of modular parts: the choice of synthesizer $g$, the target sounds of interest, the representation function $\phi$, and the heuristic for finding the optimal $\theta$.
-We discuss these components in the upcoming sections, and give a coherent typology of the sound-matching tasks in Section~\ref{sec:matching_types}.
-
-To give a formal definition of sound-matching, we expand the definition given by recent works~\cite{vahidi2023mesostructures,han2023perceptual}. The major components are as follows: 
+To give a formal definition of sound-matching, we expand the definition given by recent works~\cite{vahidi2023mesostructures,han2023perceptual}.
+The major components are as follows:
 \begin{itemize}
@@
 \begin{lstlisting}[caption={\BPNoise}, label={lst:program0}, language=Faust,
-                  float, floatplacement=htbp, xleftmargin=1em, xrightmargin=0.5em, firstnumber=0]
+                  float, floatplacement=!H, xleftmargin=1em, xrightmargin=0.5em, firstnumber=0, aboveskip=0em, belowskip=-1em]
 import("stdfaust.lib");
@@
 process = no.noise:fi.lowpass(3,lp_cut):fi.highpass(10,hp_cut);
+\end{lstlisting}
+
+\begin{lstlisting}[caption={\AddSineSaw}, label={lst:program1},language=Faust,float,floatplacement=!H,xleftmargin=1em,xrightmargin=0.5em,firstnumber=0,aboveskip=0em, belowskip=-1em]
+import("stdfaust.lib");
+saw_freq = hslider("saw_freq",800,20,1000,1);
+sine_freq = hslider("sine_freq",300,20,1000,1);
+sineOsc(f) = +(f/ma.SR) ~ ma.frac:*(2*ma.PI) : sin;
+sawOsc(f) = +(f/ma.SR) ~ ma.frac;
+process = sineOsc(sine_freq)+sawOsc(saw_freq);
+\end{lstlisting}
+
+\begin{lstlisting}[caption={\AmpMod}, label={lst:program2},language=Faust,float,floatplacement=!H,xleftmargin=1em,xrightmargin=0.5em,firstnumber=0,aboveskip=0em, belowskip=-1em]
+import("stdfaust.lib");
+amp = hslider("amp",0.5,0,5,0.01);
+modulator = hslider("modulator",0.5,0,4,0.01);
+sineOsc(f) = +(f/ma.SR) ~ ma.frac:*(2*ma.PI) : sin;
+process = no.noise*sineOsc(modulator)*amp;
+\end{lstlisting}
+
+\begin{lstlisting}[caption={\FMMod}, label={lst:program3},language=Faust,float,floatplacement=!H,xleftmargin=1em,xrightmargin=0.5em,firstnumber=0,aboveskip=0em, belowskip=-1em]
+import("stdfaust.lib");
+carrier = hslider("carrier",100,20,1000,1);
+amp = hslider("amp",6,1,20,1);
+sineOsc(f) = +(f/ma.SR) ~ ma.frac:*(2*ma.PI) : sin;
+sawOsc(f) = +(f/ma.SR) ~ ma.frac;
+process = sineOsc(amp)*sawOsc(carrier);
 \end{lstlisting}
@@
 
-\begin{lstlisting}[caption={\AddSineSaw}, label={lst:program1},language=Faust,float,floatplacement=TBH,xleftmargin=1em,xrightmargin=0.5em,firstnumber=0]
-import("stdfaust.lib");
-saw_freq = hslider("saw_freq",800,20,1000,1);
-sine_freq = hslider("sine_freq",300,20,1000,1);
-sineOsc(f) = +(f/ma.SR) ~ ma.frac:*(2*ma.PI) : sin;
-sawOsc(f) = +(f/ma.SR) ~ ma.frac;
-process = sineOsc(sine_freq)+sawOsc(saw_freq);
-\end{lstlisting}
-
 \AddSineSaw{} is an additive program that combines a saw and a sine wave function.
@@
 \label{sec:program2}
-\begin{lstlisting}[caption={\AmpMod}, label={lst:program2},language=Faust,float,floatplacement=TBH,xleftmargin=1em,xrightmargin=0.5em,firstnumber=0]
-import("stdfaust.lib");
-amp = hslider("amp",0.5,0,5,0.01);
-modulator = hslider("modulator",0.5,0,4,0.01);
-sineOsc(f) = +(f/ma.SR) ~ ma.frac:*(2*ma.PI) : sin;
-process = no.noise*sineOsc(modulator)*amp;
-\end{lstlisting}
 
@@
 \label{sec:program3}
-\begin{lstlisting}[caption={\FMMod}, label={lst:program3},language=Faust,float,floatplacement=TBH,xleftmargin=1em,xrightmargin=0.5em,firstnumber=0]
-import("stdfaust.lib");
-carrier = hslider("carrier",100,20,1000,1);
-amp = hslider("amp",6,1,20,1);
-sineOsc(f) = +(f/ma.SR) ~ ma.frac:*(2*ma.PI) : sin;
-sawOsc(f) = +(f/ma.SR) ~ ma.frac;
-process = sineOsc(amp)*sawOsc(carrier);
-\end{lstlisting}
+
 Relative to additive and subtractive synthesis, AM/FM synthesis are less common methods of sound design.
@@
 
-\begin{figure*}[htbp]
-    \centering
-    \begin{minipage}{0.03\textwidth}
-        \footnotesize
-        \raggedleft
-        \vspace{-.2cm}
-        SIMSE\\[0.6cm]
-        L1\\[0.65cm]
-        JTFS\\[0.65cm]
-        DTW
-    \end{minipage}%
-    \begin{minipage}{0.95\textwidth}
-        \centering
-        \subfloat[Likert Scores]{
-            \includegraphics[width=0.3\textwidth]{images/npsk_likert_0.png}
-            \label{fig:p0_hearing}
-        }\hspace{0.005\textwidth} % Reduced space between figures
-        \subfloat[Inverse P-Loss]{
-            \includegraphics[width=0.3\textwidth]{images/npsk_P-Loss_0.png}
-            \label{fig:p0_P-Loss}
-        }\hspace{0.005\textwidth} % Reduced space between figures
-        \subfloat[Inverse MSS]{
-            \includegraphics[width=0.3\textwidth]{images/npsk_MSS_0.png}
-            \label{fig:p0_MSS}
-        }
-    \end{minipage}
-    \vspace{-0.5em}
-    \caption{\BPNoise~bootstrapped distributions and ranks given by NPSK.}
-    \label{fig:npsk_p0}
-\end{figure*}
-\begin{figure*}[htbp]
-    \centering
-    \begin{minipage}{0.03\textwidth}
-        \footnotesize
-        \raggedleft
-        \vspace{-.2cm}
-        SIMSE\\[0.6cm]
-        L1\\[0.65cm]
-        JTFS\\[0.65cm]
-        DTW
-    \end{minipage}%
-    \begin{minipage}{0.95\textwidth}
-        \centering
-        \subfloat[Likert Scores]{
-            \includegraphics[width=0.3\textwidth]{images/npsk_likert_1.png}
-            \label{fig:p1_hearing}
-        }\hspace{0.005\textwidth} % Reduced space between figures
-        \subfloat[Inverse P-Loss]{
-            \includegraphics[width=0.3\textwidth]{images/npsk_P-Loss_1.png}
-            \label{fig:p1_P-Loss}
-        }\hspace{0.005\textwidth} % Reduced space between figures
-        \subfloat[Inverse MSS]{
-            \includegraphics[width=0.3\textwidth]{images/npsk_MSS_1.png}
-            \label{fig:p1_MSS}
-        }
-    \end{minipage}
-    \vspace{-0.5em}
-    \caption{\AddSineSaw~bootstrapped distributions and ranks given by NPSK.}
-    \label{fig:npsk_p1}
-\end{figure*}
-\begin{figure*}[htbp]
-    \centering
-    \begin{minipage}{0.03\textwidth}
-        \footnotesize
-        \raggedleft
-        \vspace{-.2cm}
-        SIMSE\\[0.6cm]
-        L1\\[0.65cm]
-        JTFS\\[0.65cm]
-        DTW
-    \end{minipage}%
-    \begin{minipage}{0.95\textwidth}
-        \centering
-        \subfloat[Likert Scores]{
-            \includegraphics[width=0.3\textwidth]{images/npsk_likert_2.png}
-            \label{fig:p2_hearing}
-        }\hspace{0.005\textwidth} % Reduced space between figures
-        \subfloat[Inverse P-Loss]{
-            \includegraphics[width=0.3\textwidth]{images/npsk_P-Loss_2.png}
-            \label{fig:p2_P-Loss}
-        }\hspace{0.005\textwidth} % Reduced space between figures
-        \subfloat[Inverse MSS]{
-            \includegraphics[width=0.3\textwidth]{images/npsk_MSS_2.png}
-            \label{fig:p2_MSS}
-        }
-    \end{minipage}
-    \vspace{-0.5em}
-    \caption{\AmpMod~bootstrapped distributions and ranks given by NPSK.}
-    \label{fig:npsk_p2}
-\end{figure*}
-\begin{figure*}[htbp]
-    \centering
-    \begin{minipage}{0.03\textwidth}
-        \footnotesize
-        \raggedleft
-        \vspace{-.2cm}
-        SIMSE\\[0.6cm]
-        L1\\[0.65cm]
-        JTFS\\[0.65cm]
-        DTW
-    \end{minipage}%
-    \begin{minipage}{0.95\textwidth}
-        \centering
-        \subfloat[Likert Scores]{
-            \includegraphics[width=0.3\textwidth]{images/npsk_likert_3.png}
-            \label{fig:p3_hearing}
-        }\hspace{0.005\textwidth} % Reduced space between figures
-        \subfloat[Inverse P-Loss]{
-            \includegraphics[width=0.3\textwidth]{images/npsk_P-Loss_3.png}
-            \label{fig:p3_P-Loss}
-        }\hspace{0.005\textwidth} % Reduced space between figures
-        \subfloat[Inverse MSS]{
-            \includegraphics[width=0.3\textwidth]{images/npsk_MSS_3.png}
-            \label{fig:p3_MSS}
-        }
-    \end{minipage}
-    \vspace{-0.5em}
-    \caption{\FMMod~bootstrapped distributions and ranks given by NPSK.}
-    \label{fig:npsk_p3}
-\end{figure*}
-
-\subsubsection{\AddSineSaw}
-As shown in Figure~\ref{fig:npsk_p1}, JTFS was the best performing loss function in all evaluation methods. Moreover, all evaluation methods produced identical results.
-
-\subsubsection{\AmpMod}
-IP-Loss and manual hearing tests yield identical rankings, and MSS results vary for for ranks 2 to 4.
-It is not surprising that MSS results differ from the hearing tests, as we saw in Table~\ref{tab:kruskal_auto}, MSS showed no significant differences between groups.
-As shown in Figure~\ref{fig:npsk_p2}, all models selected \DTWEnv~as the best performer.
-This may be due to \DTWEnv's focus on periodic changes in loudness.
-
-\subsubsection{\FMMod}
-\DTWEnv~is again the best performer here. As shown in Figure~\ref{fig:npsk_p3}, all methods of analysis gave identical rankings to all programs.
-
-\subsection{Consistency In Rankings}
-
-Automatic sound evaluation methods are unreliable, as we do not know how they would align with the often subjective opinion of human listeners.
-Here, the manual blinded rankings given by the reviewers could be treated as a ground truth, particularly since the two reviewers gave very similar rankings to the experiment results.
-To measure the agreement, Spearman's rank correlation is used.
-The Spearman test returns a p-value for the null hypothesis of no correlation, along with a $\rho$ value between -1 (perfect negative correlation) to 1 (perfect correlation), and middle value of 0 for no correlation~\cite{spearman1987proof,rebekic2015pearson}.
-For each program and loss function, the blind manual hearing test gives us 80 ranks, which are given by two of the authors to a set of 40 randomly selected experiment results.
-This gives 160 pairs of ranks per program, and 640 for all programs.
-We computed Spearman correlations between the two reviewers.
-Across all programs, the correlation was very strong ($\rho = 0.86$, $p < 10^{-180}$).
-Per-program correlations were similarly strong and statistically significant: $\rho = 0.71$ for \BPNoise{} ($p < 10^{-25}$), $\rho = 0.64$ for \AddSineSaw{} ($p < 10^{-19}$), $\rho = 0.84$ for \AmpMod{} ($p < 10^{-43}$), and $\rho = 0.85$ for \FMMod{} ($p < 10^{-44}$).
-
-Table~\ref{tab:combined_ranks} shows the rankings for each program, with ranks different from the corresponding hearing test results marked with an asterisk.
-Both MSS and P-Loss gave a different rank from the hearing results in 3 out of 16 cases, which---somewhat surprisingly---shows consistency between automatic and manual hearing tests, at least for the simple programs used in this work.
-We also observe that identical top ranks were consistent across performance evaluation methodologies, with the exception of P-Loss in \BPNoise, which narrowly picks a different spectrogram-based loss function.
-% \input{extra_figures/ranking_tables}
 \begin{table*}[htbp]
@@
 
+\input{extra_figures/snpk_figures}
+
+\subsubsection{\AddSineSaw}
+As shown in Figure~\ref{fig:npsk_p1}, JTFS was the best performing loss function in all evaluation methods. Moreover, all evaluation methods produced identical results.
+
+\subsubsection{\AmpMod}
+IP-Loss and manual hearing tests yield identical rankings, and MSS results vary for for ranks 2 to 4.
+It is not surprising that MSS results differ from the hearing tests, as we saw in Table~\ref{tab:kruskal_auto}, MSS showed no significant differences between groups.
+As shown in Figure~\ref{fig:npsk_p2}, all models selected \DTWEnv~as the best performer.
+This may be due to \DTWEnv's focus on periodic changes in loudness.
+
+\subsubsection{\FMMod}
+\DTWEnv~is again the best performer here. As shown in Figure~\ref{fig:npsk_p3}, all methods of analysis gave identical rankings to all programs.
+
+\subsection{Consistency In Rankings}
+
+Automatic sound evaluation methods are unreliable, as we do not know how they would align with the often subjective opinion of human listeners.
+Here, the manual blinded rankings given by the reviewers could be treated as a ground truth, particularly since the two reviewers gave very similar rankings to the experiment results.
+To measure the agreement, Spearman's rank correlation is used.
+The Spearman test returns a p-value for the null hypothesis of no correlation, along with a $\rho$ value between -1 (perfect negative correlation) to 1 (perfect correlation), and middle value of 0 for no correlation~\cite{spearman1987proof,rebekic2015pearson}.
+For each program and loss function, the blind manual hearing test gives us 80 ranks, which are given by two of the authors to a set of 40 randomly selected experiment results.
+This gives 160 pairs of ranks per program, and 640 for all programs.
+We computed Spearman correlations between the two reviewers.
+Across all programs, the correlation was very strong ($\rho = 0.86$, $p < 10^{-180}$).
+Per-program correlations were similarly strong and statistically significant: $\rho = 0.71$ for \BPNoise{} ($p < 10^{-25}$), $\rho = 0.64$ for \AddSineSaw{} ($p < 10^{-19}$), $\rho = 0.84$ for \AmpMod{} ($p < 10^{-43}$), and $\rho = 0.85$ for \FMMod{} ($p < 10^{-44}$).
+
+Table~\ref{tab:combined_ranks} shows the rankings for each program, with ranks different from the corresponding hearing test results marked with an asterisk.
+Both MSS and P-Loss gave a different rank from the hearing results in 3 out of 16 cases, which---somewhat surprisingly---shows consistency between automatic and manual hearing tests, at least for the simple programs used in this work.
+We also observe that identical top ranks were consistent across performance evaluation methodologies, with the exception of P-Loss in \BPNoise, which narrowly picks a different spectrogram-based loss function.
+% \input{extra_figures/ranking_tables}
+
 \subsection{Takeaways}
@@
 
-We also see that the novel \SIMSESpec{} and \DTWEnv{} loss functions (to our knowledge, not previously used in iterative sound-matching), lead to the best results in three of the four programs.
+We also see that the \DTWEnv{} and \SIMSESpec{} loss functions (to our knowledge, not previously used in iterative sound-matching), lead to the best results in three of the four programs.
 This shows room for further creative approaches to differentiable sound-similarity measures, particularly since we observed that SOTA loss functions can be ineffective depending on the design of a synthesizer.
@@
 \label{sec:summary_conclusion}
-\textbf{Conclusion:} Sound-matching is an umbrella term for the algorithmic programming of audio synthesizers, often with the goal of assisting sound designers.
-Here we provided a history of sound-matching, and discussed the importance of ``differentiable iterative sound-matching'' as a natural extension current literature.
+\textbf{Summary:} Sound-matching is an umbrella term for the algorithmic programming of audio synthesizers, often with the goal of assisting sound designers.
+Here we provided a history of sound-matching, major issues in the field, and discussed the importance of ``differentiable iterative sound-matching'' as a natural extension current literature.
 
@@
 In other words, different synthesizer programs work best with different loss functions.
-Notably, we see that our DTW and SIMSE based differentiable loss functions (\DTWEnv{} and \SIMSESpec) can outperform what are regarded as the SOTA loss functions, although this is highly synthesizer dependent.
+Notably, we see that our novel use of DTW and SIMSE based differentiable loss functions (\DTWEnv{} and \SIMSESpec) can outperform what are regarded as the SOTA loss functions, although this is highly synthesizer dependent.
 
@@
 \label{sec:future}
-\textbf{Future Work: }The analysis of different loss functions sound-matching shows that problem of periodic loss-landscapes is not limited to spectral representations using Fourier transforms~\cite{turian2020sorry,vahidi2023mesostructures,uzrad2024diffmoog,bruford2024synthesizer}.
-We can postulate that this problem emerges due to the periodic nature of sound.
+\textbf{Future Work: }The analysis of different loss functions in sound-matching shows that problem of periodic loss-landscapes is not limited to spectral representations using Fourier transforms~\cite{turian2020sorry,vahidi2023mesostructures,uzrad2024diffmoog,bruford2024synthesizer}.
+Perhaps this problem emerges due to the periodic nature of sound.
 With regard to navigation of loss function landscapes, an optimizer that is more aware of fluctuations in gradients would perhaps lead to better solutions than simple gradient descent.

### 2025-05-08 main.tex
basically nothing

@@
 \captionsetup[lstlisting]{justification=centering, singlelinecheck=false}
-
 \providecommand{\gls}[1]{#1}
@@
 \begin{document}
-\title{Measuring Sound Similarity for Iterative Sound-Matching}
+\title{Evaluating Sound Similarity Metrics for Iterative Sound-Matching}
 
@@
 
-% The paper headers
 \markboth{IEEE TRANSACTIONS ON AUDIO, SPEECH AND LANGUAGE PROCESSING, VOL -, 2025}%
@@
 
-``iterative sound-matching'' refers to works that mimic the back-and-forth process of listening to a synthesizer's output and updating the parameters until the termination conditions are met.
+``Iterative sound-matching'' refers to works that mimic the back-and-forth process of listening to a synthesizer's output and updating the parameters until the termination conditions are met.
 The two necessary components are the \textit{loss function}, which determines how close two sounds are, and the \textit{parametric synthesizer} that outputs sounds based on input parameters.

### 2025-05-12 extra_figures/evaluation_graph.tex
Updates from Overleaf
first half of abram feedback

@@
     gaussian/.style={draw, smooth, samples=100, domain=-2.5:2.5},
-    snpk/.style={rectangle, draw, rounded corners, minimum width=2.5cm, minimum height=1cm, text centered, node distance=1.5cm, fill=blue!20},
+    NPSK/.style={rectangle, draw, rounded corners, minimum width=2.5cm, minimum height=1cm, text centered, node distance=1.5cm, fill=blue!20},
     level 1/.style={sibling distance=6cm, level distance=1cm},
@@
 \node at (0,-4) {\textbf{$\longleftarrow$\ \ \large Bootstrapped Distributions\ \ $\longrightarrow$}};
-% Left side: SNPK with MSS
+% Left side: NPSK with MSS
 
-% Left side: SNPK with MSS
-\node[snpk, text height=5ex, align=center, fill=red!30, draw=black] at (-6,-6) {SNPK with 4 \\ MSS distributions};
+% Left side: NPSK with MSS
+\node[NPSK, text height=5ex, align=center, fill=red!30, draw=black] at (-6,-6) {NPSK with 4 \\ MSS distributions};
 \draw[->] (-6,-6.5) -- (-6,-7.5) node[below, below] {\textbf{Ranks 1-4}};
-\draw[red!40, thick] (-4.5, -5.4) -- (-2.75, -4.75);  % Line 1
-\draw[red!40, thick] (-5.25, -5.4) -- (-3.5, -4.75);  % Line 2
-\draw[red!40, thick] (-5.75, -5.4) -- (-5.5, -4.75);  % Line 3
-\draw[red!40, thick] (-6.25, -5.4) -- (-9, -4.75);    % Line 4
+\draw[red!40, thick] (-5, -5.4) -- (-2.95, -4.95);  % Line 1
+\draw[red!40, thick] (-5.5, -5.4) -- (-3.5, -4.75);  % Line 2
+\draw[red!40, thick] (-6, -5.4) -- (-5.5, -4.75);  % Line 3
+\draw[red!40, thick] (-6.5, -5.4) -- (-9, -4.75);    % Line 4
 
-% Middle: SNPK with P-Loss
-\node[snpk, text height=5ex, align=center, fill=blue!40, draw=black] at (0,-6) {SNPK with 4 \\ P-Loss distributions};
+% Middle: NPSK with P-Loss
+\node[NPSK, text height=5ex, align=center, fill=blue!40, draw=black] at (0,-6) {NPSK with 4 \\ P-Loss distributions};
 \draw[->] (0,-6.5) -- (0,-7.5) node[below, below] {\textbf{Ranks 1-4}};
@@
 
-% Right side: SNPK with Likert
-\node[snpk, text height=5ex, align=center, fill=green!30, draw=black] at (6,-6) {SNPK with 4 \\ Likert distributions};
+% Right side: NPSK with Likert
+\node[NPSK, text height=5ex, align=center, fill=green!30, draw=black] at (6,-6) {NPSK with 4 \\ Likert distributions};
 \draw[->] (6,-6.5) -- (6,-7.5) node[below, below] {\textbf{Ranks 1-4}};
@@
 \draw[green!50, thick] (6.75, -5.4) -- (6, -4.75);    % Line 2
-\draw[green!50, thick] (6.25, -5.4) -- (4, -4.75);    % Line 3
-\draw[green!50, thick] (5.75, -5.4) -- (3, -4.75);    % Line 4
+\draw[green!50, thick] (5.75, -5.4) -- (4, -4.75);    % Line 3
+\draw[green!50, thick] (4.95, -5.4) -- (3.25, -4.95);    % Line 4
 

### 2025-05-12 main.tex
first half of abram feedback

@@
 This approach goes beyond mere replication and uses synthesizers for artistic re-interpretation of sounds.
-Pursuing this idea, we aim to automate this process using digital audio synthesizers.
+The motivation behind this work is the automation of this ``interpretive'' approach to sound-design with digital audio synthesizers; however, not only is the ``interpretive'' aspect of automatic sound-design rarely explored, but there exists major gaps in knowledge in literature regarding the much simpler problem of ``replicating'' sounds.
+To that end, this work explores the variation in the effectiveness of different sound-similarity measures when replicating sounds with different synthesizers.
 
@@
 Since the advent of digital signal processing (\gls{DSP}) in the 1960s~\cite{stranneby2004digital}, a wide variety of parametric audio generation functions have been proposed~\cite{lyons1997understanding,russ1999sound,shier2020spiegelib}.
-Manual sound design with a synthesizer is an iterative process of trial and error; when the artist has a desired (or target) sound in mind, sound design involves listening to the output, determining whether the output is closer or farther from the target, and adjusting the parameters accordingly until the target is reached~\cite{russ1999sound}.
+Manual sound design with a synthesizer is an iterative process of trial and error; when the artist has a desired (or target) sound in mind, sound design involves listening to the output, determining whether the output is closer or farther from the target, and adjusting the parameters of the synthesizer accordingly until the target is reached~\cite{russ1999sound}.
 This paper explores the automation of this process, which we call \textit{iterative sound-matching}.
@@
 Automatic sound design has been a subject of research for decades, with reduction of tinkering time and creation of ``interesting" sounds as the main motivators \cite{krekovic2019insights,turian2020sorry,horner1993machine,salimi2020make,esling2019flow,engel2020ddsp,mitchell2007evolutionary,shier2020spiegelib}.
-\textit{Sound-matching} is a common form of automatic sound design explored in literature, where the goal is to find parameters for a synthesizer that replicate \textit{all} or \textit{some} of the characteristics of a target sound as best as possible, given some kind of similarity metric (or loss function)~\cite{horner1993machine,mitchell2007evolutionary}.
+\textit{Sound-matching} is a common form of automatic sound design explored in literature, where a similarity metric (or loss function) is given, and the goal is to find parameters for a synthesizer that replicate \textit{all} or \textit{some} of the characteristics of a target sound as best as possible~\cite{horner1993machine,mitchell2007evolutionary}.
 Our analysis of past works shows that this general goal of ``sound-matching'' can be broken down in to very different subtypes, many of them rarely or never explored, despite their relevance to practical applications.
-This leads us to a typology of sound-matching tasks, and the importance of \textit{differentiable, iterative sound-matching} as the extension of current work.
+This leads us to a typology of sound-matching tasks, and the importance of \textit{differentiable, iterative sound-matching}, rarely explored in previous literature~\cite{vahidi2023mesostructures}.
 
@@
 
+This work takes a \textit{differentiable} approach to iterative sound-matching, a natural extension of early works in iterative sound-matching combined with recent advances in Machine Learning (\gls{ML}) software~\cite{braun2024dac,paszke2019pytorch} and increase in computational power~\cite{moore2012introduction,schaller1997moore}.
+A differentiable approach requires differentiable loss and synthesizer functions, allowing the synthesizer parameters to be updated using gradient descent~\cite{goodfellow2016deep,boyd2004convex}.
+
 We discuss the history of sound-matching, and provide a nomenclature for different subtypes of sound-matching tasks.
@@
 Testing the hypothesis that the performance of a similarity measure is influenced by other factors in the environment, particularly the method of synthesis.
-We combine 4 loss functions with 4 differentiable synthesizers, and compare their performance (as measured by the similarity of the final output or parameters to the target or target parameters).
+We combine 4 loss functions with 4 differentiable synthesizers, and compare their performance as measured by the similarity of the final output or parameters to the target or target parameters.
+We measure this final similarity with two different automatic methods, as well as manual listening tests conducted by two of the authors.
+
 Our experiments show that when accounting for different synthesizers, there is no consistent choice of sound similarity measure for optimal performance in sound-matching.
@@
 
-\subsection{Differentiable Sound-Matching}
-It is important to note that this work takes a \textit{differentiable} approach to iterative sound-matching, a natural extension of early iterative sound-matching combined with recent advances in Machine Learning (\gls{ML}) software~\cite{braun2024dac,paszke2019pytorch} and increase in computational power~\cite{moore2012introduction,schaller1997moore}.
-A differentiable approach requires differentiable loss and synthesizer functions, allowing the synthesizer parameters to be updated using gradient descent~\cite{goodfellow2016deep,boyd2004convex}.
-Ideally, these updates are in a direction that improves audio similarity by reducing loss.
-
- Genetic algorithms (\gls{GA})~\cite{holland1992genetic} have been the predominant approach to iterative sound-matching, where parameters are randomly changed with hopes of increasing similarity.
-Differentiable settings have the benefit of making goal-oriented updates to the synthesizer parameters (the goal being the minimization of loss).
-There are also drawbacks: other than requiring more computation power, differentiable functions require careful implementation of operations that are continuous and numerically stable; this makes the implementation of signal processing functions quite difficult, contributing to the scarcity of works in this domain.
+% \subsection{Differentiable Sound-Matching}
 
@@
 In this Section, we give a formal definition of sound-matching and the necessary components.
-The \textit{type} of sound-matching problem is dependent on the implementation of these components and the source of target sounds.
-We give a summary of different possible component implementations and goals for sound-matching, and provide brief examples of previous works where applicable.
 These discussions lead to a nomenclature for types of sound-matching, as well as major perennial issues in the field.
@@
 In replication, the goal is to make an identical copy of the target sound.
-With imitation, the goal is to make new sounds that only retain a subset of the target's sonic features.
+Imitation is an artistic pursuit, and harder to define, where the goal is to make new sounds that only retain a subset of the target's sonic features.
 While closely related to the in-domain versus out-of-domain problem, the choice of replication versus imitation is more dependent on how the loss and representation functions are defined.
@@
 The most fit members of the group are the parameters which perform the best in the loss function, and create a new generation of parameters via mutation (random change in subset of parameters) and crossovers (combination of parameter sets); this process repeats until the goal or a maximum number of generations is reached.
+
+ Differentiable settings have the benefit of making goal-oriented updates to the synthesizer parameters (the goal being the minimization of loss).
+There are also drawbacks: other than requiring more computation power, differentiable functions require careful implementation of operations that are continuous and numerically stable; this makes the implementation of signal processing functions quite difficult, contributing to the scarcity of works in this domain.
 
@@
 
-Having looked at the current available literature, we note four major areas of weakness in sound-matching:
+Having looked at the current available literature, we note four major areas of weakness in sound-matching.
+We target the first two issues in this work, while the latter two issues are left for future work.
 \begin{enumerate}
@@
 \end{enumerate}
-
-In Section~\ref{sec:progs_methodology_results}, we discuss our work in simultaneously bridging the knowledge gap with regards to the \LossSelect~and \SynthSelect~problems.
-The latter two issues are left for future work; possible approaches are discussed in Sections~\ref{sec:summary_conclusion} and~\ref{sec:future}.
 
@@
 
-For each program, the loss functions are ranked using the non-parametric Scott-Knott (\gls{SNPK}) test~\cite{tantithamthavorn2017mvt,tantithamthavorn2018optimization}, which ranks the loss functions from rank 1 (best, or highest distribution values) to maximum rank of 4 (worst, with lowest distribution values).
-SNPK may determine that multiple functions belong to the same rank, in which case the maximum rank would be less than 4.
+For each program, the loss functions are ranked using the non-parametric Scott-Knott (\gls{NPSK}) test~\cite{tantithamthavorn2017mvt,tantithamthavorn2018optimization}, which ranks the loss functions from rank 1 (best, or highest distribution values) to maximum rank of 4 (worst, with lowest distribution values).
+NPSK may determine that multiple functions belong to the same rank, in which case the maximum rank would be less than 4.
 Lower P-Loss and MSS values indicate higher similarity, therefore (for consistency with manually assigned ranks) the inverse of their output is used, where higher values indicate greater similarity.
@@
 The first stage determines whether there is a difference in group means, with the null hypothesis being that all groups have similar mean ranks.
-If the null hypothesis is rejected, the second stage ranks the loss functions from best to worst using the non-parametric Scott-Knott algorithm (\gls{SNPK})~\cite{tantithamthavorn2017mvt,tantithamthavorn2018optimization}.
+If the null hypothesis is rejected, the second stage ranks the loss functions from best to worst using the non-parametric Scott-Knott algorithm (\gls{NPSK})~\cite{tantithamthavorn2017mvt,tantithamthavorn2018optimization}.
 The Kruskal-Wallis test is used for the first stage~\cite{kruskal1952use}.
@@
 \subsubsection{Best Performers For Each Program}
-Based on the bootstrapped distribution of scores, the SNPK algorithm ranks the loss functions from 1 (best) to a maximum of 4 (worst).
+Based on the bootstrapped distribution of scores, the NPSK algorithm ranks the loss functions from 1 (best) to a maximum of 4 (worst).
 In cases where the distributions are similar, multiple loss functions can be clustered into the same rank.
@@
 
-\input{extra_figures/snpk_figures}
+\input{extra_figures/npsk_figures}
 

### 2025-05-12 main.tex
Updates from Overleaf

@@
 \label{sec:progs_methodology_results}
-Towards addressing the first two problems, we compare the performance iterative sound-matching experiments, where four differentiable loss functions are applied to four differentiable synthesizers.
-What we mean by performance is the similarity of the final output to the target, as measured by P-Loss, MSS, and manual listening.
+Towards addressing the problems of \LossSelect{} and \SynthSelect{}, we compare the performance of iterative sound-matching experiments, where four differentiable loss functions are applied to four differentiable synthesizers.
+What we mean by performance is the similarity of the final output to the target, as measured by P-Loss, MSS, and manual listening tests, which we treat as the ground truth.
 
-\textbf{The hypothesis here is that the best performing loss function for sound-matching is influenced by the method of synthesis}.
+\textbf{The hypothesis is that the relative performance of loss functions for sound-matching is influenced by the method of synthesis}.
 The results of these experiments show that there is no universally optimal loss function, strengthening the idea that the selection of the loss function and synthesis method is not an optimization problem, but a subjective choice depending on the needs of the sound designer.
 
-In Section~\ref{sec:methodology} we lay out the experiment methodology.
+% In Section~\ref{sec:methodology} we lay out the experiment methodology.
 Section~\ref{sec:loss_implementation} discusses the details of implementation for the loss functions.
@@
 We test four different loss functions and compare their success in sound-matching using four different synthesizers.
-We verify the success of the sound-matching process using two automatic measures of sound similarity (MSS and P-Loss) and a manual blinded hearing test conducted by two of the authors where likert scores~\cite{jebb2021review} of 1-5 are assigned to a subset of targets and outputs.
+We verify the success of the sound-matching process using two automatic measures of sound similarity (MSS and P-Loss) and a manual blinded hearing test conducted by two of the authors where likert scores~\cite{jebb2021review} of 1-5 are assigned to a subset of final outputs and the corresponding targets.
 
@@
 This simplified example leads us to our methodology which has four loss functions, four synthesizer programs, and three different methods of scoring sound similarity.
+
+\subsubsection{Evaluation Methods}
 
@@
 The first stage determines whether there is a difference in group means, with the null hypothesis being that all groups have similar mean ranks.
-If the null hypothesis is rejected, the second stage ranks the loss functions from best to worst using the non-parametric Scott-Knott algorithm (\gls{NPSK})~\cite{tantithamthavorn2017mvt,tantithamthavorn2018optimization}.
+If the null hypothesis is rejected, the second stage ranks the loss functions from best to worst using the non-parametric Scott-Knott test(\gls{NPSK})~\cite{tantithamthavorn2017mvt,tantithamthavorn2018optimization}.
 The Kruskal-Wallis test is used for the first stage~\cite{kruskal1952use}.
@@
 Using the significance level of 0.05, if the H statistic is greater than the critical value of the chi-square distribution then the null hypothesis is rejected, meaning that at least one group has a mean rank significantly above others~\cite{kruskal1952use}.
-Table~\ref{tab:kruskal_auto} shows the Kruskal-Wallis results for every program and evaluation method, and shows that significant differences exist in all cases except MSS for the \AmpMod{} program.
+Table~\ref{tab:kruskal_auto} shows the Kruskal-Wallis results for every program and evaluation method, and shows that significant differences between loss function performance measures exist in all cases except MSS for the \AmpMod{} program.
 
@@
 \centering
-\caption{Kruskal-Wallis results by program and evaluation method.}
+\caption{Kruskal-Wallis results by program and evaluation method. Bootstrapped distributions of performance measures (or ranks for manual listening tests) are compared.}
 \begin{tabular}{|c|c|c|c|c|}

### 2025-05-13 main.tex
Updates from Overleaf

@@
 \begin{document}
-\title{Evaluating Sound Similarity Metrics for Iterative Sound-Matching}
+\title{Evaluating Sound Similarity Metrics for Differentiable, Iterative Sound-Matching}
 
@@
 \subsection{Digital Synthesis With Sound-Matching}
-Since the 1970s, artists such as Suzanne Ciani~\cite{ciani_life_in_waves} have explored the use of audio synthesizers to replicate natural sounds—such as a bottle cap popping off—in a way that preserves the essential qualities of the original while showcasing the characteristics of the synthesizer.
-This approach goes beyond mere replication and uses synthesizers for artistic re-interpretation of sounds.
-The motivation behind this work is the automation of this ``interpretive'' approach to sound-design with digital audio synthesizers; however, not only is the ``interpretive'' aspect of automatic sound-design rarely explored, but there exists major gaps in knowledge in literature regarding the much simpler problem of ``replicating'' sounds.
-To that end, this work explores the variation in the effectiveness of different sound-similarity measures when replicating sounds with different synthesizers.
+Since the 1970s, artists such as Suzanne Ciani~\cite{ciani_life_in_waves} have explored the use of audio synthesizers to replicate natural sounds—such as a bottle cap popping off—in ways that preserve the essential qualities of the original while showcasing the unique characteristics of the synthesizer.
+This approach goes beyond mere replication, using synthesizers for artistic reinterpretation of sounds.
+However, the automation of this interpretive approach to sound design using digital audio synthesizers has been rarely explored, and significant knowledge gaps remain even in the simpler problem of replicating sounds.
+One such gap is the lack of diversity and depth in the analysis of sound-similarity measures and their interaction with various synthesis methods.
+As a stepping stone toward automated interpretive sound design, we investigate whether the optimal choice of sound-similarity measure depends on the synthesis method.
 
@@
 Considering the interplay between the two problems, we design experiments that address them simultaneously.
-Testing the hypothesis that the performance of a similarity measure is influenced by other factors in the environment, particularly the method of synthesis.
+Testing the hypothesis that the performance of a similarity measure (or loss function) is influenced by other factors in the environment, particularly the method of synthesis.
 We combine 4 loss functions with 4 differentiable synthesizers, and compare their performance as measured by the similarity of the final output or parameters to the target or target parameters.
-We measure this final similarity with two different automatic methods, as well as manual listening tests conducted by two of the authors.
+We evaluate the final similarity with two different automatic methods, as well as manual listening tests conducted by two of the authors.
+We compare the distributions of the evaluation scores in order to rank the loss functions from best to worst.
+The results of these experiments show that when accounting for different synthesizers, there is no consistent choice of sound similarity measure for optimal performance in sound-matching.
 
-Our experiments show that when accounting for different synthesizers, there is no consistent choice of sound similarity measure for optimal performance in sound-matching.
-In some experiments, \DTWEnv, a novel loss function that only looks at envelope similarity, matches or outperforms commonly used methods of audio comparison.
+We note that for two synthesizers, our novel use of ``dynamic time warping of envelopes'' as a loss function for differentiable sound-matching outperforms the standard methods of audio comparison.
 These results imply that rather than aiming to outperform the SOTA and finding a universal best performer, more effort could be placed towards diversifying the existing similarity measures of sound as well as conducting experiments using a variety of synthesis methods.
@@
 
-\textbf{The hypothesis is that the relative performance of loss functions for sound-matching is influenced by the method of synthesis}.
+\textbf{We hypothesize that the relative performance of loss functions for sound-matching is influenced by the method of synthesis}.
 The results of these experiments show that there is no universally optimal loss function, strengthening the idea that the selection of the loss function and synthesis method is not an optimization problem, but a subjective choice depending on the needs of the sound designer.
@@
 Having 300 experiments gives us a distribution of losses to determine whether $L_1$ or $L_2$ led to better performance.
-This simplified example leads us to our methodology which has four loss functions, four synthesizer programs, and three different methods of scoring sound similarity.
+This simplified example leads us to our methodology which has four loss functions, four synthesizer programs, and three different evaluation methods of scoring sound similarity.
 
 \subsubsection{Evaluation Methods}
+The two automatic evaluation methods used here are P-Loss and MSS, as described in Section~\ref{sec:loss_funcs}.
+For P-Loss, the parameters are normalized between 0-1 based on the valid ranges defined in the Faust program, and the L1 distance is calculated.
+MSS is computed using a fixed FFT size of $512$ and hop size of $100$, with multiple window lengths $\{512, 1024, 2048, 4096\}$.
+
+Automatic sound evaluation methods are unreliable, as we do not know how they would align with the (often subjective) opinion of human listeners, for this reason, listening tests are conducted by taking a random sample of 40 experiment results for each program and loss function pair; followed by the blinded assignment of scores of 1-5 on a Likert scale (1 for no similarity, 5 for identical)~\cite{jebb2021review} by two of the authors.
+The agreement between the assigned scores is measured using Spearman's rank correlation, which returns a p-value for the null hypothesis of no correlation, along with a $\rho$ value between -1 (perfect negative correlation) to 1 (perfect correlation), and 0 for no correlation~\cite{spearman1987proof,rebekic2015pearson}.
+Across all programs, the correlation was very strong ($\rho = 0.86$, $p < 10^{-180}$).
+Per-program correlations were similarly strong and statistically significant: $\rho = 0.71$ for \BPNoise{} ($p < 10^{-25}$), $\rho = 0.64$ for \AddSineSaw{} ($p < 10^{-19}$), $\rho = 0.84$ for \AmpMod{} ($p < 10^{-43}$), and $\rho = 0.85$ for \FMMod{} ($p < 10^{-44}$).
+
+% For each program and loss function, the blind manual hearing test gives us 80 ranks, which are given by two of the authors to a set of 40 randomly selected experiment results.
+This gives 160 pairs of ranks per program, and 640 for all programs.
+We computed Spearman correlations between the two reviewers.
 
@@
 For each program and loss function pair, we have three distributions that can be used to rank the loss functions.
-Two with 300 automatically assigned similarities and one with 80 likert scores.
+Two with 300 automatically assigned similarities (P-Loss and MSS) and one with 80 Likert scores (combining the 40 ranks assigned by each author).
 For consistency and statistical robustness, the distributions are upsampled to 1000 values using bootstrapping~\cite{tibshirani1993introduction}.
@@
 \label{sec:results_posthoc}
-As summarized in Figure~\ref{fig:posthoc_evaluation}, for each program and scoring method, there are four sets of empirical results (a set of 1000 bootstrapped values for each loss function).
+As summarized in Figure~\ref{fig:posthoc_evaluation}, for each program and evaluation (or scoring) method, there are four sets 1000 bootstrapped values, corresponding to the four loss functions.
 A post-hoc test is conducted in two stages in order to determine whether there is a difference in performance between loss functions for each program, and if so, which are the best performers.
 The first stage determines whether there is a difference in group means, with the null hypothesis being that all groups have similar mean ranks.
-If the null hypothesis is rejected, the second stage ranks the loss functions from best to worst using the non-parametric Scott-Knott test(\gls{NPSK})~\cite{tantithamthavorn2017mvt,tantithamthavorn2018optimization}.
-The Kruskal-Wallis test is used for the first stage~\cite{kruskal1952use}.
-This test works by pooling and ranking all performance evaluation measures for a program, then testing whether the differences in mean rank for all loss function groups is zero.
+The second stage ranks the loss functions from best to worst using the non-parametric Scott-Knott test (\gls{NPSK})~\cite{tantithamthavorn2017mvt,tantithamthavorn2018optimization}.
+The Kruskal-Wallis test is used for the first stage~\cite{kruskal1952use}; this test pools and ranks all evaluation measures for a program, then tests whether the differences in mean rank for all loss function groups is zero.
 The Kruskal-Wallis test calculates an H statistic that is compared to a chi-square distribution with k-1 degrees of freedom (k is the number of groups, or 4, and degrees of freedom is 3).
 Using the significance level of 0.05, if the H statistic is greater than the critical value of the chi-square distribution then the null hypothesis is rejected, meaning that at least one group has a mean rank significantly above others~\cite{kruskal1952use}.
-Table~\ref{tab:kruskal_auto} shows the Kruskal-Wallis results for every program and evaluation method, and shows that significant differences between loss function performance measures exist in all cases except MSS for the \AmpMod{} program.
+Table~\ref{tab:kruskal_auto} shows the Kruskal-Wallis results for every program and evaluation method, and shows that significant differences between loss function performance measures (i.e., the bootstrapped distributions of the evaluation for each loss function) exist in all cases except MSS for the \AmpMod{} program.
 
@@
 \centering
-\caption{Kruskal-Wallis results by program and evaluation method. Bootstrapped distributions of performance measures (or ranks for manual listening tests) are compared.}
+\caption{Kruskal-Wallis results by program and evaluation method. Bootstrapped distributions of automatic evaluations (or ranks for manual listening tests) are compared.}
 \begin{tabular}{|c|c|c|c|c|}
@@
 \subsection{Consistency In Rankings}
-
-Automatic sound evaluation methods are unreliable, as we do not know how they would align with the often subjective opinion of human listeners.
-Here, the manual blinded rankings given by the reviewers could be treated as a ground truth, particularly since the two reviewers gave very similar rankings to the experiment results.
-To measure the agreement, Spearman's rank correlation is used.
-The Spearman test returns a p-value for the null hypothesis of no correlation, along with a $\rho$ value between -1 (perfect negative correlation) to 1 (perfect correlation), and middle value of 0 for no correlation~\cite{spearman1987proof,rebekic2015pearson}.
-For each program and loss function, the blind manual hearing test gives us 80 ranks, which are given by two of the authors to a set of 40 randomly selected experiment results.
-This gives 160 pairs of ranks per program, and 640 for all programs.
-We computed Spearman correlations between the two reviewers.
-Across all programs, the correlation was very strong ($\rho = 0.86$, $p < 10^{-180}$).
-Per-program correlations were similarly strong and statistically significant: $\rho = 0.71$ for \BPNoise{} ($p < 10^{-25}$), $\rho = 0.64$ for \AddSineSaw{} ($p < 10^{-19}$), $\rho = 0.84$ for \AmpMod{} ($p < 10^{-43}$), and $\rho = 0.85$ for \FMMod{} ($p < 10^{-44}$).
-
-Table~\ref{tab:combined_ranks} shows the rankings for each program, with ranks different from the corresponding hearing test results marked with an asterisk.
+\label{sec:consistency_in_rankings}
+Table~\ref{tab:combined_ranks} shows the rankings given by each evaluation method for each program, with ranks different from the corresponding hearing test results marked with an asterisk.
 Both MSS and P-Loss gave a different rank from the hearing results in 3 out of 16 cases, which---somewhat surprisingly---shows consistency between automatic and manual hearing tests, at least for the simple programs used in this work.

### 2025-05-13 main.tex
Updates from Overleaf

@@
 This approach goes beyond mere replication, using synthesizers for artistic reinterpretation of sounds.
-However, the automation of this interpretive approach to sound design using digital audio synthesizers has been rarely explored, and significant knowledge gaps remain even in the simpler problem of replicating sounds.
+However, the automation of this interpretive approach to sound design using digital audio synthesizers has rarely been explored, and significant knowledge gaps remain even in the simpler problem of replicating sounds.
 One such gap is the lack of diversity and depth in the analysis of sound-similarity measures and their interaction with various synthesis methods.
-As a stepping stone toward automated interpretive sound design, we investigate whether the optimal choice of sound-similarity measure depends on the synthesis method.
+As a stepping stone toward automated interpretive sound design, we use automatic and manual sound-evaluations to investigate whether the optimal choice of a sound-similarity measure is dependent on the synthesis method.
 
@@
 
-We note that for two synthesizers, our novel use of ``dynamic time warping of envelopes'' as a loss function for differentiable sound-matching outperforms the standard methods of audio comparison.
+We note that for some synthesizers, our novel use of ``dynamic time warping of envelopes'' as a loss function for differentiable sound-matching outperforms the standard methods of audio comparison.
 These results imply that rather than aiming to outperform the SOTA and finding a universal best performer, more effort could be placed towards diversifying the existing similarity measures of sound as well as conducting experiments using a variety of synthesis methods.
@@
 It could be argued that the implementation of these components is not an optimization problem with a correct answer, but a \textit{creative} endeavor depending on the artist's needs.
-Consider the case where the goal is to create an 8-bit version of a high quality, organic snare drum~\cite{collins2007loop}: in such a case, there is no optimal or ``correct" version of an 8-bit snare sound, and multiple correct answers exist.
+Consider the hypothetical case where the goal might be to create an 8-bit~\cite{collins2007loop} version of a high quality, organic snare drum: in such a case, there is no optimal or ``correct" version of an 8-bit snare sound.
 
@@
 
-Automatic sound evaluation methods are unreliable, as we do not know how they would align with the (often subjective) opinion of human listeners, for this reason, listening tests are conducted by taking a random sample of 40 experiment results for each program and loss function pair; followed by the blinded assignment of scores of 1-5 on a Likert scale (1 for no similarity, 5 for identical)~\cite{jebb2021review} by two of the authors.
-The agreement between the assigned scores is measured using Spearman's rank correlation, which returns a p-value for the null hypothesis of no correlation, along with a $\rho$ value between -1 (perfect negative correlation) to 1 (perfect correlation), and 0 for no correlation~\cite{spearman1987proof,rebekic2015pearson}.
+Automatic sound evaluation methods are unreliable, as we do not know how they would align with the (often subjective) opinion of human listeners, for this reason, listening tests are conducted by taking a random sample of 40 experiment results for each program and loss function pair; followed by two of the authors assigning blinded scores to the outputs and targets using a Likert scale (1 for no similarity, 5 for identical)~\cite{jebb2021review}.
+The agreement between the assigned scores by the authors is measured using Spearman's rank correlation, which returns a p-value for the null hypothesis of no correlation, along with a $\rho$ value between -1 (perfect negative correlation) to 1 (perfect correlation), and 0 for no correlation~\cite{spearman1987proof,rebekic2015pearson}.
 Across all programs, the correlation was very strong ($\rho = 0.86$, $p < 10^{-180}$).
 Per-program correlations were similarly strong and statistically significant: $\rho = 0.71$ for \BPNoise{} ($p < 10^{-25}$), $\rho = 0.64$ for \AddSineSaw{} ($p < 10^{-19}$), $\rho = 0.84$ for \AmpMod{} ($p < 10^{-43}$), and $\rho = 0.85$ for \FMMod{} ($p < 10^{-44}$).
+This high rate of agreement is a good indicator that the scores assigned by manual listeners can be taken as the ground truth.
 
@@
  For updating the synthesizer parameters (or weights), we use RMSProp, which operates similar to stochastic gradient descent (SGD)~\cite{goodfellow2016deep}, with the caveat that the gradients of each weight are scaled by the root-mean-square of past gradients of that weight.
-We used a fixed learning rate of 0.045 for all experiments.
+Based on some initial test runs, we used an arbitrary fixed learning rate of 0.045 for all experiments (learning rates used in Vahidi \textit{et al.}~\cite{vahidi2023mesostructures} is unknown).
 The maximum number of iterations is set to 200, where, based on our observations, the parameters have either irrecoverably diverged outside the acceptable ranges, or are stuck at a local minimum.

### 2025-05-14 main.tex
is it done, yuri?

@@
 Since the 1970s, artists such as Suzanne Ciani~\cite{ciani_life_in_waves} have explored the use of audio synthesizers to replicate natural sounds—such as a bottle cap popping off—in ways that preserve the essential qualities of the original while showcasing the unique characteristics of the synthesizer.
-This approach goes beyond mere replication, using synthesizers for artistic reinterpretation of sounds.
+Rather than merely replicating sounds, this approach uses synthesizers for artistic reinterpretation.
 However, the automation of this interpretive approach to sound design using digital audio synthesizers has rarely been explored, and significant knowledge gaps remain even in the simpler problem of replicating sounds.
@@
 Testing the hypothesis that the performance of a similarity measure (or loss function) is influenced by other factors in the environment, particularly the method of synthesis.
-We combine 4 loss functions with 4 differentiable synthesizers, and compare their performance as measured by the similarity of the final output or parameters to the target or target parameters.
+We combine 4 loss functions with 4 differentiable synthesizers, and compare their performance as measured by the similarity of the final output to the target.
 We evaluate the final similarity with two different automatic methods, as well as manual listening tests conducted by two of the authors.
@@
 In replication, the goal is to make an identical copy of the target sound.
-Imitation is an artistic pursuit, and harder to define, where the goal is to make new sounds that only retain a subset of the target's sonic features.
+Imitation is an artistic pursuit and harder to define, since the goal is to make new sounds that only retain a subset of the target's sonic features.
 While closely related to the in-domain versus out-of-domain problem, the choice of replication versus imitation is more dependent on how the loss and representation functions are defined.
@@
 Towards addressing the problems of \LossSelect{} and \SynthSelect{}, we compare the performance of iterative sound-matching experiments, where four differentiable loss functions are applied to four differentiable synthesizers.
-What we mean by performance is the similarity of the final output to the target, as measured by P-Loss, MSS, and manual listening tests, which we treat as the ground truth.
+What we mean by ``performance'' is the similarity of the final output to the target, as measured by P-Loss, MSS, and--most importantly--manual listening tests.
 
@@
 The two automatic evaluation methods used here are P-Loss and MSS, as described in Section~\ref{sec:loss_funcs}.
-For P-Loss, the parameters are normalized between 0-1 based on the valid ranges defined in the Faust program, and the L1 distance is calculated.
-MSS is computed using a fixed FFT size of $512$ and hop size of $100$, with multiple window lengths $\{512, 1024, 2048, 4096\}$.
+For P-Loss, the parameters are normalized between 0-1 based on the valid ranges defined in the Faust program, and the L1 distance between the normalized parameters is calculated.
+MSS is computed using a hop size of $100$ samples, and FFT window lengths of $(512, 1024, 2048, 4096)$.
 
-Automatic sound evaluation methods are unreliable, as we do not know how they would align with the (often subjective) opinion of human listeners, for this reason, listening tests are conducted by taking a random sample of 40 experiment results for each program and loss function pair; followed by two of the authors assigning blinded scores to the outputs and targets using a Likert scale (1 for no similarity, 5 for identical)~\cite{jebb2021review}.
-The agreement between the assigned scores by the authors is measured using Spearman's rank correlation, which returns a p-value for the null hypothesis of no correlation, along with a $\rho$ value between -1 (perfect negative correlation) to 1 (perfect correlation), and 0 for no correlation~\cite{spearman1987proof,rebekic2015pearson}.
+Automatic sound evaluation methods are often unreliable, as their alignment with the subjective judgments of human listeners is uncertain.
+Therefore, listening tests are conducted by randomly sampling 40 experiment results for each program and loss function pair.
+Two of the authors then assign blinded similarity scores to the outputs and targets using a 5-point Likert scale (1 = no similarity, 5 = near identical)~\cite{jebb2021review}.
+Agreement between the assigned scores is measured using Spearman’s rank correlation, which provides both a correlation coefficient ($\rho$) ranging from –1 (perfect negative correlation) to 1 (perfect positive correlation), and a p-value testing the null hypothesis of no correlation~\cite{spearman1987proof,rebekic2015pearson}.
 Across all programs, the correlation was very strong ($\rho = 0.86$, $p < 10^{-180}$).
-Per-program correlations were similarly strong and statistically significant: $\rho = 0.71$ for \BPNoise{} ($p < 10^{-25}$), $\rho = 0.64$ for \AddSineSaw{} ($p < 10^{-19}$), $\rho = 0.84$ for \AmpMod{} ($p < 10^{-43}$), and $\rho = 0.85$ for \FMMod{} ($p < 10^{-44}$).
-This high rate of agreement is a good indicator that the scores assigned by manual listeners can be taken as the ground truth.
+Per-program correlations were also very strong: $\rho = 0.71$ for \BPNoise{} ($p < 10^{-25}$), $\rho = 0.64$ for \AddSineSaw{} ($p < 10^{-19}$), $\rho = 0.84$ for \AmpMod{} ($p < 10^{-43}$), and $\rho = 0.85$ for \FMMod{} ($p < 10^{-44}$).
+This high rate of agreement is a good indicator that the scores assigned by manual listeners can be treated as the ground truth.
 
@@
 NPSK may determine that multiple functions belong to the same rank, in which case the maximum rank would be less than 4.
-Lower P-Loss and MSS values indicate higher similarity, therefore (for consistency with manually assigned ranks) the inverse of their output is used, where higher values indicate greater similarity.
+As mentioned before, lower P-Loss and MSS values indicate higher similarity, therefore (for consistency with manually assigned ranks) the reciprocal of their output is used ($x^{-1}$, if $x$ is the P-Loss or MSS value), where higher values indicate greater similarity.
 
@@
 
- This work, much like previous work, only utilizes in-domain sounds.
+ Like the majority of previous works, this work utilizes in-domain sounds.
 That is, the target sound is made by the synthesizer, and the parameters are already known.
 This simplifies the issues of measuring sound-similarity, but it is not a realistic scenario for practical sound-matching.
-This problem is left for future work, as we will discuss in the next section.
+This problem is left for future work, discussed in the next section.
 
@@
 Viewing the sound-matching problem as ``the navigation of an agent from an arbitrary point in a gradient field to a target'' closely resembles many classical problems in the field of reinforcement learning (RL)~\cite{sutton2018reinforcement}.
-Naturally, we can extend our work by applying RL and other heuristic search techniques to the problem of sound-matching, and measuring the performance changes (both quantitative and qualitative) relative to simple gradient descent updates.
+Naturally, the application of RL and other heuristic search techniques to the problem of iterative sound-matching would be an important contribution.
 

### 2025-05-14 main.tex
Updates from Overleaf

@@
     \item $t$: The target sound which we want to replicate or imitate. 
-    \item $\phi$: Representation function or feature extractor. $\phi$ is applied to $x$ and $t$ to facilitate their comparison using the loss function.
+    \item $\phi$: Representation function or feature extractor. $\phi$ is applied to $x$ and $t$ to facilitate their comparison by the loss function.
     \item $L$: Loss or error function. $L(\theta,t)$ is a measure of distance between $x$ and $t$. Often proportional to the subtraction of their representations, or $ \phi(x) - \phi(t)$
@@
 % \begin{enumerate}
-%  \item \textbf{Replicate out-of-domain sounds}: this would give us the  the parameters that could closely approximate a target sound, and we can modify the sound further by modulating the synthesizer.
+%  \item \textbf{Replicate out-of-domain sounds}: this would give us the parameters that could closely approximate a target sound, and we can modify the sound further by modulating the synthesizer. 
 % \item \textbf{Imitate in-domain or out-of-domain sounds}: This would allow for applying the characteristics of one sound to another. 
@@
 As discussed in Section~\ref{sec:matching_types}, the L1 or L2 distances are the most common methods of comparing different spectrograms, however, there are situations where this might lead to shortcomings.
-An example given by Vahidi et  al.~\cite{vahidi2023mesostructures} is two chirplets (tones that increase in pitch) starting and ending at the same frequencies.
+An example given by Vahidi \textit{et al.}~\cite{vahidi2023mesostructures} is two chirplets (tones that increase in pitch) starting and ending at the same frequencies.
 If one of these tones is slightly shifted in time, then the spectrograms will no longer overlap at any point, despite their sonic similarity.

### 2025-05-15 main.tex
Updates from Overleaf

@@
 Manual sound design with a synthesizer is inherently iterative: an artist compares the synthesized output to a mental target, adjusts parameters, and repeats until satisfied.
-Iterative sound-matching automates this workflow by continually programming a synthesizer under the guidance of a loss function that quantifies similarity to the target sound.
+Iterative sound-matching automates this workflow by continually programming a synthesizer under the guidance of a loss function (or similarity measure) towards a target sound.
 Prior comparisons of loss functions have typically favored one metric over another, but only within narrow settings: limited synthesis methods, few loss types, often without blind listening tests.
-This leaves open whether a universally optimal loss exists, or the choice of loss remains a creative decision conditioned on the synthesis method and the sound designer's preference.
+This leaves open the question of whether a universally optimal loss exists, or the choice of loss remains a creative decision conditioned on the synthesis method and the sound designer's preference.
 We propose differentiable iterative sound-matching as the natural extension of the available literature, since it combines the manual approach to sound design with modern advances in machine learning.
-To analyze the variability of loss function performance across synthesizers, we implemented differentiable subtractive, additive, and AM synthesizers, paired with four novel or established differentiable loss functions.
+To analyze the variability of loss function performance across synthesizers, we implemented a mix of four novel and established differentiable loss functions, and paired them with differentiable subtractive, additive, and AM synthesizers.
 For each of the sixteen synthesizer–loss combinations, we ran 300 randomized sound-matching trials.
-Performance is measured with parameter differences, spectrogram-distance metrics, and blind listening scores; a fair amount of consistency between the three performance measures is observed.
-A post-hoc analysis shows that each of these loss functions has the potential to be the most appropriate.
+Performance was measured using parameter differences, spectrogram-distance metrics, and manually assigned listening scores.
+We observed a moderate level of consistency among the three performance measures.
+Our post-hoc analysis shows that the loss function performance is highly dependent on the synthesizer.
 These findings underscore the value of expanding the scope of sound-matching experiments, and developing new similarity metrics tailored to specific synthesis techniques, rather than pursuing one-size-fits-all solutions.
@@
 \begin{IEEEkeywords}
-Machine Learning, Signal Processing, Sound-Matching
+Audio Synthesis, Differentiable Digital Signal Processing, Machine Learning, Sound-Matching
 \end{IEEEkeywords}
@@
 
-``Iterative sound-matching'' refers to works that mimic the back-and-forth process of listening to a synthesizer's output and updating the parameters until the termination conditions are met.
+\textit{Iterative sound-matching} refers to works that mimic the back-and-forth process of listening to a synthesizer's output and updating the parameters until the termination conditions are met.
 The two necessary components are the \textit{loss function}, which determines how close two sounds are, and the \textit{parametric synthesizer} that outputs sounds based on input parameters.
@@
 A Riemannian metric M needs to be calculated for the minimization of locally linear approximation of the ``true" spectral loss function~\cite{han2023perceptual}.
-
 \[
@@
 Towards addressing the problems of \LossSelect{} and \SynthSelect{}, we compare the performance of iterative sound-matching experiments, where four differentiable loss functions are applied to four differentiable synthesizers.
-What we mean by ``performance'' is the similarity of the final output to the target, as measured by P-Loss, MSS, and--most importantly--manual listening tests.
+What we mean by ``performance'' is the similarity of the final output to the target, as measured by P-Loss, MSS, and---most importantly---manual listening tests.
 
 \textbf{We hypothesize that the relative performance of loss functions for sound-matching is influenced by the method of synthesis}.
-The results of these experiments show that there is no universally optimal loss function, strengthening the idea that the selection of the loss function and synthesis method is not an optimization problem, but a subjective choice depending on the needs of the sound designer.
+If true, this would strengthen the idea that the selection of the loss function and synthesis method is not an optimization problem, but a subjective choice depending on the needs of the sound designer.
 
@@
 
-\subsubsection{Evaluation Methods}
+\subsubsection{Automatic and Manual Evaluation}
+\label{sec:evaluation_manual_auto}
 The two automatic evaluation methods used here are P-Loss and MSS, as described in Section~\ref{sec:loss_funcs}.
@@
 Two of the authors then assign blinded similarity scores to the outputs and targets using a 5-point Likert scale (1 = no similarity, 5 = near identical)~\cite{jebb2021review}.
-Agreement between the assigned scores is measured using Spearman’s rank correlation, which provides both a correlation coefficient ($\rho$) ranging from –1 (perfect negative correlation) to 1 (perfect positive correlation), and a p-value testing the null hypothesis of no correlation~\cite{spearman1987proof,rebekic2015pearson}.
+Using Spearmna's rank correlation~\cite{spearman1987proof,rebekic2015pearson}, we found a very strong rate of agreement between the human listeners (see Section~\ref{sec:consistency_in_rankings}), a good indicator that the scores assigned by manual listeners can be treated as the ground truth.
+We then combine the ranks assigned by both authors, giving us 80 ranks per program and loss function pair.
+
+% Agreement between the assigned scores is measured using Spearman’s rank correlation, which provides both a correlation coefficient ($\rho$) ranging from –1 (perfect negative correlation) to 1 (perfect positive correlation), and a p-value testing the null hypothesis of no correlation~\cite{spearman1987proof,rebekic2015pearson}.
 Across all programs, the correlation was very strong ($\rho = 0.86$, $p < 10^{-180}$).
@@
 This backward calculation resembles the \textit{exploding gradients} problem in recurrent neural networks~\cite{gers2000learning}.
-For this reason, gradient clipping~\cite{goodfellow2016deep} is preemptively used to ensure that the  $\ell_2$ norm of all gradientsdoes not exceed a clipping threshold 1.
+Gradient clipping~\cite{goodfellow2016deep} is preemptively used to ensure that the  $\ell_2$ norm of all gradientsdoes not exceed the threshold of 1.
 % \mathbf{\gamma}_i^{\text{clipped}} = \mathbf{\gamma}_i \cdot \mathcal{S}
@@
 
-\subsubsection{\BPNoise}
-\label{sec:program0}
-
 \begin{lstlisting}[caption={\BPNoise}, label={lst:program0}, language=Faust,
@@
 
+\subsubsection{\BPNoise}
+\label{sec:program0}
 \BPNoise{} is a bare bones example of subtractive synthesis using digital filters~\cite{smith2007introduction}.
@@
 Listing~\ref{lst:program0} shows the Faust code for \BPNoise.
+
 \subsubsection{\AddSineSaw}
 \label{sec:program1}
-
 \AddSineSaw{} is an additive program that combines a saw and a sine wave function.
@@
 \label{sec:program2}
-
 \AmpMod{} involves amplitude modulation of a noise generator by an LFO with \textit{modulator} as its frequency and a global \textit{amp} value that does not change over time and applies to the entire signal.
@@
 \subsubsection{\FMMod}
-
 \label{sec:program3}
-
 Relative to additive and subtractive synthesis, AM/FM synthesis are less common methods of sound design.
@@
 \centering
-\caption{Kruskal-Wallis results by program and evaluation method. Bootstrapped distributions of automatic evaluations (or ranks for manual listening tests) are compared.}
+\caption{Do loss functions have different median performance? Kruskal-Wallis results by program and bootstrapped evaluation results.}
 \begin{tabular}{|c|c|c|c|c|}
@@
 
+\input{extra_figures/ranking_tables}
+\input{extra_figures/npsk_figures}
+
 \subsubsection{\BPNoise}
@@
 As shown in Figure~\ref{fig:npsk_p0}, manual hearing test and MSS selected \SIMSESpec~as the best performer and \LoneSpec~as the second-best performer, while P-loss gave the reverse order.
-
-\begin{table*}[htbp]
-\centering
-\caption{Ranks for each synthesis method (rows) under the three evaluation metrics (columns), across four targets (\BPNoise, \AddSineSaw, \AmpMod, \FMMod).
-Asterisks denote difference from Likert score ranks.}
-\footnotesize
-\begin{tabular}{|c|ccc|ccc|ccc|ccc|}
-\hline
-\textbf{Function} 
-  & \multicolumn{3}{c|}{\BPNoise}
-  & \multicolumn{3}{c|}{\AddSineSaw}
-  & \multicolumn{3}{c|}{\AmpMod}
-  & \multicolumn{3}{c|}{\FMMod} \\
-\cline{2-13}
-  & MSS & P-LOSS & Hearing 
-  & MSS & P-LOSS & Hearing 
-  & MSS & P-LOSS & Hearing 
-  & MSS & P-LOSS & Hearing \\
-\hline
-\textbf{SIMSE} 
-  & 1    & 2*   & 1    
-  & 4    & 4    & 4    
-  & 3*   & 4    & 4    
-  & 2    & 2    & 2    \\
-\textbf{L1}    
-  & 2    & 1*   & 2    
-  & 2    & 2    & 2    
-  & 2    & 2    & 2    
-  & 3    & 3    & 3    \\
-\textbf{JTFS}  
-  & 3    & 4*   & 3    
-  & 1   & 1    & 1    
-  & 4*   & 3    & 3    
-  & 4    & 4    & 4    \\
-\textbf{DTW}   
-  & 4*   & 3    & 3    
-  & 3    & 3    & 3    
-  & 1    & 1    & 1    
-  & 1    & 1    & 1    \\
-\hline
-\end{tabular}
-% \vspace{0.25em}
-
-\label{tab:combined_ranks}
-\end{table*}
-
-\input{extra_figures/npsk_figures}
 
@@
 \label{sec:consistency_in_rankings}
-Table~\ref{tab:combined_ranks} shows the rankings given by each evaluation method for each program, with ranks different from the corresponding hearing test results marked with an asterisk.
-Both MSS and P-Loss gave a different rank from the hearing results in 3 out of 16 cases, which---somewhat surprisingly---shows consistency between automatic and manual hearing tests, at least for the simple programs used in this work.
+\subsubsection{Manual Ranks} Agreement between the manually assigned scores is measured using Spearman’s rank correlation, which provides both a correlation coefficient ($\rho$) ranging from –1 (perfect negative correlation) to 1 (perfect positive correlation), and a p-value testing the null hypothesis of no correlation~\cite{spearman1987proof,rebekic2015pearson}.
+Across all programs, the correlation was very strong ($\rho = 0.86$, $p < 10^{-180}$).
+Per-program correlations were also very strong: $\rho = 0.71$ for \BPNoise{} ($p < 10^{-25}$), $\rho = 0.64$ for \AddSineSaw{} ($p < 10^{-19}$), $\rho = 0.84$ for \AmpMod{} ($p < 10^{-43}$), and $\rho = 0.85$ for \FMMod{} ($p < 10^{-44}$).
+
+Table~\ref{tab:combined_ranks} shows the SNPK rankings of bootstrapped evaluation results for each program.
+Both MSS and P-Loss gave a different rank from the hearing results in 3 out of 16 cases, which shows consistency between automatic and manual hearing tests, at least for the simple programs used in this work.
 We also observe that identical top ranks were consistent across performance evaluation methodologies, with the exception of P-Loss in \BPNoise, which narrowly picks a different spectrogram-based loss function.
@@
 
-We also see that the \DTWEnv{} and \SIMSESpec{} loss functions (to our knowledge, not previously used in iterative sound-matching), lead to the best results in three of the four programs.
+We also see that the \DTWEnv{} and \SIMSESpec{} (to our knowledge, not previously used in iterative sound-matching), lead to the best results in three of the four programs.
 This shows room for further creative approaches to differentiable sound-similarity measures, particularly since we observed that SOTA loss functions can be ineffective depending on the design of a synthesizer.
@@
 In other words, different synthesizer programs work best with different loss functions.
-Notably, we see that our novel use of DTW and SIMSE based differentiable loss functions (\DTWEnv{} and \SIMSESpec) can outperform what are regarded as the SOTA loss functions, although this is highly synthesizer dependent.
+Notably, we see that our novel use of DTW and SIMSE based differentiable loss functions (\DTWEnv{} and \SIMSESpec) can outperform what are regarded as the SOTA loss functions in 3 of 4 cases, although this is highly synthesizer dependent.
 
 P-Loss and MSS have frequently been used as automatic performance measures, yet their ``preference'' has rarely been compared to human rankings.
-We observed that automatic performance measures and manual listening tests were often (though not always) in agreement, despite this, we believe that manual verification of sound-matching results remains a necessity for future works.
+We observed that automatic performance measures and manual listening tests were often in agreement, despite this, manual verification of sound-matching results remains a necessity for future works.
 
-\textbf{Weaknesses:} Compared to previous work, we presented a more cohesive approach to iterative sound-matching which utilizes and compares four loss functions and four synthesis methods.
+\textbf{Weaknesses:} Compared to previous work, we presented a more cohesive approach to iterative sound-matching which utilizes a variety of loss functions and synthesis methods.
 However, there are many other methods of synthesis and sound-similarity that can be combined in practically infinite ways.
@@
 \label{sec:future}
-\textbf{Future Work: }The analysis of different loss functions in sound-matching shows that problem of periodic loss-landscapes is not limited to spectral representations using Fourier transforms~\cite{turian2020sorry,vahidi2023mesostructures,uzrad2024diffmoog,bruford2024synthesizer}.
-Perhaps this problem emerges due to the periodic nature of sound.
-With regard to navigation of loss function landscapes, an optimizer that is more aware of fluctuations in gradients would perhaps lead to better solutions than simple gradient descent.
+\textbf{Future Work: } The problem of periodic loss-landscapes, noted by many previous works, remains unaddressed~\cite{turian2020sorry,vahidi2023mesostructures,uzrad2024diffmoog,bruford2024synthesizer}.
+Perhaps this problem emerges due to the periodic nature of sound, which requires better loss landscape navigation methods.
+An optimizer that is more aware of fluctuations in gradients would perhaps lead to better solutions than simple gradient descent.
 Viewing the sound-matching problem as ``the navigation of an agent from an arbitrary point in a gradient field to a target'' closely resembles many classical problems in the field of reinforcement learning (RL)~\cite{sutton2018reinforcement}.
@@
 
-The current trend in contemporary work appears to be the application of domain specific and computationally expensive loss functions~\cite{han2023perceptual,uzrad2024diffmoog}, use of large neural networks~\cite{hershey2017cnn,cramer2019look}, or complex ensemble methods~\cite{turian2022hear}.
-Such models are useful, but often intractable; furthermore, training such models still requires the definition of simpler loss functions, which emphasizes the need for further work on differentiable and expressive loss function implementations.
+Contemporary works often involve the application of domain specific and computationally expensive loss functions~\cite{han2023perceptual,uzrad2024diffmoog}, use of large neural networks~\cite{hershey2017cnn,cramer2019look}, or complex ensemble methods~\cite{turian2022hear}.
+Such models are useful but intractable; furthermore, training them requires the definition of simpler loss functions, which emphasizes the need for further development of differentiable and expressive loss function implementations.
 

### 2025-05-15 main.tex
Updates from Overleaf

@@
 Sound design is an inherently creative endeavor, and would likely benefit from a diversified range of solutions for different approaches.
-
-% \subsection{Differentiable Sound-Matching}
 
@@
 This backward calculation resembles the \textit{exploding gradients} problem in recurrent neural networks~\cite{gers2000learning}.
-Gradient clipping~\cite{goodfellow2016deep} is preemptively used to ensure that the  $\ell_2$ norm of all gradientsdoes not exceed the threshold of 1.
+Gradient clipping~\cite{goodfellow2016deep} is preemptively used to ensure that the  $\ell_2$ norm of all gradients does not exceed the threshold of 1.
 % \mathbf{\gamma}_i^{\text{clipped}} = \mathbf{\gamma}_i \cdot \mathcal{S}
@@
 In cases where the distributions are similar, multiple loss functions can be clustered into the same rank.
-
-We use color-coded violin plots to visualize the best performers for each program. In the violin plots, the corresponding colors for each rank are
+In Figure~\ref{fig:npsk_all}, we use color-coded violin plots to visualize the best performers for each program.
+The corresponding colors for each rank are
 \colorbox{rank1}{\textcolor{black}{\textbf{1}}} \colorbox{rank2}{\textcolor{white}{\textbf{2}}} \colorbox{rank3}{\textcolor{white}{\textbf{3}}} \colorbox{rank4}{\textcolor{black}{\textcolor{white}{\textbf{4}}}}
@@
 Both MSS and P-Loss gave a different rank from the hearing results in 3 out of 16 cases, which shows consistency between automatic and manual hearing tests, at least for the simple programs used in this work.
-We also observe that identical top ranks were consistent across performance evaluation methodologies, with the exception of P-Loss in \BPNoise, which narrowly picks a different spectrogram-based loss function.
+We also observe that top ranks were consistent across performance evaluation methodologies, with the exception of P-Loss in \BPNoise, which narrowly picks a different spectrogram-based loss function.
 % \input{extra_figures/ranking_tables}

### 2025-05-16 extra_figures/evaluation_graph.tex
Updates from Overleaf

@@
     function/.style={rectangle, draw, rounded corners, minimum width=2cm, minimum height=1cm, text centered, node distance=1cm},
-    distribution/.style={ellipse, draw, minimum width=1cm, minimum height=0cm, text centered, node distance=0.3cm},
+    distribution/.style={ellipse, draw, minimum width=1cm, minimum height=0cm, text centered, node distance=0.1cm},
     gaussian/.style={draw, smooth, samples=100, domain=-2.5:2.5},
@@
         child { node[distribution, mss] {MSS}
-            child[grow=down, level distance=1.5cm] {
+            child[grow=down, level distance=1cm] {
                 node[draw=none] {
@@
         child { node[distribution, ploss] {P-Loss}
-            child[grow=down, level distance=1.5cm] {
+            child[grow=down, level distance=1cm] {
                 node[draw=none] {
@@
         child { node[distribution, likert] {Likert}
-            child[grow=down, level distance=1.5cm] {
+            child[grow=down, level distance=1cm] {
                 node[draw=none] {
@@
 % Label for the second row (Distributions)
-\node at (0,-4) {\textbf{$\longleftarrow$\ \ \large Bootstrapped Distributions\ \ $\longrightarrow$}};
+\node at (0,-3.25) {\textbf{$\longleftarrow$\ \ \large Bootstrapped Distributions\ \ $\longrightarrow$}};
 % Left side: NPSK with MSS
 
-% Left side: NPSK with MSS
-\node[NPSK, text height=5ex, align=center, fill=red!30, draw=black] at (-6,-6) {NPSK with 4 \\ MSS distributions};
-\draw[->] (-6,-6.5) -- (-6,-7.5) node[below, below] {\textbf{Ranks 1-4}};
-\draw[red!40, thick] (-5, -5.4) -- (-2.95, -4.95);  % Line 1
-\draw[red!40, thick] (-5.5, -5.4) -- (-3.5, -4.75);  % Line 2
-\draw[red!40, thick] (-6, -5.4) -- (-5.5, -4.75);  % Line 3
-\draw[red!40, thick] (-6.5, -5.4) -- (-9, -4.75);    % Line 4
+\def\xoffsetL{-0.5}
+\def\xoffsetM{0}
+\def\xoffsetR{0.5}
+\def\yoffset{1}
+
+% Left: NPSK with MSS
+\node[NPSK, text height=5ex, align=center, fill=red!30, draw=black] at ({-6+\xoffsetL},{-6+\yoffset}) {NPSK with 4 \\ MSS distributions};
+\draw[->] ({-6+\xoffsetL},{-6.5+\yoffset}) -- ({-6+\xoffsetL},{-6.6+\yoffset}) node[below, below] {\textbf{Ranks 1-4}};
+\draw[red!40, thick] ({-5+\xoffsetL}, {-5.4+\yoffset}) -- ({-2.95+\xoffsetL}, {-5.15+\yoffset});  % Line 1
+\draw[red!40, thick] ({-5.5+\xoffsetL}, {-5.4+\yoffset}) -- ({-3.5+\xoffsetL}, {-4.75+\yoffset});  % Line 2
+\draw[red!40, thick] ({-6+\xoffsetL}, {-5.4+\yoffset}) -- ({-5.5+\xoffsetL}, {-4.75+\yoffset});  % Line 3
+\draw[red!40, thick] ({-6.5+\xoffsetL}, {-5.4+\yoffset}) -- ({-9+\xoffsetL}, {-4.75+\yoffset});  % Line 4
 
 % Middle: NPSK with P-Loss
-\node[NPSK, text height=5ex, align=center, fill=blue!40, draw=black] at (0,-6) {NPSK with 4 \\ P-Loss distributions};
-\draw[->] (0,-6.5) -- (0,-7.5) node[below, below] {\textbf{Ranks 1-4}};
-\draw[blue!40, thick] (1.5, -5.4) -- (3.25, -4.75);   % Line 1
-\draw[blue!40, thick] (0.75, -5.4) -- (1.5, -4.75);   % Line 2
-\draw[blue!40, thick] (-0.25, -5.4) -- (-1, -4.75);   % Line 3
-\draw[blue!40, thick] (-0.75, -5.4) -- (-3, -4.75);   % Line 4
+\node[NPSK, text height=5ex, align=center, fill=blue!40, draw=black] at ({0+\xoffsetM},{-6+\yoffset}) {NPSK with 4 \\ P-Loss distributions};
+\draw[->] ({0+\xoffsetM},{-6.5+\yoffset}) -- ({0+\xoffsetM},{-6.6+\yoffset}) node[below, below] {\textbf{Ranks 1-4}};
+\draw[blue!40, thick] ({1.5+\xoffsetM}, {-5.4+\yoffset}) -- ({3.25+\xoffsetM}, {-4.75+\yoffset});  % Line 1
+\draw[blue!40, thick] ({0.75+\xoffsetM}, {-5.4+\yoffset}) -- ({1.5+\xoffsetM}, {-4.75+\yoffset});  % Line 2
+\draw[blue!40, thick] ({-0.25+\xoffsetM}, {-5.4+\yoffset}) -- ({-1+\xoffsetM}, {-4.75+\yoffset});  % Line 3
+\draw[blue!40, thick] ({-0.75+\xoffsetM}, {-5.4+\yoffset}) -- ({-3+\xoffsetM}, {-4.75+\yoffset});  % Line 4
 
-% Right side: NPSK with Likert
-\node[NPSK, text height=5ex, align=center, fill=green!30, draw=black] at (6,-6) {NPSK with 4 \\ Likert distributions};
-\draw[->] (6,-6.5) -- (6,-7.5) node[below, below] {\textbf{Ranks 1-4}};
-\draw[green!50, thick] (7.5, -5.4) -- (8.75, -4.75);  % Line 1
-\draw[green!50, thick] (6.75, -5.4) -- (6, -4.75);    % Line 2
-\draw[green!50, thick] (5.75, -5.4) -- (4, -4.75);    % Line 3
-\draw[green!50, thick] (4.95, -5.4) -- (3.25, -4.95);    % Line 4
+% Right: NPSK with Likert
+\node[NPSK, text height=5ex, align=center, fill=green!30, draw=black] at ({6+\xoffsetR},{-6+\yoffset}) {NPSK with 4 \\ Likert distributions};
+\draw[->] ({6+\xoffsetR},{-6.5+\yoffset}) -- ({6+\xoffsetR},{-6.6+\yoffset}) node[below, below] {\textbf{Ranks 1-4}};
+\draw[green!50, thick] ({7.5+\xoffsetR}, {-5.4+\yoffset}) -- ({8.75+\xoffsetR}, {-5+\yoffset});  % Line 1
+\draw[green!50, thick] ({6.75+\xoffsetR}, {-5.4+\yoffset}) -- ({6+\xoffsetR}, {-4.75+\yoffset});    % Line 2
+\draw[green!50, thick] ({5.75+\xoffsetR}, {-5.4+\yoffset}) -- ({4+\xoffsetR}, {-4.85+\yoffset});    % Line 3
+\draw[green!50, thick] ({4.95+\xoffsetR}, {-5.4+\yoffset}) -- ({3.25+\xoffsetR}, {-5.1+\yoffset}); % Line 4
 

### 2025-05-16 main.tex
Updates from Overleaf

@@
 
-Automatic sound design has been a subject of research for decades, with reduction of tinkering time and creation of ``interesting" sounds as the main motivators \cite{krekovic2019insights,turian2020sorry,horner1993machine,salimi2020make,esling2019flow,engel2020ddsp,mitchell2007evolutionary,shier2020spiegelib}.
-\textit{Sound-matching} is a common form of automatic sound design explored in literature, where a similarity metric (or loss function) is given, and the goal is to find parameters for a synthesizer that replicate \textit{all} or \textit{some} of the characteristics of a target sound as best as possible~\cite{horner1993machine,mitchell2007evolutionary}.
+Automatic sound design has been a subject of research for decades, with reduction of tinkering time and creation of ``interesting" sounds as the main motivators \cite{krekovic2019insights,turian2020sorry,horner1993machine,salimi2020make,esling2019flow,engel2020ddsp,mitchell2007evolutionary,shier2020spiegelib,masuda2021soundmatch,masuda2023improving}.
+\textit{Sound-matching} is a common form of automatic sound design explored in literature, where a similarity metric (or loss function) is given, and the goal is to find parameters for a synthesizer that replicate \textit{all} or \textit{some} of the characteristics of a target sound as best as possible~\cite{horner1993machine,mitchell2007evolutionary,masuda2023improving}.
 Our analysis of past works shows that this general goal of ``sound-matching'' can be broken down in to very different subtypes, many of them rarely or never explored, despite their relevance to practical applications.
@@
 Despite the early popularity of the iterative approach~\cite{justice1979analytic,horner1993machine,mitchell2007evolutionary}, recent works in sound-matching have favored supervised learning using inference models, deep neural-networks (\gls{DNNs}), and large datasets.
-Rather than continually interacting with the environment, these models learn from a dataset of inputs (synthesizer parameters) and outputs (corresponding sound), and predict the synthesizer parameters for new sounds~\cite{yee2018automatic,han2023perceptual,han2024learning}.
+Rather than continually interacting with the environment, these models learn from a dataset of inputs (synthesizer parameters) and outputs (corresponding sound), and predict the synthesizer parameters for new sounds~\cite{yee2018automatic,han2023perceptual,han2024learning,masuda2021soundmatch,masuda2023improving}.
 
@@
 Here we discuss four methods of audio representation and the corresponding loss functions.
-Three of which have been used in previous works, and a novel method which we will make use of in our own experiments.
+Three of which have been used in previous works, and a novel method not previously used in iterative sound-matching.
 
@@
 Subjective hearing tests were conducted, showing a preference for the fine-tuned model~\cite{masuda2021soundmatch}.
+Masuda \textit{et al.} later extended this work with semi-supervised learning, highlighting significant gaps in in-domain and out-of-domain performance~\cite{masuda2023improving}.
 
@@
 The synthesizer is a customizable chain of effects, which feed one output as input to the next step of the chain; signal-chain loss compares the parameter and output difference at every output step in the chain~\cite{uzrad2024diffmoog}.
-Possible chain functionalities are FM/AM, Low Frequency Oscillators (\gls{LFO}), filters, and envelopes.
+Possible chain functionalities are FM/AM, Low-Frequency Oscillators (\gls{LFO}), filters, and envelopes.
 Like the results shown by Masuda \textit{et al.}~\cite{masuda2021soundmatch}, better out-of-domain results were achieved when pre-trained on in-domain data and fine-tuned using out-of-domain NSynth data~\cite{engel2017neural}.
@@
 We test four different loss functions and compare their success in sound-matching using four different synthesizers.
-We verify the success of the sound-matching process using two automatic measures of sound similarity (MSS and P-Loss) and a manual blinded hearing test conducted by two of the authors where likert scores~\cite{jebb2021review} of 1-5 are assigned to a subset of final outputs and the corresponding targets.
+We verify the success of the sound-matching process using two automatic measures of sound similarity (MSS and P-Loss) and a manual blinded hearing test conducted by two of the authors where Likert scores~\cite{jebb2021review} of 1-5 are assigned to a subset of final outputs and the corresponding targets.
 
@@
 Consider the following simplified scenario: we have two loss functions, $L_1$ and $L_2$, and a single synthesizer program $P$ with a target parameter set $\theta^*$ where the target sound is $t$, or $P(\theta^*) = t$.
-We run a large number (in this case, 300 for each loss function) of experiments where we randomly set the initial ($\theta_0$) and target parameters ($\theta^*$) of $P$ and run an iterative sound-matching loop for each loss function, with max number of iterations set to 200.
+We run a large number (in this case, 300 for each loss function) of experiments where we randomly set the initial ($\theta_0$) and target parameters ($\theta^*$) of $P$ and run an iterative sound-matching loop for each loss function, with maximum number of iterations set to 200.
 When the loop is terminated, we measure the similarity of the final output sound $x$ (where $P(\theta_{199}) = x$) to $t$, or if using P-Loss we compare $\theta_{199}$ to $\theta^*$.
@@
 Two of the authors then assign blinded similarity scores to the outputs and targets using a 5-point Likert scale (1 = no similarity, 5 = near identical)~\cite{jebb2021review}.
-Using Spearmna's rank correlation~\cite{spearman1987proof,rebekic2015pearson}, we found a very strong rate of agreement between the human listeners (see Section~\ref{sec:consistency_in_rankings}), a good indicator that the scores assigned by manual listeners can be treated as the ground truth.
+Using Spearman's rank correlation~\cite{spearman1987proof,rebekic2015pearson}, we found a very strong rate of agreement between the human listeners (see Section~\ref{sec:consistency_in_rankings}), a good indicator that the scores assigned by manual listeners can be treated as the ground truth.
 We then combine the ranks assigned by both authors, giving us 80 ranks per program and loss function pair.
@@
 
-\textbf{Weaknesses:} Compared to previous work, we presented a more cohesive approach to iterative sound-matching which utilizes a variety of loss functions and synthesis methods.
+\textbf{Weaknesses:} 
+While cannot prove that a universally best similarity measure does not exist, we can advocate for more creativity in the field based on our findings.
+Compared to previous work, we presented a more cohesive approach to iterative sound-matching which utilizes a variety of loss functions and synthesis methods.
 However, there are many other methods of synthesis and sound-similarity that can be combined in practically infinite ways.
-
-To limit the scope of our experiments, we had no choice but to use arbitrary parameters for the various signal processing functions.
+Due to this large search space, we set arbitrary parameters for the various signal processing functions.
 We used bare-bones versions of STFT and JTFS with fixed parameters.

### 2025-05-20 main.tex
Updates from Overleaf

@@
 \begin{IEEEkeywords}
-Audio Synthesis, Differentiable Digital Signal Processing, Machine Learning, Sound-Matching
+Audio synthesis, differentiable digital signal processing, music information retrieval, sound-matching
 \end{IEEEkeywords}
@@
 Finally, we surprisingly see somewhat consistent results between the rankings assigned by manual hearing tests (which we take as the ground truth), and automatic measures of P-Loss and MSS.
-Due to the simplicity of our programs and the occasional deviations, we cannot conclude that these automatic measures are substitutes for actual hearing.
+Due to the simplicity of our programs and the occasional deviations, we cannot conclude that these automatic measures are substitutes for human hearing tests.
 

### 2025-05-20 main.tex
Updates from Overleaf

@@
 \textbf{Summary:} Sound-matching is an umbrella term for the algorithmic programming of audio synthesizers, often with the goal of assisting sound designers.
-Here we provided a history of sound-matching, major issues in the field, and discussed the importance of ``differentiable iterative sound-matching'' as a natural extension current literature.
+Here we provided a history of sound-matching, major issues in the field, and discussed the importance of ``differentiable iterative sound-matching'' as a natural extension of current literature.
 
@@
 \textbf{Weaknesses:} 
-While cannot prove that a universally best similarity measure does not exist, we can advocate for more creativity in the field based on our findings.
+While we cannot prove that a universally best similarity measure does not exist, we can advocate for more creativity in the field based on our findings.
 Compared to previous work, we presented a more cohesive approach to iterative sound-matching which utilizes a variety of loss functions and synthesis methods.
@@
 
- Like the majority of previous works, this work utilizes in-domain sounds.
-That is, the target sound is made by the synthesizer, and the parameters are already known.
+ Like the majority of previous works, this work utilizes in-domain sounds; that is, the target sound is made by the synthesizer, and the parameters are already known.
 This simplifies the issues of measuring sound-similarity, but it is not a realistic scenario for practical sound-matching.
@@
 Perhaps this problem emerges due to the periodic nature of sound, which requires better loss landscape navigation methods.
-An optimizer that is more aware of fluctuations in gradients would perhaps lead to better solutions than simple gradient descent.
+An optimizer that is more aware of fluctuations in the gradients would perhaps lead to better solutions than simple gradient descent.
 Viewing the sound-matching problem as ``the navigation of an agent from an arbitrary point in a gradient field to a target'' closely resembles many classical problems in the field of reinforcement learning (RL)~\cite{sutton2018reinforcement}.

### 2025-05-27 main.tex
Updates from Overleaf

@@
 Faust is a functional language for audio synthesis that can succinctly define signal processing chains.
-The programs are first defined in Faust, then converted (or \textit{transpiled}) to differentiable Jax functions~\cite{braun2024dac} using the DawDreamer library\footnote{\url{https://github.com/DBraun/DawDreamer}}.
+Following Braun's methodology~\cite{braun2024dac}, programs are first defined in Faust, then converted (or \textit{transpiled}) to differentiable Jax functions using the DawDreamer library\footnote{\url{https://github.com/DBraun/DawDreamer}}.
 

### 2025-06-19 main.tex
Updates from Overleaf

@@
 
-\begin{lstlisting}[caption={\AmpMod}, label={lst:program2},language=Faust,float,floatplacement=!H,xleftmargin=1em,xrightmargin=0.5em,firstnumber=0,aboveskip=0em, belowskip=-1em]
+\begin{lstlisting}[caption={\AmpMod}, label={lst:program3},language=Faust,float,floatplacement=!H,xleftmargin=1em,xrightmargin=0.5em,firstnumber=0,aboveskip=0em, belowskip=-1em]
 import("stdfaust.lib");

### 2025-08-05 main.tex
Updates from Overleaf

--- 2025-05-27 main.tex
@@
 \newcommand{\PeriodicLoss}{\textbf{Periodic Loss}}
-\newcommand{\OutDomain}{\textbf{Out-Domain Generation}}
+\newcommand{\OutDomain}{\textbf{Out-of-Domain Generation}}
 \newcommand{\BPNoise}{\textbf{BP-Noise}}  
@@
 
-\markboth{IEEE TRANSACTIONS ON AUDIO, SPEECH AND LANGUAGE PROCESSING, VOL -, 2025}%
-{Shell \MakeLowercase{\textit{et al.}}: A Sample Article Using IEEEtran.cls for IEEE Journals}
+% \markboth{IEEE TRANSACTIONS ON AUDIO, SPEECH AND LANGUAGE PROCESSING, VOL -, 2025}%
+% {Shell \MakeLowercase{\textit{et al.}}: A Sample Article Using IEEEtran.cls for IEEE Journals}
+
+\markboth{%
+  \parbox{\textwidth}{%
+    This work has been submitted to the IEEE for possible publication. \\
+    Copyright may be transferred without notice, after which this version may no longer be accessible.%
+  }%
+}{}
 
@@
 After training, the model was fine-tuned using 20,000 out-of-domain sounds from the NSynth dataset~\cite{engel2017neural} and multi-scale spectrogram loss~\cite{engel2020ddsp}.
-This approach proved more effective---that is, lower multi-level spectral difference in out-of-domain tests---than baseline models, which were either not exposed to out-of-domain sounds or trained exclusively with P-Loss.
+This approach proved more effective---i.e, achieved lower multi-level spectral difference in out-of-domain tests---than baseline models, which were either not exposed to out-of-domain sounds or trained exclusively with P-Loss.
 Subjective hearing tests were conducted, showing a preference for the fine-tuned model~\cite{masuda2021soundmatch}.
@@
 Replication of arbitrary features of sounds is an important component of sound-design, though it is much harder to define or measure.
-Future works could explore loss functions which only measure certain characteristics of sound (such as \DTWEnv), and whether they can pave the way for better imitation sound-matching.
+Future works could explore loss functions which only measure certain characteristics of sound (such as \DTWEnv), and whether they can pave the way for better imitation in sound-matching.
 

### 2025-08-19 main.tex
Updates from Overleaf
began reviewer 1 revisions

@@
 \usepackage{caption}
+
+\newcommand{\highlight}[1]{\textcolor[RGB]{00,100,100}{#1}}
+\newcommand{\todo}[1]{\textcolor{red}{#1}}
+
 \captionsetup[lstlisting]{justification=centering, singlelinecheck=false}
 \providecommand{\gls}[1]{#1}
-\definecolor{highlight}{RGB}{00,150,00}
-\definecolor{todo}{RGB}{200,50,00}
 \definecolor{rank1}{HTML}{70FF70}
@@
 \newcommand{\PeriodicLoss}{\textbf{Periodic Loss}}
-\newcommand{\OutDomain}{\textbf{Out-of-Domain Generation}}
+\newcommand{\OutDomain}{\textbf{Out-Domain Generation}}
 \newcommand{\BPNoise}{\textbf{BP-Noise}}  
@@
 \section{Introduction}
-\subsection{Digital Synthesis With Sound-Matching}
 Since the 1970s, artists such as Suzanne Ciani~\cite{ciani_life_in_waves} have explored the use of audio synthesizers to replicate natural sounds—such as a bottle cap popping off—in ways that preserve the essential qualities of the original while showcasing the unique characteristics of the synthesizer.
@@
 
-A relatively small subset of these synthesis methods have been used for automatic sound-matching, notably additive and/or subtractive synthesizers~\cite{esling2019flow,yee2018automatic,mitchell2007evolutionary,salimi2020make}, physical modeling~\cite{riionheimo2003parameter}, and frequency/amplitude modulation (\gls{FM}/\gls{AM})~\cite{horner1993machine,vahidi2023mesostructures}.
+\todo{A relatively small subset of these synthesis methods have been used for automatic sound-matching, notably additive and/or subtractive synthesizers~\cite{esling2019flow,yee2018automatic,mitchell2007evolutionary,salimi2020make}, physical modeling~\cite{riionheimo2003parameter}, and frequency/amplitude modulation (\gls{FM}/\gls{AM})~\cite{horner1993machine,vahidi2023mesostructures}.
+}
 
@@
 \label{sec:sound_matching_definition}
-In past works the sound-matching problem has been tackled from a variety of perspectives as it involves a number of modular parts: the choice of synthesizer $g$, the target sounds of interest, the representation function $\phi$, and the heuristic for finding the optimal $\theta$.
+The problem of sound-matching has been tackled from a variety of perspectives as it involves a number of modular parts: the choice of synthesizer $g$, the target sounds of interest, the representation function $\phi$, and the heuristic for finding the optimal $\theta$.
 To give a formal definition of sound-matching, we expand the definition given by recent works~\cite{vahidi2023mesostructures,han2023perceptual}.
@@
 
-It could be argued that the implementation of these components is not an optimization problem with a correct answer, but a \textit{creative} endeavor depending on the artist's needs.
+\highlight{It could be argued that the implementation of these components is not an optimization problem with a correct answer, but a \textit{creative} endeavor depending on the artist's needs}.
 Consider the hypothetical case where the goal might be to create an 8-bit~\cite{collins2007loop} version of a high quality, organic snare drum: in such a case, there is no optimal or ``correct" version of an 8-bit snare sound.
@@
 The maximum number of iterations set to 200, and 300 experiments for each loss/synthesizer combination.
-Each synthesizer program is paired with four loss functions and 300 experiments are conducted and evaluated automatically with P-Loss and MSS.
-Two of the authors assigned blinded Likert scores to 40 randomly selected sounds, where integers from 1 to 5 are assigned, with 1 meaning no similarity and 5 for near identical.
+Each synthesizer program is paired with four loss functions and 300 experiments are conducted and evaluated automatically with P-Loss and MSS, and manually with Likert scores.
 
@@
 \subsubsection{STFT Losses}
-Due to their ubiquity and lower cost of gradient calculation, we use STFT as the basis of two of the loss functions.
+Due to their ubiquity and lower cost of gradient calculation, we use STFT as the basis of two loss functions.
 As discussed in Section~\ref{sec:matching_types}, the L1 or L2 distances are the most common methods of comparing different spectrograms, however, there are situations where this might lead to shortcomings.
 An example given by Vahidi \textit{et al.}~\cite{vahidi2023mesostructures} is two chirplets (tones that increase in pitch) starting and ending at the same frequencies.
-If one of these tones is slightly shifted in time, then the spectrograms will no longer overlap at any point, despite their sonic similarity.
+If one of these tones is slightly shifted in time, then the spectrograms will no longer overlap at any point, despite their sonic similarity (similar to the idea that a linear equation $f(x) = ax$ can never overlap with a slightly shifted version of itself, or $f(x) = ax +\epsilon$).
 In this example, we do not know if this is an issue of the spectrograms, or the L1 measure used for their comparison.
-In the computer vision field, Structural Similarity Index (SSIM)~\cite{wang2004imagesssim,wang2009mean} and Scale-Invariant Mean Squared Error (SIMSE)~\cite{barron2014shapessimse} have been used as alternatives that improved performance in various tasks relative to L1.
+In fields such as computer vision, Structural Similarity Index (SSIM)~\cite{wang2004imagesssim,wang2009mean} and Scale-Invariant Mean Squared Error (SIMSE)~\cite{barron2014shapessimse} have been used as alternatives that improved performance in various tasks relative to L1.
 

### 2025-08-20 main.tex
Updates from Overleaf

@@
 These issues have likely contributed to the limited exploration of this domain.
-Perhaps to most relevant recent work is by Vahidi \textit{et al.}, featuring the comparison of two differentiable loss functions with a simple FM synthesizer~\cite{vahidi2023mesostructures}.
+Perhaps the most relevant recent work is by Vahidi \textit{et al.}, featuring the comparison of two differentiable loss functions with a simple FM synthesizer~\cite{vahidi2023mesostructures}.
 
@@
 
+\todo{
 Genetic algorithms (\gls{GA})~\cite{holland1992genetic} can be used for direct sound optimization, and have been the earliest and most common heuristic in sound-matching~\cite{horner1993machine,mitchell2007evolutionary,yee2018automatic}.
 These algorithms start with arbitrary parameter sets that can be treated as an evolving population where the genomes are the parameter values.
-The most fit members of the group are the parameters which perform the best in the loss function, and create a new generation of parameters via mutation (random change in subset of parameters) and crossovers (combination of parameter sets); this process repeats until the goal or a maximum number of generations is reached.
+The most fit members of the group are the parameters which perform the best in the loss function, and create a new generation of parameters via mutation (random change in subset of parameters) and crossovers (combination of parameter sets); this process repeats until the goal or a maximum number of generations is reached.}
 
@@
 \label{sec:iterative_sound_matching}
- Figure~\ref{fig:sound_design_loop_iterative} shows a general model for iterative sound-matching.
+ \todo{Figure~\ref{fig:sound_design_loop_iterative} shows a general model for iterative sound-matching.
 To begin the process, the parameters are arbitrarily initialized (usually random generation), the similarity of the target and output is measured, and the parameter is \textit{optimized} with the goal of increasing the similarity, or alternatively, reducing the loss.
 This process repeats until termination.
-
+}
 \begin{figure}[ht]

### 2025-08-20 revised_main.tex
Updates from Overleaf

\documentclass[lettersize,journal]{IEEEtran}
\usepackage{amsmath,amsfonts}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{array}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{csquotes}
\usepackage{makecell}
% my added packages and commands =======================
\usepackage[numbers]{natbib} %removed cite package

\usepackage{tikz}
\usetikzlibrary{positioning,shapes.geometric}
\usepackage{adjustbox}
\usepackage{listings}
\usepackage{caption}

\newcommand{\highlight}[1]{\textcolor[RGB]{00,100,100}{#1}}
\newcommand{\todo}[1]{\textcolor{red}{#1}}

\captionsetup[lstlisting]{justification=centering, singlelinecheck=false}
\providecommand{\gls}[1]{#1}
\definecolor{rank1}{HTML}{70FF70}
\definecolor{rank2}{HTML}{858585}
\definecolor{rank3}{HTML}{454545}
\definecolor{rank4}{HTML}{000000}
\newcommand{\SIMSESpec}{\texttt{SIMSE\_Spec}}
\newcommand{\LoneSpec}{\texttt{L1\_Spec}}
\newcommand{\JTFS}{\texttt{JTFS}}
\newcommand{\DTWEnv}{\texttt{DTW\_Envelope}}
\newcommand{\LossSelect}{\textbf{Loss Selection}}
\newcommand{\SynthSelect}{\textbf{Synthesis Selection}}
\newcommand{\PeriodicLoss}{\textbf{Periodic Loss}}
\newcommand{\OutDomain}{\textbf{Out-Domain Generation}}
\newcommand{\BPNoise}{\textbf{BP-Noise}}  
\newcommand{\AddSineSaw}{\textbf{Add-SineSaw}}  
\newcommand{\AmpMod}{\textbf{Noise-AM}}  
\newcommand{\FMMod}{\textbf{SineSaw-AM}}

\input{code_style}

%  ================================
\begin{document}
\bibliographystyle{IEEEtran}

\title{Evaluating Sound Similarity Metrics for Differentiable, Iterative Sound-Matching}

\author{Amir Salimi, Abram Hindle, Osmar R. Za{\"i}ane}
\maketitle

\begin{IEEEkeywords}
Audio synthesis, differentiable digital signal processing, music information retrieval, sound-matching
\end{IEEEkeywords}
\hphantom{\vphantom{\begin{minipage}{\textwidth}\bibliography{references}\end{minipage}}}

\section{Background and Past Works}
\label{sec:background}
This section provides an introduction to digital audio synthesis, audio similarity measures, and a formal definition of sound-matching. We then provide a typology of sound-matching tasks using previous works as well as conceptual approaches that may be promising but rarely explored thus far. Finally, a discussion on the current gaps of knowledge in the field is provided.

\subsection{Digital Audio Synthesis}
\label{sec:diffSynth}
A digital audio synthesizer is any software used for the creation and manipulation of digital audio. Digital synthesizers use signal processing chains with a variety of digital signal processing (\gls{DSP}) functions to create sounds. These functions are often parametric, and the set of parameters given to the synthesizer's chain of DSP functions is called a synthesizer program.

The simplest form of DSP function is a sinusoidal tone. For example, consider:
\[ x(n) = sin( 2\pi n T)\]

where $T$ represents time in seconds, and $n$ is the frequency parameter. If $n$ is 1, then $x(n)$ would be a waveform with a frequency of 1 hertz, meaning that the value of $x(n)$ oscillates to the same point once every second. Waveform generators are called oscillators for this reason.

Since the advent of digital signal processing in the 1960s~\cite{stranneby2004digital}, a wide variety of parametric audio generation functions have been proposed~\cite{lyons1997understanding,russ1999sound,shier2020spiegelib}. This includes oscillators, filters, equalizers, and envelopes, which can be used sequentially or in parallel~\cite{lyons1997understanding,russ1999sound}. Sound design with a synthesizer is done by modifying the parameters of the functions until reaching a desired output~\cite{roads1996computer,pinch2004analog}.

Particularly in earlier works, frequency/amplitude modulators (\gls{FM}/\gls{AM}) has been the synthesizer of choice in sound-matching~\cite{horner1993machine,mitchell2007evolutionary,vahidi2023mesostructures}. This synthesis method is simple to implement, yet very expressive~\cite{chowning1973synthesis}. One of the earliest foundational sound-matching works is the analytical approach by Justice~\cite{justice1979analytic} toward the decomposition and recreation of sounds with FM synthesis using one carrier, one modulator, and the corresponding envelopes:
\[ f(t) = I_c(t) \cos(\omega_c t + I_m(t) \cos(\omega_m t))
\]
Where $f(t)$ is the output signal, $I_c(t)$ is the carrier amplitude, $\omega_c$ is the carrier frequency, and $I_m(t)$ and $\omega_m$ are the envelope and frequency for the modulator~\cite{justice1979analytic}.

% In general, relatively small subset of these synthesis methods have been used for automatic sound-matching, notably additive and/or subtractive synthesizers~\cite{esling2019flow,yee2018automatic,mitchell2007evolutionary,salimi2020make}, physical modeling~\cite{riionheimo2003parameter}.

% \subsection{Sound Representation and Loss Functions}
% \label{sec:loss_funcs}
% A digital sound (or an audio signal) is a series of numbers~\cite{smith1991viewpoints,smith2007mathematics}. To compare two digital sounds, the two corresponding series are passed to a function that measures their similarity. Two signals can sound identical to our ears, without having any values in common~\cite{moore2012introduction}. This necessitates the use of proxy representations (or feature extractors) when comparing sounds automatically. Similarity between the target sound and the synthesizer output is then measured by some form of subtraction and summation of the proxy representations.

% In sound-matching, particularly in a Deep Learning (\gls{DL}) context~\cite{goodfellow2016deep}, the similarity function can also be called a \textit{loss} function, where the emphasis is on measurement and reduction of the distance between target and output. It is important to note that there is a close relationship between the loss function $L$ and the sound representation function $\phi$. $L$ is the result of a distance measure $d$ applied to the features extracted by $\phi$.

% \[
% L(\theta, t) = d\langle\phi(t),\phi(x)\rangle
% \]

% \noindent

% A proxy representation is the output of the function \( \phi \), which can be thought of as a feature extraction function that maps the sounds \( t \) and \( x \) to their respective representations. 
% The proportionality or distance metric $d$ has typically been the L1 distance~\cite{turian2020sorry,richard2025model}, calculated as the mean of the absolute difference between every point in the proxy representation~\cite{engel2020ddsp,vahidi2023mesostructures}:
% \[
% L(\theta, t) = \left\| \phi(t) - \phi(x) \right\|_1
% \]

% Here we discuss four methods of audio representation and the corresponding loss functions. Three of which have been used in previous works, and a novel method not previously used in iterative sound-matching.

% \subsubsection{Parameter Loss}
% A common measure of similarity in sound-matching is the distance between synthesizer parameter sets, referred to as ``P-Loss"~\cite{han2023perceptual}. Typically, for the implementation of P-Loss the parameter sets are treated as vectors in space, and L1 or L2 distance is applied. There are two major limitations to this approach: First, the target and output sound must be made by the same synthesizer; otherwise the parameter sets cannot be compared (see Section~\ref{sec:matching_types}). Second, the relationship between synthesizer parameters and the audio output is not linear~\cite{shier2020spiegelib,han2023perceptual,esling2019flow}.

% \subsubsection{Fourier Spectrograms}
% \label{sec:fourier_specs}
% Fourier-based transformations such as short-time Fourier transforms (\gls{STFT}), Mel-spectrograms, and Mel-frequency cepstral coefficients have been viewed as the de facto and state-of-the-art representation of audio~\cite{beauchamp2003error,mitchell2007evolutionary,yee2018automatic}, however, there are many issues associated with their use in sound-matching~\cite{turian2020sorry,vahidi2023mesostructures,han2023perceptual,uzrad2024diffmoog}. Fourier transformations allow for the conversion of a signal from the time-domain to the frequency domain. Audio spectrograms can be generated by segmentation of a piece of audio into overlapping windows followed by the application of Fourier transforms to each window. They are costly to compute, but provide a better temporal view of changes in frequency content~\cite{muller2007dynamic,smith2007mathematics}. There are different types of spectrograms that have a basis in Fourier transformations, but the most notable and commonly used is the STFT.

% What we call \textit{Fourier-based Spectrograms} are variations on the STFT approach. For example, Mel-Spectrograms \textit{bin} frequencies on a near-logarithmic scale to better match human perception of frequencies~\cite{muller2007dynamic}. Multi-scale spectrograms (\gls{MSS}) used in recent works are a simple weighted average of multiple spectrograms with different parameters such as window size, number of frequency bins, and hop size~\cite{engel2020ddsp,vahidi2023mesostructures}; this may provide some improvements at a higher computational cost~\cite{turian2020sorry,engel2020ddsp}.

% \subsubsection{Joint-Time Frequency Spectrum}
% Recent works have focused on the limitations of parameter and spectral loss functions in sound-matching~\cite{vahidi2023mesostructures,uzrad2024diffmoog}, seeking to create more effective general solutions for the comparison of audio. 
% Noting the aforementioned weaknesses of comparing STFT spectrograms, Vahidi \textit{et al.} proposed differentiable Joint-Time Frequency Scattering (\gls{JTFS})~\cite{anden2015joint} as an alternative to spectrogram loss in sound-matching, and showed improved performance in sound-matching with differentiable chirplet synthesizers~\cite{vahidi2023mesostructures}. JTFS is the result of the application of a 2D wavelet transformation to the time-frequency representation of a signal~\cite{anden2015joint}.

% \subsubsection{Dynamic Envelope Warping}
% Dynamic Time Warping (DTW) is a method for measuring similarity between multi-dimensional time-series~\cite{rabiner1993fundamentals,muller2007dynamic,giorgino2009computing}. Given any two time-series $X = \{x_1,x_2,...,x_m\}$ and $Y = \{y_1,y_2,...,y_n\}$, we have indices $i\in\{1...m\}$ and $j\in\{1...n\}$ defining $X$ and $Y$. When the series are \textit{warped}, these indices change to expand or contract different portions of the series. To borrow the notation given by Muller~\cite{muller2007dynamic}, warped indices are a sequence $p=(p_1,...,p_L)$, where \(p_\ell = (m_\ell, n_\ell) \in [1 : m] \times [1 : n] \text{ for } \ell \in [1 : L]\), meaning that the indices for $X$ and $Y$ are reorganized under special conditions. In classical DTW, these conditions are \textit{monotonicity}, \textit{boundary matching}, and \textit{single step-size}.

% DTW measures the distance between the time-series \textit{after} alignments, typically using Euclidean distance. The distance between a time-series and shifted versions of itself would be 0, regardless of shift amount~\cite{tavenard.blog.dtw}. Additional rules can be imposed to keep alignments locally constrained~\cite{itakura1975minimum,sakoe1978dynamic}.

% \subsection{Automatic Sound Matching}
% \label{sec:sound_matching_definition}
% Sound-matching requires a number of modular parts: the choice of synthesizer $g$, the target sounds of interest, the representation function $\phi$, and the heuristic for finding the optimal $\theta$. To give a formal definition of sound-matching, we expand the definition given by recent works~\cite{vahidi2023mesostructures,han2023perceptual}. The major components are as follows: 
% \begin{itemize}
%     \item $g(\theta)$: Parametric audio synthesizer $g$ which takes on parameters $\theta$ 
%     \item $x$: The output of $g$, given a set of parameters $\theta$ or $g(\theta) = x$ 
%     \item $t$: The target sound which we want to replicate or imitate. 
%     \item $\phi$: Representation function or feature extractor. $\phi$ is applied to $x$ and $t$ to facilitate their comparison by the loss function.
%     \item $L$: Loss or error function. $L(\theta,t)$ is a measure of distance between $x$ and $t$. Often proportional to the subtraction of their representations, or $ \phi(x) - \phi(t)$
% \end{itemize}

% \highlight{It could be argued that the implementation of these components is not an optimization problem with a correct answer, but a \textit{creative} endeavor depending on the artist's needs}. Consider the hypothetical case where two sound-designers are searching for an 8-bit~\cite{collins2007loop} version of a high quality, organic snare drum: in such a case, there is no optimal solution, and the space of possible answers is extremely large. Here, it is possible (if not highly likely) that the two artists would select different samples given a number of possible samples to choose from.

% \subsection{Past Works}
% Given that sound-matching is ultimately an imitation of a creative endeavor, it might be no surprise that it can be categorized into different subtypes.

\end{document}

### 2025-08-21 revised_main.tex
Updates from Overleaf

@@
 This section provides an introduction to digital audio synthesis, audio similarity measures, and a formal definition of sound-matching.
-We then provide a typology of sound-matching tasks using previous works as well as conceptual approaches that may be promising but rarely explored thus far.
+We then provide a typology of sound-matching tasks using previous works as well as conceptual approaches that may be promising but though underexplored.
 Finally, a discussion on the current gaps of knowledge in the field is provided.
@@
 
-% In general, relatively small subset of these synthesis methods have been used for automatic sound-matching, notably additive and/or subtractive synthesizers~\cite{esling2019flow,yee2018automatic,mitchell2007evolutionary,salimi2020make}, physical modeling~\cite{riionheimo2003parameter}.
+Compared to the number of possible approaches to sound synthesis, a relatively small subset of synthesis methods have been analyzed in \textit{isolation} for sound-matching.
+Notable methods here are additive and/or subtractive synthesizers~\cite{engel2020ddsp,masuda2023improving,salimi2020make} and physical modeling~\cite{riionheimo2003parameter,han2024learning}.
+
+\highlight{
+What we mean here by \textit{isolation} is that the effect of the synthesizer's parameters (how they change the output sound and experiment outcome) is easily tractable.
+Recent computational advancements have allowed the use of complex Virtual Studio Technology (VST)\cite{steinberg1996vst} synthesizers in sound-matching~\cite{yee2018automatic,esling2019flow}, however, this has the trade-off here is the obfuscation of the interactions between the synthesizer functions, the loss, and the final outcome, among other distinctions which we will discuss in this section.
+}
+
+An important recent development in digital audio synthesis is the increase of interest in differentiable DSP (\gls{DDSP})~\cite{engel2020ddsp}.
+DDSP is a general term for applications that combine machine learning techniques such as gradient descent~\cite{goodfellow2016deep,boyd2004convex} with DSP.
+Implementing complex DSP functions in a differentiable manner can be challenging, and effective differentiable audio similarity measures require careful mathematical expression of the desired attributes of sound.
+These issues have likely contributed to the limited exploration of this domain~\cite{masuda2021soundmatch,vahidi2023mesostructures,uzrad2024diffmoog}.
 
@@
 
+\subsection{Sound Representation and Loss Functions}
+\label{sec:loss_funcs}
+A digital sound (or an audio signal) is a series of numbers~\cite{smith1991viewpoints,smith2007mathematics}.
+To compare two digital sounds, the two corresponding series are passed to a function that measures their similarity.
+Two signals can sound identical to our ears, without having any values in common~\cite{moore2012introduction}.
+This necessitates the use of proxy representations (or feature extractors) when comparing sounds automatically.
+Similarity between the target sound and the synthesizer output is then measured by some form of subtraction and summation of the proxy representations.
+
+In sound-matching, particularly in a Deep Learning (\gls{DL}) context~\cite{goodfellow2016deep}, the similarity function can also be called a \textit{loss} function, where the emphasis is on measurement and reduction of the distance between target and output.
+It is important to note that there is a close relationship between the loss function $L$ and the sound representation function $\phi$.
+$L$ is the result of a distance measure $d$ applied to the features extracted by $\phi$.
+
+\[
+L(\theta, t) = d\langle\phi(t),\phi(x)\rangle
+\]
+
+\noindent
+
+A proxy representation is the output of the function \( \phi \), which can be thought of as a feature extraction function that maps the sounds \( t \) and \( x \) to their respective representations. 
+The proportionality or distance metric $d$ has typically been the L1 distance~\cite{turian2020sorry,richard2025model}, calculated as the mean of the absolute difference between every point in the proxy representation~\cite{engel2020ddsp,vahidi2023mesostructures}:
+\[
+L(\theta, t) = \left\| \phi(t) - \phi(x) \right\|_1
+\]
+
+Here we discuss four methods of audio representation and the corresponding loss functions.
+Three of which have been used in previous works, and a novel method not previously used in iterative sound-matching.
+
+\subsubsection{Parameter Loss}
+A common measure of similarity in sound-matching is the distance between synthesizer parameter sets, referred to as ``P-Loss"~\cite{han2023perceptual}.
+Typically, for the implementation of P-Loss the parameter sets are treated as vectors in space, and L1 or L2 distance is applied.
+There are two major limitations to this approach: First, the target and output sound must be made by the same synthesizer; otherwise the parameter sets cannot be compared (see Section~\ref{sec:matching_types}).
+Second, the relationship between synthesizer parameters and the audio output is not linear~\cite{shier2020spiegelib,han2023perceptual,esling2019flow}.
+
+\subsubsection{Fourier Spectrograms}
+\label{sec:fourier_specs}
+Fourier-based transformations such as short-time Fourier transforms (\gls{STFT}), Mel-spectrograms, and Mel-frequency cepstral coefficients have been viewed as the de facto and state-of-the-art representation of audio~\cite{beauchamp2003error,mitchell2007evolutionary,yee2018automatic}, however, there are many issues associated with their use in sound-matching~\cite{turian2020sorry,vahidi2023mesostructures,han2023perceptual,uzrad2024diffmoog}.
+Fourier transformations allow for the conversion of a signal from the time-domain to the frequency domain.
+Audio spectrograms can be generated by segmentation of a piece of audio into overlapping windows followed by the application of Fourier transforms to each window.
+They are costly to compute, but provide a better temporal view of changes in frequency content~\cite{muller2007dynamic,smith2007mathematics}.
+There are different types of spectrograms that have a basis in Fourier transformations, but the most notable and commonly used is the STFT.
+
+What we call \textit{Fourier-based Spectrograms} are variations on the STFT approach.
+For example, Mel-Spectrograms \textit{bin} frequencies on a near-logarithmic scale to better match human perception of frequencies~\cite{muller2007dynamic}.
+Multi-scale spectrograms (\gls{MSS}) used in recent works are a simple weighted average of multiple spectrograms with different parameters such as window size, number of frequency bins, and hop size~\cite{engel2020ddsp,vahidi2023mesostructures}; this may provide some improvements at a higher computational cost~\cite{turian2020sorry,engel2020ddsp}.
+
+\subsubsection{Joint-Time Frequency Spectrum}
+Recent works have focused on the limitations of parameter and spectral loss functions in sound-matching~\cite{vahidi2023mesostructures,uzrad2024diffmoog}, seeking to create more effective general solutions for the comparison of audio.
+Noting the aforementioned weaknesses of comparing STFT spectrograms, Vahidi \textit{et al.} proposed differentiable Joint-Time Frequency Scattering (\gls{JTFS})~\cite{anden2015joint} as an alternative to spectrogram loss in sound-matching, and showed improved performance in sound-matching with differentiable chirplet synthesizers~\cite{vahidi2023mesostructures}.
+JTFS is the result of the application of a 2D wavelet transformation to the time-frequency representation of a signal~\cite{anden2015joint}.
+
+\subsubsection{Dynamic Envelope Warping}
+Dynamic Time Warping (DTW) is a method for measuring similarity between multi-dimensional time-series~\cite{rabiner1993fundamentals,muller2007dynamic,giorgino2009computing}.
+Given any two time-series $X = \{x_1,x_2,...,x_m\}$ and $Y = \{y_1,y_2,...,y_n\}$, we have indices $i\in\{1...m\}$ and $j\in\{1...n\}$ defining $X$ and $Y$.
+When the series are \textit{warped}, these indices change to expand or contract different portions of the series.
+To borrow the notation given by Muller~\cite{muller2007dynamic}, warped indices are a sequence $p=(p_1,...,p_L)$, where \(p_\ell = (m_\ell, n_\ell) \in [1 : m] \times [1 : n] \text{ for } \ell \in [1 : L]\), meaning that the indices for $X$ and $Y$ are reorganized under special conditions.
+In classical DTW, these conditions are \textit{monotonicity}, \textit{boundary matching}, and \textit{single step-size}.
+
+DTW measures the distance between the time-series \textit{after} alignments, typically using Euclidean distance.
+The distance between a time-series and shifted versions of itself would be 0, regardless of shift amount~\cite{tavenard.blog.dtw}.
+Additional rules can be imposed to keep alignments locally constrained~\cite{itakura1975minimum,sakoe1978dynamic}.
+% \textcolor{red}{DTW was originally used for signal comparison but using it the way we do is novel}
+
+\subsection{Automatic Sound Matching}
+\label{sec:sound_matching_definition}
+The problem of sound-matching has been tackled from a variety of perspectives as it involves a number of modular parts: the choice of synthesizer $g$, the target sounds of interest, the representation function $\phi$, and the heuristic for finding the optimal $\theta$.
+To give a formal definition of sound-matching, we expand the definition given by recent works~\cite{vahidi2023mesostructures,han2023perceptual}.
+The major components are as follows:
+\begin{itemize}
+    \item $g(\theta)$: Parametric audio synthesizer $g$ which takes on parameters $\theta$ 
+    \item $x$: The output of $g$, given a set of parameters $\theta$ or $g(\theta) = x$ 
+    \item $t$: The target sound which we want to replicate or imitate. 
+    \item $\phi$: Representation function or feature extractor. $\phi$ is applied to $x$ and $t$ to facilitate their comparison by the loss function.
+    \item $L$: Loss or error function. $L(\theta,t)$ is a measure of distance between $x$ and $t$. Often proportional to the subtraction of their representations, or $ \phi(x) - \phi(t)$
+\end{itemize}
+
+\highlight{It could be argued that the implementation of these components is not an optimization problem with a correct answer, but a \textit{creative} endeavor depending on the artist's needs}.
+Consider the hypothetical case where the goal might be to create an 8-bit~\cite{collins2007loop} version of a high quality, organic snare drum: in such a case, there is no optimal or ``correct" version of an 8-bit snare sound.
+
+\subsection{In-Domain versus Out-of-Domain}
+\label{sec:in-domain}
+The choice of domain depends on whether we want to use the same synthesizer for the target and output sounds, the scenario that is called \textit{in-domain}, or have target sounds that came from sources other than the synthesizer, or \textit{out-of-domain}.
+To paraphrase the description given by Masuda \textit{et al.}~\cite{masuda2021soundmatch}, if $g$, the synthesizer of choice, can accurately replicate the target sound $t$, or put differently, if $t$ itself is an output of $g$, then the sound-matching task is \textit{in-domain}.
+If $t$ is not an output of $g$, then the sound-matching task is \textit{out-of-domain}.
+In-domain tasks in general are simpler, and often there is a guarantee that there is a correct answer to the sound-matching problem, particularly if the goal is accurate replication of the sound.
+If the target sound is out-of-domain, replication is not guaranteed, and the goal becomes the \textit{imitation} of some aspect of sound.
+
+Regardless of the domain, the generation goal can be \textit{replication} or \textit{imitation} of the target sound.
+In replication, the goal is to make an identical copy of the target sound.
+Imitation is an artistic pursuit and harder to define, since the goal is to make new sounds that only retain a subset of the target's sonic features.
+While closely related to the in-domain versus out-of-domain problem, the choice of replication versus imitation is more dependent on how the loss and representation functions are defined.
+
+\subsection{Supervised vs Iterative}
+\label{sec:optimization}
+% The goal of sound-matching is to find the optimal parameters $\theta^*$ that minimize the loss between the synthesizer output and the target sound. 
+% \[
+% \theta^* = \arg\min_{\theta} L(\theta,t)
+% \]
+
+% The \textit{heuristics} (i.e., how the parameters are approximated) used in past works can be broadly split into two categories:
+% \begin{enumerate}
+%     \item Direct optimization methods~\cite{horner1993machine,mitchell2007evolutionary,yee2018automatic,vahidi2023mesostructures}.
+This requires a target example, and a loss function that measures the difference between the target and synthesizer outputs.
+Given these two necessary pieces, direct optimization refers to the iterative generation of a sound output, measurement the distance between target and output, and applying updates to the parameters.
+
+%     \item Supervised (or inference) methods which use datasets of sound corresponding parameters to learn the generation objective.
+Often with a large number of input and output examples and use of DNNs~\cite{engel2020ddsp,salimi2020make,yee2018automatic,esling2019flow}.
+Often, these models output their parameter predictions in a single step (or 1-shot).
+
+% \end{enumerate}
+
 \end{document}

### 2025-08-21 main.tex
Updates from Overleaf

@@
 \begin{enumerate}
-    \item Supervised (or inference) methods which use datasets of sound and parameter pairs to describe the generation objective.
-Often with a large number of input and output examples~\cite{engel2020ddsp,salimi2020make,yee2018automatic,esling2019flow}
+    \item Supervised (or inference) methods which use datasets of sound and parameter pairs to describe the generation objective, often with a large number of input and output examples~\cite{engel2020ddsp,salimi2020make,yee2018automatic,esling2019flow}
     \item Unsupervised or \textit{direct sound optimization} methods~\cite{horner1993machine,mitchell2007evolutionary,yee2018automatic,vahidi2023mesostructures}.

### 2025-08-21 revised_main.tex
Updates from Overleaf

@@
 
-\subsection{Supervised vs Iterative}
+\subsection{Heuristics and Historical Framing}
 \label{sec:optimization}
-% The goal of sound-matching is to find the optimal parameters $\theta^*$ that minimize the loss between the synthesizer output and the target sound. 
-% \[
-% \theta^* = \arg\min_{\theta} L(\theta,t)
-% \]
 
-% The \textit{heuristics} (i.e., how the parameters are approximated) used in past works can be broadly split into two categories:
-% \begin{enumerate}
-%     \item Direct optimization methods~\cite{horner1993machine,mitchell2007evolutionary,yee2018automatic,vahidi2023mesostructures}.
-This requires a target example, and a loss function that measures the difference between the target and synthesizer outputs.
-Given these two necessary pieces, direct optimization refers to the iterative generation of a sound output, measurement the distance between target and output, and applying updates to the parameters.
+The choice of \textit{heuristics} is yet another important attribute sound-matching.
+The goal of sound-matching is to find the optimal parameters $\theta^*$ that minimize the loss between the synthesizer output and the target sound.
+\[
+\theta^* = \arg\min_{\theta} L(\theta,t)
+\]
 
-%     \item Supervised (or inference) methods which use datasets of sound corresponding parameters to learn the generation objective.
-Often with a large number of input and output examples and use of DNNs~\cite{engel2020ddsp,salimi2020make,yee2018automatic,esling2019flow}.
-Often, these models output their parameter predictions in a single step (or 1-shot).
+The heuristics (i.e., how $\theta^*$ is approximated) used in past works can be broadly split into the two categories of \textit{direct optimization} and \textit{supervised}  (or inference) methods.
+Direct optimization refers to the iterative generation of a sound output, measurement of similarity between target and output, and application of updates to the parameters to maximize similarity (or minimize loss)~\cite{horner1993machine,mitchell2007evolutionary,yee2018automatic,vahidi2023mesostructures}; while supervised methods use large datasets of synthesizer sounds and corresponding parameters to learn the generation objective, commonly with the use of DNNs~\cite{engel2020ddsp,salimi2020make,yee2018automatic,esling2019flow}.
+These models often make their parameter predictions in a single step (or 1-shot).
 
-% \end{enumerate}
+Given the different attributes discussed thus far, Table~\ref{tab:summary} contains an overview of relevant literature, covering important works in sound-matching.
+A more detailed historical analysis of past and current trends in sound-matching can be discussed.
 

### 2025-08-21 revised_main.tex
Updates from Overleaf

@@
 This synthesis method is simple to implement, yet very expressive~\cite{chowning1973synthesis}.
-One of the earliest foundational sound-matching works is the analytical approach by Justice~\cite{justice1979analytic} toward the decomposition and recreation of sounds with FM synthesis using one carrier, one modulator, and the corresponding envelopes:
+A common implementation is the use of one carrier, one modulator, and the corresponding envelopes:
 \[ f(t) = I_c(t) \cos(\omega_c t + I_m(t) \cos(\omega_m t))
@@
 
-Given the different attributes discussed thus far, Table~\ref{tab:summary} contains an overview of relevant literature, covering important works in sound-matching.
-A more detailed historical analysis of past and current trends in sound-matching can be discussed.
+Given the different attributes discussed thus far, Table~\ref{tab:summary} contains an overview of important relevant literature.
+This leads us to a more detailed historical analysis of past works in sound-matching.
 

### 2025-08-21 main.tex
Updates from Overleaf

@@
 \hline
-Esling \textit{\textit{et al.}}~\cite{esling2019flow} & Non-diff VST & Spec MSS/SC P-loss & Supervised \& Modeling & Replicate & In* \& Out & 2019 & 1-shot \\
+Esling \textit{\textit{et al.}}~\cite{esling2019flow} & Non-diff VST & Spec MSS/SC P-Loss & Supervised \& Modeling & Replicate & In* \& Out & 2019 & 1-shot \\
 \hline
@@
 \hline
-Uzrad \textit{\textit{et al.}}~\cite{uzrad2024diffmoog} & DDSP chains & Synth. Chain and P-loss & Supervised & Replicate & In \& Out & 2024 & 1-shot \\
+Uzrad \textit{\textit{et al.}}~\cite{uzrad2024diffmoog} & DDSP chains & Synth. Chain and P-Loss & Supervised & Replicate & In \& Out & 2024 & 1-shot \\
 \hline
@@
 Table~\ref{tab:summary} contains an overview of relevant literature, covering important works in sound-matching.
-Earlier works in this domain focused on FM synthesis and genetic algorithms (or direct optimization methods)\cite{justice1979analytic,horner1993machine,mitchell2007evolutionary}.
+adsfEarlier works in this domain focused on FM synthesis and genetic algorithms (or direct optimization methods)\cite{justice1979analytic,horner1993machine,mitchell2007evolutionary}.
 Along with the exponential increase in computation power, later works utilized a wider range of approaches to parameter search, estimation, and audio synthesis; particularly with supervised learning and neural networks (\gls{NN})~\cite{yee2018automatic,engel2020ddsp,esling2019flow}.
@@
 For GAs and HC optimizer, MFCCs were used as a measure of performance as well as a loss function.
-For training the networks, P-loss was used, since differentiable MFCCs were not possible in their pipeline.
+For training the networks, P-Loss was used, since differentiable MFCCs were not possible in their pipeline.
 Importantly, informal hearing tests revealed that the performance of even the best NN model was unsatisfactory, possibly due to the complex nature of the synthesizer, which features 155 parameters.
@@
 This makes intuitive sense, as the visual effects of a band-pass filter on white noise are readily apparent in a spectrogram.
-As shown in Figure~\ref{fig:npsk_p0}, manual hearing test and MSS selected \SIMSESpec~as the best performer and \LoneSpec~as the second-best performer, while P-loss gave the reverse order.
+As shown in Figure~\ref{fig:npsk_p0}, manual hearing test and MSS selected \SIMSESpec~as the best performer and \LoneSpec~as the second-best performer, while P-Loss gave the reverse order.
 

### 2025-08-21 revised_main.tex
Updates from Overleaf

@@
 \subsection{Digital Audio Synthesis}
-\label{sec:diffSynth}
+\label{sec:dsp}
 A digital audio synthesizer is any software used for the creation and manipulation of digital audio.
@@
 What we mean here by \textit{isolation} is that the effect of the synthesizer's parameters (how they change the output sound and experiment outcome) is easily tractable.
-Recent computational advancements have allowed the use of complex Virtual Studio Technology (VST)\cite{steinberg1996vst} synthesizers in sound-matching~\cite{yee2018automatic,esling2019flow}, however, this has the trade-off here is the obfuscation of the interactions between the synthesizer functions, the loss, and the final outcome, among other distinctions which we will discuss in this section.
+Recent computational advancements have allowed the use of complex Virtual Studio Technology (VST)\cite{steinberg1996vst} synthesizers in sound-matching~\cite{yee2018automatic,esling2019flow}, however, the trade-off here is the obfuscation of the interactions between the synthesizer functions, the loss, and the final outcome, among other distinctions which we will discuss in this section.
 }
@@
 
-\subsection{Heuristics and Historical Framing}
+\subsection{Supvervised vs Direct Optimization}
 \label{sec:optimization}
@@
 
+Genetic algorithms (\gls{GA})~\cite{holland1992genetic} have been the earliest and most common heuristic for direct sound-matching~\cite{horner1993machine,mitchell2007evolutionary,yee2018automatic}.
+These algorithms start with arbitrary parameter sets that can be treated as an evolving population where the genomes are the parameter values.
+The most fit members of the group are the parameters which perform the best in the loss function, and create a new generation of parameters via mutation (random change in subset of parameters) and crossovers (combination of parameter sets); this process repeats until the goal or a maximum number of generations is reached.
+Rather than using random mutations, differentiable approaches allow goal-oriented updates to the synthesizer parameters (the goal being the minimization of loss), but with some drawbacks: other than requiring more computation power, differentiable functions require careful implementation of operations that are continuous and numerically stable; this makes the implementation of signal processing functions quite difficult, contributing to the scarcity of works in this domain.
+
+\subsection{Historical Framing of Sound-Matching}
 Given the different attributes discussed thus far, Table~\ref{tab:summary} contains an overview of important relevant literature.
@@
 
+Perhaps the earliest foundational work in sound-matching is the analytical approach by Justice~\cite{justice1979analytic} toward the decomposition and recreation of sounds using the simple FM synthesizer described in Section~\ref{sec:dsp}.
+The sound-matching works that followed used other heuristics for finding the synthesizer parameters, yet the simple FM synthesizer structure remained unchanged.
+Horner \textit{et al.}~\cite{horner1993machine} used GAs for re-synthesis of sounds with an FM synthesizer using one modulator and one carrier oscillator, and the McAulay-Quatieri method~\cite{mcaulay1986speech} for measuring loss.
+
+Introduction of more complex models of synthesis such as wavetables~\cite{horner2003auto} and physical modeling~\cite{riionheimo2003parameter} rendered interesting results, but led to further questions about the nature and goals of sound-matching.
+Mitchell and Creasy noted the difficulty in distinguishing between ``inefficiency of the optimization engine'' and ``synthesizer limitations'' as the cause of failure or success in sound-matching~\cite{mitchell2007evolutionary}.
+They offered a \textit{contrived methodology} for finding the best evolutionary method for sound-matching using their FM synthesizer.
+The methodology postulates that the best search heuristic for in-domain search (where an exact target exists, and a wide range of targets can be produced by sampling from different points in the parameter space) would also be the best-performing heuristic for out-of-domain search on a dataset of muted trumpet tones~\cite{opolko1989mcgill}; however, testing was not extensive and yielded some contradictory results.
+As they noted, modifications to the synthesizer, loss function, and sound domain would lead to a problem with an entirely new search space~\cite{mitchell2007evolutionary}.
+
+Recent years have seen more works in sound-matching using machine learning techniques.
+In 2018, Yee-King \textit{et al.} rendered 60,000 audio-parameter pairs from the \textit{Dexed} \gls{VST} synthesizer~\footnote{https://asb2m10.github.io/dexed/}, and showed that NNs trained on this dataset can outperform GA and hill-climber (\gls{HC})~\cite{hoffmann2000heuristic} methods in rendering speed, with slight improvements in MFCC error (used as an objective performance test).
+The speed improvement appears trivial, considering the iterative nature of GAs when compared to offline training of supervised models.
+For GAs and HC optimizer, MFCCs were used as a measure of performance as well as a loss function.
+For training the networks, P-Loss was used, since differentiable MFCCs were not possible in their pipeline.
+Importantly, informal hearing tests revealed that the performance of even the best NN model was unsatisfactory, possibly due to the complex nature of the synthesizer, which features 155 parameters.
+
+Masuda \textit{et al.} also applied supervised learning to synthesizer parameter estimation~\cite{masuda2021soundmatch}.
+Their work highlights the issue of non-linearity in parameter-to-synthesizer outputs and out-of-domain search.
+This work uses a differentiable subtractive synthesizer (two oscillators and an LP filter).
+A NN model was pre-trained using an in-domain dataset of randomly selected parameters and P-Loss.
+After training, the model was fine-tuned using 20,000 out-of-domain sounds from the NSynth dataset~\cite{engel2017neural} and multi-scale spectrogram loss~\cite{engel2020ddsp}.
+This approach proved more effective---i.e, achieved lower multi-level spectral difference in out-of-domain tests---than baseline models, which were either not exposed to out-of-domain sounds or trained exclusively with P-Loss.
+Subjective hearing tests were conducted, showing a preference for the fine-tuned model~\cite{masuda2021soundmatch}.
+Masuda \textit{et al.} later extended this work with semi-supervised learning, highlighting significant gaps in in-domain and out-of-domain performance~\cite{masuda2023improving}.
+
+Rather than focusing on a particular implementation, Shier \textit{et al.} presented Spieglib, a library for implementation of sound-matching pipelines~\cite{shier2020spiegelib}.
+This library provides different choices for DNNs, GAs, synthesizers, and feature extractors.
+Shier \textit{et al.} presented an experiment with a similar setup to Yee-King \textit{et al.}~\cite{yee2018automatic}, however, they found a genetic algorithm to be the best performing.
+
+Differentiable loss functions that use spectrogram differences can be computationally expensive.
+To mitigate this, Han \textit{et al.}~\cite{han2023perceptual} introduced ``perceptual-neural-physical loss'' (PNP).
+PNP is an approximation of loss functions; specifically, a loss function that uses the L2 norm of the difference between the features of two sounds, or $||\phi(t) - \phi(x)||^2_2$, where $\phi$ could be a spectrogram or JTFS function.
+PNP loss functions are fast and differentiable, but require training and parameter estimation.
+A Riemannian metric M needs to be calculated for the minimization of locally linear approximation of the ``true" spectral loss function~\cite{han2023perceptual}.
+\[
+\|\phi(t) - \phi(x)\|_2^2 = \langle \tilde{\theta} - \theta | \mathbf{M}(\theta) | \tilde{\theta} - \theta \rangle + O(\|\tilde{\theta} - \theta\|_2^3). \tag{4}
+\]
+
+During the training phase, this metric is calculated alongside the neural parameter estimator.
+After training, PNP can be used as a fast approximation of computationally demanding loss functions such as JTFS~\cite{han2023perceptual,han2024learning}, paired with FM synthesizers and a differentiable physical model of a drum head~\cite{smith2010physical}.
+
+Uzrad \textit{et al.}~\cite{uzrad2024diffmoog} took another unique approach to sound-matching: using a differentiable \textit{synthesis chain} of DSP generators and effects and a loss function that combines P-Loss with a \textit{signal-chain loss}.
+The synthesizer is a customizable chain of effects, which feed one output as input to the next step of the chain; signal-chain loss compares the parameter and output difference at every output step in the chain~\cite{uzrad2024diffmoog}.
+Possible chain functionalities are FM/AM, Low-Frequency Oscillators (\gls{LFO}), filters, and envelopes.
+Like the results shown by Masuda \textit{et al.}~\cite{masuda2021soundmatch}, better out-of-domain results were achieved when pre-trained on in-domain data and fine-tuned using out-of-domain NSynth data~\cite{engel2017neural}.
+
 \end{document}

### 2025-08-25 main.tex
Updates from Overleaf

@@
 
+% need to simplify the intro, and expand the hypothesis
 \section{Introduction}
-Since the 1970s, artists such as Suzanne Ciani~\cite{ciani_life_in_waves} have explored the use of audio synthesizers to replicate natural sounds—such as a bottle cap popping off—in ways that preserve the essential qualities of the original while showcasing the unique characteristics of the synthesizer.
+Since the 1970s, artists such as Suzanne Ciani~\cite{ciani_life_in_waves} have explored the use of audio synthesizers to replicate natural sounds—such as a bottle cap popping off—in ways that preserve the essential qualities of the original while showcasing the unique characteristics of the synthesizer~\cite{creativecherep2024}.
 Rather than merely replicating sounds, this approach uses synthesizers for artistic reinterpretation.
@@
 \hline
-Horner \textit{\textit{et al.}}~\cite{horner1993machine} & Non-diff FM & STFT (McAulay-Quatieri) & Direct & Replicate & In & 1991 & Iter \\
+Horner \textit{\textit{et al.}}~\cite{horner1993machine} & Non-diff FM & STFT (McAulay-Quatieri) & GA & Replicate & In & 1991 & Iter \\
 \hline
@@
 \hline
-Vahidi \textit{\textit{et al.}}~\cite{vahidi2023mesostructures} & Differentiable FM & JTFS/MSS & Direct (grad desc.) & Replicate & In & 2023 & Iter \\
+Vahidi \textit{\textit{et al.}}~\cite{vahidi2023mesostructures} & Differentiable FM & JTFS/MSS & Gradient desc. & Replicate & In & 2023 & Iter \\
+\hline
+Barkan \textit{et al.}~\cite{barkan2023inversynth}& Diff. VST Proxy& P-Loss + L1 Spec (weighted) &  Supervised (init.), Semi-Supervised & Replicate& In & 2023 & Both
+\\
 \hline
 Uzrad \textit{\textit{et al.}}~\cite{uzrad2024diffmoog} & DDSP chains & Synth. Chain and P-Loss & Supervised & Replicate & In \& Out & 2024 & 1-shot \\
+\hline
+Cherep \textit{et al.}~\cite{creativecherep2024} & DDSP & CLAP & Evolutionary & Imitate & Out & 2024 & Iter \\
 \hline

### 2025-08-26 revised_main.tex
Updates from Overleaf

@@
 \usepackage{amsmath,amsfonts}
+\DeclareMathOperator*{\argmin}{arg\,min}
 \usepackage{algorithmic}
@@
 These issues have likely contributed to the limited exploration of this domain~\cite{masuda2021soundmatch,vahidi2023mesostructures,uzrad2024diffmoog}.
-
-% \subsection{Sound Representation and Loss Functions}
-% \label{sec:loss_funcs}
-% A digital sound (or an audio signal) is a series of numbers~\cite{smith1991viewpoints,smith2007mathematics}.
-To compare two digital sounds, the two corresponding series are passed to a function that measures their similarity.
-Two signals can sound identical to our ears, without having any values in common~\cite{moore2012introduction}.
-This necessitates the use of proxy representations (or feature extractors) when comparing sounds automatically.
-Similarity between the target sound and the synthesizer output is then measured by some form of subtraction and summation of the proxy representations.
-
-% In sound-matching, particularly in a Deep Learning (\gls{DL}) context~\cite{goodfellow2016deep}, the similarity function can also be called a \textit{loss} function, where the emphasis is on measurement and reduction of the distance between target and output.
-It is important to note that there is a close relationship between the loss function $L$ and the sound representation function $\phi$.
-$L$ is the result of a distance measure $d$ applied to the features extracted by $\phi$.
-
-% \[
-% L(\theta, t) = d\langle\phi(t),\phi(x)\rangle
-% \]
-
-% \noindent
-
-% A proxy representation is the output of the function \( \phi \), which can be thought of as a feature extraction function that maps the sounds \( t \) and \( x \) to their respective representations.
-% The proportionality or distance metric $d$ has typically been the L1 distance~\cite{turian2020sorry,richard2025model}, calculated as the mean of the absolute difference between every point in the proxy representation~\cite{engel2020ddsp,vahidi2023mesostructures}:
-% \[
-% L(\theta, t) = \left\| \phi(t) - \phi(x) \right\|_1
-% \]
-
-% Here we discuss four methods of audio representation and the corresponding loss functions.
-Three of which have been used in previous works, and a novel method not previously used in iterative sound-matching.
-
-% \subsubsection{Parameter Loss}
-% A common measure of similarity in sound-matching is the distance between synthesizer parameter sets, referred to as ``P-Loss"~\cite{han2023perceptual}.
-Typically, for the implementation of P-Loss the parameter sets are treated as vectors in space, and L1 or L2 distance is applied.
-There are two major limitations to this approach: First, the target and output sound must be made by the same synthesizer; otherwise the parameter sets cannot be compared (see Section~\ref{sec:matching_types}).
-Second, the relationship between synthesizer parameters and the audio output is not linear~\cite{shier2020spiegelib,han2023perceptual,esling2019flow}.
-
-% \subsubsection{Fourier Spectrograms}
-% \label{sec:fourier_specs}
-% Fourier-based transformations such as short-time Fourier transforms (\gls{STFT}), Mel-spectrograms, and Mel-frequency cepstral coefficients have been viewed as the de facto and state-of-the-art representation of audio~\cite{beauchamp2003error,mitchell2007evolutionary,yee2018automatic}, however, there are many issues associated with their use in sound-matching~\cite{turian2020sorry,vahidi2023mesostructures,han2023perceptual,uzrad2024diffmoog}.
-Fourier transformations allow for the conversion of a signal from the time-domain to the frequency domain.
-Audio spectrograms can be generated by segmentation of a piece of audio into overlapping windows followed by the application of Fourier transforms to each window.
-They are costly to compute, but provide a better temporal view of changes in frequency content~\cite{muller2007dynamic,smith2007mathematics}.
-There are different types of spectrograms that have a basis in Fourier transformations, but the most notable and commonly used is the STFT.
-
-% What we call \textit{Fourier-based Spectrograms} are variations on the STFT approach.
-For example, Mel-Spectrograms \textit{bin} frequencies on a near-logarithmic scale to better match human perception of frequencies~\cite{muller2007dynamic}.
-Multi-scale spectrograms (\gls{MSS}) used in recent works are a simple weighted average of multiple spectrograms with different parameters such as window size, number of frequency bins, and hop size~\cite{engel2020ddsp,vahidi2023mesostructures}; this may provide some improvements at a higher computational cost~\cite{turian2020sorry,engel2020ddsp}.
-
-% \subsubsection{Joint-Time Frequency Spectrum}
-% Recent works have focused on the limitations of parameter and spectral loss functions in sound-matching~\cite{vahidi2023mesostructures,uzrad2024diffmoog}, seeking to create more effective general solutions for the comparison of audio.
-% Noting the aforementioned weaknesses of comparing STFT spectrograms, Vahidi \textit{et al.} proposed differentiable Joint-Time Frequency Scattering (\gls{JTFS})~\cite{anden2015joint} as an alternative to spectrogram loss in sound-matching, and showed improved performance in sound-matching with differentiable chirplet synthesizers~\cite{vahidi2023mesostructures}.
-JTFS is the result of the application of a 2D wavelet transformation to the time-frequency representation of a signal~\cite{anden2015joint}.
-
-% \subsubsection{Dynamic Envelope Warping}
-% Dynamic Time Warping (DTW) is a method for measuring similarity between multi-dimensional time-series~\cite{rabiner1993fundamentals,muller2007dynamic,giorgino2009computing}.
-Given any two time-series $X = \{x_1,x_2,...,x_m\}$ and $Y = \{y_1,y_2,...,y_n\}$, we have indices $i\in\{1...m\}$ and $j\in\{1...n\}$ defining $X$ and $Y$.
-When the series are \textit{warped}, these indices change to expand or contract different portions of the series.
-To borrow the notation given by Muller~\cite{muller2007dynamic}, warped indices are a sequence $p=(p_1,...,p_L)$, where \(p_\ell = (m_\ell, n_\ell) \in [1 : m] \times [1 : n] \text{ for } \ell \in [1 : L]\), meaning that the indices for $X$ and $Y$ are reorganized under special conditions.
-In classical DTW, these conditions are \textit{monotonicity}, \textit{boundary matching}, and \textit{single step-size}.
-
-% DTW measures the distance between the time-series \textit{after} alignments, typically using Euclidean distance.
-The distance between a time-series and shifted versions of itself would be 0, regardless of shift amount~\cite{tavenard.blog.dtw}.
-Additional rules can be imposed to keep alignments locally constrained~\cite{itakura1975minimum,sakoe1978dynamic}.
-
-% \subsection{Automatic Sound Matching}
-% \label{sec:sound_matching_definition}
-% Sound-matching requires a number of modular parts: the choice of synthesizer $g$, the target sounds of interest, the representation function $\phi$, and the heuristic for finding the optimal $\theta$.
-To give a formal definition of sound-matching, we expand the definition given by recent works~\cite{vahidi2023mesostructures,han2023perceptual}.
-The major components are as follows:
-% \begin{itemize}
-%     \item $g(\theta)$: Parametric audio synthesizer $g$ which takes on parameters $\theta$ 
-%     \item $x$: The output of $g$, given a set of parameters $\theta$ or $g(\theta) = x$ 
-%     \item $t$: The target sound which we want to replicate or imitate. 
-%     \item $\phi$: Representation function or feature extractor. $\phi$ is applied to $x$ and $t$ to facilitate their comparison by the loss function.
-%     \item $L$: Loss or error function. $L(\theta,t)$ is a measure of distance between $x$ and $t$. Often proportional to the subtraction of their representations, or $ \phi(x) - \phi(t)$
-% \end{itemize}
-
-% \highlight{It could be argued that the implementation of these components is not an optimization problem with a correct answer, but a \textit{creative} endeavor depending on the artist's needs}.
-Consider the hypothetical case where two sound-designers are searching for an 8-bit~\cite{collins2007loop} version of a high quality, organic snare drum: in such a case, there is no optimal solution, and the space of possible answers is extremely large.
-Here, it is possible (if not highly likely) that the two artists would select different samples given a number of possible samples to choose from.
-
-% \subsection{Past Works}
-% Given that sound-matching is ultimately an imitation of a creative endeavor, it might be no surprise that it can be categorized into different subtypes.
 
@@
 
-Recent years have seen more works in sound-matching using machine learning techniques.
+Recent years have seen more works in sound-matching using supervised machine learning techniques.
 In 2018, Yee-King \textit{et al.} rendered 60,000 audio-parameter pairs from the \textit{Dexed} \gls{VST} synthesizer~\footnote{https://asb2m10.github.io/dexed/}, and showed that NNs trained on this dataset can outperform GA and hill-climber (\gls{HC})~\cite{hoffmann2000heuristic} methods in rendering speed, with slight improvements in MFCC error (used as an objective performance test).
@@
 
+A neural approximation for another part of the supervised sound-matching chain---this time the synthesizer---was proposed by Barkan \textit{et al.}~\cite{barkan2023inversynthII}.
+As they noted, without a differentiable synthesizer, ``model-based'' (or what we call supervised) approaches cannot directly compare $x_o$ to $x_t$, often opting for P-Loss, which may not correctly map the parameters of sound to the output audio~\cite{esling2019flow,han2023perceptual,masuda2023improving}.
+Given a synthesizer and a large dataset of sounds and corresponding parameters, a model which maps sounds to parameters, and another model which approximates the synthesizer can be trained, allowing the implementation of a loss function which combines P-Loss with STFT differences.
+This ``Inversynth II'' (IS2) approach yielded significant improvements to previous works which did not use the STFT approximations for loss~\cite{esling2019flow,barkan2019inversynth}.
+
+In addition, Barkan \textit{et al.} attemped to  improve the IS2 model with Inference-Time
+Finetuning (ITF).
+For ITF, the synthesizer approximation is frozen and the encoder is iteratively finetuned for the sample sample using gradient descent in order to minimize P-Loss, stopping at convergence or when maximum number of iterations has reached~\cite{barkan2023inversynthII}.
+Barkan \textit{et al.} report that perhaps due to the discrepancy between the true and approximated synthesizer or overfitting, the application of ITF does not improve IS2 in spectral and manual performance metrics.
+
 Uzrad \textit{et al.}~\cite{uzrad2024diffmoog} took another unique approach to sound-matching: using a differentiable \textit{synthesis chain} of DSP generators and effects and a loss function that combines P-Loss with a \textit{signal-chain loss}.
@@
 
+% masuda and esling both noted out-of-domain drop in performance
+% yee king noted performance drop in manual surveys 
+%p-loss bad : masuda-improving, han2023, esling-flowsynth
+
 \end{document}

### 2025-08-26 main.tex
Updates from Overleaf

@@
 \hline
-Horner \textit{\textit{et al.}}~\cite{horner1993machine} & Non-diff FM & STFT (McAulay-Quatieri) & GA & Replicate & In & 1991 & Iter \\
+Horner \textit{\textit{et al.}}~\cite{horner1993machine} & Non-differentiable FM & STFT (McAulay-Quatieri) & GA & Replicate & In & 1991 & Iter \\
 \hline
-Mitchel \textit{\textit{et al.}}~\cite{mitchell2007evolutionary} & Non-diff FM & Spectral Relative & Evolutionary & Replicate & In (Contrived) & 2007 & Iter \\
+Mitchel \textit{\textit{et al.}}~\cite{mitchell2007evolutionary} & Non-differentiable FM & Spectral Relative & Evolutionary & Replicate & In (Contrived) & 2007 & Iter \\
 \hline
-Yee-King \textit{\textit{et al.}}~\cite{yee2018automatic} & Non-diff VST & P-Loss (for Supervised), Spectral (MFCC) & Supervised \& Evolutionary & Replicate & In & 2018 & Both \\
+Yee-King \textit{\textit{et al.}}~\cite{yee2018automatic} & Non-differentiable VST & P-Loss (for Supervised), Spectral (MFCC) & Supervised \& Evolutionary & Replicate & In & 2018 & Both \\
 \hline
-Esling \textit{\textit{et al.}}~\cite{esling2019flow} & Non-diff VST & Spec MSS/SC P-Loss & Supervised \& Modeling & Replicate & In* \& Out & 2019 & 1-shot \\
+Esling \textit{\textit{et al.}}~\cite{esling2019flow} & Non-differentiable VST & Spec MSS/SC P-Loss & Supervised \& Modeling & Replicate & In* \& Out & 2019 & 1-shot \\
 \hline
@@
 \hline
-Salimi \textit{\textit{et al.}}~\cite{salimi2020make} & Non-diff & STFT and Envelope & Supervised & Imitate & Out & 2020 & 1-shot \\
+Salimi \textit{\textit{et al.}}~\cite{salimi2020make} & DSP & STFT and Envelope & Supervised & Imitate & Out & 2020 & 1-shot \\
 \hline
@@
 \hline
-Barkan \textit{et al.}~\cite{barkan2023inversynth}& Diff. VST Proxy& P-Loss + L1 Spec (weighted) &  Supervised (init.), Semi-Supervised & Replicate& In & 2023 & Both
+Barkan \textit{et al.}~\cite{barkan2023inversynthII}& Diff. Synthesizer Proxy& P-Loss + L1 Spec (weighted) &  Supervised (init.), Semi-Supervised & Replicate& In & 2023 & Both
 \\
 \hline
-Uzrad \textit{\textit{et al.}}~\cite{uzrad2024diffmoog} & DDSP chains & Synth. Chain and P-Loss & Supervised & Replicate & In \& Out & 2024 & 1-shot \\
+Uzrad \textit{\textit{et al.}}~\cite{uzrad2024diffmoog} & DDSP & Synth. Chain and P-Loss & Supervised & Replicate & In \& Out & 2024 & 1-shot \\
 \hline

### 2025-08-26 revised_main.tex
Updates from Overleaf

@@
 
-\section{Background and Past Works}
-\label{sec:background}
-This section provides an introduction to digital audio synthesis, audio similarity measures, and a formal definition of sound-matching.
-We then provide a typology of sound-matching tasks using previous works as well as conceptual approaches that may be promising but though underexplored.
-Finally, a discussion on the current gaps of knowledge in the field is provided.
+% need to simplify the intro, and expand the hypothesis
+\section{Introduction}
+Since the 1970s, artists such as Suzanne Ciani~\cite{ciani_life_in_waves} have explored the use of audio synthesizers to replicate natural sounds—such as a bottle cap popping off—in ways that preserve the essential qualities of the original while showcasing the unique characteristics of the synthesizer~\cite{creativecherep2024}.
+Rather than merely replicating sounds, this approach uses synthesizers for artistic reinterpretation.
+However, the automation of this interpretive approach to sound design using digital audio synthesizers has rarely been explored, and significant knowledge gaps remain even in the simpler problem of replicating sounds.
+One such gap is the lack of diversity and depth in the analysis of sound-similarity measures and their interaction with various synthesis methods.
+As a stepping stone toward automated interpretive sound design, we use automatic and manual sound-evaluations to investigate whether the optimal choice of a sound-similarity measure is dependent on the synthesis method.
 
@@
 
-In addition, Barkan \textit{et al.} attemped to  improve the IS2 model with Inference-Time
+Barkan \textit{et al.} attemped to  improve the IS2 model with Inference-Time
 Finetuning (ITF).
 For ITF, the synthesizer approximation is frozen and the encoder is iteratively finetuned for the sample sample using gradient descent in order to minimize P-Loss, stopping at convergence or when maximum number of iterations has reached~\cite{barkan2023inversynthII}.
-Barkan \textit{et al.} report that perhaps due to the discrepancy between the true and approximated synthesizer or overfitting, the application of ITF does not improve IS2 in spectral and manual performance metrics.
+Barkan \textit{et al.} report that perhaps due to the discrepancy between the true and approximated synthesizer or overfitting, the application of ITF to IS2 generally worsened the spectral and manual performance metrics.
 
@@
 
+Some recent works have opted for the use of audio embeddings as a similarity metric.
+In a recent work, Cherep \textit{et al.} used latent representations from the CLAP model~\cite{wu2023large} along with a differentiable synthesizer~\cite{synthhaxcherep2023} to create creative interpretations of sound effects.
+In their approach, a desired sound-effect is described in text and embedded using clap, and gradient free optimizer~\cite{evosax2022github} iteratively updates the synthesizer's parameters to minimize difference in the text and output sound embeddings.
+Based on manual hearing tests, this approach did not produce the ``correct'' sounds more frequently than previous works~\cite{kreuk2022audiogen}, however, it did yield better scores for ``artistic-interpretation'' (or what we refer to as imitation).
+
 % masuda and esling both noted out-of-domain drop in performance
@@
 %p-loss bad : masuda-improving, han2023, esling-flowsynth
+% manual performance measure: barkan, masuda2021soundmatch, yee-king
 

### 2025-08-27 revised_main.tex
Updates from Overleaf

@@
 As a stepping stone toward automated interpretive sound design, we use automatic and manual sound-evaluations to investigate whether the optimal choice of a sound-similarity measure is dependent on the synthesis method.
+
+A digital audio synthesizer is any software used for the creation and manipulation of audio.
+The possibilities afforded by these synthesizers is infinite, leading to their widespread adaption by artists and sound-designers~\cite{lyons1997understanding,russ1999sound,stranneby2004digital}.
+A typical synthesizer has a number of parameterizable functions which affect the output sound in various ways, and the manual approach involves modifying the parameters until a desired sound is reached.
+
+The automation of the aforementioned approach is often referred to as "sound-matching", with reduction of tinkering time and creation of ``interesting" sounds as the main motivators \cite{krekovic2019insights,turian2020sorry,horner1993machine,salimi2020make,esling2019flow,engel2020ddsp,mitchell2007evolutionary,shier2020spiegelib,masuda2021soundmatch,masuda2023improving}.
+Sound-matching involves a target sound and a similarity metric (or loss function), and a heuristic to find parameters for a synthesizer that replicate \textit{all} or \textit{some} of the characteristics of a target sound as best as possible~\cite{horner1993machine,mitchell2007evolutionary,masuda2023improving}.
+Despite decades of research, major gaps of knowledge prevent sound-matching as practical approach for sound-designers.
+Here we highlight major issues in the field, leading us to
 

### 2025-08-28 revised_main.tex
Updates from Overleaf

@@
 Sound-matching involves a target sound and a similarity metric (or loss function), and a heuristic to find parameters for a synthesizer that replicate \textit{all} or \textit{some} of the characteristics of a target sound as best as possible~\cite{horner1993machine,mitchell2007evolutionary,masuda2023improving}.
-Despite decades of research, major gaps of knowledge prevent sound-matching as practical approach for sound-designers.
-Here we highlight major issues in the field, leading us to
+
+Through our review of the history of sound-matching, we provide a nomenclature for its different subtypes, and illuminate major perennial issues that prevent sound-matching as a practical approach for sound-designers.
+Here we are concerned with the issues of \LossSelect~and \SynthSelect.
+The former refers to the search for an optimal loss function, lack of novel algorithms, and the---perhaps needless---focus on outperforming state of the art (\gls{SOTA}), while the latter refers to the lack of diversity in synthesis methods.
+As a consequence of these problems, not much is known about the interaction between different methods of synthesis and different loss functions: Should we be searching for a universally best performing loss function across synthesizers, or
+
+While there have been many works and experiments comparing the accuracy of loss functions and choosing the best performers~\cite{vahidi2023mesostructures,turian2020sorry,engel2020ddsp,uzrad2024diffmoog,han2023perceptual,masuda2021soundmatch,turian2020sorry,bruford2024synthesizer}, we note that these claims regarding the effectiveness of one function versus another have consistently been made in limited contexts that may not generalize to other settings as typically only a single (and often, simple) synthesizer has been used for testing the effectiveness of loss functions.
+
+%   A primary conclusion of the work is that there is no one best sound similarity measure capable of working across all synthesizers.
+% The core problem that no single loss function performs optimally across synthesizer types
 

### 2025-08-28 revised_main.tex
Updates from Overleaf

@@
 The former refers to the search for an optimal loss function, lack of novel algorithms, and the---perhaps needless---focus on outperforming state of the art (\gls{SOTA}), while the latter refers to the lack of diversity in synthesis methods.
-As a consequence of these problems, not much is known about the interaction between different methods of synthesis and different loss functions: Should we be searching for a universally best performing loss function across synthesizers, or
+As a consequence of these problems, not much is known about the interaction between different methods of synthesis and different loss functions.
+While there have been many works and experiments comparing the accuracy of loss functions and choosing the best performers~\cite{vahidi2023mesostructures,turian2020sorry,engel2020ddsp,uzrad2024diffmoog,han2023perceptual,masuda2021soundmatch,turian2020sorry,bruford2024synthesizer}, we note that these claims regarding the effectiveness of one function versus another have consistently been made in limited contexts that may not generalize to other settings as typically only a single (and often, simple) synthesizer has been used for testing the effectiveness of loss functions.
+Should we be searching for a universally best performing loss function across synthesizers?
+Does the method of synthesis influence the selection of the ``optimal" loss function, and if so, do current, SOTA loss functions satisfy the creative needs of sound-designers?
 
-While there have been many works and experiments comparing the accuracy of loss functions and choosing the best performers~\cite{vahidi2023mesostructures,turian2020sorry,engel2020ddsp,uzrad2024diffmoog,han2023perceptual,masuda2021soundmatch,turian2020sorry,bruford2024synthesizer}, we note that these claims regarding the effectiveness of one function versus another have consistently been made in limited contexts that may not generalize to other settings as typically only a single (and often, simple) synthesizer has been used for testing the effectiveness of loss functions.
+The main hypothesis here is that the performance of a similarity measure (or loss function) is influenced by other factors in the environment, particularly the method of synthesis.
+For the experiments, a \textit{differentiable}, \textit{iterative} approach to sound-matching is used.
+The iterative approach  better mimics the manual process of sound design (recursively listening to synthesizer outputs and adjusting the parameters accordingly) while allowing access to the gradients that can be used to better understand the nature of the problem.
+
+% The simple synthesizers tested here represent fundimental building blocks of more complex implementations, and direct use of loss function gradients assuming
+
+We combine 4 loss functions with 4 differentiable synthesizers, and compare their performance as measured by the similarity of the final output to the target.
+We evaluate the final similarity with two different automatic methods, as well as manual listening tests conducted by two of the authors.
+We compare the distributions of the evaluation scores in order to rank the loss functions from best to worst.
+The results of these experiments show that when accounting for different synthesizers, there is no consistent choice of sound similarity measure for optimal performance in sound-matching.
+
+Here, several common characteristics of previous work are avoided in order to gain fundamental, practical takeaways from the experiments.
+In order to directly analyze the interaction between the loss and method of synthesis, the synthesizer here are implemented with simple, functions that are used as the building blocks of complex synthesizers, and the gradients flow from the loss function to the synthesizer parameters with no extra layers between the two.
+Furthermore, since the true target parameters are unknown in any realistic sound-design scenario, parameter loss (P-Loss) is only used as a secondary measure of performance, and not used as a loss function.
 
@@
 
+\section{Background and Related Works}
 \subsection{Digital Audio Synthesis}

### 2025-09-04 revised_main.tex
Updates from Overleaf

@@
 Rather than merely replicating sounds, this approach uses synthesizers for artistic reinterpretation.
-However, the automation of this interpretive approach to sound design using digital audio synthesizers has rarely been explored, and significant knowledge gaps remain even in the simpler problem of replicating sounds.
+However, the automation of this interpretive approach to sound design using digital audio synthesizers remains impractical for sound-designers, and significant knowledge gaps remain even in the simpler problem of replicating sounds.
 One such gap is the lack of diversity and depth in the analysis of sound-similarity measures and their interaction with various synthesis methods.
@@
 The possibilities afforded by these synthesizers is infinite, leading to their widespread adaption by artists and sound-designers~\cite{lyons1997understanding,russ1999sound,stranneby2004digital}.
-A typical synthesizer has a number of parameterizable functions which affect the output sound in various ways, and the manual approach involves modifying the parameters until a desired sound is reached.
+A typical synthesizer has a number of parameterizable functions which affect the output sound in various ways, and the manual approach often involves modifying the parameters until a conceptualized target sound is reached.
 
-The automation of the aforementioned approach is often referred to as "sound-matching", with reduction of tinkering time and creation of ``interesting" sounds as the main motivators \cite{krekovic2019insights,turian2020sorry,horner1993machine,salimi2020make,esling2019flow,engel2020ddsp,mitchell2007evolutionary,shier2020spiegelib,masuda2021soundmatch,masuda2023improving}.
+The automation of the aforementioned approach has been commonly referred to as "sound-matching", with reduction of tinkering time and creation of ``interesting" sounds as the main motivators \cite{krekovic2019insights,turian2020sorry,horner1993machine,salimi2020make,esling2019flow,engel2020ddsp,mitchell2007evolutionary,shier2020spiegelib,masuda2021soundmatch,masuda2023improving}.
 Sound-matching involves a target sound and a similarity metric (or loss function), and a heuristic to find parameters for a synthesizer that replicate \textit{all} or \textit{some} of the characteristics of a target sound as best as possible~\cite{horner1993machine,mitchell2007evolutionary,masuda2023improving}.
 
+We provide a review of past works in order to understand what is lacking in the field, and how it can be improved.
 Through our review of the history of sound-matching, we provide a nomenclature for its different subtypes, and illuminate major perennial issues that prevent sound-matching as a practical approach for sound-designers.
@@
 The former refers to the search for an optimal loss function, lack of novel algorithms, and the---perhaps needless---focus on outperforming state of the art (\gls{SOTA}), while the latter refers to the lack of diversity in synthesis methods.
-As a consequence of these problems, not much is known about the interaction between different methods of synthesis and different loss functions.
+As a consequence of these problems, we find that not much is known about the interaction between different methods of synthesis and different loss functions.
 While there have been many works and experiments comparing the accuracy of loss functions and choosing the best performers~\cite{vahidi2023mesostructures,turian2020sorry,engel2020ddsp,uzrad2024diffmoog,han2023perceptual,masuda2021soundmatch,turian2020sorry,bruford2024synthesizer}, we note that these claims regarding the effectiveness of one function versus another have consistently been made in limited contexts that may not generalize to other settings as typically only a single (and often, simple) synthesizer has been used for testing the effectiveness of loss functions.
@@
 
-The main hypothesis here is that the performance of a similarity measure (or loss function) is influenced by other factors in the environment, particularly the method of synthesis.
+The main hypothesis here is that \textit{the performance of a similarity measure (or loss function) is influenced by other factors in the environment, particularly the method of synthesis}.
 For the experiments, a \textit{differentiable}, \textit{iterative} approach to sound-matching is used.

### 2025-09-04 revised_main.tex
Updates from Overleaf

@@
 As a consequence of these problems, we find that not much is known about the interaction between different methods of synthesis and different loss functions.
-While there have been many works and experiments comparing the accuracy of loss functions and choosing the best performers~\cite{vahidi2023mesostructures,turian2020sorry,engel2020ddsp,uzrad2024diffmoog,han2023perceptual,masuda2021soundmatch,turian2020sorry,bruford2024synthesizer}, we note that these claims regarding the effectiveness of one function versus another have consistently been made in limited contexts that may not generalize to other settings as typically only a single (and often, simple) synthesizer has been used for testing the effectiveness of loss functions.
-Should we be searching for a universally best performing loss function across synthesizers?
+We believe that clarifying the relationship between these two components is a necessary step for the prioritization of future works towards practical sound-matching.
+
+The main hypothesis here is that \textit{the performance of a similarity measure (or loss function) is influenced by other factors in the environment, particularly the method of synthesis}.
+While there have been many works and experiments comparing the accuracy of loss functions~\cite{vahidi2023mesostructures,turian2020sorry,engel2020ddsp,uzrad2024diffmoog,han2023perceptual,masuda2021soundmatch,turian2020sorry,bruford2024synthesizer}, we note that claims regarding the effectiveness of one function versus another have consistently been made in limited contexts that may not generalize to other settings; For example, it is typical for only a single (and often, simple) synthesizer to be used for testing the effectiveness of loss functions.
+This main hypothesis would help with important questions regarding the path of future research: Should we be searching for a universally best performing loss function across synthesizers?
 Does the method of synthesis influence the selection of the ``optimal" loss function, and if so, do current, SOTA loss functions satisfy the creative needs of sound-designers?
 
-The main hypothesis here is that \textit{the performance of a similarity measure (or loss function) is influenced by other factors in the environment, particularly the method of synthesis}.
-For the experiments, a \textit{differentiable}, \textit{iterative} approach to sound-matching is used.
-The iterative approach  better mimics the manual process of sound design (recursively listening to synthesizer outputs and adjusting the parameters accordingly) while allowing access to the gradients that can be used to better understand the nature of the problem.
-
-% The simple synthesizers tested here represent fundimental building blocks of more complex implementations, and direct use of loss function gradients assuming
-
-We combine 4 loss functions with 4 differentiable synthesizers, and compare their performance as measured by the similarity of the final output to the target.
+For the experiments, an \textit{iterative} and \textit{differentiable}, approach to sound-matching is used.
+The iterative approach better mimics the manual process of sound design (recursively listening to synthesizer outputs and adjusting the parameters accordingly) while a differentiable environment allows access to the loss function gradients that can be used to better understand the nature of the problem.
+We define four differentiable synthesizers (each showcasing a fundamental method of synthesis used in modern synthesizers) and pair them with four different loss functions (two established and two novel methods).
 We evaluate the final similarity with two different automatic methods, as well as manual listening tests conducted by two of the authors.
@@
 
-Here, several common characteristics of previous work are avoided in order to gain fundamental, practical takeaways from the experiments.
+% Here, several common characteristics of previous work are avoided in order to gain fundamental, practical takeaways from the experiments.
 In order to directly analyze the interaction between the loss and method of synthesis, the synthesizer here are implemented with simple, functions that are used as the building blocks of complex synthesizers, and the gradients flow from the loss function to the synthesizer parameters with no extra layers between the two.
@@
 
+% The simple synthesizers tested here represent fundimental building blocks of more complex implementations, and direct use of loss function gradients assuming 
 %   A primary conclusion of the work is that there is no one best sound similarity measure capable of working across all synthesizers.

### 2025-09-05 revised_main.tex
Updates from Overleaf

@@
 %  ================================
+
 \begin{document}
@@
 
+\begin{abstract}
+Manual sound design with a synthesizer is inherently iterative: an artist compares the synthesized output to a mental target, adjusts parameters, and repeats until satisfied.
+Iterative sound-matching automates this workflow by continually programming a synthesizer under the guidance of a loss function (or similarity measure) towards a target sound.
+Prior comparisons of loss functions have typically favored one metric over another, but only within narrow settings: limited synthesis methods, few loss types, often without blind listening tests.
+This leaves open the question of whether a universally optimal loss exists, or the choice of loss remains a creative decision conditioned on the synthesis method and the sound designer's preference.
+We propose differentiable iterative sound-matching as the natural extension of the available literature, since it combines the manual approach to sound design with modern advances in machine learning.
+To analyze the variability of loss function performance across synthesizers, we implemented a mix of four novel and established differentiable loss functions, and paired them with differentiable subtractive, additive, and AM synthesizers.
+For each of the sixteen synthesizer–loss combinations, we ran 300 randomized sound-matching trials.
+Performance was measured using parameter differences, spectrogram-distance metrics, and manually assigned listening scores.
+We observed a moderate level of consistency among the three performance measures.
+Our post-hoc analysis shows that the loss function performance is highly dependent on the synthesizer.
+These findings underscore the value of expanding the scope of sound-matching experiments, and developing new similarity metrics tailored to specific synthesis techniques, rather than pursuing one-size-fits-all solutions.
+\end{abstract}
 \begin{IEEEkeywords}
@@
 \section{Introduction}
-Since the 1970s, artists such as Suzanne Ciani~\cite{ciani_life_in_waves} have explored the use of audio synthesizers to replicate natural sounds—such as a bottle cap popping off—in ways that preserve the essential qualities of the original while showcasing the unique characteristics of the synthesizer~\cite{creativecherep2024}.
+
+% Since the 1970s, artists such as Suzanne Ciani~\cite{ciani_life_in_waves} have explored the use of audio synthesizers to replicate natural sounds—such as a bottle cap popping off—in ways that preserve the essential qualities of the original while showcasing the unique characteristics of the synthesizer~\cite{creativecherep2024}.
 Rather than merely replicating sounds, this approach uses synthesizers for artistic reinterpretation.
@@
 The possibilities afforded by these synthesizers is infinite, leading to their widespread adaption by artists and sound-designers~\cite{lyons1997understanding,russ1999sound,stranneby2004digital}.
-A typical synthesizer has a number of parameterizable functions which affect the output sound in various ways, and the manual approach often involves modifying the parameters until a conceptualized target sound is reached.
+A typical synthesizer has a number of parameterizable functions which affect the output sound in various ways, and manual sound-design often involves modifying the parameters until a conceptualized target sound is reached.
+The automation of this approach has commonly been referred to as "sound-matching", with reduction of tinkering time and creation of ``interesting" sounds as the main motivators ~\cite{krekovic2019insights,turian2020sorry,horner1993machine,salimi2020make,esling2019flow,engel2020ddsp,mitchell2007evolutionary,shier2020spiegelib,masuda2021soundmatch,masuda2023improving}.
+The main requirements of sound-matching are a target sound, a similarity metric (or loss function), and a heuristic to find parameters for a synthesizer that replicate \textit{all} or \textit{some} of the characteristics of the target sound as best as possible~\cite{horner1993machine,mitchell2007evolutionary,masuda2023improving}.
+However, despite the seeming simplicity of the problem and decades of research, sound-matching is not yet practical solution for sound-designers.
+Here we are motivated to understand why this is the case, and what future research should prioritize in order to improve the field.
 
-The automation of the aforementioned approach has been commonly referred to as "sound-matching", with reduction of tinkering time and creation of ``interesting" sounds as the main motivators \cite{krekovic2019insights,turian2020sorry,horner1993machine,salimi2020make,esling2019flow,engel2020ddsp,mitchell2007evolutionary,shier2020spiegelib,masuda2021soundmatch,masuda2023improving}.
-Sound-matching involves a target sound and a similarity metric (or loss function), and a heuristic to find parameters for a synthesizer that replicate \textit{all} or \textit{some} of the characteristics of a target sound as best as possible~\cite{horner1993machine,mitchell2007evolutionary,masuda2023improving}.
-
-We provide a review of past works in order to understand what is lacking in the field, and how it can be improved.
-Through our review of the history of sound-matching, we provide a nomenclature for its different subtypes, and illuminate major perennial issues that prevent sound-matching as a practical approach for sound-designers.
-Here we are concerned with the issues of \LossSelect~and \SynthSelect.
+We provide a review of past works in order to understand what is lacking, a nomenclature for different subtypes of sound-matching, and illuminate major perennial issues.
+In particular, we are concerned with the issues of \LossSelect~and \SynthSelect.
 The former refers to the search for an optimal loss function, lack of novel algorithms, and the---perhaps needless---focus on outperforming state of the art (\gls{SOTA}), while the latter refers to the lack of diversity in synthesis methods.
 As a consequence of these problems, we find that not much is known about the interaction between different methods of synthesis and different loss functions.
-We believe that clarifying the relationship between these two components is a necessary step for the prioritization of future works towards practical sound-matching.
+How much performance variation can we expect?
 
 The main hypothesis here is that \textit{the performance of a similarity measure (or loss function) is influenced by other factors in the environment, particularly the method of synthesis}.
-While there have been many works and experiments comparing the accuracy of loss functions~\cite{vahidi2023mesostructures,turian2020sorry,engel2020ddsp,uzrad2024diffmoog,han2023perceptual,masuda2021soundmatch,turian2020sorry,bruford2024synthesizer}, we note that claims regarding the effectiveness of one function versus another have consistently been made in limited contexts that may not generalize to other settings; For example, it is typical for only a single (and often, simple) synthesizer to be used for testing the effectiveness of loss functions.
-This main hypothesis would help with important questions regarding the path of future research: Should we be searching for a universally best performing loss function across synthesizers?
-Does the method of synthesis influence the selection of the ``optimal" loss function, and if so, do current, SOTA loss functions satisfy the creative needs of sound-designers?
+While there have been many works and experiments comparing the accuracy of loss functions~\cite{vahidi2023mesostructures,turian2020sorry,engel2020ddsp,uzrad2024diffmoog,han2023perceptual,masuda2021soundmatch,turian2020sorry,bruford2024synthesizer}, we note that claims regarding the effectiveness of one function versus another have consistently been made in limited contexts that may not generalize to other settings.
+Beyond this central hypothesis, we also investigate three additional research questions: (Q1) To what extent do automatic evaluation metrics agree with manual listening tests?
+(Q2) Can novel loss functions utilizing Dynamic Time Warping (DTW) and Scale-Invariant Mean Squared Error (SIMSE) provide advantages over standard loss functions?
+(Q3) Is an iterative differentiable optimization a viable strategy for design of sound-matching experiments?
 
-For the experiments, an \textit{iterative} and \textit{differentiable}, approach to sound-matching is used.
-The iterative approach better mimics the manual process of sound design (recursively listening to synthesizer outputs and adjusting the parameters accordingly) while a differentiable environment allows access to the loss function gradients that can be used to better understand the nature of the problem.
-We define four differentiable synthesizers (each showcasing a fundamental method of synthesis used in modern synthesizers) and pair them with four different loss functions (two established and two novel methods).
+% This main hypothesis (along with our forthcoming research questions) would help with important issues regarding the path of future research.
+Should we be searching for a universally best performing loss function across synthesizers?
+Do current, SOTA loss functions satisfy the creative needs of sound-designers?
+
+ The iterative approach (commonly used in earlier works) better mimics the manual process of sound design (recursively listening to synthesizer outputs and adjusting the parameters accordingly) while a differentiable environment allows access to the loss function gradients that can be used to better understand the nature of the problem.
+We define four differentiable synthesizers (each showcasing a fundamental method of synthesis used in modern synthesizers) and pair them with four different loss functions (two established methods, one utilizing DTW, and one utilizing SIMSE).
 We evaluate the final similarity with two different automatic methods, as well as manual listening tests conducted by two of the authors.
 We compare the distributions of the evaluation scores in order to rank the loss functions from best to worst.
-The results of these experiments show that when accounting for different synthesizers, there is no consistent choice of sound similarity measure for optimal performance in sound-matching.
+
+The results of these experiments show that when accounting for different synthesizers, there is no consistent choice of sound similarity measure for optimal performance.
+Regarding Q1, We see a large overlap between automatic and manual evaluation methods, but continue to encourage manual listening tests as the gold standard for audio evaluation as extreme disagreements can occur.
+In the case of Q2, we confirm that \DTWEnv outperforms other methods when envelope modulation plays a central role, and \SIMSESpec is the best performer in simple subtractive synthesis.
+For Q3, we found the declaration of differentiable sound-matching experiments using Jax and transpiled Faust code to be a convenient and resource efficient, and and overall viable general strategy for sound-matching in differentiable settings.
+These results imply that rather than aiming to outperform the SOTA and finding a universal best performer, more effort could be placed towards diversifying the existing similarity measures of sound as well as conducting experiments using a variety of synthesis methods.
+Sound design is an inherently creative endeavor, and would likely benefit from a diversified range of solutions for different approaches.
 

### 2025-09-08 revised_main.tex
Updates from Overleaf

@@
 As a consequence of these problems, we find that not much is known about the interaction between different methods of synthesis and different loss functions.
-How much performance variation can we expect?
+Is there a universally best performing loss function for audio?
 
@@
 Beyond this central hypothesis, we also investigate three additional research questions: (Q1) To what extent do automatic evaluation metrics agree with manual listening tests?
-(Q2) Can novel loss functions utilizing Dynamic Time Warping (DTW) and Scale-Invariant Mean Squared Error (SIMSE) provide advantages over standard loss functions?
+(Q2) Can novel loss functions utilizing Dynamic Time Warping (DTW) and Scale-Invariant Mean Squared Error (SIMSE) provide advantages over standard loss functions, and if so, when?
 (Q3) Is an iterative differentiable optimization a viable strategy for design of sound-matching experiments?
@@
 
+ We take an \textit{iterative} and \textit{differentiable} approach to defining our sound-matching experiments.
 The iterative approach (commonly used in earlier works) better mimics the manual process of sound design (recursively listening to synthesizer outputs and adjusting the parameters accordingly) while a differentiable environment allows access to the loss function gradients that can be used to better understand the nature of the problem.
@@
 We evaluate the final similarity with two different automatic methods, as well as manual listening tests conducted by two of the authors.
-We compare the distributions of the evaluation scores in order to rank the loss functions from best to worst.
+Using the Non-parametric Scott-Knott tests~\cite{tantithamthavorn2017mvt,tantithamthavorn2018optimization}, compare the distributions of the evaluation scores in order to rank the loss functions from best to worst.
 
@@
 Regarding Q1, We see a large overlap between automatic and manual evaluation methods, but continue to encourage manual listening tests as the gold standard for audio evaluation as extreme disagreements can occur.
-In the case of Q2, we confirm that \DTWEnv outperforms other methods when envelope modulation plays a central role, and \SIMSESpec is the best performer in simple subtractive synthesis.
+In the case of Q2, we confirm that \DTWEnv~outperforms other methods when envelope modulation plays a central role, and \SIMSESpec~is the best performer in simple subtractive synthesis.
 For Q3, we found the declaration of differentiable sound-matching experiments using Jax and transpiled Faust code to be a convenient and resource efficient, and and overall viable general strategy for sound-matching in differentiable settings.

### 2025-09-09 revised_main.tex
Updates from Overleaf

--- 2025-09-05 revised_main.tex
@@
 % need to simplify the intro, and expand the hypothesis
-\section{Introduction}
-
-% Since the 1970s, artists such as Suzanne Ciani~\cite{ciani_life_in_waves} have explored the use of audio synthesizers to replicate natural sounds—such as a bottle cap popping off—in ways that preserve the essential qualities of the original while showcasing the unique characteristics of the synthesizer~\cite{creativecherep2024}.
-Rather than merely replicating sounds, this approach uses synthesizers for artistic reinterpretation.
-However, the automation of this interpretive approach to sound design using digital audio synthesizers remains impractical for sound-designers, and significant knowledge gaps remain even in the simpler problem of replicating sounds.
-One such gap is the lack of diversity and depth in the analysis of sound-similarity measures and their interaction with various synthesis methods.
-As a stepping stone toward automated interpretive sound design, we use automatic and manual sound-evaluations to investigate whether the optimal choice of a sound-similarity measure is dependent on the synthesis method.
-
-A digital audio synthesizer is any software used for the creation and manipulation of audio.
-The possibilities afforded by these synthesizers is infinite, leading to their widespread adaption by artists and sound-designers~\cite{lyons1997understanding,russ1999sound,stranneby2004digital}.
-A typical synthesizer has a number of parameterizable functions which affect the output sound in various ways, and manual sound-design often involves modifying the parameters until a conceptualized target sound is reached.
-The automation of this approach has commonly been referred to as "sound-matching", with reduction of tinkering time and creation of ``interesting" sounds as the main motivators ~\cite{krekovic2019insights,turian2020sorry,horner1993machine,salimi2020make,esling2019flow,engel2020ddsp,mitchell2007evolutionary,shier2020spiegelib,masuda2021soundmatch,masuda2023improving}.
-The main requirements of sound-matching are a target sound, a similarity metric (or loss function), and a heuristic to find parameters for a synthesizer that replicate \textit{all} or \textit{some} of the characteristics of the target sound as best as possible~\cite{horner1993machine,mitchell2007evolutionary,masuda2023improving}.
-However, despite the seeming simplicity of the problem and decades of research, sound-matching is not yet practical solution for sound-designers.
-Here we are motivated to understand why this is the case, and what future research should prioritize in order to improve the field.
-
-We provide a review of past works in order to understand what is lacking, a nomenclature for different subtypes of sound-matching, and illuminate major perennial issues.
-In particular, we are concerned with the issues of \LossSelect~and \SynthSelect.
-The former refers to the search for an optimal loss function, lack of novel algorithms, and the---perhaps needless---focus on outperforming state of the art (\gls{SOTA}), while the latter refers to the lack of diversity in synthesis methods.
-As a consequence of these problems, we find that not much is known about the interaction between different methods of synthesis and different loss functions.
-How much performance variation can we expect?
-
-The main hypothesis here is that \textit{the performance of a similarity measure (or loss function) is influenced by other factors in the environment, particularly the method of synthesis}.
-While there have been many works and experiments comparing the accuracy of loss functions~\cite{vahidi2023mesostructures,turian2020sorry,engel2020ddsp,uzrad2024diffmoog,han2023perceptual,masuda2021soundmatch,turian2020sorry,bruford2024synthesizer}, we note that claims regarding the effectiveness of one function versus another have consistently been made in limited contexts that may not generalize to other settings.
-Beyond this central hypothesis, we also investigate three additional research questions: (Q1) To what extent do automatic evaluation metrics agree with manual listening tests?
-(Q2) Can novel loss functions utilizing Dynamic Time Warping (DTW) and Scale-Invariant Mean Squared Error (SIMSE) provide advantages over standard loss functions?
-(Q3) Is an iterative differentiable optimization a viable strategy for design of sound-matching experiments?
-
-% This main hypothesis (along with our forthcoming research questions) would help with important issues regarding the path of future research.
-Should we be searching for a universally best performing loss function across synthesizers?
-Do current, SOTA loss functions satisfy the creative needs of sound-designers?
-
- The iterative approach (commonly used in earlier works) better mimics the manual process of sound design (recursively listening to synthesizer outputs and adjusting the parameters accordingly) while a differentiable environment allows access to the loss function gradients that can be used to better understand the nature of the problem.
-We define four differentiable synthesizers (each showcasing a fundamental method of synthesis used in modern synthesizers) and pair them with four different loss functions (two established methods, one utilizing DTW, and one utilizing SIMSE).
-We evaluate the final similarity with two different automatic methods, as well as manual listening tests conducted by two of the authors.
-We compare the distributions of the evaluation scores in order to rank the loss functions from best to worst.
-
-The results of these experiments show that when accounting for different synthesizers, there is no consistent choice of sound similarity measure for optimal performance.
-Regarding Q1, We see a large overlap between automatic and manual evaluation methods, but continue to encourage manual listening tests as the gold standard for audio evaluation as extreme disagreements can occur.
-In the case of Q2, we confirm that \DTWEnv outperforms other methods when envelope modulation plays a central role, and \SIMSESpec is the best performer in simple subtractive synthesis.
-For Q3, we found the declaration of differentiable sound-matching experiments using Jax and transpiled Faust code to be a convenient and resource efficient, and and overall viable general strategy for sound-matching in differentiable settings.
-These results imply that rather than aiming to outperform the SOTA and finding a universal best performer, more effort could be placed towards diversifying the existing similarity measures of sound as well as conducting experiments using a variety of synthesis methods.
-Sound design is an inherently creative endeavor, and would likely benefit from a diversified range of solutions for different approaches.
-
-% Here, several common characteristics of previous work are avoided in order to gain fundamental, practical takeaways from the experiments.
-In order to directly analyze the interaction between the loss and method of synthesis, the synthesizer here are implemented with simple, functions that are used as the building blocks of complex synthesizers, and the gradients flow from the loss function to the synthesizer parameters with no extra layers between the two.
-Furthermore, since the true target parameters are unknown in any realistic sound-design scenario, parameter loss (P-Loss) is only used as a secondary measure of performance, and not used as a loss function.
-
-% The simple synthesizers tested here represent fundimental building blocks of more complex implementations, and direct use of loss function gradients assuming 
-%   A primary conclusion of the work is that there is no one best sound similarity measure capable of working across all synthesizers.
-% The core problem that no single loss function performs optimally across synthesizer types
-
-\section{Background and Related Works}
-\subsection{Digital Audio Synthesis}
-\label{sec:dsp}
-A digital audio synthesizer is any software used for the creation and manipulation of digital audio.
-Digital synthesizers use signal processing chains with a variety of digital signal processing (\gls{DSP}) functions to create sounds.
-These functions are often parametric, and the set of parameters given to the synthesizer's chain of DSP functions is called a synthesizer program.
-
-The simplest form of DSP function is a sinusoidal tone. For example, consider:
-\[ x(n) = sin( 2\pi n T)\]
-
-where $T$ represents time in seconds, and $n$ is the frequency parameter.
-If $n$ is 1, then $x(n)$ would be a waveform with a frequency of 1 hertz, meaning that the value of $x(n)$ oscillates to the same point once every second.
-Waveform generators are called oscillators for this reason.
-
-Since the advent of digital signal processing in the 1960s~\cite{stranneby2004digital}, a wide variety of parametric audio generation functions have been proposed~\cite{lyons1997understanding,russ1999sound,shier2020spiegelib}.
-This includes oscillators, filters, equalizers, and envelopes, which can be used sequentially or in parallel~\cite{lyons1997understanding,russ1999sound}.
-Sound design with a synthesizer is done by modifying the parameters of the functions until reaching a desired output~\cite{roads1996computer,pinch2004analog}.
-
-Particularly in earlier works, frequency/amplitude modulators (\gls{FM}/\gls{AM}) has been the synthesizer of choice in sound-matching~\cite{horner1993machine,mitchell2007evolutionary,vahidi2023mesostructures}.
-This synthesis method is simple to implement, yet very expressive~\cite{chowning1973synthesis}.
-A common implementation is the use of one carrier, one modulator, and the corresponding envelopes:
-\[ f(t) = I_c(t) \cos(\omega_c t + I_m(t) \cos(\omega_m t))
-\]
-Where $f(t)$ is the output signal, $I_c(t)$ is the carrier amplitude, $\omega_c$ is the carrier frequency, and $I_m(t)$ and $\omega_m$ are the envelope and frequency for the modulator~\cite{justice1979analytic}.
-
-Compared to the number of possible approaches to sound synthesis, a relatively small subset of synthesis methods have been analyzed in \textit{isolation} for sound-matching.
-Notable methods here are additive and/or subtractive synthesizers~\cite{engel2020ddsp,masuda2023improving,salimi2020make} and physical modeling~\cite{riionheimo2003parameter,han2024learning}.
-
-\highlight{
-What we mean here by \textit{isolation} is that the effect of the synthesizer's parameters (how they change the output sound and experiment outcome) is easily tractable.
-Recent computational advancements have allowed the use of complex Virtual Studio Technology (VST)\cite{steinberg1996vst} synthesizers in sound-matching~\cite{yee2018automatic,esling2019flow}, however, the trade-off here is the obfuscation of the interactions between the synthesizer functions, the loss, and the final outcome, among other distinctions which we will discuss in this section.
-}
-
-An important recent development in digital audio synthesis is the increase of interest in differentiable DSP (\gls{DDSP})~\cite{engel2020ddsp}.
-DDSP is a general term for applications that combine machine learning techniques such as gradient descent~\cite{goodfellow2016deep,boyd2004convex} with DSP.
-Implementing complex DSP functions in a differentiable manner can be challenging, and effective differentiable audio similarity measures require careful mathematical expression of the desired attributes of sound.
-These issues have likely contributed to the limited exploration of this domain~\cite{masuda2021soundmatch,vahidi2023mesostructures,uzrad2024diffmoog}.
-
-\subsection{Sound Representation and Loss Functions}
-\label{sec:loss_funcs}
-A digital sound (or an audio signal) is a series of numbers~\cite{smith1991viewpoints,smith2007mathematics}.
-To compare two digital sounds, the two corresponding series are passed to a function that measures their similarity.
-Two signals can sound identical to our ears, without having any values in common~\cite{moore2012introduction}.
-This necessitates the use of proxy representations (or feature extractors) when comparing sounds automatically.
-Similarity between the target sound and the synthesizer output is then measured by some form of subtraction and summation of the proxy representations.
-
-In sound-matching, particularly in a Deep Learning (\gls{DL}) context~\cite{goodfellow2016deep}, the similarity function can also be called a \textit{loss} function, where the emphasis is on measurement and reduction of the distance between target and output.
-It is important to note that there is a close relationship between the loss function $L$ and the sound representation function $\phi$.
-$L$ is the result of a distance measure $d$ applied to the features extracted by $\phi$.
-
-\[
-L(\theta, t) = d\langle\phi(t),\phi(x)\rangle
-\]
-
-\noindent
-
-A proxy representation is the output of the function \( \phi \), which can be thought of as a feature extraction function that maps the sounds \( t \) and \( x \) to their respective representations. 
-The proportionality or distance metric $d$ has typically been the L1 distance~\cite{turian2020sorry,richard2025model}, calculated as the mean of the absolute difference between every point in the proxy representation~\cite{engel2020ddsp,vahidi2023mesostructures}:
-\[
-L(\theta, t) = \left\| \phi(t) - \phi(x) \right\|_1
-\]
-
-Here we discuss four methods of audio representation and the corresponding loss functions.
-Three of which have been used in previous works, and a novel method not previously used in iterative sound-matching.
-
-\subsubsection{Parameter Loss}
-A common measure of similarity in sound-matching is the distance between synthesizer parameter sets, referred to as ``P-Loss"~\cite{han2023perceptual}.
-Typically, for the implementation of P-Loss the parameter sets are treated as vectors in space, and L1 or L2 distance is applied.
-There are two major limitations to this approach: First, the target and output sound must be made by the same synthesizer; otherwise the parameter sets cannot be compared (see Section~\ref{sec:matching_types}).
-Second, the relationship between synthesizer parameters and the audio output is not linear~\cite{shier2020spiegelib,han2023perceptual,esling2019flow}.
-
-\subsubsection{Fourier Spectrograms}
-\label{sec:fourier_specs}
-Fourier-based transformations such as short-time Fourier transforms (\gls{STFT}), Mel-spectrograms, and Mel-frequency cepstral coefficients have been viewed as the de facto and state-of-the-art representation of audio~\cite{beauchamp2003error,mitchell2007evolutionary,yee2018automatic}, however, there are many issues associated with their use in sound-matching~\cite{turian2020sorry,vahidi2023mesostructures,han2023perceptual,uzrad2024diffmoog}.
-Fourier transformations allow for the conversion of a signal from the time-domain to the frequency domain.
-Audio spectrograms can be generated by segmentation of a piece of audio into overlapping windows followed by the application of Fourier transforms to each window.
-They are costly to compute, but provide a better temporal view of changes in frequency content~\cite{muller2007dynamic,smith2007mathematics}.
-There are different types of spectrograms that have a basis in Fourier transformations, but the most notable and commonly used is the STFT.
-
-What we call \textit{Fourier-based Spectrograms} are variations on the STFT approach.
-For example, Mel-Spectrograms \textit{bin} frequencies on a near-logarithmic scale to better match human perception of frequencies~\cite{muller2007dynamic}.
-Multi-scale spectrograms (\gls{MSS}) used in recent works are a simple weighted average of multiple spectrograms with different parameters such as window size, number of frequency bins, and hop size~\cite{engel2020ddsp,vahidi2023mesostructures}; this may provide some improvements at a higher computational cost~\cite{turian2020sorry,engel2020ddsp}.
-
-\subsubsection{Joint-Time Frequency Spectrum}
-Recent works have focused on the limitations of parameter and spectral loss functions in sound-matching~\cite{vahidi2023mesostructures,uzrad2024diffmoog}, seeking to create more effective general solutions for the comparison of audio.
-Noting the aforementioned weaknesses of comparing STFT spectrograms, Vahidi \textit{et al.} proposed differentiable Joint-Time Frequency Scattering (\gls{JTFS})~\cite{anden2015joint} as an alternative to spectrogram loss in sound-matching, and showed improved performance in sound-matching with differentiable chirplet synthesizers~\cite{vahidi2023mesostructures}.
-JTFS is the result of the application of a 2D wavelet transformation to the time-frequency representation of a signal~\cite{anden2015joint}.
-
-\subsubsection{Dynamic Envelope Warping}
-Dynamic Time Warping (DTW) is a method for measuring similarity between multi-dimensional time-series~\cite{rabiner1993fundamentals,muller2007dynamic,giorgino2009computing}.
-Given any two time-series $X = \{x_1,x_2,...,x_m\}$ and $Y = \{y_1,y_2,...,y_n\}$, we have indices $i\in\{1...m\}$ and $j\in\{1...n\}$ defining $X$ and $Y$.
-When the series are \textit{warped}, these indices change to expand or contract different portions of the series.
-To borrow the notation given by Muller~\cite{muller2007dynamic}, warped indices are a sequence $p=(p_1,...,p_L)$, where \(p_\ell = (m_\ell, n_\ell) \in [1 : m] \times [1 : n] \text{ for } \ell \in [1 : L]\), meaning that the indices for $X$ and $Y$ are reorganized under special conditions.
-In classical DTW, these conditions are \textit{monotonicity}, \textit{boundary matching}, and \textit{single step-size}.
-
-DTW measures the distance between the time-series \textit{after} alignments, typically using Euclidean distance.
-The distance between a time-series and shifted versions of itself would be 0, regardless of shift amount~\cite{tavenard.blog.dtw}.
-Additional rules can be imposed to keep alignments locally constrained~\cite{itakura1975minimum,sakoe1978dynamic}.
-% \textcolor{red}{DTW was originally used for signal comparison but using it the way we do is novel}
-
-\subsection{Automatic Sound Matching}
-\label{sec:sound_matching_definition}
-The problem of sound-matching has been tackled from a variety of perspectives as it involves a number of modular parts: the choice of synthesizer $g$, the target sounds of interest, the representation function $\phi$, and the heuristic for finding the optimal $\theta$.
-To give a formal definition of sound-matching, we expand the definition given by recent works~\cite{vahidi2023mesostructures,han2023perceptual}.
-The major components are as follows:
-\begin{itemize}
-    \item $g(\theta)$: Parametric audio synthesizer $g$ which takes on parameters $\theta$ 
-    \item $x$: The output of $g$, given a set of parameters $\theta$ or $g(\theta) = x$ 
-    \item $t$: The target sound which we want to replicate or imitate. 
-    \item $\phi$: Representation function or feature extractor. $\phi$ is applied to $x$ and $t$ to facilitate their comparison by the loss function.
-    \item $L$: Loss or error function. $L(\theta,t)$ is a measure of distance between $x$ and $t$. Often proportional to the subtraction of their representations, or $ \phi(x) - \phi(t)$
-\end{itemize}
-
-\highlight{It could be argued that the implementation of these components is not an optimization problem with a correct answer, but a \textit{creative} endeavor depending on the artist's needs}.
-Consider the hypothetical case where the goal might be to create an 8-bit~\cite{collins2007loop} version of a high quality, organic snare drum: in such a case, there is no optimal or ``correct" version of an 8-bit snare sound.
-
-\subsection{In-Domain versus Out-of-Domain}
-\label{sec:in-domain}
-The choice of domain depends on whether we want to use the same synthesizer for the target and output sounds, the scenario that is called \textit{in-domain}, or have target sounds that came from sources other than the synthesizer, or \textit{out-of-domain}.
-To paraphrase the description given by Masuda \textit{et al.}~\cite{masuda2021soundmatch}, if $g$, the synthesizer of choice, can accurately replicate the target sound $t$, or put differently, if $t$ itself is an output of $g$, then the sound-matching task is \textit{in-domain}.
-If $t$ is not an output of $g$, then the sound-matching task is \textit{out-of-domain}.
-In-domain tasks in general are simpler, and often there is a guarantee that there is a correct answer to the sound-matching problem, particularly if the goal is accurate replication of the sound.
-If the target sound is out-of-domain, replication is not guaranteed, and the goal becomes the \textit{imitation} of some aspect of sound.
-
-Regardless of the domain, the generation goal can be \textit{replication} or \textit{imitation} of the target sound.
-In replication, the goal is to make an identical copy of the target sound.
-Imitation is an artistic pursuit and harder to define, since the goal is to make new sounds that only retain a subset of the target's sonic features.
-While closely related to the in-domain versus out-of-domain problem, the choice of replication versus imitation is more dependent on how the loss and representation functions are defined.
-
-\subsection{Supvervised vs Direct Optimization}
-\label{sec:optimization}
-
-The choice of \textit{heuristics} is yet another important attribute sound-matching.
-The goal of sound-matching is to find the optimal parameters $\theta^*$ that minimize the loss between the synthesizer output and the target sound.
-\[
-\theta^* = \arg\min_{\theta} L(\theta,t)
-\]
-
-The heuristics (i.e., how $\theta^*$ is approximated) used in past works can be broadly split into the two categories of \textit{direct optimization} and \textit{supervised}  (or inference) methods.
-Direct optimization refers to the iterative generation of a sound output, measurement of similarity between target and output, and application of updates to the parameters to maximize similarity (or minimize loss)~\cite{horner1993machine,mitchell2007evolutionary,yee2018automatic,vahidi2023mesostructures}; while supervised methods use large datasets of synthesizer sounds and corresponding parameters to learn the generation objective, commonly with the use of DNNs~\cite{engel2020ddsp,salimi2020make,yee2018automatic,esling2019flow}.
-These models often make their parameter predictions in a single step (or 1-shot).
-
-Genetic algorithms (\gls{GA})~\cite{holland1992genetic} have been the earliest and most common heuristic for direct sound-matching~\cite{horner1993machine,mitchell2007evolutionary,yee2018automatic}.
-These algorithms start with arbitrary parameter sets that can be treated as an evolving population where the genomes are the parameter values.
-The most fit members of the group are the parameters which perform the best in the loss function, and create a new generation of parameters via mutation (random change in subset of parameters) and crossovers (combination of parameter sets); this process repeats until the goal or a maximum number of generations is reached.
-Rather than using random mutations, differentiable approaches allow goal-oriented updates to the synthesizer parameters (the goal being the minimization of loss), but with some drawbacks: other than requiring more computation power, differentiable functions require careful implementation of operations that are continuous and numerically stable; this makes the implementation of signal processing functions quite difficult, contributing to the scarcity of works in this domain.
-
-\subsection{Historical Framing of Sound-Matching}
-Given the different attributes discussed thus far, Table~\ref{tab:summary} contains an overview of important relevant literature.
-This leads us to a more detailed historical analysis of past works in sound-matching.
-
-Perhaps the earliest foundational work in sound-matching is the analytical approach by Justice~\cite{justice1979analytic} toward the decomposition and recreation of sounds using the simple FM synthesizer described in Section~\ref{sec:dsp}.
-The sound-matching works that followed used other heuristics for finding the synthesizer parameters, yet the simple FM synthesizer structure remained unchanged.
-Horner \textit{et al.}~\cite{horner1993machine} used GAs for re-synthesis of sounds with an FM synthesizer using one modulator and one carrier oscillator, and the McAulay-Quatieri method~\cite{mcaulay1986speech} for measuring loss.
-
-Introduction of more complex models of synthesis such as wavetables~\cite{horner2003auto} and physical modeling~\cite{riionheimo2003parameter} rendered interesting results, but led to further questions about the nature and goals of sound-matching.
-Mitchell and Creasy noted the difficulty in distinguishing between ``inefficiency of the optimization engine'' and ``synthesizer limitations'' as the cause of failure or success in sound-matching~\cite{mitchell2007evolutionary}.
-They offered a \textit{contrived methodology} for finding the best evolutionary method for sound-matching using their FM synthesizer.
-The methodology postulates that the best search heuristic for in-domain search (where an exact target exists, and a wide range of targets can be produced by sampling from different points in the parameter space) would also be the best-performing heuristic for out-of-domain search on a dataset of muted trumpet tones~\cite{opolko1989mcgill}; however, testing was not extensive and yielded some contradictory results.
-As they noted, modifications to the synthesizer, loss function, and sound domain would lead to a problem with an entirely new search space~\cite{mitchell2007evolutionary}.
-
-Recent years have seen more works in sound-matching using supervised machine learning techniques.
-In 2018, Yee-King \textit{et al.} rendered 60,000 audio-parameter pairs from the \textit{Dexed} \gls{VST} synthesizer~\footnote{https://asb2m10.github.io/dexed/}, and showed that NNs trained on this dataset can outperform GA and hill-climber (\gls{HC})~\cite{hoffmann2000heuristic} methods in rendering speed, with slight improvements in MFCC error (used as an objective performance test).
-The speed improvement appears trivial, considering the iterative nature of GAs when compared to offline training of supervised models.
-For GAs and HC optimizer, MFCCs were used as a measure of performance as well as a loss function.
-For training the networks, P-Loss was used, since differentiable MFCCs were not possible in their pipeline.
-Importantly, informal hearing tests revealed that the performance of even the best NN model was unsatisfactory, possibly due to the complex nature of the synthesizer, which features 155 parameters.
-
-Masuda \textit{et al.} also applied supervised learning to synthesizer parameter estimation~\cite{masuda2021soundmatch}.
-Their work highlights the issue of non-linearity in parameter-to-synthesizer outputs and out-of-domain search.
-This work uses a differentiable subtractive synthesizer (two oscillators and an LP filter).
-A NN model was pre-trained using an in-domain dataset of randomly selected parameters and P-Loss.
-After training, the model was fine-tuned using 20,000 out-of-domain sounds from the NSynth dataset~\cite{engel2017neural} and multi-scale spectrogram loss~\cite{engel2020ddsp}.
-This approach proved more effective---i.e, achieved lower multi-level spectral difference in out-of-domain tests---than baseline models, which were either not exposed to out-of-domain sounds or trained exclusively with P-Loss.
-Subjective hearing tests were conducted, showing a preference for the fine-tuned model~\cite{masuda2021soundmatch}.
-Masuda \textit{et al.} later extended this work with semi-supervised learning, highlighting significant gaps in in-domain and out-of-domain performance~\cite{masuda2023improving}.
-
-Rather than focusing on a particular implementation, Shier \textit{et al.} presented Spieglib, a library for implementation of sound-matching pipelines~\cite{shier2020spiegelib}.
-This library provides different choices for DNNs, GAs, synthesizers, and feature extractors.
-Shier \textit{et al.} presented an experiment with a similar setup to Yee-King \textit{et al.}~\cite{yee2018automatic}, however, they found a genetic algorithm to be the best performing.
-
-Differentiable loss functions that use spectrogram differences can be computationally expensive.
-To mitigate this, Han \textit{et al.}~\cite{han2023perceptual} introduced ``perceptual-neural-physical loss'' (PNP).
-PNP is an approximation of loss functions; specifically, a loss function that uses the L2 norm of the difference between the features of two sounds, or $||\phi(t) - \phi(x)||^2_2$, where $\phi$ could be a spectrogram or JTFS function.
-PNP loss functions are fast and differentiable, but require training and parameter estimation.
-A Riemannian metric M needs to be calculated for the minimization of locally linear approximation of the ``true" spectral loss function~\cite{han2023perceptual}.
-\[
-\|\phi(t) - \phi(x)\|_2^2 = \langle \tilde{\theta} - \theta | \mathbf{M}(\theta) | \tilde{\theta} - \theta \rangle + O(\|\tilde{\theta} - \theta\|_2^3). \tag{4}
-\]
-
-During the training phase, this metric is calculated alongside the neural parameter estimator.
-After training, PNP can be used as a fast approximation of computationally demanding loss functions such as JTFS~\cite{han2023perceptual,han2024learning}, paired with FM synthesizers and a differentiable physical model of a drum head~\cite{smith2010physical}.
-
-A neural approximation for another part of the supervised sound-matching chain---this time the synthesizer---was proposed by Barkan \textit{et al.}~\cite{barkan2023inversynthII}.
-As they noted, without a differentiable synthesizer, ``model-based'' (or what we call supervised) approaches cannot directly compare $x_o$ to $x_t$, often opting for P-Loss, which may not correctly map the parameters of sound to the output audio~\cite{esling2019flow,han2023perceptual,masuda2023improving}.
-Given a synthesizer and a large dataset of sounds and corresponding parameters, a model which maps sounds to parameters, and another model which approximates the synthesizer can be trained, allowing the implementation of a loss function which combines P-Loss with STFT differences.
-This ``Inversynth II'' (IS2) approach yielded significant improvements to previous works which did not use the STFT approximations for loss~\cite{esling2019flow,barkan2019inversynth}.
-
-Barkan \textit{et al.} attemped to  improve the IS2 model with Inference-Time
-Finetuning (ITF).
-For ITF, the synthesizer approximation is frozen and the encoder is iteratively finetuned for the sample sample using gradient descent in order to minimize P-Loss, stopping at convergence or when maximum number of iterations has reached~\cite{barkan2023inversynthII}.
-Barkan \textit{et al.} report that perhaps due to the discrepancy between the true and approximated synthesizer or overfitting, the application of ITF to IS2 generally worsened the spectral and manual performance metrics.
-
-Uzrad \textit{et al.}~\cite{uzrad2024diffmoog} took another unique approach to sound-matching: using a differentiable \textit{synthesis chain} of DSP generators and effects and a loss function that combines P-Loss with a \textit{signal-chain loss}.
-The synthesizer is a customizable chain of effects, which feed one output as input to the next step of the chain; signal-chain loss compares the parameter and output difference at every output step in the chain~\cite{uzrad2024diffmoog}.
-Possible chain functionalities are FM/AM, Low-Frequency Oscillators (\gls{LFO}), filters, and envelopes.
-Like the results shown by Masuda \textit{et al.}~\cite{masuda2021soundmatch}, better out-of-domain results were achieved when pre-trained on in-domain data and fine-tuned using out-of-domain NSynth data~\cite{engel2017neural}.
-
-Some recent works have opted for the use of audio embeddings as a similarity metric.
-In a recent work, Cherep \textit{et al.} used latent representations from the CLAP model~\cite{wu2023large} along with a differentiable synthesizer~\cite{synthhaxcherep2023} to create creative interpretations of sound effects.
-In their approach, a desired sound-effect is described in text and embedded using clap, and gradient free optimizer~\cite{evosax2022github} iteratively updates the synthesizer's parameters to minimize difference in the text and output sound embeddings.
-Based on manual hearing tests, this approach did not produce the ``correct'' sounds more frequently than previous works~\cite{kreuk2022audiogen}, however, it did yield better scores for ``artistic-interpretation'' (or what we refer to as imitation).
-
-% masuda and esling both noted out-of-domain drop in performance
-% yee king noted performance drop in manual surveys 
-%p-loss bad : masuda-improving, han2023, esling-flowsynth
-% manual performance measure: barkan, masuda2021soundmatch, yee-king
+% \input{sections/introduction.tex}
+\input{sections/background_related}
+\input{sections/setup_and_posthoc_analysis}
 

### 2025-09-09 sections/background_related.tex
Updates from Overleaf

--- 2025-08-21 main.tex
@@
-\documentclass[lettersize,journal]{IEEEtran}
-\usepackage{amsmath,amsfonts}
-\usepackage{algorithmic}
-\usepackage{algorithm}
-\usepackage{array}
-\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-\usepackage{textcomp}
-\usepackage{stfloats}
-\usepackage{url}
-\usepackage{verbatim}
-\usepackage{graphicx}
-\usepackage{xcolor}
-\usepackage{csquotes}
-\usepackage{makecell}
-% my added packages and commands =======================
-\usepackage[numbers]{natbib} %removed cite package
-\usepackage{tikz}
-\usetikzlibrary{positioning,shapes.geometric}
-\usepackage{adjustbox}
-\usepackage{listings}
-\usepackage{caption}
+\section{Background and Related Works}
 
-\newcommand{\highlight}[1]{\textcolor[RGB]{00,100,100}{#1}}
-\newcommand{\todo}[1]{\textcolor{red}{#1}}
-
-\captionsetup[lstlisting]{justification=centering, singlelinecheck=false}
-\providecommand{\gls}[1]{#1}
-\definecolor{rank1}{HTML}{70FF70}
-\definecolor{rank2}{HTML}{858585}
-\definecolor{rank3}{HTML}{454545}
-\definecolor{rank4}{HTML}{000000}
-\newcommand{\SIMSESpec}{\texttt{SIMSE\_Spec}}
-\newcommand{\LoneSpec}{\texttt{L1\_Spec}}
-\newcommand{\JTFS}{\texttt{JTFS}}
-\newcommand{\DTWEnv}{\texttt{DTW\_Envelope}}
-\newcommand{\LossSelect}{\textbf{Loss Selection}}
-\newcommand{\SynthSelect}{\textbf{Synthesis Selection}}
-\newcommand{\PeriodicLoss}{\textbf{Periodic Loss}}
-\newcommand{\OutDomain}{\textbf{Out-Domain Generation}}
-\newcommand{\BPNoise}{\textbf{BP-Noise}}  
-\newcommand{\AddSineSaw}{\textbf{Add-SineSaw}}  
-\newcommand{\AmpMod}{\textbf{Noise-AM}}  
-\newcommand{\FMMod}{\textbf{SineSaw-AM}}
-
-\input{code_style}
-
-%  ================================
-\begin{document}
-\title{Evaluating Sound Similarity Metrics for Differentiable, Iterative Sound-Matching}
-
-\author{Amir Salimi, Abram Hindle, Osmar R. Za{\"i}ane}
-
-% \markboth{IEEE TRANSACTIONS ON AUDIO, SPEECH AND LANGUAGE PROCESSING, VOL -, 2025}%
-% {Shell \MakeLowercase{\textit{et al.}}: A Sample Article Using IEEEtran.cls for IEEE Journals}
-
-\markboth{%
-  \parbox{\textwidth}{%
-    This work has been submitted to the IEEE for possible publication. \\
-    Copyright may be transferred without notice, after which this version may no longer be accessible.%
-  }%
-}{}
-
-\maketitle
-
-\begin{abstract}
-Manual sound design with a synthesizer is inherently iterative: an artist compares the synthesized output to a mental target, adjusts parameters, and repeats until satisfied.
-Iterative sound-matching automates this workflow by continually programming a synthesizer under the guidance of a loss function (or similarity measure) towards a target sound.
-Prior comparisons of loss functions have typically favored one metric over another, but only within narrow settings: limited synthesis methods, few loss types, often without blind listening tests.
-This leaves open the question of whether a universally optimal loss exists, or the choice of loss remains a creative decision conditioned on the synthesis method and the sound designer's preference.
-We propose differentiable iterative sound-matching as the natural extension of the available literature, since it combines the manual approach to sound design with modern advances in machine learning.
-To analyze the variability of loss function performance across synthesizers, we implemented a mix of four novel and established differentiable loss functions, and paired them with differentiable subtractive, additive, and AM synthesizers.
-For each of the sixteen synthesizer–loss combinations, we ran 300 randomized sound-matching trials.
-Performance was measured using parameter differences, spectrogram-distance metrics, and manually assigned listening scores.
-We observed a moderate level of consistency among the three performance measures.
-Our post-hoc analysis shows that the loss function performance is highly dependent on the synthesizer.
-These findings underscore the value of expanding the scope of sound-matching experiments, and developing new similarity metrics tailored to specific synthesis techniques, rather than pursuing one-size-fits-all solutions.
-\end{abstract}
-
-% the automatic performance measures have not been compared to human hearing
-% lack of creativity in loss function definitions 
-\begin{IEEEkeywords}
-Audio synthesis, differentiable digital signal processing, music information retrieval, sound-matching
-\end{IEEEkeywords}
-
-\section{Introduction}
-Since the 1970s, artists such as Suzanne Ciani~\cite{ciani_life_in_waves} have explored the use of audio synthesizers to replicate natural sounds—such as a bottle cap popping off—in ways that preserve the essential qualities of the original while showcasing the unique characteristics of the synthesizer.
-Rather than merely replicating sounds, this approach uses synthesizers for artistic reinterpretation.
-However, the automation of this interpretive approach to sound design using digital audio synthesizers has rarely been explored, and significant knowledge gaps remain even in the simpler problem of replicating sounds.
-One such gap is the lack of diversity and depth in the analysis of sound-similarity measures and their interaction with various synthesis methods.
-As a stepping stone toward automated interpretive sound design, we use automatic and manual sound-evaluations to investigate whether the optimal choice of a sound-similarity measure is dependent on the synthesis method.
-
-A digital audio synthesizer is any software used for the creation and manipulation of audio.
-Since the advent of digital signal processing (\gls{DSP}) in the 1960s~\cite{stranneby2004digital}, a wide variety of parametric audio generation functions have been proposed~\cite{lyons1997understanding,russ1999sound,shier2020spiegelib}.
-Manual sound design with a synthesizer is an iterative process of trial and error; when the artist has a desired (or target) sound in mind, sound design involves listening to the output, determining whether the output is closer or farther from the target, and adjusting the parameters of the synthesizer accordingly until the target is reached~\cite{russ1999sound}.
-This paper explores the automation of this process, which we call \textit{iterative sound-matching}.
-
-Automatic sound design has been a subject of research for decades, with reduction of tinkering time and creation of ``interesting" sounds as the main motivators \cite{krekovic2019insights,turian2020sorry,horner1993machine,salimi2020make,esling2019flow,engel2020ddsp,mitchell2007evolutionary,shier2020spiegelib,masuda2021soundmatch,masuda2023improving}.
-\textit{Sound-matching} is a common form of automatic sound design explored in literature, where a similarity metric (or loss function) is given, and the goal is to find parameters for a synthesizer that replicate \textit{all} or \textit{some} of the characteristics of a target sound as best as possible~\cite{horner1993machine,mitchell2007evolutionary,masuda2023improving}.
-Our analysis of past works shows that this general goal of ``sound-matching'' can be broken down in to very different subtypes, many of them rarely or never explored, despite their relevance to practical applications.
-This leads us to a typology of sound-matching tasks, and the importance of \textit{differentiable, iterative sound-matching}, rarely explored in previous literature~\cite{vahidi2023mesostructures}.
-
-\textit{Iterative sound-matching} refers to works that mimic the back-and-forth process of listening to a synthesizer's output and updating the parameters until the termination conditions are met.
-The two necessary components are the \textit{loss function}, which determines how close two sounds are, and the \textit{parametric synthesizer} that outputs sounds based on input parameters.
-In the iterative workflow, a target sound is selected, the synthesizer outputs sounds based on its parameters, and the loss function determines the proximity of the output to the target.
-From there, the synthesizer parameters are updated, the output and target are compared, and so on.
-
-Despite the early popularity of the iterative approach~\cite{justice1979analytic,horner1993machine,mitchell2007evolutionary}, recent works in sound-matching have favored supervised learning using inference models, deep neural-networks (\gls{DNNs}), and large datasets.
-Rather than continually interacting with the environment, these models learn from a dataset of inputs (synthesizer parameters) and outputs (corresponding sound), and predict the synthesizer parameters for new sounds~\cite{yee2018automatic,han2023perceptual,han2024learning,masuda2021soundmatch,masuda2023improving}.
-
-Iterative sound-matching is interesting for multiple reasons.
-First, it better resembles the human approach to sound-design, which requires continually interacting with the environment and making decisions according to a feedback signal~\cite{sutton2018reinforcement}.
-Second, the trial-and-error approach leaves a trail of parameter sets from the starting point to the end goal, which could be of use to sound designers.
-Third, relative to the DNN-based inference models, it provides a more tractable environment with room for creative approaches to sound-similarity definitions, and more amenable to deployment on different synthesizers as it does not require pre-training on large datasets.
-
-This work takes a \textit{differentiable} approach to iterative sound-matching, a natural extension of early works in iterative sound-matching combined with recent advances in Machine Learning (\gls{ML}) software~\cite{braun2024dac,paszke2019pytorch} and increase in computational power~\cite{moore2012introduction,schaller1997moore}.
-A differentiable approach requires differentiable loss and synthesizer functions, allowing the synthesizer parameters to be updated using gradient descent~\cite{goodfellow2016deep,boyd2004convex}.
-
-We discuss the history of sound-matching, and provide a nomenclature for different subtypes of sound-matching tasks.
-A number of weaknesses in current literature (see Section~\ref{sec:lacking}) are identified.
-The experiments in this work focus on two of these core issues, referred to as \LossSelect~and \SynthSelect.
-The former refers to the search for an optimal loss function, lack of novel algorithms, and the---perhaps needless---focus on outperforming state of the art (\gls{SOTA}), while the latter refers to the lack of diversity in synthesis methods.
-
-In the case of \LossSelect, there have been many works and experiments comparing the accuracy of loss functions and choosing the best performers~\cite{vahidi2023mesostructures,turian2020sorry,engel2020ddsp,uzrad2024diffmoog,han2023perceptual,masuda2021soundmatch,turian2020sorry,bruford2024synthesizer}.
-We note that these claims regarding the effectiveness of one function versus another have consistently been made in limited contexts that may not generalize to other settings.
-One such limitation is the synthesizers used, and ties into the problem of \SynthSelect, as typically only a single (and often, simple) synthesizer has been used for testing the effectiveness of loss functions.
-Furthermore, there is a lack of diversity in the underlying functions used by the synthesizers, with simple Frequency Modulation (\gls{FM}) being the most common method of synthesis~\cite{horner1993machine,vahidi2023mesostructures,mitchell2007evolutionary,caspe2023envelope}.
-
-Considering the interplay between the two problems, we design experiments that address them simultaneously.
-Testing the hypothesis that the performance of a similarity measure (or loss function) is influenced by other factors in the environment, particularly the method of synthesis.
-We combine 4 loss functions with 4 differentiable synthesizers, and compare their performance as measured by the similarity of the final output to the target.
-We evaluate the final similarity with two different automatic methods, as well as manual listening tests conducted by two of the authors.
-We compare the distributions of the evaluation scores in order to rank the loss functions from best to worst.
-The results of these experiments show that when accounting for different synthesizers, there is no consistent choice of sound similarity measure for optimal performance in sound-matching.
-
-We note that for some synthesizers, our novel use of ``dynamic time warping of envelopes'' as a loss function for differentiable sound-matching outperforms the standard methods of audio comparison.
-These results imply that rather than aiming to outperform the SOTA and finding a universal best performer, more effort could be placed towards diversifying the existing similarity measures of sound as well as conducting experiments using a variety of synthesis methods.
-Sound design is an inherently creative endeavor, and would likely benefit from a diversified range of solutions for different approaches.
-
-\section{Background}
-\label{sec:background}
-In this Section, we give a formal definition of sound-matching and the necessary components.
-These discussions lead to a nomenclature for types of sound-matching, as well as major perennial issues in the field.
+\todo{Sound-matching is an amalgamation of various fields and  requires knowledge of digital signal processing, audio representation, and heuristic search.
+Here we introduce the various fields and ideas behind sound-matching, before we use these ideas in a historical framing of past sound-matching works.}
 
 \subsection{Digital Audio Synthesis}
-\label{sec:diffSynth}
+\label{sec:dsp}
 A digital audio synthesizer is any software used for the creation and manipulation of digital audio.
@@
 
-% If $n$ is in the range of 30 to 20,000 (the upper limit is often lower for some), $x(n)$ could be played as an audible tone~\cite{smith2007mathematics,jilek2014reference}.
-
 Since the advent of digital signal processing in the 1960s~\cite{stranneby2004digital}, a wide variety of parametric audio generation functions have been proposed~\cite{lyons1997understanding,russ1999sound,shier2020spiegelib}.
@@
 
-\todo{A relatively small subset of these synthesis methods have been used for automatic sound-matching, notably additive and/or subtractive synthesizers~\cite{esling2019flow,yee2018automatic,mitchell2007evolutionary,salimi2020make}, physical modeling~\cite{riionheimo2003parameter}, and frequency/amplitude modulation (\gls{FM}/\gls{AM})~\cite{horner1993machine,vahidi2023mesostructures}.
+Particularly in earlier works, frequency/amplitude modulators (\gls{FM}/\gls{AM}) has been the synthesizer of choice in sound-matching~\cite{horner1993machine,mitchell2007evolutionary,vahidi2023mesostructures}.
+This synthesis method is simple to implement, yet very expressive~\cite{chowning1973synthesis}.
+A common implementation is the use of one carrier, one modulator, and the corresponding envelopes:
+\[ f(t) = I_c(t) \cos(\omega_c t + I_m(t) \cos(\omega_m t))
+\]
+Where $f(t)$ is the output signal, $I_c(t)$ is the carrier amplitude, $\omega_c$ is the carrier frequency, and $I_m(t)$ and $\omega_m$ are the envelope and frequency for the modulator~\cite{justice1979analytic}.
+
+Compared to the number of possible approaches to sound synthesis, a relatively small subset of synthesis methods have been analyzed in \textit{isolation} for sound-matching.
+Notable methods here are additive and/or subtractive synthesizers~\cite{engel2020ddsp,masuda2023improving,salimi2020make} and physical modeling~\cite{riionheimo2003parameter,han2024learning}.
+
+\highlight{
+What we mean here by \textit{isolation} is that the effect of the synthesizer's parameters (how they change the output sound and experiment outcome) is easily tractable.
+Recent computational advancements have allowed the use of complex Virtual Studio Technology (VST)\cite{steinberg1996vst} synthesizers in sound-matching~\cite{yee2018automatic,esling2019flow}, however, the trade-off here is the obfuscation of the interactions between the synthesizer functions, the loss, and the final outcome, among other distinctions which we will discuss in this section.
 }
 
-Differential DSP (\gls{DDSP}) is a general term for applications that combine machine learning techniques such as gradient descent~\cite{goodfellow2016deep,boyd2004convex} with DSP~\cite{engel2020ddsp}.
+An important recent development in digital audio synthesis is the increase of interest in differentiable DSP (\gls{DDSP})~\cite{engel2020ddsp}.
+DDSP is a general term for applications that combine machine learning techniques such as gradient descent~\cite{goodfellow2016deep,boyd2004convex} with DSP.
 Implementing complex DSP functions in a differentiable manner can be challenging, and effective differentiable audio similarity measures require careful mathematical expression of the desired attributes of sound.
-These issues have likely contributed to the limited exploration of this domain.
-Perhaps the most relevant recent work is by Vahidi \textit{et al.}, featuring the comparison of two differentiable loss functions with a simple FM synthesizer~\cite{vahidi2023mesostructures}.
+These issues have likely contributed to the limited exploration of this domain~\cite{masuda2021soundmatch,vahidi2023mesostructures,uzrad2024diffmoog}.
 
@@
 A proxy representation is the output of the function \( \phi \), which can be thought of as a feature extraction function that maps the sounds \( t \) and \( x \) to their respective representations. 
-The proportionality or distance metric $d$ has typically been the L1 distance~\cite{turian2020sorry,richard2025model}, calculated as the mean of the absolute difference between every point in the proxy representation~\cite{engel2020ddsp,vahidi2023mesostructures}:
+The proportionality or distance metric $d$ has typically been the L1 or L2 distance~\cite{turian2020sorry,richard2025model}, with L1 being calculated as the mean of the absolute difference between every point in the proxy representation:
 \[
@@
 Here we discuss four methods of audio representation and the corresponding loss functions.
-Three of which have been used in previous works, and a novel method not previously used in iterative sound-matching.
+We also include a technical justification for our novel methods of defining sound similarity using the Scale-Invariant Spectrogram Error (SIMSE) and Dynamic Time Warping (DTW) techniques.
 
@@
 
+\subsubsection{Novelty of Scale-Invariant Spectrogram Error}
+A limitation of L1/L2 distances on spectrograms is their sensitivity to global gain: 
+two sounds with identical spectral shape but different amplitude can yield large losses. 
+Scale-Invariant Mean Squared Error (SIMSE) addresses this by normalizing the energy of the compared signals, thereby focusing on proportional differences in spectral envelopes. 
+This can be advantageous for subtractive synthesis (used by one of our synthesizers), where filter cutoffs reshape the spectral balance which can cause changes in total energy. 
+\todo{Perceptually, human listeners are often invariant to global amplitude changes, attending instead to relative spectral shapes}. 
+Although SIMSE has been used in other contexts (e.g., image/audio reconstruction), its application as a differentiable spectrogram loss in iterative sound-matching is, to our knowledge, novel.
+
 \subsubsection{Joint-Time Frequency Spectrum}
@@
 
+\paragraph*{Justification and Novelty}
 DTW measures the distance between the time-series \textit{after} alignments, typically using Euclidean distance.
@@
 Additional rules can be imposed to keep alignments locally constrained~\cite{itakura1975minimum,sakoe1978dynamic}.
-% \textcolor{red}{DTW was originally used for signal comparison but using it the way we do is novel}
+DTW provides robustness to local temporal shifts, making it well suited for comparing modulated signals where the perceptual similarity lies in envelope dynamics rather than precise onset alignment.
+For example, two tremolo signals with slightly different phase are perceptually similar but would appear distant under spectrogram L1. 
+By applying DTW specifically to amplitude envelopes, we align with perceptual cues of loudness modulation while maintaining differentiability. 
+To our knowledge, DTW applied in this way has not been used as a loss in sound-matching (whether differentiable and iterative or not).
 
@@
 
-\subsection{Sound Matching Types}
-\label{sec:matching_types}
-
-Sound-matching tasks can be categorized based on the choices of target sound $t$, the loss $L$ and representation function $\phi$, and the heuristic used for minimizing $\theta$.
-We will discuss the subtypes of sound-matching tasks in the upcoming sub-sections.
-
-\subsubsection{In-Domain vs Out-of-Domain}
+\subsection{In-Domain versus Out-of-Domain}
 \label{sec:in-domain}
-The choice of domain depends on whether we want to use the same synthesizer for the target and output sounds, which we call \textit{in-domain}, or have target sounds that came from sources other than the synthesizer, or \textit{out-of-domain}.
+The choice of domain depends on whether we want to use the same synthesizer for the target and output sounds, the scenario that is called \textit{in-domain}, or have target sounds that came from sources other than the synthesizer, or \textit{out-of-domain}.
 To paraphrase the description given by Masuda \textit{et al.}~\cite{masuda2021soundmatch}, if $g$, the synthesizer of choice, can accurately replicate the target sound $t$, or put differently, if $t$ itself is an output of $g$, then the sound-matching task is \textit{in-domain}.
@@
 
-\subsubsection{Replication vs Imitation}
 Regardless of the domain, the generation goal can be \textit{replication} or \textit{imitation} of the target sound.
@@
 
-% From a signal perspective, replication is a scenario where the goal is to find a $\theta$ such that $t-x=0$ where $x=g(\theta)$.
-From a sound perspective, this definition of replication is unnecessarily restrictive, as $t$ and $x$ can sound identical but contain very different values.
-For example, if $t$ and $x$ were periodic waveforms at an audible frequency $f$ with different phase values, their subtraction would not be 0, yet they would sound identical to our ears, as our ears are insensitive to phase values~\cite{moore2012introduction}.
-Representation functions such as STFT spectrograms or Mel-frequency cepstral coefficients (\gls{MFCC}) become viable in these cases, as they allow for separation and removal of the phase information within an audio signal.
-The \textit{sound replication} goal can be defined, in layman's terms, as ``$x$ sounds exactly like $t$".
-More formally, the goal is to find parameters such that $\phi(t)-\phi(x)=0$, where $\phi$ transforms an audio into a spectrogram that removes the phase information from sounds, but otherwise represents the sonic information as accurately as possible.
-
-% With \textit{sound imitation}, the goal is the replication of some aspect of the target sound.
-Instrument conversion is a common example of imitation.
-As an example, a note played by a violin is converted to the same note played by a piano.
-Conversion of even longer pieces has been made possible with large enough datasets~\cite{engel2020ddsp}.
-Imitation of sound envelopes (how the loudness changes over time) is another interesting idea~\cite{caspe2023envelope}.
-Transfer of other sound properties such as reverb and delay has also been a subject of research~\cite{engel2020ddsp}.
-
-% A target sound can have many properties of interest, and the imitation task is the transfer of a subset of those properties to other sounds.
-Relative to replication, imitation is a very open ended goal.
-From a research point of view, success in the replication task can be measured more reliably, often without the need for manual evaluation.
-With imitation, listening tests are needed to ensure quality of outputs, particularly in a supervised learning setting where learning is done by examples~\cite{salimi2020make,masuda2021soundmatch}.
-The need for subjective evaluations has perhaps contributed to the lack of research in imitation and out-of-domain sound-matching.
-
-\subsubsection{Supervised Versus Direct Optimization}
+\subsection{Supvervised vs Direct Optimization}
 \label{sec:optimization}
 
+The choice of \textit{heuristics} is yet another important attribute sound-matching.
 The goal of sound-matching is to find the optimal parameters $\theta^*$ that minimize the loss between the synthesizer output and the target sound.
@@
 
-The heuristics in past works can be broadly split into two categories:
-\begin{enumerate}
-    \item Supervised (or inference) methods which use datasets of sound and parameter pairs to describe the generation objective, often with a large number of input and output examples~\cite{engel2020ddsp,salimi2020make,yee2018automatic,esling2019flow}
-    \item Unsupervised or \textit{direct sound optimization} methods~\cite{horner1993machine,mitchell2007evolutionary,yee2018automatic,vahidi2023mesostructures}.
-Usually require at most one target example, and a loss function that measures the difference between the target and synthesizer outputs.
-\end{enumerate}
+The heuristics (i.e., how $\theta^*$ is approximated) used in past works can be broadly split into the two categories of \textit{direct optimization} and \textit{supervised}  (or inference) methods.
+Direct optimization refers to the iterative generation of a sound output, measurement of similarity between target and output, and application of updates to the parameters to maximize similarity (or minimize loss)~\cite{horner1993machine,mitchell2007evolutionary,yee2018automatic,vahidi2023mesostructures}; while supervised methods use large datasets of synthesizer sounds and corresponding parameters to learn the generation objective, commonly with the use of DNNs~\cite{engel2020ddsp,salimi2020make,yee2018automatic,esling2019flow}.
+These models often make their parameter predictions in a single step (or 1-shot).
 
-\todo{
-Genetic algorithms (\gls{GA})~\cite{holland1992genetic} can be used for direct sound optimization, and have been the earliest and most common heuristic in sound-matching~\cite{horner1993machine,mitchell2007evolutionary,yee2018automatic}.
+Genetic algorithms (\gls{GA})~\cite{holland1992genetic} have been the earliest and most common heuristic for direct sound-matching~\cite{horner1993machine,mitchell2007evolutionary,yee2018automatic}.
 These algorithms start with arbitrary parameter sets that can be treated as an evolving population where the genomes are the parameter values.
-The most fit members of the group are the parameters which perform the best in the loss function, and create a new generation of parameters via mutation (random change in subset of parameters) and crossovers (combination of parameter sets); this process repeats until the goal or a maximum number of generations is reached.}
+The most fit members of the group are the parameters which perform the best in the loss function, and create a new generation of parameters via mutation (random change in subset of parameters) and crossovers (combination of parameter sets); this process repeats until the goal or a maximum number of generations is reached.
+Rather than using random mutations, differentiable approaches allow goal-oriented updates to the synthesizer parameters (the goal being the minimization of loss), but with some drawbacks: other than requiring more computation power, differentiable functions require careful implementation of operations that are continuous and numerically stable; this makes the implementation of signal processing functions quite difficult, contributing to the scarcity of works in this domain.
 
- Differentiable settings have the benefit of making goal-oriented updates to the synthesizer parameters (the goal being the minimization of loss).
-There are also drawbacks: other than requiring more computation power, differentiable functions require careful implementation of operations that are continuous and numerically stable; this makes the implementation of signal processing functions quite difficult, contributing to the scarcity of works in this domain.
+\subsection{Historical Framing of Sound-Matching}
+Given the different attributes discussed thus far, Table~\ref{tab:summary} contains an overview of important relevant literature.
+This leads us to a more detailed historical analysis of past works in sound-matching.
 
-% One important aspect of the heuristics is whether they imitate the iterative nature of manual sound design---that is, listening to the output and adjusting the synthesizer parameters recursively until the desired goal is reached.
-Supervised methods generate their final output in a single step, while direct sound optimization methods refine their outputs iteratively until termination, which resembles manual sound design more closely.
-
-\subsection{Iterative Sound Matching}
-\label{sec:iterative_sound_matching}
- \todo{Figure~\ref{fig:sound_design_loop_iterative} shows a general model for iterative sound-matching.
-To begin the process, the parameters are arbitrarily initialized (usually random generation), the similarity of the target and output is measured, and the parameter is \textit{optimized} with the goal of increasing the similarity, or alternatively, reducing the loss.
-This process repeats until termination.
-}
-\begin{figure}[ht]
-    \centering
-\begin{tikzpicture}[node distance=2cm, auto]
-
-    % Nodes
-    \node (start) [text centered] {\( g (\hat{\theta}) = x \)};
-    \node (L) [above of=start,right of=start, text centered] {\( L(\hat{\theta},t) {\ \propto \ } \phi(x) - \phi(t) \)};
-    \node (optimize) [below of=L,right of=L, text centered] {Optimize $\hat{\theta}$};
-    \node (new_theta) [below of=optimize,right of = start, text centered] {New \( \hat{\theta} \)};
-
-    % Highlight and arrow for the start node
-    \draw[->, very thick, red] (start) ++(-2,1.5) -- (start)
-        node[midway, below, align=center, sloped, color=red] {Start}
-        node[midway, above, align=center, sloped, color=red] {Random $\hat{\theta}$};
-
-    % Arrows with multi-line labels
-    \draw[->, bend left] (start) to node[midway, right, align=center] {} (L);
-    \draw[->, bend left] (L) to node[midway, right, align=center] {} (optimize);
-    \draw[->, bend left] (optimize) to node[midway, below, align=center] {} (new_theta);
-    \draw[->, bend left] (new_theta) to node[midway, left, align=center] {} (start);
-
-\end{tikzpicture}
-    % \caption{ Iterative approach to sound design.
-Synthesizer $g$ with arbitrary initialized parameters $\hat{\theta}$, creates sound $x$.
-The target sound $t$ is used as the goal.
-Parameters $\hat{\theta}$ are adjusted to minimize error (or loss) $L(\hat{\theta},t)$, where $L$ is proportional to the difference between the representations of $x$ and $t$.
-$\phi$ is the audio feature extractor, or representation function.}
-    \caption{ Iterative approach to sound design.}
-    \label{fig:sound_design_loop_iterative}
-\end{figure}
-
-In the upcoming ``Past Works'' section, we will discuss the evolution of prior sound-matching research, with \textit{differentiable iterative sound-matching} being the latest extension of the field.
-
-\begin{table*}[t]
-\centering
-\caption{Summary of select works in sound-matching.
-Gen.
-refers to whether the generations are done in 1-shot (generally, output by a neural network) or a search is conducted with the goal of iteratively getting closer to the target sound.}
-\setlength{\tabcolsep}{3pt} % Reduce horizontal spacing
-\renewcommand{\arraystretch}{1.05} % Reduce vertical spacing
-\scriptsize % Reduce font size
-\resizebox{\textwidth}{!}{
-\begin{tabular}{|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{1.2cm}|>{\raggedright\arraybackslash}m{1.2cm}|>{\raggedright\arraybackslash}m{1cm}|>{\raggedright\arraybackslash}m{1cm}|}
-\hline
-\textbf{Work} & \textbf{Synthesis} & \textbf{Loss} & \textbf{Heuristics} & \textbf{Goal} & \textbf{Domain} & \textbf{Year} & \textbf{Gen. } \\
-\hline
-Horner \textit{\textit{et al.}}~\cite{horner1993machine} & Non-diff FM & STFT (McAulay-Quatieri) & Direct & Replicate & In & 1991 & Iter \\
-\hline
-Mitchel \textit{\textit{et al.}}~\cite{mitchell2007evolutionary} & Non-diff FM & Spectral Relative & Evolutionary & Replicate & In (Contrived) & 2007 & Iter \\
-\hline
-Yee-King \textit{\textit{et al.}}~\cite{yee2018automatic} & Non-diff VST & P-Loss (for Supervised), Spectral (MFCC) & Supervised \& Evolutionary & Replicate & In & 2018 & Both \\
-\hline
-Esling \textit{\textit{et al.}}~\cite{esling2019flow} & Non-diff VST & Spec MSS/SC P-Loss & Supervised \& Modeling & Replicate & In* \& Out & 2019 & 1-shot \\
-\hline
-Shier \textit{\textit{et al.}}~\cite{shier2020spiegelib} & Custom, non-diff & FFT \& Spectral & Supervised \& Direct & Replicate & In & 2020 & Both \\
-\hline
-Salimi \textit{\textit{et al.}}~\cite{salimi2020make} & Non-diff & STFT and Envelope & Supervised & Imitate & Out & 2020 & 1-shot \\
-\hline
-Masuda \textit{\textit{et al.}}~\cite{masuda2021soundmatch} & DDSP & P-Loss initially, fine-tune with Spec. & Supervised & Both & In \& Out & 2021 & 1-shot \\
-\hline
-Han \textit{\textit{et al.}}~\cite{han2023perceptual} & NN-encoder $\rightarrow$ physical model & PNP (approx. of STFT) & Supervised & Replicate & In & 2023 & 1-shot \\
-\hline
-Vahidi \textit{\textit{et al.}}~\cite{vahidi2023mesostructures} & Differentiable FM & JTFS/MSS & Direct (grad desc.) & Replicate & In & 2023 & Iter \\
-\hline
-Uzrad \textit{\textit{et al.}}~\cite{uzrad2024diffmoog} & DDSP chains & Synth. Chain and P-Loss & Supervised & Replicate & In \& Out & 2024 & 1-shot \\
-\hline
-\end{tabular}
-}
-
-\label{tab:summary}
-\end{table*}
-
-\section{Past Works}
-\label{sec:related_works}
-Table~\ref{tab:summary} contains an overview of relevant literature, covering important works in sound-matching.
-adsfEarlier works in this domain focused on FM synthesis and genetic algorithms (or direct optimization methods)\cite{justice1979analytic,horner1993machine,mitchell2007evolutionary}.
-Along with the exponential increase in computation power, later works utilized a wider range of approaches to parameter search, estimation, and audio synthesis; particularly with supervised learning and neural networks (\gls{NN})~\cite{yee2018automatic,engel2020ddsp,esling2019flow}.
-
-\subsection{Earlier Works}
-FM synthesizers are simple to implement, yet very expressive~\cite{chowning1973synthesis}.
-Perhaps the earliest foundational work in sound-matching is the analytical approach by Justice~\cite{justice1979analytic} toward the decomposition and recreation of sounds with FM synthesis using one carrier, one modulator, and the corresponding envelopes:
-\[ f(t) = I_c(t) \cos(\omega_c t + I_m(t) \cos(\omega_m t))
-\]
-Where $f(t)$ is the output signal, $I_c(t)$ is the carrier amplitude, $\omega_c$ is the carrier frequency, and $I_m(t)$ and $\omega_m$ are the envelope and frequency for the modulator~\cite{justice1979analytic}.
-
+Perhaps the earliest foundational work in sound-matching is the analytical approach by Justice~\cite{justice1979analytic} toward the decomposition and recreation of sounds using the simple FM synthesizer described in Section~\ref{sec:dsp}.
 The sound-matching works that followed used other heuristics for finding the synthesizer parameters, yet the simple FM synthesizer structure remained unchanged.
 Horner \textit{et al.}~\cite{horner1993machine} used GAs for re-synthesis of sounds with an FM synthesizer using one modulator and one carrier oscillator, and the McAulay-Quatieri method~\cite{mcaulay1986speech} for measuring loss.
-
- % MQ calculates the STFT and marks the start and end of \textit{trends} in a sound with the assumption that the harmonic content belongs to trends that have a beginning and an end.
 
@@
 
-% Testing sound-matching experiments can be difficult, in part due to the large number of parametric components and navigation of complex non-linear systems.
-The issue with finding the best approaches to sound-matching is exacerbated when slight changes to the pipeline render previous results moot.
-Perhaps due to these reasons, the contrived methodology was not extensively tested by Mitchell and Creasy, who only provided a single experiment where the best evolutionary strategy found by in-domain search would also perform well in out-of-domain synthesis.
-\subsection{Introduction of Deep-Learning}
-Recent years have seen more works in sound-matching using machine learning techniques.
+Recent years have seen more works in sound-matching using supervised machine learning techniques.
 In 2018, Yee-King \textit{et al.} rendered 60,000 audio-parameter pairs from the \textit{Dexed} \gls{VST} synthesizer~\footnote{https://asb2m10.github.io/dexed/}, and showed that NNs trained on this dataset can outperform GA and hill-climber (\gls{HC})~\cite{hoffmann2000heuristic} methods in rendering speed, with slight improvements in MFCC error (used as an objective performance test).
@@
 Importantly, informal hearing tests revealed that the performance of even the best NN model was unsatisfactory, possibly due to the complex nature of the synthesizer, which features 155 parameters.
-
-% Programming synthesizers and querying sounds by imitating human vocals has also been a subject of interest~\cite{Cartwright2014SynthAssistQA,siamesevocalimitation2018}.
-For example, Cartwright and Pardo created a framework for searching pre-made patches for a synthesizer to best imitate vocals using audio features such as pitch, spectral centroid, and loudness~\cite{Cartwright2014SynthAssistQA}.
-
-Esling \textit{et al.} noted the issue of using P-Loss in sound-matching~\cite{esling2019flow}, considering the non-linear mapping between parameters of a synth and the audio output.
-In their approach, dubbed as ``flowsynth'', they used a large parameter-to-audio dataset for a commercial \gls{VST} synthesizer.
-In flowsynth, audio is encoded to an auditory space, using a NN autoencoding approach~\cite{hinton2006reducing}.
-The encodings in the auditory space can then be converted to the parameter space, but only for that particular synthesizer.
-Using P-Loss and Spectral Convergence (Frobenius norm of the spectrograms) as metrics, flowsynth outperformed other NN approaches that did not use an auditory to parameter space conversion, but formal hearing tests were not conducted.
 
@@
 
-% Rather than focusing on a particular implementation, Shier \textit{et al.} presented Spieglib, a library for implementation of sound-matching pipelines (which Shier \textit{et al.} referred to as ``automatic parameter search'')~\cite{shier2020spiegelib}.
-This library provides extensible choices of deep-learning estimators with \textit{tensorflow}~\cite{abadi2016tensorflow}, genetic estimators with \textit{DEAP}~\cite{DEAP_JMLR2012}, feature extractors with \textit{librosa}~\cite{mcfee2015librosa}, and different choices for synthesizers, evaluators, dataloaders.
-Shier \textit{et al.} provide an example experiment using this library.
-This example has a setup similar to Yee-King \textit{et al.}~\cite{yee2018automatic}, where the \textit{Dexed} VST was used and different genetic and NN estimators were compared by using MFCC difference as the minimization objective.
-Depending on the NN architecture, MFCC or STFT spectrograms were used as the input.
-Unlike previous comparisons of NN and genetic algorithms, the best estimator was the multi-objective non-dominated sorting genetic algorithm III (NSGA-III).
-This algorithm had multiple minimization objectives: 13-band MFCC, STFT, and five spectral features.
-Shier \textit{et al.} report that all models performed relatively well at imitation of harmonic distributions, but the multi-objective function was the only optimizer which did \textit{not} struggle with matching temporal envelopes of the audio spectrum~\cite{shier2020spiegelib}.
-
 Differentiable loss functions that use spectrogram differences can be computationally expensive.
@@
 
+A neural approximation for another part of the supervised sound-matching chain---this time the synthesizer---was proposed by Barkan \textit{et al.}~\cite{barkan2023inversynthII}.
+As they noted, without a differentiable synthesizer, ``model-based'' (or what we call supervised) approaches cannot directly compare $x_o$ to $x_t$, often opting for P-Loss, which may not correctly map the parameters of sound to the output audio~\cite{esling2019flow,han2023perceptual,masuda2023improving}.
+Given a synthesizer and a large dataset of sounds and corresponding parameters, a model which maps sounds to parameters, and another model which approximates the synthesizer can be trained, allowing the implementation of a loss function which combines P-Loss with STFT differences.
+This ``Inversynth II'' (IS2) approach yielded significant improvements to previous works which did not use the STFT approximations for loss~\cite{esling2019flow,barkan2019inversynth}.
+
+Barkan \textit{et al.} attemped to  improve the IS2 model with Inference-Time
+Finetuning (ITF).
+For ITF, the synthesizer approximation is frozen and the encoder is iteratively finetuned for the sample sample using gradient descent in order to minimize P-Loss, stopping at convergence or when maximum number of iterations has reached~\cite{barkan2023inversynthII}.
+Barkan \textit{et al.} report that perhaps due to the discrepancy between the true and approximated synthesizer or overfitting, the application of ITF to IS2 generally worsened the spectral and manual performance metrics.
+
 Uzrad \textit{et al.}~\cite{uzrad2024diffmoog} took another unique approach to sound-matching: using a differentiable \textit{synthesis chain} of DSP generators and effects and a loss function that combines P-Loss with a \textit{signal-chain loss}.
@@
 
-% Uzrad \textit{et al.}\cite{uzrad2024diffmoog} used log-spectral distance (LSD) as their loss function, where   
-% $\text{LSD} = \| \log(S(t)) - \log(S(x)) \|_F$, S is the mel-spectrogram of the sounds and F is the Frobenius norm~\cite{golub2013matrix}. The Frobenius norm of a matrix $A$ is defined as
-% $ \|A\|_F = \sqrt{\sum_{i,j} |a_{ij}|^2}$.
-Here, $A$ would be the 2D matrix resulting from subtracting two log-spectrograms.
-Along with other works, \textbf{non-convexity of loss surfaces when applying gradient optimization} is a noted issue~\cite{turian2020sorry,vahidi2023mesostructures,uzrad2024diffmoog}.
+Some recent works have opted for the use of audio embeddings as a similarity metric.
+In a recent work, Cherep \textit{et al.} used latent representations from the CLAP model~\cite{wu2023large} along with a differentiable synthesizer~\cite{synthhaxcherep2023} to create creative interpretations of sound effects.
+In their approach, a desired sound-effect is described in text and embedded using clap, and gradient free optimizer~\cite{evosax2022github} iteratively updates the synthesizer's parameters to minimize difference in the text and output sound embeddings.
+Based on manual hearing tests, this approach did not produce the ``correct'' sounds more frequently than previous works~\cite{kreuk2022audiogen}, however, it did yield better scores for ``artistic-interpretation'' (or what we refer to as imitation).
 
-\subsection{What is Lacking In the Field}
-\label{sec:lacking}
-%   % First, we need to clearly define the characteristics we are looking for with non-generic loss functions, otherwise the chances of finding the sound(s) of interest is slim.
-Second, we need to move beyond direct replication, and focus on imitating target sounds that are outside the synthesizers capabilities.
-Third, more complex methods of synthesis need to be explored for iterative sound-matching.
-Fourth, we need to find optimization methods that can effectively navigate the the sinusoidal nature of loss function landscapes.
+% masuda and esling both noted out-of-domain drop in performance
+% yee king noted performance drop in manual surveys 
+%p-loss bad : masuda-improving, han2023, esling-flowsynth
+% manual performance measure: barkan, masuda2021soundmatch, yee-king
 
-% Based on the literature reviewed, we find several gaps in past research that makes the application of sound-matching for sound-designers difficult.
-Thus far, replication of in-domain sounds has been the main subject of research.
-With in-domain replication, we already have the target sound \textit{and} the program that makes it.
-This simplifies the problem of measuring performance and is the reason why this type of sound-matching research is by far the most common.
-However, practical application of sound-matching requires answers to much harder problems.
-Sound-matching would be an interesting creative tool if we could use it to either:
-% \begin{enumerate}
-%  \item \textbf{Replicate out-of-domain sounds}: this would give us the parameters that could closely approximate a target sound, and we can modify the sound further by modulating the synthesizer. 
-% \item \textbf{Imitate in-domain or out-of-domain sounds}: This would allow for applying the characteristics of one sound to another. 
-% \end{enumerate}
-
-% We need to be aware of the ``open set recognition'' (\gls{OSR}) problem here \cite{mundt2019open,gers2000learning}.
-OSR is a well documented issue in DL that manifests when the set or category that the network is supposed to reject or accept cannot be explained via examples.
-With imitation or replication of out-of-domain sounds, we can expect issues related to OSR to occur, since target and output sounds would be from different domains~\cite{salimi2021percussive}.
-
-% Perhaps one method of addressing the OSR problem is the use of representations (or feature extraction) methods that reduce sounds to a feature vector in the same domain.
-For example, a representation that only looks at the envelope of sounds is far less likely to be subject to the OSR problem.
-\textcolor{highlight}{We highlight the lack of non-generic loss functions as a major weakness in previous works and an interesting area to explore}.
-
-% Another weakness \textcolor{highlight}{is that the synthesis methods in past works are not representative of the common methods used by sound designers and modern synthesizers}.
-Simple FM/AM synthesizers have generally been the go-to methods of sound-matching.
-We believe that general statements regarding the effectiveness of an approach to iterative sound-matching (particularly in regards to a loss function) can only be reasonably made in the context of the subset of the tested methods of synthesis.
-% \textcolor{highlight}{The simplicity of parameter optimization approaches} is another area of weakness in past works.
-Classic genetic algorithms update the parameters randomly, while in recent differentiable settings, simple gradient descent methods are used.
-Past works have provided issues which arise in differentiable parameter estimation due to periodic loss functions.
-Perhaps the use of reinforcement learning (\gls{RL}), or other algorithms that continually learn from their environment, could help with this problem.
-
-% \subsection{Defining The Major Issues}
-% \label{sec:addressing_problems}
-
-Having looked at the current available literature, we note four major areas of weakness in sound-matching.
-We target the first two issues in this work, while the latter two issues are left for future work.
-\begin{enumerate}
-    \item \LossSelect: Is there a \textit{best performing} loss function, or is the selection of the loss function dependent on factors such as the sound domain, synthesizer capability and the desired characteristics of the output?
-    \item \SynthSelect: There is a lack of diversity in the synthesis methods used in sound-matching.
-The use of different DSP functions for iterative sound-matching and their effects on the outcome or the interactions with the loss function have not been tested.
-    \item \PeriodicLoss: The sinusoidal nature of loss function landscapes is a problem that appears frequently.
-In differentiable settings, this problem causes gradient descent updates to local minima.
-    \item \OutDomain: Sound-matching with out-of-domain sounds is an unexplored area, yet a necessary one for practical applications for sound designers.
-\end{enumerate}
-
-\section{Programs, Methodology, and Results}
-\label{sec:progs_methodology_results}
-Towards addressing the problems of \LossSelect{} and \SynthSelect{}, we compare the performance of iterative sound-matching experiments, where four differentiable loss functions are applied to four differentiable synthesizers.
-What we mean by ``performance'' is the similarity of the final output to the target, as measured by P-Loss, MSS, and---most importantly---manual listening tests.
-
-\textbf{We hypothesize that the relative performance of loss functions for sound-matching is influenced by the method of synthesis}.
-If true, this would strengthen the idea that the selection of the loss function and synthesis method is not an optimization problem, but a subjective choice depending on the needs of the sound designer.
-
-% In Section~\ref{sec:methodology} we lay out the experiment methodology.
-Section~\ref{sec:loss_implementation} discusses the details of implementation for the loss functions.
-Section~\ref{sec:programs} covers the characteristics of the programs, and finally Section~\ref{sec:results_posthoc} shows that the effect of loss function choice on the success of iterative sound-matching is dependent on the signal process chain of interest.
-
-\subsection{Experiment Methodology}
-\label{sec:methodology}
-% We have highlighted the lack of attention given to the selection of loss functions in sound-matching, and noted that the few existing works addressing this issue evaluate their proposed losses using only simplistic sound synthesis methods.
-\textcolor{highlight}{We hypothesize that the best performing loss function is influenced by the synthesis techniques utilized in the synthesizer}.
-This hypothesis targets the issues of \LossSelect~and \SynthSelect.
-
-We test four different loss functions and compare their success in sound-matching using four different synthesizers.
-We verify the success of the sound-matching process using two automatic measures of sound similarity (MSS and P-Loss) and a manual blinded hearing test conducted by two of the authors where Likert scores~\cite{jebb2021review} of 1-5 are assigned to a subset of final outputs and the corresponding targets.
-
-An \textit{experiment} is one iteration of sound-matching given a loss function, target sound,
-and synthesizer program.
-When the final iteration is complete, a score is given (automatically with an algorithm or with manual listening) to an experiment that measures its success in replicating the target sound.
-Consider the following simplified scenario: we have two loss functions, $L_1$ and $L_2$, and a single synthesizer program $P$ with a target parameter set $\theta^*$ where the target sound is $t$, or $P(\theta^*) = t$.
-We run a large number (in this case, 300 for each loss function) of experiments where we randomly set the initial ($\theta_0$) and target parameters ($\theta^*$) of $P$ and run an iterative sound-matching loop for each loss function, with maximum number of iterations set to 200.
-When the loop is terminated, we measure the similarity of the final output sound $x$ (where $P(\theta_{199}) = x$) to $t$, or if using P-Loss we compare $\theta_{199}$ to $\theta^*$.
-Having 300 experiments gives us a distribution of losses to determine whether $L_1$ or $L_2$ led to better performance.
-This simplified example leads us to our methodology which has four loss functions, four synthesizer programs, and three different evaluation methods of scoring sound similarity.
-
-\subsubsection{Automatic and Manual Evaluation}
-\label{sec:evaluation_manual_auto}
-The two automatic evaluation methods used here are P-Loss and MSS, as described in Section~\ref{sec:loss_funcs}.
-For P-Loss, the parameters are normalized between 0-1 based on the valid ranges defined in the Faust program, and the L1 distance between the normalized parameters is calculated.
-MSS is computed using a hop size of $100$ samples, and FFT window lengths of $(512, 1024, 2048, 4096)$.
-
-Automatic sound evaluation methods are often unreliable, as their alignment with the subjective judgments of human listeners is uncertain.
-Therefore, listening tests are conducted by randomly sampling 40 experiment results for each program and loss function pair.
-Two of the authors then assign blinded similarity scores to the outputs and targets using a 5-point Likert scale (1 = no similarity, 5 = near identical)~\cite{jebb2021review}.
-Using Spearman's rank correlation~\cite{spearman1987proof,rebekic2015pearson}, we found a very strong rate of agreement between the human listeners (see Section~\ref{sec:consistency_in_rankings}), a good indicator that the scores assigned by manual listeners can be treated as the ground truth.
-We then combine the ranks assigned by both authors, giving us 80 ranks per program and loss function pair.
-
-% Agreement between the assigned scores is measured using Spearman’s rank correlation, which provides both a correlation coefficient ($\rho$) ranging from –1 (perfect negative correlation) to 1 (perfect positive correlation), and a p-value testing the null hypothesis of no correlation~\cite{spearman1987proof,rebekic2015pearson}.
-Across all programs, the correlation was very strong ($\rho = 0.86$, $p < 10^{-180}$).
-Per-program correlations were also very strong: $\rho = 0.71$ for \BPNoise{} ($p < 10^{-25}$), $\rho = 0.64$ for \AddSineSaw{} ($p < 10^{-19}$), $\rho = 0.84$ for \AmpMod{} ($p < 10^{-43}$), and $\rho = 0.85$ for \FMMod{} ($p < 10^{-44}$).
-This high rate of agreement is a good indicator that the scores assigned by manual listeners can be treated as the ground truth.
-
-% For each program and loss function, the blind manual hearing test gives us 80 ranks, which are given by two of the authors to a set of 40 randomly selected experiment results.
-This gives 160 pairs of ranks per program, and 640 for all programs.
-We computed Spearman correlations between the two reviewers.
-
-\subsubsection{Determining Best Performers}
- Figure~\ref{fig:posthoc_evaluation} is a visualized summary of how the loss functions are ranked.
-We have four loss functions and four synthesizer programs.
-The maximum number of iterations set to 200, and 300 experiments for each loss/synthesizer combination.
-Each synthesizer program is paired with four loss functions and 300 experiments are conducted and evaluated automatically with P-Loss and MSS, and manually with Likert scores.
-
-For each program and loss function pair, we have three distributions that can be used to rank the loss functions.
-Two with 300 automatically assigned similarities (P-Loss and MSS) and one with 80 Likert scores (combining the 40 ranks assigned by each author).
-For consistency and statistical robustness, the distributions are upsampled to 1000 values using bootstrapping~\cite{tibshirani1993introduction}.
-Bootstrapping gives an estimation of the distribution for the mean performance of each experiment.
-$k$ (set to 1000) samples of $n$ values (set to 100\% of the values) are taken with replacement from the empirical distribution (list of 300 performance values for MSS and P-Loss or 80 for manual rankings).
-The mean performance is calculated for each of the $k$ samples.
-The $k$ estimates of the mean performance give us a bootstrapped distribution for each loss function.
-
-For each program, the loss functions are ranked using the non-parametric Scott-Knott (\gls{NPSK}) test~\cite{tantithamthavorn2017mvt,tantithamthavorn2018optimization}, which ranks the loss functions from rank 1 (best, or highest distribution values) to maximum rank of 4 (worst, with lowest distribution values).
-NPSK may determine that multiple functions belong to the same rank, in which case the maximum rank would be less than 4.
-As mentioned before, lower P-Loss and MSS values indicate higher similarity, therefore (for consistency with manually assigned ranks) the reciprocal of their output is used ($x^{-1}$, if $x$ is the P-Loss or MSS value), where higher values indicate greater similarity.
-
-\input{extra_figures/evaluation_graph.tex}
-
-\subsubsection{Training Parameters and Gradient Calculations}
- % Previous work utilized gradient fields for analysis of the algorithm performance~\cite{vahidi2023mesostructures}.
-Gradient fields are not the center of our analysis, but we include them in Appendix~\ref{appendix:gradient_fields}.
- For updating the synthesizer parameters (or weights), we use RMSProp, which operates similar to stochastic gradient descent (SGD)~\cite{goodfellow2016deep}, with the caveat that the gradients of each weight are scaled by the root-mean-square of past gradients of that weight.
-Based on some initial test runs, we used an arbitrary fixed learning rate of 0.045 for all experiments (learning rates used in Vahidi \textit{et al.}~\cite{vahidi2023mesostructures} is unknown).
-The maximum number of iterations is set to 200, where, based on our observations, the parameters have either irrecoverably diverged outside the acceptable ranges, or are stuck at a local minimum.
-Additionally, the gradients are large, and calculated backwards sequentially through the signal processing chain.
-This backward calculation resembles the \textit{exploding gradients} problem in recurrent neural networks~\cite{gers2000learning}.
-Gradient clipping~\cite{goodfellow2016deep} is preemptively used to ensure that the  $\ell_2$ norm of all gradients does not exceed the threshold of 1.
-% \mathbf{\gamma}_i^{\text{clipped}} = \mathbf{\gamma}_i \cdot \mathcal{S}
-% \]
-% Where $\mathcal{S}$ is a value less than or equal to 1, depending on whether the norm of the gradients exceeds the threshold $c$.  
-% \[
-% \mathcal{S} = \frac{c}{\max(\|\mathbf{\gamma}\|_2, c)}
-% \]
-
-\subsection{Loss Function Implementation Details}
-\label{sec:loss_implementation}
-\subsubsection{STFT Losses}
-Due to their ubiquity and lower cost of gradient calculation, we use STFT as the basis of two loss functions.
-As discussed in Section~\ref{sec:matching_types}, the L1 or L2 distances are the most common methods of comparing different spectrograms, however, there are situations where this might lead to shortcomings.
-An example given by Vahidi \textit{et al.}~\cite{vahidi2023mesostructures} is two chirplets (tones that increase in pitch) starting and ending at the same frequencies.
-If one of these tones is slightly shifted in time, then the spectrograms will no longer overlap at any point, despite their sonic similarity (similar to the idea that a linear equation $f(x) = ax$ can never overlap with a slightly shifted version of itself, or $f(x) = ax +\epsilon$).
-In this example, we do not know if this is an issue of the spectrograms, or the L1 measure used for their comparison.
-In fields such as computer vision, Structural Similarity Index (SSIM)~\cite{wang2004imagesssim,wang2009mean} and Scale-Invariant Mean Squared Error (SIMSE)~\cite{barron2014shapessimse} have been used as alternatives that improved performance in various tasks relative to L1.
-
-We define the \LoneSpec~and \SIMSESpec~functions as the application of L1 and SIMSE to the STFT spectrograms.
-The STFT spectrogram function uses 512 FFT bins, window size of 600 samples, and hop length (how many samples the window shifts) of 100 samples.
-The L1 and SIMSE implementations are differentiable.
-
-\subsubsection{JTFS Loss}
-The \JTFS~loss function is the application of L1 difference to the JTFS representations of two sounds~\cite{vahidi2023mesostructures}.
-The code used for the JTFS transformation is the differentiable implementation of a 1-dimensional JTFS function provided by Andreux \textit{et al.}~\cite{kymatio}.
-
-\subsubsection{Soft-DTW Loss}
-We use the soft-DTW function, which is differentiable and --- depending on its parameters --- not shift-invariant~\cite{cuturi2017soft,janati2020spatio,tavenard.blog.softdtw}.
-The loss function \DTWEnv~is the application of the soft-DTW function to the envelope of the two sounds being compared~\cite{lyons1997understanding}.
-The envelope is calculated by creating the STFT spectrogram of a sound (the same process used in Section~\ref{sec:fourier_specs}) and summing the values at each timestep.
-
-\subsection{The Synthesizers}
-\label{sec:programs}
-The synthesizer programs are meant to be simple examples that test the building blocks of digital sound synthesis.
-Subtractive, additive, and FM/AM synthesis are three of the most common techniques in sound design~\cite{smith1991viewpoints}.
-In subtractive synthesis, frequencies are removed from a sound via digital filters.
-In additive synthesis, complex sounds are created via the linear combination of simpler sounds~\cite{lyons1997understanding,smith2007introduction}.
-As discussed previously, FM/AM synthesis refers to the general technique of modulating the frequency or amplitude of a waveform (the carrier) by another waveform (the modulator).
-
-For each program, we provide the Faust code~\cite{orlarey2009faust}, which can be run in the online IDE~\footnote{\url{https://faustide.grame.fr/}}.
-Faust is a functional language for audio synthesis that can succinctly define signal processing chains.
-Following Braun's methodology~\cite{braun2024dac}, programs are first defined in Faust, then converted (or \textit{transpiled}) to differentiable Jax functions using the DawDreamer library\footnote{\url{https://github.com/DBraun/DawDreamer}}.
-
- Once the programs are transpiled to differentiable synthesizers, the iterative sound-matching procedure is:
- \begin{enumerate}
-    \item Initialize $\theta^*$ and $\hat{\theta}$, i.e., random generation of target and initial parameters uniformly over a predefined range
-    \item Generating the output of the synthesizer with $\hat{\theta}$ (the length of the output is set to 1 second with sample-rate of 44100 Hz) 
-    \item Calculating the loss between the target and output
-    \item Applying gradient updates to the synthesizer parameters
-    \item Repeating the second step with the updated parameters $\hat{\theta}$, until maximum number of iterations has been reached
- \end{enumerate}
-
-% A general overview of the programs and the predefined parameter ranges is given in Table~\ref{tab:programs}.
-
-In the following subsections, we describe the programs and their parameters in more detail, followed by the methodology in Section~\ref{sec:methodology} and results in Section~\ref{sec:results_posthoc}.
-
-% \begin{table*}[htbp]
-% \centering
-% \begin{adjustbox}{max width=\textwidth}
-% \begin{tabular}{|c|c|c|c|l|}
-% \hline
-% \textbf{ID} & \textbf{Parameters} & \makecell{\textbf{Param. 1}\\\textbf{Range}} & \makecell{\textbf{Param. 2}\\\textbf{Range}} & \textbf{Description} \\
-% \hline
-% \BPNoise & hp\_cut, lp\_cut & 50-1000 & 1-120 & Band-pass of Noise signal \\
-% \hline
-% \AddSineSaw & saw\_freq, sine\_freq & 20-1000 & 20-1000 & Addition of sine and saw waves \\
-% \hline
-% \AmpMod & amp, carrier & 1-20 & 20-1000 & LFO on amplitude of noise osc \\
-% \hline
-% \FMMod & amp, carrier & 0-5 & 0-4 & LFO on amplitude of saw osc \\
-% \hline
-% \end{tabular}
-% \end{adjustbox}
-% \caption{Overview of Program Parameters and the range of values each parameter can take.}
-% \label{tab:programs}
-% \end{table*}
-
-\begin{lstlisting}[caption={\BPNoise}, label={lst:program0}, language=Faust,
-                  float, floatplacement=!H, xleftmargin=1em, xrightmargin=0.5em, firstnumber=0, aboveskip=0em, belowskip=-1em]
-import("stdfaust.lib");
-lp_cut = hslider("lp_cut",900,50,1000,1);
-hp_cut = hslider("hp_cut",100,1,120,1);
-process = no.noise:fi.lowpass(3,lp_cut):fi.highpass(10,hp_cut);
-\end{lstlisting}
-
-\begin{lstlisting}[caption={\AddSineSaw}, label={lst:program1},language=Faust,float,floatplacement=!H,xleftmargin=1em,xrightmargin=0.5em,firstnumber=0,aboveskip=0em, belowskip=-1em]
-import("stdfaust.lib");
-saw_freq = hslider("saw_freq",800,20,1000,1);
-sine_freq = hslider("sine_freq",300,20,1000,1);
-sineOsc(f) = +(f/ma.SR) ~ ma.frac:*(2*ma.PI) : sin;
-sawOsc(f) = +(f/ma.SR) ~ ma.frac;
-process = sineOsc(sine_freq)+sawOsc(saw_freq);
-\end{lstlisting}
-
-\begin{lstlisting}[caption={\AmpMod}, label={lst:program2},language=Faust,float,floatplacement=!H,xleftmargin=1em,xrightmargin=0.5em,firstnumber=0,aboveskip=0em, belowskip=-1em]
-import("stdfaust.lib");
-amp = hslider("amp",0.5,0,5,0.01);
-modulator = hslider("modulator",0.5,0,4,0.01);
-sineOsc(f) = +(f/ma.SR) ~ ma.frac:*(2*ma.PI) : sin;
-process = no.noise*sineOsc(modulator)*amp;
-\end{lstlisting}
-
-\begin{lstlisting}[caption={\FMMod}, label={lst:program3},language=Faust,float,floatplacement=!H,xleftmargin=1em,xrightmargin=0.5em,firstnumber=0,aboveskip=0em, belowskip=-1em]
-import("stdfaust.lib");
-carrier = hslider("carrier",100,20,1000,1);
-amp = hslider("amp",6,1,20,1);
-sineOsc(f) = +(f/ma.SR) ~ ma.frac:*(2*ma.PI) : sin;
-sawOsc(f) = +(f/ma.SR) ~ ma.frac;
-process = sineOsc(amp)*sawOsc(carrier);
-\end{lstlisting}
-
-\subsubsection{\BPNoise}
-\label{sec:program0}
-\BPNoise{} is a bare bones example of subtractive synthesis using digital filters~\cite{smith2007introduction}.
-Despite the ubiquity of subtractive synthesis in practical sound design, it has rarely been tested in differentiable sound-matching.
-In this program, a noise signal is fed through a band-pass (BP) filter.
-This removes the frequencies outside the low-pass (LP) and high-pass (HP) cutoffs.
-The noise generator produces all frequencies with random variation over time.
-The LP filter removes frequencies over its threshold frequency, and the HP removes frequencies lower than its threshold~\cite{smith2007introduction}.
-The search parameters are the cutoff thresholds for the LP and HP filters.
-Listing~\ref{lst:program0} shows the Faust code for \BPNoise.
-
-\subsubsection{\AddSineSaw}
-\label{sec:program1}
-\AddSineSaw{} is an additive program that combines a saw and a sine wave function.
-Like subtractive synthesis, additive synthesis is a common approach to sound design that has not been extensively tested as a benchmark in sound-matching.
-The search parameters here are the frequency of the sine and saw oscillators.
-Listing~\ref{lst:program1} is the Faust code for \AddSineSaw.
-
-\subsubsection{\AmpMod}
-\label{sec:program2}
-\AmpMod{} involves amplitude modulation of a noise generator by an LFO with \textit{modulator} as its frequency and a global \textit{amp} value that does not change over time and applies to the entire signal.
-The search parameters for this program are the LFO frequency and the amp value.
-The amp value acts as a global volume control, and since the sounds are normalized before analysis, it is likely inconsequential to the performance.
-This program is meant to be a stepping stone to \FMMod, which utilizes proper AM synthesis.
-Listing~\ref{lst:program2} is the Faust code for \AmpMod.
-
-\subsubsection{\FMMod}
-\label{sec:program3}
-Relative to additive and subtractive synthesis, AM/FM synthesis are less common methods of sound design.
-Due to their frequent use in previous works, we also tested an AM synthesizer.
-The synthesizer program is the multiplication of a low frequency sine oscillator with frequency parameter \textit{amp}, and a saw oscillator with frequency parameter \textit{carrier}.
-Listing~\ref{lst:program3} is the Faust code for \FMMod.
-
-\subsection{Post hoc Analysis of Loss Performance}
-\label{sec:results_posthoc}
-As summarized in Figure~\ref{fig:posthoc_evaluation}, for each program and evaluation (or scoring) method, there are four sets 1000 bootstrapped values, corresponding to the four loss functions.
-A post-hoc test is conducted in two stages in order to determine whether there is a difference in performance between loss functions for each program, and if so, which are the best performers.
-The first stage determines whether there is a difference in group means, with the null hypothesis being that all groups have similar mean ranks.
-The second stage ranks the loss functions from best to worst using the non-parametric Scott-Knott test (\gls{NPSK})~\cite{tantithamthavorn2017mvt,tantithamthavorn2018optimization}.
-The Kruskal-Wallis test is used for the first stage~\cite{kruskal1952use}; this test pools and ranks all evaluation measures for a program, then tests whether the differences in mean rank for all loss function groups is zero.
-The Kruskal-Wallis test calculates an H statistic that is compared to a chi-square distribution with k-1 degrees of freedom (k is the number of groups, or 4, and degrees of freedom is 3).
-Using the significance level of 0.05, if the H statistic is greater than the critical value of the chi-square distribution then the null hypothesis is rejected, meaning that at least one group has a mean rank significantly above others~\cite{kruskal1952use}.
-Table~\ref{tab:kruskal_auto} shows the Kruskal-Wallis results for every program and evaluation method, and shows that significant differences between loss function performance measures (i.e., the bootstrapped distributions of the evaluation for each loss function) exist in all cases except MSS for the \AmpMod{} program.
-
-\begin{table}[ht]
-\centering
-\caption{Do loss functions have different median performance? Kruskal-Wallis results by program and bootstrapped evaluation results.}
-\begin{tabular}{|c|c|c|c|c|}
-\hline
-\textbf{Program} & \textbf{Eval. Method} & \textbf{H-Stat.} & \textbf{P-value} & \textbf{Reject} \\
-\hline
-\BPNoise & MSS      & 81.19  & $1.70 \times 10^{-17}$ & Yes \\
-\BPNoise & P-Loss  & 163.60 & $3.07 \times 10^{-35}$ & Yes \\
-\BPNoise & Manual & 9.45  & $2.39 \times 10^{-2}$ & Yes \\
-\AddSineSaw & MSS      & 308.97 & $1.14 \times 10^{-66}$ & Yes \\
-\AddSineSaw & P-Loss  & 348.42 & $3.28 \times 10^{-75}$ & Yes \\
-\AddSineSaw & Manual & 9.45  & $2.39 \times 10^{-2}$ & Yes \\
-\AmpMod & MSS      & 1.35   & $0.7171$               & No \\
-\AmpMod & P-Loss  & 366.76 & $3.50 \times 10^{-79}$ & Yes \\
-\AmpMod & Manual & 32.71 & $3.70 \times 10^{-7}$ & Yes \\
-\FMMod & MSS      & 564.65 & $4.65 \times 10^{-122}$ & Yes \\
-\FMMod & P-Loss  & 229.19 & $2.07 \times 10^{-49}$ & Yes \\
-\FMMod & Manual &  207.58 & $9.69 \times 10^{-45}$ & Yes \\
-\hline
-\end{tabular}
-\label{tab:kruskal_auto}
-\end{table}
-
-% \begin{table}[ht]
-% \centering
-% \begin{tabular}{|c|c|c|c|}
-% \hline
-% \textbf{Program} & \textbf{H-Stat.} & \textbf{P-value} & \textbf{Reject} \\
-% \hline
-% \BPNoise & 9.45  & $2.39 \times 10^{-2}$ & Yes \\
-% \AddSineSaw & 9.45  & $2.39 \times 10^{-2}$ & Yes \\
-% \AmpMod & 32.71 & $3.70 \times 10^{-7}$ & Yes \\
-% \FMMod & 207.58 & $9.69 \times 10^{-45}$ & Yes \\
-% \hline
-% \end{tabular}
-% \caption{Manual ranking Kruskal-Wallis results. Significant differences are observed for all programs.}
-% \label{tab:kruskal_manual}
-% \end{table}
-
-\subsubsection{Best Performers For Each Program}
-Based on the bootstrapped distribution of scores, the NPSK algorithm ranks the loss functions from 1 (best) to a maximum of 4 (worst).
-In cases where the distributions are similar, multiple loss functions can be clustered into the same rank.
-In Figure~\ref{fig:npsk_all}, we use color-coded violin plots to visualize the best performers for each program.
-The corresponding colors for each rank are
-\colorbox{rank1}{\textcolor{black}{\textbf{1}}} \colorbox{rank2}{\textcolor{white}{\textbf{2}}} \colorbox{rank3}{\textcolor{white}{\textbf{3}}} \colorbox{rank4}{\textcolor{black}{\textcolor{white}{\textbf{4}}}}
-
-\input{extra_figures/ranking_tables}
-\input{extra_figures/npsk_figures}
-
-\subsubsection{\BPNoise}
-For this synthesizer program, the spectrogram-based models performed the best.
-This makes intuitive sense, as the visual effects of a band-pass filter on white noise are readily apparent in a spectrogram.
-As shown in Figure~\ref{fig:npsk_p0}, manual hearing test and MSS selected \SIMSESpec~as the best performer and \LoneSpec~as the second-best performer, while P-Loss gave the reverse order.
-
-\subsubsection{\AddSineSaw}
-As shown in Figure~\ref{fig:npsk_p1}, JTFS was the best performing loss function in all evaluation methods. Moreover, all evaluation methods produced identical results.
-
-\subsubsection{\AmpMod}
-IP-Loss and manual hearing tests yield identical rankings, and MSS results vary for for ranks 2 to 4.
-It is not surprising that MSS results differ from the hearing tests, as we saw in Table~\ref{tab:kruskal_auto}, MSS showed no significant differences between groups.
-As shown in Figure~\ref{fig:npsk_p2}, all models selected \DTWEnv~as the best performer.
-This may be due to \DTWEnv's focus on periodic changes in loudness.
-
-\subsubsection{\FMMod}
-\DTWEnv~is again the best performer here. As shown in Figure~\ref{fig:npsk_p3}, all methods of analysis gave identical rankings to all programs.
-
-\subsection{Consistency In Rankings}
-\label{sec:consistency_in_rankings}
-\subsubsection{Manual Ranks} Agreement between the manually assigned scores is measured using Spearman’s rank correlation, which provides both a correlation coefficient ($\rho$) ranging from –1 (perfect negative correlation) to 1 (perfect positive correlation), and a p-value testing the null hypothesis of no correlation~\cite{spearman1987proof,rebekic2015pearson}.
-Across all programs, the correlation was very strong ($\rho = 0.86$, $p < 10^{-180}$).
-Per-program correlations were also very strong: $\rho = 0.71$ for \BPNoise{} ($p < 10^{-25}$), $\rho = 0.64$ for \AddSineSaw{} ($p < 10^{-19}$), $\rho = 0.84$ for \AmpMod{} ($p < 10^{-43}$), and $\rho = 0.85$ for \FMMod{} ($p < 10^{-44}$).
-
-Table~\ref{tab:combined_ranks} shows the SNPK rankings of bootstrapped evaluation results for each program.
-Both MSS and P-Loss gave a different rank from the hearing results in 3 out of 16 cases, which shows consistency between automatic and manual hearing tests, at least for the simple programs used in this work.
-We also observe that top ranks were consistent across performance evaluation methodologies, with the exception of P-Loss in \BPNoise, which narrowly picks a different spectrogram-based loss function.
-% \input{extra_figures/ranking_tables}
-
-\subsection{Takeaways}
-An important takeaway is that the loss function that yields the best sound-matching outcomes varies depending on the synthesizer program.
-This is perhaps due to the interaction between how the parameters of the synthesizer influence the sound, and the core sonic features of the loss function.
-
-We also see that the \DTWEnv{} and \SIMSESpec{} (to our knowledge, not previously used in iterative sound-matching), lead to the best results in three of the four programs.
-This shows room for further creative approaches to differentiable sound-similarity measures, particularly since we observed that SOTA loss functions can be ineffective depending on the design of a synthesizer.
-
-Finally, we surprisingly see somewhat consistent results between the rankings assigned by manual hearing tests (which we take as the ground truth), and automatic measures of P-Loss and MSS.
-Due to the simplicity of our programs and the occasional deviations, we cannot conclude that these automatic measures are substitutes for human hearing tests.
-
-\section{Summary, Weaknesses, and Conclusion}
-\label{sec:summary_conclusion}
-\textbf{Summary:} Sound-matching is an umbrella term for the algorithmic programming of audio synthesizers, often with the goal of assisting sound designers.
-Here we provided a history of sound-matching, major issues in the field, and discussed the importance of ``differentiable iterative sound-matching'' as a natural extension of current literature.
-
-The main hypothesis tested here is whether the performance of differentiable loss functions is influenced by the synthesis techniques used for sound-matching.
-We created iterative sound-matching experiments by combining four different loss functions with four different sound synthesis programs.
-We ranked the performance of the iterative sound-matching pipelines for every loss function and program, and observed that the success of the pipeline (that is, how closely the output sound matches the target sound) is program dependent.
-In other words, different synthesizer programs work best with different loss functions.
-Notably, we see that our novel use of DTW and SIMSE based differentiable loss functions (\DTWEnv{} and \SIMSESpec) can outperform what are regarded as the SOTA loss functions in 3 of 4 cases, although this is highly synthesizer dependent.
-
-P-Loss and MSS have frequently been used as automatic performance measures, yet their ``preference'' has rarely been compared to human rankings.
-We observed that automatic performance measures and manual listening tests were often in agreement, despite this, manual verification of sound-matching results remains a necessity for future works.
-
-\textbf{Weaknesses:} 
-While we cannot prove that a universally best similarity measure does not exist, we can advocate for more creativity in the field based on our findings.
-Compared to previous work, we presented a more cohesive approach to iterative sound-matching which utilizes a variety of loss functions and synthesis methods.
-However, there are many other methods of synthesis and sound-similarity that can be combined in practically infinite ways.
-Due to this large search space, we set arbitrary parameters for the various signal processing functions.
-We used bare-bones versions of STFT and JTFS with fixed parameters.
-We did not test complex synthesizers using parallel and sequential DSP functions.
-Arbitrary hyperparameters such as learning rate and max number of iterations were selected for the DL pipeline, and only the RMSProp optimizer was tested.
-
- Like the majority of previous works, this work utilizes in-domain sounds; that is, the target sound is made by the synthesizer, and the parameters are already known.
-This simplifies the issues of measuring sound-similarity, but it is not a realistic scenario for practical sound-matching.
-This problem is left for future work, discussed in the next section.
-
-\label{sec:future}
-\textbf{Future Work: } The problem of periodic loss-landscapes, noted by many previous works, remains unaddressed~\cite{turian2020sorry,vahidi2023mesostructures,uzrad2024diffmoog,bruford2024synthesizer}.
-Perhaps this problem emerges due to the periodic nature of sound, which requires better loss landscape navigation methods.
-An optimizer that is more aware of fluctuations in the gradients would perhaps lead to better solutions than simple gradient descent.
-Viewing the sound-matching problem as ``the navigation of an agent from an arbitrary point in a gradient field to a target'' closely resembles many classical problems in the field of reinforcement learning (RL)~\cite{sutton2018reinforcement}.
-Naturally, the application of RL and other heuristic search techniques to the problem of iterative sound-matching would be an important contribution.
-
-Contemporary works often involve the application of domain specific and computationally expensive loss functions~\cite{han2023perceptual,uzrad2024diffmoog}, use of large neural networks~\cite{hershey2017cnn,cramer2019look}, or complex ensemble methods~\cite{turian2022hear}.
-Such models are useful but intractable; furthermore, training them requires the definition of simpler loss functions, which emphasizes the need for further development of differentiable and expressive loss function implementations.
-
-Like nearly all previous work in sound-matching, the main measure of success here was the accurate \textit{replication} of sounds (or synth parameters) rather than imitation of sounds outside of the synthesizer's domain.
-Replication of arbitrary features of sounds is an important component of sound-design, though it is much harder to define or measure.
-Future works could explore loss functions which only measure certain characteristics of sound (such as \DTWEnv), and whether they can pave the way for better imitation in sound-matching.
-
-\bibliographystyle{IEEEtran}
-\bibliography{references}
-\end{document}
+% Section IV titled "Programs, Methodology, and Results" is overloaded and blends experimental setup with results and interpretation.
+For clear clarity and alignment with structural norms, we recommend splitting this into three sections: (1) Experimental Setup (including synthesizer programs, loss definitions, optimization procedure); (2) Experimental Results (performance metrics, statistical tests, and rankings); and (3) Discussion (qualitative interpretation, synthesis-specific observations, and practical implications).

### 2025-09-09 sections/introduction.tex
Updates from Overleaf

\section{Introduction}

% Since the 1970s, artists such as Suzanne Ciani~\cite{ciani_life_in_waves} have explored the use of audio synthesizers to replicate natural sounds—such as a bottle cap popping off—in ways that preserve the essential qualities of the original while showcasing the unique characteristics of the synthesizer~\cite{creativecherep2024}. Rather than merely replicating sounds, this approach uses synthesizers for artistic reinterpretation. However, the automation of this interpretive approach to sound design using digital audio synthesizers remains impractical for sound-designers, and significant knowledge gaps remain even in the simpler problem of replicating sounds. One such gap is the lack of diversity and depth in the analysis of sound-similarity measures and their interaction with various synthesis methods. As a stepping stone toward automated interpretive sound design, we use automatic and manual sound-evaluations to investigate whether the optimal choice of a sound-similarity measure is dependent on the synthesis method.

A digital audio synthesizer is any software used for the creation and manipulation of audio. The possibilities afforded by these synthesizers is infinite, leading to their widespread adaption by artists and sound-designers~\cite{lyons1997understanding,russ1999sound,stranneby2004digital}. A typical synthesizer has a number of parameterizable functions which affect the output sound in various ways, and manual sound-design often involves modifying the parameters until a conceptualized target sound is reached. The automation of this approach has commonly been referred to as "sound-matching", with reduction of tinkering time and creation of ``interesting" sounds as the main motivators ~\cite{krekovic2019insights,turian2020sorry,horner1993machine,salimi2020make,esling2019flow,engel2020ddsp,mitchell2007evolutionary,shier2020spiegelib,masuda2021soundmatch,masuda2023improving}. The main requirements of sound-matching are a target sound, a similarity metric (or loss function), and a heuristic to find parameters for a synthesizer that replicate \textit{all} or \textit{some} of the characteristics of the target sound as best as possible~\cite{horner1993machine,mitchell2007evolutionary,masuda2023improving}. However, despite the seeming simplicity of the problem and decades of research, sound-matching is not yet practical solution for sound-designers. Here we are motivated to understand why this is the case, and what future research should prioritize in order to improve the field.

We provide a review of past works in order to understand what is lacking, a nomenclature for different subtypes of sound-matching, and illuminate major perennial issues. In particular, we are concerned with the issues of \LossSelect~and \SynthSelect. The former refers to the search for an optimal loss function, lack of novel algorithms, and the---perhaps needless---focus on outperforming state of the art (\gls{SOTA}), while the latter refers to the lack of diversity in synthesis methods. As a consequence of these problems, we find that not much is known about the interaction between different methods of synthesis and different loss functions. Is there a universally best performing loss function for audio?

The main hypothesis here is that \textit{the performance of a similarity measure (or loss function) is influenced by other factors in the environment, particularly the method of synthesis}. While there have been many works and experiments comparing the accuracy of loss functions~\cite{vahidi2023mesostructures,turian2020sorry,engel2020ddsp,uzrad2024diffmoog,han2023perceptual,masuda2021soundmatch,turian2020sorry,bruford2024synthesizer}, we note that claims regarding the effectiveness of one function versus another have consistently been made in limited contexts that may not generalize to other settings. Beyond this central hypothesis, we also investigate three additional research questions: (Q1) To what extent do automatic evaluation metrics agree with manual listening tests? (Q2) Can novel loss functions based on Dynamic Time Warping (DTW) and Scale-Invariant Mean Squared Error (SIMSE) provide advantages over standard loss functions, and under what conditions?
(Q3) Is an iterative differentiable optimization a viable strategy for design of sound-matching experiments?

% This main hypothesis (along with our forthcoming research questions) would help with important issues regarding the path of future research. Should we be searching for a universally best performing loss function across synthesizers? Do current, SOTA loss functions satisfy the creative needs of sound-designers?

 We take an \textit{iterative} and \textit{differentiable} approach to defining our sound-matching experiments. The iterative approach better mimics the manual process of recursive listening and parameter adjustments towards sound design, while a differentiable environment allows access to the loss function gradients that can be used to better understand the nature of the problem. We define four differentiable synthesizers (each showcasing a fundamental method of synthesis used in modern synthesizers) and pair them with four different loss functions (two established methods, one utilizing DTW, and one utilizing SIMSE). We evaluate the final similarity with two different automatic methods, as well as manual listening tests conducted by two of the authors. Using the Non-parametric Scott-Knott tests~\cite{tantithamthavorn2017mvt,tantithamthavorn2018optimization}, we compare the distributions of the evaluation scores in order to rank the loss functions from best to worst.

The results of these experiments show that when accounting for different synthesizers, there is no consistent choice of sound similarity measure for optimal performance. Regarding Q1, We see a large overlap between automatic and manual evaluation methods, but continue to encourage manual listening tests as the gold standard for audio evaluation as extreme disagreements can occur. In the case of Q2, we confirm that \DTWEnv~outperforms other methods when envelope modulation plays a central role, and \SIMSESpec~is the best performer in simple subtractive synthesis. For Q3, we found the declaration of differentiable sound-matching experiments using Jax and transpiled Faust code to be a convenient and resource efficient, and and overall viable general strategy for sound-matching in differentiable settings. These results imply that rather than aiming to outperform the SOTA and finding a universal best performer, more effort could be placed towards diversifying the existing similarity measures of sound as well as conducting experiments using a variety of synthesis methods. Sound design is an inherently creative endeavor, and would likely benefit from a diversified range of solutions for different approaches.

% Here, several common characteristics of previous work are avoided in order to gain fundamental, practical takeaways from the experiments. In order to directly analyze the interaction between the loss and method of synthesis, the synthesizer here are implemented with simple, functions that are used as the building blocks of complex synthesizers, and the gradients flow from the loss function to the synthesizer parameters with no extra layers between the two. Furthermore, since the true target parameters are unknown in any realistic sound-design scenario, parameter loss (P-Loss) is only used as a secondary measure of performance, and not used as a loss function.

% The simple synthesizers tested here represent fundimental building blocks of more complex implementations, and direct use of loss function gradients assuming 
%   A primary conclusion of the work is that there is no one best sound similarity measure capable of working across all synthesizers.
% The core problem that no single loss function performs optimally across synthesizer types

### 2025-09-09 sections/setup_and_posthoc_analysis.tex
Updates from Overleaf

--- 2025-08-21 main.tex
@@
-\documentclass[lettersize,journal]{IEEEtran}
-\usepackage{amsmath,amsfonts}
-\usepackage{algorithmic}
-\usepackage{algorithm}
-\usepackage{array}
-\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-\usepackage{textcomp}
-\usepackage{stfloats}
-\usepackage{url}
-\usepackage{verbatim}
-\usepackage{graphicx}
-\usepackage{xcolor}
-\usepackage{csquotes}
-\usepackage{makecell}
-% my added packages and commands =======================
-\usepackage[numbers]{natbib} %removed cite package
-\usepackage{tikz}
-\usetikzlibrary{positioning,shapes.geometric}
-\usepackage{adjustbox}
-\usepackage{listings}
-\usepackage{caption}
-
-\newcommand{\highlight}[1]{\textcolor[RGB]{00,100,100}{#1}}
-\newcommand{\todo}[1]{\textcolor{red}{#1}}
-
-\captionsetup[lstlisting]{justification=centering, singlelinecheck=false}
-\providecommand{\gls}[1]{#1}
-\definecolor{rank1}{HTML}{70FF70}
-\definecolor{rank2}{HTML}{858585}
-\definecolor{rank3}{HTML}{454545}
-\definecolor{rank4}{HTML}{000000}
-\newcommand{\SIMSESpec}{\texttt{SIMSE\_Spec}}
-\newcommand{\LoneSpec}{\texttt{L1\_Spec}}
-\newcommand{\JTFS}{\texttt{JTFS}}
-\newcommand{\DTWEnv}{\texttt{DTW\_Envelope}}
-\newcommand{\LossSelect}{\textbf{Loss Selection}}
-\newcommand{\SynthSelect}{\textbf{Synthesis Selection}}
-\newcommand{\PeriodicLoss}{\textbf{Periodic Loss}}
-\newcommand{\OutDomain}{\textbf{Out-Domain Generation}}
-\newcommand{\BPNoise}{\textbf{BP-Noise}}  
-\newcommand{\AddSineSaw}{\textbf{Add-SineSaw}}  
-\newcommand{\AmpMod}{\textbf{Noise-AM}}  
-\newcommand{\FMMod}{\textbf{SineSaw-AM}}
-
-\input{code_style}
-
-%  ================================
-\begin{document}
-\title{Evaluating Sound Similarity Metrics for Differentiable, Iterative Sound-Matching}
-
-\author{Amir Salimi, Abram Hindle, Osmar R. Za{\"i}ane}
-
-% \markboth{IEEE TRANSACTIONS ON AUDIO, SPEECH AND LANGUAGE PROCESSING, VOL -, 2025}%
-% {Shell \MakeLowercase{\textit{et al.}}: A Sample Article Using IEEEtran.cls for IEEE Journals}
-
-\markboth{%
-  \parbox{\textwidth}{%
-    This work has been submitted to the IEEE for possible publication. \\
-    Copyright may be transferred without notice, after which this version may no longer be accessible.%
-  }%
-}{}
-
-\maketitle
-
-\begin{abstract}
-Manual sound design with a synthesizer is inherently iterative: an artist compares the synthesized output to a mental target, adjusts parameters, and repeats until satisfied.
-Iterative sound-matching automates this workflow by continually programming a synthesizer under the guidance of a loss function (or similarity measure) towards a target sound.
-Prior comparisons of loss functions have typically favored one metric over another, but only within narrow settings: limited synthesis methods, few loss types, often without blind listening tests.
-This leaves open the question of whether a universally optimal loss exists, or the choice of loss remains a creative decision conditioned on the synthesis method and the sound designer's preference.
-We propose differentiable iterative sound-matching as the natural extension of the available literature, since it combines the manual approach to sound design with modern advances in machine learning.
-To analyze the variability of loss function performance across synthesizers, we implemented a mix of four novel and established differentiable loss functions, and paired them with differentiable subtractive, additive, and AM synthesizers.
-For each of the sixteen synthesizer–loss combinations, we ran 300 randomized sound-matching trials.
-Performance was measured using parameter differences, spectrogram-distance metrics, and manually assigned listening scores.
-We observed a moderate level of consistency among the three performance measures.
-Our post-hoc analysis shows that the loss function performance is highly dependent on the synthesizer.
-These findings underscore the value of expanding the scope of sound-matching experiments, and developing new similarity metrics tailored to specific synthesis techniques, rather than pursuing one-size-fits-all solutions.
-\end{abstract}
-
-% the automatic performance measures have not been compared to human hearing
-% lack of creativity in loss function definitions 
-\begin{IEEEkeywords}
-Audio synthesis, differentiable digital signal processing, music information retrieval, sound-matching
-\end{IEEEkeywords}
-
-\section{Introduction}
-Since the 1970s, artists such as Suzanne Ciani~\cite{ciani_life_in_waves} have explored the use of audio synthesizers to replicate natural sounds—such as a bottle cap popping off—in ways that preserve the essential qualities of the original while showcasing the unique characteristics of the synthesizer.
-Rather than merely replicating sounds, this approach uses synthesizers for artistic reinterpretation.
-However, the automation of this interpretive approach to sound design using digital audio synthesizers has rarely been explored, and significant knowledge gaps remain even in the simpler problem of replicating sounds.
-One such gap is the lack of diversity and depth in the analysis of sound-similarity measures and their interaction with various synthesis methods.
-As a stepping stone toward automated interpretive sound design, we use automatic and manual sound-evaluations to investigate whether the optimal choice of a sound-similarity measure is dependent on the synthesis method.
-
-A digital audio synthesizer is any software used for the creation and manipulation of audio.
-Since the advent of digital signal processing (\gls{DSP}) in the 1960s~\cite{stranneby2004digital}, a wide variety of parametric audio generation functions have been proposed~\cite{lyons1997understanding,russ1999sound,shier2020spiegelib}.
-Manual sound design with a synthesizer is an iterative process of trial and error; when the artist has a desired (or target) sound in mind, sound design involves listening to the output, determining whether the output is closer or farther from the target, and adjusting the parameters of the synthesizer accordingly until the target is reached~\cite{russ1999sound}.
-This paper explores the automation of this process, which we call \textit{iterative sound-matching}.
-
-Automatic sound design has been a subject of research for decades, with reduction of tinkering time and creation of ``interesting" sounds as the main motivators \cite{krekovic2019insights,turian2020sorry,horner1993machine,salimi2020make,esling2019flow,engel2020ddsp,mitchell2007evolutionary,shier2020spiegelib,masuda2021soundmatch,masuda2023improving}.
-\textit{Sound-matching} is a common form of automatic sound design explored in literature, where a similarity metric (or loss function) is given, and the goal is to find parameters for a synthesizer that replicate \textit{all} or \textit{some} of the characteristics of a target sound as best as possible~\cite{horner1993machine,mitchell2007evolutionary,masuda2023improving}.
-Our analysis of past works shows that this general goal of ``sound-matching'' can be broken down in to very different subtypes, many of them rarely or never explored, despite their relevance to practical applications.
-This leads us to a typology of sound-matching tasks, and the importance of \textit{differentiable, iterative sound-matching}, rarely explored in previous literature~\cite{vahidi2023mesostructures}.
-
-\textit{Iterative sound-matching} refers to works that mimic the back-and-forth process of listening to a synthesizer's output and updating the parameters until the termination conditions are met.
-The two necessary components are the \textit{loss function}, which determines how close two sounds are, and the \textit{parametric synthesizer} that outputs sounds based on input parameters.
-In the iterative workflow, a target sound is selected, the synthesizer outputs sounds based on its parameters, and the loss function determines the proximity of the output to the target.
-From there, the synthesizer parameters are updated, the output and target are compared, and so on.
-
-Despite the early popularity of the iterative approach~\cite{justice1979analytic,horner1993machine,mitchell2007evolutionary}, recent works in sound-matching have favored supervised learning using inference models, deep neural-networks (\gls{DNNs}), and large datasets.
-Rather than continually interacting with the environment, these models learn from a dataset of inputs (synthesizer parameters) and outputs (corresponding sound), and predict the synthesizer parameters for new sounds~\cite{yee2018automatic,han2023perceptual,han2024learning,masuda2021soundmatch,masuda2023improving}.
-
-Iterative sound-matching is interesting for multiple reasons.
-First, it better resembles the human approach to sound-design, which requires continually interacting with the environment and making decisions according to a feedback signal~\cite{sutton2018reinforcement}.
-Second, the trial-and-error approach leaves a trail of parameter sets from the starting point to the end goal, which could be of use to sound designers.
-Third, relative to the DNN-based inference models, it provides a more tractable environment with room for creative approaches to sound-similarity definitions, and more amenable to deployment on different synthesizers as it does not require pre-training on large datasets.
-
-This work takes a \textit{differentiable} approach to iterative sound-matching, a natural extension of early works in iterative sound-matching combined with recent advances in Machine Learning (\gls{ML}) software~\cite{braun2024dac,paszke2019pytorch} and increase in computational power~\cite{moore2012introduction,schaller1997moore}.
-A differentiable approach requires differentiable loss and synthesizer functions, allowing the synthesizer parameters to be updated using gradient descent~\cite{goodfellow2016deep,boyd2004convex}.
-
-We discuss the history of sound-matching, and provide a nomenclature for different subtypes of sound-matching tasks.
-A number of weaknesses in current literature (see Section~\ref{sec:lacking}) are identified.
-The experiments in this work focus on two of these core issues, referred to as \LossSelect~and \SynthSelect.
-The former refers to the search for an optimal loss function, lack of novel algorithms, and the---perhaps needless---focus on outperforming state of the art (\gls{SOTA}), while the latter refers to the lack of diversity in synthesis methods.
-
-In the case of \LossSelect, there have been many works and experiments comparing the accuracy of loss functions and choosing the best performers~\cite{vahidi2023mesostructures,turian2020sorry,engel2020ddsp,uzrad2024diffmoog,han2023perceptual,masuda2021soundmatch,turian2020sorry,bruford2024synthesizer}.
-We note that these claims regarding the effectiveness of one function versus another have consistently been made in limited contexts that may not generalize to other settings.
-One such limitation is the synthesizers used, and ties into the problem of \SynthSelect, as typically only a single (and often, simple) synthesizer has been used for testing the effectiveness of loss functions.
-Furthermore, there is a lack of diversity in the underlying functions used by the synthesizers, with simple Frequency Modulation (\gls{FM}) being the most common method of synthesis~\cite{horner1993machine,vahidi2023mesostructures,mitchell2007evolutionary,caspe2023envelope}.
-
-Considering the interplay between the two problems, we design experiments that address them simultaneously.
-Testing the hypothesis that the performance of a similarity measure (or loss function) is influenced by other factors in the environment, particularly the method of synthesis.
-We combine 4 loss functions with 4 differentiable synthesizers, and compare their performance as measured by the similarity of the final output to the target.
-We evaluate the final similarity with two different automatic methods, as well as manual listening tests conducted by two of the authors.
-We compare the distributions of the evaluation scores in order to rank the loss functions from best to worst.
-The results of these experiments show that when accounting for different synthesizers, there is no consistent choice of sound similarity measure for optimal performance in sound-matching.
-
-We note that for some synthesizers, our novel use of ``dynamic time warping of envelopes'' as a loss function for differentiable sound-matching outperforms the standard methods of audio comparison.
-These results imply that rather than aiming to outperform the SOTA and finding a universal best performer, more effort could be placed towards diversifying the existing similarity measures of sound as well as conducting experiments using a variety of synthesis methods.
-Sound design is an inherently creative endeavor, and would likely benefit from a diversified range of solutions for different approaches.
-
-\section{Background}
-\label{sec:background}
-In this Section, we give a formal definition of sound-matching and the necessary components.
-These discussions lead to a nomenclature for types of sound-matching, as well as major perennial issues in the field.
-
-\subsection{Digital Audio Synthesis}
-\label{sec:diffSynth}
-A digital audio synthesizer is any software used for the creation and manipulation of digital audio.
-Digital synthesizers use signal processing chains with a variety of digital signal processing (\gls{DSP}) functions to create sounds.
-These functions are often parametric, and the set of parameters given to the synthesizer's chain of DSP functions is called a synthesizer program.
-
-The simplest form of DSP function is a sinusoidal tone. For example, consider:
-\[ x(n) = sin( 2\pi n T)\]
-
-where $T$ represents time in seconds, and $n$ is the frequency parameter.
-If $n$ is 1, then $x(n)$ would be a waveform with a frequency of 1 hertz, meaning that the value of $x(n)$ oscillates to the same point once every second.
-Waveform generators are called oscillators for this reason.
-
-% If $n$ is in the range of 30 to 20,000 (the upper limit is often lower for some), $x(n)$ could be played as an audible tone~\cite{smith2007mathematics,jilek2014reference}.
-
-Since the advent of digital signal processing in the 1960s~\cite{stranneby2004digital}, a wide variety of parametric audio generation functions have been proposed~\cite{lyons1997understanding,russ1999sound,shier2020spiegelib}.
-This includes oscillators, filters, equalizers, and envelopes, which can be used sequentially or in parallel~\cite{lyons1997understanding,russ1999sound}.
-Sound design with a synthesizer is done by modifying the parameters of the functions until reaching a desired output~\cite{roads1996computer,pinch2004analog}.
-
-\todo{A relatively small subset of these synthesis methods have been used for automatic sound-matching, notably additive and/or subtractive synthesizers~\cite{esling2019flow,yee2018automatic,mitchell2007evolutionary,salimi2020make}, physical modeling~\cite{riionheimo2003parameter}, and frequency/amplitude modulation (\gls{FM}/\gls{AM})~\cite{horner1993machine,vahidi2023mesostructures}.
-}
-
-Differential DSP (\gls{DDSP}) is a general term for applications that combine machine learning techniques such as gradient descent~\cite{goodfellow2016deep,boyd2004convex} with DSP~\cite{engel2020ddsp}.
-Implementing complex DSP functions in a differentiable manner can be challenging, and effective differentiable audio similarity measures require careful mathematical expression of the desired attributes of sound.
-These issues have likely contributed to the limited exploration of this domain.
-Perhaps the most relevant recent work is by Vahidi \textit{et al.}, featuring the comparison of two differentiable loss functions with a simple FM synthesizer~\cite{vahidi2023mesostructures}.
-
-\subsection{Sound Representation and Loss Functions}
-\label{sec:loss_funcs}
-A digital sound (or an audio signal) is a series of numbers~\cite{smith1991viewpoints,smith2007mathematics}.
-To compare two digital sounds, the two corresponding series are passed to a function that measures their similarity.
-Two signals can sound identical to our ears, without having any values in common~\cite{moore2012introduction}.
-This necessitates the use of proxy representations (or feature extractors) when comparing sounds automatically.
-Similarity between the target sound and the synthesizer output is then measured by some form of subtraction and summation of the proxy representations.
-
-In sound-matching, particularly in a Deep Learning (\gls{DL}) context~\cite{goodfellow2016deep}, the similarity function can also be called a \textit{loss} function, where the emphasis is on measurement and reduction of the distance between target and output.
-It is important to note that there is a close relationship between the loss function $L$ and the sound representation function $\phi$.
-$L$ is the result of a distance measure $d$ applied to the features extracted by $\phi$.
-
-\[
-L(\theta, t) = d\langle\phi(t),\phi(x)\rangle
-\]
-
-\noindent
-
-A proxy representation is the output of the function \( \phi \), which can be thought of as a feature extraction function that maps the sounds \( t \) and \( x \) to their respective representations. 
-The proportionality or distance metric $d$ has typically been the L1 distance~\cite{turian2020sorry,richard2025model}, calculated as the mean of the absolute difference between every point in the proxy representation~\cite{engel2020ddsp,vahidi2023mesostructures}:
-\[
-L(\theta, t) = \left\| \phi(t) - \phi(x) \right\|_1
-\]
-
-Here we discuss four methods of audio representation and the corresponding loss functions.
-Three of which have been used in previous works, and a novel method not previously used in iterative sound-matching.
-
-\subsubsection{Parameter Loss}
-A common measure of similarity in sound-matching is the distance between synthesizer parameter sets, referred to as ``P-Loss"~\cite{han2023perceptual}.
-Typically, for the implementation of P-Loss the parameter sets are treated as vectors in space, and L1 or L2 distance is applied.
-There are two major limitations to this approach: First, the target and output sound must be made by the same synthesizer; otherwise the parameter sets cannot be compared (see Section~\ref{sec:matching_types}).
-Second, the relationship between synthesizer parameters and the audio output is not linear~\cite{shier2020spiegelib,han2023perceptual,esling2019flow}.
-
-\subsubsection{Fourier Spectrograms}
-\label{sec:fourier_specs}
-Fourier-based transformations such as short-time Fourier transforms (\gls{STFT}), Mel-spectrograms, and Mel-frequency cepstral coefficients have been viewed as the de facto and state-of-the-art representation of audio~\cite{beauchamp2003error,mitchell2007evolutionary,yee2018automatic}, however, there are many issues associated with their use in sound-matching~\cite{turian2020sorry,vahidi2023mesostructures,han2023perceptual,uzrad2024diffmoog}.
-Fourier transformations allow for the conversion of a signal from the time-domain to the frequency domain.
-Audio spectrograms can be generated by segmentation of a piece of audio into overlapping windows followed by the application of Fourier transforms to each window.
-They are costly to compute, but provide a better temporal view of changes in frequency content~\cite{muller2007dynamic,smith2007mathematics}.
-There are different types of spectrograms that have a basis in Fourier transformations, but the most notable and commonly used is the STFT.
-
-What we call \textit{Fourier-based Spectrograms} are variations on the STFT approach.
-For example, Mel-Spectrograms \textit{bin} frequencies on a near-logarithmic scale to better match human perception of frequencies~\cite{muller2007dynamic}.
-Multi-scale spectrograms (\gls{MSS}) used in recent works are a simple weighted average of multiple spectrograms with different parameters such as window size, number of frequency bins, and hop size~\cite{engel2020ddsp,vahidi2023mesostructures}; this may provide some improvements at a higher computational cost~\cite{turian2020sorry,engel2020ddsp}.
-
-\subsubsection{Joint-Time Frequency Spectrum}
-Recent works have focused on the limitations of parameter and spectral loss functions in sound-matching~\cite{vahidi2023mesostructures,uzrad2024diffmoog}, seeking to create more effective general solutions for the comparison of audio.
-Noting the aforementioned weaknesses of comparing STFT spectrograms, Vahidi \textit{et al.} proposed differentiable Joint-Time Frequency Scattering (\gls{JTFS})~\cite{anden2015joint} as an alternative to spectrogram loss in sound-matching, and showed improved performance in sound-matching with differentiable chirplet synthesizers~\cite{vahidi2023mesostructures}.
-JTFS is the result of the application of a 2D wavelet transformation to the time-frequency representation of a signal~\cite{anden2015joint}.
-
-\subsubsection{Dynamic Envelope Warping}
-Dynamic Time Warping (DTW) is a method for measuring similarity between multi-dimensional time-series~\cite{rabiner1993fundamentals,muller2007dynamic,giorgino2009computing}.
-Given any two time-series $X = \{x_1,x_2,...,x_m\}$ and $Y = \{y_1,y_2,...,y_n\}$, we have indices $i\in\{1...m\}$ and $j\in\{1...n\}$ defining $X$ and $Y$.
-When the series are \textit{warped}, these indices change to expand or contract different portions of the series.
-To borrow the notation given by Muller~\cite{muller2007dynamic}, warped indices are a sequence $p=(p_1,...,p_L)$, where \(p_\ell = (m_\ell, n_\ell) \in [1 : m] \times [1 : n] \text{ for } \ell \in [1 : L]\), meaning that the indices for $X$ and $Y$ are reorganized under special conditions.
-In classical DTW, these conditions are \textit{monotonicity}, \textit{boundary matching}, and \textit{single step-size}.
-
-DTW measures the distance between the time-series \textit{after} alignments, typically using Euclidean distance.
-The distance between a time-series and shifted versions of itself would be 0, regardless of shift amount~\cite{tavenard.blog.dtw}.
-Additional rules can be imposed to keep alignments locally constrained~\cite{itakura1975minimum,sakoe1978dynamic}.
-% \textcolor{red}{DTW was originally used for signal comparison but using it the way we do is novel}
-
-\subsection{Automatic Sound Matching}
-\label{sec:sound_matching_definition}
-The problem of sound-matching has been tackled from a variety of perspectives as it involves a number of modular parts: the choice of synthesizer $g$, the target sounds of interest, the representation function $\phi$, and the heuristic for finding the optimal $\theta$.
-To give a formal definition of sound-matching, we expand the definition given by recent works~\cite{vahidi2023mesostructures,han2023perceptual}.
-The major components are as follows:
-\begin{itemize}
-    \item $g(\theta)$: Parametric audio synthesizer $g$ which takes on parameters $\theta$ 
-    \item $x$: The output of $g$, given a set of parameters $\theta$ or $g(\theta) = x$ 
-    \item $t$: The target sound which we want to replicate or imitate. 
-    \item $\phi$: Representation function or feature extractor. $\phi$ is applied to $x$ and $t$ to facilitate their comparison by the loss function.
-    \item $L$: Loss or error function. $L(\theta,t)$ is a measure of distance between $x$ and $t$. Often proportional to the subtraction of their representations, or $ \phi(x) - \phi(t)$
-\end{itemize}
-
-\highlight{It could be argued that the implementation of these components is not an optimization problem with a correct answer, but a \textit{creative} endeavor depending on the artist's needs}.
-Consider the hypothetical case where the goal might be to create an 8-bit~\cite{collins2007loop} version of a high quality, organic snare drum: in such a case, there is no optimal or ``correct" version of an 8-bit snare sound.
-
-\subsection{Sound Matching Types}
-\label{sec:matching_types}
-
-Sound-matching tasks can be categorized based on the choices of target sound $t$, the loss $L$ and representation function $\phi$, and the heuristic used for minimizing $\theta$.
-We will discuss the subtypes of sound-matching tasks in the upcoming sub-sections.
-
-\subsubsection{In-Domain vs Out-of-Domain}
-\label{sec:in-domain}
-The choice of domain depends on whether we want to use the same synthesizer for the target and output sounds, which we call \textit{in-domain}, or have target sounds that came from sources other than the synthesizer, or \textit{out-of-domain}.
-To paraphrase the description given by Masuda \textit{et al.}~\cite{masuda2021soundmatch}, if $g$, the synthesizer of choice, can accurately replicate the target sound $t$, or put differently, if $t$ itself is an output of $g$, then the sound-matching task is \textit{in-domain}.
-If $t$ is not an output of $g$, then the sound-matching task is \textit{out-of-domain}.
-In-domain tasks in general are simpler, and often there is a guarantee that there is a correct answer to the sound-matching problem, particularly if the goal is accurate replication of the sound.
-If the target sound is out-of-domain, replication is not guaranteed, and the goal becomes the \textit{imitation} of some aspect of sound.
-
-\subsubsection{Replication vs Imitation}
-Regardless of the domain, the generation goal can be \textit{replication} or \textit{imitation} of the target sound.
-In replication, the goal is to make an identical copy of the target sound.
-Imitation is an artistic pursuit and harder to define, since the goal is to make new sounds that only retain a subset of the target's sonic features.
-While closely related to the in-domain versus out-of-domain problem, the choice of replication versus imitation is more dependent on how the loss and representation functions are defined.
-
-% From a signal perspective, replication is a scenario where the goal is to find a $\theta$ such that $t-x=0$ where $x=g(\theta)$.
-From a sound perspective, this definition of replication is unnecessarily restrictive, as $t$ and $x$ can sound identical but contain very different values.
-For example, if $t$ and $x$ were periodic waveforms at an audible frequency $f$ with different phase values, their subtraction would not be 0, yet they would sound identical to our ears, as our ears are insensitive to phase values~\cite{moore2012introduction}.
-Representation functions such as STFT spectrograms or Mel-frequency cepstral coefficients (\gls{MFCC}) become viable in these cases, as they allow for separation and removal of the phase information within an audio signal.
-The \textit{sound replication} goal can be defined, in layman's terms, as ``$x$ sounds exactly like $t$".
-More formally, the goal is to find parameters such that $\phi(t)-\phi(x)=0$, where $\phi$ transforms an audio into a spectrogram that removes the phase information from sounds, but otherwise represents the sonic information as accurately as possible.
-
-% With \textit{sound imitation}, the goal is the replication of some aspect of the target sound.
-Instrument conversion is a common example of imitation.
-As an example, a note played by a violin is converted to the same note played by a piano.
-Conversion of even longer pieces has been made possible with large enough datasets~\cite{engel2020ddsp}.
-Imitation of sound envelopes (how the loudness changes over time) is another interesting idea~\cite{caspe2023envelope}.
-Transfer of other sound properties such as reverb and delay has also been a subject of research~\cite{engel2020ddsp}.
-
-% A target sound can have many properties of interest, and the imitation task is the transfer of a subset of those properties to other sounds.
-Relative to replication, imitation is a very open ended goal.
-From a research point of view, success in the replication task can be measured more reliably, often without the need for manual evaluation.
-With imitation, listening tests are needed to ensure quality of outputs, particularly in a supervised learning setting where learning is done by examples~\cite{salimi2020make,masuda2021soundmatch}.
-The need for subjective evaluations has perhaps contributed to the lack of research in imitation and out-of-domain sound-matching.
-
-\subsubsection{Supervised Versus Direct Optimization}
-\label{sec:optimization}
-
-The goal of sound-matching is to find the optimal parameters $\theta^*$ that minimize the loss between the synthesizer output and the target sound. 
-\[
-\theta^* = \arg\min_{\theta} L(\theta,t)
-\]
-
-The heuristics in past works can be broadly split into two categories:
-\begin{enumerate}
-    \item Supervised (or inference) methods which use datasets of sound and parameter pairs to describe the generation objective, often with a large number of input and output examples~\cite{engel2020ddsp,salimi2020make,yee2018automatic,esling2019flow}
-    \item Unsupervised or \textit{direct sound optimization} methods~\cite{horner1993machine,mitchell2007evolutionary,yee2018automatic,vahidi2023mesostructures}.
-Usually require at most one target example, and a loss function that measures the difference between the target and synthesizer outputs.
-\end{enumerate}
-
-\todo{
-Genetic algorithms (\gls{GA})~\cite{holland1992genetic} can be used for direct sound optimization, and have been the earliest and most common heuristic in sound-matching~\cite{horner1993machine,mitchell2007evolutionary,yee2018automatic}.
-These algorithms start with arbitrary parameter sets that can be treated as an evolving population where the genomes are the parameter values.
-The most fit members of the group are the parameters which perform the best in the loss function, and create a new generation of parameters via mutation (random change in subset of parameters) and crossovers (combination of parameter sets); this process repeats until the goal or a maximum number of generations is reached.}
-
- Differentiable settings have the benefit of making goal-oriented updates to the synthesizer parameters (the goal being the minimization of loss).
-There are also drawbacks: other than requiring more computation power, differentiable functions require careful implementation of operations that are continuous and numerically stable; this makes the implementation of signal processing functions quite difficult, contributing to the scarcity of works in this domain.
-
-% One important aspect of the heuristics is whether they imitate the iterative nature of manual sound design---that is, listening to the output and adjusting the synthesizer parameters recursively until the desired goal is reached.
-Supervised methods generate their final output in a single step, while direct sound optimization methods refine their outputs iteratively until termination, which resembles manual sound design more closely.
-
-\subsection{Iterative Sound Matching}
-\label{sec:iterative_sound_matching}
- \todo{Figure~\ref{fig:sound_design_loop_iterative} shows a general model for iterative sound-matching.
-To begin the process, the parameters are arbitrarily initialized (usually random generation), the similarity of the target and output is measured, and the parameter is \textit{optimized} with the goal of increasing the similarity, or alternatively, reducing the loss.
-This process repeats until termination.
-}
-\begin{figure}[ht]
-    \centering
-\begin{tikzpicture}[node distance=2cm, auto]
-
-    % Nodes
-    \node (start) [text centered] {\( g (\hat{\theta}) = x \)};
-    \node (L) [above of=start,right of=start, text centered] {\( L(\hat{\theta},t) {\ \propto \ } \phi(x) - \phi(t) \)};
-    \node (optimize) [below of=L,right of=L, text centered] {Optimize $\hat{\theta}$};
-    \node (new_theta) [below of=optimize,right of = start, text centered] {New \( \hat{\theta} \)};
-
-    % Highlight and arrow for the start node
-    \draw[->, very thick, red] (start) ++(-2,1.5) -- (start)
-        node[midway, below, align=center, sloped, color=red] {Start}
-        node[midway, above, align=center, sloped, color=red] {Random $\hat{\theta}$};
-
-    % Arrows with multi-line labels
-    \draw[->, bend left] (start) to node[midway, right, align=center] {} (L);
-    \draw[->, bend left] (L) to node[midway, right, align=center] {} (optimize);
-    \draw[->, bend left] (optimize) to node[midway, below, align=center] {} (new_theta);
-    \draw[->, bend left] (new_theta) to node[midway, left, align=center] {} (start);
-
-\end{tikzpicture}
-    % \caption{ Iterative approach to sound design.
-Synthesizer $g$ with arbitrary initialized parameters $\hat{\theta}$, creates sound $x$.
-The target sound $t$ is used as the goal.
-Parameters $\hat{\theta}$ are adjusted to minimize error (or loss) $L(\hat{\theta},t)$, where $L$ is proportional to the difference between the representations of $x$ and $t$.
-$\phi$ is the audio feature extractor, or representation function.}
-    \caption{ Iterative approach to sound design.}
-    \label{fig:sound_design_loop_iterative}
-\end{figure}
-
-In the upcoming ``Past Works'' section, we will discuss the evolution of prior sound-matching research, with \textit{differentiable iterative sound-matching} being the latest extension of the field.
-
-\begin{table*}[t]
-\centering
-\caption{Summary of select works in sound-matching.
-Gen.
-refers to whether the generations are done in 1-shot (generally, output by a neural network) or a search is conducted with the goal of iteratively getting closer to the target sound.}
-\setlength{\tabcolsep}{3pt} % Reduce horizontal spacing
-\renewcommand{\arraystretch}{1.05} % Reduce vertical spacing
-\scriptsize % Reduce font size
-\resizebox{\textwidth}{!}{
-\begin{tabular}{|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{1.2cm}|>{\raggedright\arraybackslash}m{1.2cm}|>{\raggedright\arraybackslash}m{1cm}|>{\raggedright\arraybackslash}m{1cm}|}
-\hline
-\textbf{Work} & \textbf{Synthesis} & \textbf{Loss} & \textbf{Heuristics} & \textbf{Goal} & \textbf{Domain} & \textbf{Year} & \textbf{Gen. } \\
-\hline
-Horner \textit{\textit{et al.}}~\cite{horner1993machine} & Non-diff FM & STFT (McAulay-Quatieri) & Direct & Replicate & In & 1991 & Iter \\
-\hline
-Mitchel \textit{\textit{et al.}}~\cite{mitchell2007evolutionary} & Non-diff FM & Spectral Relative & Evolutionary & Replicate & In (Contrived) & 2007 & Iter \\
-\hline
-Yee-King \textit{\textit{et al.}}~\cite{yee2018automatic} & Non-diff VST & P-Loss (for Supervised), Spectral (MFCC) & Supervised \& Evolutionary & Replicate & In & 2018 & Both \\
-\hline
-Esling \textit{\textit{et al.}}~\cite{esling2019flow} & Non-diff VST & Spec MSS/SC P-Loss & Supervised \& Modeling & Replicate & In* \& Out & 2019 & 1-shot \\
-\hline
-Shier \textit{\textit{et al.}}~\cite{shier2020spiegelib} & Custom, non-diff & FFT \& Spectral & Supervised \& Direct & Replicate & In & 2020 & Both \\
-\hline
-Salimi \textit{\textit{et al.}}~\cite{salimi2020make} & Non-diff & STFT and Envelope & Supervised & Imitate & Out & 2020 & 1-shot \\
-\hline
-Masuda \textit{\textit{et al.}}~\cite{masuda2021soundmatch} & DDSP & P-Loss initially, fine-tune with Spec. & Supervised & Both & In \& Out & 2021 & 1-shot \\
-\hline
-Han \textit{\textit{et al.}}~\cite{han2023perceptual} & NN-encoder $\rightarrow$ physical model & PNP (approx. of STFT) & Supervised & Replicate & In & 2023 & 1-shot \\
-\hline
-Vahidi \textit{\textit{et al.}}~\cite{vahidi2023mesostructures} & Differentiable FM & JTFS/MSS & Direct (grad desc.) & Replicate & In & 2023 & Iter \\
-\hline
-Uzrad \textit{\textit{et al.}}~\cite{uzrad2024diffmoog} & DDSP chains & Synth. Chain and P-Loss & Supervised & Replicate & In \& Out & 2024 & 1-shot \\
-\hline
-\end{tabular}
-}
-
-\label{tab:summary}
-\end{table*}
-
-\section{Past Works}
-\label{sec:related_works}
-Table~\ref{tab:summary} contains an overview of relevant literature, covering important works in sound-matching.
-adsfEarlier works in this domain focused on FM synthesis and genetic algorithms (or direct optimization methods)\cite{justice1979analytic,horner1993machine,mitchell2007evolutionary}.
-Along with the exponential increase in computation power, later works utilized a wider range of approaches to parameter search, estimation, and audio synthesis; particularly with supervised learning and neural networks (\gls{NN})~\cite{yee2018automatic,engel2020ddsp,esling2019flow}.
-
-\subsection{Earlier Works}
-FM synthesizers are simple to implement, yet very expressive~\cite{chowning1973synthesis}.
-Perhaps the earliest foundational work in sound-matching is the analytical approach by Justice~\cite{justice1979analytic} toward the decomposition and recreation of sounds with FM synthesis using one carrier, one modulator, and the corresponding envelopes:
-\[ f(t) = I_c(t) \cos(\omega_c t + I_m(t) \cos(\omega_m t))
-\]
-Where $f(t)$ is the output signal, $I_c(t)$ is the carrier amplitude, $\omega_c$ is the carrier frequency, and $I_m(t)$ and $\omega_m$ are the envelope and frequency for the modulator~\cite{justice1979analytic}.
-
- The sound-matching works that followed used other heuristics for finding the synthesizer parameters, yet the simple FM synthesizer structure remained unchanged.
-Horner \textit{et al.}~\cite{horner1993machine} used GAs for re-synthesis of sounds with an FM synthesizer using one modulator and one carrier oscillator, and the McAulay-Quatieri method~\cite{mcaulay1986speech} for measuring loss.
-
- % MQ calculates the STFT and marks the start and end of \textit{trends} in a sound with the assumption that the harmonic content belongs to trends that have a beginning and an end.
-
-Introduction of more complex models of synthesis such as wavetables~\cite{horner2003auto} and physical modeling~\cite{riionheimo2003parameter} rendered interesting results, but led to further questions about the nature and goals of sound-matching.
-Mitchell and Creasy noted the difficulty in distinguishing between ``inefficiency of the optimization engine'' and ``synthesizer limitations'' as the cause of failure or success in sound-matching~\cite{mitchell2007evolutionary}.
-They offered a \textit{contrived methodology} for finding the best evolutionary method for sound-matching using their FM synthesizer.
-The methodology postulates that the best search heuristic for in-domain search (where an exact target exists, and a wide range of targets can be produced by sampling from different points in the parameter space) would also be the best-performing heuristic for out-of-domain search on a dataset of muted trumpet tones~\cite{opolko1989mcgill}; however, testing was not extensive and yielded some contradictory results.
-As they noted, modifications to the synthesizer, loss function, and sound domain would lead to a problem with an entirely new search space~\cite{mitchell2007evolutionary}.
-
-% Testing sound-matching experiments can be difficult, in part due to the large number of parametric components and navigation of complex non-linear systems.
-The issue with finding the best approaches to sound-matching is exacerbated when slight changes to the pipeline render previous results moot.
-Perhaps due to these reasons, the contrived methodology was not extensively tested by Mitchell and Creasy, who only provided a single experiment where the best evolutionary strategy found by in-domain search would also perform well in out-of-domain synthesis.
-\subsection{Introduction of Deep-Learning}
-Recent years have seen more works in sound-matching using machine learning techniques.
-In 2018, Yee-King \textit{et al.} rendered 60,000 audio-parameter pairs from the \textit{Dexed} \gls{VST} synthesizer~\footnote{https://asb2m10.github.io/dexed/}, and showed that NNs trained on this dataset can outperform GA and hill-climber (\gls{HC})~\cite{hoffmann2000heuristic} methods in rendering speed, with slight improvements in MFCC error (used as an objective performance test).
-The speed improvement appears trivial, considering the iterative nature of GAs when compared to offline training of supervised models.
-For GAs and HC optimizer, MFCCs were used as a measure of performance as well as a loss function.
-For training the networks, P-Loss was used, since differentiable MFCCs were not possible in their pipeline.
-Importantly, informal hearing tests revealed that the performance of even the best NN model was unsatisfactory, possibly due to the complex nature of the synthesizer, which features 155 parameters.
-
-% Programming synthesizers and querying sounds by imitating human vocals has also been a subject of interest~\cite{Cartwright2014SynthAssistQA,siamesevocalimitation2018}.
-For example, Cartwright and Pardo created a framework for searching pre-made patches for a synthesizer to best imitate vocals using audio features such as pitch, spectral centroid, and loudness~\cite{Cartwright2014SynthAssistQA}.
-
-Esling \textit{et al.} noted the issue of using P-Loss in sound-matching~\cite{esling2019flow}, considering the non-linear mapping between parameters of a synth and the audio output.
-In their approach, dubbed as ``flowsynth'', they used a large parameter-to-audio dataset for a commercial \gls{VST} synthesizer.
-In flowsynth, audio is encoded to an auditory space, using a NN autoencoding approach~\cite{hinton2006reducing}.
-The encodings in the auditory space can then be converted to the parameter space, but only for that particular synthesizer.
-Using P-Loss and Spectral Convergence (Frobenius norm of the spectrograms) as metrics, flowsynth outperformed other NN approaches that did not use an auditory to parameter space conversion, but formal hearing tests were not conducted.
-
-Masuda \textit{et al.} also applied supervised learning to synthesizer parameter estimation~\cite{masuda2021soundmatch}.
-Their work highlights the issue of non-linearity in parameter-to-synthesizer outputs and out-of-domain search.
-This work uses a differentiable subtractive synthesizer (two oscillators and an LP filter).
-A NN model was pre-trained using an in-domain dataset of randomly selected parameters and P-Loss.
-After training, the model was fine-tuned using 20,000 out-of-domain sounds from the NSynth dataset~\cite{engel2017neural} and multi-scale spectrogram loss~\cite{engel2020ddsp}.
-This approach proved more effective---i.e, achieved lower multi-level spectral difference in out-of-domain tests---than baseline models, which were either not exposed to out-of-domain sounds or trained exclusively with P-Loss.
-Subjective hearing tests were conducted, showing a preference for the fine-tuned model~\cite{masuda2021soundmatch}.
-Masuda \textit{et al.} later extended this work with semi-supervised learning, highlighting significant gaps in in-domain and out-of-domain performance~\cite{masuda2023improving}.
-
-Rather than focusing on a particular implementation, Shier \textit{et al.} presented Spieglib, a library for implementation of sound-matching pipelines~\cite{shier2020spiegelib}.
-This library provides different choices for DNNs, GAs, synthesizers, and feature extractors.
-Shier \textit{et al.} presented an experiment with a similar setup to Yee-King \textit{et al.}~\cite{yee2018automatic}, however, they found a genetic algorithm to be the best performing.
-
-% Rather than focusing on a particular implementation, Shier \textit{et al.} presented Spieglib, a library for implementation of sound-matching pipelines (which Shier \textit{et al.} referred to as ``automatic parameter search'')~\cite{shier2020spiegelib}.
-This library provides extensible choices of deep-learning estimators with \textit{tensorflow}~\cite{abadi2016tensorflow}, genetic estimators with \textit{DEAP}~\cite{DEAP_JMLR2012}, feature extractors with \textit{librosa}~\cite{mcfee2015librosa}, and different choices for synthesizers, evaluators, dataloaders.
-Shier \textit{et al.} provide an example experiment using this library.
-This example has a setup similar to Yee-King \textit{et al.}~\cite{yee2018automatic}, where the \textit{Dexed} VST was used and different genetic and NN estimators were compared by using MFCC difference as the minimization objective.
-Depending on the NN architecture, MFCC or STFT spectrograms were used as the input.
-Unlike previous comparisons of NN and genetic algorithms, the best estimator was the multi-objective non-dominated sorting genetic algorithm III (NSGA-III).
-This algorithm had multiple minimization objectives: 13-band MFCC, STFT, and five spectral features.
-Shier \textit{et al.} report that all models performed relatively well at imitation of harmonic distributions, but the multi-objective function was the only optimizer which did \textit{not} struggle with matching temporal envelopes of the audio spectrum~\cite{shier2020spiegelib}.
-
-Differentiable loss functions that use spectrogram differences can be computationally expensive.
-To mitigate this, Han \textit{et al.}~\cite{han2023perceptual} introduced ``perceptual-neural-physical loss'' (PNP).
-PNP is an approximation of loss functions; specifically, a loss function that uses the L2 norm of the difference between the features of two sounds, or $||\phi(t) - \phi(x)||^2_2$, where $\phi$ could be a spectrogram or JTFS function.
-PNP loss functions are fast and differentiable, but require training and parameter estimation.
-A Riemannian metric M needs to be calculated for the minimization of locally linear approximation of the ``true" spectral loss function~\cite{han2023perceptual}.
-\[
-\|\phi(t) - \phi(x)\|_2^2 = \langle \tilde{\theta} - \theta | \mathbf{M}(\theta) | \tilde{\theta} - \theta \rangle + O(\|\tilde{\theta} - \theta\|_2^3). \tag{4}
-\]
-
-During the training phase, this metric is calculated alongside the neural parameter estimator.
-After training, PNP can be used as a fast approximation of computationally demanding loss functions such as JTFS~\cite{han2023perceptual,han2024learning}, paired with FM synthesizers and a differentiable physical model of a drum head~\cite{smith2010physical}.
-
-Uzrad \textit{et al.}~\cite{uzrad2024diffmoog} took another unique approach to sound-matching: using a differentiable \textit{synthesis chain} of DSP generators and effects and a loss function that combines P-Loss with a \textit{signal-chain loss}.
-The synthesizer is a customizable chain of effects, which feed one output as input to the next step of the chain; signal-chain loss compares the parameter and output difference at every output step in the chain~\cite{uzrad2024diffmoog}.
-Possible chain functionalities are FM/AM, Low-Frequency Oscillators (\gls{LFO}), filters, and envelopes.
-Like the results shown by Masuda \textit{et al.}~\cite{masuda2021soundmatch}, better out-of-domain results were achieved when pre-trained on in-domain data and fine-tuned using out-of-domain NSynth data~\cite{engel2017neural}.
-
-% Uzrad \textit{et al.}\cite{uzrad2024diffmoog} used log-spectral distance (LSD) as their loss function, where   
-% $\text{LSD} = \| \log(S(t)) - \log(S(x)) \|_F$, S is the mel-spectrogram of the sounds and F is the Frobenius norm~\cite{golub2013matrix}. The Frobenius norm of a matrix $A$ is defined as
-% $ \|A\|_F = \sqrt{\sum_{i,j} |a_{ij}|^2}$.
-Here, $A$ would be the 2D matrix resulting from subtracting two log-spectrograms.
-Along with other works, \textbf{non-convexity of loss surfaces when applying gradient optimization} is a noted issue~\cite{turian2020sorry,vahidi2023mesostructures,uzrad2024diffmoog}.
-
-\subsection{What is Lacking In the Field}
-\label{sec:lacking}
-%   % First, we need to clearly define the characteristics we are looking for with non-generic loss functions, otherwise the chances of finding the sound(s) of interest is slim.
-Second, we need to move beyond direct replication, and focus on imitating target sounds that are outside the synthesizers capabilities.
-Third, more complex methods of synthesis need to be explored for iterative sound-matching.
-Fourth, we need to find optimization methods that can effectively navigate the the sinusoidal nature of loss function landscapes.
-
-% Based on the literature reviewed, we find several gaps in past research that makes the application of sound-matching for sound-designers difficult.
-Thus far, replication of in-domain sounds has been the main subject of research.
-With in-domain replication, we already have the target sound \textit{and} the program that makes it.
-This simplifies the problem of measuring performance and is the reason why this type of sound-matching research is by far the most common.
-However, practical application of sound-matching requires answers to much harder problems.
-Sound-matching would be an interesting creative tool if we could use it to either:
-% \begin{enumerate}
-%  \item \textbf{Replicate out-of-domain sounds}: this would give us the parameters that could closely approximate a target sound, and we can modify the sound further by modulating the synthesizer. 
-% \item \textbf{Imitate in-domain or out-of-domain sounds}: This would allow for applying the characteristics of one sound to another. 
-% \end{enumerate}
-
-% We need to be aware of the ``open set recognition'' (\gls{OSR}) problem here \cite{mundt2019open,gers2000learning}.
-OSR is a well documented issue in DL that manifests when the set or category that the network is supposed to reject or accept cannot be explained via examples.
-With imitation or replication of out-of-domain sounds, we can expect issues related to OSR to occur, since target and output sounds would be from different domains~\cite{salimi2021percussive}.
-
-% Perhaps one method of addressing the OSR problem is the use of representations (or feature extraction) methods that reduce sounds to a feature vector in the same domain.
-For example, a representation that only looks at the envelope of sounds is far less likely to be subject to the OSR problem.
-\textcolor{highlight}{We highlight the lack of non-generic loss functions as a major weakness in previous works and an interesting area to explore}.
-
-% Another weakness \textcolor{highlight}{is that the synthesis methods in past works are not representative of the common methods used by sound designers and modern synthesizers}.
-Simple FM/AM synthesizers have generally been the go-to methods of sound-matching.
-We believe that general statements regarding the effectiveness of an approach to iterative sound-matching (particularly in regards to a loss function) can only be reasonably made in the context of the subset of the tested methods of synthesis.
-% \textcolor{highlight}{The simplicity of parameter optimization approaches} is another area of weakness in past works.
-Classic genetic algorithms update the parameters randomly, while in recent differentiable settings, simple gradient descent methods are used.
-Past works have provided issues which arise in differentiable parameter estimation due to periodic loss functions.
-Perhaps the use of reinforcement learning (\gls{RL}), or other algorithms that continually learn from their environment, could help with this problem.
-
-% \subsection{Defining The Major Issues}
-% \label{sec:addressing_problems}
-
-Having looked at the current available literature, we note four major areas of weakness in sound-matching.
-We target the first two issues in this work, while the latter two issues are left for future work.
-\begin{enumerate}
-    \item \LossSelect: Is there a \textit{best performing} loss function, or is the selection of the loss function dependent on factors such as the sound domain, synthesizer capability and the desired characteristics of the output?
-    \item \SynthSelect: There is a lack of diversity in the synthesis methods used in sound-matching.
-The use of different DSP functions for iterative sound-matching and their effects on the outcome or the interactions with the loss function have not been tested.
-    \item \PeriodicLoss: The sinusoidal nature of loss function landscapes is a problem that appears frequently.
-In differentiable settings, this problem causes gradient descent updates to local minima.
-    \item \OutDomain: Sound-matching with out-of-domain sounds is an unexplored area, yet a necessary one for practical applications for sound designers.
-\end{enumerate}
-
-\section{Programs, Methodology, and Results}
-\label{sec:progs_methodology_results}
+\section{Experimental Setup}
+\label{sec:experiment_setup}
 Towards addressing the problems of \LossSelect{} and \SynthSelect{}, we compare the performance of iterative sound-matching experiments, where four differentiable loss functions are applied to four differentiable synthesizers.
@@
 If true, this would strengthen the idea that the selection of the loss function and synthesis method is not an optimization problem, but a subjective choice depending on the needs of the sound designer.
+% For clear clarity and alignment with structural norms, we recommend splitting this into three sections: (1) Experimental Setup (including synthesizer programs, loss definitions, optimization procedure); (2) Experimental Results (performance metrics, statistical tests, and rankings); and (3) Discussion (qualitative interpretation, synthesis-specific observations, and practical implications).
 
-% In Section~\ref{sec:methodology} we lay out the experiment methodology.
-Section~\ref{sec:loss_implementation} discusses the details of implementation for the loss functions.
-Section~\ref{sec:programs} covers the characteristics of the programs, and finally Section~\ref{sec:results_posthoc} shows that the effect of loss function choice on the success of iterative sound-matching is dependent on the signal process chain of interest.
-
-\subsection{Experiment Methodology}
-\label{sec:methodology}
 % We have highlighted the lack of attention given to the selection of loss functions in sound-matching, and noted that the few existing works addressing this issue evaluate their proposed losses using only simplistic sound synthesis methods.
@@
 
-We test four different loss functions and compare their success in sound-matching using four different synthesizers.
+We test four different loss functions (described in Section~\ref{sec:loss_implementation}) and compare their success in sound-matching using four different synthesizers (Section~\ref{sec:programs}).
 We verify the success of the sound-matching process using two automatic measures of sound similarity (MSS and P-Loss) and a manual blinded hearing test conducted by two of the authors where Likert scores~\cite{jebb2021review} of 1-5 are assigned to a subset of final outputs and the corresponding targets.
+Details regarding the evaluation methods are described in Section~\ref{sec:evaluation_manual_auto}.
 
@@
 This simplified example leads us to our methodology which has four loss functions, four synthesizer programs, and three different evaluation methods of scoring sound similarity.
-
-\subsubsection{Automatic and Manual Evaluation}
-\label{sec:evaluation_manual_auto}
-The two automatic evaluation methods used here are P-Loss and MSS, as described in Section~\ref{sec:loss_funcs}.
-For P-Loss, the parameters are normalized between 0-1 based on the valid ranges defined in the Faust program, and the L1 distance between the normalized parameters is calculated.
-MSS is computed using a hop size of $100$ samples, and FFT window lengths of $(512, 1024, 2048, 4096)$.
-
-Automatic sound evaluation methods are often unreliable, as their alignment with the subjective judgments of human listeners is uncertain.
-Therefore, listening tests are conducted by randomly sampling 40 experiment results for each program and loss function pair.
-Two of the authors then assign blinded similarity scores to the outputs and targets using a 5-point Likert scale (1 = no similarity, 5 = near identical)~\cite{jebb2021review}.
-Using Spearman's rank correlation~\cite{spearman1987proof,rebekic2015pearson}, we found a very strong rate of agreement between the human listeners (see Section~\ref{sec:consistency_in_rankings}), a good indicator that the scores assigned by manual listeners can be treated as the ground truth.
-We then combine the ranks assigned by both authors, giving us 80 ranks per program and loss function pair.
-
-% Agreement between the assigned scores is measured using Spearman’s rank correlation, which provides both a correlation coefficient ($\rho$) ranging from –1 (perfect negative correlation) to 1 (perfect positive correlation), and a p-value testing the null hypothesis of no correlation~\cite{spearman1987proof,rebekic2015pearson}.
-Across all programs, the correlation was very strong ($\rho = 0.86$, $p < 10^{-180}$).
-Per-program correlations were also very strong: $\rho = 0.71$ for \BPNoise{} ($p < 10^{-25}$), $\rho = 0.64$ for \AddSineSaw{} ($p < 10^{-19}$), $\rho = 0.84$ for \AmpMod{} ($p < 10^{-43}$), and $\rho = 0.85$ for \FMMod{} ($p < 10^{-44}$).
-This high rate of agreement is a good indicator that the scores assigned by manual listeners can be treated as the ground truth.
-
-% For each program and loss function, the blind manual hearing test gives us 80 ranks, which are given by two of the authors to a set of 40 randomly selected experiment results.
-This gives 160 pairs of ranks per program, and 640 for all programs.
-We computed Spearman correlations between the two reviewers.
-
-\subsubsection{Determining Best Performers}
- Figure~\ref{fig:posthoc_evaluation} is a visualized summary of how the loss functions are ranked.
-We have four loss functions and four synthesizer programs.
-The maximum number of iterations set to 200, and 300 experiments for each loss/synthesizer combination.
-Each synthesizer program is paired with four loss functions and 300 experiments are conducted and evaluated automatically with P-Loss and MSS, and manually with Likert scores.
-
-For each program and loss function pair, we have three distributions that can be used to rank the loss functions.
-Two with 300 automatically assigned similarities (P-Loss and MSS) and one with 80 Likert scores (combining the 40 ranks assigned by each author).
-For consistency and statistical robustness, the distributions are upsampled to 1000 values using bootstrapping~\cite{tibshirani1993introduction}.
-Bootstrapping gives an estimation of the distribution for the mean performance of each experiment.
-$k$ (set to 1000) samples of $n$ values (set to 100\% of the values) are taken with replacement from the empirical distribution (list of 300 performance values for MSS and P-Loss or 80 for manual rankings).
-The mean performance is calculated for each of the $k$ samples.
-The $k$ estimates of the mean performance give us a bootstrapped distribution for each loss function.
-
-For each program, the loss functions are ranked using the non-parametric Scott-Knott (\gls{NPSK}) test~\cite{tantithamthavorn2017mvt,tantithamthavorn2018optimization}, which ranks the loss functions from rank 1 (best, or highest distribution values) to maximum rank of 4 (worst, with lowest distribution values).
-NPSK may determine that multiple functions belong to the same rank, in which case the maximum rank would be less than 4.
-As mentioned before, lower P-Loss and MSS values indicate higher similarity, therefore (for consistency with manually assigned ranks) the reciprocal of their output is used ($x^{-1}$, if $x$ is the P-Loss or MSS value), where higher values indicate greater similarity.
-
-\input{extra_figures/evaluation_graph.tex}
-
-\subsubsection{Training Parameters and Gradient Calculations}
- % Previous work utilized gradient fields for analysis of the algorithm performance~\cite{vahidi2023mesostructures}.
-Gradient fields are not the center of our analysis, but we include them in Appendix~\ref{appendix:gradient_fields}.
- For updating the synthesizer parameters (or weights), we use RMSProp, which operates similar to stochastic gradient descent (SGD)~\cite{goodfellow2016deep}, with the caveat that the gradients of each weight are scaled by the root-mean-square of past gradients of that weight.
-Based on some initial test runs, we used an arbitrary fixed learning rate of 0.045 for all experiments (learning rates used in Vahidi \textit{et al.}~\cite{vahidi2023mesostructures} is unknown).
-The maximum number of iterations is set to 200, where, based on our observations, the parameters have either irrecoverably diverged outside the acceptable ranges, or are stuck at a local minimum.
-Additionally, the gradients are large, and calculated backwards sequentially through the signal processing chain.
-This backward calculation resembles the \textit{exploding gradients} problem in recurrent neural networks~\cite{gers2000learning}.
-Gradient clipping~\cite{goodfellow2016deep} is preemptively used to ensure that the  $\ell_2$ norm of all gradients does not exceed the threshold of 1.
-% \mathbf{\gamma}_i^{\text{clipped}} = \mathbf{\gamma}_i \cdot \mathcal{S}
-% \]
-% Where $\mathcal{S}$ is a value less than or equal to 1, depending on whether the norm of the gradients exceeds the threshold $c$.  
-% \[
-% \mathcal{S} = \frac{c}{\max(\|\mathbf{\gamma}\|_2, c)}
-% \]
 
@@
 
- Once the programs are transpiled to differentiable synthesizers, the iterative sound-matching procedure is:
- \begin{enumerate}
-    \item Initialize $\theta^*$ and $\hat{\theta}$, i.e., random generation of target and initial parameters uniformly over a predefined range
-    \item Generating the output of the synthesizer with $\hat{\theta}$ (the length of the output is set to 1 second with sample-rate of 44100 Hz) 
-    \item Calculating the loss between the target and output
-    \item Applying gradient updates to the synthesizer parameters
-    \item Repeating the second step with the updated parameters $\hat{\theta}$, until maximum number of iterations has been reached
- \end{enumerate}
-
 % A general overview of the programs and the predefined parameter ranges is given in Table~\ref{tab:programs}.
@@
 In the following subsections, we describe the programs and their parameters in more detail, followed by the methodology in Section~\ref{sec:methodology} and results in Section~\ref{sec:results_posthoc}.
-
-% \begin{table*}[htbp]
-% \centering
-% \begin{adjustbox}{max width=\textwidth}
-% \begin{tabular}{|c|c|c|c|l|}
-% \hline
-% \textbf{ID} & \textbf{Parameters} & \makecell{\textbf{Param. 1}\\\textbf{Range}} & \makecell{\textbf{Param. 2}\\\textbf{Range}} & \textbf{Description} \\
-% \hline
-% \BPNoise & hp\_cut, lp\_cut & 50-1000 & 1-120 & Band-pass of Noise signal \\
-% \hline
-% \AddSineSaw & saw\_freq, sine\_freq & 20-1000 & 20-1000 & Addition of sine and saw waves \\
-% \hline
-% \AmpMod & amp, carrier & 1-20 & 20-1000 & LFO on amplitude of noise osc \\
-% \hline
-% \FMMod & amp, carrier & 0-5 & 0-4 & LFO on amplitude of saw osc \\
-% \hline
-% \end{tabular}
-% \end{adjustbox}
-% \caption{Overview of Program Parameters and the range of values each parameter can take.}
-% \label{tab:programs}
-% \end{table*}
 
@@
 
-\subsection{Post hoc Analysis of Loss Performance}
+\subsection{Evaluation Methods}
+\label{sec:evaluation_manual_auto}
+The two automatic evaluation methods used here are P-Loss and MSS, as described in Section~\ref{sec:loss_funcs}.
+For P-Loss, the parameters are normalized between 0-1 based on the valid ranges defined in the Faust program, and the L1 distance between the normalized parameters is calculated.
+MSS is computed using a hop size of $100$ samples, and FFT window lengths of $(512, 1024, 2048, 4096)$.
+
+Automatic sound evaluation methods are often unreliable, as their alignment with the subjective judgments of human listeners is uncertain.
+Therefore, listening tests are conducted by randomly sampling 40 experiment results for each program and loss function pair.
+Two of the authors then assign blinded similarity scores to the outputs and targets using a 5-point Likert scale (1 = no similarity, 5 = near identical)~\cite{jebb2021review}.
+Using Spearman's rank correlation~\cite{spearman1987proof,rebekic2015pearson}, we found a very strong rate of agreement between the human listeners (see Section~\ref{sec:consistency_in_rankings}), a good indicator that the scores assigned by manual listeners can be treated as the ground truth.
+We then combine the ranks assigned by both authors, giving us 80 ranks per program and loss function pair.
+
+% Agreement between the assigned scores is measured using Spearman’s rank correlation, which provides both a correlation coefficient ($\rho$) ranging from –1 (perfect negative correlation) to 1 (perfect positive correlation), and a p-value testing the null hypothesis of no correlation~\cite{spearman1987proof,rebekic2015pearson}.
+Across all programs, the correlation was very strong ($\rho = 0.86$, $p < 10^{-180}$).
+Per-program correlations were also very strong: $\rho = 0.71$ for \BPNoise{} ($p < 10^{-25}$), $\rho = 0.64$ for \AddSineSaw{} ($p < 10^{-19}$), $\rho = 0.84$ for \AmpMod{} ($p < 10^{-43}$), and $\rho = 0.85$ for \FMMod{} ($p < 10^{-44}$).
+This high rate of agreement is a good indicator that the scores assigned by manual listeners can be treated as the ground truth.
+
+% For each program and loss function, the blind manual hearing test gives us 80 ranks, which are given by two of the authors to a set of 40 randomly selected experiment results.
+This gives 160 pairs of ranks per program, and 640 for all programs.
+We computed Spearman correlations between the two reviewers.
+
+\subsection{Training Parameters and Gradient Calculations}
+ % Previous work utilized gradient fields for analysis of the algorithm performance~\cite{vahidi2023mesostructures}.
+Gradient fields are not the center of our analysis, but we include them in Appendix~\ref{appendix:gradient_fields}.
+ For updating the synthesizer parameters (or weights), we use RMSProp, which operates similar to stochastic gradient descent (SGD)~\cite{goodfellow2016deep}, with the caveat that the gradients of each weight are scaled by the root-mean-square of past gradients of that weight.
+Based on some initial test runs, we used an arbitrary fixed learning rate of 0.045 for all experiments (learning rates used in Vahidi \textit{et al.}~\cite{vahidi2023mesostructures} is unknown).
+The maximum number of iterations is set to 200, where, based on our observations, the parameters have either irrecoverably diverged outside the acceptable ranges, or are stuck at a local minimum.
+Additionally, the gradients are large, and calculated backwards sequentially through the signal processing chain.
+This backward calculation resembles the \textit{exploding gradients} problem in recurrent neural networks~\cite{gers2000learning}.
+Gradient clipping~\cite{goodfellow2016deep} is preemptively used to ensure that the  $\ell_2$ norm of all gradients does not exceed the threshold of 1.
+% \mathbf{\gamma}_i^{\text{clipped}} = \mathbf{\gamma}_i \cdot \mathcal{S}
+% \]
+% Where $\mathcal{S}$ is a value less than or equal to 1, depending on whether the norm of the gradients exceeds the threshold $c$.  
+% \[
+% \mathcal{S} = \frac{c}{\max(\|\mathbf{\gamma}\|_2, c)}
+% \]
+
+\subsection{Determining Best Performers}
+ Once the programs are transpiled to differentiable synthesizers, the iterative sound-matching procedure is:
+ \begin{enumerate}
+    \item Initialize $\theta^*$ and $\hat{\theta}$, i.e., random generation of target and initial parameters uniformly over a predefined range
+    \item Generating the output of the synthesizer with $\hat{\theta}$ (the length of the output is set to 1 second with sample-rate of 44100 Hz) 
+    \item Calculating the loss between the target and output
+    \item Applying gradient updates to the synthesizer parameters
+    \item Repeating the second step with the updated parameters $\hat{\theta}$, until maximum number of iterations has been reached
+ \end{enumerate}
+
+ Figure~\ref{fig:posthoc_evaluation} is a visualized summary of how the loss functions are ranked.
+We have four loss functions and four synthesizer programs.
+The maximum number of iterations set to 200, and 300 experiments for each loss/synthesizer combination.
+Each synthesizer program is paired with four loss functions and 300 experiments are conducted and evaluated automatically with P-Loss and MSS, and manually with Likert scores.
+
+For each program and loss function pair, we have three distributions that can be used to rank the loss functions.
+Two with 300 automatically assigned similarities (P-Loss and MSS) and one with 80 Likert scores (combining the 40 ranks assigned by each author).
+For consistency and statistical robustness, the distributions are upsampled to 1000 values using bootstrapping~\cite{tibshirani1993introduction}.
+Bootstrapping gives an estimation of the distribution for the mean performance of each experiment.
+$k$ (set to 1000) samples of $n$ values (set to 100\% of the values) are taken with replacement from the empirical distribution (list of 300 performance values for MSS and P-Loss or 80 for manual rankings).
+The mean performance is calculated for each of the $k$ samples.
+The $k$ estimates of the mean performance give us a bootstrapped distribution for each loss function.
+
+% For each program, the loss functions are ranked using the non-parametric Scott-Knott (\gls{NPSK}) test~\cite{tantithamthavorn2017mvt,tantithamthavorn2018optimization}, which ranks the loss functions from rank 1 (best, or highest distribution values) to maximum rank of 4 (worst, with lowest distribution values).
+NPSK may determine that multiple functions belong to the same rank, in which case the maximum rank would be less than 4.
+As mentioned before, lower P-Loss and MSS values indicate higher similarity, therefore (for consistency with manually assigned ranks) the reciprocal of their output is used ($x^{-1}$, if $x$ is the P-Loss or MSS value), where higher values indicate greater similarity.
+
+\input{extra_figures/evaluation_graph.tex}
+
+\section{Post hoc Analysis of Loss Performance}
 \label{sec:results_posthoc}
-As summarized in Figure~\ref{fig:posthoc_evaluation}, for each program and evaluation (or scoring) method, there are four sets 1000 bootstrapped values, corresponding to the four loss functions.
+As discussed in the previous section and summarized in Figure~\ref{fig:posthoc_evaluation}, for each program and evaluation (or scoring) method, there are four sets of 1000 bootstrapped values corresponding to the four loss functions.
 A post-hoc test is conducted in two stages in order to determine whether there is a difference in performance between loss functions for each program, and if so, which are the best performers.
@@
 
-\subsubsection{Best Performers For Each Program}
+\subsection{Best Performers For Each Program}
 Based on the bootstrapped distribution of scores, the NPSK algorithm ranks the loss functions from 1 (best) to a maximum of 4 (worst).
@@
 The corresponding colors for each rank are
-\colorbox{rank1}{\textcolor{black}{\textbf{1}}} \colorbox{rank2}{\textcolor{white}{\textbf{2}}} \colorbox{rank3}{\textcolor{white}{\textbf{3}}} \colorbox{rank4}{\textcolor{black}{\textcolor{white}{\textbf{4}}}}
+\colorbox{rank1}{\textcolor{black}{\textbf{1}}} \colorbox{rank2}{\textcolor{white}{\textbf{2}}} \colorbox{rank3}{\textcolor{white}{\textbf{3}}} \colorbox{rank4}{\textcolor{black}{\textcolor{white}{\textbf{4}}}}.
 
@@
 
+\section{Takeaways, Practical Implications, and Caveats}
+% (Q1) To what extent do automatic evaluation metrics agree with manual listening tests?
+(Q2) Can novel loss functions utilizing Dynamic Time Warping (DTW) and Scale-Invariant Mean Squared Error (SIMSE) provide advantages over standard loss functions, and if so, when?
+(Q3) Is an iterative differentiable optimization a viable strategy for design of sound-matching experiments?
+The experiment results provide answers regarding the main hypothesis of this work as well as the secondary questions. Based on these answers, we can make practical recommendations for future research.
+
 \subsection{Takeaways}
-An important takeaway is that the loss function that yields the best sound-matching outcomes varies depending on the synthesizer program.
-This is perhaps due to the interaction between how the parameters of the synthesizer influence the sound, and the core sonic features of the loss function.
+Perhaps the most important takeaway regarding our main hypothesis is that the loss function that yields the best sound-matching outcomes varies depending on the synthesizer program.
+This is likely due to the interaction between how the parameters of the synthesizer influence the sound, and the core sonic features used by the loss function.
 
-We also see that the \DTWEnv{} and \SIMSESpec{} (to our knowledge, not previously used in iterative sound-matching), lead to the best results in three of the four programs.
-This shows room for further creative approaches to differentiable sound-similarity measures, particularly since we observed that SOTA loss functions can be ineffective depending on the design of a synthesizer.
+\textbf{Q1}: To what extent do automatic evaluation metrics agree with manual listening tests?
+we see somewhat consistent results between the rankings assigned by manual hearing tests (which we take as the ground truth), and automatic measures of P-Loss and MSS.
+With the exception of P-Loss in \BPNoise{} narrowly selecting a different spectrogram based loss, all measures of performance selected the same top performer.
+Due to the simplicity of our programs and the occasional deviations, we cannot conclude that these automatic measures are substitutes for human hearing tests.
+As many previous works have done\todo{cite}, the use of MSS and/or P-Loss as general loss functions does seem appropriate, however, we advise the use of listening tests or custom made loss functions for more specific analysis.
 
-Finally, we surprisingly see somewhat consistent results between the rankings assigned by manual hearing tests (which we take as the ground truth), and automatic measures of P-Loss and MSS.
-Due to the simplicity of our programs and the occasional deviations, we cannot conclude that these automatic measures are substitutes for human hearing tests.
+\textbf{Q2}: Are DTW and SIMSE effective measures of loss? If so,
 
-\section{Summary, Weaknesses, and Conclusion}
-\label{sec:summary_conclusion}
-\textbf{Summary:} Sound-matching is an umbrella term for the algorithmic programming of audio synthesizers, often with the goal of assisting sound designers.
-Here we provided a history of sound-matching, major issues in the field, and discussed the importance of ``differentiable iterative sound-matching'' as a natural extension of current literature.
+ \DTWEnv{} consistently outperformed other measures in amplitude-modulated synthesis, where envelope alignment is critical.
+\SIMSESpec{} performed best in subtractive synthesis, where scale-invariance captures noise filtering effects more robustly than L1-based measures.
+These results indicate that the advantages of novel losses are context-dependent, and the necessity of further creative approaches to differentiable sound-similarity measures.
 
-The main hypothesis tested here is whether the performance of differentiable loss functions is influenced by the synthesis techniques used for sound-matching.
-We created iterative sound-matching experiments by combining four different loss functions with four different sound synthesis programs.
-We ranked the performance of the iterative sound-matching pipelines for every loss function and program, and observed that the success of the pipeline (that is, how closely the output sound matches the target sound) is program dependent.
-In other words, different synthesizer programs work best with different loss functions.
-Notably, we see that our novel use of DTW and SIMSE based differentiable loss functions (\DTWEnv{} and \SIMSESpec) can outperform what are regarded as the SOTA loss functions in 3 of 4 cases, although this is highly synthesizer dependent.
+\textbf{Q3}: Can iterative differentiable optimization be applied effectively in synthesizers using classical DSP functions?
+The approach of designing DSP functions in Faust and transpiling to differentiable Faust code yielded meaningful outputs and enabled comparative evaluation of losses.
+The use of Faust for synthesizer definition was convenient, and transpiling to Jax enabled quick gradient optimization.
+We were able to run our 1200 experiments within 72 hours on a laptop without a GPU\footnote{Lenovo ThinkPad T480, i5-8350U CPU at 1.70GHz, 32 GB RAM.}.
 
-P-Loss and MSS have frequently been used as automatic performance measures, yet their ``preference'' has rarely been compared to human rankings.
-We observed that automatic performance measures and manual listening tests were often in agreement, despite this, manual verification of sound-matching results remains a necessity for future works.
+\subsection{Practical Recommendations:}
+Based on our findings, several actionable guidelines emerge.
+These heuristics can guide practitioners in selecting appropriate similarity measures and synthesis methods for future sound-matching experiments.
+\begin{itemize}
+    \item In leu of listening tests, MSS or P-Loss are generally useful for large-scale benchmarking, but confirm conclusions with manual listening when accuracy in isolated cases is important. 
+    \item For amplitude-modulated synthesis, DTW-based losses (e.g., \DTWEnv{}) are most effective.
+    \item For subtractive/noise-filtered synthesis, spectrogram based losses are most effective. We can recommend the use of SIMSE over L1, particularly when relative amplitudes are less important. 
+    \item Iterative differentiable optimization via Faust-to-JAX is a viable strategy for defining various differentiable DSP functions, requiring only modest hardware resources.
+\end{itemize}
 
-\textbf{Weaknesses:} 
-While we cannot prove that a universally best similarity measure does not exist, we can advocate for more creativity in the field based on our findings.
-Compared to previous work, we presented a more cohesive approach to iterative sound-matching which utilizes a variety of loss functions and synthesis methods.
-However, there are many other methods of synthesis and sound-similarity that can be combined in practically infinite ways.
-Due to this large search space, we set arbitrary parameters for the various signal processing functions.
-We used bare-bones versions of STFT and JTFS with fixed parameters.
-We did not test complex synthesizers using parallel and sequential DSP functions.
-Arbitrary hyperparameters such as learning rate and max number of iterations were selected for the DL pipeline, and only the RMSProp optimizer was tested.
+\subsection{Applicability to More Complex Synthesizers}
+The experiments in this study were conducted with differentiable synthesizers of modest complexity (2 parameters each), chosen to isolate fundamental synthesis principles. 
+While this level of simplicity facilitates controlled comparisons, it does not fully capture the parameter richness of real-world synthesizers, which can have hundreds of parameters routed in various ways.
+The generalizability of our findings does require further research.
 
- Like the majority of previous works, this work utilizes in-domain sounds; that is, the target sound is made by the synthesizer, and the parameters are already known.
-This simplifies the issues of measuring sound-similarity, but it is not a realistic scenario for practical sound-matching.
-This problem is left for future work, discussed in the next section.
+Nonetheless, we believe that the insights here can extend naturally to more complex domains.
+First, the observed dependence of loss performance on synthesis method is will likely remain stable even with added complexities such as layers of modulation, filtering, or nonlinearity.
+\DTWEnv{} and \SIMSESpec{} measure very specific features of audio, and their use in the recommended settings would likely yield the desired results regardless of synthesizer complexity.
 
-\label{sec:future}
-\textbf{Future Work: } The problem of periodic loss-landscapes, noted by many previous works, remains unaddressed~\cite{turian2020sorry,vahidi2023mesostructures,uzrad2024diffmoog,bruford2024synthesizer}.
-Perhaps this problem emerges due to the periodic nature of sound, which requires better loss landscape navigation methods.
-An optimizer that is more aware of fluctuations in the gradients would perhaps lead to better solutions than simple gradient descent.
-Viewing the sound-matching problem as ``the navigation of an agent from an arbitrary point in a gradient field to a target'' closely resembles many classical problems in the field of reinforcement learning (RL)~\cite{sutton2018reinforcement}.
-Naturally, the application of RL and other heuristic search techniques to the problem of iterative sound-matching would be an important contribution.
+(same as previous sentence) Second, the heuristics identified here (e.g., DTW\_Envelope excelling in amplitude-modulated contexts, SIMSE\_Spec in subtractive filtering) map directly onto common building blocks found in commercial and modular synthesizers, suggesting that these recommendations can guide loss selection in practical scenarios.
 
-Contemporary works often involve the application of domain specific and computationally expensive loss functions~\cite{han2023perceptual,uzrad2024diffmoog}, use of large neural networks~\cite{hershey2017cnn,cramer2019look}, or complex ensemble methods~\cite{turian2022hear}.
-Such models are useful but intractable; furthermore, training them requires the definition of simpler loss functions, which emphasizes the need for further development of differentiable and expressive loss function implementations.
+% Third, the iterative differentiable optimization framework proved viable even on limited hardware, which indicates that similar pipelines could be applied to larger differentiable synthesizers if computational resources allow.
 
-Like nearly all previous work in sound-matching, the main measure of success here was the accurate \textit{replication} of sounds (or synth parameters) rather than imitation of sounds outside of the synthesizer's domain.
-Replication of arbitrary features of sounds is an important component of sound-design, though it is much harder to define or measure.
-Future works could explore loss functions which only measure certain characteristics of sound (such as \DTWEnv), and whether they can pave the way for better imitation in sound-matching.
+% At the same time, real-world synthesizers present new challenges. 
+% High-dimensional parameter spaces may exacerbate issues of local minima and make optimization less stable, particularly when loss functions are only partially informative.
 
-\bibliographystyle{IEEEtran}
-\bibliography{references}
-\end{document}
+% Non-differentiable components (e.g., physical modeling, complex routing) also limit the direct applicability of gradient-based methods.
+
+% Future work should therefore investigate hybrid approaches, combining differentiable blocks with surrogate models or gradient-free optimization for non-differentiable components. 
+% Out-of-domain targets, such as recorded instruments or environmental sounds, represent another direction where the loss–synth interaction may behave differently than in controlled in-domain experiments.

### 2025-09-10 sections/background_related.tex
Updates from Overleaf

@@
 
-\subsubsection{Novelty of Scale-Invariant Spectrogram Error}
+\textbf{Why use Scale-Invariant Spectrogram for audio loss}
 A limitation of L1/L2 distances on spectrograms is their sensitivity to global gain: 
@@
 
-\paragraph*{Justification and Novelty}
+\textbf{Why use DTW for audio loss?}
 DTW measures the distance between the time-series \textit{after} alignments, typically using Euclidean distance.
@@
 
-% Section IV titled "Programs, Methodology, and Results" is overloaded and blends experimental setup with results and interpretation.
-For clear clarity and alignment with structural norms, we recommend splitting this into three sections: (1) Experimental Setup (including synthesizer programs, loss definitions, optimization procedure); (2) Experimental Results (performance metrics, statistical tests, and rankings); and (3) Discussion (qualitative interpretation, synthesis-specific observations, and practical implications).
+\subsection{What is Lacking In the Field}
+\label{sec:lacking}
+%   % First, we need to clearly define the characteristics we are looking for with non-generic loss functions, otherwise the chances of finding the sound(s) of interest is slim.
+Second, we need to move beyond direct replication, and focus on imitating target sounds that are outside the synthesizers capabilities.
+Third, more complex methods of synthesis need to be explored for iterative sound-matching.
+Fourth, we need to find optimization methods that can effectively navigate the the sinusoidal nature of loss function landscapes.
+
+% Based on the literature reviewed, we find several gaps in past research that makes the application of sound-matching for sound-designers difficult.
+Thus far, replication of in-domain sounds has been the main subject of research.
+With in-domain replication, we already have the target sound \textit{and} the program that makes it.
+This simplifies the problem of measuring performance and is the reason why this type of sound-matching research is by far the most common.
+However, practical application of sound-matching requires answers to much harder problems.
+Sound-matching would be an interesting creative tool if we could use it to either:
+% \begin{enumerate}
+%  \item \textbf{Replicate out-of-domain sounds}: this would give us the parameters that could closely approximate a target sound, and we can modify the sound further by modulating the synthesizer. 
+% \item \textbf{Imitate in-domain or out-of-domain sounds}: This would allow for applying the characteristics of one sound to another. 
+% \end{enumerate}
+
+% We need to be aware of the ``open set recognition'' (\gls{OSR}) problem here \cite{mundt2019open,gers2000learning}.
+OSR is a well documented issue in DL that manifests when the set or category that the network is supposed to reject or accept cannot be explained via examples.
+With imitation or replication of out-of-domain sounds, we can expect issues related to OSR to occur, since target and output sounds would be from different domains~\cite{salimi2021percussive}.
+
+% Perhaps one method of addressing the OSR problem is the use of representations (or feature extraction) methods that reduce sounds to a feature vector in the same domain.
+For example, a representation that only looks at the envelope of sounds is far less likely to be subject to the OSR problem.
+\textcolor{highlight}{We highlight the lack of non-generic loss functions as a major weakness in previous works and an interesting area to explore}.
+
+% Another weakness \textcolor{highlight}{is that the synthesis methods in past works are not representative of the common methods used by sound designers and modern synthesizers}.
+Simple FM/AM synthesizers have generally been the go-to methods of sound-matching.
+We believe that general statements regarding the effectiveness of an approach to iterative sound-matching (particularly in regards to a loss function) can only be reasonably made in the context of the subset of the tested methods of synthesis.
+% \textcolor{highlight}{The simplicity of parameter optimization approaches} is another area of weakness in past works.
+Classic genetic algorithms update the parameters randomly, while in recent differentiable settings, simple gradient descent methods are used.
+Past works have provided issues which arise in differentiable parameter estimation due to periodic loss functions.
+Perhaps the use of reinforcement learning (\gls{RL}), or other algorithms that continually learn from their environment, could help with this problem.
+
+% \subsection{Defining The Major Issues}
+% \label{sec:addressing_problems}
+
+Having looked at the current available literature, we note four major areas of weakness in sound-matching.
+We target the first two issues in this work, while the latter two issues are left for future work.
+\begin{enumerate}
+    \item \LossSelect: Is there a \textit{best performing} loss function, or is the selection of the loss function dependent on factors such as the sound domain, synthesizer capability and the desired characteristics of the output?
+    \item \SynthSelect: There is a lack of diversity in the synthesis methods used in sound-matching.
+The use of different DSP functions for iterative sound-matching and their effects on the outcome or the interactions with the loss function have not been tested.
+    \item \PeriodicLoss: The sinusoidal nature of loss function landscapes is a problem that appears frequently.
+In differentiable settings, this problem causes gradient descent updates to local minima.
+    \item \OutDomain: Sound-matching with out-of-domain sounds is an unexplored area, yet a necessary one for practical applications for sound designers.
+\end{enumerate}

### 2025-09-10 sections/setup_and_posthoc_analysis.tex
Updates from Overleaf

@@
     \item In leu of listening tests, MSS or P-Loss are generally useful for large-scale benchmarking, but confirm conclusions with manual listening when accuracy in isolated cases is important. 
-    \item For amplitude-modulated synthesis, DTW-based losses (e.g., \DTWEnv{}) are most effective.
-    \item For subtractive/noise-filtered synthesis, spectrogram based losses are most effective. We can recommend the use of SIMSE over L1, particularly when relative amplitudes are less important. 
-    \item Iterative differentiable optimization via Faust-to-JAX is a viable strategy for defining various differentiable DSP functions, requiring only modest hardware resources.
+    \item For amplitude-modulated synthesis, DTW-based losses (e.g., \DTWEnv{}) are recommended over spectrogram based and JTFS. 
+    \item For subtractive/noise-filtered synthesis, spectrogram based losses are most effective.
+Based on our results, SIMSE appears to be the better method of spectrogram comparisons relative to L1, but the amplitude agnosticism of SIMSE likely plays a role in its poor performance in additive synthesis.
+    \item In our amplitude modulation experiments, \JTFS{} did not outperform other measures, contradicting prior claims of its superiority for mesostructures~\cite{vahidi2023mesostructures}.
+Thus, our results call into question its effectiveness as a general mesostructure measure
+    \item Iterative differentiable optimization via Faust-to-JAX is a viable strategy for defining various differentiable DSP functions, requiring only modest hardware resource and providing a more natural approach to synthesizer definition.
 \end{itemize}
@@
 \subsection{Applicability to More Complex Synthesizers}
-The experiments in this study were conducted with differentiable synthesizers of modest complexity (2 parameters each), chosen to isolate fundamental synthesis principles. 
-While this level of simplicity facilitates controlled comparisons, it does not fully capture the parameter richness of real-world synthesizers, which can have hundreds of parameters routed in various ways.
-The generalizability of our findings does require further research.
+The experiments in this study were conducted with differentiable synthesizers chosen to isolate fundamental synthesis principle with 2 parameters each.
+While this level of simplicity facilitates controlled comparisons, it does not fully capture the richness of real-world synthesizers, which can have hundreds of parameters with various routes.
+As a result, the extent of the generalizability of our findings does require further research.
 
 Nonetheless, we believe that the insights here can extend naturally to more complex domains.
-First, the observed dependence of loss performance on synthesis method is will likely remain stable even with added complexities such as layers of modulation, filtering, or nonlinearity.
-\DTWEnv{} and \SIMSESpec{} measure very specific features of audio, and their use in the recommended settings would likely yield the desired results regardless of synthesizer complexity.
+Depending on the feature of interest, the observed dependence of loss performance on synthesis method is will likely remain stable even with added complexities such as layers of modulation, filtering, or nonlinearity.
 
-(same as previous sentence) Second, the heuristics identified here (e.g., DTW\_Envelope excelling in amplitude-modulated contexts, SIMSE\_Spec in subtractive filtering) map directly onto common building blocks found in commercial and modular synthesizers, suggesting that these recommendations can guide loss selection in practical scenarios.
+\DTWEnv{} measures very specific features of audio, and its use in the recommended settings---where low frequency amplitude modulation matching is required---would likely yield the desired results regardless of synthesizer complexity.
+Likewise, the use of \SIMSESpec{} or \LoneSpec{} for matching filter cut-offs is likely to be successful regardless of the underlying sound.
 

### 2025-09-10 sections/summary_weakness_conclusion.tex
Updates from Overleaf

--- 2025-08-21 main.tex
@@
-\documentclass[lettersize,journal]{IEEEtran}
-\usepackage{amsmath,amsfonts}
-\usepackage{algorithmic}
-\usepackage{algorithm}
-\usepackage{array}
-\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-\usepackage{textcomp}
-\usepackage{stfloats}
-\usepackage{url}
-\usepackage{verbatim}
-\usepackage{graphicx}
-\usepackage{xcolor}
-\usepackage{csquotes}
-\usepackage{makecell}
-% my added packages and commands =======================
-\usepackage[numbers]{natbib} %removed cite package
-\usepackage{tikz}
-\usetikzlibrary{positioning,shapes.geometric}
-\usepackage{adjustbox}
-\usepackage{listings}
-\usepackage{caption}
-
-\newcommand{\highlight}[1]{\textcolor[RGB]{00,100,100}{#1}}
-\newcommand{\todo}[1]{\textcolor{red}{#1}}
-
-\captionsetup[lstlisting]{justification=centering, singlelinecheck=false}
-\providecommand{\gls}[1]{#1}
-\definecolor{rank1}{HTML}{70FF70}
-\definecolor{rank2}{HTML}{858585}
-\definecolor{rank3}{HTML}{454545}
-\definecolor{rank4}{HTML}{000000}
-\newcommand{\SIMSESpec}{\texttt{SIMSE\_Spec}}
-\newcommand{\LoneSpec}{\texttt{L1\_Spec}}
-\newcommand{\JTFS}{\texttt{JTFS}}
-\newcommand{\DTWEnv}{\texttt{DTW\_Envelope}}
-\newcommand{\LossSelect}{\textbf{Loss Selection}}
-\newcommand{\SynthSelect}{\textbf{Synthesis Selection}}
-\newcommand{\PeriodicLoss}{\textbf{Periodic Loss}}
-\newcommand{\OutDomain}{\textbf{Out-Domain Generation}}
-\newcommand{\BPNoise}{\textbf{BP-Noise}}  
-\newcommand{\AddSineSaw}{\textbf{Add-SineSaw}}  
-\newcommand{\AmpMod}{\textbf{Noise-AM}}  
-\newcommand{\FMMod}{\textbf{SineSaw-AM}}
-
-\input{code_style}
-
-%  ================================
-\begin{document}
-\title{Evaluating Sound Similarity Metrics for Differentiable, Iterative Sound-Matching}
-
-\author{Amir Salimi, Abram Hindle, Osmar R. Za{\"i}ane}
-
-% \markboth{IEEE TRANSACTIONS ON AUDIO, SPEECH AND LANGUAGE PROCESSING, VOL -, 2025}%
-% {Shell \MakeLowercase{\textit{et al.}}: A Sample Article Using IEEEtran.cls for IEEE Journals}
-
-\markboth{%
-  \parbox{\textwidth}{%
-    This work has been submitted to the IEEE for possible publication. \\
-    Copyright may be transferred without notice, after which this version may no longer be accessible.%
-  }%
-}{}
-
-\maketitle
-
-\begin{abstract}
-Manual sound design with a synthesizer is inherently iterative: an artist compares the synthesized output to a mental target, adjusts parameters, and repeats until satisfied.
-Iterative sound-matching automates this workflow by continually programming a synthesizer under the guidance of a loss function (or similarity measure) towards a target sound.
-Prior comparisons of loss functions have typically favored one metric over another, but only within narrow settings: limited synthesis methods, few loss types, often without blind listening tests.
-This leaves open the question of whether a universally optimal loss exists, or the choice of loss remains a creative decision conditioned on the synthesis method and the sound designer's preference.
-We propose differentiable iterative sound-matching as the natural extension of the available literature, since it combines the manual approach to sound design with modern advances in machine learning.
-To analyze the variability of loss function performance across synthesizers, we implemented a mix of four novel and established differentiable loss functions, and paired them with differentiable subtractive, additive, and AM synthesizers.
-For each of the sixteen synthesizer–loss combinations, we ran 300 randomized sound-matching trials.
-Performance was measured using parameter differences, spectrogram-distance metrics, and manually assigned listening scores.
-We observed a moderate level of consistency among the three performance measures.
-Our post-hoc analysis shows that the loss function performance is highly dependent on the synthesizer.
-These findings underscore the value of expanding the scope of sound-matching experiments, and developing new similarity metrics tailored to specific synthesis techniques, rather than pursuing one-size-fits-all solutions.
-\end{abstract}
-
-% the automatic performance measures have not been compared to human hearing
-% lack of creativity in loss function definitions 
-\begin{IEEEkeywords}
-Audio synthesis, differentiable digital signal processing, music information retrieval, sound-matching
-\end{IEEEkeywords}
-
-\section{Introduction}
-Since the 1970s, artists such as Suzanne Ciani~\cite{ciani_life_in_waves} have explored the use of audio synthesizers to replicate natural sounds—such as a bottle cap popping off—in ways that preserve the essential qualities of the original while showcasing the unique characteristics of the synthesizer.
-Rather than merely replicating sounds, this approach uses synthesizers for artistic reinterpretation.
-However, the automation of this interpretive approach to sound design using digital audio synthesizers has rarely been explored, and significant knowledge gaps remain even in the simpler problem of replicating sounds.
-One such gap is the lack of diversity and depth in the analysis of sound-similarity measures and their interaction with various synthesis methods.
-As a stepping stone toward automated interpretive sound design, we use automatic and manual sound-evaluations to investigate whether the optimal choice of a sound-similarity measure is dependent on the synthesis method.
-
-A digital audio synthesizer is any software used for the creation and manipulation of audio.
-Since the advent of digital signal processing (\gls{DSP}) in the 1960s~\cite{stranneby2004digital}, a wide variety of parametric audio generation functions have been proposed~\cite{lyons1997understanding,russ1999sound,shier2020spiegelib}.
-Manual sound design with a synthesizer is an iterative process of trial and error; when the artist has a desired (or target) sound in mind, sound design involves listening to the output, determining whether the output is closer or farther from the target, and adjusting the parameters of the synthesizer accordingly until the target is reached~\cite{russ1999sound}.
-This paper explores the automation of this process, which we call \textit{iterative sound-matching}.
-
-Automatic sound design has been a subject of research for decades, with reduction of tinkering time and creation of ``interesting" sounds as the main motivators \cite{krekovic2019insights,turian2020sorry,horner1993machine,salimi2020make,esling2019flow,engel2020ddsp,mitchell2007evolutionary,shier2020spiegelib,masuda2021soundmatch,masuda2023improving}.
-\textit{Sound-matching} is a common form of automatic sound design explored in literature, where a similarity metric (or loss function) is given, and the goal is to find parameters for a synthesizer that replicate \textit{all} or \textit{some} of the characteristics of a target sound as best as possible~\cite{horner1993machine,mitchell2007evolutionary,masuda2023improving}.
-Our analysis of past works shows that this general goal of ``sound-matching'' can be broken down in to very different subtypes, many of them rarely or never explored, despite their relevance to practical applications.
-This leads us to a typology of sound-matching tasks, and the importance of \textit{differentiable, iterative sound-matching}, rarely explored in previous literature~\cite{vahidi2023mesostructures}.
-
-\textit{Iterative sound-matching} refers to works that mimic the back-and-forth process of listening to a synthesizer's output and updating the parameters until the termination conditions are met.
-The two necessary components are the \textit{loss function}, which determines how close two sounds are, and the \textit{parametric synthesizer} that outputs sounds based on input parameters.
-In the iterative workflow, a target sound is selected, the synthesizer outputs sounds based on its parameters, and the loss function determines the proximity of the output to the target.
-From there, the synthesizer parameters are updated, the output and target are compared, and so on.
-
-Despite the early popularity of the iterative approach~\cite{justice1979analytic,horner1993machine,mitchell2007evolutionary}, recent works in sound-matching have favored supervised learning using inference models, deep neural-networks (\gls{DNNs}), and large datasets.
-Rather than continually interacting with the environment, these models learn from a dataset of inputs (synthesizer parameters) and outputs (corresponding sound), and predict the synthesizer parameters for new sounds~\cite{yee2018automatic,han2023perceptual,han2024learning,masuda2021soundmatch,masuda2023improving}.
-
-Iterative sound-matching is interesting for multiple reasons.
-First, it better resembles the human approach to sound-design, which requires continually interacting with the environment and making decisions according to a feedback signal~\cite{sutton2018reinforcement}.
-Second, the trial-and-error approach leaves a trail of parameter sets from the starting point to the end goal, which could be of use to sound designers.
-Third, relative to the DNN-based inference models, it provides a more tractable environment with room for creative approaches to sound-similarity definitions, and more amenable to deployment on different synthesizers as it does not require pre-training on large datasets.
-
-This work takes a \textit{differentiable} approach to iterative sound-matching, a natural extension of early works in iterative sound-matching combined with recent advances in Machine Learning (\gls{ML}) software~\cite{braun2024dac,paszke2019pytorch} and increase in computational power~\cite{moore2012introduction,schaller1997moore}.
-A differentiable approach requires differentiable loss and synthesizer functions, allowing the synthesizer parameters to be updated using gradient descent~\cite{goodfellow2016deep,boyd2004convex}.
-
-We discuss the history of sound-matching, and provide a nomenclature for different subtypes of sound-matching tasks.
-A number of weaknesses in current literature (see Section~\ref{sec:lacking}) are identified.
-The experiments in this work focus on two of these core issues, referred to as \LossSelect~and \SynthSelect.
-The former refers to the search for an optimal loss function, lack of novel algorithms, and the---perhaps needless---focus on outperforming state of the art (\gls{SOTA}), while the latter refers to the lack of diversity in synthesis methods.
-
-In the case of \LossSelect, there have been many works and experiments comparing the accuracy of loss functions and choosing the best performers~\cite{vahidi2023mesostructures,turian2020sorry,engel2020ddsp,uzrad2024diffmoog,han2023perceptual,masuda2021soundmatch,turian2020sorry,bruford2024synthesizer}.
-We note that these claims regarding the effectiveness of one function versus another have consistently been made in limited contexts that may not generalize to other settings.
-One such limitation is the synthesizers used, and ties into the problem of \SynthSelect, as typically only a single (and often, simple) synthesizer has been used for testing the effectiveness of loss functions.
-Furthermore, there is a lack of diversity in the underlying functions used by the synthesizers, with simple Frequency Modulation (\gls{FM}) being the most common method of synthesis~\cite{horner1993machine,vahidi2023mesostructures,mitchell2007evolutionary,caspe2023envelope}.
-
-Considering the interplay between the two problems, we design experiments that address them simultaneously.
-Testing the hypothesis that the performance of a similarity measure (or loss function) is influenced by other factors in the environment, particularly the method of synthesis.
-We combine 4 loss functions with 4 differentiable synthesizers, and compare their performance as measured by the similarity of the final output to the target.
-We evaluate the final similarity with two different automatic methods, as well as manual listening tests conducted by two of the authors.
-We compare the distributions of the evaluation scores in order to rank the loss functions from best to worst.
-The results of these experiments show that when accounting for different synthesizers, there is no consistent choice of sound similarity measure for optimal performance in sound-matching.
-
-We note that for some synthesizers, our novel use of ``dynamic time warping of envelopes'' as a loss function for differentiable sound-matching outperforms the standard methods of audio comparison.
-These results imply that rather than aiming to outperform the SOTA and finding a universal best performer, more effort could be placed towards diversifying the existing similarity measures of sound as well as conducting experiments using a variety of synthesis methods.
-Sound design is an inherently creative endeavor, and would likely benefit from a diversified range of solutions for different approaches.
-
-\section{Background}
-\label{sec:background}
-In this Section, we give a formal definition of sound-matching and the necessary components.
-These discussions lead to a nomenclature for types of sound-matching, as well as major perennial issues in the field.
-
-\subsection{Digital Audio Synthesis}
-\label{sec:diffSynth}
-A digital audio synthesizer is any software used for the creation and manipulation of digital audio.
-Digital synthesizers use signal processing chains with a variety of digital signal processing (\gls{DSP}) functions to create sounds.
-These functions are often parametric, and the set of parameters given to the synthesizer's chain of DSP functions is called a synthesizer program.
-
-The simplest form of DSP function is a sinusoidal tone. For example, consider:
-\[ x(n) = sin( 2\pi n T)\]
-
-where $T$ represents time in seconds, and $n$ is the frequency parameter.
-If $n$ is 1, then $x(n)$ would be a waveform with a frequency of 1 hertz, meaning that the value of $x(n)$ oscillates to the same point once every second.
-Waveform generators are called oscillators for this reason.
-
-% If $n$ is in the range of 30 to 20,000 (the upper limit is often lower for some), $x(n)$ could be played as an audible tone~\cite{smith2007mathematics,jilek2014reference}.
-
-Since the advent of digital signal processing in the 1960s~\cite{stranneby2004digital}, a wide variety of parametric audio generation functions have been proposed~\cite{lyons1997understanding,russ1999sound,shier2020spiegelib}.
-This includes oscillators, filters, equalizers, and envelopes, which can be used sequentially or in parallel~\cite{lyons1997understanding,russ1999sound}.
-Sound design with a synthesizer is done by modifying the parameters of the functions until reaching a desired output~\cite{roads1996computer,pinch2004analog}.
-
-\todo{A relatively small subset of these synthesis methods have been used for automatic sound-matching, notably additive and/or subtractive synthesizers~\cite{esling2019flow,yee2018automatic,mitchell2007evolutionary,salimi2020make}, physical modeling~\cite{riionheimo2003parameter}, and frequency/amplitude modulation (\gls{FM}/\gls{AM})~\cite{horner1993machine,vahidi2023mesostructures}.
-}
-
-Differential DSP (\gls{DDSP}) is a general term for applications that combine machine learning techniques such as gradient descent~\cite{goodfellow2016deep,boyd2004convex} with DSP~\cite{engel2020ddsp}.
-Implementing complex DSP functions in a differentiable manner can be challenging, and effective differentiable audio similarity measures require careful mathematical expression of the desired attributes of sound.
-These issues have likely contributed to the limited exploration of this domain.
-Perhaps the most relevant recent work is by Vahidi \textit{et al.}, featuring the comparison of two differentiable loss functions with a simple FM synthesizer~\cite{vahidi2023mesostructures}.
-
-\subsection{Sound Representation and Loss Functions}
-\label{sec:loss_funcs}
-A digital sound (or an audio signal) is a series of numbers~\cite{smith1991viewpoints,smith2007mathematics}.
-To compare two digital sounds, the two corresponding series are passed to a function that measures their similarity.
-Two signals can sound identical to our ears, without having any values in common~\cite{moore2012introduction}.
-This necessitates the use of proxy representations (or feature extractors) when comparing sounds automatically.
-Similarity between the target sound and the synthesizer output is then measured by some form of subtraction and summation of the proxy representations.
-
-In sound-matching, particularly in a Deep Learning (\gls{DL}) context~\cite{goodfellow2016deep}, the similarity function can also be called a \textit{loss} function, where the emphasis is on measurement and reduction of the distance between target and output.
-It is important to note that there is a close relationship between the loss function $L$ and the sound representation function $\phi$.
-$L$ is the result of a distance measure $d$ applied to the features extracted by $\phi$.
-
-\[
-L(\theta, t) = d\langle\phi(t),\phi(x)\rangle
-\]
-
-\noindent
-
-A proxy representation is the output of the function \( \phi \), which can be thought of as a feature extraction function that maps the sounds \( t \) and \( x \) to their respective representations. 
-The proportionality or distance metric $d$ has typically been the L1 distance~\cite{turian2020sorry,richard2025model}, calculated as the mean of the absolute difference between every point in the proxy representation~\cite{engel2020ddsp,vahidi2023mesostructures}:
-\[
-L(\theta, t) = \left\| \phi(t) - \phi(x) \right\|_1
-\]
-
-Here we discuss four methods of audio representation and the corresponding loss functions.
-Three of which have been used in previous works, and a novel method not previously used in iterative sound-matching.
-
-\subsubsection{Parameter Loss}
-A common measure of similarity in sound-matching is the distance between synthesizer parameter sets, referred to as ``P-Loss"~\cite{han2023perceptual}.
-Typically, for the implementation of P-Loss the parameter sets are treated as vectors in space, and L1 or L2 distance is applied.
-There are two major limitations to this approach: First, the target and output sound must be made by the same synthesizer; otherwise the parameter sets cannot be compared (see Section~\ref{sec:matching_types}).
-Second, the relationship between synthesizer parameters and the audio output is not linear~\cite{shier2020spiegelib,han2023perceptual,esling2019flow}.
-
-\subsubsection{Fourier Spectrograms}
-\label{sec:fourier_specs}
-Fourier-based transformations such as short-time Fourier transforms (\gls{STFT}), Mel-spectrograms, and Mel-frequency cepstral coefficients have been viewed as the de facto and state-of-the-art representation of audio~\cite{beauchamp2003error,mitchell2007evolutionary,yee2018automatic}, however, there are many issues associated with their use in sound-matching~\cite{turian2020sorry,vahidi2023mesostructures,han2023perceptual,uzrad2024diffmoog}.
-Fourier transformations allow for the conversion of a signal from the time-domain to the frequency domain.
-Audio spectrograms can be generated by segmentation of a piece of audio into overlapping windows followed by the application of Fourier transforms to each window.
-They are costly to compute, but provide a better temporal view of changes in frequency content~\cite{muller2007dynamic,smith2007mathematics}.
-There are different types of spectrograms that have a basis in Fourier transformations, but the most notable and commonly used is the STFT.
-
-What we call \textit{Fourier-based Spectrograms} are variations on the STFT approach.
-For example, Mel-Spectrograms \textit{bin} frequencies on a near-logarithmic scale to better match human perception of frequencies~\cite{muller2007dynamic}.
-Multi-scale spectrograms (\gls{MSS}) used in recent works are a simple weighted average of multiple spectrograms with different parameters such as window size, number of frequency bins, and hop size~\cite{engel2020ddsp,vahidi2023mesostructures}; this may provide some improvements at a higher computational cost~\cite{turian2020sorry,engel2020ddsp}.
-
-\subsubsection{Joint-Time Frequency Spectrum}
-Recent works have focused on the limitations of parameter and spectral loss functions in sound-matching~\cite{vahidi2023mesostructures,uzrad2024diffmoog}, seeking to create more effective general solutions for the comparison of audio.
-Noting the aforementioned weaknesses of comparing STFT spectrograms, Vahidi \textit{et al.} proposed differentiable Joint-Time Frequency Scattering (\gls{JTFS})~\cite{anden2015joint} as an alternative to spectrogram loss in sound-matching, and showed improved performance in sound-matching with differentiable chirplet synthesizers~\cite{vahidi2023mesostructures}.
-JTFS is the result of the application of a 2D wavelet transformation to the time-frequency representation of a signal~\cite{anden2015joint}.
-
-\subsubsection{Dynamic Envelope Warping}
-Dynamic Time Warping (DTW) is a method for measuring similarity between multi-dimensional time-series~\cite{rabiner1993fundamentals,muller2007dynamic,giorgino2009computing}.
-Given any two time-series $X = \{x_1,x_2,...,x_m\}$ and $Y = \{y_1,y_2,...,y_n\}$, we have indices $i\in\{1...m\}$ and $j\in\{1...n\}$ defining $X$ and $Y$.
-When the series are \textit{warped}, these indices change to expand or contract different portions of the series.
-To borrow the notation given by Muller~\cite{muller2007dynamic}, warped indices are a sequence $p=(p_1,...,p_L)$, where \(p_\ell = (m_\ell, n_\ell) \in [1 : m] \times [1 : n] \text{ for } \ell \in [1 : L]\), meaning that the indices for $X$ and $Y$ are reorganized under special conditions.
-In classical DTW, these conditions are \textit{monotonicity}, \textit{boundary matching}, and \textit{single step-size}.
-
-DTW measures the distance between the time-series \textit{after} alignments, typically using Euclidean distance.
-The distance between a time-series and shifted versions of itself would be 0, regardless of shift amount~\cite{tavenard.blog.dtw}.
-Additional rules can be imposed to keep alignments locally constrained~\cite{itakura1975minimum,sakoe1978dynamic}.
-% \textcolor{red}{DTW was originally used for signal comparison but using it the way we do is novel}
-
-\subsection{Automatic Sound Matching}
-\label{sec:sound_matching_definition}
-The problem of sound-matching has been tackled from a variety of perspectives as it involves a number of modular parts: the choice of synthesizer $g$, the target sounds of interest, the representation function $\phi$, and the heuristic for finding the optimal $\theta$.
-To give a formal definition of sound-matching, we expand the definition given by recent works~\cite{vahidi2023mesostructures,han2023perceptual}.
-The major components are as follows:
-\begin{itemize}
-    \item $g(\theta)$: Parametric audio synthesizer $g$ which takes on parameters $\theta$ 
-    \item $x$: The output of $g$, given a set of parameters $\theta$ or $g(\theta) = x$ 
-    \item $t$: The target sound which we want to replicate or imitate. 
-    \item $\phi$: Representation function or feature extractor. $\phi$ is applied to $x$ and $t$ to facilitate their comparison by the loss function.
-    \item $L$: Loss or error function. $L(\theta,t)$ is a measure of distance between $x$ and $t$. Often proportional to the subtraction of their representations, or $ \phi(x) - \phi(t)$
-\end{itemize}
-
-\highlight{It could be argued that the implementation of these components is not an optimization problem with a correct answer, but a \textit{creative} endeavor depending on the artist's needs}.
-Consider the hypothetical case where the goal might be to create an 8-bit~\cite{collins2007loop} version of a high quality, organic snare drum: in such a case, there is no optimal or ``correct" version of an 8-bit snare sound.
-
-\subsection{Sound Matching Types}
-\label{sec:matching_types}
-
-Sound-matching tasks can be categorized based on the choices of target sound $t$, the loss $L$ and representation function $\phi$, and the heuristic used for minimizing $\theta$.
-We will discuss the subtypes of sound-matching tasks in the upcoming sub-sections.
-
-\subsubsection{In-Domain vs Out-of-Domain}
-\label{sec:in-domain}
-The choice of domain depends on whether we want to use the same synthesizer for the target and output sounds, which we call \textit{in-domain}, or have target sounds that came from sources other than the synthesizer, or \textit{out-of-domain}.
-To paraphrase the description given by Masuda \textit{et al.}~\cite{masuda2021soundmatch}, if $g$, the synthesizer of choice, can accurately replicate the target sound $t$, or put differently, if $t$ itself is an output of $g$, then the sound-matching task is \textit{in-domain}.
-If $t$ is not an output of $g$, then the sound-matching task is \textit{out-of-domain}.
-In-domain tasks in general are simpler, and often there is a guarantee that there is a correct answer to the sound-matching problem, particularly if the goal is accurate replication of the sound.
-If the target sound is out-of-domain, replication is not guaranteed, and the goal becomes the \textit{imitation} of some aspect of sound.
-
-\subsubsection{Replication vs Imitation}
-Regardless of the domain, the generation goal can be \textit{replication} or \textit{imitation} of the target sound.
-In replication, the goal is to make an identical copy of the target sound.
-Imitation is an artistic pursuit and harder to define, since the goal is to make new sounds that only retain a subset of the target's sonic features.
-While closely related to the in-domain versus out-of-domain problem, the choice of replication versus imitation is more dependent on how the loss and representation functions are defined.
-
-% From a signal perspective, replication is a scenario where the goal is to find a $\theta$ such that $t-x=0$ where $x=g(\theta)$.
-From a sound perspective, this definition of replication is unnecessarily restrictive, as $t$ and $x$ can sound identical but contain very different values.
-For example, if $t$ and $x$ were periodic waveforms at an audible frequency $f$ with different phase values, their subtraction would not be 0, yet they would sound identical to our ears, as our ears are insensitive to phase values~\cite{moore2012introduction}.
-Representation functions such as STFT spectrograms or Mel-frequency cepstral coefficients (\gls{MFCC}) become viable in these cases, as they allow for separation and removal of the phase information within an audio signal.
-The \textit{sound replication} goal can be defined, in layman's terms, as ``$x$ sounds exactly like $t$".
-More formally, the goal is to find parameters such that $\phi(t)-\phi(x)=0$, where $\phi$ transforms an audio into a spectrogram that removes the phase information from sounds, but otherwise represents the sonic information as accurately as possible.
-
-% With \textit{sound imitation}, the goal is the replication of some aspect of the target sound.
-Instrument conversion is a common example of imitation.
-As an example, a note played by a violin is converted to the same note played by a piano.
-Conversion of even longer pieces has been made possible with large enough datasets~\cite{engel2020ddsp}.
-Imitation of sound envelopes (how the loudness changes over time) is another interesting idea~\cite{caspe2023envelope}.
-Transfer of other sound properties such as reverb and delay has also been a subject of research~\cite{engel2020ddsp}.
-
-% A target sound can have many properties of interest, and the imitation task is the transfer of a subset of those properties to other sounds.
-Relative to replication, imitation is a very open ended goal.
-From a research point of view, success in the replication task can be measured more reliably, often without the need for manual evaluation.
-With imitation, listening tests are needed to ensure quality of outputs, particularly in a supervised learning setting where learning is done by examples~\cite{salimi2020make,masuda2021soundmatch}.
-The need for subjective evaluations has perhaps contributed to the lack of research in imitation and out-of-domain sound-matching.
-
-\subsubsection{Supervised Versus Direct Optimization}
-\label{sec:optimization}
-
-The goal of sound-matching is to find the optimal parameters $\theta^*$ that minimize the loss between the synthesizer output and the target sound. 
-\[
-\theta^* = \arg\min_{\theta} L(\theta,t)
-\]
-
-The heuristics in past works can be broadly split into two categories:
-\begin{enumerate}
-    \item Supervised (or inference) methods which use datasets of sound and parameter pairs to describe the generation objective, often with a large number of input and output examples~\cite{engel2020ddsp,salimi2020make,yee2018automatic,esling2019flow}
-    \item Unsupervised or \textit{direct sound optimization} methods~\cite{horner1993machine,mitchell2007evolutionary,yee2018automatic,vahidi2023mesostructures}.
-Usually require at most one target example, and a loss function that measures the difference between the target and synthesizer outputs.
-\end{enumerate}
-
-\todo{
-Genetic algorithms (\gls{GA})~\cite{holland1992genetic} can be used for direct sound optimization, and have been the earliest and most common heuristic in sound-matching~\cite{horner1993machine,mitchell2007evolutionary,yee2018automatic}.
-These algorithms start with arbitrary parameter sets that can be treated as an evolving population where the genomes are the parameter values.
-The most fit members of the group are the parameters which perform the best in the loss function, and create a new generation of parameters via mutation (random change in subset of parameters) and crossovers (combination of parameter sets); this process repeats until the goal or a maximum number of generations is reached.}
-
- Differentiable settings have the benefit of making goal-oriented updates to the synthesizer parameters (the goal being the minimization of loss).
-There are also drawbacks: other than requiring more computation power, differentiable functions require careful implementation of operations that are continuous and numerically stable; this makes the implementation of signal processing functions quite difficult, contributing to the scarcity of works in this domain.
-
-% One important aspect of the heuristics is whether they imitate the iterative nature of manual sound design---that is, listening to the output and adjusting the synthesizer parameters recursively until the desired goal is reached.
-Supervised methods generate their final output in a single step, while direct sound optimization methods refine their outputs iteratively until termination, which resembles manual sound design more closely.
-
-\subsection{Iterative Sound Matching}
-\label{sec:iterative_sound_matching}
- \todo{Figure~\ref{fig:sound_design_loop_iterative} shows a general model for iterative sound-matching.
-To begin the process, the parameters are arbitrarily initialized (usually random generation), the similarity of the target and output is measured, and the parameter is \textit{optimized} with the goal of increasing the similarity, or alternatively, reducing the loss.
-This process repeats until termination.
-}
-\begin{figure}[ht]
-    \centering
-\begin{tikzpicture}[node distance=2cm, auto]
-
-    % Nodes
-    \node (start) [text centered] {\( g (\hat{\theta}) = x \)};
-    \node (L) [above of=start,right of=start, text centered] {\( L(\hat{\theta},t) {\ \propto \ } \phi(x) - \phi(t) \)};
-    \node (optimize) [below of=L,right of=L, text centered] {Optimize $\hat{\theta}$};
-    \node (new_theta) [below of=optimize,right of = start, text centered] {New \( \hat{\theta} \)};
-
-    % Highlight and arrow for the start node
-    \draw[->, very thick, red] (start) ++(-2,1.5) -- (start)
-        node[midway, below, align=center, sloped, color=red] {Start}
-        node[midway, above, align=center, sloped, color=red] {Random $\hat{\theta}$};
-
-    % Arrows with multi-line labels
-    \draw[->, bend left] (start) to node[midway, right, align=center] {} (L);
-    \draw[->, bend left] (L) to node[midway, right, align=center] {} (optimize);
-    \draw[->, bend left] (optimize) to node[midway, below, align=center] {} (new_theta);
-    \draw[->, bend left] (new_theta) to node[midway, left, align=center] {} (start);
-
-\end{tikzpicture}
-    % \caption{ Iterative approach to sound design.
-Synthesizer $g$ with arbitrary initialized parameters $\hat{\theta}$, creates sound $x$.
-The target sound $t$ is used as the goal.
-Parameters $\hat{\theta}$ are adjusted to minimize error (or loss) $L(\hat{\theta},t)$, where $L$ is proportional to the difference between the representations of $x$ and $t$.
-$\phi$ is the audio feature extractor, or representation function.}
-    \caption{ Iterative approach to sound design.}
-    \label{fig:sound_design_loop_iterative}
-\end{figure}
-
-In the upcoming ``Past Works'' section, we will discuss the evolution of prior sound-matching research, with \textit{differentiable iterative sound-matching} being the latest extension of the field.
-
-\begin{table*}[t]
-\centering
-\caption{Summary of select works in sound-matching.
-Gen.
-refers to whether the generations are done in 1-shot (generally, output by a neural network) or a search is conducted with the goal of iteratively getting closer to the target sound.}
-\setlength{\tabcolsep}{3pt} % Reduce horizontal spacing
-\renewcommand{\arraystretch}{1.05} % Reduce vertical spacing
-\scriptsize % Reduce font size
-\resizebox{\textwidth}{!}{
-\begin{tabular}{|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{1.2cm}|>{\raggedright\arraybackslash}m{1.2cm}|>{\raggedright\arraybackslash}m{1cm}|>{\raggedright\arraybackslash}m{1cm}|}
-\hline
-\textbf{Work} & \textbf{Synthesis} & \textbf{Loss} & \textbf{Heuristics} & \textbf{Goal} & \textbf{Domain} & \textbf{Year} & \textbf{Gen. } \\
-\hline
-Horner \textit{\textit{et al.}}~\cite{horner1993machine} & Non-diff FM & STFT (McAulay-Quatieri) & Direct & Replicate & In & 1991 & Iter \\
-\hline
-Mitchel \textit{\textit{et al.}}~\cite{mitchell2007evolutionary} & Non-diff FM & Spectral Relative & Evolutionary & Replicate & In (Contrived) & 2007 & Iter \\
-\hline
-Yee-King \textit{\textit{et al.}}~\cite{yee2018automatic} & Non-diff VST & P-Loss (for Supervised), Spectral (MFCC) & Supervised \& Evolutionary & Replicate & In & 2018 & Both \\
-\hline
-Esling \textit{\textit{et al.}}~\cite{esling2019flow} & Non-diff VST & Spec MSS/SC P-Loss & Supervised \& Modeling & Replicate & In* \& Out & 2019 & 1-shot \\
-\hline
-Shier \textit{\textit{et al.}}~\cite{shier2020spiegelib} & Custom, non-diff & FFT \& Spectral & Supervised \& Direct & Replicate & In & 2020 & Both \\
-\hline
-Salimi \textit{\textit{et al.}}~\cite{salimi2020make} & Non-diff & STFT and Envelope & Supervised & Imitate & Out & 2020 & 1-shot \\
-\hline
-Masuda \textit{\textit{et al.}}~\cite{masuda2021soundmatch} & DDSP & P-Loss initially, fine-tune with Spec. & Supervised & Both & In \& Out & 2021 & 1-shot \\
-\hline
-Han \textit{\textit{et al.}}~\cite{han2023perceptual} & NN-encoder $\rightarrow$ physical model & PNP (approx. of STFT) & Supervised & Replicate & In & 2023 & 1-shot \\
-\hline
-Vahidi \textit{\textit{et al.}}~\cite{vahidi2023mesostructures} & Differentiable FM & JTFS/MSS & Direct (grad desc.) & Replicate & In & 2023 & Iter \\
-\hline
-Uzrad \textit{\textit{et al.}}~\cite{uzrad2024diffmoog} & DDSP chains & Synth. Chain and P-Loss & Supervised & Replicate & In \& Out & 2024 & 1-shot \\
-\hline
-\end{tabular}
-}
-
-\label{tab:summary}
-\end{table*}
-
-\section{Past Works}
-\label{sec:related_works}
-Table~\ref{tab:summary} contains an overview of relevant literature, covering important works in sound-matching.
-adsfEarlier works in this domain focused on FM synthesis and genetic algorithms (or direct optimization methods)\cite{justice1979analytic,horner1993machine,mitchell2007evolutionary}.
-Along with the exponential increase in computation power, later works utilized a wider range of approaches to parameter search, estimation, and audio synthesis; particularly with supervised learning and neural networks (\gls{NN})~\cite{yee2018automatic,engel2020ddsp,esling2019flow}.
-
-\subsection{Earlier Works}
-FM synthesizers are simple to implement, yet very expressive~\cite{chowning1973synthesis}.
-Perhaps the earliest foundational work in sound-matching is the analytical approach by Justice~\cite{justice1979analytic} toward the decomposition and recreation of sounds with FM synthesis using one carrier, one modulator, and the corresponding envelopes:
-\[ f(t) = I_c(t) \cos(\omega_c t + I_m(t) \cos(\omega_m t))
-\]
-Where $f(t)$ is the output signal, $I_c(t)$ is the carrier amplitude, $\omega_c$ is the carrier frequency, and $I_m(t)$ and $\omega_m$ are the envelope and frequency for the modulator~\cite{justice1979analytic}.
-
- The sound-matching works that followed used other heuristics for finding the synthesizer parameters, yet the simple FM synthesizer structure remained unchanged.
-Horner \textit{et al.}~\cite{horner1993machine} used GAs for re-synthesis of sounds with an FM synthesizer using one modulator and one carrier oscillator, and the McAulay-Quatieri method~\cite{mcaulay1986speech} for measuring loss.
-
- % MQ calculates the STFT and marks the start and end of \textit{trends} in a sound with the assumption that the harmonic content belongs to trends that have a beginning and an end.
-
-Introduction of more complex models of synthesis such as wavetables~\cite{horner2003auto} and physical modeling~\cite{riionheimo2003parameter} rendered interesting results, but led to further questions about the nature and goals of sound-matching.
-Mitchell and Creasy noted the difficulty in distinguishing between ``inefficiency of the optimization engine'' and ``synthesizer limitations'' as the cause of failure or success in sound-matching~\cite{mitchell2007evolutionary}.
-They offered a \textit{contrived methodology} for finding the best evolutionary method for sound-matching using their FM synthesizer.
-The methodology postulates that the best search heuristic for in-domain search (where an exact target exists, and a wide range of targets can be produced by sampling from different points in the parameter space) would also be the best-performing heuristic for out-of-domain search on a dataset of muted trumpet tones~\cite{opolko1989mcgill}; however, testing was not extensive and yielded some contradictory results.
-As they noted, modifications to the synthesizer, loss function, and sound domain would lead to a problem with an entirely new search space~\cite{mitchell2007evolutionary}.
-
-% Testing sound-matching experiments can be difficult, in part due to the large number of parametric components and navigation of complex non-linear systems.
-The issue with finding the best approaches to sound-matching is exacerbated when slight changes to the pipeline render previous results moot.
-Perhaps due to these reasons, the contrived methodology was not extensively tested by Mitchell and Creasy, who only provided a single experiment where the best evolutionary strategy found by in-domain search would also perform well in out-of-domain synthesis.
-\subsection{Introduction of Deep-Learning}
-Recent years have seen more works in sound-matching using machine learning techniques.
-In 2018, Yee-King \textit{et al.} rendered 60,000 audio-parameter pairs from the \textit{Dexed} \gls{VST} synthesizer~\footnote{https://asb2m10.github.io/dexed/}, and showed that NNs trained on this dataset can outperform GA and hill-climber (\gls{HC})~\cite{hoffmann2000heuristic} methods in rendering speed, with slight improvements in MFCC error (used as an objective performance test).
-The speed improvement appears trivial, considering the iterative nature of GAs when compared to offline training of supervised models.
-For GAs and HC optimizer, MFCCs were used as a measure of performance as well as a loss function.
-For training the networks, P-Loss was used, since differentiable MFCCs were not possible in their pipeline.
-Importantly, informal hearing tests revealed that the performance of even the best NN model was unsatisfactory, possibly due to the complex nature of the synthesizer, which features 155 parameters.
-
-% Programming synthesizers and querying sounds by imitating human vocals has also been a subject of interest~\cite{Cartwright2014SynthAssistQA,siamesevocalimitation2018}.
-For example, Cartwright and Pardo created a framework for searching pre-made patches for a synthesizer to best imitate vocals using audio features such as pitch, spectral centroid, and loudness~\cite{Cartwright2014SynthAssistQA}.
-
-Esling \textit{et al.} noted the issue of using P-Loss in sound-matching~\cite{esling2019flow}, considering the non-linear mapping between parameters of a synth and the audio output.
-In their approach, dubbed as ``flowsynth'', they used a large parameter-to-audio dataset for a commercial \gls{VST} synthesizer.
-In flowsynth, audio is encoded to an auditory space, using a NN autoencoding approach~\cite{hinton2006reducing}.
-The encodings in the auditory space can then be converted to the parameter space, but only for that particular synthesizer.
-Using P-Loss and Spectral Convergence (Frobenius norm of the spectrograms) as metrics, flowsynth outperformed other NN approaches that did not use an auditory to parameter space conversion, but formal hearing tests were not conducted.
-
-Masuda \textit{et al.} also applied supervised learning to synthesizer parameter estimation~\cite{masuda2021soundmatch}.
-Their work highlights the issue of non-linearity in parameter-to-synthesizer outputs and out-of-domain search.
-This work uses a differentiable subtractive synthesizer (two oscillators and an LP filter).
-A NN model was pre-trained using an in-domain dataset of randomly selected parameters and P-Loss.
-After training, the model was fine-tuned using 20,000 out-of-domain sounds from the NSynth dataset~\cite{engel2017neural} and multi-scale spectrogram loss~\cite{engel2020ddsp}.
-This approach proved more effective---i.e, achieved lower multi-level spectral difference in out-of-domain tests---than baseline models, which were either not exposed to out-of-domain sounds or trained exclusively with P-Loss.
-Subjective hearing tests were conducted, showing a preference for the fine-tuned model~\cite{masuda2021soundmatch}.
-Masuda \textit{et al.} later extended this work with semi-supervised learning, highlighting significant gaps in in-domain and out-of-domain performance~\cite{masuda2023improving}.
-
-Rather than focusing on a particular implementation, Shier \textit{et al.} presented Spieglib, a library for implementation of sound-matching pipelines~\cite{shier2020spiegelib}.
-This library provides different choices for DNNs, GAs, synthesizers, and feature extractors.
-Shier \textit{et al.} presented an experiment with a similar setup to Yee-King \textit{et al.}~\cite{yee2018automatic}, however, they found a genetic algorithm to be the best performing.
-
-% Rather than focusing on a particular implementation, Shier \textit{et al.} presented Spieglib, a library for implementation of sound-matching pipelines (which Shier \textit{et al.} referred to as ``automatic parameter search'')~\cite{shier2020spiegelib}.
-This library provides extensible choices of deep-learning estimators with \textit{tensorflow}~\cite{abadi2016tensorflow}, genetic estimators with \textit{DEAP}~\cite{DEAP_JMLR2012}, feature extractors with \textit{librosa}~\cite{mcfee2015librosa}, and different choices for synthesizers, evaluators, dataloaders.
-Shier \textit{et al.} provide an example experiment using this library.
-This example has a setup similar to Yee-King \textit{et al.}~\cite{yee2018automatic}, where the \textit{Dexed} VST was used and different genetic and NN estimators were compared by using MFCC difference as the minimization objective.
-Depending on the NN architecture, MFCC or STFT spectrograms were used as the input.
-Unlike previous comparisons of NN and genetic algorithms, the best estimator was the multi-objective non-dominated sorting genetic algorithm III (NSGA-III).
-This algorithm had multiple minimization objectives: 13-band MFCC, STFT, and five spectral features.
-Shier \textit{et al.} report that all models performed relatively well at imitation of harmonic distributions, but the multi-objective function was the only optimizer which did \textit{not} struggle with matching temporal envelopes of the audio spectrum~\cite{shier2020spiegelib}.
-
-Differentiable loss functions that use spectrogram differences can be computationally expensive.
-To mitigate this, Han \textit{et al.}~\cite{han2023perceptual} introduced ``perceptual-neural-physical loss'' (PNP).
-PNP is an approximation of loss functions; specifically, a loss function that uses the L2 norm of the difference between the features of two sounds, or $||\phi(t) - \phi(x)||^2_2$, where $\phi$ could be a spectrogram or JTFS function.
-PNP loss functions are fast and differentiable, but require training and parameter estimation.
-A Riemannian metric M needs to be calculated for the minimization of locally linear approximation of the ``true" spectral loss function~\cite{han2023perceptual}.
-\[
-\|\phi(t) - \phi(x)\|_2^2 = \langle \tilde{\theta} - \theta | \mathbf{M}(\theta) | \tilde{\theta} - \theta \rangle + O(\|\tilde{\theta} - \theta\|_2^3). \tag{4}
-\]
-
-During the training phase, this metric is calculated alongside the neural parameter estimator.
-After training, PNP can be used as a fast approximation of computationally demanding loss functions such as JTFS~\cite{han2023perceptual,han2024learning}, paired with FM synthesizers and a differentiable physical model of a drum head~\cite{smith2010physical}.
-
-Uzrad \textit{et al.}~\cite{uzrad2024diffmoog} took another unique approach to sound-matching: using a differentiable \textit{synthesis chain} of DSP generators and effects and a loss function that combines P-Loss with a \textit{signal-chain loss}.
-The synthesizer is a customizable chain of effects, which feed one output as input to the next step of the chain; signal-chain loss compares the parameter and output difference at every output step in the chain~\cite{uzrad2024diffmoog}.
-Possible chain functionalities are FM/AM, Low-Frequency Oscillators (\gls{LFO}), filters, and envelopes.
-Like the results shown by Masuda \textit{et al.}~\cite{masuda2021soundmatch}, better out-of-domain results were achieved when pre-trained on in-domain data and fine-tuned using out-of-domain NSynth data~\cite{engel2017neural}.
-
-% Uzrad \textit{et al.}\cite{uzrad2024diffmoog} used log-spectral distance (LSD) as their loss function, where   
-% $\text{LSD} = \| \log(S(t)) - \log(S(x)) \|_F$, S is the mel-spectrogram of the sounds and F is the Frobenius norm~\cite{golub2013matrix}. The Frobenius norm of a matrix $A$ is defined as
-% $ \|A\|_F = \sqrt{\sum_{i,j} |a_{ij}|^2}$.
-Here, $A$ would be the 2D matrix resulting from subtracting two log-spectrograms.
-Along with other works, \textbf{non-convexity of loss surfaces when applying gradient optimization} is a noted issue~\cite{turian2020sorry,vahidi2023mesostructures,uzrad2024diffmoog}.
-
-\subsection{What is Lacking In the Field}
-\label{sec:lacking}
-%   % First, we need to clearly define the characteristics we are looking for with non-generic loss functions, otherwise the chances of finding the sound(s) of interest is slim.
-Second, we need to move beyond direct replication, and focus on imitating target sounds that are outside the synthesizers capabilities.
-Third, more complex methods of synthesis need to be explored for iterative sound-matching.
-Fourth, we need to find optimization methods that can effectively navigate the the sinusoidal nature of loss function landscapes.
-
-% Based on the literature reviewed, we find several gaps in past research that makes the application of sound-matching for sound-designers difficult.
-Thus far, replication of in-domain sounds has been the main subject of research.
-With in-domain replication, we already have the target sound \textit{and} the program that makes it.
-This simplifies the problem of measuring performance and is the reason why this type of sound-matching research is by far the most common.
-However, practical application of sound-matching requires answers to much harder problems.
-Sound-matching would be an interesting creative tool if we could use it to either:
-% \begin{enumerate}
-%  \item \textbf{Replicate out-of-domain sounds}: this would give us the parameters that could closely approximate a target sound, and we can modify the sound further by modulating the synthesizer. 
-% \item \textbf{Imitate in-domain or out-of-domain sounds}: This would allow for applying the characteristics of one sound to another. 
-% \end{enumerate}
-
-% We need to be aware of the ``open set recognition'' (\gls{OSR}) problem here \cite{mundt2019open,gers2000learning}.
-OSR is a well documented issue in DL that manifests when the set or category that the network is supposed to reject or accept cannot be explained via examples.
-With imitation or replication of out-of-domain sounds, we can expect issues related to OSR to occur, since target and output sounds would be from different domains~\cite{salimi2021percussive}.
-
-% Perhaps one method of addressing the OSR problem is the use of representations (or feature extraction) methods that reduce sounds to a feature vector in the same domain.
-For example, a representation that only looks at the envelope of sounds is far less likely to be subject to the OSR problem.
-\textcolor{highlight}{We highlight the lack of non-generic loss functions as a major weakness in previous works and an interesting area to explore}.
-
-% Another weakness \textcolor{highlight}{is that the synthesis methods in past works are not representative of the common methods used by sound designers and modern synthesizers}.
-Simple FM/AM synthesizers have generally been the go-to methods of sound-matching.
-We believe that general statements regarding the effectiveness of an approach to iterative sound-matching (particularly in regards to a loss function) can only be reasonably made in the context of the subset of the tested methods of synthesis.
-% \textcolor{highlight}{The simplicity of parameter optimization approaches} is another area of weakness in past works.
-Classic genetic algorithms update the parameters randomly, while in recent differentiable settings, simple gradient descent methods are used.
-Past works have provided issues which arise in differentiable parameter estimation due to periodic loss functions.
-Perhaps the use of reinforcement learning (\gls{RL}), or other algorithms that continually learn from their environment, could help with this problem.
-
-% \subsection{Defining The Major Issues}
-% \label{sec:addressing_problems}
-
-Having looked at the current available literature, we note four major areas of weakness in sound-matching.
-We target the first two issues in this work, while the latter two issues are left for future work.
-\begin{enumerate}
-    \item \LossSelect: Is there a \textit{best performing} loss function, or is the selection of the loss function dependent on factors such as the sound domain, synthesizer capability and the desired characteristics of the output?
-    \item \SynthSelect: There is a lack of diversity in the synthesis methods used in sound-matching.
-The use of different DSP functions for iterative sound-matching and their effects on the outcome or the interactions with the loss function have not been tested.
-    \item \PeriodicLoss: The sinusoidal nature of loss function landscapes is a problem that appears frequently.
-In differentiable settings, this problem causes gradient descent updates to local minima.
-    \item \OutDomain: Sound-matching with out-of-domain sounds is an unexplored area, yet a necessary one for practical applications for sound designers.
-\end{enumerate}
-
-\section{Programs, Methodology, and Results}
-\label{sec:progs_methodology_results}
-Towards addressing the problems of \LossSelect{} and \SynthSelect{}, we compare the performance of iterative sound-matching experiments, where four differentiable loss functions are applied to four differentiable synthesizers.
-What we mean by ``performance'' is the similarity of the final output to the target, as measured by P-Loss, MSS, and---most importantly---manual listening tests.
-
-\textbf{We hypothesize that the relative performance of loss functions for sound-matching is influenced by the method of synthesis}.
-If true, this would strengthen the idea that the selection of the loss function and synthesis method is not an optimization problem, but a subjective choice depending on the needs of the sound designer.
-
-% In Section~\ref{sec:methodology} we lay out the experiment methodology.
-Section~\ref{sec:loss_implementation} discusses the details of implementation for the loss functions.
-Section~\ref{sec:programs} covers the characteristics of the programs, and finally Section~\ref{sec:results_posthoc} shows that the effect of loss function choice on the success of iterative sound-matching is dependent on the signal process chain of interest.
-
-\subsection{Experiment Methodology}
-\label{sec:methodology}
-% We have highlighted the lack of attention given to the selection of loss functions in sound-matching, and noted that the few existing works addressing this issue evaluate their proposed losses using only simplistic sound synthesis methods.
-\textcolor{highlight}{We hypothesize that the best performing loss function is influenced by the synthesis techniques utilized in the synthesizer}.
-This hypothesis targets the issues of \LossSelect~and \SynthSelect.
-
-We test four different loss functions and compare their success in sound-matching using four different synthesizers.
-We verify the success of the sound-matching process using two automatic measures of sound similarity (MSS and P-Loss) and a manual blinded hearing test conducted by two of the authors where Likert scores~\cite{jebb2021review} of 1-5 are assigned to a subset of final outputs and the corresponding targets.
-
-An \textit{experiment} is one iteration of sound-matching given a loss function, target sound,
-and synthesizer program.
-When the final iteration is complete, a score is given (automatically with an algorithm or with manual listening) to an experiment that measures its success in replicating the target sound.
-Consider the following simplified scenario: we have two loss functions, $L_1$ and $L_2$, and a single synthesizer program $P$ with a target parameter set $\theta^*$ where the target sound is $t$, or $P(\theta^*) = t$.
-We run a large number (in this case, 300 for each loss function) of experiments where we randomly set the initial ($\theta_0$) and target parameters ($\theta^*$) of $P$ and run an iterative sound-matching loop for each loss function, with maximum number of iterations set to 200.
-When the loop is terminated, we measure the similarity of the final output sound $x$ (where $P(\theta_{199}) = x$) to $t$, or if using P-Loss we compare $\theta_{199}$ to $\theta^*$.
-Having 300 experiments gives us a distribution of losses to determine whether $L_1$ or $L_2$ led to better performance.
-This simplified example leads us to our methodology which has four loss functions, four synthesizer programs, and three different evaluation methods of scoring sound similarity.
-
-\subsubsection{Automatic and Manual Evaluation}
-\label{sec:evaluation_manual_auto}
-The two automatic evaluation methods used here are P-Loss and MSS, as described in Section~\ref{sec:loss_funcs}.
-For P-Loss, the parameters are normalized between 0-1 based on the valid ranges defined in the Faust program, and the L1 distance between the normalized parameters is calculated.
-MSS is computed using a hop size of $100$ samples, and FFT window lengths of $(512, 1024, 2048, 4096)$.
-
-Automatic sound evaluation methods are often unreliable, as their alignment with the subjective judgments of human listeners is uncertain.
-Therefore, listening tests are conducted by randomly sampling 40 experiment results for each program and loss function pair.
-Two of the authors then assign blinded similarity scores to the outputs and targets using a 5-point Likert scale (1 = no similarity, 5 = near identical)~\cite{jebb2021review}.
-Using Spearman's rank correlation~\cite{spearman1987proof,rebekic2015pearson}, we found a very strong rate of agreement between the human listeners (see Section~\ref{sec:consistency_in_rankings}), a good indicator that the scores assigned by manual listeners can be treated as the ground truth.
-We then combine the ranks assigned by both authors, giving us 80 ranks per program and loss function pair.
-
-% Agreement between the assigned scores is measured using Spearman’s rank correlation, which provides both a correlation coefficient ($\rho$) ranging from –1 (perfect negative correlation) to 1 (perfect positive correlation), and a p-value testing the null hypothesis of no correlation~\cite{spearman1987proof,rebekic2015pearson}.
-Across all programs, the correlation was very strong ($\rho = 0.86$, $p < 10^{-180}$).
-Per-program correlations were also very strong: $\rho = 0.71$ for \BPNoise{} ($p < 10^{-25}$), $\rho = 0.64$ for \AddSineSaw{} ($p < 10^{-19}$), $\rho = 0.84$ for \AmpMod{} ($p < 10^{-43}$), and $\rho = 0.85$ for \FMMod{} ($p < 10^{-44}$).
-This high rate of agreement is a good indicator that the scores assigned by manual listeners can be treated as the ground truth.
-
-% For each program and loss function, the blind manual hearing test gives us 80 ranks, which are given by two of the authors to a set of 40 randomly selected experiment results.
-This gives 160 pairs of ranks per program, and 640 for all programs.
-We computed Spearman correlations between the two reviewers.
-
-\subsubsection{Determining Best Performers}
- Figure~\ref{fig:posthoc_evaluation} is a visualized summary of how the loss functions are ranked.
-We have four loss functions and four synthesizer programs.
-The maximum number of iterations set to 200, and 300 experiments for each loss/synthesizer combination.
-Each synthesizer program is paired with four loss functions and 300 experiments are conducted and evaluated automatically with P-Loss and MSS, and manually with Likert scores.
-
-For each program and loss function pair, we have three distributions that can be used to rank the loss functions.
-Two with 300 automatically assigned similarities (P-Loss and MSS) and one with 80 Likert scores (combining the 40 ranks assigned by each author).
-For consistency and statistical robustness, the distributions are upsampled to 1000 values using bootstrapping~\cite{tibshirani1993introduction}.
-Bootstrapping gives an estimation of the distribution for the mean performance of each experiment.
-$k$ (set to 1000) samples of $n$ values (set to 100\% of the values) are taken with replacement from the empirical distribution (list of 300 performance values for MSS and P-Loss or 80 for manual rankings).
-The mean performance is calculated for each of the $k$ samples.
-The $k$ estimates of the mean performance give us a bootstrapped distribution for each loss function.
-
-For each program, the loss functions are ranked using the non-parametric Scott-Knott (\gls{NPSK}) test~\cite{tantithamthavorn2017mvt,tantithamthavorn2018optimization}, which ranks the loss functions from rank 1 (best, or highest distribution values) to maximum rank of 4 (worst, with lowest distribution values).
-NPSK may determine that multiple functions belong to the same rank, in which case the maximum rank would be less than 4.
-As mentioned before, lower P-Loss and MSS values indicate higher similarity, therefore (for consistency with manually assigned ranks) the reciprocal of their output is used ($x^{-1}$, if $x$ is the P-Loss or MSS value), where higher values indicate greater similarity.
-
-\input{extra_figures/evaluation_graph.tex}
-
-\subsubsection{Training Parameters and Gradient Calculations}
- % Previous work utilized gradient fields for analysis of the algorithm performance~\cite{vahidi2023mesostructures}.
-Gradient fields are not the center of our analysis, but we include them in Appendix~\ref{appendix:gradient_fields}.
- For updating the synthesizer parameters (or weights), we use RMSProp, which operates similar to stochastic gradient descent (SGD)~\cite{goodfellow2016deep}, with the caveat that the gradients of each weight are scaled by the root-mean-square of past gradients of that weight.
-Based on some initial test runs, we used an arbitrary fixed learning rate of 0.045 for all experiments (learning rates used in Vahidi \textit{et al.}~\cite{vahidi2023mesostructures} is unknown).
-The maximum number of iterations is set to 200, where, based on our observations, the parameters have either irrecoverably diverged outside the acceptable ranges, or are stuck at a local minimum.
-Additionally, the gradients are large, and calculated backwards sequentially through the signal processing chain.
-This backward calculation resembles the \textit{exploding gradients} problem in recurrent neural networks~\cite{gers2000learning}.
-Gradient clipping~\cite{goodfellow2016deep} is preemptively used to ensure that the  $\ell_2$ norm of all gradients does not exceed the threshold of 1.
-% \mathbf{\gamma}_i^{\text{clipped}} = \mathbf{\gamma}_i \cdot \mathcal{S}
-% \]
-% Where $\mathcal{S}$ is a value less than or equal to 1, depending on whether the norm of the gradients exceeds the threshold $c$.  
-% \[
-% \mathcal{S} = \frac{c}{\max(\|\mathbf{\gamma}\|_2, c)}
-% \]
-
-\subsection{Loss Function Implementation Details}
-\label{sec:loss_implementation}
-\subsubsection{STFT Losses}
-Due to their ubiquity and lower cost of gradient calculation, we use STFT as the basis of two loss functions.
-As discussed in Section~\ref{sec:matching_types}, the L1 or L2 distances are the most common methods of comparing different spectrograms, however, there are situations where this might lead to shortcomings.
-An example given by Vahidi \textit{et al.}~\cite{vahidi2023mesostructures} is two chirplets (tones that increase in pitch) starting and ending at the same frequencies.
-If one of these tones is slightly shifted in time, then the spectrograms will no longer overlap at any point, despite their sonic similarity (similar to the idea that a linear equation $f(x) = ax$ can never overlap with a slightly shifted version of itself, or $f(x) = ax +\epsilon$).
-In this example, we do not know if this is an issue of the spectrograms, or the L1 measure used for their comparison.
-In fields such as computer vision, Structural Similarity Index (SSIM)~\cite{wang2004imagesssim,wang2009mean} and Scale-Invariant Mean Squared Error (SIMSE)~\cite{barron2014shapessimse} have been used as alternatives that improved performance in various tasks relative to L1.
-
-We define the \LoneSpec~and \SIMSESpec~functions as the application of L1 and SIMSE to the STFT spectrograms.
-The STFT spectrogram function uses 512 FFT bins, window size of 600 samples, and hop length (how many samples the window shifts) of 100 samples.
-The L1 and SIMSE implementations are differentiable.
-
-\subsubsection{JTFS Loss}
-The \JTFS~loss function is the application of L1 difference to the JTFS representations of two sounds~\cite{vahidi2023mesostructures}.
-The code used for the JTFS transformation is the differentiable implementation of a 1-dimensional JTFS function provided by Andreux \textit{et al.}~\cite{kymatio}.
-
-\subsubsection{Soft-DTW Loss}
-We use the soft-DTW function, which is differentiable and --- depending on its parameters --- not shift-invariant~\cite{cuturi2017soft,janati2020spatio,tavenard.blog.softdtw}.
-The loss function \DTWEnv~is the application of the soft-DTW function to the envelope of the two sounds being compared~\cite{lyons1997understanding}.
-The envelope is calculated by creating the STFT spectrogram of a sound (the same process used in Section~\ref{sec:fourier_specs}) and summing the values at each timestep.
-
-\subsection{The Synthesizers}
-\label{sec:programs}
-The synthesizer programs are meant to be simple examples that test the building blocks of digital sound synthesis.
-Subtractive, additive, and FM/AM synthesis are three of the most common techniques in sound design~\cite{smith1991viewpoints}.
-In subtractive synthesis, frequencies are removed from a sound via digital filters.
-In additive synthesis, complex sounds are created via the linear combination of simpler sounds~\cite{lyons1997understanding,smith2007introduction}.
-As discussed previously, FM/AM synthesis refers to the general technique of modulating the frequency or amplitude of a waveform (the carrier) by another waveform (the modulator).
-
-For each program, we provide the Faust code~\cite{orlarey2009faust}, which can be run in the online IDE~\footnote{\url{https://faustide.grame.fr/}}.
-Faust is a functional language for audio synthesis that can succinctly define signal processing chains.
-Following Braun's methodology~\cite{braun2024dac}, programs are first defined in Faust, then converted (or \textit{transpiled}) to differentiable Jax functions using the DawDreamer library\footnote{\url{https://github.com/DBraun/DawDreamer}}.
-
- Once the programs are transpiled to differentiable synthesizers, the iterative sound-matching procedure is:
- \begin{enumerate}
-    \item Initialize $\theta^*$ and $\hat{\theta}$, i.e., random generation of target and initial parameters uniformly over a predefined range
-    \item Generating the output of the synthesizer with $\hat{\theta}$ (the length of the output is set to 1 second with sample-rate of 44100 Hz) 
-    \item Calculating the loss between the target and output
-    \item Applying gradient updates to the synthesizer parameters
-    \item Repeating the second step with the updated parameters $\hat{\theta}$, until maximum number of iterations has been reached
- \end{enumerate}
-
-% A general overview of the programs and the predefined parameter ranges is given in Table~\ref{tab:programs}.
-
-In the following subsections, we describe the programs and their parameters in more detail, followed by the methodology in Section~\ref{sec:methodology} and results in Section~\ref{sec:results_posthoc}.
-
-% \begin{table*}[htbp]
-% \centering
-% \begin{adjustbox}{max width=\textwidth}
-% \begin{tabular}{|c|c|c|c|l|}
-% \hline
-% \textbf{ID} & \textbf{Parameters} & \makecell{\textbf{Param. 1}\\\textbf{Range}} & \makecell{\textbf{Param. 2}\\\textbf{Range}} & \textbf{Description} \\
-% \hline
-% \BPNoise & hp\_cut, lp\_cut & 50-1000 & 1-120 & Band-pass of Noise signal \\
-% \hline
-% \AddSineSaw & saw\_freq, sine\_freq & 20-1000 & 20-1000 & Addition of sine and saw waves \\
-% \hline
-% \AmpMod & amp, carrier & 1-20 & 20-1000 & LFO on amplitude of noise osc \\
-% \hline
-% \FMMod & amp, carrier & 0-5 & 0-4 & LFO on amplitude of saw osc \\
-% \hline
-% \end{tabular}
-% \end{adjustbox}
-% \caption{Overview of Program Parameters and the range of values each parameter can take.}
-% \label{tab:programs}
-% \end{table*}
-
-\begin{lstlisting}[caption={\BPNoise}, label={lst:program0}, language=Faust,
-                  float, floatplacement=!H, xleftmargin=1em, xrightmargin=0.5em, firstnumber=0, aboveskip=0em, belowskip=-1em]
-import("stdfaust.lib");
-lp_cut = hslider("lp_cut",900,50,1000,1);
-hp_cut = hslider("hp_cut",100,1,120,1);
-process = no.noise:fi.lowpass(3,lp_cut):fi.highpass(10,hp_cut);
-\end{lstlisting}
-
-\begin{lstlisting}[caption={\AddSineSaw}, label={lst:program1},language=Faust,float,floatplacement=!H,xleftmargin=1em,xrightmargin=0.5em,firstnumber=0,aboveskip=0em, belowskip=-1em]
-import("stdfaust.lib");
-saw_freq = hslider("saw_freq",800,20,1000,1);
-sine_freq = hslider("sine_freq",300,20,1000,1);
-sineOsc(f) = +(f/ma.SR) ~ ma.frac:*(2*ma.PI) : sin;
-sawOsc(f) = +(f/ma.SR) ~ ma.frac;
-process = sineOsc(sine_freq)+sawOsc(saw_freq);
-\end{lstlisting}
-
-\begin{lstlisting}[caption={\AmpMod}, label={lst:program2},language=Faust,float,floatplacement=!H,xleftmargin=1em,xrightmargin=0.5em,firstnumber=0,aboveskip=0em, belowskip=-1em]
-import("stdfaust.lib");
-amp = hslider("amp",0.5,0,5,0.01);
-modulator = hslider("modulator",0.5,0,4,0.01);
-sineOsc(f) = +(f/ma.SR) ~ ma.frac:*(2*ma.PI) : sin;
-process = no.noise*sineOsc(modulator)*amp;
-\end{lstlisting}
-
-\begin{lstlisting}[caption={\FMMod}, label={lst:program3},language=Faust,float,floatplacement=!H,xleftmargin=1em,xrightmargin=0.5em,firstnumber=0,aboveskip=0em, belowskip=-1em]
-import("stdfaust.lib");
-carrier = hslider("carrier",100,20,1000,1);
-amp = hslider("amp",6,1,20,1);
-sineOsc(f) = +(f/ma.SR) ~ ma.frac:*(2*ma.PI) : sin;
-sawOsc(f) = +(f/ma.SR) ~ ma.frac;
-process = sineOsc(amp)*sawOsc(carrier);
-\end{lstlisting}
-
-\subsubsection{\BPNoise}
-\label{sec:program0}
-\BPNoise{} is a bare bones example of subtractive synthesis using digital filters~\cite{smith2007introduction}.
-Despite the ubiquity of subtractive synthesis in practical sound design, it has rarely been tested in differentiable sound-matching.
-In this program, a noise signal is fed through a band-pass (BP) filter.
-This removes the frequencies outside the low-pass (LP) and high-pass (HP) cutoffs.
-The noise generator produces all frequencies with random variation over time.
-The LP filter removes frequencies over its threshold frequency, and the HP removes frequencies lower than its threshold~\cite{smith2007introduction}.
-The search parameters are the cutoff thresholds for the LP and HP filters.
-Listing~\ref{lst:program0} shows the Faust code for \BPNoise.
-
-\subsubsection{\AddSineSaw}
-\label{sec:program1}
-\AddSineSaw{} is an additive program that combines a saw and a sine wave function.
-Like subtractive synthesis, additive synthesis is a common approach to sound design that has not been extensively tested as a benchmark in sound-matching.
-The search parameters here are the frequency of the sine and saw oscillators.
-Listing~\ref{lst:program1} is the Faust code for \AddSineSaw.
-
-\subsubsection{\AmpMod}
-\label{sec:program2}
-\AmpMod{} involves amplitude modulation of a noise generator by an LFO with \textit{modulator} as its frequency and a global \textit{amp} value that does not change over time and applies to the entire signal.
-The search parameters for this program are the LFO frequency and the amp value.
-The amp value acts as a global volume control, and since the sounds are normalized before analysis, it is likely inconsequential to the performance.
-This program is meant to be a stepping stone to \FMMod, which utilizes proper AM synthesis.
-Listing~\ref{lst:program2} is the Faust code for \AmpMod.
-
-\subsubsection{\FMMod}
-\label{sec:program3}
-Relative to additive and subtractive synthesis, AM/FM synthesis are less common methods of sound design.
-Due to their frequent use in previous works, we also tested an AM synthesizer.
-The synthesizer program is the multiplication of a low frequency sine oscillator with frequency parameter \textit{amp}, and a saw oscillator with frequency parameter \textit{carrier}.
-Listing~\ref{lst:program3} is the Faust code for \FMMod.
-
-\subsection{Post hoc Analysis of Loss Performance}
-\label{sec:results_posthoc}
-As summarized in Figure~\ref{fig:posthoc_evaluation}, for each program and evaluation (or scoring) method, there are four sets 1000 bootstrapped values, corresponding to the four loss functions.
-A post-hoc test is conducted in two stages in order to determine whether there is a difference in performance between loss functions for each program, and if so, which are the best performers.
-The first stage determines whether there is a difference in group means, with the null hypothesis being that all groups have similar mean ranks.
-The second stage ranks the loss functions from best to worst using the non-parametric Scott-Knott test (\gls{NPSK})~\cite{tantithamthavorn2017mvt,tantithamthavorn2018optimization}.
-The Kruskal-Wallis test is used for the first stage~\cite{kruskal1952use}; this test pools and ranks all evaluation measures for a program, then tests whether the differences in mean rank for all loss function groups is zero.
-The Kruskal-Wallis test calculates an H statistic that is compared to a chi-square distribution with k-1 degrees of freedom (k is the number of groups, or 4, and degrees of freedom is 3).
-Using the significance level of 0.05, if the H statistic is greater than the critical value of the chi-square distribution then the null hypothesis is rejected, meaning that at least one group has a mean rank significantly above others~\cite{kruskal1952use}.
-Table~\ref{tab:kruskal_auto} shows the Kruskal-Wallis results for every program and evaluation method, and shows that significant differences between loss function performance measures (i.e., the bootstrapped distributions of the evaluation for each loss function) exist in all cases except MSS for the \AmpMod{} program.
-
-\begin{table}[ht]
-\centering
-\caption{Do loss functions have different median performance? Kruskal-Wallis results by program and bootstrapped evaluation results.}
-\begin{tabular}{|c|c|c|c|c|}
-\hline
-\textbf{Program} & \textbf{Eval. Method} & \textbf{H-Stat.} & \textbf{P-value} & \textbf{Reject} \\
-\hline
-\BPNoise & MSS      & 81.19  & $1.70 \times 10^{-17}$ & Yes \\
-\BPNoise & P-Loss  & 163.60 & $3.07 \times 10^{-35}$ & Yes \\
-\BPNoise & Manual & 9.45  & $2.39 \times 10^{-2}$ & Yes \\
-\AddSineSaw & MSS      & 308.97 & $1.14 \times 10^{-66}$ & Yes \\
-\AddSineSaw & P-Loss  & 348.42 & $3.28 \times 10^{-75}$ & Yes \\
-\AddSineSaw & Manual & 9.45  & $2.39 \times 10^{-2}$ & Yes \\
-\AmpMod & MSS      & 1.35   & $0.7171$               & No \\
-\AmpMod & P-Loss  & 366.76 & $3.50 \times 10^{-79}$ & Yes \\
-\AmpMod & Manual & 32.71 & $3.70 \times 10^{-7}$ & Yes \\
-\FMMod & MSS      & 564.65 & $4.65 \times 10^{-122}$ & Yes \\
-\FMMod & P-Loss  & 229.19 & $2.07 \times 10^{-49}$ & Yes \\
-\FMMod & Manual &  207.58 & $9.69 \times 10^{-45}$ & Yes \\
-\hline
-\end{tabular}
-\label{tab:kruskal_auto}
-\end{table}
-
-% \begin{table}[ht]
-% \centering
-% \begin{tabular}{|c|c|c|c|}
-% \hline
-% \textbf{Program} & \textbf{H-Stat.} & \textbf{P-value} & \textbf{Reject} \\
-% \hline
-% \BPNoise & 9.45  & $2.39 \times 10^{-2}$ & Yes \\
-% \AddSineSaw & 9.45  & $2.39 \times 10^{-2}$ & Yes \\
-% \AmpMod & 32.71 & $3.70 \times 10^{-7}$ & Yes \\
-% \FMMod & 207.58 & $9.69 \times 10^{-45}$ & Yes \\
-% \hline
-% \end{tabular}
-% \caption{Manual ranking Kruskal-Wallis results. Significant differences are observed for all programs.}
-% \label{tab:kruskal_manual}
-% \end{table}
-
-\subsubsection{Best Performers For Each Program}
-Based on the bootstrapped distribution of scores, the NPSK algorithm ranks the loss functions from 1 (best) to a maximum of 4 (worst).
-In cases where the distributions are similar, multiple loss functions can be clustered into the same rank.
-In Figure~\ref{fig:npsk_all}, we use color-coded violin plots to visualize the best performers for each program.
-The corresponding colors for each rank are
-\colorbox{rank1}{\textcolor{black}{\textbf{1}}} \colorbox{rank2}{\textcolor{white}{\textbf{2}}} \colorbox{rank3}{\textcolor{white}{\textbf{3}}} \colorbox{rank4}{\textcolor{black}{\textcolor{white}{\textbf{4}}}}
-
-\input{extra_figures/ranking_tables}
-\input{extra_figures/npsk_figures}
-
-\subsubsection{\BPNoise}
-For this synthesizer program, the spectrogram-based models performed the best.
-This makes intuitive sense, as the visual effects of a band-pass filter on white noise are readily apparent in a spectrogram.
-As shown in Figure~\ref{fig:npsk_p0}, manual hearing test and MSS selected \SIMSESpec~as the best performer and \LoneSpec~as the second-best performer, while P-Loss gave the reverse order.
-
-\subsubsection{\AddSineSaw}
-As shown in Figure~\ref{fig:npsk_p1}, JTFS was the best performing loss function in all evaluation methods. Moreover, all evaluation methods produced identical results.
-
-\subsubsection{\AmpMod}
-IP-Loss and manual hearing tests yield identical rankings, and MSS results vary for for ranks 2 to 4.
-It is not surprising that MSS results differ from the hearing tests, as we saw in Table~\ref{tab:kruskal_auto}, MSS showed no significant differences between groups.
-As shown in Figure~\ref{fig:npsk_p2}, all models selected \DTWEnv~as the best performer.
-This may be due to \DTWEnv's focus on periodic changes in loudness.
-
-\subsubsection{\FMMod}
-\DTWEnv~is again the best performer here. As shown in Figure~\ref{fig:npsk_p3}, all methods of analysis gave identical rankings to all programs.
-
-\subsection{Consistency In Rankings}
-\label{sec:consistency_in_rankings}
-\subsubsection{Manual Ranks} Agreement between the manually assigned scores is measured using Spearman’s rank correlation, which provides both a correlation coefficient ($\rho$) ranging from –1 (perfect negative correlation) to 1 (perfect positive correlation), and a p-value testing the null hypothesis of no correlation~\cite{spearman1987proof,rebekic2015pearson}.
-Across all programs, the correlation was very strong ($\rho = 0.86$, $p < 10^{-180}$).
-Per-program correlations were also very strong: $\rho = 0.71$ for \BPNoise{} ($p < 10^{-25}$), $\rho = 0.64$ for \AddSineSaw{} ($p < 10^{-19}$), $\rho = 0.84$ for \AmpMod{} ($p < 10^{-43}$), and $\rho = 0.85$ for \FMMod{} ($p < 10^{-44}$).
-
-Table~\ref{tab:combined_ranks} shows the SNPK rankings of bootstrapped evaluation results for each program.
-Both MSS and P-Loss gave a different rank from the hearing results in 3 out of 16 cases, which shows consistency between automatic and manual hearing tests, at least for the simple programs used in this work.
-We also observe that top ranks were consistent across performance evaluation methodologies, with the exception of P-Loss in \BPNoise, which narrowly picks a different spectrogram-based loss function.
-% \input{extra_figures/ranking_tables}
-
-\subsection{Takeaways}
-An important takeaway is that the loss function that yields the best sound-matching outcomes varies depending on the synthesizer program.
-This is perhaps due to the interaction between how the parameters of the synthesizer influence the sound, and the core sonic features of the loss function.
-
-We also see that the \DTWEnv{} and \SIMSESpec{} (to our knowledge, not previously used in iterative sound-matching), lead to the best results in three of the four programs.
-This shows room for further creative approaches to differentiable sound-similarity measures, particularly since we observed that SOTA loss functions can be ineffective depending on the design of a synthesizer.
-
-Finally, we surprisingly see somewhat consistent results between the rankings assigned by manual hearing tests (which we take as the ground truth), and automatic measures of P-Loss and MSS.
-Due to the simplicity of our programs and the occasional deviations, we cannot conclude that these automatic measures are substitutes for human hearing tests.
-
 \section{Summary, Weaknesses, and Conclusion}
@@
 Future works could explore loss functions which only measure certain characteristics of sound (such as \DTWEnv), and whether they can pave the way for better imitation in sound-matching.
-
-\bibliographystyle{IEEEtran}
-\bibliography{references}
-\end{document}

### 2025-09-11 sections/background_related.tex
Updates from Overleaf

@@
 In a recent work, Cherep \textit{et al.} used latent representations from the CLAP model~\cite{wu2023large} along with a differentiable synthesizer~\cite{synthhaxcherep2023} to create creative interpretations of sound effects.
-In their approach, a desired sound-effect is described in text and embedded using clap, and gradient free optimizer~\cite{evosax2022github} iteratively updates the synthesizer's parameters to minimize difference in the text and output sound embeddings.
-Based on manual hearing tests, this approach did not produce the ``correct'' sounds more frequently than previous works~\cite{kreuk2022audiogen}, however, it did yield better scores for ``artistic-interpretation'' (or what we refer to as imitation).
+In their approach, a desired sound-effect is described in text and embedded using CLAP before a gradient free optimizer~\cite{evosax2022github} iteratively updates the synthesizer's parameters to minimize the difference between the text and output sound embeddings.
+Based on manual hearing tests, this approach did not produce the ``correct'' sounds more frequently than previous works~\cite{kreuk2022audiogen}.
+However, it did yield better scores for ``artistic-interpretation'' (or what this works refers to as ``imitation'').
 
@@
 \end{enumerate}
+
+\subsection{What needs to be done?Moving the field forward}

### 2025-09-12 sections/background_related.tex
Updates from Overleaf

--- 2025-09-09 sections/background_related.tex
@@
 For ITF, the synthesizer approximation is frozen and the encoder is iteratively finetuned for the sample sample using gradient descent in order to minimize P-Loss, stopping at convergence or when maximum number of iterations has reached~\cite{barkan2023inversynthII}.
-Barkan \textit{et al.} report that perhaps due to the discrepancy between the true and approximated synthesizer or overfitting, the application of ITF to IS2 generally worsened the spectral and manual performance metrics.
+Barkan \textit{et al.} report that the application of ITF generally worsened the spectral and manual performance metrics, perhaps due to the discrepancy between the true and approximated synthesizer or overfitting.
 
@@
 
-% Section IV titled "Programs, Methodology, and Results" is overloaded and blends experimental setup with results and interpretation.
-For clear clarity and alignment with structural norms, we recommend splitting this into three sections: (1) Experimental Setup (including synthesizer programs, loss definitions, optimization procedure); (2) Experimental Results (performance metrics, statistical tests, and rankings); and (3) Discussion (qualitative interpretation, synthesis-specific observations, and practical implications).
+\subsection{What is Lacking In the Field}
+\label{sec:lacking}
+%-----> much of the original text (available in older version) is removed for brevity)
+
+Having looked at the current available literature, we note four major areas of weakness in sound-matching.
+We target the first two issues in this work, while the latter two issues are left for future work.
+\begin{enumerate}
+    \item \LossSelect: Is there a \textit{best performing} loss function, or is the selection of the loss function dependent on factors such as the sound domain, synthesizer capability and the desired characteristics of the output?
+    \item \SynthSelect: There is a lack of diversity in the synthesis methods used in sound-matching.
+The use of different DSP functions for iterative sound-matching and their effects on the outcome or the interactions with the loss function have not been tested.
+    \item \PeriodicLoss: The sinusoidal nature of loss function landscapes is a problem that appears frequently.
+In differentiable settings, this problem causes gradient descent updates to local minima.
+    \item \OutDomain: Sound-matching with out-of-domain sounds is an unexplored area, yet a necessary one for practical applications for sound designers.
+\end{enumerate}
+
+ % We have highlighted the lack of attention given to the selection of loss functions in sound-matching, and noted that the few existing works addressing this issue evaluate their proposed losses using only simplistic sound synthesis methods.
+We hypothesize that the best performing loss function is influenced by the synthesis techniques utilized in the synthesizer.
+This hypothesis targets the issues of \LossSelect~and \SynthSelect.
+
+  % Towards addressing the problems of \LossSelect{} and \SynthSelect{}, we compare the performance of iterative sound-matching experiments, where four differentiable loss functions are applied to four differentiable synthesizers.
+What we mean by ``performance'' is the similarity of the final output to the target, as measured by P-Loss, MSS, and---most importantly---manual listening tests.
+\textbf{We hypothesize that the relative performance of loss functions for sound-matching is influenced by the method of synthesis}.
+If true, this would strengthen the idea that the selection of the loss function and synthesis method is not an optimization problem, but a subjective choice depending on the needs of the sound designer.

### 2025-09-13 sections/background_related.tex
Updates from Overleaf

@@
     \item \LossSelect: Is there a \textit{best performing} loss function, or is the selection of the loss function dependent on factors such as the sound domain, synthesizer capability and the desired characteristics of the output?
+The few existing works addressing this issue evaluate their proposed losses using only simplistic sound synthesis methods\todo{cite?}.
     \item \SynthSelect: There is a lack of diversity in the synthesis methods used in sound-matching.
 The use of different DSP functions for iterative sound-matching and their effects on the outcome or the interactions with the loss function have not been tested.
-    \item \PeriodicLoss: The sinusoidal nature of loss function landscapes is a problem that appears frequently.
+    \item \PeriodicLoss: The sinusoidal nature of loss function landscapes is a problem that appears frequently\todo{cite?}.
 In differentiable settings, this problem causes gradient descent updates to local minima.
@@
 \end{enumerate}
+
+\subsection{Why This Methodology?}
+\label{sec:why_method}
+Our study takes a deliberately controlled approach: simple differentiable DSP synthesizers, optimized directly with gradient descent, and evaluated under multiple loss functions.
+
+We adopt this design for three reasons.
+
+First, isolating small-scale, low-parameter synthesizers avoids the confounding effects of large neural proxies or embedding models, making it possible to directly analyze the interaction between synthesis structure and loss function.
+This fills a gap in prior work where there is little insight into the comparison of loss functions under different synthesis conditions.
+
+2.
+by comparing multiple synthesis methods side by side (subtractive, additive, amplitude modulation, etc.), we explicitly address the lack of diversity in \SynthSelect, and show that loss effectiveness is not universal but context-dependent.
+This comparative methodology highlights scenarios where specific losses (e.g., DTW\_Envelope for modulated signals, SIMSE\_Spec for subtractive filtering) are most appropriate.
+
+3. Our methodology emphasizes reproducibility and scalability. 
+By using differentiable Faust-to-JAX synthesizers and modest computational resources, we create a lightweight experimental platform that can be extended to more complex synthesis techniques in future work, while still being transparent and interpretable in its current form.
+
+4.
+Past synths are either simple differentiable functions (commonly FM), or complex VSTs where we can't isolate the interactions between loss and functions, here we use isolated building blocks of DSP functions in a differentiable environment.
+
+In this way, our approach complements prior neural and proxy-based methods by providing the missing “low-level experiments” that clarify fundamental interactions between DSP functions and similarity measures.
+We argue that such isolated benchmarking is a necessary precursor to building robust, generalizable sound-matching systems.
 

### 2025-09-15 sections/background_related.tex
Updates from Overleaf

--- 2025-09-12 sections/background_related.tex
@@
 \end{enumerate}
+
+We argue that isolated benchmarking is a necessary precursor to building robust, generalizable sound-matching systems.
+This fills a gap in prior work where there is little insight into the comparison of loss functions under different synthesis conditions, and how the two interact.
+To address the problems of \LossSelect{} and \SynthSelect{}, we pursue a methodology with 3 key characteristics that do not appear together in other works.
+
+1.
+Simple differentiable synthesizers using common DSP functions, filters, and oscillators would be beneficial in addressing the \SynthSelect{} problem.
+Isolating small-scale, low-parameter synthesizers avoids the confounding effects of large neural proxies or embedding models, making it possible to directly analyze the interaction between synthesis structure and loss function.
+
+2.
+Evaluation of experiments under multiple loss functions: Addressing the \LossSelect{} problem requires that multiple loss functions are tested.
+A comparison of performance of different loss functions is rare in previous works~\todo{cite}.
+
+3.
+Parameters optimized directly with gradient descent without additional neural network approximations.
+This would complement prior neural and proxy-based methods by providing the missing “low-level experiments” that clarify fundamental interactions between DSP functions and similarity measures.
+
+% 2.
+by comparing multiple synthesis methods side by side (subtractive, additive, amplitude modulation, etc.), we explicitly address the lack of diversity in \SynthSelect, and show that loss effectiveness is not universal but context-dependent.
+% This comparative methodology highlights scenarios where specific losses (e.g., DTW\_Envelope for modulated signals, SIMSE\_Spec for subtractive filtering) are most appropriate.
+
+% 3. Our methodology emphasizes reproducibility and scalability. 
+% By using differentiable Faust-to-JAX synthesizers and modest computational resources, we create a lightweight experimental platform that can be extended to more complex synthesis techniques in future work, while still being transparent and interpretable in its current form.
+
+% 4.
+Past synths are either simple differentiable functions (commonly FM), or complex VSTs where we can't isolate the interactions between loss and functions, here we use isolated building blocks of DSP functions in a differentiable environment.
 

### 2025-09-16 sections/background_related.tex
Updates from Overleaf

--- 2025-09-12 sections/background_related.tex
@@
-\section{Background and Related Works}
+\section{Background and Related Work}
+\label{sec:background_related}
 
-\todo{Sound-matching is an amalgamation of various fields and  requires knowledge of digital signal processing, audio representation, and heuristic search.
-Here we introduce the various fields and ideas behind sound-matching, before we use these ideas in a historical framing of past sound-matching works.}
+Sound-matching sits at the intersection of digital signal processing, audio representation, and optimization. 
+In this section we first formalize the sound-matching task, then review core synthesis methods, similarity measures, and optimization approaches. 
+We conclude with a survey of related work and a discussion of gaps in the field.
 
-\subsection{Digital Audio Synthesis}
+\subsection{Formalization of Sound-Matching}
+\label{sec:sound_matching_definition}
+Following prior works~\cite{vahidi2023mesostructures,han2023perceptual}, we define sound-matching in terms of a parametric synthesizer, a target sound, a representation function, and a similarity measure.
+The key components are:
+\begin{itemize}
+    \item $g(\theta)$: A parametric audio synthesizer $g$ with parameters $\theta$.
+    \item $x$: The synthesizer output, $x = g(\theta)$.
+    \item $t$: The target sound to be replicated or imitated. 
+    \item $\phi(\cdot)$: A representation (feature extraction) function mapping signals into a comparison space.
+    \item $L$: A loss function that measures distance between $x$ and $t$, typically via $L(\theta,t) = d(\phi(x), \phi(t))$ for some metric $d$.
+\end{itemize}
+
+This formalization highlights that sound-matching is not a single well-defined optimization problem, but rather depends on design choices for $g$, $\phi$, and $L$. 
+Moreover, the ``correct'' solution may not exist in a strict sense: depending on the artistic goal, multiple parameter settings may yield acceptable or even preferable outputs.
+In the following sections we will discuss the various components of sound-matching, the subjectivity of ``correct'' solutions, and important issues in the field.
+
+\subsection{Digital Signal Processing and Synthesis}
 \label{sec:dsp}
-A digital audio synthesizer is any software used for the creation and manipulation of digital audio.
-Digital synthesizers use signal processing chains with a variety of digital signal processing (\gls{DSP}) functions to create sounds.
-These functions are often parametric, and the set of parameters given to the synthesizer's chain of DSP functions is called a synthesizer program.
 
-The simplest form of DSP function is a sinusoidal tone. For example, consider:
-\[ x(n) = sin( 2\pi n T)\]
+A digital audio synthesizer generates or processes audio by chaining digital signal processing (\gls{DSP}) functions. 
+Each function is parameterized, and the set of parameters defining a chain of DSP functions is called a synthesizer program.
 
-where $T$ represents time in seconds, and $n$ is the frequency parameter.
-If $n$ is 1, then $x(n)$ would be a waveform with a frequency of 1 hertz, meaning that the value of $x(n)$ oscillates to the same point once every second.
-Waveform generators are called oscillators for this reason.
+The simplest DSP function is a sinusoidal oscillator: 
+\[
+x[n] = \sin\!\left(2 \pi f \frac{n}{SR}\right),
+\]
+where $f$ is frequency in hertz, $SR$ is the sampling rate, and $n$ is the discrete time index. 
+At a sampling rate of $SR$ samples per second, a 1\,Hz oscillator completes one cycle per second. 
+Such computations form the foundation of digital audio synthesis~\cite{lyons1997understanding}.
 
-Since the advent of digital signal processing in the 1960s~\cite{stranneby2004digital}, a wide variety of parametric audio generation functions have been proposed~\cite{lyons1997understanding,russ1999sound,shier2020spiegelib}.
-This includes oscillators, filters, equalizers, and envelopes, which can be used sequentially or in parallel~\cite{lyons1997understanding,russ1999sound}.
-Sound design with a synthesizer is done by modifying the parameters of the functions until reaching a desired output~\cite{roads1996computer,pinch2004analog}.
+Since the advent of DSP in the 1960s~\cite{stranneby2004digital}, a wide variety of parametric functions have been proposed, including oscillators, filters, equalizers, and envelopes~\cite{lyons1997understanding,russ1999sound}.
+Sound design involves modifying these parameters until a desired output is reached~\cite{roads1996computer,pinch2004analog}.
 
-Particularly in earlier works, frequency/amplitude modulators (\gls{FM}/\gls{AM}) has been the synthesizer of choice in sound-matching~\cite{horner1993machine,mitchell2007evolutionary,vahidi2023mesostructures}.
-This synthesis method is simple to implement, yet very expressive~\cite{chowning1973synthesis}.
-A common implementation is the use of one carrier, one modulator, and the corresponding envelopes:
-\[ f(t) = I_c(t) \cos(\omega_c t + I_m(t) \cos(\omega_m t))
-\]
-Where $f(t)$ is the output signal, $I_c(t)$ is the carrier amplitude, $\omega_c$ is the carrier frequency, and $I_m(t)$ and $\omega_m$ are the envelope and frequency for the modulator~\cite{justice1979analytic}.
+Early sound-matching research focused heavily on frequency and amplitude modulation (\gls{FM}/\gls{AM}) synthesis~\cite{horner1993machine,mitchell2007evolutionary,vahidi2023mesostructures}, which is simple to implement yet expressive~\cite{chowning1973synthesis}.
+Other approaches studied in isolation include additive and subtractive synthesis~\cite{engel2020ddsp,masuda2023improving,salimi2020make} and physical modeling~\cite{riionheimo2003parameter,han2024learning}.
 
-Compared to the number of possible approaches to sound synthesis, a relatively small subset of synthesis methods have been analyzed in \textit{isolation} for sound-matching.
-Notable methods here are additive and/or subtractive synthesizers~\cite{engel2020ddsp,masuda2023improving,salimi2020make} and physical modeling~\cite{riionheimo2003parameter,han2024learning}.
+\highlight{By \textit{isolation}, we mean settings where the effect of individual parameters on the output sound remains tractable. 
+Recent work has also explored commercial Virtual Studio Technology (VST) synthesizers~\cite{yee2018automatic,esling2019flow}. 
+While powerful, such systems obscure the interactions between modules, losses, and outputs. 
+In contrast, small-scale differentiable synthesizers built from isolated DSP blocks offer interpretability, making them particularly useful for benchmarking. }
 
-\highlight{
-What we mean here by \textit{isolation} is that the effect of the synthesizer's parameters (how they change the output sound and experiment outcome) is easily tractable.
-Recent computational advancements have allowed the use of complex Virtual Studio Technology (VST)\cite{steinberg1996vst} synthesizers in sound-matching~\cite{yee2018automatic,esling2019flow}, however, the trade-off here is the obfuscation of the interactions between the synthesizer functions, the loss, and the final outcome, among other distinctions which we will discuss in this section.
-}
-
-An important recent development in digital audio synthesis is the increase of interest in differentiable DSP (\gls{DDSP})~\cite{engel2020ddsp}.
-DDSP is a general term for applications that combine machine learning techniques such as gradient descent~\cite{goodfellow2016deep,boyd2004convex} with DSP.
-Implementing complex DSP functions in a differentiable manner can be challenging, and effective differentiable audio similarity measures require careful mathematical expression of the desired attributes of sound.
-These issues have likely contributed to the limited exploration of this domain~\cite{masuda2021soundmatch,vahidi2023mesostructures,uzrad2024diffmoog}.
+Finally, recent years have seen a growing interest in differentiable DSP (\gls{DDSP})~\cite{engel2020ddsp}, which integrates gradient-based optimization~\cite{goodfellow2016deep,boyd2004convex} with DSP building blocks.
+Implementing complex DSP functions in a differentiable manner remains challenging, and robust differentiable audio similarity measures are still under active investigation~\cite{masuda2021soundmatch,vahidi2023mesostructures,uzrad2024diffmoog}.
 
@@
 
-Here we discuss four methods of audio representation and the corresponding loss functions.
-We also include a technical justification for our novel methods of defining sound similarity using the Scale-Invariant Spectrogram Error (SIMSE) and Dynamic Time Warping (DTW) techniques.
+Here we discuss four methods of audio representation and the corresponding loss functions, including technical justifications for our novel methods.
 
@@
 There are different types of spectrograms that have a basis in Fourier transformations, but the most notable and commonly used is the STFT.
-
 What we call \textit{Fourier-based Spectrograms} are variations on the STFT approach.
@@
 
-\textbf{Why use Scale-Invariant Spectrogram for audio loss}
-A limitation of L1/L2 distances on spectrograms is their sensitivity to global gain: 
-two sounds with identical spectral shape but different amplitude can yield large losses. 
-Scale-Invariant Mean Squared Error (SIMSE) addresses this by normalizing the energy of the compared signals, thereby focusing on proportional differences in spectral envelopes. 
-This can be advantageous for subtractive synthesis (used by one of our synthesizers), where filter cutoffs reshape the spectral balance which can cause changes in total energy. 
-\todo{Perceptually, human listeners are often invariant to global amplitude changes, attending instead to relative spectral shapes}. 
-Although SIMSE has been used in other contexts (e.g., image/audio reconstruction), its application as a differentiable spectrogram loss in iterative sound-matching is, to our knowledge, novel.
+A common limitation of spectrogram-based losses is their sensitivity to global gain: two signals with identical spectral shape but different overall amplitude can yield large errors under L1/L2 norms.
+Here we propose and test Scale-Invariant Mean Squared Error (SIMSE) as an alternative to L1/L2 norms in spectrogram comparisons.
+SIMSE normalizes the amplitudes before comparison and emphasizes proportional differences in spectral envelopes.
+Perceptually, listeners are often tolerant of loudness changes while remaining sensitive to timbral shape, therefore this property could be advantageous for subtractive synthesis, where filter cutoffs reshape spectral balance without predictable changes in total energy.
+Although SIMSE has been applied in other contexts such as audio reconstruction, to our knowledge its use as a differentiable spectrogram loss in iterative sound-matching is novel.
 
@@
 In classical DTW, these conditions are \textit{monotonicity}, \textit{boundary matching}, and \textit{single step-size}.
+DTW measures the distance between the time-series \textit{after} alignments, typically using Euclidean distance, such that the distance between a time-series and shifted versions of itself would be 0, regardless of shift amount~\cite{tavenard.blog.dtw}.
+Additional rules can be imposed to keep alignments locally constrained~\cite{itakura1975minimum,sakoe1978dynamic}.
 
-\textbf{Why use DTW for audio loss?}
-DTW measures the distance between the time-series \textit{after} alignments, typically using Euclidean distance.
-The distance between a time-series and shifted versions of itself would be 0, regardless of shift amount~\cite{tavenard.blog.dtw}.
-Additional rules can be imposed to keep alignments locally constrained~\cite{itakura1975minimum,sakoe1978dynamic}.
 DTW provides robustness to local temporal shifts, making it well suited for comparing modulated signals where the perceptual similarity lies in envelope dynamics rather than precise onset alignment. 
 For example, two tremolo signals with slightly different phase are perceptually similar but would appear distant under spectrogram L1. 
-By applying DTW specifically to amplitude envelopes, we align with perceptual cues of loudness modulation while maintaining differentiability. 
+By applying DTW specifically to amplitude envelopes, we can match two sounds with identical rates ofloudness modulation regardless of alignments. 
 To our knowledge, DTW applied in this way has not been used as a loss in sound-matching (whether differentiable and iterative or not).
-
-\subsection{Automatic Sound Matching}
-\label{sec:sound_matching_definition}
-The problem of sound-matching has been tackled from a variety of perspectives as it involves a number of modular parts: the choice of synthesizer $g$, the target sounds of interest, the representation function $\phi$, and the heuristic for finding the optimal $\theta$.
-To give a formal definition of sound-matching, we expand the definition given by recent works~\cite{vahidi2023mesostructures,han2023perceptual}.
-The major components are as follows:
-\begin{itemize}
-    \item $g(\theta)$: Parametric audio synthesizer $g$ which takes on parameters $\theta$ 
-    \item $x$: The output of $g$, given a set of parameters $\theta$ or $g(\theta) = x$ 
-    \item $t$: The target sound which we want to replicate or imitate. 
-    \item $\phi$: Representation function or feature extractor. $\phi$ is applied to $x$ and $t$ to facilitate their comparison by the loss function.
-    \item $L$: Loss or error function. $L(\theta,t)$ is a measure of distance between $x$ and $t$. Often proportional to the subtraction of their representations, or $ \phi(x) - \phi(t)$
-\end{itemize}
-
-\highlight{It could be argued that the implementation of these components is not an optimization problem with a correct answer, but a \textit{creative} endeavor depending on the artist's needs}.
-Consider the hypothetical case where the goal might be to create an 8-bit~\cite{collins2007loop} version of a high quality, organic snare drum: in such a case, there is no optimal or ``correct" version of an 8-bit snare sound.
 
@@
 \subsection{Historical Framing of Sound-Matching}
-Given the different attributes discussed thus far, Table~\ref{tab:summary} contains an overview of important relevant literature.
-This leads us to a more detailed historical analysis of past works in sound-matching.
+Given the attributes discussed above, Table~\ref{tab:summary} summarizes relevant literature. Here we expand with a historical analysis of past sound-matching work.
 
-Perhaps the earliest foundational work in sound-matching is the analytical approach by Justice~\cite{justice1979analytic} toward the decomposition and recreation of sounds using the simple FM synthesizer described in Section~\ref{sec:dsp}.
-The sound-matching works that followed used other heuristics for finding the synthesizer parameters, yet the simple FM synthesizer structure remained unchanged.
-Horner \textit{et al.}~\cite{horner1993machine} used GAs for re-synthesis of sounds with an FM synthesizer using one modulator and one carrier oscillator, and the McAulay-Quatieri method~\cite{mcaulay1986speech} for measuring loss.
+Perhaps the earliest foundational study is Justice~\cite{justice1979analytic}, who analytically decomposed and recreated sounds using a simple FM synthesizer (see Section~\ref{sec:dsp}). 
+Subsequent work retained this FM structure but explored new heuristics for parameter search. 
+For example, Horner \textit{et al.}~\cite{horner1993machine} applied genetic algorithms (GAs) to resynthesize sounds with one modulator and one carrier, measuring similarity via the McAulay–Quatieri method~\cite{mcaulay1986speech}.
 
-Introduction of more complex models of synthesis such as wavetables~\cite{horner2003auto} and physical modeling~\cite{riionheimo2003parameter} rendered interesting results, but led to further questions about the nature and goals of sound-matching.
-Mitchell and Creasy noted the difficulty in distinguishing between ``inefficiency of the optimization engine'' and ``synthesizer limitations'' as the cause of failure or success in sound-matching~\cite{mitchell2007evolutionary}.
-They offered a \textit{contrived methodology} for finding the best evolutionary method for sound-matching using their FM synthesizer.
-The methodology postulates that the best search heuristic for in-domain search (where an exact target exists, and a wide range of targets can be produced by sampling from different points in the parameter space) would also be the best-performing heuristic for out-of-domain search on a dataset of muted trumpet tones~\cite{opolko1989mcgill}; however, testing was not extensive and yielded some contradictory results.
-As they noted, modifications to the synthesizer, loss function, and sound domain would lead to a problem with an entirely new search space~\cite{mitchell2007evolutionary}.
+Later studies introduced more complex synthesis models, such as wavetables~\cite{horner2003auto} and physical modeling~\cite{riionheimo2003parameter}. 
+Mitchell and Creasy~\cite{mitchell2007evolutionary} highlighted the difficulty of disentangling synthesizer limitations from optimization inefficiency. 
+They proposed a \textit{contrived methodology} in which the best heuristic for in-domain FM resynthesis should also generalize to out-of-domain targets (e.g., muted trumpet tones~\cite{opolko1989mcgill}).
+However, limited testing produced contradictory results, and they concluded that changes to the synthesizer, loss, or sound domain effectively redefine the search space~\cite{mitchell2007evolutionary}.
 
@@
 For training the networks, P-Loss was used, since differentiable MFCCs were not possible in their pipeline.
-Importantly, informal hearing tests revealed that the performance of even the best NN model was unsatisfactory, possibly due to the complex nature of the synthesizer, which features 155 parameters.
+Importantly, informal listening tests found the results unsatisfactory, likely due to the synthesizer’s 155-parameter complexity.
 
-Masuda \textit{et al.} also applied supervised learning to synthesizer parameter estimation~\cite{masuda2021soundmatch}.
+Masuda \textit{et al.} also applied supervised learning~\cite{masuda2021soundmatch}.
 Their work highlights the issue of non-linearity in parameter-to-synthesizer outputs and out-of-domain search.
-This work uses a differentiable subtractive synthesizer (two oscillators and an LP filter).
-A NN model was pre-trained using an in-domain dataset of randomly selected parameters and P-Loss.
+This work uses a differentiable subtractive synthesizer (two oscillators and an LP filter) alongside a NN model pre-trained with P-Loss on an in-domain dataset of randomly selected parameters.
 After training, the model was fine-tuned using 20,000 out-of-domain sounds from the NSynth dataset~\cite{engel2017neural} and multi-scale spectrogram loss~\cite{engel2020ddsp}.
@@
 
-During the training phase, this metric is calculated alongside the neural parameter estimator.
-After training, PNP can be used as a fast approximation of computationally demanding loss functions such as JTFS~\cite{han2023perceptual,han2024learning}, paired with FM synthesizers and a differentiable physical model of a drum head~\cite{smith2010physical}.
+This metric is calculated alongside the neural parameter estimator.
+Once trained, PNP enabled the fast differentiable optimization of computationally expensive loss functions such as JTFS with FM and physical models~\cite{han2023perceptual,han2024learning}.
 
@@
 
-Barkan \textit{et al.} attemped to  improve the IS2 model with Inference-Time
+Barkan \textit{et al.} attempted to  improve the IS2 model with Inference-Time
 Finetuning (ITF).
-For ITF, the synthesizer approximation is frozen and the encoder is iteratively finetuned for the sample sample using gradient descent in order to minimize P-Loss, stopping at convergence or when maximum number of iterations has reached~\cite{barkan2023inversynthII}.
+For ITF, the synthesizer approximation is frozen and the encoder is iteratively finetuned for the sample using gradient descent in order to minimize P-Loss, stopping at convergence or when maximum number of iterations has reached~\cite{barkan2023inversynthII}.
 Barkan \textit{et al.} report that the application of ITF generally worsened the spectral and manual performance metrics, perhaps due to the discrepancy between the true and approximated synthesizer or overfitting.
@@
 
-\subsection{What is Lacking In the Field}
-\label{sec:lacking}
-%-----> much of the original text (available in older version) is removed for brevity)
+% \subsection{What is Lacking In the Field}
+% \label{sec:lacking}
+% %-----> much of the original text (available in older version) is removed for brevity)
 
-Having looked at the current available literature, we note four major areas of weakness in sound-matching.
+% Having looked at the current available literature, we note four major areas of weakness in sound-matching.
 We target the first two issues in this work, while the latter two issues are left for future work.
-\begin{enumerate}
-    \item \LossSelect: Is there a \textit{best performing} loss function, or is the selection of the loss function dependent on factors such as the sound domain, synthesizer capability and the desired characteristics of the output?
-    \item \SynthSelect: There is a lack of diversity in the synthesis methods used in sound-matching.
+% \begin{enumerate}
+%     \item \LossSelect: Is there a \textit{best performing} loss function, or is the selection of the loss function dependent on factors such as the sound domain, synthesizer capability and the desired characteristics of the output?
+The few existing works addressing this issue evaluate their proposed losses using only simplistic sound synthesis methods\todo{cite?}.
+%     \item \SynthSelect: There is a lack of diversity in the synthesis methods used in sound-matching.
 The use of different DSP functions for iterative sound-matching and their effects on the outcome or the interactions with the loss function have not been tested.
-    \item \PeriodicLoss: The sinusoidal nature of loss function landscapes is a problem that appears frequently.
+%     \item \PeriodicLoss: The sinusoidal nature of loss function landscapes is a problem that appears frequently\todo{cite?}.
 In differentiable settings, this problem causes gradient descent updates to local minima.
-    \item \OutDomain: Sound-matching with out-of-domain sounds is an unexplored area, yet a necessary one for practical applications for sound designers.
-\end{enumerate}
+%     \item \OutDomain: Sound-matching with out-of-domain sounds is an unexplored area, yet a necessary one for practical applications for sound designers.
+% \end{enumerate}
 
- % We have highlighted the lack of attention given to the selection of loss functions in sound-matching, and noted that the few existing works addressing this issue evaluate their proposed losses using only simplistic sound synthesis methods.
+% \subsection{What Approach Should We Take?}
+% We argue that isolated benchmarking is a necessary precursor to building robust, generalizable sound-matching systems.
+To address the problems of \LossSelect{} and \SynthSelect{}, we pursue a methodology with 3 key characteristics that do not appear together in other works.
+
+% 1.
+Simple differentiable synthesizers using common DSP functions, filters, and oscillators would be beneficial in addressing the \SynthSelect{} problem.
+Isolating small-scale, low-parameter synthesizers avoids the confounding effects of large neural proxies or embedding models, making it possible to directly analyze the interaction between synthesis structure and loss function.
+
+% 2.
+Evaluation of experiments under multiple loss functions: Addressing the \LossSelect{} problem requires that multiple loss functions are tested.
+A comparison of performance of different loss functions is rare in previous works~\todo{cite}.
+
+% 3.
+Parameters optimized directly with gradient descent without additional neural network approximations.
+This would complement prior neural and proxy-based methods by providing the missing “low-level experiments” that clarify fundamental interactions between DSP functions and similarity measures.
+
+% These characteristics would fill a gap in prior work where there is little insight into the comparison of loss functions under different synthesis conditions, and how the two interact.
+
+% % 4.
+Past synths are either simple differentiable functions (commonly FM), or complex VSTs where we can't isolate the interactions between loss and functions, here we use isolated building blocks of DSP functions in a differentiable environment.
+
+%  % We have highlighted the lack of attention given to the selection of loss functions in sound-matching, and noted that the few existing works addressing this issue evaluate their proposed losses using only simplistic sound synthesis methods.
 We hypothesize that the best performing loss function is influenced by the synthesis techniques utilized in the synthesizer.
 This hypothesis targets the issues of \LossSelect~and \SynthSelect.
-
-  % Towards addressing the problems of \LossSelect{} and \SynthSelect{}, we compare the performance of iterative sound-matching experiments, where four differentiable loss functions are applied to four differentiable synthesizers.
-What we mean by ``performance'' is the similarity of the final output to the target, as measured by P-Loss, MSS, and---most importantly---manual listening tests.
-\textbf{We hypothesize that the relative performance of loss functions for sound-matching is influenced by the method of synthesis}.
-If true, this would strengthen the idea that the selection of the loss function and synthesis method is not an optimization problem, but a subjective choice depending on the needs of the sound designer.

### 2025-09-16 sections/introduction.tex
Updates from Overleaf

@@
 The main requirements of sound-matching are a target sound, a similarity metric (or loss function), and a heuristic to find parameters for a synthesizer that replicate \textit{all} or \textit{some} of the characteristics of the target sound as best as possible~\cite{horner1993machine,mitchell2007evolutionary,masuda2023improving}.
-However, despite the seeming simplicity of the problem and decades of research, sound-matching is not yet practical solution for sound-designers.
+However, despite the seeming simplicity of the problem and decades of research, sound-matching is not yet a practical solution for sound-designers.
 Here we are motivated to understand why this is the case, and what future research should prioritize in order to improve the field.
 
-We provide a review of past works in order to understand what is lacking, a nomenclature for different subtypes of sound-matching, and illuminate major perennial issues.
+We provide a review of past works and identify major perennial issues in the field.
 In particular, we are concerned with the issues of \LossSelect~and \SynthSelect.
@@
 Beyond this central hypothesis, we also investigate three additional research questions: (Q1) To what extent do automatic evaluation metrics agree with manual listening tests?
-(Q2) Can novel loss functions based on Dynamic Time Warping (DTW) and Scale-Invariant Mean Squared Error (SIMSE) provide advantages over standard loss functions, and under what conditions?
+(Q2) Can loss functions based on Dynamic Time Warping (DTW) and Scale-Invariant Mean Squared Error (SIMSE) provide advantages over SOTA loss functions, and under what conditions?
 (Q3) Is an iterative differentiable optimization a viable strategy for design of sound-matching experiments?
 
-% This main hypothesis (along with our forthcoming research questions) would help with important issues regarding the path of future research.
-Should we be searching for a universally best performing loss function across synthesizers?
-Do current, SOTA loss functions satisfy the creative needs of sound-designers?
-
- We take an \textit{iterative} and \textit{differentiable} approach to defining our sound-matching experiments.
+ We adopt an \textit{iterative} and \textit{differentiable} approach to defining our sound-matching experiments.
 The iterative approach better mimics the manual process of recursive listening and parameter adjustments towards sound design, while a differentiable environment allows access to the loss function gradients that can be used to better understand the nature of the problem.
@@
 We evaluate the final similarity with two different automatic methods, as well as manual listening tests conducted by two of the authors.
-Using the Non-parametric Scott-Knott tests~\cite{tantithamthavorn2017mvt,tantithamthavorn2018optimization}, we compare the distributions of the evaluation scores in order to rank the loss functions from best to worst.
+We apply statistical ranking applied to the evaluation scores to compare outcomes across synthesizer–loss combinations and measure the similarity of different evaluation measures.
 
-The results of these experiments show that when accounting for different synthesizers, there is no consistent choice of sound similarity measure for optimal performance.
-Regarding Q1, We see a large overlap between automatic and manual evaluation methods, but continue to encourage manual listening tests as the gold standard for audio evaluation as extreme disagreements can occur.
-In the case of Q2, we confirm that \DTWEnv~outperforms other methods when envelope modulation plays a central role, and \SIMSESpec~is the best performer in simple subtractive synthesis.
-For Q3, we found the declaration of differentiable sound-matching experiments using Jax and transpiled Faust code to be a convenient and resource efficient, and and overall viable general strategy for sound-matching in differentiable settings.
-These results imply that rather than aiming to outperform the SOTA and finding a universal best performer, more effort could be placed towards diversifying the existing similarity measures of sound as well as conducting experiments using a variety of synthesis methods.
-Sound design is an inherently creative endeavor, and would likely benefit from a diversified range of solutions for different approaches.
+\textbf{Contributions.} The contributions of this paper include (1) an evaluation of multiple differentiable losses across multiple synthesis methods, (2) the introduction and justification of loss functions utilizing DTW and SIMSE (3) a discussion on the utility and agreement between manual and automatic evaluation metrics (4) a nomenclature of different sound-matching approaches and unsolved issues in the field.
 

### 2025-09-16 sections/methodology_and_analysis.tex
Updates from Overleaf

--- 2025-09-10 sections/setup_and_posthoc_analysis.tex
@@
 \label{sec:experiment_setup}
+
+% For clear clarity and alignment with structural norms, we recommend splitting this into three sections: (1) Experimental Setup (including synthesizer programs, loss definitions, optimization procedure); (2) Experimental Results (performance metrics, statistical tests, and rankings); and (3) Discussion (qualitative interpretation, synthesis-specific observations, and practical implications).
+
   Towards addressing the problems of \LossSelect{} and \SynthSelect{}, we compare the performance of iterative sound-matching experiments, where four differentiable loss functions are applied to four differentiable synthesizers.
 What we mean by ``performance'' is the similarity of the final output to the target, as measured by P-Loss, MSS, and---most importantly---manual listening tests.
-
 \textbf{We hypothesize that the relative performance of loss functions for sound-matching is influenced by the method of synthesis}.
 If true, this would strengthen the idea that the selection of the loss function and synthesis method is not an optimization problem, but a subjective choice depending on the needs of the sound designer.
-% For clear clarity and alignment with structural norms, we recommend splitting this into three sections: (1) Experimental Setup (including synthesizer programs, loss definitions, optimization procedure); (2) Experimental Results (performance metrics, statistical tests, and rankings); and (3) Discussion (qualitative interpretation, synthesis-specific observations, and practical implications).
-
-% We have highlighted the lack of attention given to the selection of loss functions in sound-matching, and noted that the few existing works addressing this issue evaluate their proposed losses using only simplistic sound synthesis methods.
-\textcolor{highlight}{We hypothesize that the best performing loss function is influenced by the synthesis techniques utilized in the synthesizer}.
-This hypothesis targets the issues of \LossSelect~and \SynthSelect.
 

### 2025-09-16 sections/background_related.tex
Updates from Overleaf

@@
 After training, the model was fine-tuned using 20,000 out-of-domain sounds from the NSynth dataset~\cite{engel2017neural} and multi-scale spectrogram loss~\cite{engel2020ddsp}.
-This approach proved more effective---i.e, achieved lower multi-level spectral difference in out-of-domain tests---than baseline models, which were either not exposed to out-of-domain sounds or trained exclusively with P-Loss.
-Subjective hearing tests were conducted, showing a preference for the fine-tuned model~\cite{masuda2021soundmatch}.
+This approach proved more effective---i.e., achieved lower multi-level spectral difference in out-of-domain tests---than baseline models, which were either not exposed to out-of-domain sounds or trained exclusively with P-Loss.
+Subjective hearing tests were conducted, showing a preference for the fine-tuned model~\cite{masuda2022soundmatch}.
 Masuda \textit{et al.} later extended this work with semi-supervised learning, highlighting significant gaps in in-domain and out-of-domain performance~\cite{masuda2023improving}.
@@
 This library provides different choices for DNNs, GAs, synthesizers, and feature extractors.
-Shier \textit{et al.} presented an experiment with a similar setup to Yee-King \textit{et al.}~\cite{yee2018automatic}, however, they found a genetic algorithm to be the best performing.
+Shier \textit{et al.} presented an experiment with a similar setup to Yee-King \textit{et al.}~\cite{yee2018automatic}, however, they found a genetic algorithm as the best performing.
 
@@
 As they noted, without a differentiable synthesizer, ``model-based'' (or what we call supervised) approaches cannot directly compare $x_o$ to $x_t$, often opting for P-Loss, which may not correctly map the parameters of sound to the output audio~\cite{esling2019flow,han2023perceptual,masuda2023improving}.
-Given a synthesizer and a large dataset of sounds and corresponding parameters, a model which maps sounds to parameters, and another model which approximates the synthesizer can be trained, allowing the implementation of a loss function which combines P-Loss with STFT differences.
+This approach used a model mapping sounds to parameters, another approximating the synthesizer, and a loss function which combines P-Loss with STFT differences.
 This ``Inversynth II'' (IS2) approach yielded significant improvements to previous works which did not use the STFT approximations for loss~\cite{esling2019flow,barkan2019inversynth}.
-
-Barkan \textit{et al.} attempted to  improve the IS2 model with Inference-Time
-Finetuning (ITF).
-For ITF, the synthesizer approximation is frozen and the encoder is iteratively finetuned for the sample using gradient descent in order to minimize P-Loss, stopping at convergence or when maximum number of iterations has reached~\cite{barkan2023inversynthII}.
-Barkan \textit{et al.} report that the application of ITF generally worsened the spectral and manual performance metrics, perhaps due to the discrepancy between the true and approximated synthesizer or overfitting.
+Barkan \textit{et al.} then attempted to improve the IS2 model with Inference-Time
+Fine-tuning (ITF).
+For ITF, the synthesizer approximation is frozen and the encoder is iteratively fine-tuned for a particular sample.
+However, ITF often degraded performance, possibly due to proxy–synthesizer mismatch or overfitting.
 
@@
 
-Some recent works have opted for the use of audio embeddings as a similarity metric.
+ Audio embeddings can also be used as a similarity metric.
 In a recent work, Cherep \textit{et al.} used latent representations from the CLAP model~\cite{wu2023large} along with a differentiable synthesizer~\cite{synthhaxcherep2023} to create creative interpretations of sound effects.
-In their approach, a desired sound-effect is described in text and embedded using CLAP before a gradient free optimizer~\cite{evosax2022github} iteratively updates the synthesizer's parameters to minimize the difference between the text and output sound embeddings.
+In their approach, a desired sound-effect is described in text and embedded using CLAP, followed by iterative updates from a gradient-free optimizer~\cite{evosax2022github} to the synthesizer's parameters to minimize embedding differences.
 Based on manual hearing tests, this approach did not produce the ``correct'' sounds more frequently than previous works~\cite{kreuk2022audiogen}.
@@
 
-% \subsection{What is Lacking In the Field}
-% \label{sec:lacking}
-% %-----> much of the original text (available in older version) is removed for brevity)
+\subsection{What is Lacking In the Field}
+\label{sec:lacking}
+%-----> much of the original text (available in older version) is removed for brevity)
 
-% Having looked at the current available literature, we note four major areas of weakness in sound-matching.
+Having looked at the current available literature, we note four major areas of weakness in sound-matching.
 We target the first two issues in this work, while the latter two issues are left for future work.
-% \begin{enumerate}
-%     \item \LossSelect: Is there a \textit{best performing} loss function, or is the selection of the loss function dependent on factors such as the sound domain, synthesizer capability and the desired characteristics of the output?
+\begin{enumerate}
+    \item \LossSelect: Is there a \textit{best performing} loss function, or is the selection of the loss function dependent on factors such as the sound domain, synthesizer capability and the desired characteristics of the output?
 The few existing works addressing this issue evaluate their proposed losses using only simplistic sound synthesis methods\todo{cite?}.
-%     \item \SynthSelect: There is a lack of diversity in the synthesis methods used in sound-matching.
+    \item \SynthSelect: There is a lack of diversity in the synthesis methods used in sound-matching.
 The use of different DSP functions for iterative sound-matching and their effects on the outcome or the interactions with the loss function have not been tested.
-%     \item \PeriodicLoss: The sinusoidal nature of loss function landscapes is a problem that appears frequently\todo{cite?}.
+    \item \PeriodicLoss: The sinusoidal nature of loss function landscapes is a problem that appears frequently\todo{cite?}.
 In differentiable settings, this problem causes gradient descent updates to local minima.
-%     \item \OutDomain: Sound-matching with out-of-domain sounds is an unexplored area, yet a necessary one for practical applications for sound designers.
-% \end{enumerate}
+    \item \OutDomain: Sound-matching with out-of-domain sounds is an unexplored area, yet a necessary one for practical applications for sound designers.
+\end{enumerate}
 
-% \subsection{What Approach Should We Take?}
-% We argue that isolated benchmarking is a necessary precursor to building robust, generalizable sound-matching systems.
+\subsection{What Approach Should We Take?}
+We argue that isolated benchmarking is a necessary precursor to building robust, generalizable sound-matching systems.
 To address the problems of \LossSelect{} and \SynthSelect{}, we pursue a methodology with 3 key characteristics that do not appear together in other works.
 
-% 1.
+1.
 Simple differentiable synthesizers using common DSP functions, filters, and oscillators would be beneficial in addressing the \SynthSelect{} problem.
@@
 
-% 2.
+2.
 Evaluation of experiments under multiple loss functions: Addressing the \LossSelect{} problem requires that multiple loss functions are tested.
@@
 
-% 3.
+3.
 Parameters optimized directly with gradient descent without additional neural network approximations.
@@
 
-% These characteristics would fill a gap in prior work where there is little insight into the comparison of loss functions under different synthesis conditions, and how the two interact.
+These characteristics would fill a gap in prior work where there is little insight into the comparison of loss functions under different synthesis conditions, and how the two interact.
 
-% % 4.
+% 4.
 Past synths are either simple differentiable functions (commonly FM), or complex VSTs where we can't isolate the interactions between loss and functions, here we use isolated building blocks of DSP functions in a differentiable environment.

### 2025-09-17 sections/background_related.tex
Updates from Overleaf

@@
 
-Having looked at the current available literature, we note four major areas of weakness in sound-matching.
-We target the first two issues in this work, while the latter two issues are left for future work.
+Having looked at the current literature, we identify four major areas of weakness in sound-matching. We target the first two issues in this work, while the latter two issues are left for future work.
 \begin{enumerate}
-    \item \LossSelect: Is there a \textit{best performing} loss function, or is the selection of the loss function dependent on factors such as the sound domain, synthesizer capability and the desired characteristics of the output?
-The few existing works addressing this issue evaluate their proposed losses using only simplistic sound synthesis methods\todo{cite?}.
+    \item \LossSelect: It remains unclear whether there exists a universally best-performing loss function, 
+    or whether performance depends on factors such as the sound domain, synthesizer architecture, 
+    and desired output characteristics. Existing studies typically evaluate losses only with simple synthesis 
+    setups~\cite{vahidi2023mesostructures}, leaving open questions about their generality. 
     \item \SynthSelect: There is a lack of diversity in the synthesis methods used in sound-matching.
-The use of different DSP functions for iterative sound-matching and their effects on the outcome or the interactions with the loss function have not been tested.
-    \item \PeriodicLoss: The sinusoidal nature of loss function landscapes is a problem that appears frequently\todo{cite?}.
+The effects of using different isolated DSP functions in iterative sound-matching and how these interact with loss functions remains largely untested.
+    \item \PeriodicLoss: The sinusoidal nature of loss function landscapes is a problem that appears frequently\todo{cite and make clearer}.
 In differentiable settings, this problem causes gradient descent updates to local minima.
-    \item \OutDomain: Sound-matching with out-of-domain sounds is an unexplored area, yet a necessary one for practical applications for sound designers.
+    \item \OutDomain: Sound-matching with out-of-domain sounds is a largely unexplored area, yet a necessary one for practical applications for sound designers.
 \end{enumerate}
@@
 We argue that isolated benchmarking is a necessary precursor to building robust, generalizable sound-matching systems.
-To address the problems of \LossSelect{} and \SynthSelect{}, we pursue a methodology with 3 key characteristics that do not appear together in other works.
+To address the problems of \LossSelect{} and \SynthSelect{}, we adopt a methodology with three key characteristics that have not previously appeared together.
 
-1.
-Simple differentiable synthesizers using common DSP functions, filters, and oscillators would be beneficial in addressing the \SynthSelect{} problem.
-Isolating small-scale, low-parameter synthesizers avoids the confounding effects of large neural proxies or embedding models, making it possible to directly analyze the interaction between synthesis structure and loss function.
+\begin{enumerate}
+    \item Use simple differentiable synthesizers built from common DSP functions (oscillators, filters, envelopes) to address the \SynthSelect{} problem.
+Restricting to small-scale, low-parameter models avoids the confounding effects of neural proxies or embedding models, enabling direct analysis of how synthesis method interacts with loss functions.
+    \item Evaluate experiments under multiple loss functions to address the \LossSelect{} problem. Prior work has typically tested losses in isolation, making direct comparisons rare~\todo{cite}.  
+    \item Optimize synthesizer parameters directly with gradient descent, without neural approximations.
+This complements prior proxy-based methods by providing the missing “low-level experiments” that clarify fundamental interactions between DSP functions and similarity measures.
+\end{enumerate}
 
-2.
-Evaluation of experiments under multiple loss functions: Addressing the \LossSelect{} problem requires that multiple loss functions are tested.
-A comparison of performance of different loss functions is rare in previous works~\todo{cite}.
-
-3.
-Parameters optimized directly with gradient descent without additional neural network approximations.
-This would complement prior neural and proxy-based methods by providing the missing “low-level experiments” that clarify fundamental interactions between DSP functions and similarity measures.
-
-These characteristics would fill a gap in prior work where there is little insight into the comparison of loss functions under different synthesis conditions, and how the two interact.
+Together, these design choices supply the controlled experimental evidence missing from the literature, 
+clarifying how losses and synthesis methods interact in differentiable, iterative settings, 
+and offering insights likely to generalize to more complex domains.
 

### 2025-09-17 sections/methodology_and_analysis.tex
Updates from Overleaf

@@
+% Section IV titled "Programs, Methodology, and Results" is overloaded and blends experimental setup with results and interpretation.
+For clear clarity and alignment with structural norms, we recommend splitting this into three sections: (1) Experimental Setup (including synthesizer programs, loss definitions, optimization procedure); (2) Experimental Results (performance metrics, statistical tests, and rankings); and (3) Discussion (qualitative interpretation, synthesis-specific observations, and practical implications).
+
 \section{Experimental Setup}
 \label{sec:experiment_setup}
-
-% For clear clarity and alignment with structural norms, we recommend splitting this into three sections: (1) Experimental Setup (including synthesizer programs, loss definitions, optimization procedure); (2) Experimental Results (performance metrics, statistical tests, and rankings); and (3) Discussion (qualitative interpretation, synthesis-specific observations, and practical implications).
-
-  Towards addressing the problems of \LossSelect{} and \SynthSelect{}, we compare the performance of iterative sound-matching experiments, where four differentiable loss functions are applied to four differentiable synthesizers.
+  We want to compare the performance of iterative sound-matching experiments, where different differentiable loss functions are applied to different differentiable synthesizers.
 What we mean by ``performance'' is the similarity of the final output to the target, as measured by P-Loss, MSS, and---most importantly---manual listening tests.
-\textbf{We hypothesize that the relative performance of loss functions for sound-matching is influenced by the method of synthesis}.
-If true, this would strengthen the idea that the selection of the loss function and synthesis method is not an optimization problem, but a subjective choice depending on the needs of the sound designer.
+The results here can provide answers for our stated hypothesis and secondary questions\todo{reference}.
+\todo{Do I remove?}If our hypothesis is true, this would strengthen the idea that the selection of the loss function and synthesis method is not an optimization problem, but a subjective choice depending on the needs of the sound designer.
 
@@
 
-An \textit{experiment} is one iteration of sound-matching given a loss function, target sound,
+What we call an \textit{experiment} is one iteration of sound-matching given a loss function, target sound,
 and synthesizer program.
@@
 Due to their ubiquity and lower cost of gradient calculation, we use STFT as the basis of two loss functions.
-As discussed in Section~\ref{sec:matching_types}, the L1 or L2 distances are the most common methods of comparing different spectrograms, however, there are situations where this might lead to shortcomings.
-An example given by Vahidi \textit{et al.}~\cite{vahidi2023mesostructures} is two chirplets (tones that increase in pitch) starting and ending at the same frequencies.
-If one of these tones is slightly shifted in time, then the spectrograms will no longer overlap at any point, despite their sonic similarity (similar to the idea that a linear equation $f(x) = ax$ can never overlap with a slightly shifted version of itself, or $f(x) = ax +\epsilon$).
+L1 or L2 distances are the most common methods of comparing different spectrograms~\cite{turian2020sorry,richard2025model}, however, there are situations where this might lead to shortcomings.
+\todo{(this is not addressed by SIMSE so why is it here?) An example given by Vahidi \textit{et al.}~\cite{vahidi2023mesostructures} is two chirplets (tones that increase in pitch) starting and ending at the same frequencies.
+If one of these tones is slightly shifted in time, then the spectrograms will no longer overlap at any point, despite their sonic similarity (similar to the idea that a 2D plot of say, $f(x) = ax$, can never overlap with a slightly vertically shifted version of itself, or $f(x) = ax +\epsilon$).
 In this example, we do not know if this is an issue of the spectrograms, or the L1 measure used for their comparison.
-In fields such as computer vision, Structural Similarity Index (SSIM)~\cite{wang2004imagesssim,wang2009mean} and Scale-Invariant Mean Squared Error (SIMSE)~\cite{barron2014shapessimse} have been used as alternatives that improved performance in various tasks relative to L1.
+In fields such as computer vision, Structural Similarity Index (SSIM)~\cite{wang2004imagesssim,wang2009mean} and Scale-Invariant Mean Squared Error (SIMSE)~\cite{barron2014shapessimse} have been used as alternatives that improved performance in various tasks relative to L1.}
 
@@
 
-\section{Post hoc Analysis of Loss Performance}
+% \section{Post hoc Analysis of Loss Performance}
 \label{sec:results_posthoc}
@@
 
-\section{Takeaways, Practical Implications, and Caveats}
+% \section{Takeaways, Practical Implications, and Caveats}
 % (Q1) To what extent do automatic evaluation metrics agree with manual listening tests?

### 2025-09-18 sections/background_related.tex
Updates from Overleaf

@@
 Here we propose and test Scale-Invariant Mean Squared Error (SIMSE) as an alternative to L1/L2 norms in spectrogram comparisons.
-SIMSE normalizes the amplitudes before comparison and emphasizes proportional differences in spectral envelopes.
+SIMSE normalizes the amplitudes before comparison and emphasizes proportional differences in spectral envelopes~\cite{barron2014shapessimse}.
 Perceptually, listeners are often tolerant of loudness changes while remaining sensitive to timbral shape, therefore this property could be advantageous for subtractive synthesis, where filter cutoffs reshape spectral balance without predictable changes in total energy.
@@
 
+ % \todo{(this is not addressed by SIMSE so why is it here?) An example given by Vahidi \textit{et al.}~\cite{vahidi2023mesostructures} is two chirplets (tones that increase in pitch) starting and ending at the same frequencies.
+If one of these tones is slightly shifted in time, then the spectrograms will no longer overlap at any point, despite their sonic similarity (similar to the idea that a 2D plot of say, $f(x) = ax$, can never overlap with a slightly vertically shifted version of itself, or $f(x) = ax +\epsilon$).
+In this example, we do not know if this is an issue of the spectrograms, or the L1 measure used for their comparison.
+In fields such as computer vision, Structural Similarity Index (SSIM)~\cite{wang2004imagesssim,wang2009mean} and Scale-Invariant Mean Squared Error (SIMSE)~\cite{barron2014shapessimse} have been used as alternatives that improved performance in various tasks relative to L1.}
+
 \subsubsection{Joint-Time Frequency Spectrum}
 Recent works have focused on the limitations of parameter and spectral loss functions in sound-matching~\cite{vahidi2023mesostructures,uzrad2024diffmoog}, seeking to create more effective general solutions for the comparison of audio.
-Noting the aforementioned weaknesses of comparing STFT spectrograms, Vahidi \textit{et al.} proposed differentiable Joint-Time Frequency Scattering (\gls{JTFS})~\cite{anden2015joint} as an alternative to spectrogram loss in sound-matching, and showed improved performance in sound-matching with differentiable chirplet synthesizers~\cite{vahidi2023mesostructures}.
-JTFS is the result of the application of a 2D wavelet transformation to the time-frequency representation of a signal~\cite{anden2015joint}.
+Noting the aforementioned weaknesses of comparing STFT spectrograms\todo{you discussed loudness only and not misalignments}, Vahidi \textit{et al.} proposed differentiable Joint-Time Frequency Scattering (\gls{JTFS})~\cite{anden2015joint} as an alternative to spectrogram loss in sound-matching, and showed improved performance in sound-matching with differentiable chirplet synthesizers~\cite{vahidi2023mesostructures}.
+JTFS is the result of the application of a 2D wavelet transformation to the time-frequency representation of a signal~\cite{anden2015joint} and has been reported as more sensitive to mesostructural features such \todo{list}~\cite{vahidi2023mesostructures}.
 

### 2025-09-18 sections/methodology_and_analysis.tex
Updates from Overleaf

@@
 \label{sec:experiment_setup}
-  We want to compare the performance of iterative sound-matching experiments, where different differentiable loss functions are applied to different differentiable synthesizers.
-What we mean by ``performance'' is the similarity of the final output to the target, as measured by P-Loss, MSS, and---most importantly---manual listening tests.
-The results here can provide answers for our stated hypothesis and secondary questions\todo{reference}.
-\todo{Do I remove?}If our hypothesis is true, this would strengthen the idea that the selection of the loss function and synthesis method is not an optimization problem, but a subjective choice depending on the needs of the sound designer.
+We systematically evaluate iterative sound-matching pipelines that combine
+different differentiable loss functions with differentiable synthesizers.
+By ``performance'' we mean the similarity of the synthesized output to a
+target, assessed via P-Loss, MSS, and---most importantly---manual listening
+tests. This setup allows us to test our central hypothesis: that the relative
+performance of a loss function is contingent on the synthesis method, and
+thus no universal ``best'' similarity measure exists.
 
-We test four different loss functions (described in Section~\ref{sec:loss_implementation}) and compare their success in sound-matching using four different synthesizers (Section~\ref{sec:programs}).
-We verify the success of the sound-matching process using two automatic measures of sound similarity (MSS and P-Loss) and a manual blinded hearing test conducted by two of the authors where Likert scores~\cite{jebb2021review} of 1-5 are assigned to a subset of final outputs and the corresponding targets.
-Details regarding the evaluation methods are described in Section~\ref{sec:evaluation_manual_auto}.
-
-What we call an \textit{experiment} is one iteration of sound-matching given a loss function, target sound,
-and synthesizer program.
-When the final iteration is complete, a score is given (automatically with an algorithm or with manual listening) to an experiment that measures its success in replicating the target sound.
-Consider the following simplified scenario: we have two loss functions, $L_1$ and $L_2$, and a single synthesizer program $P$ with a target parameter set $\theta^*$ where the target sound is $t$, or $P(\theta^*) = t$.
-We run a large number (in this case, 300 for each loss function) of experiments where we randomly set the initial ($\theta_0$) and target parameters ($\theta^*$) of $P$ and run an iterative sound-matching loop for each loss function, with maximum number of iterations set to 200.
-When the loop is terminated, we measure the similarity of the final output sound $x$ (where $P(\theta_{199}) = x$) to $t$, or if using P-Loss we compare $\theta_{199}$ to $\theta^*$.
-Having 300 experiments gives us a distribution of losses to determine whether $L_1$ or $L_2$ led to better performance.
-This simplified example leads us to our methodology which has four loss functions, four synthesizer programs, and three different evaluation methods of scoring sound similarity.
+We evaluate four loss functions (Section~\ref{sec:loss_implementation})
+across four differentiable synthesizers (Section~\ref{sec:programs}). For
+each loss–synthesizer pair, we run a large number of trials and evaluate
+their outcomes automatically and manually
+(Section~\ref{sec:evaluation_manual_auto}). ``Experiments'' in this context
+refer to one complete iterative sound-matching run, from random parameter
+initialization through gradient-based optimization until termination. Scores
+from these experiments are then aggregated and statistically compared to
+determine best-performing losses and synth–loss pairings.
 
@@
 Due to their ubiquity and lower cost of gradient calculation, we use STFT as the basis of two loss functions. 
-L1 or L2 distances are the most common methods of comparing different spectrograms~\cite{turian2020sorry,richard2025model}, however, there are situations where this might lead to shortcomings.
-\todo{(this is not addressed by SIMSE so why is it here?) An example given by Vahidi \textit{et al.}~\cite{vahidi2023mesostructures} is two chirplets (tones that increase in pitch) starting and ending at the same frequencies.
-If one of these tones is slightly shifted in time, then the spectrograms will no longer overlap at any point, despite their sonic similarity (similar to the idea that a 2D plot of say, $f(x) = ax$, can never overlap with a slightly vertically shifted version of itself, or $f(x) = ax +\epsilon$).
-In this example, we do not know if this is an issue of the spectrograms, or the L1 measure used for their comparison.
-In fields such as computer vision, Structural Similarity Index (SSIM)~\cite{wang2004imagesssim,wang2009mean} and Scale-Invariant Mean Squared Error (SIMSE)~\cite{barron2014shapessimse} have been used as alternatives that improved performance in various tasks relative to L1.}
+L1 or L2 distances are the most common methods of comparing spectrograms~\cite{turian2020sorry,richard2025model}. 
+However, these measures can be overly sensitive to global gain or minor misalignments, potentially overstating perceptual differences. 
+To address this, we also test Scale-Invariant Mean Squared Error (SIMSE) as an alternative to L1 for comparing STFT spectrograms.
 
@@
 \subsubsection{Soft-DTW Loss}
-We use the soft-DTW function, which is differentiable and --- depending on its parameters --- not shift-invariant~\cite{cuturi2017soft,janati2020spatio,tavenard.blog.softdtw}.
+We use the soft-DTW function, which is differentiable and---depending on its parameters---not shift-invariant~\cite{cuturi2017soft,janati2020spatio,tavenard.blog.softdtw}.
 The loss function \DTWEnv~is the application of the soft-DTW function to the envelope of the two sounds being compared~\cite{lyons1997understanding}.
+This loss uses the similarity of amplitude modulation patterns, which are perceptually important in many sound-design scenarios.
 The envelope is calculated by creating the STFT spectrogram of a sound (the same process used in Section~\ref{sec:fourier_specs}) and summing the values at each timestep.
@@
 Following Braun's methodology~\cite{braun2024dac}, programs are first defined in Faust, then converted (or \textit{transpiled}) to differentiable Jax functions using the DawDreamer library\footnote{\url{https://github.com/DBraun/DawDreamer}}.
-
-% A general overview of the programs and the predefined parameter ranges is given in Table~\ref{tab:programs}.
-
-In the following subsections, we describe the programs and their parameters in more detail, followed by the methodology in Section~\ref{sec:methodology} and results in Section~\ref{sec:results_posthoc}.
-
 \begin{lstlisting}[caption={\BPNoise}, label={lst:program0}, language=Faust,
@@
  \end{enumerate}
+
+ As an example, consider the following simplified scenario: we have two loss functions, $L_1$ and $L_2$, and a single synthesizer program $P$ with a target parameter set $\theta^*$ where the target sound is $t$, or $P(\theta^*) = t$.
+We run a large number (in this case, 300 for each loss function) of experiments where we randomly set the initial ($\theta_0$) and target parameters ($\theta^*$) of $P$ and run an iterative sound-matching loop for each loss function, with maximum number of iterations set to 200.
+When the loop is terminated, we measure the similarity of the final output sound $x$ (where $P(\theta_{199}) = x$) to $t$, or if using P-Loss we compare $\theta_{199}$ to $\theta^*$.
+Having 300 experiments gives us a distribution of losses to determine whether $L_1$ or $L_2$ led to better performance.
+This simplified example leads us to our methodology which has four loss functions, four synthesizer programs, and three different evaluation methods of scoring sound similarity.
 

### 2025-09-18 main.tex
Updates from Overleaf

@@
 \label{sec:results_posthoc}
-As summarized in Figure~\ref{fig:posthoc_evaluation}, for each program and evaluation (or scoring) method, there are four sets 1000 bootstrapped values, corresponding to the four loss functions.
+This should be earlier \todo{As summarized in Figure~\ref{fig:posthoc_evaluation}, for each program and evaluation (or scoring) method, there are four sets 1000 bootstrapped values, corresponding to the four loss functions.
 A post-hoc test is conducted in two stages in order to determine whether there is a difference in performance between loss functions for each program, and if so, which are the best performers.
@@
 Using the significance level of 0.05, if the H statistic is greater than the critical value of the chi-square distribution then the null hypothesis is rejected, meaning that at least one group has a mean rank significantly above others~\cite{kruskal1952use}.
-Table~\ref{tab:kruskal_auto} shows the Kruskal-Wallis results for every program and evaluation method, and shows that significant differences between loss function performance measures (i.e., the bootstrapped distributions of the evaluation for each loss function) exist in all cases except MSS for the \AmpMod{} program.
+Table~\ref{tab:kruskal_auto} shows the Kruskal-Wallis results for every program and evaluation method, and shows that significant differences between loss function performance measures (i.e., the bootstrapped distributions of the evaluation for each loss function) exist in all cases except MSS for the \AmpMod{} program.}
 

### 2025-09-19 main.tex
Updates from Overleaf

--- 2025-08-26 main.tex
@@
 \label{sec:results_posthoc}
-As summarized in Figure~\ref{fig:posthoc_evaluation}, for each program and evaluation (or scoring) method, there are four sets 1000 bootstrapped values, corresponding to the four loss functions.
+As summarized in Figure~\ref{fig:posthoc_evaluation}, for each program and evaluation (or scoring) method, there are four sets of 1000 bootstrapped values, corresponding to the four loss functions.
 A post-hoc test is conducted in two stages in order to determine whether there is a difference in performance between loss functions for each program, and if so, which are the best performers.

### 2025-09-19 sections/methodology_and_analysis.tex
Updates from Overleaf

@@
 
-\subsection{Training Parameters and Gradient Calculations}
- % Previous work utilized gradient fields for analysis of the algorithm performance~\cite{vahidi2023mesostructures}.
-Gradient fields are not the center of our analysis, but we include them in Appendix~\ref{appendix:gradient_fields}.
- For updating the synthesizer parameters (or weights), we use RMSProp, which operates similar to stochastic gradient descent (SGD)~\cite{goodfellow2016deep}, with the caveat that the gradients of each weight are scaled by the root-mean-square of past gradients of that weight.
-Based on some initial test runs, we used an arbitrary fixed learning rate of 0.045 for all experiments (learning rates used in Vahidi \textit{et al.}~\cite{vahidi2023mesostructures} is unknown).
-The maximum number of iterations is set to 200, where, based on our observations, the parameters have either irrecoverably diverged outside the acceptable ranges, or are stuck at a local minimum.
-Additionally, the gradients are large, and calculated backwards sequentially through the signal processing chain.
-This backward calculation resembles the \textit{exploding gradients} problem in recurrent neural networks~\cite{gers2000learning}.
-Gradient clipping~\cite{goodfellow2016deep} is preemptively used to ensure that the  $\ell_2$ norm of all gradients does not exceed the threshold of 1.
-% \mathbf{\gamma}_i^{\text{clipped}} = \mathbf{\gamma}_i \cdot \mathcal{S}
-% \]
-% Where $\mathcal{S}$ is a value less than or equal to 1, depending on whether the norm of the gradients exceeds the threshold $c$.  
-% \[
-% \mathcal{S} = \frac{c}{\max(\|\mathbf{\gamma}\|_2, c)}
-% \]
-
-\subsection{Determining Best Performers}
- Once the programs are transpiled to differentiable synthesizers, the iterative sound-matching procedure is:
+\subsection{Training Loop and Gradient Calculations}
+Given a differentiable loss and synthesizer, the iterative sound-matching procedure is as follows:
  \begin{enumerate}
@@
 
- As an example, consider the following simplified scenario: we have two loss functions, $L_1$ and $L_2$, and a single synthesizer program $P$ with a target parameter set $\theta^*$ where the target sound is $t$, or $P(\theta^*) = t$.
+ For updating the synthesizer parameters (or weights), we use RMSProp, which operates similar to stochastic gradient descent (SGD)~\cite{goodfellow2016deep}, with the caveat that the gradients of each weight are scaled by the root-mean-square of past gradients of that weight.
+Based on some initial test runs, we used an arbitrary fixed learning rate of 0.045 for all experiments (learning rates used in Vahidi \textit{et al.}~\cite{vahidi2023mesostructures} is unknown).
+The maximum number of iterations is set to 200, where, based on our observations, the parameters have either irrecoverably diverged outside the acceptable ranges, or are stuck at a local minimum.
+Additionally, the gradients are large, and calculated backwards sequentially through the signal processing chain.
+This backward calculation resembles the \textit{exploding gradients} problem in recurrent neural networks~\cite{gers2000learning}.
+Gradient clipping~\cite{goodfellow2016deep} is preemptively used to ensure that the  $\ell_2$ norm of all gradients does not exceed the threshold of 1.
+%
+\subsection{Ranking Loss Functions}
+
+ % As an example, consider the following simplified scenario: we have two loss functions, $L_1$ and $L_2$, and a single synthesizer program $P$ with a target parameter set $\theta^*$ where the target sound is $t$, or $P(\theta^*) = t$.
 We run a large number (in this case, 300 for each loss function) of experiments where we randomly set the initial ($\theta_0$) and target parameters ($\theta^*$) of $P$ and run an iterative sound-matching loop for each loss function, with maximum number of iterations set to 200.
@@
 Each synthesizer program is paired with four loss functions and 300 experiments are conducted and evaluated automatically with P-Loss and MSS, and manually with Likert scores.
+With this approach, for each program and loss function pair, there are three distributions that can be used to rank the loss functions.
+Two distributions with 300 automatically assigned similarities (P-Loss and MSS) and one distribution with 80 Likert scores (combining the 40 ranks assigned by each author).
 
-For each program and loss function pair, we have three distributions that can be used to rank the loss functions.
-Two with 300 automatically assigned similarities (P-Loss and MSS) and one with 80 Likert scores (combining the 40 ranks assigned by each author).
 For consistency and statistical robustness, the distributions are upsampled to 1000 values using bootstrapping~\cite{tibshirani1993introduction}.
@@
 
-% For each program, the loss functions are ranked using the non-parametric Scott-Knott (\gls{NPSK}) test~\cite{tantithamthavorn2017mvt,tantithamthavorn2018optimization}, which ranks the loss functions from rank 1 (best, or highest distribution values) to maximum rank of 4 (worst, with lowest distribution values).
-NPSK may determine that multiple functions belong to the same rank, in which case the maximum rank would be less than 4.
-As mentioned before, lower P-Loss and MSS values indicate higher similarity, therefore (for consistency with manually assigned ranks) the reciprocal of their output is used ($x^{-1}$, if $x$ is the P-Loss or MSS value), where higher values indicate greater similarity.
-
 \input{extra_figures/evaluation_graph.tex}
-
-% \section{Post hoc Analysis of Loss Performance}
-\label{sec:results_posthoc}
-As discussed in the previous section and summarized in Figure~\ref{fig:posthoc_evaluation}, for each program and evaluation (or scoring) method, there are four sets of 1000 bootstrapped values corresponding to the four loss functions.
  A post-hoc test is conducted in two stages in order to determine whether there is a difference in performance between loss functions for each program, and if so, which are the best performers.
@@
 Using the significance level of 0.05, if the H statistic is greater than the critical value of the chi-square distribution then the null hypothesis is rejected, meaning that at least one group has a mean rank significantly above others~\cite{kruskal1952use}.
-Table~\ref{tab:kruskal_auto} shows the Kruskal-Wallis results for every program and evaluation method, and shows that significant differences between loss function performance measures (i.e., the bootstrapped distributions of the evaluation for each loss function) exist in all cases except MSS for the \AmpMod{} program.
+\todo{More detail for NPSK and leadup to results}
 
@@
 \subsection{Best Performers For Each Program}
+
+Table~\ref{tab:kruskal_auto} shows the Kruskal-Wallis results for every program and evaluation method, we see a significant differences between loss function performance measures (i.e., the bootstrapped distributions of the evaluation for each loss function) in 11 of the 12 cases, with the exception of  MSS for the \AmpMod{} program.
+
 Based on the bootstrapped distribution of scores, the NPSK algorithm ranks the loss functions from 1 (best) to a maximum of 4 (worst).
@@
 
-\subsection{Takeaways}
-Perhaps the most important takeaway regarding our main hypothesis is that the loss function that yields the best sound-matching outcomes varies depending on the synthesizer program.
-This is likely due to the interaction between how the parameters of the synthesizer influence the sound, and the core sonic features used by the loss function.
-
-\textbf{Q1}: To what extent do automatic evaluation metrics agree with manual listening tests?
-we see somewhat consistent results between the rankings assigned by manual hearing tests (which we take as the ground truth), and automatic measures of P-Loss and MSS.
-With the exception of P-Loss in \BPNoise{} narrowly selecting a different spectrogram based loss, all measures of performance selected the same top performer.
-Due to the simplicity of our programs and the occasional deviations, we cannot conclude that these automatic measures are substitutes for human hearing tests.
-As many previous works have done\todo{cite}, the use of MSS and/or P-Loss as general loss functions does seem appropriate, however, we advise the use of listening tests or custom made loss functions for more specific analysis.
-
-\textbf{Q2}: Are DTW and SIMSE effective measures of loss? If so,
-
- \DTWEnv{} consistently outperformed other measures in amplitude-modulated synthesis, where envelope alignment is critical.
-\SIMSESpec{} performed best in subtractive synthesis, where scale-invariance captures noise filtering effects more robustly than L1-based measures.
-These results indicate that the advantages of novel losses are context-dependent, and the necessity of further creative approaches to differentiable sound-similarity measures.
-
-\textbf{Q3}: Can iterative differentiable optimization be applied effectively in synthesizers using classical DSP functions?
-The approach of designing DSP functions in Faust and transpiling to differentiable Faust code yielded meaningful outputs and enabled comparative evaluation of losses.
-The use of Faust for synthesizer definition was convenient, and transpiling to Jax enabled quick gradient optimization.
-We were able to run our 1200 experiments within 72 hours on a laptop without a GPU\footnote{Lenovo ThinkPad T480, i5-8350U CPU at 1.70GHz, 32 GB RAM.}.
-
-\subsection{Practical Recommendations:}
-Based on our findings, several actionable guidelines emerge.
-These heuristics can guide practitioners in selecting appropriate similarity measures and synthesis methods for future sound-matching experiments.
-\begin{itemize}
-    \item In leu of listening tests, MSS or P-Loss are generally useful for large-scale benchmarking, but confirm conclusions with manual listening when accuracy in isolated cases is important. 
-    \item For amplitude-modulated synthesis, DTW-based losses (e.g., \DTWEnv{}) are recommended over spectrogram based and JTFS. 
-    \item For subtractive/noise-filtered synthesis, spectrogram based losses are most effective.
-Based on our results, SIMSE appears to be the better method of spectrogram comparisons relative to L1, but the amplitude agnosticism of SIMSE likely plays a role in its poor performance in additive synthesis.
-    \item In our amplitude modulation experiments, \JTFS{} did not outperform other measures, contradicting prior claims of its superiority for mesostructures~\cite{vahidi2023mesostructures}.
-Thus, our results call into question its effectiveness as a general mesostructure measure
-    \item Iterative differentiable optimization via Faust-to-JAX is a viable strategy for defining various differentiable DSP functions, requiring only modest hardware resource and providing a more natural approach to synthesizer definition.
-\end{itemize}
-
-\subsection{Applicability to More Complex Synthesizers}
-The experiments in this study were conducted with differentiable synthesizers chosen to isolate fundamental synthesis principle with 2 parameters each.
-While this level of simplicity facilitates controlled comparisons, it does not fully capture the richness of real-world synthesizers, which can have hundreds of parameters with various routes.
-As a result, the extent of the generalizability of our findings does require further research.
-
-Nonetheless, we believe that the insights here can extend naturally to more complex domains.
-Depending on the feature of interest, the observed dependence of loss performance on synthesis method is will likely remain stable even with added complexities such as layers of modulation, filtering, or nonlinearity.
-
-\DTWEnv{} measures very specific features of audio, and its use in the recommended settings---where low frequency amplitude modulation matching is required---would likely yield the desired results regardless of synthesizer complexity.
-Likewise, the use of \SIMSESpec{} or \LoneSpec{} for matching filter cut-offs is likely to be successful regardless of the underlying sound.
-
-% Third, the iterative differentiable optimization framework proved viable even on limited hardware, which indicates that similar pipelines could be applied to larger differentiable synthesizers if computational resources allow.
-
-% At the same time, real-world synthesizers present new challenges. 
-% High-dimensional parameter spaces may exacerbate issues of local minima and make optimization less stable, particularly when loss functions are only partially informative.
-
-% Non-differentiable components (e.g., physical modeling, complex routing) also limit the direct applicability of gradient-based methods.
-
 % Future work should therefore investigate hybrid approaches, combining differentiable blocks with surrogate models or gradient-free optimization for non-differentiable components. 

### 2025-09-19 sections/summary_weakness_conclusion.tex
Updates from Overleaf

@@
+\subsection{Takeaways}
+Perhaps the most important takeaway regarding our main hypothesis is that the loss function that yields the best sound-matching outcomes varies depending on the synthesizer program.
+This is likely due to the interaction between how the parameters of the synthesizer influence the sound, and the core sonic features used by the loss function.
+
+\textbf{Q1}: To what extent do automatic evaluation metrics agree with manual listening tests?
+we see somewhat consistent results between the rankings assigned by manual hearing tests (which we take as the ground truth), and automatic measures of P-Loss and MSS.
+With the exception of P-Loss in \BPNoise{} narrowly selecting a different spectrogram based loss, all measures of performance selected the same top performer.
+Due to the simplicity of our programs and the occasional deviations, we cannot conclude that these automatic measures are substitutes for human hearing tests.
+As many previous works have done\todo{cite}, the use of MSS and/or P-Loss as general loss functions does seem appropriate, however, we advise the use of listening tests or custom made loss functions for more specific analysis.
+
+\textbf{Q2}: Are DTW and SIMSE effective measures of loss? If so,
+
+ \DTWEnv{} consistently outperformed other measures in amplitude-modulated synthesis, where envelope alignment is critical.
+\SIMSESpec{} performed best in subtractive synthesis, where scale-invariance captures noise filtering effects more robustly than L1-based measures.
+These results indicate that the advantages of novel losses are context-dependent, and the necessity of further creative approaches to differentiable sound-similarity measures.
+
+\textbf{Q3}: Can iterative differentiable optimization be applied effectively in synthesizers using classical DSP functions?
+The approach of designing DSP functions in Faust and transpiling to differentiable Faust code yielded meaningful outputs and enabled comparative evaluation of losses.
+The use of Faust for synthesizer definition was convenient, and transpiling to Jax enabled quick gradient optimization.
+We were able to run our 1200 experiments within 72 hours on a laptop without a GPU\footnote{Lenovo ThinkPad T480, i5-8350U CPU at 1.70GHz, 32 GB RAM.}.
+
+\subsection{Practical Recommendations:}
+Based on our findings, several actionable guidelines emerge.
+These heuristics can guide practitioners in selecting appropriate similarity measures and synthesis methods for future sound-matching experiments.
+\begin{itemize}
+    \item In leu of listening tests, MSS or P-Loss are generally useful for large-scale benchmarking, but confirm conclusions with manual listening when accuracy in isolated cases is important. 
+    \item For amplitude-modulated synthesis, DTW-based losses (e.g., \DTWEnv{}) are recommended over spectrogram based and JTFS. 
+    \item For subtractive/noise-filtered synthesis, spectrogram based losses are most effective.
+Based on our results, SIMSE appears to be the better method of spectrogram comparisons relative to L1, but the amplitude agnosticism of SIMSE likely plays a role in its poor performance in additive synthesis.
+    \item In our amplitude modulation experiments, \JTFS{} did not outperform other measures, contradicting prior claims of its superiority for mesostructures~\cite{vahidi2023mesostructures}.
+Thus, our results call into question its effectiveness as a general mesostructure measure
+    \item Iterative differentiable optimization via Faust-to-JAX is a viable strategy for defining various differentiable DSP functions, requiring only modest hardware resource and providing a more natural approach to synthesizer definition.
+\end{itemize}
+
+\subsection{Applicability to More Complex Synthesizers}
+The experiments in this study were conducted with differentiable synthesizers chosen to isolate fundamental synthesis principle with 2 parameters each.
+While this level of simplicity facilitates controlled comparisons, it does not fully capture the richness of real-world synthesizers, which can have hundreds of parameters with various routes.
+As a result, the extent of the generalizability of our findings does require further research.
+
+Nonetheless, we believe that the insights here can extend naturally to more complex domains.
+Depending on the feature of interest, the observed dependence of loss performance on synthesis method is will likely remain stable even with added complexities such as layers of modulation, filtering, or nonlinearity.
+
+\DTWEnv{} measures very specific features of audio, and its use in the recommended settings---where low frequency amplitude modulation matching is required---would likely yield the desired results regardless of synthesizer complexity.
+Likewise, the use of \SIMSESpec{} or \LoneSpec{} for matching filter cut-offs is likely to be successful regardless of the underlying sound.
+
+% Third, the iterative differentiable optimization framework proved viable even on limited hardware, which indicates that similar pipelines could be applied to larger differentiable synthesizers if computational resources allow.
+
+% At the same time, real-world synthesizers present new challenges. 
+% High-dimensional parameter spaces may exacerbate issues of local minima and make optimization less stable, particularly when loss functions are only partially informative.
+
+% Non-differentiable components (e.g., physical modeling, complex routing) also limit the direct applicability of gradient-based methods.
+
 \section{Summary, Weaknesses, and Conclusion}

### 2025-09-21 sections/methodology_and_analysis.tex
Updates from Overleaf

@@
 \label{sec:experiment_setup}
-We systematically evaluate iterative sound-matching pipelines that combine
-different differentiable loss functions with differentiable synthesizers.
-By ``performance'' we mean the similarity of the synthesized output to a
-target, assessed via P-Loss, MSS, and---most importantly---manual listening
-tests. This setup allows us to test our central hypothesis: that the relative
-performance of a loss function is contingent on the synthesis method, and
-thus no universal ``best'' similarity measure exists.
+At a glance, our methodology is to conduct controlled, low-level differentiable experiments—without proxy networks—by pairing four DSP-based synthesizers with four differentiable loss functions.
+We optimize parameters directly via gradients in order to isolate loss–synthesizer interactions.
+This controlled setup systematically evaluates the performance of iterative sound-matching pipelines, where “performance” is defined as the similarity of the synthesized output to a target.
+Similarity is assessed with P-Loss, MSS, and—most importantly—manual listening tests.
+This design allows us to directly test our central hypothesis: that the effectiveness of a loss function depends on the synthesis method, and that no universal “best” similarity measure exists.
 
-We evaluate four loss functions (Section~\ref{sec:loss_implementation})
+Here we discuss the implementation of the four loss functions (Section~\ref{sec:loss_implementation})
 across four differentiable synthesizers (Section~\ref{sec:programs}). For
@@
 Faust is a functional language for audio synthesis that can succinctly define signal processing chains.
-Following Braun's methodology~\cite{braun2024dac}, programs are first defined in Faust, then converted (or \textit{transpiled}) to differentiable Jax functions using the DawDreamer library\footnote{\url{https://github.com/DBraun/DawDreamer}}.
+Following Braun's methodology~\cite{braun2024dac}, programs are first defined in Faust, then converted (or \textit{transpiled}) to differentiable JAX functions using the DawDreamer library\footnote{\url{https://github.com/DBraun/DawDreamer}}.
 \begin{lstlisting}[caption={\BPNoise}, label={lst:program0}, language=Faust,
@@
 Using the significance level of 0.05, if the H statistic is greater than the critical value of the chi-square distribution then the null hypothesis is rejected, meaning that at least one group has a mean rank significantly above others~\cite{kruskal1952use}.
-\todo{More detail for NPSK and leadup to results}
 
@@
 \subsection{Best Performers For Each Program}
+Table~\ref{tab:kruskal_auto} shows the Kruskal-Wallis results for every program and evaluation method, we see significant differences between loss function performance measures (i.e., the bootstrapped distributions of the evaluation for each loss function) in 11 of the 12 cases, with the exception of  MSS for the \AmpMod{} program.
 
-Table~\ref{tab:kruskal_auto} shows the Kruskal-Wallis results for every program and evaluation method, we see a significant differences between loss function performance measures (i.e., the bootstrapped distributions of the evaluation for each loss function) in 11 of the 12 cases, with the exception of  MSS for the \AmpMod{} program.
+Based on the bootstrapped distribution of the scores, the NPSK algorithm ranks the loss functions from 1 (best) to a maximum of 4 (worst).
+In cases where the distributions are similar, multiple loss functions can be clustered into the same rank.
 
-Based on the bootstrapped distribution of scores, the NPSK algorithm ranks the loss functions from 1 (best) to a maximum of 4 (worst).
-In cases where the distributions are similar, multiple loss functions can be clustered into the same rank.
-In Figure~\ref{fig:npsk_all}, we use color-coded violin plots to visualize the best performers for each program.
-The corresponding colors for each rank are
+In Figure~\ref{fig:npsk_all}, we use color-coded violin plots to visualize the best performers for each program. The corresponding colors for each rank are
 \colorbox{rank1}{\textcolor{black}{\textbf{1}}} \colorbox{rank2}{\textcolor{white}{\textbf{2}}} \colorbox{rank3}{\textcolor{white}{\textbf{3}}} \colorbox{rank4}{\textcolor{black}{\textcolor{white}{\textbf{4}}}}.
@@
 \subsubsection{\AmpMod}
-IP-Loss and manual hearing tests yield identical rankings, and MSS results vary for for ranks 2 to 4.
+P-Loss and manual hearing tests yield identical rankings, and MSS results vary for ranks 2 to 4.
 It is not surprising that MSS results differ from the hearing tests, as we saw in Table~\ref{tab:kruskal_auto}, MSS showed no significant differences between groups.

### 2025-09-21 sections/summary_weakness_conclusion.tex
Updates from Overleaf

@@
-\subsection{Takeaways}
-Perhaps the most important takeaway regarding our main hypothesis is that the loss function that yields the best sound-matching outcomes varies depending on the synthesizer program.
+\subsection{Key Findings}
+The most important takeaway regarding our main hypothesis is that the loss function that yields the best sound-matching outcomes varies depending on the synthesizer program.
 This is likely due to the interaction between how the parameters of the synthesizer influence the sound, and the core sonic features used by the loss function.
@@
 \textbf{Q1}: To what extent do automatic evaluation metrics agree with manual listening tests? 
-we see somewhat consistent results between the rankings assigned by manual hearing tests (which we take as the ground truth), and automatic measures of P-Loss and MSS.
+\\We see somewhat consistent results between the rankings assigned by manual hearing tests (which we take as the ground truth), and automatic measures of P-Loss and MSS.
 With the exception of P-Loss in \BPNoise{} narrowly selecting a different spectrogram based loss, all measures of performance selected the same top performer.
-Due to the simplicity of our programs and the occasional deviations, we cannot conclude that these automatic measures are substitutes for human hearing tests.
+However, given the simplicity of our programs and the occasional deviations, we cannot conclude that these automatic measures are substitutes for human hearing tests.
 As many previous works have done\todo{cite}, the use of MSS and/or P-Loss as general loss functions does seem appropriate, however, we advise the use of listening tests or custom made loss functions for more specific analysis.
 
-\textbf{Q2}: Are DTW and SIMSE effective measures of loss? If so,
+\textbf{Q2}: Are DTW and SIMSE effective measures of loss? If so, in what contexts?\\
+\DTWEnv{} consistently outperformed other measures in amplitude-modulated synthesis, where envelope alignment is critical.
+\SIMSESpec{} performed best in subtractive synthesis, where scale-invariance captures noise-filtering effects more robustly than L1-based measures.
+These results suggest that the advantages of these novel losses are context-dependent, underscoring the need for further exploration of creative differentiable similarity measures.
 
- \DTWEnv{} consistently outperformed other measures in amplitude-modulated synthesis, where envelope alignment is critical.
-\SIMSESpec{} performed best in subtractive synthesis, where scale-invariance captures noise filtering effects more robustly than L1-based measures.
-These results indicate that the advantages of novel losses are context-dependent, and the necessity of further creative approaches to differentiable sound-similarity measures.
-
-\textbf{Q3}: Can iterative differentiable optimization be applied effectively in synthesizers using classical DSP functions?
+\textbf{Q3}: Can iterative differentiable optimization be applied effectively in synthesizers using classical DSP functions? \\
 The approach of designing DSP functions in Faust and transpiling to differentiable Faust code yielded meaningful outputs and enabled comparative evaluation of losses.
@@
 \begin{itemize}
-    \item In leu of listening tests, MSS or P-Loss are generally useful for large-scale benchmarking, but confirm conclusions with manual listening when accuracy in isolated cases is important. 
+    \item In lieu of listening tests, MSS or P-Loss are generally useful for large-scale benchmarking, but confirm conclusions with manual listening when accuracy in isolated cases is important. 
     \item For amplitude-modulated synthesis, DTW-based losses (e.g., \DTWEnv{}) are recommended over spectrogram based and JTFS. 

### 2025-09-22 sections/introduction.tex
Updates from Overleaf

@@
 The possibilities afforded by these synthesizers is infinite, leading to their widespread adaption by artists and sound-designers~\cite{lyons1997understanding,russ1999sound,stranneby2004digital}.
-A typical synthesizer has a number of parameterizable functions which affect the output sound in various ways, and manual sound-design often involves modifying the parameters until a conceptualized target sound is reached.
+A typical synthesizer has a number of parameterizable functions which affect the sound output in various ways, and manual sound-design often involves modifying the parameters until a conceptualized target sound is reached.
 The automation of this approach has commonly been referred to as "sound-matching", with reduction of tinkering time and creation of ``interesting" sounds as the main motivators ~\cite{krekovic2019insights,turian2020sorry,horner1993machine,salimi2020make,esling2019flow,engel2020ddsp,mitchell2007evolutionary,shier2020spiegelib,masuda2021soundmatch,masuda2023improving}.
@@
 Is there a universally best performing loss function for audio?
+Is there a need for further development of bespoke loss functions?
 
@@
 (Q2) Can loss functions based on Dynamic Time Warping (DTW) and Scale-Invariant Mean Squared Error (SIMSE) provide advantages over SOTA loss functions, and under what conditions?
-(Q3) Is an iterative differentiable optimization a viable strategy for design of sound-matching experiments?
+(Q3) Is iterative differentiable optimization a viable strategy for design of sound-matching experiments?
 
- We adopt an \textit{iterative} and \textit{differentiable} approach to defining our sound-matching experiments.
+ We adopt a lesser used \textit{iterative} and \textit{differentiable} approach to defining sound-matching experiments.
 The iterative approach better mimics the manual process of recursive listening and parameter adjustments towards sound design, while a differentiable environment allows access to the loss function gradients that can be used to better understand the nature of the problem.

### 2025-09-23 sections/background_related.tex
Updates from Overleaf

@@
 \label{sec:background_related}
-
  Sound-matching sits at the intersection of digital signal processing, audio representation, and optimization.
-In this section we first formalize the sound-matching task, then review core synthesis methods, similarity measures, and optimization approaches. 
+Here we first formalize the sound-matching task, then review core synthesis methods, similarity measures, and optimization approaches.
 We conclude with a survey of related work and a discussion of gaps in the field.
@@
 In the following sections we will discuss the various components of sound-matching, the subjectivity of ``correct'' solutions, and important issues in the field.
+
+ Figure~\ref{fig:sound_design_loop_iterative} shows a general model for iterative sound-matching.
+To begin the process, the parameters are arbitrarily initialized (usually random generation), the similarity of the target and output is measured, and the parameter is \textit{optimized} with the goal of increasing the similarity, or alternatively, reducing the loss.
+This process repeats until termination.
+\begin{figure}[ht]
+    \centering
+\begin{tikzpicture}[node distance=2cm, auto]
+
+    % Nodes
+    \node (start) [text centered] {\( g (\hat{\theta}) = x \)};
+    \node (L) [above of=start,right of=start, text centered] {\( L(\hat{\theta},t) {\ \propto \ } \phi(x) - \phi(t) \)};
+    \node (optimize) [below of=L,right of=L, text centered] {Optimize $\hat{\theta}$};
+    \node (new_theta) [below of=optimize,right of = start, text centered] {New \( \hat{\theta} \)};
+
+    % Highlight and arrow for the start node
+    \draw[->, very thick, red] (start) ++(-2,1.5) -- (start)
+        node[midway, below, align=center, sloped, color=red] {Start}
+        node[midway, above, align=center, sloped, color=red] {Random $\hat{\theta}$};
+
+    % Arrows with multi-line labels
+    \draw[->, bend left] (start) to node[midway, right, align=center] {} (L);
+    \draw[->, bend left] (L) to node[midway, right, align=center] {} (optimize);
+    \draw[->, bend left] (optimize) to node[midway, below, align=center] {} (new_theta);
+    \draw[->, bend left] (new_theta) to node[midway, left, align=center] {} (start);
+
+\end{tikzpicture}
+    % \caption{ Iterative approach to sound design.
+Synthesizer $g$ with arbitrary initialized parameters $\hat{\theta}$, creates sound $x$.
+The target sound $t$ is used as the goal.
+Parameters $\hat{\theta}$ are adjusted to minimize error (or loss) $L(\hat{\theta},t)$, where $L$ is proportional to the difference between the representations of $x$ and $t$.
+$\phi$ is the audio feature extractor, or representation function.}
+    \caption{ Iterative approach to sound design.}
+    \label{fig:sound_design_loop_iterative}
+\end{figure}
 
@@
 Early sound-matching research focused heavily on frequency and amplitude modulation (\gls{FM}/\gls{AM}) synthesis~\cite{horner1993machine,mitchell2007evolutionary,vahidi2023mesostructures}, which is simple to implement yet expressive~\cite{chowning1973synthesis}.
-Other approaches studied in isolation include additive and subtractive synthesis~\cite{engel2020ddsp,masuda2023improving,salimi2020make} and physical modeling~\cite{riionheimo2003parameter,han2024learning}.
-
-\highlight{By \textit{isolation}, we mean settings where the effect of individual parameters on the output sound remains tractable. 
-Recent work has also explored commercial Virtual Studio Technology (VST) synthesizers~\cite{yee2018automatic,esling2019flow}. 
-While powerful, such systems obscure the interactions between modules, losses, and outputs. 
-In contrast, small-scale differentiable synthesizers built from isolated DSP blocks offer interpretability, making them particularly useful for benchmarking. }
+Other synthesis methods studied in \textit{isolation} include additive and subtractive synthesis~\cite{engel2020ddsp,masuda2023improving,salimi2020make} and physical modeling~\cite{riionheimo2003parameter,han2024learning}.
+By \textit{isolation}, we mean settings where the effect of individual parameters on the output sound remains tractable.
+For example, we exclude studies using commercial Virtual Studio Technology (VST) which can obscure the interactions between modules, losses, and outputs.
 
@@
 Typically, for the implementation of P-Loss the parameter sets are treated as vectors in space, and L1 or L2 distance is applied.
-There are two major limitations to this approach: First, the target and output sound must be made by the same synthesizer; otherwise the parameter sets cannot be compared (see Section~\ref{sec:matching_types}).
+There are two major limitations to this approach: First, the target and output sound must be made by the same synthesizer; otherwise the parameter sets cannot be compared (see Section~\ref{sec:in-domain}).
 Second, the relationship between synthesizer parameters and the audio output is not linear~\cite{shier2020spiegelib,han2023perceptual,esling2019flow}.
@@
 
+\begin{table*}[t]
+\centering
+\caption{Summary of select works in sound-matching.
+Gen.
+refers to whether the generations are done in 1-shot (generally, output by a neural network) or a search is conducted with the goal of iteratively getting closer to the target sound.}
+\setlength{\tabcolsep}{3pt} % Reduce horizontal spacing
+\renewcommand{\arraystretch}{1.05} % Reduce vertical spacing
+\scriptsize % Reduce font size
+\resizebox{\textwidth}{!}{
+\begin{tabular}{|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{1.2cm}|>{\raggedright\arraybackslash}m{1.2cm}|>{\raggedright\arraybackslash}m{1cm}|>{\raggedright\arraybackslash}m{1cm}|}
+\hline
+\textbf{Work} & \textbf{Synthesis} & \textbf{Loss} & \textbf{Heuristics} & \textbf{Goal} & \textbf{Domain} & \textbf{Year} & \textbf{Gen. } \\
+\hline
+Horner \textit{\textit{et al.}}~\cite{horner1993machine} & Non-differentiable FM & STFT (McAulay-Quatieri) & GA & Replicate & In & 1991 & Iter \\
+\hline
+Mitchel \textit{\textit{et al.}}~\cite{mitchell2007evolutionary} & Non-differentiable FM & Spectral Relative & Evolutionary & Replicate & In (Contrived) & 2007 & Iter \\
+\hline
+Yee-King \textit{\textit{et al.}}~\cite{yee2018automatic} & Non-differentiable VST & P-Loss (for Supervised), Spectral (MFCC) & Supervised \& Evolutionary & Replicate & In & 2018 & Both \\
+\hline
+Esling \textit{\textit{et al.}}~\cite{esling2019flow} & Non-differentiable VST & Spec MSS/SC P-Loss & Supervised \& Modeling & Replicate & In* \& Out & 2019 & 1-shot \\
+\hline
+Shier \textit{\textit{et al.}}~\cite{shier2020spiegelib} & Custom, non-diff & FFT \& Spectral & Supervised \& Direct & Replicate & In & 2020 & Both \\
+\hline
+Salimi \textit{\textit{et al.}}~\cite{salimi2020make} & DSP & STFT and Envelope & Supervised & Imitate & Out & 2020 & 1-shot \\
+\hline
+Masuda \textit{\textit{et al.}}~\cite{masuda2021soundmatch} & DDSP & P-Loss initially, fine-tune with Spec. & Supervised & Both & In \& Out & 2021 & 1-shot \\
+\hline
+Han \textit{\textit{et al.}}~\cite{han2023perceptual} & NN-encoder $\rightarrow$ physical model & PNP (approx. of STFT) & Supervised & Replicate & In & 2023 & 1-shot \\
+\hline
+Vahidi \textit{\textit{et al.}}~\cite{vahidi2023mesostructures} & Differentiable FM & JTFS/MSS & Gradient desc. & Replicate & In & 2023 & Iter \\
+\hline
+Barkan \textit{et al.}~\cite{barkan2023inversynthII}& Diff. Synthesizer Proxy& P-Loss + L1 Spec (weighted) &  Supervised (init.), Semi-Supervised & Replicate& In & 2023 & Both
+\\
+\hline
+Uzrad \textit{\textit{et al.}}~\cite{uzrad2024diffmoog} & DDSP & Synth. Chain and P-Loss & Supervised & Replicate & In \& Out & 2024 & 1-shot \\
+\hline
+Cherep \textit{et al.}~\cite{creativecherep2024} & DDSP & CLAP & Evolutionary & Imitate & Out & 2024 & Iter \\
+\hline
+\end{tabular}
+}
+
+\label{tab:summary}
+\end{table*}
+
 \subsection{Historical Framing of Sound-Matching}
@@
 This approach proved more effective---i.e., achieved lower multi-level spectral difference in out-of-domain tests---than baseline models, which were either not exposed to out-of-domain sounds or trained exclusively with P-Loss.
-Subjective hearing tests were conducted, showing a preference for the fine-tuned model~\cite{masuda2022soundmatch}.
+Subjective hearing tests were conducted, showing a preference for the fine-tuned model~\cite{masuda2021soundmatch}.
 Masuda \textit{et al.} later extended this work with semi-supervised learning, highlighting significant gaps in in-domain and out-of-domain performance~\cite{masuda2023improving}.

### 2025-09-23 sections/summary_weakness_conclusion.tex
Updates from Overleaf

@@
 \end{itemize}
+
+\subsection{Illustrative Loss Landscapes}
+
+To complement our quantitative, statistically robust analysis, we provide qualitative examples of loss landscapes to illustrate why specific losses succeed or fail in particular synthesis contexts.
+These examples should not be seen as conclusive evidence, but rather as illustrative aids to understanding the mechanisms behind the numerical results.
+
+Figure~\ref{fig:loss_landscape_noisebp} shows the case of \BPNoise{} modified to use only a high-pass filter parameter.
+Here, \LoneSpec{}, \SIMSESpec{}, and \JTFS{} exhibit clear global minima near the correct parameter (red dashed line), whereas \DTWEnv{} remains relatively flat around the target, making gradient-based exploration more challenging.
+This observation explains DTW’s weaker performance in this context.
+
+In contrast, Figure~\ref{fig:loss_landscape_ampmod} depicts a simplified version of \AmpMod{} with only the amplitude-modulation parameter (range: 0.1–20\,Hz).
+In this case, \DTWEnv{} yields the smoothest and most informative landscape, strongly correlating with proximity to the target.
+This aligns with its superior performance in amplitude-modulated synthesis observed in our experiments.
+
+These examples underscore the importance of loss landscapes in shaping optimization dynamics and further motivate future work on landscape-aware loss design.
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.9\linewidth]{images/experiment_plots/comparing_loss_landscapes_normalized.png}
+    \caption{Loss landscapes for \BPNoise{} with only a high-pass filter parameter. 
+    \LoneSpec{}, \SIMSESpec{}, and \JTFS{} show clear global minima near the correct parameter, while \DTWEnv{} remains flat around the target.}
+    \label{fig:loss_landscape_noisebp}
+\end{figure}
+
+\begin{figure}[ht]
+    \centering
+    \includegraphics[width=0.9\linewidth]{images/experiment_plots/comparing_lanscapes_1_1d_normalized.png}
+    \caption{Loss landscapes for a simplified \AmpMod{} synthesizer with only amplitude modulation parameter. 
+    \DTWEnv{} exhibits the smoothest and most informative landscape, explaining its superior performance in amplitude-modulated synthesis.}
+    \label{fig:loss_landscape_ampmod}
+\end{figure}
 

### 2025-09-24 sections/background_related.tex
Updates from Overleaf

@@
 The effects of using different isolated DSP functions in iterative sound-matching and how these interact with loss functions remains largely untested.
-    \item \PeriodicLoss: The sinusoidal nature of loss function landscapes is a problem that appears frequently\todo{cite and make clearer}.
-In differentiable settings, this problem causes gradient descent updates to local minima.
+    \item \PeriodicLoss: Loss function landscapes are not easy to navigate, as there can be many local minima, or large flat areas~\cite{turian2020sorry,vahidi2023mesostructures}.
+In differentiable settings, this can cause gradient descent updates to not reach the global minima.
     \item \OutDomain: Sound-matching with out-of-domain sounds is a largely unexplored area, yet a necessary one for practical applications for sound designers.

### 2025-09-24 sections/summary_weakness_conclusion.tex
Updates from Overleaf

--- 2025-09-21 sections/summary_weakness_conclusion.tex
@@
+\section{Discussion}
 \subsection{Key Findings}
@@
 \end{itemize}
+
+\subsection{Illustrating Loss Landscapes}
+\label{sec:loss_landscape_examples}
+To complement our quantitative analysis, we provide examples of loss landscapes that illustrate why specific losses succeed or fail in certain synthesis contexts.
+These are illustrative aids rather than conclusive evidence, and highlight the importance of \PeriodicLoss{} discussed earlier in Section~\ref{sec:lacking}.
+
+We show the landscapes of the loss functions with simplified versions of \BPNoise{} (HP-Noise) and \AmpMod{} (S-Noise-AM), each reduced to a single parameter for one-dimensional visualization.
+HP-Noise uses only a high-pass filter with cutoff between 100–20,000 Hz, while S-Noise-AM varies only the modulation rate between 0.1–20 Hz.
+Loss values are normalized to 0–1 to allow comparison across functions.
+
+Figure~\ref{fig:loss_landscape_noisebp} shows the HP-Noise results.
+\LoneSpec{}, \SIMSESpec{}, and \JTFS{} exhibit clear minima at the correct parameter (red dashed line), whereas \DTWEnv{} remains relatively flat, making gradient-based optimization more difficult.
+
+Figure~\ref{fig:loss_landscape_ampmod} shows the S-Noise-AM results.
+Here, \DTWEnv{} produces the smoothest and most informative landscape, while spectrogram losses are minimized near the target but lack consistent correlation with parameter distance.
+\JTFS{} remains largely flat.
+This aligns with \DTWEnv's superior performance in amplitude-modulated synthesis observed in our experiments.
+
+\begin{figure*}[ht]
+    \centering
+    \begin{minipage}[t]{0.48\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{images/experiment_plots/comparing_loss_landscapes_normalized.png}
+        \caption{Loss landscapes for \BPNoise{} with only a high-pass filter parameter. 
+        \LoneSpec{}, \SIMSESpec{}, and \JTFS{} show clear global minima near the correct parameter, while \DTWEnv{} remains flat around the target.}
+        \label{fig:loss_landscape_noisebp}
+    \end{minipage}%
+    \hfill
+    \begin{minipage}[t]{0.48\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{images/experiment_plots/comparing_lanscapes_1_1d_normalized.png}
+        \caption{Loss landscapes for a simplified \AmpMod{} synthesizer with only amplitude modulation parameter. 
+        \DTWEnv{} exhibits the smoothest and most informative landscape, explaining its superior performance in amplitude-modulated synthesis.}
+        \label{fig:loss_landscape_ampmod}
+    \end{minipage}
+\end{figure*}
 

### 2025-09-24 reviews.tex
Updates from Overleaf

\documentclass[11pt]{article}

\usepackage[letterpaper,margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage[strings]{underscore} % allow _ in text
\usepackage{natbib}

\hypersetup{
  colorlinks=true,
  linkcolor=blue!60!black,
  urlcolor=blue!60!black,
  citecolor=blue!60!black
}

% Compact lists
\setlist[itemize]{leftmargin=1.2em,itemsep=0.3em,topsep=0.3em}
\setlist[enumerate]{leftmargin=1.4em,itemsep=0.3em,topsep=0.3em}

% Handy macros used in your text
\newcommand{\DTWEnv}{\textsc{DTW\_Envelope}}
\newcommand{\SIMSESpec}{\textsc{SIMSE\_Spec}}
\newcommand{\LoneSpec}{\textsc{L1\_Spec}}
\newcommand{\JTFS}{\textsc{JTFS}}

\title{\vspace{-0.5em}Response to Reviewers\\
\large Manuscript: \emph{Evaluating Sound Similarity Metrics for Differentiable, Iterative Sound-Matching}}
\author{}
\date{}

\begin{document}
\maketitle
\vspace{-1.25em}

\noindent
We sincerely thank you and the reviewers for the thoughtful and constructive feedback on our manuscript titled~\emph{``Evaluating Sound Similarity Metrics for Differentiable, Iterative Sound-Matching''}. We have carefully considered all comments and revised the manuscript accordingly. Below, we provide a detailed point-by-point response to each comment. 
For each reviewer, we first provide the review in full, with the key concerns highlighted. Below the full review, we address the concerns in sections. Each section contains the reviewer's key requests or issues, followed by the changes necessary. Some of our revisions are not direct responses to the concrete, point by point suggestions made by the reviewers, but address more general concerns expressed in the body of the reviews (simplification of the introduction and background requested by reviewer 1 for example). As such, for the sake of brevity we took the liberty to paraphrase and aggregate the reviewer's requests and concerns as accurately as possible when necessary.

\medskip
\noindent
We hope the revisions address all concerns and improve the clarity and quality of the manuscript.

\section*{Reviewer 1}

\noindent
The Introduction section lacks a clear articulation of the central research question early on. The core problem that no single loss function performs optimally across synthesizer types is buried within extended historical context and repeated definitions. Moreover, the use of a single subsection label (A. Digital Synthesis with Sound-Matching) without any follow-up sections is structurally inconsistent. \textbf{We recommend tightening the Introduction to focus more directly on problem motivation and contribution, and removing or completing the subsection structure.} \\
\textbf{The current division of the manuscript into Introduction, Background, and Past Works results in significant conceptual overlap and repetition.} Historical framing, loss function taxonomy, and method classification are spread across all three sections. \textbf{A clearer organizational approach would be to merge Background, and Past Works into a unified Background and Related Work section, where formal definitions, theoretical framing, and literature review are presented coherently.} This would allow the Introduction to focus solely on the motivation, problem framing, and contribution. \\
Section IV titled ``Programs, Methodology, and Results'' is overloaded and blends experimental setup with results and interpretation. \textbf{For clear clarity and alignment with structural norms, we recommend splitting this into three sections: (1) Experimental Setup (including synthesizer programs, loss definitions, optimization procedure); (2) Experimental Results (performance metrics, statistical tests, and rankings); and (3) Discussion (qualitative interpretation, synthesis-specific observations, and practical implications).} \\
Conceptually, the work serves primarily as a benchmarking study rather than a novel methodological contribution. While the comparison is comprehensive, the novel losses (DTW\_Envelope, SIMSE\_Spec) are adaptations of known techniques and are not framed with sufficient theoretical justification. \textbf{There is also limited interpretability analysis; for example, the paper does not explore why DTW performs well on modulated noise or why SIMSE outperforms L1 in certain scenarios.} Additionally, the synthesizers used are very simple (only 2--3 parameters), and all experiments are conducted in-domain, limiting the generalizability of the findings to real-world synthesis environments with more complex parameter spaces or out-of-domain targets. \\
\textbf{The manuscript would also benefit from clearer practical implications.} While the authors argue that sound-matching is a creative process and reject the idea of a universal loss, they stop short of offering actionable heuristics. \textbf{Some guidance (e.g., Use DTW for AM synthesis; SIMSE for subtractive synthesis) would strengthen the paper’s relevance to practitioners.} \\
\textbf{The paper would be improved by formally stating additional hypotheses that are currently only implicit.} For example, the observed agreement between manual and automatic evaluation metrics, the effectiveness of the novel losses, and the usability of iterative optimization as a general strategy are all important insights, but are not positioned as testable hypotheses upfront. \\
\textbf{Suggested Actions for Revision:}
\begin{enumerate}
  \item \textbf{Clearly state additional hypotheses regarding metric agreement, perceptual evaluation, and loss novelty.}
  \item \textbf{Deepen the technical and perceptual justification for DTW\_Envelope and SIMSE\_Spec.}
  \item \textbf{Add qualitative or spectral analyses of where specific losses succeed or fail.}
  \item \textbf{Expand discussion to address applicability to more complex or real-world synthesizers.}
  \item \textbf{Provide practical takeaways or usage guidelines for different loss-synth combinations.}
  \item \textbf{Restructure the manuscript to clarify the boundary between methodology, results, and discussion.}
\end{enumerate}

\subsection*{Introduction clarity}
\noindent\textit{Paraphrase of key requests:}
\begin{itemize}
  \item Central research question and core problem not stated early enough.
  \item Historical context and repeated definitions obscure the main motivation.
  \item Introduction should focus solely on the motivation, problem framing, and contribution.
  \item We recommend tightening the Introduction to focus more directly on problem motivation and contribution, and removing or completing the subsection structure.
\end{itemize}

\noindent\textbf{Response:} \\
We agree with the reviewer’s comment regarding lack of clarity and focus in the Introduction. The Introduction has been revised accordingly: 
\begin{itemize}
  \item The Introduction is now a single continuous section, without subsection headings.  
  \item We have removed unnecessary historical context and repeated definitions, reducing the length by roughly 25\%. We moved some of this material to the new \emph{Background and Related Work} section.  
  \item The central hypothesis now appears at the beginning of the third paragraph (previously it appeared mid-way through paragraph 10).  
  \item In addition to the main hypothesis, we have added three explicit research questions, covering metric agreement (Q1), the utility of novel DTW and SIMSE losses (Q2), and the viability of iterative optimization (Q3).  These are stated directly after the central hypothesis.
  \item The Introduction concludes with a concise statement of our contributions, listing our systematic evaluation across synthesizers, the introduction of DTW and SIMSE losses, the discussion of automatic versus manual metrics, a proposed nomenclature for categorizing sound-matching works, and a list of important issues in the field that require further research.  
\end{itemize}

\subsection*{Improve Background and Related Work}
\noindent\textit{Paraphrase of key request:} A clearer organizational approach would be to merge Background and Past Works into a unified Background and Related Work section, where formal definitions, theoretical framing, and literature review are presented coherently.

\noindent\textbf{Response:} \\
We agree that the organization and writing in these sections could be improved. We merged \emph{Background} and \emph{Past Works} into a unified \emph{Background and Related Work} section. We edited these sections to improve reading and remove redundancies between the various subsections while taking precautions in compromising the flow.
The \emph{background and Related Work} section contains the following topics in order:  
\begin{itemize}
  \item A formal definition of sound-matching and its components
  \item Definitions of synthesizers, losses, and optimization heuristics,  
  \item A chronological review of prior sound-matching approaches
  \item A discussion of gaps in the field and four critical issues.
  \item A justification for our methodology, and how our experiments address two of the critical issues (Loss and Synthesizer selection).
\end{itemize}

\subsection*{Deepen the technical and perceptual justification for DTW\_Envelope and SIMSE\_Spec.}
\noindent While the comparison is comprehensive, the novel losses (DTW\_Envelope, SIMSE\_Spec) are adaptations of known techniques and are not framed with sufficient theoretical justification.

\noindent\textbf{Response} \\
We have expanded the theoretical and perceptual justifications for DTW\_Envelope and SIMSE\_Spec in subsection \textbf{II.C Sound Representation and Loss Functions} under \textbf{Background and Previous Works}. For SIMSE\_Spec, we emphasize that scale-invariance removes sensitivity to global gain, making it particularly appropriate for subtractive synthesis where filter cutoffs alter spectral energy but preserve timbral shape. For DTW\_Envelope, we clarify its perceptual motivation: aligning amplitude envelopes is crucial for AM synthesis, where traditional spectrogram L1 distances fail to capture similarity. We explicitly frame these losses not as wholly new techniques, but as carefully adapted methods tested in an iterative differentiable setting, a contribution that has not previously been made.

We did not modify the defintions in \textbf{III.A Loss Function Implementation Details}, dedicating this section to only technical explanations. However, we discuss the empirical effectiveness of these losses in the \textbf{IV.A Key Findings} subsection, as an important takeaway.

\subsection*{Analyses of where specific losses succeed or fail.}
\noindent ``Add qualitative or spectral analyses of where specific losses succeed or fail''

\noindent\textbf{Response} \\
This is an interesting suggestion. Our original experiments were designed to emphasize quantitative, statistically robust measurements of loss function performance, which is why we initially avoided focusing on individual examples of sounds, spectrograms, or loss landscapes (to reduce the risk of cherry-picking). However, we think that illustrative examples can provide helpful intuition for readers and motivate future research.

To this end, we have added qualitative visualizations of loss landscapes (see \textbf{Figures 4 and 5} in the revised manuscript). These examples highlight why specific losses may succeed or fail and reinforce the broader importance of loss landscapes in assessing the appropriateness of similarity measures~\cite{vahidi2023mesostructures,turian2020sorry}. This addition also supports our argument that effective navigation of loss landscapes represents an important gap in current knowledge and a promising direction for future work.

\begin{itemize}
  \item For the HP-Noise synthesizer, \LoneSpec{}, \SIMSESpec{}, and \JTFS{} exhibit clear minima at the target parameter, while \DTWEnv{} remains relatively flat around the target, consistent with its weaker performance in this setting.
  \item For the S-Noise-AM synthesizer, \DTWEnv{} produces the smoothest and most informative landscape, while spectrogram-based losses and \JTFS{} provide less reliable gradients. This aligns with the superior performance of \DTWEnv{} observed in amplitude-modulated synthesis.
\end{itemize}

These additions serve as illustrative examples that reinforce the mechanisms underlying our quantitative findings and highlight the importance of loss landscape navigation as an open area for future research.

\subsection*{Restructure the manuscript to clarify the boundary between methodology, results, and discussion.}
\noindent Section IV titled ``Programs, Methodology, and Results'' is overloaded and blends experimental setup with results and interpretation. For clear clarity and alignment with structural norms, we recommend splitting this into three sections: (1) Experimental Setup (including synthesizer programs, loss definitions, optimization procedure); (2) Experimental Results (performance metrics, statistical tests, and rankings); and (3) Discussion (qualitative interpretation, synthesis-specific observations, and practical implications).

\noindent\textbf{Response} \\
In the revised manuscript we have:
\begin{itemize}
  \item Split the original Section IV into three distinct sections:
    \begin{enumerate}
      \item \textbf{Experimental Setup} -- describes synthesizer programs, loss definitions, optimization procedure, and evaluation metrics.  (The only partial overlap is at the beginning of the subsection \textbf{Best Performers For Each Program}, which transitions between results and discussion.)
      \item \textbf{Experimental Results} -- presents performance metrics, statistical tests (Kruskal--Wallis, NPSK), and rankings of losses across synthesizers.
      \item \textbf{Discussion} -- provides qualitative interpretation, synthesis-specific observations, applicability to complex synthesizers, and practical implications.
    \end{enumerate}
  \item Moved explanatory material regarding experiment setup out of the \textbf{Experimental Results} and into the \textbf{Experimental Setup} section where appropriate, ensuring results sections are now strictly empirical.
  \item Added new subsections (\emph{Takeaways}, \emph{Practical Recommendations}, \emph{Applicability to More Complex Synthesizers}) to make the discussion more explicit and actionable.
\end{itemize}
We believe this major restructuring and rewrite of appropriate sections has clarified the flow of the manuscript and made the boundary between methodology, results, and discussion much clearer.

\subsection*{Expand discussion to address applicability to more complex or real-world synthesizers.}
\noindent 4. Expand discussion to address applicability to more complex or real-world synthesizers

\noindent\textbf{Response} \\
We added a new subsection, \emph{Applicability to More Complex Synthesizers}, that explicitly discusses how the insights from our controlled two-parameter DSP-based synthesizers generalize to richer synthesizers with many parameters and complex routing. While acknowledging the limitations of our simplified test bed, we argue that the observed dependencies between synthesis method and loss function remain informative even in more complex settings.

\subsection*{Provide practical takeaways or usage guidelines for different loss--synth combinations.}
\noindent 5. Provide practical takeaways or usage guidelines for different loss-synth combinations.

\noindent\textbf{Response} \\
We introduced a new subsection, \emph{Practical Recommendations}, which distills our findings into actionable heuristics for practitioners. DTWEnv is recommended for amplitude-modulated synthesis, SIMSE for subtractive/noise-filtered synthesis, and MSS or P-Loss as large-scale proxies for sound-matching evaluation (with manual validation advised). We also note that our results challenge prior claims about JTFS’s universal superiority for mesostructures.

\section*{Reviewer 2}

\noindent
The paper considers the evaluation of different audio similarity metrics for sound matching applications.  The authors categorize prior methods largely as direct optimization via gradient-free iterative search or open-loop inference via a supervised deep learning model.  Moreover, different approaches which may or may not accommodate out-of-domain sounds.  Accordingly, the authors propose the combined use of differentiable synthesizers and closed-loop iteration as a sound means of achieving synthesizer sound matching, effectively combining the advantages of otherwise disparate methods (i.e. access to gradients while iteratively improving matching results).  Additionally, within their evaluation of different similarity metrics (parameter losses, spectral distances, etc.), they also introduce a DTW-based similarity metric to partially neutralize the effect of time alignment on similarity scores.  A primary conclusion of the work is that there is no one best sound similarity measure capable of working across all synthesizers.  \\
The subject matter of the work is relevant, the experimental setup is satisfactory, and I do not have reason to doubt the correctness of their analysis.  However, my primary concerns with this work are that:  
\begin{enumerate}
  \item \textbf{It is not clear what the proposed method really is, and the paper reads largely as a readout of evaluations that were conducted by the authors.}  
  \item \textbf{The authors claim that differentiable iterative sound matching is largely unexplored, encouraging development of techniques in this area.}  
    \begin{itemize}
      \item \textbf{Barkan et. al, ``Inversynth II: Sound Matching via Self-Supervised Synthesizer-Proxy and Inference-Time Finetuning'' explored precisely this, training an initial ``one-shot'' supervised model, and then performing inference-time optimization, updating model weights to improve the matching against test audio.  Moreover, their use of a synthesizer-proxy negates the explicit need of implementing the synthesizer math differentiably, while still falling under the category of differentiable iterative sound matching.}  
      \item \textbf{Cherep et. al, ``Creative Text-to-Audio Generation via Synthesizer Programming'' uses a differentiable synthesizer and the CLAP score as a similarity metric for iterative optimization.  While the authors lean into synthesizer sound matching against a text query, the method would be readily applicable to audio-to-audio sound matching using the CLAP audio tower alone.}  
    \end{itemize}
  \item \textbf{This work addresses the problems on ``Loss Selection'' and ``Synthesis Selection.''  However, the results as presented here are largely inconclusive (i.e. it is not imminently clear how to select losses/synthesizer pairings), leaving very little in the way of clear takeaways sparking future research.}
\end{enumerate}

\subsection*{Clarity of proposed method}
\noindent It is not clear what the proposed method really is, and the paper reads largely as a readout of evaluations.

\noindent\textbf{Response:} \\
As the reviewer points out, we wanted the revised paper to ``not read merely as a set of experiments''.  In the revised version of the paper, we take steps to ensure that our contributions are clear: (i) controlled benchmarking of loss--synth pairings, (ii) novel, context-dependent benefits of DTW and SIMSE, and (iii) practical heuristics that can guide both future research and applied sound design. These contributions are first stated at the end of the introduction, and expanded on in later in the appropriate sections:

\medskip
\noindent In \textbf{II.B Digital Signal Processing and Synthesis}, we state the importance of isolated experiments:
\begin{quote}\small\ttfamily
Other synthesis methods studied in \textit{isolation} include additive and subtractive synthesis~\cite{engel2020ddsp,masuda2023improving,salimi2020make} and physical modeling~\cite{riionheimo2003parameter,han2024learning}. By \textit{isolation}, we mean settings where the effect of individual parameters on the output sound remains tractable. For example, we exclude studies using commercial Virtual Studio Technology (VST) which can obscure the interactions between modules, losses, and outputs. 
\end{quote}

\noindent To further clarify the methodology and its importance, we have added a new subsection, \emph{What Approach Should We Take?}, placed immediately before the Methodology. This subsection clarifies our experimental design and explains why our contribution is unique. Specifically, we use:  
\begin{enumerate}
  \item Simple differentiable DSP synthesizers to isolate synthesis--loss interactions,  
  \item Multiple losses tested side by side within the same framework, and  
  \item Direct gradient descent optimization without neural proxies.  
\end{enumerate}
This section makes clear that our work is not simply a report of evaluations, but a controlled benchmarking study designed to reveal how losses and synthesis methods interact.

\medskip
\noindent We restate our methodology in the opening paragraph of the \textbf{Methodology} section. Our empirical survey, uses a \emph{controlled experimental design} intended to isolate the interactions between differentiable loss functions and DSP-based synthesizers. Unlike proxy- or embedding-based approaches (e.g., Inversynth II, CLAP), our methodology avoids neural intermediaries and tests fundamental building blocks directly via gradients. This allows us to provide low-level empirical evidence that loss function performance depends strongly on the synthesis method, which we argue is a necessary precursor for generalizable, real-world sound-matching systems.

\subsection*{Relation to prior work}
\noindent The authors claim that differentiable iterative sound matching is largely unexplored, but prior work (Barkan et al., Cherep et al.) could be considered in this category.

\noindent\textbf{Response:} \\
We thank the reviewer for pointing out sound-matching works which fall under the ``differentiable and iterative'' description that were not covered in our literature review. We believe that the term ``under-explored'' is perhaps more appropriate. We've added these works to \textbf{Table 1} which provides a summary of important previous works, and discussed these works in more detail in the \textbf{Background and Related works} subsection titled: \textbf{F. Historical Framing of Sound-Matching}. 
\begin{itemize}
  \item \textbf{Barkan et al. (Inversynth II):} Their work relies on neural proxies for the synthesizer, with inference-time finetuning of the encoder. This constitutes iterative optimization, but the optimization occurs on an approximate model rather than directly on differentiable DSP functions. By contrast, our approach avoids proxy error and directly analyzes how losses behave relative to known synthesis structures.  
  \item \textbf{Cherep et al.:} Their work uses CLAP embeddings with evolutionary search for text-to-audio tasks. While applicable to audio-to-audio matching, CLAP acts as a black-box representation and does not isolate loss--synth interactions. Our approach is complementary, focusing on controlled experiments with simple differentiable DSP synthesizers to expose the strengths and weaknesses of different similarity measures.  In addition, embedding models such as CLAP use simpler representations at their core, commonly the Fourier transformations, and our work explores the low level functions that are often used for the training of embedding models.
\end{itemize}
Looking at these works, we see that our contribution is distinct: we provide the missing \textbf{low-level benchmarking} of loss functions across multiple synthesis methods, which is not addressed by proxy-based or embedding-based approaches.

\subsection*{This work addresses the problems on ``Loss Selection'' and ``Synthesis Selection.''  However, the results as presented here are largely inconclusive (i.e. it is not imminently clear how to select losses/synthesizer pairings), leaving very little in the way of clear takeaways sparking future research.}

\noindent\textbf{Response} \\
In the revised paper, we've added two sections which address the valid issues raised here and by Reviewer 1:
\begin{itemize}
  \item We introduced a new subsection, \emph{Practical Recommendations}, which distills our findings into actionable heuristics for practitioners. DTWEnv is recommended for amplitude-modulated synthesis, SIMSE for subtractive/noise-filtered synthesis, and MSS or P-Loss as large-scale proxies for sound-matching evaluation (with manual validation being strongly advised). We also note that our results challenge prior claims about JTFS’s universal superiority for mesostructures.
  \item We added a new subsection, \emph{Applicability to More Complex Synthesizers}, that explicitly discusses how the insights from our controlled two-parameter DSP-based synthesizers generalize to richer synthesizers with many parameters and complex routing. While acknowledging the limitations of our simplified testbed, we argue that the observed dependencies between synthesis method and loss function remain informative even in more complex settings.
\end{itemize}

\end{document}

### 2025-09-25 review_responses.tex
Updates from Overleaf

\documentclass[11pt]{article}

\usepackage[letterpaper,margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage[strings]{underscore} % allow _ in text

\usepackage[numbers]{natbib} %removed cite package
\usepackage{csquotes}
\usepackage{adjustbox}
\usepackage{listings}
\usepackage{caption}

\usepackage{graphicx}   % for \includegraphics
\usepackage{caption}    % better caption control inside minipage
\usepackage{subcaption} % optional, but useful if you want subfigures later

\hypersetup{
  colorlinks=true,
  linkcolor=blue!60!black,
  urlcolor=blue!60!black,
  citecolor=blue!60!black
}
\setcounter{secnumdepth}{1}
% Compact lists
\setlist[itemize]{leftmargin=1.2em,itemsep=0.3em,topsep=0.3em}
\setlist[enumerate]{leftmargin=1.4em,itemsep=0.3em,topsep=0.3em}

% Handy macros used in your text
\newcommand{\BPNoise}{\textsc{BP-Noise}}
\newcommand{\AmpMod}{\textsc{AmpMod}}
\newcommand{\LossSelect}{\textbf{Loss Selection}}
\newcommand{\SynthSelect}{\textbf{Synthesis Selection}}
\newcommand{\PeriodicLoss}{\textbf{Loss Landscape Navigation}}
\newcommand{\OutDomain}{\textbf{Out-Domain Generation}}
\newcommand{\DTWEnv}{\textsc{DTW\_Envelope}}
\newcommand{\SIMSESpec}{\textsc{SIMSE\_Spec}}
\newcommand{\LoneSpec}{\textsc{L1\_Spec}}
\newcommand{\JTFS}{\textsc{JTFS}}

\title{\vspace{-0.5em}Response to Reviewers\\
\large Manuscript: \emph{Evaluating Sound Similarity Metrics for Differentiable, Iterative Sound-Matching}}
\author{}
\date{}

\begin{document}
\maketitle
\vspace{-1.25em}

\noindent
We sincerely thank you and the reviewers for the thoughtful and constructive feedback on our manuscript titled~\emph{``Evaluating Sound Similarity Metrics for Differentiable, Iterative Sound-Matching''}. We have carefully considered all comments and revised the manuscript accordingly. Below, we provide a detailed point-by-point response to each comment. 
For each reviewer, we first provide the review in full, with the key concerns highlighted. Below the full review, we address the concerns in sections. Each section contains the reviewer's key requests or issues, followed by the changes necessary. Some of our revisions are not direct responses to the concrete, point by point suggestions made by the reviewers, but address more general concerns expressed in the body of the reviews (simplification of the introduction and background requested by reviewer 1 for example). As such, for the sake of brevity we took the liberty to paraphrase and aggregate the reviewer's requests and concerns as accurately as possible when necessary.

\medskip
\noindent
We hope the revisions address all concerns and improve the clarity and quality of the manuscript.

\section*{Reviewer 1}

\noindent
The Introduction section lacks a clear articulation of the central research question early on. The core problem that no single loss function performs optimally across synthesizer types is buried within extended historical context and repeated definitions. Moreover, the use of a single subsection label (A. Digital Synthesis with Sound-Matching) without any follow-up sections is structurally inconsistent. \textbf{(R1.1) We recommend tightening the Introduction to focus more directly on problem motivation and contribution, and removing or completing the subsection structure.} \\
\textbf{(R1.2) The current division of the manuscript into Introduction, Background, and Past Works results in significant conceptual overlap and repetition.} Historical framing, loss function taxonomy, and method classification are spread across all three sections. \textbf{A clearer organizational approach would be to merge Background, and Past Works into a unified Background and Related Work section, where formal definitions, theoretical framing, and literature review are presented coherently.} This would allow the Introduction to focus solely on the motivation, problem framing, and contribution. \\
Section IV titled ``Programs, Methodology, and Results'' is overloaded and blends experimental setup with results and interpretation. \textbf{(R1.3) For clear clarity and alignment with structural norms, we recommend splitting this into three sections: (1) Experimental Setup (including synthesizer programs, loss definitions, optimization procedure); (2) Experimental Results (performance metrics, statistical tests, and rankings); and (3) Discussion (qualitative interpretation, synthesis-specific observations, and practical implications).} \\
Conceptually, the work serves primarily as a benchmarking study rather than a novel methodological contribution. While the comparison is comprehensive, the novel losses (DTW\_Envelope, SIMSE\_Spec) are adaptations of known techniques and are not framed with sufficient theoretical justification. \textbf{(R1.4) There is also limited interpretability analysis; for example, the paper does not explore why DTW performs well on modulated noise or why SIMSE outperforms L1 in certain scenarios.} Additionally, the synthesizers used are very simple (only 2--3 parameters), and all experiments are conducted in-domain, limiting the generalizability of the findings to real-world synthesis environments with more complex parameter spaces or out-of-domain targets. \\
\textbf{(R1.5) The manuscript would also benefit from clearer practical implications.} While the authors argue that sound-matching is a creative process and reject the idea of a universal loss, they stop short of offering actionable heuristics. \textbf{Some guidance (e.g., Use DTW for AM synthesis; SIMSE for subtractive synthesis) would strengthen the paper’s relevance to practitioners.} \\
\textbf{(R1.1) The paper would be improved by formally stating additional hypotheses that are currently only implicit.} For example, the observed agreement between manual and automatic evaluation metrics, the effectiveness of the novel losses, and the usability of iterative optimization as a general strategy are all important insights, but are not positioned as testable hypotheses upfront. \\
\textbf{Suggested Actions for Revision:}
\begin{enumerate}
  \item \textbf{(R1.1) Clearly state additional hypotheses regarding metric agreement, perceptual evaluation, and loss novelty.}
  \item \textbf{(R1.4) Deepen the technical and perceptual justification for DTW\_Envelope and SIMSE\_Spec.}
  \item \textbf{(R1.5) Add qualitative or spectral analyses of where specific losses succeed or fail.}
  \item \textbf{(R1.6) Expand discussion to address applicability to more complex or real-world synthesizers.}
  \item \textbf{(R1.7) Provide practical takeaways or usage guidelines for different loss-synth combinations.}
  \item \textbf{(R1.3) Restructure the manuscript to clarify the boundary between methodology, results, and discussion.}
\end{enumerate}

\subsection*{R1.1: Introduction Clarity}
\label{R1.1}
\noindent\textbf{Reviewer:}\\
\textbf{We recommend tightening the Introduction to focus more directly on problem motivation and contribution, and removing or completing the subsection structure.
\\}

\noindent\textbf{Response:} \\
We agree with the reviewer’s comment regarding lack of clarity and focus in the Introduction. The Introduction has been revised accordingly: 
\begin{itemize}
  \item The Introduction is now a single continuous section, without subsection headings.  
  \item We have removed unnecessary historical context and repeated definitions, reducing the length by roughly 25\%. We moved some of this material to the new \emph{Background and Related Work} section.  
  \item The central hypothesis now appears at the beginning of the third paragraph (previously it appeared mid-way through paragraph 10).  
  \item In addition to the main hypothesis, we have added three explicit research questions, covering metric agreement (Q1), the utility of novel DTW and SIMSE losses (Q2), and the viability of iterative optimization (Q3).  These are stated directly after the central hypothesis:

  \begin{displayquote}
     Beyond this central hypothesis, we also investigate three additional research questions: (Q1) To what extent do automatic evaluation metrics agree with manual listening tests? (Q2) Can loss functions based on Dynamic Time Warping (DTW) and Scale-Invariant Mean Squared Error (SIMSE) provide advantages over SOTA loss functions, and under what conditions?
    (Q3) Is iterative differentiable optimization a viable strategy for design of sound-matching experiments?
\end{displayquote}

  \item The Introduction concludes with a concise statement of our contributions:

  \begin{displayquote}
      The contributions of this paper include (1) an evaluation of multiple differentiable losses across multiple synthesis methods, (2) the introduction and justification of loss functions utilizing DTW and SIMSE (3) a discussion on the utility and agreement between manual and automatic evaluation metrics (4) a nomenclature of different sound-matching approaches and unsolved issues in the field. 
  \end{displayquote}
\end{itemize}

\subsection{R1.2 Improve Background and Related Work}
\label{R1.2}
\noindent\textbf{Reviewer:}\\
\noindent\textbf{The current division of the manuscript into Introduction, Background, and Past Works results in significant conceptual overlap and repetition. Historical framing, loss function taxonomy, and method classification are spread across all three sections. A clearer organizational approach would be to merge Background, and Past Works into a unified Background and Related Work section, where formal definitions, theoretical framing, and literature review are presented coherently. This would allow the Introduction to focus solely on the motivation, problem framing, and contribution.}\\

\noindent\textbf{Response:} \\
We agree that the organization and writing in these sections could be improved. We merged \emph{Background} and \emph{Past Works} into a unified \emph{Background and Related Work} section. Major applied major edits and rewrites to these sections to improve reading and remove redundancies between the various subsections while taking precautions in compromising the flow.
The \emph{background and Related Work} section contains the following topics in order:  
\begin{itemize}
  \item A formal definition of sound-matching and its components
  \item Definitions of synthesizers, losses, and optimization heuristics,  
  \item A chronological review of prior sound-matching approaches
  \item A discussion of gaps in the field and four critical issues.
  \item A justification for our methodology, and how our experiments address two of the critical issues (Loss and Synthesizer selection).
\end{itemize}

\subsection*{R1.3 Clarify the boundary between methodology, results, and discussion.}
\label{R1.3}
\noindent\textbf{Reviewer:}\\
\textbf{Section IV titled ``Programs, Methodology, and Results'' is overloaded and blends experimental setup with results and interpretation. For clear clarity and alignment with structural norms, we recommend splitting this into three sections: (1) Experimental Setup (including synthesizer programs, loss definitions, optimization procedure); (2) Experimental Results (performance metrics, statistical tests, and rankings); and (3) Discussion (qualitative interpretation, synthesis-specific observations, and practical implications).}
\\

\noindent\textbf{Response:} \\
We agree that the earlier draft blurred the boundaries between setup, results, and discussion. In the revised manuscript we implemented the reviewer’s suggestion as follows:

\begin{itemize}
  \item \textbf{III. Experimental Setup} (Sections III-A–E) now exclusively describes the methodological components: the synthesizer programs, the definitions of losses, the optimization procedure, evaluation metrics, and statistical tests. Each subsection is focused on describing how the experiments are designed, without interpretation of outcomes. For instance, the synthesizer section defines parameter ranges, while the evaluation section specifies how statistical significance is tested.  (he only partial overlap is at the beginning of the \textbf{III-F Best Performers for Each Program}, which transitions between results and discussion.

  \item Experimental findings are presented in \textbf{III-F Best Performers for Each Program} and \textbf{III-G Consistency in Rankings}, which together form the \emph{Results} component. Section III-F reports which losses performed best for each synthesizer, while Section III-G quantifies the consistency of rankings across bootstrap resamples. Keeping all numerical findings here avoids overlap with later interpretation.

  \item \textbf{IV. Discussion} then provides the qualitative interpretation, structured into four dedicated subsections:  
    \begin{enumerate}
      \item \textbf{IV-A. Key Findings}: summarizes the central empirical trends, such as the dependency of loss performance on synthesis method.  
      \item \textbf{IV-B. Practical Recommendations} : translates results into concrete recommendations for practitioners and researchers.  
      \item \textbf{IV-C. Illustrating Loss Landscapes} : adds visualizations that explain why certain losses succeed or fail in specific contexts.  
      \item \textbf{IV-D. Applicability to More Complex Synthesizers} : discussion on how observed patterns extend beyond the simplified testbed to richer, real-world synthesizers.  
    \end{enumerate}

  \item To improve clarity, we also relocated explanatory material about experiment design (e.g., parameter settings, optimizer configuration) previously scattered across the results section into Sections III-A–E, so that the results sections are strictly empirical while the discussion is interpretive.  
\end{itemize}

Overall, this restructuring produces a clearer separation of topics: methodology is isolated in Section III-A–E, empirical findings are concentrated in Section III-F–G, and broader interpretation is reserved for Section IV. We believe this major restructuring and rewrites of appropriate sections has improved flow of the manuscript and made the boundary between methodology, results, and discussion much clearer.

\subsection*{R1.4 Justification for \DTWEnv{} and \SIMSESpec}
\label{R1.4}
\noindent\textbf{Reviewer:} \\
\noindent While the comparison is comprehensive, the novel losses (DTW\_Envelope, SIMSE\_Spec) are adaptations of known techniques and are not framed with sufficient theoretical justification.

\noindent\textbf{Response:} \\
We have expanded the theoretical justifications for DTW\_Envelope and \SIMSESpec{} in subsection \textbf{II.C Sound Representation and Loss Functions}. In \textbf{II.C.2 Fourier Spectrograms}, we theorize the possible application of SIMSE as an alternative to L1:
\begin{displayquote}
    A common limitation of spectrogram-based losses is their sensitivity to global gain: two signals with identical spectral shape but different overall amplitude can yield large errors under L1/L2 norms. 
    Here we propose and test Scale-Invariant Mean Squared Error (SIMSE) as an alternative to L1/L2 norms in spectrogram comparisons. SIMSE normalizes the amplitudes before comparison and emphasizes proportional differences in spectral envelopes~\cite{barron2014shapessimse}. 
    Perceptually, listeners are often tolerant of loudness changes while remaining sensitive to timbral shape, therefore this property could be advantageous for subtractive synthesis, where filter cutoffs reshape spectral balance without predictable changes in total energy. Although SIMSE has been applied in other contexts such as audio reconstruction, to our knowledge its use as a differentiable spectrogram loss in iterative sound-matching is novel.
\end{displayquote}

 In \textbf{II.C.4 Dynamic Time Warping} For DTW\_Envelope, we clarify its perceptual motivation: aligning amplitude envelopes is crucial for AM synthesis, where traditional spectrogram L1 distances fail to capture similarity: 
 \begin{displayquote}
     DTW provides robustness to local temporal shifts, making it well suited for comparing modulated signals where the perceptual similarity lies in envelope dynamics rather than precise onset alignment. 
    For example, two tremolo signals with slightly different phase are perceptually similar but would appear distant under spectrogram L1. 
    A possible use case to consider is the application of DTW to amplitude envelopes, which may be able to match two sounds with identical rates of loudness modulation regardless of alignments. 
    To our knowledge, DTW applied in this way has not been used as a loss in sound-matching (whether differentiable and iterative or not).
 \end{displayquote}

 We explicitly frame these losses not as wholly new techniques, but as carefully adapted methods tested in an iterative differentiable setting, a contribution that has not previously been made.

We did not modify the definitions in \textbf{III.A Loss Function Implementation Details}, dedicating this section to only technical explanations.

\subsection*{R1.5 Analyses of where specific losses succeed or fail.}
\label{R1.5}
\noindent\textbf{Reviewer:} \\
\textbf{Add qualitative or spectral analyses of where specific losses succeed or fail}\\

\noindent\textbf{Response:} \\
This is an interesting suggestion. Our original experiments were designed to emphasize quantitative, statistically robust measurements of loss function performance, which is why we initially avoided focusing on individual examples of sounds, spectrograms, or loss landscapes (to reduce the risk of cherry-picking). However, illustrative examples can provide helpful intuition for readers and motivate future research. However, we believe that an analysis of loss landscapes (rather than spectrograms or qualitative tests) is a more useful approach to a discussion on where and \textit{why} losses might succeed or fail.

To this end, we have added a new Section \textbf{IV.C llustrating Loss Landscapes} where qualitative visualizations of loss landscapes  are added. These examples highlight why specific losses may succeed or fail and reinforce the broader importance of loss landscapes in assessing the appropriateness of similarity measures~\cite{vahidi2023mesostructures,turian2020sorry}, supporting our argument that effective navigation of loss landscapes represents an important gap in current knowledge and a promising direction for future work:

\begin{displayquote}
    Figure~\ref{fig:loss_landscape_noisebp} shows the HP-Noise results. \LoneSpec{}, \SIMSESpec{}, and \JTFS{} exhibit clear minima at the correct parameter (red dashed line), whereas \DTWEnv{} remains relatively flat, making gradient-based optimization more difficult.

    Figure~\ref{fig:loss_landscape_ampmod} shows the S-Noise-AM results. Here, \DTWEnv{} produces the smoothest and most informative landscape, while spectrogram losses are minimized near the target but lack consistent correlation with parameter distance. \JTFS{} remains largely flat. This aligns with \DTWEnv's superior performance in amplitude-modulated synthesis observed in our experiments.
\end{displayquote}

\begin{figure*}[ht]
    \centering
    \begin{minipage}[t]{0.48\textwidth}
        \centering
        \includegraphics[width=\linewidth]{images/experiment_plots/comparing_loss_landscapes_normalized.png}
        \caption{Loss landscapes for \BPNoise{} with only a high-pass filter parameter. 
        \LoneSpec{}, \SIMSESpec{}, and \JTFS{} show clear global minima near the correct parameter, while \DTWEnv{} remains flat around the target.}
        \label{fig:loss_landscape_noisebp}
    \end{minipage}%
    \hfill
    \begin{minipage}[t]{0.48\textwidth}
        \centering
        \includegraphics[width=\linewidth]{images/experiment_plots/comparing_lanscapes_1_1d_normalized.png}
        \caption{Loss landscapes for a simplified \AmpMod{} synthesizer with only amplitude modulation parameter. 
        \DTWEnv{} exhibits the smoothest and most informative landscape, explaining its superior performance in amplitude-modulated synthesis.}
        \label{fig:loss_landscape_ampmod}
    \end{minipage}
\end{figure*}

While this may not be the "qualitative" or "spectral" analysis, we believe it is instructive on the mechanisms underlying our quantitative findings and highlight the importance of loss landscape navigation as an open area for future research.

\subsection*{R1.6 Applicability to more complex or real-world synthesizers.}
\label{R1.6}
\noindent\textbf{Review:} \\
\noindent \textbf{Expand the discussion to address applicability to more complex or real-world synthesizers}
\\

\noindent\textbf{Response:} \\
We added a new subsection, \textbf{IV.D Applicability to More Complex Synthesizers}, that explicitly discusses how the insights from our controlled two-parameter DSP-based synthesizers generalize to richer synthesizers with many parameters and complex routing. While acknowledging the limitations of our simplified test bed, we argue that the observed dependencies between synthesis method and loss function remain informative even in more complex settings:

\begin{displayquote}
    The experiments in this study were conducted with differentiable synthesizers chosen to isolate fundamental synthesis principle with 2 parameters each. While this level of simplicity facilitates controlled comparisons, it does not fully capture the richness of real-world synthesizers, which can have hundreds of parameters with various routes. As a result, the extent of the generalizability of our findings does require further research.

    Nonetheless, we believe that the insights here can extend naturally to more complex domains. Depending on the feature of interest, the observed dependence of loss performance on synthesis method is will likely remain stable even with added complexities such as layers of modulation, filtering, or nonlinearity.

    \DTWEnv{} measures very specific features of audio, and its use in the recommended settings---where low frequency amplitude modulation matching is required---would likely yield the desired results regardless of synthesizer complexity. Likewise, the use of \SIMSESpec{} or \LoneSpec{} for matching filter cut-offs is likely to be successful regardless of the underlying sound. 
    \end{displayquote}

\subsection{R1.7 Provide practical takeaways}
\label{R1.7}
\noindent\textbf{Review:} \\
\noindent \textbf{Provide practical takeaways or usage guidelines for different loss-synth combinations. } 
\\

\noindent\textbf{Response} \\
We introduced a new subsection, \textbf{IV.B Practical Recommendations}, which distills our findings into actionable heuristics for practitioners.

\begin{displayquote}
    Based on our findings, several actionable guidelines emerge. These heuristics can guide practitioners in selecting appropriate similarity measures and synthesis methods for future sound-matching experiments.
\begin{itemize}
    \item In lieu of listening tests, MSS or P-Loss are generally useful for large-scale benchmarking, but confirm conclusions with manual listening when accuracy in isolated cases is important. 
    \item For amplitude-modulated synthesis, DTW-based losses (e.g., \DTWEnv{}) are recommended over spectrogram based and JTFS. 
    \item For subtractive/noise-filtered synthesis, spectrogram based losses are most effective. Based on our results, SIMSE appears to be the better method of spectrogram comparisons relative to L1, but the amplitude agnosticism of SIMSE likely plays a role in its poor performance in additive synthesis.  
    \item In our amplitude modulation experiments, \JTFS{} did not outperform other measures, contradicting prior claims of its superiority for mesostructures~\cite{vahidi2023mesostructures}. Thus, our results call into question its effectiveness as a general mesostructure measure
    \item Iterative differentiable optimization via Faust-to-JAX is a viable strategy for defining various differentiable DSP functions, requiring only modest hardware resource and providing a more natural approach to synthesizer definition.
\end{itemize}

\end{displayquote}
\section*{Reviewer 2}

\noindent
The paper considers the evaluation of different audio similarity metrics for sound matching applications.  The authors categorize prior methods largely as direct optimization via gradient-free iterative search or open-loop inference via a supervised deep learning model.  Moreover, different approaches which may or may not accommodate out-of-domain sounds.  Accordingly, the authors propose the combined use of differentiable synthesizers and closed-loop iteration as a sound means of achieving synthesizer sound matching, effectively combining the advantages of otherwise disparate methods (i.e. access to gradients while iteratively improving matching results).  Additionally, within their evaluation of different similarity metrics (parameter losses, spectral distances, etc.), they also introduce a DTW-based similarity metric to partially neutralize the effect of time alignment on similarity scores.  A primary conclusion of the work is that there is no one best sound similarity measure capable of working across all synthesizers.  \\
The subject matter of the work is relevant, the experimental setup is satisfactory, and I do not have reason to doubt the correctness of their analysis.  However, my primary concerns with this work are that:  
\begin{enumerate}
  \item \textbf{(R2.1) It is not clear what the proposed method really is, and the paper reads largely as a readout of evaluations that were conducted by the authors.}  
  \item \textbf{(R2.2) The authors claim that differentiable iterative sound matching is largely unexplored, encouraging development of techniques in this area.}  
    \begin{itemize}
      \item \textbf{Barkan et. al, ``Inversynth II: Sound Matching via Self-Supervised Synthesizer-Proxy and Inference-Time Finetuning'' explored precisely this, training an initial ``one-shot'' supervised model, and then performing inference-time optimization, updating model weights to improve the matching against test audio.  Moreover, their use of a synthesizer-proxy negates the explicit need of implementing the synthesizer math differentiably, while still falling under the category of differentiable iterative sound matching.}  
      \item \textbf{Cherep et. al, ``Creative Text-to-Audio Generation via Synthesizer Programming'' uses a differentiable synthesizer and the CLAP score as a similarity metric for iterative optimization.  While the authors lean into synthesizer sound matching against a text query, the method would be readily applicable to audio-to-audio sound matching using the CLAP audio tower alone.}  
    \end{itemize}
  \item \textbf{(R2.3) This work addresses the problems on ``Loss Selection'' and ``Synthesis Selection.''  However, the results as presented here are largely inconclusive (i.e. it is not imminently clear how to select losses/synthesizer pairings), leaving very little in the way of clear takeaways sparking future research.}
\end{enumerate}

\subsection{R2.1 Clarity of proposed method}
\label{R2.1}
\noindent \textbf{It is not clear what the proposed method really is, and the paper reads largely as a readout of evaluations. } 
\\

\noindent\textbf{Response:} \\
We have taken several measures to clarify the methodology and ensure that the paper is not read ``merely as a set of experiments''.

\medskip
\noindent In \textbf{II.B Digital Signal Processing and Synthesis}, we state the importance of isolated experiments:
\begin{displayquote}

Other synthesis methods studied in \textit{isolation} include additive and subtractive synthesis~\cite{engel2020ddsp,masuda2023improving,salimi2020make} and physical modeling~\cite{riionheimo2003parameter,han2024learning}. By \textit{isolation}, we mean settings where the effect of individual parameters on the output sound remains tractable. For example, we exclude studies using commercial Virtual Studio Technology (VST) which can obscure the interactions between modules, losses, and outputs. 
\end{displayquote}

\noindent To further clarify the methodology and its importance, we have added a new subsection, \textbf{II.H What Approach Should We Take?} placed immediately before the Methodology section. This subsection clarifies our experimental design and explains why our contribution is unique. Specifically, we use:  
\begin{enumerate}
  \item Simple differentiable DSP synthesizers to isolate synthesis--loss interactions,  
  \item Multiple losses tested side by side within the same framework, and  
  \item Direct gradient descent optimization without neural proxies.  
\end{enumerate}

This quote from the aforementioned section in the paper makes clear that our work is not simply a report of evaluations, but a controlled benchmarking study designed to reveal how losses and synthesis methods interact:

\begin{displayquote}
    We argue that isolated benchmarking is a necessary precursor to building robust, generalizable sound-matching systems. To address the problems of \LossSelect{} and \SynthSelect{}, we adopt a methodology with three key characteristics that have not previously appeared together.

\begin{enumerate}
    \item Use simple differentiable synthesizers built from common DSP functions (oscillators, filters, envelopes) to address the \SynthSelect{} problem. Restricting to small-scale, low-parameter models avoids the confounding effects of neural proxies or embedding models, enabling direct analysis of how synthesis method interacts with loss functions. 
    \item Evaluate experiments under multiple loss functions to address the \LossSelect{} problem. Prior work has typically tested losses in isolation, making direct comparisons rare \textbf{(((CITE)))}.  
    \item Optimize synthesizer parameters directly with gradient descent, without neural approximations. This complements prior proxy-based methods by providing the missing “low-level experiments” that clarify fundamental interactions between DSP functions and similarity measures. 
\end{enumerate}

Together, these design choices supply the controlled experimental evidence missing from the literature, 
clarifying how losses and synthesis methods interact in differentiable, iterative settings, 
and offering insights likely to generalize to more complex domains.

\end{displayquote}

\medskip
\noindent We restate our methodology in the opening paragraph of the \textbf{III. Experimental Setup} section:

\begin{displayquote}
    At a glance, our methodology is to conduct controlled, low-level differentiable experiments—without proxy networks—by pairing four DSP-based synthesizers with four differentiable loss functions. We optimize parameters directly via gradients in order to isolate loss–synthesizer interactions. This controlled setup systematically evaluates the performance of iterative sound-matching pipelines, where “performance” is defined as the similarity of the synthesized output to a target. Similarity is assessed with P-Loss, MSS, and—most importantly—manual listening tests. This design allows us to directly test our central hypothesis: that the effectiveness of a loss function depends on the synthesis method, and that no universal “best” similarity measure exists.
\end{displayquote}

\subsection*{R2.2 Relation to prior work}
\noindent \textbf{The authors claim that differentiable iterative sound matching is largely unexplored, but prior work (Barkan et al., Cherep et al.) could be considered in this category.}\\

\noindent\textbf{Response:} \\
We thank the reviewer for pointing out sound-matching works which fall under the ``differentiable and iterative'' description that were not covered in our literature review. We believe that the term ``under-explored'' is perhaps more appropriate. We've added these works to \textbf{Table 1} which provides a summary of important previous works, and discussed these works in more detail in the \textbf{II.F Historical Framing of Sound-Matching} section: 
\begin{itemize}
  \item \textbf{Barkan et al. (Inversynth II):} Their work relies on neural proxies for the synthesizer, with inference-time finetuning of the encoder. This constitutes iterative optimization, but the optimization occurs on an approximate model rather than directly on differentiable DSP functions. By contrast, our approach avoids proxy error and directly analyzes how losses behave relative to known synthesis structures.  
  \item \textbf{Cherep et al.:} Their work uses CLAP embeddings with evolutionary search for text-to-audio tasks. While applicable to audio-to-audio matching, CLAP acts as a black-box representation and does not isolate loss--synth interactions. Our approach is complementary, focusing on controlled experiments with simple differentiable DSP synthesizers to expose the strengths and weaknesses of different similarity measures.  In addition, embedding models such as CLAP use simpler representations at their core, commonly the Fourier transformations, and our work explores the low level functions that are often used for the training of embedding models.
\end{itemize}
Looking at these works, we see that our contribution is distinct: we provide the missing \textbf{low-level benchmarking} of loss functions across multiple synthesis methods, which is not addressed by proxy-based or embedding-based approaches.

\subsection*{R2.3 Clear takeaways}
\noindent\textbf{Review:}
\textbf{This work addresses the problems on ``Loss Selection'' and ``Synthesis Selection.''  However, the results as presented here are largely inconclusive (i.e. it is not imminently clear how to select losses/synthesizer pairings), leaving very little in the way of clear takeaways sparking future research.\\}

\noindent\textbf{Response:} \\
In the revised paper, we've added two sections which address the valid issues raised here and by Reviewer 1:
\begin{itemize}
  \item We introduced a new subsection, \emph{Practical Recommendations}, which distills our findings into actionable heuristics for practitioners. DTWEnv is recommended for amplitude-modulated synthesis, SIMSE for subtractive/noise-filtered synthesis, and MSS or P-Loss as large-scale proxies for sound-matching evaluation (with manual validation being strongly advised). We also note that our results challenge prior claims about JTFS’s universal superiority for mesostructures.
  \item We added a new subsection, \emph{Applicability to More Complex Synthesizers}, that explicitly discusses how the insights from our controlled two-parameter DSP-based synthesizers generalize to richer synthesizers with many parameters and complex routing. While acknowledging the limitations of our simplified testbed, we argue that the observed dependencies between synthesis method and loss function remain informative even in more complex settings.
\end{itemize}
Please see our response to the first Reviwer in Sections~\ref{R1.7}
\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}

### 2025-09-25 sections/background_related.tex
Updates from Overleaf

@@
 For example, two tremolo signals with slightly different phase are perceptually similar but would appear distant under spectrogram L1. 
-By applying DTW specifically to amplitude envelopes, we can match two sounds with identical rates ofloudness modulation regardless of alignments. 
+A possible use case to consider is the application of DTW to amplitude envelopes, which may be able to match two sounds with identical rates of loudness modulation regardless of alignments. 
 To our knowledge, DTW applied in this way has not been used as a loss in sound-matching (whether differentiable and iterative or not).

### 2025-09-26 review_responses.tex
Updates from Overleaf

@@
 
-\title{\vspace{-0.5em}Response to Reviewers\\
+\title{\vspace{-0.5em}Author Response for T-ASL-11868-2025\\
 \large Manuscript: \emph{Evaluating Sound Similarity Metrics for Differentiable, Iterative Sound-Matching}}
@@
 
+\noindent Dear Editor Alberto Bernardini, \\
 \noindent
-We sincerely thank you and the reviewers for the thoughtful and constructive feedback on our manuscript titled~\emph{``Evaluating Sound Similarity Metrics for Differentiable, Iterative Sound-Matching''}.
+We'd like to thank you and the reviewers for the thoughtful and constructive feedback on our manuscript.
 We have carefully considered all comments and revised the manuscript accordingly.
-Below, we provide a detailed point-by-point response to each comment.
-For each reviewer, we first provide the review in full, with the key concerns highlighted.
-Below the full review, we address the concerns in sections.
-Each section contains the reviewer's key requests or issues, followed by the changes necessary.
-Some of our revisions are not direct responses to the concrete, point by point suggestions made by the reviewers, but address more general concerns expressed in the body of the reviews (simplification of the introduction and background requested by reviewer 1 for example).
-As such, for the sake of brevity we took the liberty to paraphrase and aggregate the reviewer's requests and concerns as accurately as possible when necessary.
+Here you will find our detailed point-by-point response to the feedback.
+
+For each reviewer, we first provide the review in full, with the key concerns numbered and in bold.
+Below the full review, we address the concerns in dedicated sections.
+Each section contains the reviewer's key requests or issues, followed by our response.
+Some revisions address more general concerns expressed in the body of the reviews (\nameref{R1.1} and \nameref{R1.2} for example).
+As such, for the sake of brevity we took the liberty to paraphrase the reviewer's requests and concerns as accurately as possible when necessary.
 
@@
 
-\section*{Reviewer 1}
+\noindent Sincerely, Amir Salimi, Abram Hindle, Osmar R. Za{\"i}ane
 
+\section{Reviewer 1}
+\subsection{Original Review}
 \noindent
@@
 \end{displayquote}
-\section*{Reviewer 2}
-
+\section{Reviewer 2}
+\subsection{Original Review}
 \noindent
@@
 \label{R2.1}
+\noindent\textbf{Review:}
 \noindent \textbf{It is not clear what the proposed method really is, and the paper reads largely as a readout of evaluations. } 
@@
     \item Evaluate experiments under multiple loss functions to address the \LossSelect{} problem.
-Prior work has typically tested losses in isolation, making direct comparisons rare \textbf{(((CITE)))}.
+Prior work has typically tested losses in isolation, making direct comparisons rare\cite{vahidi2023mesostructures}.
     \item Optimize synthesizer parameters directly with gradient descent, without neural approximations.
@@
 \noindent\textbf{Response:} \\
-In the revised paper, we've added two sections which address the valid issues raised here and by Reviewer 1:
+In the revised paper, we've added two sections which address the valid issues raised here and by Reviewer 1.
+
 \begin{itemize}
+  \item We added a new subsection, \emph{Applicability to More Complex Synthesizers}, that explicitly discusses how the insights from our controlled two-parameter DSP-based synthesizers generalize to richer synthesizers with many parameters and complex routing.
+While acknowledging the limitations of our simplified testbed, we argue that the observed dependencies between synthesis method and loss function remain informative even in more complex settings.
   \item We introduced a new subsection, \emph{Practical Recommendations}, which distills our findings into actionable heuristics for practitioners.
@@
 We also note that our results challenge prior claims about JTFS’s universal superiority for mesostructures.
-  \item We added a new subsection, \emph{Applicability to More Complex Synthesizers}, that explicitly discusses how the insights from our controlled two-parameter DSP-based synthesizers generalize to richer synthesizers with many parameters and complex routing.
-While acknowledging the limitations of our simplified testbed, we argue that the observed dependencies between synthesis method and loss function remain informative even in more complex settings.
+
 \end{itemize}
-Please see our response to the first Reviwer in Sections~\ref{R1.7}
+
+For a more detailed response, see our response to the Reviewer.1 in Sections~\nameref{R1.6} and \nameref{R1.7}.
+
 \bibliographystyle{IEEEtran}

### 2025-09-26 review_responses.tex
Updates from Overleaf

@@
 As such, for the sake of brevity we took the liberty to paraphrase the reviewer's requests and concerns as accurately as possible when necessary.
+We hope these revisions have addressed your concerns and improved the clarity and quality of the manuscript.
 
 \medskip
-\noindent
-We hope the revisions address all concerns and improve the clarity and quality of the manuscript.
-
 \noindent Sincerely, Amir Salimi, Abram Hindle, Osmar R. Za{\"i}ane
@@
   \item We have removed unnecessary historical context and repeated definitions, reducing the length by roughly 25\%.
-We moved some of this material to the new \emph{Background and Related Work} section.
+We moved some of this material to the new \textbf{II.
+Background and Related Work} section.
   \item The central hypothesis now appears at the beginning of the third paragraph (previously it appeared mid-way through paragraph 10).  
-  \item In addition to the main hypothesis, we have added three explicit research questions, covering metric agreement (Q1), the utility of novel DTW and SIMSE losses (Q2), and the viability of iterative optimization (Q3).
+  \item In addition to the main hypothesis, we have added three explicit research questions  (answered in \textbf{IV.B Practical Recommendations}), covering metric agreement (Q1), the utility of DTW and SIMSE in implementing novel losses (Q2), and the viability of iterative optimization (Q3).
 These are stated directly after the central hypothesis:
@@
 We agree that the organization and writing in these sections could be improved.
-We merged \emph{Background} and \emph{Past Works} into a unified \emph{Background and Related Work} section.
-Major applied major edits and rewrites to these sections to improve reading and remove redundancies between the various subsections while taking precautions in compromising the flow.
-The \emph{background and Related Work} section contains the following topics in order:  
+We merged \emph{Background} and \emph{Past Works} into a unified section, \textbf{II.
+Background and Related Work}.
+We applied major edits and rewrites to these sections to improve reading and remove redundancies between the various subsections while taking precautions in compromising the flow.
+\textbf{II. Background and Related Work} contains the following topics in order:  
 \begin{itemize}
-  \item A formal definition of sound-matching and its components
-  \item Definitions of synthesizers, losses, and optimization heuristics,  
-  \item A chronological review of prior sound-matching approaches
+  \item A formal definition of sound-matching and its components.
+  \item Definitions of synthesizers, losses, and optimization heuristics. 
+  \item A chronological review of prior sound-matching approaches.
   \item A discussion of gaps in the field and four critical issues.
-  \item A justification for our methodology, and how our experiments address two of the critical issues (Loss and Synthesizer selection).
+  \item A justification for our methodology, and how our experiments address the two critical issues of \emph{Loss and Synthesizer Selection}.
 \end{itemize}
 
-\subsection*{R1.3 Clarify the boundary between methodology, results, and discussion.}
+\subsection*{R1.3 Organize Methodology, Results, and Discussion.}
 \label{R1.3}
@@
   \item \textbf{III.
-Experimental Setup} (Sections III-A–E) now exclusively describes the methodological components: the synthesizer programs, the definitions of losses, the optimization procedure, evaluation metrics, and statistical tests.
+Experimental Setup and Results A-E}: Sections III.A–E now exclusively describe the methodological components: the synthesizer programs, the definitions of losses, the optimization procedure, evaluation metrics, and statistical tests.
 Each subsection is focused on describing how the experiments are designed, without interpretation of outcomes.
 For instance, the synthesizer section defines parameter ranges, while the evaluation section specifies how statistical significance is tested.
-(he only partial overlap is at the beginning of the \textbf{III-F Best Performers for Each Program}, which transitions between results and discussion.
+The only partial overlap is at the beginning of the \textbf{III-F Best Performers for Each Program}, which serves as a transition between results and discussion.
 
@@
       \item \textbf{IV-A. Key Findings}: summarizes the central empirical trends, such as the dependency of loss performance on synthesis method.  
-      \item \textbf{IV-B. Practical Recommendations} : translates results into concrete recommendations for practitioners and researchers.  
-      \item \textbf{IV-C. Illustrating Loss Landscapes} : adds visualizations that explain why certain losses succeed or fail in specific contexts.  
-      \item \textbf{IV-D. Applicability to More Complex Synthesizers} : discussion on how observed patterns extend beyond the simplified testbed to richer, real-world synthesizers.  
+      \item \textbf{IV-B. Practical Recommendations}: translates results into concrete recommendations for practitioners and researchers.  
+      \item \textbf{IV-C. Illustrating Loss Landscapes}: adds visualizations that explain why certain losses succeed or fail in specific contexts.  
+      \item \textbf{IV-D. Applicability to More Complex Synthesizers}: discussion on how observed patterns extend beyond the simplified testbed to richer, real-world synthesizers.  
     \end{enumerate}
@@
 We have expanded the theoretical justifications for DTW\_Envelope and \SIMSESpec{} in subsection \textbf{II.C Sound Representation and Loss Functions}.
-In \textbf{II.C.2 Fourier Spectrograms}, we theorize the possible application of SIMSE as an alternative to L1:
+In \textbf{II.C.2 Fourier Spectrograms}, we discuss the possible application of SIMSE as an alternative to L1:
 \begin{displayquote}
@@
 
- In \textbf{II.C.4 Dynamic Time Warping} For DTW\_Envelope, we clarify its perceptual motivation: aligning amplitude envelopes is crucial for AM synthesis, where traditional spectrogram L1 distances fail to capture similarity:
+ In \textbf{II.C.4 Dynamic Time Warping} For DTW\_Envelope, we clarify its perceptual motivation: aligning amplitude envelopes is crucial for AM synthesis, where traditional spectrogram distances fail to capture similarity:
  \begin{displayquote}
@@
     For example, two tremolo signals with slightly different phase are perceptually similar but would appear distant under spectrogram L1. 
-    A possible use case to consider is the application of DTW to amplitude envelopes, which may be able to match two sounds with identical rates of loudness modulation regardless of alignments. 
+    A possible use case to consider is the application of DTW to amplitude envelopes, which may be able to match two sounds with similar rates of loudness modulation regardless of alignments. 
     To our knowledge, DTW applied in this way has not been used as a loss in sound-matching (whether differentiable and iterative or not).
@@
 
- We explicitly frame these losses not as wholly new techniques, but as carefully adapted methods tested in an iterative differentiable setting, a contribution that has not previously been made.
+ We explicitly frame these losses not as wholly new techniques, but as carefully adapted methods to be tested in an iterative differentiable setting, a contribution that has not previously been made.
 
@@
 
-\subsection*{R1.5 Analyses of where specific losses succeed or fail.}
+\subsection*{R1.5 Analysis of Failure or Success.}
 \label{R1.5}
@@
 This is an interesting suggestion.
-Our original experiments were designed to emphasize quantitative, statistically robust measurements of loss function performance, which is why we initially avoided focusing on individual examples of sounds, spectrograms, or loss landscapes (to reduce the risk of cherry-picking).
-However, illustrative examples can provide helpful intuition for readers and motivate future research.
-However, we believe that an analysis of loss landscapes (rather than spectrograms or qualitative tests) is a more useful approach to a discussion on where and \textit{why} losses might succeed or fail.
+Our original experiments were designed to emphasize quantitative, statistically robust measurements of loss function performance, which is why we initially avoided focusing on individual examples of sounds, spectrograms, or loss landscapes (in part to reduce the risk of cherry-picking).
+We do concede that illustrative examples can provide helpful intuition for readers and motivate future research, however, we believe that an analysis of loss landscapes (rather than spectrograms or qualitative tests) is a more useful approach to an explanation for \textit{where} and \textit{why} losses might succeed or fail.
 
 To this end, we have added a new Section \textbf{IV.C llustrating Loss Landscapes} where qualitative visualizations of loss landscapes  are added.
-These examples highlight why specific losses may succeed or fail and reinforce the broader importance of loss landscapes in assessing the appropriateness of similarity measures~\cite{vahidi2023mesostructures,turian2020sorry}, supporting our argument that effective navigation of loss landscapes represents an important gap in current knowledge and a promising direction for future work:
+These examples highlight why specific losses may succeed or fail and reinforce the broader importance of loss landscapes in assessing the appropriateness of similarity measures~\cite{vahidi2023mesostructures,turian2020sorry}, supporting our argument that effective navigation of loss landscapes represents an important gap in current knowledge and a possible direction for future work:
 
 \begin{displayquote}
-    Figure~\ref{fig:loss_landscape_noisebp} shows the HP-Noise results.
+We show the landscapes of the loss functions with simplified versions of \BPNoise{} (HP-Noise) and \AmpMod{} (S-Noise-AM), each reduced to a single parameter for one-dimensional visualization.
+HP-Noise uses only a high-pass filter with cutoff between 100–20,000 Hz, while S-Noise-AM varies only the modulation rate between 0.1–20 Hz.
+Loss values are normalized to 0–1 to allow comparison across functions.
+
+    Figure~\ref{fig:loss_landscape_noisebp} [\textbf{Figure 4 in paper}] shows the HP-Noise results.
 \LoneSpec{}, \SIMSESpec{}, and \JTFS{} exhibit clear minima at the correct parameter (red dashed line), whereas \DTWEnv{} remains relatively flat, making gradient-based optimization more difficult.
 
-    Figure~\ref{fig:loss_landscape_ampmod} shows the S-Noise-AM results.
+    Figure~\ref{fig:loss_landscape_ampmod} [\textbf{Figure 5 in paper}]  shows the S-Noise-AM results.
 Here, \DTWEnv{} produces the smoothest and most informative landscape, while spectrogram losses are minimized near the target but lack consistent correlation with parameter distance.
@@
 
-\subsection*{R1.6 Applicability to more complex or real-world synthesizers.}
+\subsection*{R1.6 Applicability to More Complex or Real-world Synthesizers.}
 \label{R1.6}
@@
 
-\subsection{R1.7 Provide practical takeaways}
+\subsection{R1.7 Provide Practical Takeaways}
 \label{R1.7}
@@
 \noindent\textbf{Response} \\
-We introduced a new subsection, \textbf{IV.B Practical Recommendations}, which distills our findings into actionable heuristics for practitioners.
+
+In \textbf{IV.A Key Findings}, we discuss important findings and provide answers to our research questions.
+The following section, \textbf{IV.B Practical Recommendations}, distills our findings into actionable heuristics for practitioners.
 
 \begin{displayquote}
-    Based on our findings, several actionable guidelines emerge.
-These heuristics can guide practitioners in selecting appropriate similarity measures and synthesis methods for future sound-matching experiments.
+    Our findings not only provide answers to the research questions proposed in the introduction, but also serve as guidelines for practitioners in selecting appropriate similarity measures and synthesis methods for future sound-matching experiments.
 \begin{itemize}
-    \item In lieu of listening tests, MSS or P-Loss are generally useful for large-scale benchmarking, but confirm conclusions with manual listening when accuracy in isolated cases is important. 
-    \item For amplitude-modulated synthesis, DTW-based losses (e.g., \DTWEnv{}) are recommended over spectrogram based and JTFS. 
+    \item In lieu of listening tests, MSS or P-Loss are generally useful for large-scale benchmarking.
+However, we encourage confirmation of conclusions with manual listening, particularly if specificity is important.
+    \item For amplitude-modulated synthesis, DTW-based losses such as \DTWEnv{} are recommended over spectrogram differences and JTFS.
     \item  For subtractive/noise-filtered synthesis, spectrogram based losses are most effective.
 Based on our results, SIMSE appears to be the better method of spectrogram comparisons relative to L1, but the amplitude agnosticism of SIMSE likely plays a role in its poor performance in additive synthesis.
+    \item Iterative differentiable optimization via the Faust-to-JAX pipeline is a viable strategy for defining various differentiable DSP functions, requiring only modest hardware resource and providing a more natural approach to synthesizer definition.
     \item In our amplitude modulation experiments, \JTFS{} did not outperform other measures, contradicting prior claims of its superiority for mesostructures~\cite{vahidi2023mesostructures}.
-Thus, our results call into question its effectiveness as a general mesostructure measure
-    \item Iterative differentiable optimization via Faust-to-JAX is a viable strategy for defining various differentiable DSP functions, requiring only modest hardware resource and providing a more natural approach to synthesizer definition.
+Thus, our results call into question its effectiveness as a general mesostructure measure, and highlight a need for further research into its utility.
 \end{itemize}
@@
 
-\subsection{R2.1 Clarity of proposed method}
+\subsection{R2.1 Clarity of Proposed Method}
 \label{R2.1}
@@
 \noindent\textbf{Response:} \\
-We have taken several measures to clarify the methodology and ensure that the paper is not read ``merely as a set of experiments''.
+We have taken several measures to emphasize the importance and reasoning behind our methodology design and ensure that the paper is not read as a set of evaluation results.
 
 \medskip
-\noindent In \textbf{II.B Digital Signal Processing and Synthesis}, we state the importance of isolated experiments:
+\noindent In \textbf{II.B Digital Signal Processing and Synthesis}, we state the importance of isolated sound-matching experiments for analysis of loss-synthesizer interactions:
 \begin{displayquote}
@@
     \item Use simple differentiable synthesizers built from common DSP functions (oscillators, filters, envelopes) to address the \SynthSelect{} problem.
-Restricting to small-scale, low-parameter models avoids the confounding effects of neural proxies or embedding models, enabling direct analysis of how synthesis method interacts with loss functions.
+Restriction to small-scale, low-parameter models avoids the confounding effects of neural proxies or embedding models, enabling direct analysis of how synthesis method interacts with loss functions.
     \item Evaluate experiments under multiple loss functions to address the \LossSelect{} problem.
-Prior work has typically tested losses in isolation, making direct comparisons rare\cite{vahidi2023mesostructures}.
+Prior work has typically tested losses in isolation, making direct comparisons of loss functions rare~\cite{vahidi2023mesostructures,han2023perceptual,uzrad2024diffmoog}, particularly if comparisons are done in the context of different methods of synthesis.
     \item Optimize synthesizer parameters directly with gradient descent, without neural approximations.
-This complements prior proxy-based methods by providing the missing “low-level experiments” that clarify fundamental interactions between DSP functions and similarity measures.
+This complements prior proxy-based methods by providing the missing “low-level experiments” that aim to analyze fundamental interactions between DSP functions and similarity measures.
 \end{enumerate}
@@
 
-\subsection*{R2.2 Relation to prior work}
+\subsection*{R2.2 Relation to Prior Work}
 \noindent \textbf{The authors claim that differentiable iterative sound matching is largely unexplored, but prior work (Barkan et al., Cherep et al.) could be considered in this category.}\\
@@
 \noindent\textbf{Response:} \\
-We thank the reviewer for pointing out sound-matching works which fall under the ``differentiable and iterative'' description that were not covered in our literature review.
+We thank the reviewer for pointing out sound-matching works that match the ``differentiable and iterative'' description, but were not covered in our literature review.
 We believe that the term ``under-explored'' is perhaps more appropriate.
-We've added these works to \textbf{Table 1} which provides a summary of important previous works, and discussed these works in more detail in the \textbf{II.F Historical Framing of Sound-Matching} section:
+We've added these works to \textbf{Table 1} which provides a summary of notable previous work, and discussed them in more detail in the final paragraphs of \textbf{II.F Historical Framing of Sound-Matching}:
 \begin{itemize}
-  \item \textbf{Barkan et al.
-(Inversynth II):} Their work relies on neural proxies for the synthesizer, with inference-time finetuning of the encoder.
+  \item \textbf{Barkan et al.:} Their work relies on neural proxies for the synthesizer, with inference-time finetuning of the encoder.
 This constitutes iterative optimization, but the optimization occurs on an approximate model rather than directly on differentiable DSP functions.
@@
 Our approach is complementary, focusing on controlled experiments with simple differentiable DSP synthesizers to expose the strengths and weaknesses of different similarity measures.
-In addition, embedding models such as CLAP use simpler representations at their core, commonly the Fourier transformations, and our work explores the low level functions that are often used for the training of embedding models.
+In addition, embedding models such as CLAP use simpler representations at their core---commonly Fourier transformations---and our work explores the low level functions that are often used for the training of embedding models.
 \end{itemize}
@@
 
-\subsection*{R2.3 Clear takeaways}
+\subsection*{R2.3 Clear Takeaways}
 \noindent\textbf{Review:}
@@
 \noindent\textbf{Response:} \\
-In the revised paper, we've added two sections which address the valid issues raised here and by Reviewer 1.
+
+We acknowledge the reviewer’s concern that our main result may appear inconclusive.
+However, we see this as an important finding in itself: much of the prior literature implicitly over-claims general effectiveness of particular loss functions based on narrow or in-domain settings.
+Our revisions now explicitly emphasize that the lack of a “silver bullet” for loss functions is a core contribution of this work.
+By showing that loss effectiveness depends on the synthesis context, the inconclusiveness can be seen as evidence against over-generalization, and an argument for more thorough, context-sensitive benchmarks in measuring sound similarity.\\
+
+As for clear takeaways, we've added two new sections which address the valid issues raised here and by Reviewer 1.
 
 \begin{itemize}
-  \item We added a new subsection, \emph{Applicability to More Complex Synthesizers}, that explicitly discusses how the insights from our controlled two-parameter DSP-based synthesizers generalize to richer synthesizers with many parameters and complex routing.
-While acknowledging the limitations of our simplified testbed, we argue that the observed dependencies between synthesis method and loss function remain informative even in more complex settings.
-  \item We introduced a new subsection, \emph{Practical Recommendations}, which distills our findings into actionable heuristics for practitioners.
+  \item \textbf{IV.B Practical Recommendations} which distills our findings into actionable guidlines for practitioners.
 DTWEnv is recommended for amplitude-modulated synthesis, SIMSE for subtractive/noise-filtered synthesis, and MSS or P-Loss as large-scale proxies for sound-matching evaluation (with manual validation being strongly advised).
 We also note that our results challenge prior claims about JTFS’s universal superiority for mesostructures.
+  \item \textbf{IV.D Applicability to More Complex Synthesizers} that explicitly discusses how the insights from our controlled two-parameter DSP-based synthesizers generalize to richer synthesizers with many parameters and complex routing.
+While acknowledging the limitations of our simplified testbed, we argue that the observed dependencies between synthesis method and loss function remain informative even in more complex settings.
+\end{itemize}
+For a more detailed response, see our response to the Reviewer.1 in Sections~\nameref{R1.6} and \nameref{R1.7}. Our response to \nameref{R1.5} may also be of interest.
 
-\end{itemize}
+\section{Reviewer 4}
+\subsection{Use of Appropriate Symbols}
+\noindent\textbf{Review:}
+The notation \textbf{[for the equation introduced in II.A]} is confusing. In the manuscript, $t$ is used to denote the target sound, but in audio/signal processing $t$ usually represents time.  
+A clearer alternative would be to write $x_\theta$ and $x_0$, so that $x$ always denotes a sound waveform, with the subscript indicating whether it is generated from parameters $\theta$ or given as the target.
+This avoids overloading $t$ and improves readability.
 
-For a more detailed response, see our response to the Reviewer.1 in Sections~\nameref{R1.6} and \nameref{R1.7}.
+\noindent\textbf{Response:}
+This is a helpful suggestion, and we have made the switch from $x$ and $t$ to $x_{\theta}$ and $x_0$.
+
+\section{Optimal Targets}
+\noindent\textbf{Review:}
+You mentioned "There is no optimal or correct version of an 8-bit snare sound".
+This statement is true when you are given only the text description of the sound.
+However, if you have the target sound $x_0$  the target sound itself is "correct", isn't it?
+So your hypothesis only holds if you don't have the target sound but the text description, am I understanding it correctly?
+
+\noindent\textbf{Response:}
+
+\subsection{STFT Losses and Time-Shifts}
+\noindent\textbf{Review:}
+"If one of these tones is slightly shifted in time, then the spectrogram will no longer overlap". This is true if we don't apply any techniques like DTW to realign them, right?
+So the STFT losses doesn't work for the cases where there's time shift?
+
+\noindent\textbf{Response:}
+
+\subsection{Can We Overfit?}
+\noindent\textbf{Review:}
+One final question that is just out of curiosity.
+In your iterative algorithm, you always run it for 200 steps, right?
+In my field of machine learning, if we set 200 steps, the result in the final step might not be the optimal one due to overfitting.
+So sometimes we obtain the optimal results at around 100 steps, and the model gets worse after this.
+Do you have the same thing in your algorithm? Or the final step is always the best step?
+
+\noindent\textbf{Response}
+
+\subsection{Minor Edits}
+\noindent\textbf{Review}
+Isn't it better to move the symbol explanation earlier to where the first equation is? I could only under the equations after reading section C.
+
+\noindent\textbf{Response:}
+We have moved symbol explanations right after the introduction in 
+\textbf{II. A Formalization of Sound-Matching}.
+
+\noindent\textbf{Review:}
+"out-of-domain sounds is an unexplored area".
+Unexplored means no one has done it before to me.
+But in Table 1, you have already cited a few work that tried to do it.
+So, should it be "underexplored" instead of "unexplored"?
+
+\noindent\textbf{Response:}
+
+\noindent\textbf{Review:}
+You mentioned "four different synthesizers" here, but you never elaborate on which four, not until page 8. Probably you can specify the four synthesizers earlier?
+
+\noindent\textbf{Review:}
+You described the Likert scale from 1 to 5 here.
+But in the next paragraph, you describe the Likert scores again.
+Not sure if I misunderstood anything, but It looks a bit redundant to me.
+Would it be possible to combine both to reduce redundancy?
+
+\noindent\textbf{Review:}
+Just a minor grammar question, Is there a missing subject in the first sentence?While [we] cannot prove ...
+
+\noindent\textbf{Review:}
 

### 2025-09-26 sections/background_related.tex
Updates from Overleaf

--- 2025-09-24 sections/background_related.tex
@@
     \item $g(\theta)$: A parametric audio synthesizer $g$ with parameters $\theta$.
-    \item $x$: The synthesizer output, $x = g(\theta)$.
-    \item $t$: The target sound to be replicated or imitated. 
+    \item $x_\theta$: The synthesizer output, $x_\theta = g(\theta)$.”.
+    \item $\xzero$: The target sound to be replicated or imitated. 
     \item $\phi(\cdot)$: A representation (feature extraction) function mapping signals into a comparison space.
-    \item $L$: A loss function that measures distance between $x$ and $t$, typically via $L(\theta,t) = d(\phi(x), \phi(t))$ for some metric $d$.
+    \item $L$: A loss function that measures distance between $\xtheta$ and $\xzero$, typically via $L(\theta,\xzero) = d(\phi(\xtheta), \phi(\xzero))$ for some metric $d$.
 \end{itemize}
@@
 % Nodes
-    \node (start) [text centered] {\( g (\hat{\theta}) = x \)};
-    \node (L) [above of=start,right of=start, text centered] {\( L(\hat{\theta},t) {\ \propto \ } \phi(x) - \phi(t) \)};
-    \node (optimize) [below of=L,right of=L, text centered] {Optimize $\hat{\theta}$};
-    \node (new_theta) [below of=optimize,right of = start, text centered] {New \( \hat{\theta} \)};
+\node (start) [text centered] {\( g(\hat{\theta}) = x_{\theta} \)};
+\node (L) [above of=start, right of=start, text centered] 
+    {\( L(\hat{\theta}, x_{0}) \propto \phi(x_{\theta}) - \phi(x_{0}) \)};
+\node (optimize) [below of=L, right of=L, text centered] {Optimize $\hat{\theta}$};
+\node (new_theta) [below of=optimize, right of=start, text centered] {New \( \hat{\theta} \)};
 
@@
 \[
-L(\theta, t) = d\langle\phi(t),\phi(x)\rangle
+L(\theta, \xzero) = d(\phi(\xtheta), \phi(\xzero))
 \]
@@
 
-A proxy representation is the output of the function \( \phi \), which can be thought of as a feature extraction function that maps the sounds \( t \) and \( x \) to their respective representations. 
+A proxy representation is the output of the function \( \phi \), which can be thought of as a feature extraction function that maps the sounds  $\xzero$ and $\xtheta$  to their respective representations.
 The proportionality or distance metric $d$ has typically been the L1 or L2 distance~\cite{turian2020sorry,richard2025model}, with L1 being calculated as the mean of the absolute difference between every point in the proxy representation:
 \[
-L(\theta, t) = \left\| \phi(t) - \phi(x) \right\|_1
+L(\theta, \xzero) = \left\| \phi(\xzero) - \phi(\xtheta) \right\|_1
 \]
@@
 For example, two tremolo signals with slightly different phase are perceptually similar but would appear distant under spectrogram L1. 
-By applying DTW specifically to amplitude envelopes, we can match two sounds with identical rates ofloudness modulation regardless of alignments. 
+A possible use case to consider is the application of DTW to amplitude envelopes, which may be able to match two sounds with similar rates of loudness modulation regardless of alignments. 
 To our knowledge, DTW applied in this way has not been used as a loss in sound-matching (whether differentiable and iterative or not).
@@
 The choice of domain depends on whether we want to use the same synthesizer for the target and output sounds, the scenario that is called \textit{in-domain}, or have target sounds that came from sources other than the synthesizer, or \textit{out-of-domain}.
-To paraphrase the description given by Masuda \textit{et al.}~\cite{masuda2021soundmatch}, if $g$, the synthesizer of choice, can accurately replicate the target sound $t$, or put differently, if $t$ itself is an output of $g$, then the sound-matching task is \textit{in-domain}.
-If $t$ is not an output of $g$, then the sound-matching task is \textit{out-of-domain}.
+To paraphrase the description given by Masuda \textit{et al.}~\cite{masuda2021soundmatch}, if $g$, the synthesizer of choice, can accurately replicate the target sound $\xzero$, or put differently, if $\xzero$ itself is an output of $g$, then the sound-matching task is \textit{in-domain}.
+If $\xzero$ is not an output of $g$, then the sound-matching task is \textit{out-of-domain}.
 In-domain tasks in general are simpler, and often there is a guarantee that there is a correct answer to the sound-matching problem, particularly if the goal is accurate replication of the sound.
@@
 
-\subsection{Supvervised vs Direct Optimization}
+\subsection{Supvervised versus Direct Optimization}
 \label{sec:optimization}
@@
 \[
-\theta^* = \arg\min_{\theta} L(\theta,t)
+\theta^* = \arg\min_{\theta} L(\theta,\xzero)
 \]
@@
 A Riemannian metric M needs to be calculated for the minimization of locally linear approximation of the ``true" spectral loss function~\cite{han2023perceptual}.
+
 \[
-\|\phi(t) - \phi(x)\|_2^2 = \langle \tilde{\theta} - \theta | \mathbf{M}(\theta) | \tilde{\theta} - \theta \rangle + O(\|\tilde{\theta} - \theta\|_2^3). \tag{4}
+\| \varphi(\xzero) - \varphi(\xtheta) \|_2^2
+= \langle \tilde{\theta} - \theta \,|\, M(\theta) \,|\, \tilde{\theta} - \theta \rangle
++ O(\|\tilde{\theta} - \theta\|_3^2).
 \]
@@
 A neural approximation for another part of the supervised sound-matching chain---this time the synthesizer---was proposed by Barkan \textit{et al.}~\cite{barkan2023inversynthII}.
-As they noted, without a differentiable synthesizer, ``model-based'' (or what we call supervised) approaches cannot directly compare $x_o$ to $x_t$, often opting for P-Loss, which may not correctly map the parameters of sound to the output audio~\cite{esling2019flow,han2023perceptual,masuda2023improving}.
+As they noted, without a differentiable synthesizer, ``model-based'' (or what we call supervised) approaches cannot directly compare $\xtheta$ to $\xzero$, often opting for P-Loss, which may not correctly map the parameters of sound to the output audio~\cite{esling2019flow,han2023perceptual,masuda2023improving}.
 This approach used a model mapping sounds to parameters, another approximating the synthesizer, and a loss function which combines P-Loss with STFT differences.
@@
     \item Use simple differentiable synthesizers built from common DSP functions (oscillators, filters, envelopes) to address the \SynthSelect{} problem.
-Restricting to small-scale, low-parameter models avoids the confounding effects of neural proxies or embedding models, enabling direct analysis of how synthesis method interacts with loss functions.
-    \item Evaluate experiments under multiple loss functions to address the \LossSelect{} problem. Prior work has typically tested losses in isolation, making direct comparisons rare~\todo{cite}.  
+Restriction to small-scale, low-parameter models avoids the confounding effects of neural proxies or embedding models, enabling direct analysis of how synthesis method interacts with loss functions.
+    \item Evaluate experiments under multiple loss functions to address the \LossSelect{} problem.
+Prior work has typically tested losses in isolation, making direct comparisons of loss functions difficult~\cite{vahidi2023mesostructures,han2023perceptual,uzrad2024diffmoog}.
+Particularly rare are comparisons which are done in the context of different methods of synthesis.
     \item Optimize synthesizer parameters directly with gradient descent, without neural approximations.

### 2025-09-26 sections/summary_weakness_conclusion.tex
Updates from Overleaf

@@
 
-\subsection{Practical Recommendations:}
-Based on our findings, several actionable guidelines emerge.
-These heuristics can guide practitioners in selecting appropriate similarity measures and synthesis methods for future sound-matching experiments.
+\subsection{Practical Recommendations}
+    Our findings not only provide answers to the research questions proposed in the introduction, but also serve as guidelines for practitioners in selecting appropriate similarity measures and synthesis methods for future sound-matching experiments.
 \begin{itemize}
-    \item In lieu of listening tests, MSS or P-Loss are generally useful for large-scale benchmarking, but confirm conclusions with manual listening when accuracy in isolated cases is important. 
-    \item For amplitude-modulated synthesis, DTW-based losses (e.g., \DTWEnv{}) are recommended over spectrogram based and JTFS. 
+    \item In lieu of listening tests, MSS or P-Loss are generally useful for large-scale benchmarking.
+However, we encourage confirmation of conclusions with manual listening, particularly if specificity is important.
+    \item For amplitude-modulated synthesis, DTW-based losses such as \DTWEnv{} are recommended over spectrogram differences and JTFS.
     \item  For subtractive/noise-filtered synthesis, spectrogram based losses are most effective.
 Based on our results, SIMSE appears to be the better method of spectrogram comparisons relative to L1, but the amplitude agnosticism of SIMSE likely plays a role in its poor performance in additive synthesis.
+    \item Iterative differentiable optimization via the Faust-to-JAX pipeline is a viable strategy for defining various differentiable DSP functions, requiring only modest hardware resource and providing a more natural approach to synthesizer definition.
     \item In our amplitude modulation experiments, \JTFS{} did not outperform other measures, contradicting prior claims of its superiority for mesostructures~\cite{vahidi2023mesostructures}.
-Thus, our results call into question its effectiveness as a general mesostructure measure
-    \item Iterative differentiable optimization via Faust-to-JAX is a viable strategy for defining various differentiable DSP functions, requiring only modest hardware resource and providing a more natural approach to synthesizer definition.
+Thus, our results call into question its effectiveness as a general mesostructure measure, and highlight a need for further research into its utility.
 \end{itemize}
@@
 We show the landscapes of the loss functions with simplified versions of \BPNoise{} (HP-Noise) and \AmpMod{} (S-Noise-AM), each reduced to a single parameter for one-dimensional visualization.
-HP-Noise uses only a high-pass filter with cutoff between 100–20,000 Hz, while S-Noise-AM varies only the modulation rate between 0.1–20 Hz.
+HP-Noise uses only a high-pass filter with cut-off between 100–20,000 Hz, while S-Noise-AM varies only the modulation rate between 0.1–20 Hz.
 Loss values are normalized to 0–1 to allow comparison across functions.

### 2025-09-26 review_responses.tex
Updates from Overleaf

@@
 
-\subsection*{R2.2 Relation to Prior Work}
+\subsection{R2.2 Relation to Prior Work}
+\label{R2.2}
 \noindent \textbf{The authors claim that differentiable iterative sound matching is largely unexplored, but prior work (Barkan et al., Cherep et al.) could be considered in this category.}\\
@@
 \section{Reviewer 4}
-\subsection{Use of Appropriate Symbols}
+Here we've addressed the four bigger concerns first, leaving the minor edits for \nameref{r4.minor}.
+
+\subsection{Original Review}
+Well written paper with comprehensive literature review and extensive experiments to bridge existing research gap.
+Although this paper does not have any ground-breaking new methods that could achieve SOTA performance, the experimental results could still serve as a useful reference for researchers who want to try this field.
+So I believe it should be accepted with some minor modifications.
+
+(\textbf{R4.1}) Page 3, Left Column, the first equation.
+I am not sure if `t` is a common symbol to represent `target` in sound marking, but my field of audio/signal processing, `t` usually represents `time`.
+So it is a bit confusing to me when reading it for the first time.
+Would something like $x_\theta$ and $x_t$ be better?
+So x always represent the sound.
+And the subscript indicated if the sound is from the output given the parameter theta, or the target sound.
+Even better, probably we can use $x_0$ instead of $x_t$ to represent the target sound?
+
+Page 3 Right column section C.
+Isn't it better to move the symbol explanation earlier to where the first equation is? I could only under the equations after reading section C.
+
+(\textbf{R4.2}) Page 3 Right column at the end
+You mentioned "There is no optimal or correct version of an 8-bit snare sound".
+This statement is true when you are given only the text description of the sound.
+However, if you have the target sound $x_0$ (or t in your current draft), the target sound itself is "correct", isn't it?
+So your hypothesis only holds if you don't have the target sound but the text description, am I understanding it correctly?
+
+Page 6 Right column Section C Point 4.
+"out-of-domain sounds is an unexplored area".
+Unexplored means no one has done it before to me.
+But in Table 1, you have already cited a few work that tried to do it.
+So, should it be "underexplored" instead of "unexplored"?
+
+Page 6 Right column Section IV A
+You mentioned "four different synthesizers" here, but you never elaborate on which four, not until page 8. Probably you can specify the four synthesizers earlier?
+
+(\textbf{R4.3}) Also, in the blinded hearing test, the two listeners are the authors, it seems a bit to biased to me.
+Not sure how difficult is it to add one more listener, but I believe having an extra sound engineer/musician who is not one of the author would yield a more convincing listening results.
+
+Page 7 Left Column 1st Paragraph
+You described the Likert scale from 1 to 5 here.
+But in the next paragraph, you describe the Likert scores again.
+Not sure if I misunderstood anything, but It looks a bit redundant to me.
+Would it be possible to combine both to reduce redundancy?
+
+(\textbf{R4.4}) Page 7 Right Column section B
+"If one of these tones is slightly shifted in time, then the spectrogram will no longer overlap". This is true if we don't apply any techniques like DTW to realign them, right?
+So the STFT losses doesn't work for the cases where there's time shift?
+
+Page 11, the "weakness" section under Summary
+Just a minor grammar question, Is there a missing subject in the first sentence?
+While [we] cannot prove ...
+
+(\textbf{R4.5}) One final question that is just out of curiosity.
+In your iterative algorithm, you always run it for 200 steps, right?
+In my field of machine learning, if we set 200 steps, the result in the final step might not be the optimal one due to overfitting.
+So sometimes we obtain the optimal results at around 100 steps, and the model gets worse after this.
+Do you have the same thing in your algorithm? Or the final step is always the best step?
+
+\subsection{\textbf{R4.1} Use of Appropriate Symbols}
 \noindent\textbf{Review:}
@@
 This avoids overloading $t$ and improves readability.
+\\
 
 \noindent\textbf{Response:}
-This is a helpful suggestion, and we have made the switch from $x$ and $t$ to $x_{\theta}$ and $x_0$.
+This is a helpful suggestion, and we have made the switch from $x$ and $t$ to $x_{\theta}$ and $x_0$ throughout the manuscript.
 
-\section{Optimal Targets}
+\subsection{\textbf{R4.2} Optimal Targets}
 \noindent\textbf{Review:}
@@
 However, if you have the target sound $x_0$  the target sound itself is "correct", isn't it?
-So your hypothesis only holds if you don't have the target sound but the text description, am I understanding it correctly?
+So your hypothesis only holds if you don't have the target sound but the text description, am I understanding it correctly?\\
 
 \noindent\textbf{Response:}
+The original statement referred to how realistically in sound-design, a "correct" solution rarely exists due to our subjective experience of sound.
+To reduce confusion, the section has been rephrased as follows:
+\begin{displayquote}
+    Moreover, the ``correct'' solution may not exist in a strict sense: depending on the artistic goal, multiple parameter settings may yield acceptable or even preferable outputs.
+In the following sections we will discuss the various components of sound-matching, the subjectivity of ``correct'' solutions, and important issues in the field.
+\end{displayquote}
 
-\subsection{STFT Losses and Time-Shifts}
+\subsection{Hearing Tests}
+\noindent{\textbf{Review:}}
+Also, in the blinded hearing test, the two listeners are the authors, it seems a bit to biased to me.
+Not sure how difficult is it to add one more listener, but I believe having an extra sound engineer/musician who is not one of the author would yield a more convincing listening results.
+\\
+
+\noindent{\textbf{Response:}}
+
+\subsection{\textbf{R4.4} STFT Losses and Time-Shifts}
 \noindent\textbf{Review:}
@@
 \noindent\textbf{Response:}
+What we refer to as STFT losses refers to the near ubiquitous approach of applying L1/L2 to STFT spectrograms.
+This approach does not account for two sounds that are nearly identical but slightly time-shifted~\cite{vahidi2023mesostructures,turian2020sorry}.
 
-\subsection{Can We Overfit?}
+\subsection{\textbf{R4.5} Can We Overfit? Is it Bad}
 \noindent\textbf{Review:}
@@
 \noindent\textbf{Response} 
+Since we are not training a model that will be applied to other scenarios, overfitting here just means that the parameter updates are stuck around a local minimum.
+This means that \textit{if} the target parameters have been reached, subsequent updates will not diverge from it (or at least stay very close).
+We can share our visualization of the gradient update paths in 2D, which we felt was not statistically robust enough for the paper.
+Figures~\ref{fig:p0_simse_iterations} and~\ref{fig:p2_dtw_iterations}, we see that once the gradient updates (small red dots connected by lines) take the parameters from the starting position (green circle) to the target (green star), subsequent updates move sligtly and randomly around the target.
 
-\subsection{Minor Edits}
+\begin{figure*}[h]
+    \centering
+    \begin{minipage}[t]{0.48\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{images/experiment_plots/p0_SIMSE_Spec.png}
+        \caption{SIMSE\_Spec reaching target and not diverging in \BPNoise}
+        \label{fig:p0_simse_iterations}
+    \end{minipage}%
+    \hfill
+    \begin{minipage}[t]{0.48\textwidth}
+        \centering
+        \includegraphics[width=\linewidth]{images/experiment_plots/p2_DTW_Onset.png}
+        \caption{DTW\_Env reaching target and not diverging in \AmpMod}
+        \label{fig:p2_dtw_iterations}
+    \end{minipage}
+\end{figure*}
+
+\subsection{R4 Minor Edits}
+\label{r4.minor}
 \noindent\textbf{Review}
@@
 \textbf{II. A Formalization of Sound-Matching}. 
+\\
 
@@
 \noindent\textbf{Response:}
+We've changed our phrasing where necessary. See also \nameref{R2.2}.
+\\
 
@@
 
-\noindent\textbf{Review:}
-You described the Likert scale from 1 to 5 here.
-But in the next paragraph, you describe the Likert scores again.
-Not sure if I misunderstood anything, but It looks a bit redundant to me.
-Would it be possible to combine both to reduce redundancy?
+\noindent\textbf{Response:}
+We've modified the beginning of Section \textbf{III. Experimental Setup and Results}
+
+\begin{displayquote}
+    At a glance, our methodology is to conduct controlled, low-level differentiable experiments without the use of proxy networks.
+We pair four differentiable losses with four differentiable DSP-based synthesizers: BP-Noise (band-pass noise with LP/HP filters), Add-SineSaw (additive sine and saw oscillators), Noise-AM (noise modulated by an LFO), and SineSaw-AM (sine-modulated saw oscillator).
+\end{displayquote}
 
 \noindent\textbf{Review:}
-Just a minor grammar question, Is there a missing subject in the first sentence?While [we] cannot prove ...
+You described the Likert scale from 1 to 5 here. But in the next paragraph, you describe the Likert scores again.
+
+\noindent\textbf{Response:}
+This has been fixed in our rewrite of the methodology section (\textbf{III. Experimental Setup and Results}).
+\\
 
 \noindent\textbf{Review:}
+Fix "While [we] cannot prove ..."
+
+\noindent\textbf{Response:}
+Fixed, thank you.
+\\
 

### 2025-09-26 sections/methodology_and_analysis.tex
Updates from Overleaf

@@
 
-\section{Experimental Setup}
+\section{Experimental Setup and Results}
 \label{sec:experiment_setup}
-At a glance, our methodology is to conduct controlled, low-level differentiable experiments—without proxy networks—by pairing four DSP-based synthesizers with four differentiable loss functions.
+At a glance, our methodology is to conduct controlled, low-level differentiable experiments without the use of proxy networks.
+We pair four differentiable losses with four differentiable DSP-based synthesizers: BP-Noise (band-pass noise with LP/HP filters), Add-SineSaw (additive sine and saw oscillators), Noise-AM (noise modulated by an LFO), and SineSaw-AM (sine-modulated saw oscillator).
 We optimize parameters directly via gradients in order to isolate loss–synthesizer interactions.
@@
 (Q3) Is an iterative differentiable optimization a viable strategy for design of sound-matching experiments?
-The experiment results provide answers regarding the main hypothesis of this work as well as the secondary questions. Based on these answers, we can make practical recommendations for future research.
+The experiment results provide answers regarding the main hypothesis of this work as well as the secondary questions.
+In the following sections, we will discuss our key findings, caveats, and make recommendations for future research.
 

### 2025-09-28 review_responses.tex
Updates from Overleaf

@@
 
-\subsection{Hearing Tests}
+\subsection{R4.3 Hearing Tests}
+\label{}
 \noindent{\textbf{Review:}}
-Also, in the blinded hearing test, the two listeners are the authors, it seems a bit to biased to me.
+Also, in the blinded hearing test, the two listeners are the authors, it seems a bit too biased to me.
 Not sure how difficult is it to add one more listener, but I believe having an extra sound engineer/musician who is not one of the author would yield a more convincing listening results.
 \\
-
 \noindent{\textbf{Response:}}
+The listening tests were done without use of any labels or previous exposure to outputs, which we believe minimizes the risk of bias.
+Unfortunately, adding another listener was not possible given the allotted time for the revisions.
+If the reviewer feels that this is necessary, it can be done for the next round of reviews.
 

### 2025-09-28 sections/background_related.tex
Updates from Overleaf

@@
 Recent works have focused on the limitations of parameter and spectral loss functions in sound-matching~\cite{vahidi2023mesostructures,uzrad2024diffmoog}, seeking to create more effective general solutions for the comparison of audio.
-Noting the aforementioned weaknesses of comparing STFT spectrograms\todo{you discussed loudness only and not misalignments}, Vahidi \textit{et al.} proposed differentiable Joint-Time Frequency Scattering (\gls{JTFS})~\cite{anden2015joint} as an alternative to spectrogram loss in sound-matching, and showed improved performance in sound-matching with differentiable chirplet synthesizers~\cite{vahidi2023mesostructures}.
-JTFS is the result of the application of a 2D wavelet transformation to the time-frequency representation of a signal~\cite{anden2015joint} and has been reported as more sensitive to mesostructural features such \todo{list}~\cite{vahidi2023mesostructures}.
+Noting the weaknesses of comparing STFT spectrograms (such as alignment and loudness distances), Vahidi \textit{et al.} proposed differentiable Joint-Time Frequency Scattering (\gls{JTFS})~\cite{anden2015joint} as an alternative to spectrogram loss in sound-matching, and showed improved performance in sound-matching with differentiable chirplet synthesizers~\cite{vahidi2023mesostructures}.
+JTFS is the result of the application of a 2D wavelet transformation to the time-frequency representation of a signal~\cite{anden2015joint} and has been reported as more sensitive to \textit{mesostructural} features such as melody, syncopation, and textural
+contrast~\cite{vahidi2023mesostructures}.
 
@@
 
-\subsection{In-Domain versus Out-of-Domain}
+\subsection{In-Domain Versus Out-of-Domain}
 \label{sec:in-domain}
@@
 
-\subsection{Supvervised versus Direct Optimization}
+\subsection{Supvervised Versus Direct Optimization}
 \label{sec:optimization}

### 2025-09-28 sections/summary_weakness_conclusion.tex
Updates from Overleaf

@@
 However, given the simplicity of our programs and the occasional deviations, we cannot conclude that these automatic measures are substitutes for human hearing tests.
-As many previous works have done\todo{cite}, the use of MSS and/or P-Loss as general loss functions does seem appropriate, however, we advise the use of listening tests or custom made loss functions for more specific analysis.
+As many previous works have done (see Table~\ref{tab:summary}, the use of MSS and/or P-Loss as general loss functions does seem appropriate, however, we advise the use of listening tests or custom made loss functions for more specific analysis.
 

### 2025-09-29 review_responses.tex
Updates from Overleaf

@@
 
-\noindent Dear Editor Alberto Bernardini, \\
+\noindent Dear Editor, Alberto Bernardini, \\
 \noindent
@@
 
-  \item To improve clarity, we also relocated explanatory material about experiment design (e.g., parameter settings, optimizer configuration) previously scattered across the results section into Sections III-A–E, so that the results sections are strictly empirical while the discussion is interpretive.
+  \item To improve clarity, we also relocated explanatory material about experiment design (e.g., parameter settings, optimizer configuration) previously scattered across the results section into Sections III-A–E, ensuring that the results sections remain strictly empirical, while the discussion is reserved for interpretation.
 \end{itemize}
@@
 
-To this end, we have added a new Section \textbf{IV.C llustrating Loss Landscapes} where qualitative visualizations of loss landscapes  are added.
+To this end, we have added a new Section \textbf{IV.C Illustrating Loss Landscapes} where qualitative visualizations of loss landscapes  are added.
 These examples highlight why specific losses may succeed or fail and reinforce the broader importance of loss landscapes in assessing the appropriateness of similarity measures~\cite{vahidi2023mesostructures,turian2020sorry}, supporting our argument that effective navigation of loss landscapes represents an important gap in current knowledge and a possible direction for future work:
@@
     Nonetheless, we believe that the insights here can extend naturally to more complex domains.
-Depending on the feature of interest, the observed dependence of loss performance on synthesis method is will likely remain stable even with added complexities such as layers of modulation, filtering, or nonlinearity.
+Depending on the feature of interest, the observed dependence of loss performance on synthesis method will likely remain stable even with added complexities such as layers of modulation, filtering, or nonlinearity.
 
@@
 \noindent
-The paper considers the evaluation of different audio similarity metrics for sound matching applications.
+The paper considers the evaluation of different audio similarity metrics for sound-matching applications.
 The authors categorize prior methods largely as direct optimization via gradient-free iterative search or open-loop inference via a supervised deep learning model.
 Moreover, different approaches which may or may not accommodate out-of-domain sounds.
-Accordingly, the authors propose the combined use of differentiable synthesizers and closed-loop iteration as a sound means of achieving synthesizer sound matching, effectively combining the advantages of otherwise disparate methods (i.e.
+Accordingly, the authors propose the combined use of differentiable synthesizers and closed-loop iteration as a sound means of achieving synthesizer sound-matching, effectively combining the advantages of otherwise disparate methods (i.e.
 access to gradients while iteratively improving matching results).
@@
   \item \textbf{(R2.1) It is not clear what the proposed method really is, and the paper reads largely as a readout of evaluations that were conducted by the authors.}  
-  \item \textbf{(R2.2) The authors claim that differentiable iterative sound matching is largely unexplored, encouraging development of techniques in this area.}  
+  \item \textbf{(R2.2) The authors claim that differentiable iterative sound-matching is largely unexplored, encouraging development of techniques in this area.}  
     \begin{itemize}
       \item \textbf{Barkan et.
-al, ``Inversynth II: Sound Matching via Self-Supervised Synthesizer-Proxy and Inference-Time Finetuning'' explored precisely this, training an initial ``one-shot'' supervised model, and then performing inference-time optimization, updating model weights to improve the matching against test audio.
-Moreover, their use of a synthesizer-proxy negates the explicit need of implementing the synthesizer math differentiably, while still falling under the category of differentiable iterative sound matching.}
+al, ``Inversynth II: sound-matching via Self-Supervised Synthesizer-Proxy and Inference-Time Finetuning'' explored precisely this, training an initial ``one-shot'' supervised model, and then performing inference-time optimization, updating model weights to improve the matching against test audio.
+Moreover, their use of a synthesizer-proxy negates the explicit need of implementing the synthesizer math differentiably, while still falling under the category of differentiable iterative sound-matching.}
       \item \textbf{Cherep et.
 al, ``Creative Text-to-Audio Generation via Synthesizer Programming'' uses a differentiable synthesizer and the CLAP score as a similarity metric for iterative optimization.
-While the authors lean into synthesizer sound matching against a text query, the method would be readily applicable to audio-to-audio sound matching using the CLAP audio tower alone.}
+While the authors lean into synthesizer sound-matching against a text query, the method would be readily applicable to audio-to-audio sound-matching using the CLAP audio tower alone.}
     \end{itemize}
@@
 \label{R2.1}
-\noindent\textbf{Review:}
+\noindent\textbf{Reviewer:}\\
 \noindent \textbf{It is not clear what the proposed method really is, and the paper reads largely as a readout of evaluations. } 
@@
 \label{R2.2}
-\noindent \textbf{The authors claim that differentiable iterative sound matching is largely unexplored, but prior work (Barkan et al., Cherep et al.) could be considered in this category.}\\
+\noindent\textbf{Reviewer:}\\
+\noindent \textbf{Review: The authors claim that differentiable iterative sound-matching is largely unexplored, but prior work (Barkan et al., Cherep et al.) could be considered in this category.}\\
 
@@
 \subsection*{R2.3 Clear Takeaways}
-\noindent\textbf{Review:}
+\noindent\textbf{Reviewer:}\\
 \textbf{This work addresses the problems on ``Loss Selection'' and ``Synthesis Selection.''  However, the results as presented here are largely inconclusive (i.e.
@@
 \noindent\textbf{Response:} \\
-
 We acknowledge the reviewer’s concern that our main result may appear inconclusive.
@@
 \begin{itemize}
-  \item \textbf{IV.B Practical Recommendations} which distills our findings into actionable guidlines for practitioners.
-DTWEnv is recommended for amplitude-modulated synthesis, SIMSE for subtractive/noise-filtered synthesis, and MSS or P-Loss as large-scale proxies for sound-matching evaluation (with manual validation being strongly advised).
+  \item \textbf{IV.B Practical Recommendations} which distills our findings into actionable guidelines for practitioners.
+DTW_Env is recommended for amplitude-modulated synthesis, SIMSE for subtractive/noise-filtered synthesis, and MSS or P-Loss as large-scale proxies for sound-matching evaluation (with manual validation being strongly advised).
 We also note that our results challenge prior claims about JTFS’s universal superiority for mesostructures.
@@
 (\textbf{R4.1}) Page 3, Left Column, the first equation.
-I am not sure if `t` is a common symbol to represent `target` in sound marking, but my field of audio/signal processing, `t` usually represents `time`.
+\textbf{I am not sure if `t` is a common symbol to represent `target` in sound marking, but my field of audio/signal processing, `t` usually represents `time`.
 So it is a bit confusing to me when reading it for the first time.
-Would something like $x_\theta$ and $x_t$ be better?
-So x always represent the sound.
+Would something like $x_\theta$ and $x_t$ be better?} So x always represent the sound.
 And the subscript indicated if the sound is from the output given the parameter theta, or the target sound.
@@
 
-(\textbf{R4.2}) Page 3 Right column at the end
+(\textbf{R4.2}) \textbf{Page 3 Right column at the end
 You mentioned "There is no optimal or correct version of an 8-bit snare sound".
 This statement is true when you are given only the text description of the sound.
-However, if you have the target sound $x_0$ (or t in your current draft), the target sound itself is "correct", isn't it?
-So your hypothesis only holds if you don't have the target sound but the text description, am I understanding it correctly?
+However, if you have the target sound $x_0$ (or t in your current draft), the target sound itself is "correct", isn't it?} So your hypothesis only holds if you don't have the target sound but the text description, am I understanding it correctly?
 
@@
 
-(\textbf{R4.3}) Also, in the blinded hearing test, the two listeners are the authors, it seems a bit to biased to me.
-Not sure how difficult is it to add one more listener, but I believe having an extra sound engineer/musician who is not one of the author would yield a more convincing listening results.
+(\textbf{R4.3}) \textbf{Also, in the blinded hearing test, the two listeners are the authors, it seems a bit too biased to me.} Not sure how difficult is it to add one more listener, but I believe having an extra sound engineer/musician who is not one of the authors would yield a more convincing listening results.
 
@@
 
-(\textbf{R4.4}) Page 7 Right Column section B
-"If one of these tones is slightly shifted in time, then the spectrogram will no longer overlap". This is true if we don't apply any techniques like DTW to realign them, right?
+(\textbf{R4.4}) \textbf{Page 7 Right Column section B
+"If one of these tones is slightly shifted in time, then the spectrogram will no longer overlap". This is true if we don't apply any techniques like DTW to realign them, right?}
 So the STFT losses doesn't work for the cases where there's time shift?
@@
 
-(\textbf{R4.5}) One final question that is just out of curiosity.
+(\textbf{R4.5}) O\textbf{ne final question that is just out of curiosity.
 In your iterative algorithm, you always run it for 200 steps, right?
@@
 So sometimes we obtain the optimal results at around 100 steps, and the model gets worse after this.
-Do you have the same thing in your algorithm? Or the final step is always the best step?
+Do you have the same thing in your algorithm? Or the final step is always the best step?}
 
 \subsection{\textbf{R4.1} Use of Appropriate Symbols}
-\noindent\textbf{Review:}
-The notation \textbf{[for the equation introduced in II.A]} is confusing. In the manuscript, $t$ is used to denote the target sound, but in audio/signal processing $t$ usually represents time.  
+\noindent\textbf{Reviewer:}\\
+\textbf{The notation \textbf{[for the equation introduced in II.A]} is confusing.
+In the manuscript, $t$ is used to denote the target sound, but in audio/signal processing $t$ usually represents time.
 A clearer alternative would be to write $x_\theta$ and $x_0$, so that $x$ always denotes a sound waveform, with the subscript indicating whether it is generated from parameters $\theta$ or given as the target.
-This avoids overloading $t$ and improves readability.
+This avoids overloading $t$ and improves readability.}
 \\
@@
 \subsection{\textbf{R4.2} Optimal Targets}
-\noindent\textbf{Review:}
-You mentioned "There is no optimal or correct version of an 8-bit snare sound".
+\noindent\textbf{Reviewer:}\\
+\textbf{You mentioned "There is no optimal or correct version of an 8-bit snare sound".
 This statement is true when you are given only the text description of the sound.
 However, if you have the target sound $x_0$  the target sound itself is "correct", isn't it?
-So your hypothesis only holds if you don't have the target sound but the text description, am I understanding it correctly?\\
+So your hypothesis only holds if you don't have the target sound but the text description, am I understanding it correctly?\\}
 
@@
 \label{}
-\noindent{\textbf{Review:}}
-Also, in the blinded hearing test, the two listeners are the authors, it seems a bit too biased to me.
+\noindent\textbf{Reviewer:}\\
+\textbf{Also, in the blinded hearing test, the two listeners are the authors, it seems a bit too biased to me.
 Not sure how difficult is it to add one more listener, but I believe having an extra sound engineer/musician who is not one of the author would yield a more convincing listening results.
+}
 \\
+
 \noindent{\textbf{Response:}}
@@
 \subsection{\textbf{R4.4} STFT Losses and Time-Shifts}
-\noindent\textbf{Review:}
-"If one of these tones is slightly shifted in time, then the spectrogram will no longer overlap". This is true if we don't apply any techniques like DTW to realign them, right?
-So the STFT losses doesn't work for the cases where there's time shift?
+\noindent\textbf{Reviewer:}\\
+\textbf{"If one of these tones is slightly shifted in time, then the spectrogram will no longer overlap". This is true if we don't apply any techniques like DTW to realign them, right?
+So the STFT losses doesn't work for the cases where there's time shift?}
+\\
 
@@
 
-\subsection{\textbf{R4.5} Can We Overfit? Is it Bad}
-\noindent\textbf{Review:}
-One final question that is just out of curiosity.
+\subsection{\textbf{R4.5} Can We Overfit? Is it a Concern?}
+\noindent\textbf{Reviewer:}\\
+\textbf{One final question that is just out of curiosity.
 In your iterative algorithm, you always run it for 200 steps, right?
@@
 Do you have the same thing in your algorithm? Or the final step is always the best step?
+}
+\\
 
-\noindent\textbf{Response} 
-Since we are not training a model that will be applied to other scenarios, overfitting here just means that the parameter updates are stuck around a local minimum.
+\noindent\textbf{Response:} 
+Since we are not training a model that will be applied to other scenarios, overfitting here just means that the parameter updates are stuck around a local or global minimum.
 This means that \textit{if} the target parameters have been reached, subsequent updates will not diverge from it (or at least stay very close).
 We can share our visualization of the gradient update paths in 2D, which we felt was not statistically robust enough for the paper.
-Figures~\ref{fig:p0_simse_iterations} and~\ref{fig:p2_dtw_iterations}, we see that once the gradient updates (small red dots connected by lines) take the parameters from the starting position (green circle) to the target (green star), subsequent updates move sligtly and randomly around the target.
+Figures~\ref{fig:p0_simse_iterations} and~\ref{fig:p2_dtw_iterations}, we see that once the gradient updates (small red dots connected by lines) take the parameters from the starting position (green circle) to the target (green star), subsequent updates fluctuate slightly and randomly around the target.
 
@@
 \label{r4.minor}
-\noindent\textbf{Review}
+\noindent\textbf{Reviewer:}\\
 Isn't it better to move the symbol explanation earlier to where the first equation is? I could only under the equations after reading section C.
@@
 
-\noindent\textbf{Review:}
+\noindent\textbf{Reviewer:}\\
 "out-of-domain sounds is an unexplored area".
@@
 
-\noindent\textbf{Review:}
+\noindent\textbf{Reviewer:}\\
 You mentioned "four different synthesizers" here, but you never elaborate on which four, not until page 8. Probably you can specify the four synthesizers earlier?
@@
 
-\noindent\textbf{Review:}
+\noindent\textbf{Reviewer:}\\
 You described the Likert scale from 1 to 5 here. But in the next paragraph, you describe the Likert scores again.
@@
 
-\noindent\textbf{Review:}
+\noindent\textbf{Reviewer:}\\
 Fix "While [we] cannot prove ..."

### 2025-09-29 main.tex
Updates from Overleaf

@@
 
-\markboth{%
-  \parbox{\textwidth}{%
-    This work has been submitted to the IEEE for possible publication. \\
-    Copyright may be transferred without notice, after which this version may no longer be accessible.%
-  }%
-}{}
+% \markboth{%
+%   \parbox{\textwidth}{%
+%     This work has been submitted to the IEEE for possible publication. \\
+%     Copyright may be transferred without notice, after which this version may no longer be accessible.%
+%   }%
+% }{}
 
@@
 
-\todo{A relatively small subset of these synthesis methods have been used for automatic sound-matching, notably additive and/or subtractive synthesizers~\cite{esling2019flow,yee2018automatic,mitchell2007evolutionary,salimi2020make}, physical modeling~\cite{riionheimo2003parameter}, and frequency/amplitude modulation (\gls{FM}/\gls{AM})~\cite{horner1993machine,vahidi2023mesostructures}.
-}
+A relatively small subset of these synthesis methods have been used for automatic sound-matching, notably additive and/or subtractive synthesizers~\cite{esling2019flow,yee2018automatic,mitchell2007evolutionary,salimi2020make}, physical modeling~\cite{riionheimo2003parameter}, and frequency/amplitude modulation (\gls{FM}/\gls{AM})~\cite{horner1993machine,vahidi2023mesostructures}.
 
@@
 
-\todo{
 Genetic algorithms (\gls{GA})~\cite{holland1992genetic} can be used for direct sound optimization, and have been the earliest and most common heuristic in sound-matching~\cite{horner1993machine,mitchell2007evolutionary,yee2018automatic}.
 These algorithms start with arbitrary parameter sets that can be treated as an evolving population where the genomes are the parameter values.
-The most fit members of the group are the parameters which perform the best in the loss function, and create a new generation of parameters via mutation (random change in subset of parameters) and crossovers (combination of parameter sets); this process repeats until the goal or a maximum number of generations is reached.}
+The most fit members of the group are the parameters which perform the best in the loss function, and create a new generation of parameters via mutation (random change in subset of parameters) and crossovers (combination of parameter sets); this process repeats until the goal or a maximum number of generations is reached.
 

### 2025-09-29 sections/background_related.tex
Updates from Overleaf

@@
 
-\subsection{Supvervised Versus Direct Optimization}
+\subsection{Supervised Versus Direct Optimization}
 \label{sec:optimization}
@@
 In differentiable settings, this can cause gradient descent updates to not reach the global minima.
-    \item \OutDomain: Sound-matching with out-of-domain sounds is a largely unexplored area, yet a necessary one for practical applications for sound designers.
+    \item \OutDomain: Sound-matching with out-of-domain sounds is an under-explored area, yet a necessary one for practical applications for sound designers.
 \end{enumerate}

### 2025-09-29 sections/introduction.tex
Updates from Overleaf

@@
 A digital audio synthesizer is any software used for the creation and manipulation of audio.
-The possibilities afforded by these synthesizers is infinite, leading to their widespread adaption by artists and sound-designers~\cite{lyons1997understanding,russ1999sound,stranneby2004digital}.
+The possibilities afforded by these synthesizers are infinite, leading to their widespread adoption by artists and sound-designers~\cite{lyons1997understanding,russ1999sound,stranneby2004digital}.
 A typical synthesizer has a number of parameterizable functions which affect the sound output in various ways, and manual sound-design often involves modifying the parameters until a conceptualized target sound is reached.
@@
 The main hypothesis here is that \textit{the performance of a similarity measure (or loss function) is influenced by other factors in the environment, particularly the method of synthesis}.
-While there have been many works and experiments comparing the accuracy of loss functions~\cite{vahidi2023mesostructures,turian2020sorry,engel2020ddsp,uzrad2024diffmoog,han2023perceptual,masuda2021soundmatch,turian2020sorry,bruford2024synthesizer}, we note that claims regarding the effectiveness of one function versus another have consistently been made in limited contexts that may not generalize to other settings.
+Testing this hypothesis requires a variety of sound-matching experiments that measure whether a single loss function proves most effective.
+While there have been many works and experiments comparing the accuracy of loss functions~\cite{vahidi2023mesostructures,turian2020sorry,engel2020ddsp,uzrad2024diffmoog,han2023perceptual,masuda2021soundmatch,turian2020sorry,bruford2024synthesizer}, we note that claims regarding the effectiveness of one function versus another have often been made in limited contexts that may not generalize to other settings.
 Beyond this central hypothesis, we also investigate three additional research questions: (Q1) To what extent do automatic evaluation metrics agree with manual listening tests?
@@
 We evaluate the final similarity with two different automatic methods, as well as manual listening tests conducted by two of the authors.
-We apply statistical ranking applied to the evaluation scores to compare outcomes across synthesizer–loss combinations and measure the similarity of different evaluation measures.
+We apply statistical ranking to the evaluation scores to compare outcomes across synthesizer–loss combinations and measure the similarity of different evaluation measures.
 

### 2025-09-29 sections/summary_weakness_conclusion.tex
Updates from Overleaf

@@
 Based on our results, SIMSE appears to be the better method of spectrogram comparisons relative to L1, but the amplitude agnosticism of SIMSE likely plays a role in its poor performance in additive synthesis.
-    \item Iterative differentiable optimization via the Faust-to-JAX pipeline is a viable strategy for defining various differentiable DSP functions, requiring only modest hardware resource and providing a more natural approach to synthesizer definition.
+    \item Iterative differentiable optimization via the Faust-to-JAX pipeline is a viable strategy for defining various differentiable DSP functions, requiring only modest hardware resources and providing a more natural approach to synthesizer definition.
     \item In our amplitude modulation experiments, \JTFS{} did not outperform other measures, contradicting prior claims of its superiority for mesostructures~\cite{vahidi2023mesostructures}.
@@
 \subsection{Applicability to More Complex Synthesizers}
-The experiments in this study were conducted with differentiable synthesizers chosen to isolate fundamental synthesis principle with 2 parameters each.
+The experiments in this study were conducted with differentiable synthesizers chosen to isolate fundamental synthesis principles, each with two parameters.
 While this level of simplicity facilitates controlled comparisons, it does not fully capture the richness of real-world synthesizers, which can have hundreds of parameters with various routes.
@@
 Nonetheless, we believe that the insights here can extend naturally to more complex domains.
-Depending on the feature of interest, the observed dependence of loss performance on synthesis method is will likely remain stable even with added complexities such as layers of modulation, filtering, or nonlinearity.
+Depending on the feature of interest, the observed dependence of loss performance on synthesis method is likely to remain stable even with added complexities such as layers of modulation, filtering, or nonlinearity.
 
@@
 The main hypothesis tested here is whether the performance of differentiable loss functions is influenced by the synthesis techniques used for sound-matching.
-We created iterative sound-matching experiments by combining four different loss functions with four different sound synthesis programs.
+We conducted systematic iterative sound-matching experiments by combining four different loss functions with four different sound synthesis programs.
 We ranked the performance of the iterative sound-matching pipelines for every loss function and program, and observed that the success of the pipeline (that is, how closely the output sound matches the target sound) is program dependent.

### 2025-09-29 review_responses.tex
Updates from Overleaf

@@
     \item In lieu of listening tests, MSS or P-Loss are generally useful for large-scale benchmarking.
-However, we encourage confirmation of conclusions with manual listening, particularly if specificity is important.
+However, we encourage confirming the results with manual listening, particularly if specificity is important.
     \item For amplitude-modulated synthesis, DTW-based losses such as \DTWEnv{} are recommended over spectrogram differences and JTFS.
@@
   \item \textbf{IV.B Practical Recommendations} which distills our findings into actionable guidelines for practitioners.
-DTW_Env is recommended for amplitude-modulated synthesis, SIMSE for subtractive/noise-filtered synthesis, and MSS or P-Loss as large-scale proxies for sound-matching evaluation (with manual validation being strongly advised).
+DTW\_Env is recommended for amplitude-modulated synthesis, SIMSE for subtractive/noise-filtered synthesis, and MSS or P-Loss as large-scale proxies for sound-matching evaluation (with manual validation being strongly advised).
 We also note that our results challenge prior claims about JTFS’s universal superiority for mesostructures.

### 2025-09-29 sections/background_related.tex
Updates from Overleaf

@@
 As they noted, without a differentiable synthesizer, ``model-based'' (or what we call supervised) approaches cannot directly compare $\xtheta$ to $\xzero$, often opting for P-Loss, which may not correctly map the parameters of sound to the output audio~\cite{esling2019flow,han2023perceptual,masuda2023improving}.
-This approach used a model mapping sounds to parameters, another approximating the synthesizer, and a loss function which combines P-Loss with STFT differences.
+This approach used a model for mapping sounds to parameters, another for approximating the synthesizer, and a loss function which combines P-Loss with STFT differences.
 This ``Inversynth II'' (IS2) approach yielded significant improvements to previous works which did not use the STFT approximations for loss~\cite{esling2019flow,barkan2019inversynth}.
@@
 Based on manual hearing tests, this approach did not produce the ``correct'' sounds more frequently than previous works~\cite{kreuk2022audiogen}.
-However, it did yield better scores for ``artistic-interpretation'' (or what this works refers to as ``imitation'').
+However, it did yield better scores for ``artistic-interpretation'' (what we referred to as ``imitation'').
+It should be noted that audio embedding models are often trained using simpler loss functions---particularly those which utilize Fourier transforms---therefore improving such simpler loss functions is likely to improve the embedding models.
 
@@
 
-Having looked at the current literature, we identify four major areas of weakness in sound-matching. We target the first two issues in this work, while the latter two issues are left for future work.
+Having looked at the current literature, we identify four major areas of weakness in sound-matching. We target the first two issues in this work, while the latter two are left for future work.
 \begin{enumerate}
@@
 Together, these design choices supply the controlled experimental evidence missing from the literature, 
-clarifying how losses and synthesis methods interact in differentiable, iterative settings, 
+clarifying how losses and synthesis methods interact in iterative sound-matching,
 and offering insights likely to generalize to more complex domains.

### 2025-09-29 sections/methodology_and_analysis.tex
Updates from Overleaf

@@
 their outcomes automatically and manually
-(Section~\ref{sec:evaluation_manual_auto}). ``Experiments'' in this context
-refer to one complete iterative sound-matching run, from random parameter
-initialization through gradient-based optimization until termination. Scores
-from these experiments are then aggregated and statistically compared to
+(Section~\ref{sec:evaluation_manual_auto}). ``Experiment'' in this context
+refers to one complete iterative sound-matching run, from random parameter
+initialization through iterative gradient-based optimization until termination. Scores
+from these experiments are then aggregated and compared to
 determine best-performing losses and synth–loss pairings.

### 2025-09-29 sections/summary_weakness_conclusion.tex
Updates from Overleaf

@@
 However, given the simplicity of our programs and the occasional deviations, we cannot conclude that these automatic measures are substitutes for human hearing tests.
-As many previous works have done (see Table~\ref{tab:summary}, the use of MSS and/or P-Loss as general loss functions does seem appropriate, however, we advise the use of listening tests or custom made loss functions for more specific analysis.
+As many previous works have done (see Table~\ref{tab:summary}), the use of MSS and/or P-Loss as general loss functions does seem appropriate in general cases, however, the use of listening tests or custom made loss functions in specific contexts remains necessary.
 
@@
 \SIMSESpec{} performed best in subtractive synthesis, where scale-invariance captures noise-filtering effects more robustly than L1-based measures.
-These results suggest that the advantages of these novel losses are context-dependent, underscoring the need for further exploration of creative differentiable similarity measures.
+These results suggest that the advantages of loss functions are context-dependent, underscoring the need for further exploration of creative differentiable similarity measures.
 
-\textbf{Q3}: Can iterative differentiable optimization be applied effectively in synthesizers using classical DSP functions? \\
-The approach of designing DSP functions in Faust and transpiling to differentiable Faust code yielded meaningful outputs and enabled comparative evaluation of losses.
-The use of Faust for synthesizer definition was convenient, and transpiling to Jax enabled quick gradient optimization.
+\textbf{Q3}: Can iterative differentiable optimization be applied effectively with synthesizers using classical DSP functions? \\
+The approach of designing DSP functions in Faust and transpiling to differentiable Faust code was convenient, yielded meaningful outputs, and enabled comparative evaluation of losses.
 We were able to run our 1200 experiments within 72 hours on a laptop without a GPU\footnote{Lenovo ThinkPad T480, i5-8350U CPU at 1.70GHz, 32 GB RAM.}.
@@
 \subsection{Practical Recommendations}
-    Our findings not only provide answers to the research questions proposed in the introduction, but also serve as guidelines for practitioners in selecting appropriate similarity measures and synthesis methods for future sound-matching experiments.
+    Our findings not only provide answers to the research questions proposed in the introduction, but also serve as guidelines for practitioners in selecting appropriate similarity measures and synthesis methods for future experiments.
 \begin{itemize}
     \item In lieu of listening tests, MSS or P-Loss are generally useful for large-scale benchmarking.
-However, we encourage confirmation of conclusions with manual listening, particularly if specificity is important.
+However, we encourage confirmation of such results with manual listening, particularly if specificity is important.
     \item For amplitude-modulated synthesis, DTW-based losses such as \DTWEnv{} are recommended over spectrogram differences and JTFS.
@@
 In other words, different synthesizer programs work best with different loss functions.
-Notably, we see that our novel use of DTW and SIMSE based differentiable loss functions (\DTWEnv{} and \SIMSESpec) can outperform what are regarded as the SOTA loss functions in 3 of 4 cases, although this is highly synthesizer dependent.
+Notably, we see that our novel use of DTW and SIMSE based differentiable loss functions (\DTWEnv{} and \SIMSESpec) can outperform what are regarded as the SOTA loss functions in 3 of 4 cases, although their success is highly synthesizer dependent.
 
 P-Loss and MSS have frequently been used as automatic performance measures, yet their ``preference'' has rarely been compared to human rankings.
-We observed that automatic performance measures and manual listening tests were often in agreement, despite this, manual verification of sound-matching results remains a necessity for future works.
+We observed that automatic performance measures and manual listening tests were generally in agreement, despite this, manual verification of sound-matching results remains a necessity in non-general experiments due to occasional divergences between automatic and manual tests.
 
@@
 
-Like nearly all previous work in sound-matching, the main measure of success here was the accurate \textit{replication} of sounds (or synth parameters) rather than imitation of sounds outside of the synthesizer's domain.
+Like nearly all previous work in sound-matching, the main measure of success here was the accurate \textit{replication} of sounds (or synthesizer parameters) rather than \textit{imitation} of sounds outside of the synthesizer's domain.
 Replication of arbitrary features of sounds is an important component of sound-design, though it is much harder to define or measure.

### 2025-10-01 review_responses.tex
Updates from Overleaf

@@
 \textbf{Some guidance (e.g., Use DTW for AM synthesis; SIMSE for subtractive synthesis) would strengthen the paper’s relevance to practitioners.} \\
-\textbf{(R1.1) The paper would be improved by formally stating additional hypotheses that are currently only implicit.} For example, the observed agreement between manual and automatic evaluation metrics, the effectiveness of the novel losses, and the usability of iterative optimization as a general strategy are all important insights, but are not positioned as testable hypotheses upfront.
+\textbf{(R1.1) The paper would be improved by formally stating additional hypotheses that are currently only implicit.} For example, the observed agreement between manual and automatic evaluation metrics, the effectiveness of the novel losses, and the usability of iterative differentiable optimization as a general strategy are all important insights, but are not positioned as testable hypotheses upfront.
 \\
@@
 Background and Related Work}.
-We applied major edits and rewrites to these sections to improve reading and remove redundancies between the various subsections while taking precautions in compromising the flow.
+We applied major edits and rewrites to these sections to improve reading and remove redundancies between the various subsections while taking precautions against compromising the flow.
 \textbf{II. Background and Related Work} contains the following topics in order:  
@@
   \item \textbf{III.
-Experimental Setup and Results A-E}: Sections III.A–E now exclusively describe the methodological components: the synthesizer programs, the definitions of losses, the optimization procedure, evaluation metrics, and statistical tests.
+Experimental Setup and Results A-E} sections now exclusively describe the methodological components: the synthesizer programs, the definitions of losses, the optimization procedure, evaluation metrics, and statistical tests.
 Each subsection is focused on describing how the experiments are designed, without interpretation of outcomes.
@@
 
-  \item Experimental findings are presented in \textbf{III-F Best Performers for Each Program} and \textbf{III-G Consistency in Rankings}, which together form the \emph{Results} component.
+  \item Experimental analysis are presented in \textbf{III-F Best Performers for Each Program} and \textbf{III-G Consistency in Rankings}, which together form the \emph{Results} component.
 Section III-F reports which losses performed best for each synthesizer, while Section III-G quantifies the consistency of rankings across bootstrap resamples.
@@
 
-  \item \textbf{IV. Discussion} then provides the qualitative interpretation, structured into four dedicated subsections:  
+  \item \textbf{IV. Discussion} is structured into four dedicated subsections:  
     \begin{enumerate}
@@
 \noindent\textbf{Reviewer:} \\
-\noindent While the comparison is comprehensive, the novel losses (DTW\_Envelope, SIMSE\_Spec) are adaptations of known techniques and are not framed with sufficient theoretical justification.
+\textbf{\noindent While the comparison is comprehensive, the novel losses (DTW\_Envelope, SIMSE\_Spec) are adaptations of known techniques and are not framed with sufficient theoretical justification.
+}\\
 
@@
 
-While this may not be the "qualitative" or "spectral" analysis, we believe it is instructive on the mechanisms underlying our quantitative findings and highlight the importance of loss landscape navigation as an open area for future research.
+While this may not be the "qualitative" or "spectral" analysis, we believe it is instructive on the mechanisms underlying our quantitative findings and highlight the importance of loss landscape navigation as an open area for future research (See also~\nameref{R4.5}).
 
@@
 \noindent\textbf{Response} \\
-
 In \textbf{IV.A Key Findings}, we discuss important findings and provide answers to our research questions.
@@
 \end{itemize}
-For a more detailed response, see our response to the Reviewer.1 in Sections~\nameref{R1.6} and \nameref{R1.7}. Our response to \nameref{R1.5} may also be of interest.
+For a more detailed response, see our response to Reviewer 1 in Sections~\nameref{R1.6} and \nameref{R1.7}. Our response to \nameref{R1.5} may also be of interest.
 
@@
 \noindent\textbf{Response:}
-The original statement referred to how realistically in sound-design, a "correct" solution rarely exists due to our subjective experience of sound.
+The original statement referred to how realistically in sound-design, a ``correct" solution rarely exists due to our subjective experience of sound and personal goals.
+The existence of an objective sound-matching goal is only true in \textit{in-domain replication} scenarios, which are far removed from real life applications.
 To reduce confusion, the section has been rephrased as follows:
@@
 The listening tests were done without use of any labels or previous exposure to outputs, which we believe minimizes the risk of bias.
+Additionally the strong correlation between the two listeners adds to our confidence in using the listening test results as the ground truth.
 Unfortunately, adding another listener was not possible given the allotted time for the revisions.
@@
 This approach does not account for two sounds that are nearly identical but slightly time-shifted~\cite{vahidi2023mesostructures,turian2020sorry}.
+We've shown in this work that DTW can be used in loss functions to account for envelope time-shifts.
 
 \subsection{\textbf{R4.5} Can We Overfit? Is it a Concern?}
+\label{R4.5}
 \noindent\textbf{Reviewer:}\\
@@
 This means that \textit{if} the target parameters have been reached, subsequent updates will not diverge from it (or at least stay very close).
-We can share our visualization of the gradient update paths in 2D, which we felt was not statistically robust enough for the paper.
+We share some visualizations of the gradient update paths in 2D, which we felt was not statistically robust enough for the paper.
 Figures~\ref{fig:p0_simse_iterations} and~\ref{fig:p2_dtw_iterations}, we see that once the gradient updates (small red dots connected by lines) take the parameters from the starting position (green circle) to the target (green star), subsequent updates fluctuate slightly and randomly around the target.

### 2025-10-10 review_responses.tex
Updates from Overleaf

@@
 For instance, the synthesizer section defines parameter ranges, while the evaluation section specifies how statistical significance is tested.
-The only partial overlap is at the beginning of the \textbf{III-F Best Performers for Each Program}, which serves as a transition between results and discussion.
+The only partial overlap is at the beginning of the \textbf{III-F Best Performers for Each Program} section, which serves as a transition between results and discussion.
 
@@
 We share some visualizations of the gradient update paths in 2D, which we felt was not statistically robust enough for the paper.
-Figures~\ref{fig:p0_simse_iterations} and~\ref{fig:p2_dtw_iterations}, we see that once the gradient updates (small red dots connected by lines) take the parameters from the starting position (green circle) to the target (green star), subsequent updates fluctuate slightly and randomly around the target.
+In Figures~\ref{fig:p0_simse_iterations} and~\ref{fig:p2_dtw_iterations}, we see that once the gradient updates (small red dots connected by lines) take the parameters from the starting position (green circle) to the target (green star), subsequent updates fluctuate slightly and randomly around the target.
 

### 2025-10-10 sections/summary_weakness_conclusion.tex
Updates from Overleaf

@@
 \textbf{Weaknesses:} 
-While we cannot prove that a universally best similarity measure does not exist, we can advocate for more creativity in the field based on our findings.
+While we cannot prove that a universally best similarity measure does not exist, we can advocate for more creativity in the field.
 Compared to previous work, we presented a more cohesive approach to iterative sound-matching which utilizes a variety of loss functions and synthesis methods.

### 2025-10-27 main.tex
final files

@@
 \usepackage[numbers]{natbib} %removed cite package
+
 \usepackage{tikz}
@@
 \newcommand{\todo}[1]{\textcolor{red}{#1}}
+
+\newcommand{\xtheta}{x_{\theta}} % synthesizer output
+\newcommand{\xzero}{x_{0}}       % target sound
 
@@
 \newcommand{\SynthSelect}{\textbf{Synthesis Selection}}
-\newcommand{\PeriodicLoss}{\textbf{Periodic Loss}}
-\newcommand{\OutDomain}{\textbf{Out-Domain Generation}}
+\newcommand{\PeriodicLoss}{\textbf{Loss Landscape Navigation}}
+\newcommand{\OutDomain}{\textbf{Out-of-Domain Generation}}
 \newcommand{\BPNoise}{\textbf{BP-Noise}}  
@@
 %  ================================
+
 \begin{document}
+\bibliographystyle{IEEEtran}
+
 \title{Evaluating Sound Similarity Metrics for Differentiable, Iterative Sound-Matching}
@@
 \author{Amir Salimi, Abram Hindle, Osmar R. Za{\"i}ane}
-
+\maketitle
 % \markboth{IEEE TRANSACTIONS ON AUDIO, SPEECH AND LANGUAGE PROCESSING, VOL -, 2025}%
@@
 % }{}
-
-\maketitle
 
@@
 We observed a moderate level of consistency among the three performance measures.
-Our post-hoc analysis shows that the loss function performance is highly dependent on the synthesizer.
+Our post hoc analysis shows that the loss function performance is highly dependent on the synthesizer.
 These findings underscore the value of expanding the scope of sound-matching experiments, and developing new similarity metrics tailored to specific synthesis techniques, rather than pursuing one-size-fits-all solutions.
 \end{abstract}
-
-% the automatic performance measures have not been compared to human hearing
-% lack of creativity in loss function definitions 
 \begin{IEEEkeywords}
@@
 \end{IEEEkeywords}
+% \hphantom{\vphantom{\begin{minipage}{\textwidth}\bibliography{references}\end{minipage}}}
 
 % need to simplify the intro, and expand the hypothesis
-\section{Introduction}
-Since the 1970s, artists such as Suzanne Ciani~\cite{ciani_life_in_waves} have explored the use of audio synthesizers to replicate natural sounds—such as a bottle cap popping off—in ways that preserve the essential qualities of the original while showcasing the unique characteristics of the synthesizer~\cite{creativecherep2024}.
-Rather than merely replicating sounds, this approach uses synthesizers for artistic reinterpretation.
-However, the automation of this interpretive approach to sound design using digital audio synthesizers has rarely been explored, and significant knowledge gaps remain even in the simpler problem of replicating sounds.
-One such gap is the lack of diversity and depth in the analysis of sound-similarity measures and their interaction with various synthesis methods.
-As a stepping stone toward automated interpretive sound design, we use automatic and manual sound-evaluations to investigate whether the optimal choice of a sound-similarity measure is dependent on the synthesis method.
+\input{sections/introduction.tex}
+\input{sections/background_related}
+\input{sections/methodology_and_analysis}
+\input{sections/summary_weakness_conclusion}
 
-A digital audio synthesizer is any software used for the creation and manipulation of audio.
-Since the advent of digital signal processing (\gls{DSP}) in the 1960s~\cite{stranneby2004digital}, a wide variety of parametric audio generation functions have been proposed~\cite{lyons1997understanding,russ1999sound,shier2020spiegelib}.
-Manual sound design with a synthesizer is an iterative process of trial and error; when the artist has a desired (or target) sound in mind, sound design involves listening to the output, determining whether the output is closer or farther from the target, and adjusting the parameters of the synthesizer accordingly until the target is reached~\cite{russ1999sound}.
-This paper explores the automation of this process, which we call \textit{iterative sound-matching}.
+% \bibliographystyle{IEEEtran}
+\bibliography{references}
 
-Automatic sound design has been a subject of research for decades, with reduction of tinkering time and creation of ``interesting" sounds as the main motivators \cite{krekovic2019insights,turian2020sorry,horner1993machine,salimi2020make,esling2019flow,engel2020ddsp,mitchell2007evolutionary,shier2020spiegelib,masuda2021soundmatch,masuda2023improving}.
-\textit{Sound-matching} is a common form of automatic sound design explored in literature, where a similarity metric (or loss function) is given, and the goal is to find parameters for a synthesizer that replicate \textit{all} or \textit{some} of the characteristics of a target sound as best as possible~\cite{horner1993machine,mitchell2007evolutionary,masuda2023improving}.
-Our analysis of past works shows that this general goal of ``sound-matching'' can be broken down in to very different subtypes, many of them rarely or never explored, despite their relevance to practical applications.
-This leads us to a typology of sound-matching tasks, and the importance of \textit{differentiable, iterative sound-matching}, rarely explored in previous literature~\cite{vahidi2023mesostructures}.
-
-\textit{Iterative sound-matching} refers to works that mimic the back-and-forth process of listening to a synthesizer's output and updating the parameters until the termination conditions are met.
-The two necessary components are the \textit{loss function}, which determines how close two sounds are, and the \textit{parametric synthesizer} that outputs sounds based on input parameters.
-In the iterative workflow, a target sound is selected, the synthesizer outputs sounds based on its parameters, and the loss function determines the proximity of the output to the target.
-From there, the synthesizer parameters are updated, the output and target are compared, and so on.
-
-Despite the early popularity of the iterative approach~\cite{justice1979analytic,horner1993machine,mitchell2007evolutionary}, recent works in sound-matching have favored supervised learning using inference models, deep neural-networks (\gls{DNNs}), and large datasets.
-Rather than continually interacting with the environment, these models learn from a dataset of inputs (synthesizer parameters) and outputs (corresponding sound), and predict the synthesizer parameters for new sounds~\cite{yee2018automatic,han2023perceptual,han2024learning,masuda2021soundmatch,masuda2023improving}.
-
-Iterative sound-matching is interesting for multiple reasons.
-First, it better resembles the human approach to sound-design, which requires continually interacting with the environment and making decisions according to a feedback signal~\cite{sutton2018reinforcement}.
-Second, the trial-and-error approach leaves a trail of parameter sets from the starting point to the end goal, which could be of use to sound designers.
-Third, relative to the DNN-based inference models, it provides a more tractable environment with room for creative approaches to sound-similarity definitions, and more amenable to deployment on different synthesizers as it does not require pre-training on large datasets.
-
-This work takes a \textit{differentiable} approach to iterative sound-matching, a natural extension of early works in iterative sound-matching combined with recent advances in Machine Learning (\gls{ML}) software~\cite{braun2024dac,paszke2019pytorch} and increase in computational power~\cite{moore2012introduction,schaller1997moore}.
-A differentiable approach requires differentiable loss and synthesizer functions, allowing the synthesizer parameters to be updated using gradient descent~\cite{goodfellow2016deep,boyd2004convex}.
-
-We discuss the history of sound-matching, and provide a nomenclature for different subtypes of sound-matching tasks.
-A number of weaknesses in current literature (see Section~\ref{sec:lacking}) are identified.
-The experiments in this work focus on two of these core issues, referred to as \LossSelect~and \SynthSelect.
-The former refers to the search for an optimal loss function, lack of novel algorithms, and the---perhaps needless---focus on outperforming state of the art (\gls{SOTA}), while the latter refers to the lack of diversity in synthesis methods.
-
-In the case of \LossSelect, there have been many works and experiments comparing the accuracy of loss functions and choosing the best performers~\cite{vahidi2023mesostructures,turian2020sorry,engel2020ddsp,uzrad2024diffmoog,han2023perceptual,masuda2021soundmatch,turian2020sorry,bruford2024synthesizer}.
-We note that these claims regarding the effectiveness of one function versus another have consistently been made in limited contexts that may not generalize to other settings.
-One such limitation is the synthesizers used, and ties into the problem of \SynthSelect, as typically only a single (and often, simple) synthesizer has been used for testing the effectiveness of loss functions.
-Furthermore, there is a lack of diversity in the underlying functions used by the synthesizers, with simple Frequency Modulation (\gls{FM}) being the most common method of synthesis~\cite{horner1993machine,vahidi2023mesostructures,mitchell2007evolutionary,caspe2023envelope}.
-
-Considering the interplay between the two problems, we design experiments that address them simultaneously.
-Testing the hypothesis that the performance of a similarity measure (or loss function) is influenced by other factors in the environment, particularly the method of synthesis.
-We combine 4 loss functions with 4 differentiable synthesizers, and compare their performance as measured by the similarity of the final output to the target.
-We evaluate the final similarity with two different automatic methods, as well as manual listening tests conducted by two of the authors.
-We compare the distributions of the evaluation scores in order to rank the loss functions from best to worst.
-The results of these experiments show that when accounting for different synthesizers, there is no consistent choice of sound similarity measure for optimal performance in sound-matching.
-
-We note that for some synthesizers, our novel use of ``dynamic time warping of envelopes'' as a loss function for differentiable sound-matching outperforms the standard methods of audio comparison.
-These results imply that rather than aiming to outperform the SOTA and finding a universal best performer, more effort could be placed towards diversifying the existing similarity measures of sound as well as conducting experiments using a variety of synthesis methods.
-Sound design is an inherently creative endeavor, and would likely benefit from a diversified range of solutions for different approaches.
-
-\section{Background}
-\label{sec:background}
-In this Section, we give a formal definition of sound-matching and the necessary components.
-These discussions lead to a nomenclature for types of sound-matching, as well as major perennial issues in the field.
-
-\subsection{Digital Audio Synthesis}
-\label{sec:diffSynth}
-A digital audio synthesizer is any software used for the creation and manipulation of digital audio.
-Digital synthesizers use signal processing chains with a variety of digital signal processing (\gls{DSP}) functions to create sounds.
-These functions are often parametric, and the set of parameters given to the synthesizer's chain of DSP functions is called a synthesizer program.
-
-The simplest form of DSP function is a sinusoidal tone. For example, consider:
-\[ x(n) = sin( 2\pi n T)\]
-
-where $T$ represents time in seconds, and $n$ is the frequency parameter.
-If $n$ is 1, then $x(n)$ would be a waveform with a frequency of 1 hertz, meaning that the value of $x(n)$ oscillates to the same point once every second.
-Waveform generators are called oscillators for this reason.
-
-% If $n$ is in the range of 30 to 20,000 (the upper limit is often lower for some), $x(n)$ could be played as an audible tone~\cite{smith2007mathematics,jilek2014reference}.
-
-Since the advent of digital signal processing in the 1960s~\cite{stranneby2004digital}, a wide variety of parametric audio generation functions have been proposed~\cite{lyons1997understanding,russ1999sound,shier2020spiegelib}.
-This includes oscillators, filters, equalizers, and envelopes, which can be used sequentially or in parallel~\cite{lyons1997understanding,russ1999sound}.
-Sound design with a synthesizer is done by modifying the parameters of the functions until reaching a desired output~\cite{roads1996computer,pinch2004analog}.
-
-A relatively small subset of these synthesis methods have been used for automatic sound-matching, notably additive and/or subtractive synthesizers~\cite{esling2019flow,yee2018automatic,mitchell2007evolutionary,salimi2020make}, physical modeling~\cite{riionheimo2003parameter}, and frequency/amplitude modulation (\gls{FM}/\gls{AM})~\cite{horner1993machine,vahidi2023mesostructures}.
-
-Differential DSP (\gls{DDSP}) is a general term for applications that combine machine learning techniques such as gradient descent~\cite{goodfellow2016deep,boyd2004convex} with DSP~\cite{engel2020ddsp}.
-Implementing complex DSP functions in a differentiable manner can be challenging, and effective differentiable audio similarity measures require careful mathematical expression of the desired attributes of sound.
-These issues have likely contributed to the limited exploration of this domain.
-Perhaps the most relevant recent work is by Vahidi \textit{et al.}, featuring the comparison of two differentiable loss functions with a simple FM synthesizer~\cite{vahidi2023mesostructures}.
-
-\subsection{Sound Representation and Loss Functions}
-\label{sec:loss_funcs}
-A digital sound (or an audio signal) is a series of numbers~\cite{smith1991viewpoints,smith2007mathematics}.
-To compare two digital sounds, the two corresponding series are passed to a function that measures their similarity.
-Two signals can sound identical to our ears, without having any values in common~\cite{moore2012introduction}.
-This necessitates the use of proxy representations (or feature extractors) when comparing sounds automatically.
-Similarity between the target sound and the synthesizer output is then measured by some form of subtraction and summation of the proxy representations.
-
-In sound-matching, particularly in a Deep Learning (\gls{DL}) context~\cite{goodfellow2016deep}, the similarity function can also be called a \textit{loss} function, where the emphasis is on measurement and reduction of the distance between target and output.
-It is important to note that there is a close relationship between the loss function $L$ and the sound representation function $\phi$.
-$L$ is the result of a distance measure $d$ applied to the features extracted by $\phi$.
-
-\[
-L(\theta, t) = d\langle\phi(t),\phi(x)\rangle
-\]
-
-\noindent
-
-A proxy representation is the output of the function \( \phi \), which can be thought of as a feature extraction function that maps the sounds \( t \) and \( x \) to their respective representations. 
-The proportionality or distance metric $d$ has typically been the L1 distance~\cite{turian2020sorry,richard2025model}, calculated as the mean of the absolute difference between every point in the proxy representation~\cite{engel2020ddsp,vahidi2023mesostructures}:
-\[
-L(\theta, t) = \left\| \phi(t) - \phi(x) \right\|_1
-\]
-
-Here we discuss four methods of audio representation and the corresponding loss functions.
-Three of which have been used in previous works, and a novel method not previously used in iterative sound-matching.
-
-\subsubsection{Parameter Loss}
-A common measure of similarity in sound-matching is the distance between synthesizer parameter sets, referred to as ``P-Loss"~\cite{han2023perceptual}.
-Typically, for the implementation of P-Loss the parameter sets are treated as vectors in space, and L1 or L2 distance is applied.
-There are two major limitations to this approach: First, the target and output sound must be made by the same synthesizer; otherwise the parameter sets cannot be compared (see Section~\ref{sec:matching_types}).
-Second, the relationship between synthesizer parameters and the audio output is not linear~\cite{shier2020spiegelib,han2023perceptual,esling2019flow}.
-
-\subsubsection{Fourier Spectrograms}
-\label{sec:fourier_specs}
-Fourier-based transformations such as short-time Fourier transforms (\gls{STFT}), Mel-spectrograms, and Mel-frequency cepstral coefficients have been viewed as the de facto and state-of-the-art representation of audio~\cite{beauchamp2003error,mitchell2007evolutionary,yee2018automatic}, however, there are many issues associated with their use in sound-matching~\cite{turian2020sorry,vahidi2023mesostructures,han2023perceptual,uzrad2024diffmoog}.
-Fourier transformations allow for the conversion of a signal from the time-domain to the frequency domain.
-Audio spectrograms can be generated by segmentation of a piece of audio into overlapping windows followed by the application of Fourier transforms to each window.
-They are costly to compute, but provide a better temporal view of changes in frequency content~\cite{muller2007dynamic,smith2007mathematics}.
-There are different types of spectrograms that have a basis in Fourier transformations, but the most notable and commonly used is the STFT.
-
-What we call \textit{Fourier-based Spectrograms} are variations on the STFT approach.
-For example, Mel-Spectrograms \textit{bin} frequencies on a near-logarithmic scale to better match human perception of frequencies~\cite{muller2007dynamic}.
-Multi-scale spectrograms (\gls{MSS}) used in recent works are a simple weighted average of multiple spectrograms with different parameters such as window size, number of frequency bins, and hop size~\cite{engel2020ddsp,vahidi2023mesostructures}; this may provide some improvements at a higher computational cost~\cite{turian2020sorry,engel2020ddsp}.
-
-\subsubsection{Joint-Time Frequency Spectrum}
-Recent works have focused on the limitations of parameter and spectral loss functions in sound-matching~\cite{vahidi2023mesostructures,uzrad2024diffmoog}, seeking to create more effective general solutions for the comparison of audio.
-Noting the aforementioned weaknesses of comparing STFT spectrograms, Vahidi \textit{et al.} proposed differentiable Joint-Time Frequency Scattering (\gls{JTFS})~\cite{anden2015joint} as an alternative to spectrogram loss in sound-matching, and showed improved performance in sound-matching with differentiable chirplet synthesizers~\cite{vahidi2023mesostructures}.
-JTFS is the result of the application of a 2D wavelet transformation to the time-frequency representation of a signal~\cite{anden2015joint}.
-
-\subsubsection{Dynamic Envelope Warping}
-Dynamic Time Warping (DTW) is a method for measuring similarity between multi-dimensional time-series~\cite{rabiner1993fundamentals,muller2007dynamic,giorgino2009computing}.
-Given any two time-series $X = \{x_1,x_2,...,x_m\}$ and $Y = \{y_1,y_2,...,y_n\}$, we have indices $i\in\{1...m\}$ and $j\in\{1...n\}$ defining $X$ and $Y$.
-When the series are \textit{warped}, these indices change to expand or contract different portions of the series.
-To borrow the notation given by Muller~\cite{muller2007dynamic}, warped indices are a sequence $p=(p_1,...,p_L)$, where \(p_\ell = (m_\ell, n_\ell) \in [1 : m] \times [1 : n] \text{ for } \ell \in [1 : L]\), meaning that the indices for $X$ and $Y$ are reorganized under special conditions.
-In classical DTW, these conditions are \textit{monotonicity}, \textit{boundary matching}, and \textit{single step-size}.
-
-DTW measures the distance between the time-series \textit{after} alignments, typically using Euclidean distance.
-The distance between a time-series and shifted versions of itself would be 0, regardless of shift amount~\cite{tavenard.blog.dtw}.
-Additional rules can be imposed to keep alignments locally constrained~\cite{itakura1975minimum,sakoe1978dynamic}.
-% \textcolor{red}{DTW was originally used for signal comparison but using it the way we do is novel}
-
-\subsection{Automatic Sound Matching}
-\label{sec:sound_matching_definition}
-The problem of sound-matching has been tackled from a variety of perspectives as it involves a number of modular parts: the choice of synthesizer $g$, the target sounds of interest, the representation function $\phi$, and the heuristic for finding the optimal $\theta$.
-To give a formal definition of sound-matching, we expand the definition given by recent works~\cite{vahidi2023mesostructures,han2023perceptual}.
-The major components are as follows:
-\begin{itemize}
-    \item $g(\theta)$: Parametric audio synthesizer $g$ which takes on parameters $\theta$ 
-    \item $x$: The output of $g$, given a set of parameters $\theta$ or $g(\theta) = x$ 
-    \item $t$: The target sound which we want to replicate or imitate. 
-    \item $\phi$: Representation function or feature extractor. $\phi$ is applied to $x$ and $t$ to facilitate their comparison by the loss function.
-    \item $L$: Loss or error function. $L(\theta,t)$ is a measure of distance between $x$ and $t$. Often proportional to the subtraction of their representations, or $ \phi(x) - \phi(t)$
-\end{itemize}
-
-\highlight{It could be argued that the implementation of these components is not an optimization problem with a correct answer, but a \textit{creative} endeavor depending on the artist's needs}.
-Consider the hypothetical case where the goal might be to create an 8-bit~\cite{collins2007loop} version of a high quality, organic snare drum: in such a case, there is no optimal or ``correct" version of an 8-bit snare sound.
-
-\subsection{Sound Matching Types}
-\label{sec:matching_types}
-
-Sound-matching tasks can be categorized based on the choices of target sound $t$, the loss $L$ and representation function $\phi$, and the heuristic used for minimizing $\theta$.
-We will discuss the subtypes of sound-matching tasks in the upcoming sub-sections.
-
-\subsubsection{In-Domain vs Out-of-Domain}
-\label{sec:in-domain}
-The choice of domain depends on whether we want to use the same synthesizer for the target and output sounds, which we call \textit{in-domain}, or have target sounds that came from sources other than the synthesizer, or \textit{out-of-domain}.
-To paraphrase the description given by Masuda \textit{et al.}~\cite{masuda2021soundmatch}, if $g$, the synthesizer of choice, can accurately replicate the target sound $t$, or put differently, if $t$ itself is an output of $g$, then the sound-matching task is \textit{in-domain}.
-If $t$ is not an output of $g$, then the sound-matching task is \textit{out-of-domain}.
-In-domain tasks in general are simpler, and often there is a guarantee that there is a correct answer to the sound-matching problem, particularly if the goal is accurate replication of the sound.
-If the target sound is out-of-domain, replication is not guaranteed, and the goal becomes the \textit{imitation} of some aspect of sound.
-
-\subsubsection{Replication vs Imitation}
-Regardless of the domain, the generation goal can be \textit{replication} or \textit{imitation} of the target sound.
-In replication, the goal is to make an identical copy of the target sound.
-Imitation is an artistic pursuit and harder to define, since the goal is to make new sounds that only retain a subset of the target's sonic features.
-While closely related to the in-domain versus out-of-domain problem, the choice of replication versus imitation is more dependent on how the loss and representation functions are defined.
-
-% From a signal perspective, replication is a scenario where the goal is to find a $\theta$ such that $t-x=0$ where $x=g(\theta)$.
-From a sound perspective, this definition of replication is unnecessarily restrictive, as $t$ and $x$ can sound identical but contain very different values.
-For example, if $t$ and $x$ were periodic waveforms at an audible frequency $f$ with different phase values, their subtraction would not be 0, yet they would sound identical to our ears, as our ears are insensitive to phase values~\cite{moore2012introduction}.
-Representation functions such as STFT spectrograms or Mel-frequency cepstral coefficients (\gls{MFCC}) become viable in these cases, as they allow for separation and removal of the phase information within an audio signal.
-The \textit{sound replication} goal can be defined, in layman's terms, as ``$x$ sounds exactly like $t$".
-More formally, the goal is to find parameters such that $\phi(t)-\phi(x)=0$, where $\phi$ transforms an audio into a spectrogram that removes the phase information from sounds, but otherwise represents the sonic information as accurately as possible.
-
-% With \textit{sound imitation}, the goal is the replication of some aspect of the target sound.
-Instrument conversion is a common example of imitation.
-As an example, a note played by a violin is converted to the same note played by a piano.
-Conversion of even longer pieces has been made possible with large enough datasets~\cite{engel2020ddsp}.
-Imitation of sound envelopes (how the loudness changes over time) is another interesting idea~\cite{caspe2023envelope}.
-Transfer of other sound properties such as reverb and delay has also been a subject of research~\cite{engel2020ddsp}.
-
-% A target sound can have many properties of interest, and the imitation task is the transfer of a subset of those properties to other sounds.
-Relative to replication, imitation is a very open ended goal.
-From a research point of view, success in the replication task can be measured more reliably, often without the need for manual evaluation.
-With imitation, listening tests are needed to ensure quality of outputs, particularly in a supervised learning setting where learning is done by examples~\cite{salimi2020make,masuda2021soundmatch}.
-The need for subjective evaluations has perhaps contributed to the lack of research in imitation and out-of-domain sound-matching.
-
-\subsubsection{Supervised Versus Direct Optimization}
-\label{sec:optimization}
-
-The goal of sound-matching is to find the optimal parameters $\theta^*$ that minimize the loss between the synthesizer output and the target sound. 
-\[
-\theta^* = \arg\min_{\theta} L(\theta,t)
-\]
-
-The heuristics in past works can be broadly split into two categories:
-\begin{enumerate}
-    \item Supervised (or inference) methods which use datasets of sound and parameter pairs to describe the generation objective, often with a large number of input and output examples~\cite{engel2020ddsp,salimi2020make,yee2018automatic,esling2019flow}
-    \item Unsupervised or \textit{direct sound optimization} methods~\cite{horner1993machine,mitchell2007evolutionary,yee2018automatic,vahidi2023mesostructures}.
-Usually require at most one target example, and a loss function that measures the difference between the target and synthesizer outputs.
-\end{enumerate}
-
-Genetic algorithms (\gls{GA})~\cite{holland1992genetic} can be used for direct sound optimization, and have been the earliest and most common heuristic in sound-matching~\cite{horner1993machine,mitchell2007evolutionary,yee2018automatic}.
-These algorithms start with arbitrary parameter sets that can be treated as an evolving population where the genomes are the parameter values.
-The most fit members of the group are the parameters which perform the best in the loss function, and create a new generation of parameters via mutation (random change in subset of parameters) and crossovers (combination of parameter sets); this process repeats until the goal or a maximum number of generations is reached.
-
- Differentiable settings have the benefit of making goal-oriented updates to the synthesizer parameters (the goal being the minimization of loss).
-There are also drawbacks: other than requiring more computation power, differentiable functions require careful implementation of operations that are continuous and numerically stable; this makes the implementation of signal processing functions quite difficult, contributing to the scarcity of works in this domain.
-
-% One important aspect of the heuristics is whether they imitate the iterative nature of manual sound design---that is, listening to the output and adjusting the synthesizer parameters recursively until the desired goal is reached.
-Supervised methods generate their final output in a single step, while direct sound optimization methods refine their outputs iteratively until termination, which resembles manual sound design more closely.
-
-\subsection{Iterative Sound Matching}
-\label{sec:iterative_sound_matching}
- \todo{Figure~\ref{fig:sound_design_loop_iterative} shows a general model for iterative sound-matching.
-To begin the process, the parameters are arbitrarily initialized (usually random generation), the similarity of the target and output is measured, and the parameter is \textit{optimized} with the goal of increasing the similarity, or alternatively, reducing the loss.
-This process repeats until termination.
-}
-\begin{figure}[ht]
-    \centering
-\begin{tikzpicture}[node distance=2cm, auto]
-
-    % Nodes
-    \node (start) [text centered] {\( g (\hat{\theta}) = x \)};
-    \node (L) [above of=start,right of=start, text centered] {\( L(\hat{\theta},t) {\ \propto \ } \phi(x) - \phi(t) \)};
-    \node (optimize) [below of=L,right of=L, text centered] {Optimize $\hat{\theta}$};
-    \node (new_theta) [below of=optimize,right of = start, text centered] {New \( \hat{\theta} \)};
-
-    % Highlight and arrow for the start node
-    \draw[->, very thick, red] (start) ++(-2,1.5) -- (start)
-        node[midway, below, align=center, sloped, color=red] {Start}
-        node[midway, above, align=center, sloped, color=red] {Random $\hat{\theta}$};
-
-    % Arrows with multi-line labels
-    \draw[->, bend left] (start) to node[midway, right, align=center] {} (L);
-    \draw[->, bend left] (L) to node[midway, right, align=center] {} (optimize);
-    \draw[->, bend left] (optimize) to node[midway, below, align=center] {} (new_theta);
-    \draw[->, bend left] (new_theta) to node[midway, left, align=center] {} (start);
-
-\end{tikzpicture}
-    % \caption{ Iterative approach to sound design.
-Synthesizer $g$ with arbitrary initialized parameters $\hat{\theta}$, creates sound $x$.
-The target sound $t$ is used as the goal.
-Parameters $\hat{\theta}$ are adjusted to minimize error (or loss) $L(\hat{\theta},t)$, where $L$ is proportional to the difference between the representations of $x$ and $t$.
-$\phi$ is the audio feature extractor, or representation function.}
-    \caption{ Iterative approach to sound design.}
-    \label{fig:sound_design_loop_iterative}
-\end{figure}
-
-In the upcoming ``Past Works'' section, we will discuss the evolution of prior sound-matching research, with \textit{differentiable iterative sound-matching} being the latest extension of the field.
-
-\begin{table*}[t]
-\centering
-\caption{Summary of select works in sound-matching.
-Gen.
-refers to whether the generations are done in 1-shot (generally, output by a neural network) or a search is conducted with the goal of iteratively getting closer to the target sound.}
-\setlength{\tabcolsep}{3pt} % Reduce horizontal spacing
-\renewcommand{\arraystretch}{1.05} % Reduce vertical spacing
-\scriptsize % Reduce font size
-\resizebox{\textwidth}{!}{
-\begin{tabular}{|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{2cm}|>{\raggedright\arraybackslash}m{1.2cm}|>{\raggedright\arraybackslash}m{1.2cm}|>{\raggedright\arraybackslash}m{1cm}|>{\raggedright\arraybackslash}m{1cm}|}
-\hline
-\textbf{Work} & \textbf{Synthesis} & \textbf{Loss} & \textbf{Heuristics} & \textbf{Goal} & \textbf{Domain} & \textbf{Year} & \textbf{Gen. } \\
-\hline
-Horner \textit{\textit{et al.}}~\cite{horner1993machine} & Non-differentiable FM & STFT (McAulay-Quatieri) & GA & Replicate & In & 1991 & Iter \\
-\hline
-Mitchel \textit{\textit{et al.}}~\cite{mitchell2007evolutionary} & Non-differentiable FM & Spectral Relative & Evolutionary & Replicate & In (Contrived) & 2007 & Iter \\
-\hline
-Yee-King \textit{\textit{et al.}}~\cite{yee2018automatic} & Non-differentiable VST & P-Loss (for Supervised), Spectral (MFCC) & Supervised \& Evolutionary & Replicate & In & 2018 & Both \\
-\hline
-Esling \textit{\textit{et al.}}~\cite{esling2019flow} & Non-differentiable VST & Spec MSS/SC P-Loss & Supervised \& Modeling & Replicate & In* \& Out & 2019 & 1-shot \\
-\hline
-Shier \textit{\textit{et al.}}~\cite{shier2020spiegelib} & Custom, non-diff & FFT \& Spectral & Supervised \& Direct & Replicate & In & 2020 & Both \\
-\hline
-Salimi \textit{\textit{et al.}}~\cite{salimi2020make} & DSP & STFT and Envelope & Supervised & Imitate & Out & 2020 & 1-shot \\
-\hline
-Masuda \textit{\textit{et al.}}~\cite{masuda2021soundmatch} & DDSP & P-Loss initially, fine-tune with Spec. & Supervised & Both & In \& Out & 2021 & 1-shot \\
-\hline
-Han \textit{\textit{et al.}}~\cite{han2023perceptual} & NN-encoder $\rightarrow$ physical model & PNP (approx. of STFT) & Supervised & Replicate & In & 2023 & 1-shot \\
-\hline
-Vahidi \textit{\textit{et al.}}~\cite{vahidi2023mesostructures} & Differentiable FM & JTFS/MSS & Gradient desc. & Replicate & In & 2023 & Iter \\
-\hline
-Barkan \textit{et al.}~\cite{barkan2023inversynthII}& Diff. Synthesizer Proxy& P-Loss + L1 Spec (weighted) &  Supervised (init.), Semi-Supervised & Replicate& In & 2023 & Both
-\\
-\hline
-Uzrad \textit{\textit{et al.}}~\cite{uzrad2024diffmoog} & DDSP & Synth. Chain and P-Loss & Supervised & Replicate & In \& Out & 2024 & 1-shot \\
-\hline
-Cherep \textit{et al.}~\cite{creativecherep2024} & DDSP & CLAP & Evolutionary & Imitate & Out & 2024 & Iter \\
-\hline
-\end{tabular}
-}
-
-\label{tab:summary}
-\end{table*}
-
-\section{Past Works}
-\label{sec:related_works}
-Table~\ref{tab:summary} contains an overview of relevant literature, covering important works in sound-matching.
-adsfEarlier works in this domain focused on FM synthesis and genetic algorithms (or direct optimization methods)\cite{justice1979analytic,horner1993machine,mitchell2007evolutionary}.
-Along with the exponential increase in computation power, later works utilized a wider range of approaches to parameter search, estimation, and audio synthesis; particularly with supervised learning and neural networks (\gls{NN})~\cite{yee2018automatic,engel2020ddsp,esling2019flow}.
-
-\subsection{Earlier Works}
-FM synthesizers are simple to implement, yet very expressive~\cite{chowning1973synthesis}.
-Perhaps the earliest foundational work in sound-matching is the analytical approach by Justice~\cite{justice1979analytic} toward the decomposition and recreation of sounds with FM synthesis using one carrier, one modulator, and the corresponding envelopes:
-\[ f(t) = I_c(t) \cos(\omega_c t + I_m(t) \cos(\omega_m t))
-\]
-Where $f(t)$ is the output signal, $I_c(t)$ is the carrier amplitude, $\omega_c$ is the carrier frequency, and $I_m(t)$ and $\omega_m$ are the envelope and frequency for the modulator~\cite{justice1979analytic}.
-
- The sound-matching works that followed used other heuristics for finding the synthesizer parameters, yet the simple FM synthesizer structure remained unchanged.
-Horner \textit{et al.}~\cite{horner1993machine} used GAs for re-synthesis of sounds with an FM synthesizer using one modulator and one carrier oscillator, and the McAulay-Quatieri method~\cite{mcaulay1986speech} for measuring loss.
-
- % MQ calculates the STFT and marks the start and end of \textit{trends} in a sound with the assumption that the harmonic content belongs to trends that have a beginning and an end.
-
-Introduction of more complex models of synthesis such as wavetables~\cite{horner2003auto} and physical modeling~\cite{riionheimo2003parameter} rendered interesting results, but led to further questions about the nature and goals of sound-matching.
-Mitchell and Creasy noted the difficulty in distinguishing between ``inefficiency of the optimization engine'' and ``synthesizer limitations'' as the cause of failure or success in sound-matching~\cite{mitchell2007evolutionary}.
-They offered a \textit{contrived methodology} for finding the best evolutionary method for sound-matching using their FM synthesizer.
-The methodology postulates that the best search heuristic for in-domain search (where an exact target exists, and a wide range of targets can be produced by sampling from different points in the parameter space) would also be the best-performing heuristic for out-of-domain search on a dataset of muted trumpet tones~\cite{opolko1989mcgill}; however, testing was not extensive and yielded some contradictory results.
-As they noted, modifications to the synthesizer, loss function, and sound domain would lead to a problem with an entirely new search space~\cite{mitchell2007evolutionary}.
-
-% Testing sound-matching experiments can be difficult, in part due to the large number of parametric components and navigation of complex non-linear systems.
-The issue with finding the best approaches to sound-matching is exacerbated when slight changes to the pipeline render previous results moot.
-Perhaps due to these reasons, the contrived methodology was not extensively tested by Mitchell and Creasy, who only provided a single experiment where the best evolutionary strategy found by in-domain search would also perform well in out-of-domain synthesis.
-\subsection{Introduction of Deep-Learning}
-Recent years have seen more works in sound-matching using machine learning techniques.
-In 2018, Yee-King \textit{et al.} rendered 60,000 audio-parameter pairs from the \textit{Dexed} \gls{VST} synthesizer~\footnote{https://asb2m10.github.io/dexed/}, and showed that NNs trained on this dataset can outperform GA and hill-climber (\gls{HC})~\cite{hoffmann2000heuristic} methods in rendering speed, with slight improvements in MFCC error (used as an objective performance test).
-The speed improvement appears trivial, considering the iterative nature of GAs when compared to offline training of supervised models.
-For GAs and HC optimizer, MFCCs were used as a measure of performance as well as a loss function.
-For training the networks, P-Loss was used, since differentiable MFCCs were not possible in their pipeline.
-Importantly, informal hearing tests revealed that the performance of even the best NN model was unsatisfactory, possibly due to the complex nature of the synthesizer, which features 155 parameters.
-
-% Programming synthesizers and querying sounds by imitating human vocals has also been a subject of interest~\cite{Cartwright2014SynthAssistQA,siamesevocalimitation2018}.
-For example, Cartwright and Pardo created a framework for searching pre-made patches for a synthesizer to best imitate vocals using audio features such as pitch, spectral centroid, and loudness~\cite{Cartwright2014SynthAssistQA}.
-
-Esling \textit{et al.} noted the issue of using P-Loss in sound-matching~\cite{esling2019flow}, considering the non-linear mapping between parameters of a synth and the audio output.
-In their approach, dubbed as ``flowsynth'', they used a large parameter-to-audio dataset for a commercial \gls{VST} synthesizer.
-In flowsynth, audio is encoded to an auditory space, using a NN autoencoding approach~\cite{hinton2006reducing}.
-The encodings in the auditory space can then be converted to the parameter space, but only for that particular synthesizer.
-Using P-Loss and Spectral Convergence (Frobenius norm of the spectrograms) as metrics, flowsynth outperformed other NN approaches that did not use an auditory to parameter space conversion, but formal hearing tests were not conducted.
-
-Masuda \textit{et al.} also applied supervised learning to synthesizer parameter estimation~\cite{masuda2021soundmatch}.
-Their work highlights the issue of non-linearity in parameter-to-synthesizer outputs and out-of-domain search.
-This work uses a differentiable subtractive synthesizer (two oscillators and an LP filter).
-A NN model was pre-trained using an in-domain dataset of randomly selected parameters and P-Loss.
-After training, the model was fine-tuned using 20,000 out-of-domain sounds from the NSynth dataset~\cite{engel2017neural} and multi-scale spectrogram loss~\cite{engel2020ddsp}.
-This approach proved more effective---i.e, achieved lower multi-level spectral difference in out-of-domain tests---than baseline models, which were either not exposed to out-of-domain sounds or trained exclusively with P-Loss.
-Subjective hearing tests were conducted, showing a preference for the fine-tuned model~\cite{masuda2021soundmatch}.
-Masuda \textit{et al.} later extended this work with semi-supervised learning, highlighting significant gaps in in-domain and out-of-domain performance~\cite{masuda2023improving}.
-
-Rather than focusing on a particular implementation, Shier \textit{et al.} presented Spieglib, a library for implementation of sound-matching pipelines~\cite{shier2020spiegelib}.
-This library provides different choices for DNNs, GAs, synthesizers, and feature extractors.
-Shier \textit{et al.} presented an experiment with a similar setup to Yee-King \textit{et al.}~\cite{yee2018automatic}, however, they found a genetic algorithm to be the best performing.
-
-% Rather than focusing on a particular implementation, Shier \textit{et al.} presented Spieglib, a library for implementation of sound-matching pipelines (which Shier \textit{et al.} referred to as ``automatic parameter search'')~\cite{shier2020spiegelib}.
-This library provides extensible choices of deep-learning estimators with \textit{tensorflow}~\cite{abadi2016tensorflow}, genetic estimators with \textit{DEAP}~\cite{DEAP_JMLR2012}, feature extractors with \textit{librosa}~\cite{mcfee2015librosa}, and different choices for synthesizers, evaluators, dataloaders.
-Shier \textit{et al.} provide an example experiment using this library.
-This example has a setup similar to Yee-King \textit{et al.}~\cite{yee2018automatic}, where the \textit{Dexed} VST was used and different genetic and NN estimators were compared by using MFCC difference as the minimization objective.
-Depending on the NN architecture, MFCC or STFT spectrograms were used as the input.
-Unlike previous comparisons of NN and genetic algorithms, the best estimator was the multi-objective non-dominated sorting genetic algorithm III (NSGA-III).
-This algorithm had multiple minimization objectives: 13-band MFCC, STFT, and five spectral features.
-Shier \textit{et al.} report that all models performed relatively well at imitation of harmonic distributions, but the multi-objective function was the only optimizer which did \textit{not} struggle with matching temporal envelopes of the audio spectrum~\cite{shier2020spiegelib}.
-
-Differentiable loss functions that use spectrogram differences can be computationally expensive.
-To mitigate this, Han \textit{et al.}~\cite{han2023perceptual} introduced ``perceptual-neural-physical loss'' (PNP).
-PNP is an approximation of loss functions; specifically, a loss function that uses the L2 norm of the difference between the features of two sounds, or $||\phi(t) - \phi(x)||^2_2$, where $\phi$ could be a spectrogram or JTFS function.
-PNP loss functions are fast and differentiable, but require training and parameter estimation.
-A Riemannian metric M needs to be calculated for the minimization of locally linear approximation of the ``true" spectral loss function~\cite{han2023perceptual}.
-\[
-\|\phi(t) - \phi(x)\|_2^2 = \langle \tilde{\theta} - \theta | \mathbf{M}(\theta) | \tilde{\theta} - \theta \rangle + O(\|\tilde{\theta} - \theta\|_2^3). \tag{4}
-\]
-
-During the training phase, this metric is calculated alongside the neural parameter estimator.
-After training, PNP can be used as a fast approximation of computationally demanding loss functions such as JTFS~\cite{han2023perceptual,han2024learning}, paired with FM synthesizers and a differentiable physical model of a drum head~\cite{smith2010physical}.
-
-Uzrad \textit{et al.}~\cite{uzrad2024diffmoog} took another unique approach to sound-matching: using a differentiable \textit{synthesis chain} of DSP generators and effects and a loss function that combines P-Loss with a \textit{signal-chain loss}.
-The synthesizer is a customizable chain of effects, which feed one output as input to the next step of the chain; signal-chain loss compares the parameter and output difference at every output step in the chain~\cite{uzrad2024diffmoog}.
-Possible chain functionalities are FM/AM, Low-Frequency Oscillators (\gls{LFO}), filters, and envelopes.
-Like the results shown by Masuda \textit{et al.}~\cite{masuda2021soundmatch}, better out-of-domain results were achieved when pre-trained on in-domain data and fine-tuned using out-of-domain NSynth data~\cite{engel2017neural}.
-
-% Uzrad \textit{et al.}\cite{uzrad2024diffmoog} used log-spectral distance (LSD) as their loss function, where   
-% $\text{LSD} = \| \log(S(t)) - \log(S(x)) \|_F$, S is the mel-spectrogram of the sounds and F is the Frobenius norm~\cite{golub2013matrix}. The Frobenius norm of a matrix $A$ is defined as
-% $ \|A\|_F = \sqrt{\sum_{i,j} |a_{ij}|^2}$.
-Here, $A$ would be the 2D matrix resulting from subtracting two log-spectrograms.
-Along with other works, \textbf{non-convexity of loss surfaces when applying gradient optimization} is a noted issue~\cite{turian2020sorry,vahidi2023mesostructures,uzrad2024diffmoog}.
-
-\subsection{What is Lacking In the Field}
-\label{sec:lacking}
-%   % First, we need to clearly define the characteristics we are looking for with non-generic loss functions, otherwise the chances of finding the sound(s) of interest is slim.
-Second, we need to move beyond direct replication, and focus on imitating target sounds that are outside the synthesizers capabilities.
-Third, more complex methods of synthesis need to be explored for iterative sound-matching.
-Fourth, we need to find optimization methods that can effectively navigate the the sinusoidal nature of loss function landscapes.
-
-% Based on the literature reviewed, we find several gaps in past research that makes the application of sound-matching for sound-designers difficult.
-Thus far, replication of in-domain sounds has been the main subject of research.
-With in-domain replication, we already have the target sound \textit{and} the program that makes it.
-This simplifies the problem of measuring performance and is the reason why this type of sound-matching research is by far the most common.
-However, practical application of sound-matching requires answers to much harder problems.
-Sound-matching would be an interesting creative tool if we could use it to either:
-% \begin{enumerate}
-%  \item \textbf{Replicate out-of-domain sounds}: this would give us the parameters that could closely approximate a target sound, and we can modify the sound further by modulating the synthesizer. 
-% \item \textbf{Imitate in-domain or out-of-domain sounds}: This would allow for applying the characteristics of one sound to another. 
-% \end{enumerate}
-
-% We need to be aware of the ``open set recognition'' (\gls{OSR}) problem here \cite{mundt2019open,gers2000learning}.
-OSR is a well documented issue in DL that manifests when the set or category that the network is supposed to reject or accept cannot be explained via examples.
-With imitation or replication of out-of-domain sounds, we can expect issues related to OSR to occur, since target and output sounds would be from different domains~\cite{salimi2021percussive}.
-
-% Perhaps one method of addressing the OSR problem is the use of representations (or feature extraction) methods that reduce sounds to a feature vector in the same domain.
-For example, a representation that only looks at the envelope of sounds is far less likely to be subject to the OSR problem.
-\textcolor{highlight}{We highlight the lack of non-generic loss functions as a major weakness in previous works and an interesting area to explore}.
-
-% Another weakness \textcolor{highlight}{is that the synthesis methods in past works are not representative of the common methods used by sound designers and modern synthesizers}.
-Simple FM/AM synthesizers have generally been the go-to methods of sound-matching.
-We believe that general statements regarding the effectiveness of an approach to iterative sound-matching (particularly in regards to a loss function) can only be reasonably made in the context of the subset of the tested methods of synthesis.
-% \textcolor{highlight}{The simplicity of parameter optimization approaches} is another area of weakness in past works.
-Classic genetic algorithms update the parameters randomly, while in recent differentiable settings, simple gradient descent methods are used.
-Past works have provided issues which arise in differentiable parameter estimation due to periodic loss functions.
-Perhaps the use of reinforcement learning (\gls{RL}), or other algorithms that continually learn from their environment, could help with this problem.
-
-% \subsection{Defining The Major Issues}
-% \label{sec:addressing_problems}
-
-Having looked at the current available literature, we note four major areas of weakness in sound-matching.
-We target the first two issues in this work, while the latter two issues are left for future work.
-\begin{enumerate}
-    \item \LossSelect: Is there a \textit{best performing} loss function, or is the selection of the loss function dependent on factors such as the sound domain, synthesizer capability and the desired characteristics of the output?
-    \item \SynthSelect: There is a lack of diversity in the synthesis methods used in sound-matching.
-The use of different DSP functions for iterative sound-matching and their effects on the outcome or the interactions with the loss function have not been tested.
-    \item \PeriodicLoss: The sinusoidal nature of loss function landscapes is a problem that appears frequently.
-In differentiable settings, this problem causes gradient descent updates to local minima.
-    \item \OutDomain: Sound-matching with out-of-domain sounds is an unexplored area, yet a necessary one for practical applications for sound designers.
-\end{enumerate}
-
-\section{Programs, Methodology, and Results}
-\label{sec:progs_methodology_results}
-Towards addressing the problems of \LossSelect{} and \SynthSelect{}, we compare the performance of iterative sound-matching experiments, where four differentiable loss functions are applied to four differentiable synthesizers.
-What we mean by ``performance'' is the similarity of the final output to the target, as measured by P-Loss, MSS, and---most importantly---manual listening tests.
-
-\textbf{We hypothesize that the relative performance of loss functions for sound-matching is influenced by the method of synthesis}.
-If true, this would strengthen the idea that the selection of the loss function and synthesis method is not an optimization problem, but a subjective choice depending on the needs of the sound designer.
-
-% In Section~\ref{sec:methodology} we lay out the experiment methodology.
-Section~\ref{sec:loss_implementation} discusses the details of implementation for the loss functions.
-Section~\ref{sec:programs} covers the characteristics of the programs, and finally Section~\ref{sec:results_posthoc} shows that the effect of loss function choice on the success of iterative sound-matching is dependent on the signal process chain of interest.
-
-\subsection{Experiment Methodology}
-\label{sec:methodology}
-% We have highlighted the lack of attention given to the selection of loss functions in sound-matching, and noted that the few existing works addressing this issue evaluate their proposed losses using only simplistic sound synthesis methods.
-\textcolor{highlight}{We hypothesize that the best performing loss function is influenced by the synthesis techniques utilized in the synthesizer}.
-This hypothesis targets the issues of \LossSelect~and \SynthSelect.
-
-We test four different loss functions and compare their success in sound-matching using four different synthesizers.
-We verify the success of the sound-matching process using two automatic measures of sound similarity (MSS and P-Loss) and a manual blinded hearing test conducted by two of the authors where Likert scores~\cite{jebb2021review} of 1-5 are assigned to a subset of final outputs and the corresponding targets.
-
-An \textit{experiment} is one iteration of sound-matching given a loss function, target sound,
-and synthesizer program.
-When the final iteration is complete, a score is given (automatically with an algorithm or with manual listening) to an experiment that measures its success in replicating the target sound.
-Consider the following simplified scenario: we have two loss functions, $L_1$ and $L_2$, and a single synthesizer program $P$ with a target parameter set $\theta^*$ where the target sound is $t$, or $P(\theta^*) = t$.
-We run a large number (in this case, 300 for each loss function) of experiments where we randomly set the initial ($\theta_0$) and target parameters ($\theta^*$) of $P$ and run an iterative sound-matching loop for each loss function, with maximum number of iterations set to 200.
-When the loop is terminated, we measure the similarity of the final output sound $x$ (where $P(\theta_{199}) = x$) to $t$, or if using P-Loss we compare $\theta_{199}$ to $\theta^*$.
-Having 300 experiments gives us a distribution of losses to determine whether $L_1$ or $L_2$ led to better performance.
-This simplified example leads us to our methodology which has four loss functions, four synthesizer programs, and three different evaluation methods of scoring sound similarity.
-
-\subsubsection{Automatic and Manual Evaluation}
-\label{sec:evaluation_manual_auto}
-The two automatic evaluation methods used here are P-Loss and MSS, as described in Section~\ref{sec:loss_funcs}.
-For P-Loss, the parameters are normalized between 0-1 based on the valid ranges defined in the Faust program, and the L1 distance between the normalized parameters is calculated.
-MSS is computed using a hop size of $100$ samples, and FFT window lengths of $(512, 1024, 2048, 4096)$.
-
-Automatic sound evaluation methods are often unreliable, as their alignment with the subjective judgments of human listeners is uncertain.
-Therefore, listening tests are conducted by randomly sampling 40 experiment results for each program and loss function pair.
-Two of the authors then assign blinded similarity scores to the outputs and targets using a 5-point Likert scale (1 = no similarity, 5 = near identical)~\cite{jebb2021review}.
-Using Spearman's rank correlation~\cite{spearman1987proof,rebekic2015pearson}, we found a very strong rate of agreement between the human listeners (see Section~\ref{sec:consistency_in_rankings}), a good indicator that the scores assigned by manual listeners can be treated as the ground truth.
-We then combine the ranks assigned by both authors, giving us 80 ranks per program and loss function pair.
-
-% Agreement between the assigned scores is measured using Spearman’s rank correlation, which provides both a correlation coefficient ($\rho$) ranging from –1 (perfect negative correlation) to 1 (perfect positive correlation), and a p-value testing the null hypothesis of no correlation~\cite{spearman1987proof,rebekic2015pearson}.
-Across all programs, the correlation was very strong ($\rho = 0.86$, $p < 10^{-180}$).
-Per-program correlations were also very strong: $\rho = 0.71$ for \BPNoise{} ($p < 10^{-25}$), $\rho = 0.64$ for \AddSineSaw{} ($p < 10^{-19}$), $\rho = 0.84$ for \AmpMod{} ($p < 10^{-43}$), and $\rho = 0.85$ for \FMMod{} ($p < 10^{-44}$).
-This high rate of agreement is a good indicator that the scores assigned by manual listeners can be treated as the ground truth.
-
-% For each program and loss function, the blind manual hearing test gives us 80 ranks, which are given by two of the authors to a set of 40 randomly selected experiment results.
-This gives 160 pairs of ranks per program, and 640 for all programs.
-We computed Spearman correlations between the two reviewers.
-
-\subsubsection{Determining Best Performers}
- Figure~\ref{fig:posthoc_evaluation} is a visualized summary of how the loss functions are ranked.
-We have four loss functions and four synthesizer programs.
-The maximum number of iterations set to 200, and 300 experiments for each loss/synthesizer combination.
-Each synthesizer program is paired with four loss functions and 300 experiments are conducted and evaluated automatically with P-Loss and MSS, and manually with Likert scores.
-
-For each program and loss function pair, we have three distributions that can be used to rank the loss functions.
-Two with 300 automatically assigned similarities (P-Loss and MSS) and one with 80 Likert scores (combining the 40 ranks assigned by each author).
-For consistency and statistical robustness, the distributions are upsampled to 1000 values using bootstrapping~\cite{tibshirani1993introduction}.
-Bootstrapping gives an estimation of the distribution for the mean performance of each experiment.
-$k$ (set to 1000) samples of $n$ values (set to 100\% of the values) are taken with replacement from the empirical distribution (list of 300 performance values for MSS and P-Loss or 80 for manual rankings).
-The mean performance is calculated for each of the $k$ samples.
-The $k$ estimates of the mean performance give us a bootstrapped distribution for each loss function.
-
-For each program, the loss functions are ranked using the non-parametric Scott-Knott (\gls{NPSK}) test~\cite{tantithamthavorn2017mvt,tantithamthavorn2018optimization}, which ranks the loss functions from rank 1 (best, or highest distribution values) to maximum rank of 4 (worst, with lowest distribution values).
-NPSK may determine that multiple functions belong to the same rank, in which case the maximum rank would be less than 4.
-As mentioned before, lower P-Loss and MSS values indicate higher similarity, therefore (for consistency with manually assigned ranks) the reciprocal of their output is used ($x^{-1}$, if $x$ is the P-Loss or MSS value), where higher values indicate greater similarity.
-
-\input{extra_figures/evaluation_graph.tex}
-
-\subsubsection{Training Parameters and Gradient Calculations}
- % Previous work utilized gradient fields for analysis of the algorithm performance~\cite{vahidi2023mesostructures}.
-Gradient fields are not the center of our analysis, but we include them in Appendix~\ref{appendix:gradient_fields}.
- For updating the synthesizer parameters (or weights), we use RMSProp, which operates similar to stochastic gradient descent (SGD)~\cite{goodfellow2016deep}, with the caveat that the gradients of each weight are scaled by the root-mean-square of past gradients of that weight.
-Based on some initial test runs, we used an arbitrary fixed learning rate of 0.045 for all experiments (learning rates used in Vahidi \textit{et al.}~\cite{vahidi2023mesostructures} is unknown).
-The maximum number of iterations is set to 200, where, based on our observations, the parameters have either irrecoverably diverged outside the acceptable ranges, or are stuck at a local minimum.
-Additionally, the gradients are large, and calculated backwards sequentially through the signal processing chain.
-This backward calculation resembles the \textit{exploding gradients} problem in recurrent neural networks~\cite{gers2000learning}.
-Gradient clipping~\cite{goodfellow2016deep} is preemptively used to ensure that the  $\ell_2$ norm of all gradients does not exceed the threshold of 1.
-% \mathbf{\gamma}_i^{\text{clipped}} = \mathbf{\gamma}_i \cdot \mathcal{S}
-% \]
-% Where $\mathcal{S}$ is a value less than or equal to 1, depending on whether the norm of the gradients exceeds the threshold $c$.  
-% \[
-% \mathcal{S} = \frac{c}{\max(\|\mathbf{\gamma}\|_2, c)}
-% \]
-
-\subsection{Loss Function Implementation Details}
-\label{sec:loss_implementation}
-\subsubsection{STFT Losses}
-Due to their ubiquity and lower cost of gradient calculation, we use STFT as the basis of two loss functions.
-As discussed in Section~\ref{sec:matching_types}, the L1 or L2 distances are the most common methods of comparing different spectrograms, however, there are situations where this might lead to shortcomings.
-An example given by Vahidi \textit{et al.}~\cite{vahidi2023mesostructures} is two chirplets (tones that increase in pitch) starting and ending at the same frequencies.
-If one of these tones is slightly shifted in time, then the spectrograms will no longer overlap at any point, despite their sonic similarity (similar to the idea that a linear equation $f(x) = ax$ can never overlap with a slightly shifted version of itself, or $f(x) = ax +\epsilon$).
-In this example, we do not know if this is an issue of the spectrograms, or the L1 measure used for their comparison.
-In fields such as computer vision, Structural Similarity Index (SSIM)~\cite{wang2004imagesssim,wang2009mean} and Scale-Invariant Mean Squared Error (SIMSE)~\cite{barron2014shapessimse} have been used as alternatives that improved performance in various tasks relative to L1.
-
-We define the \LoneSpec~and \SIMSESpec~functions as the application of L1 and SIMSE to the STFT spectrograms.
-The STFT spectrogram function uses 512 FFT bins, window size of 600 samples, and hop length (how many samples the window shifts) of 100 samples.
-The L1 and SIMSE implementations are differentiable.
-
-\subsubsection{JTFS Loss}
-The \JTFS~loss function is the application of L1 difference to the JTFS representations of two sounds~\cite{vahidi2023mesostructures}.
-The code used for the JTFS transformation is the differentiable implementation of a 1-dimensional JTFS function provided by Andreux \textit{et al.}~\cite{kymatio}.
-
-\subsubsection{Soft-DTW Loss}
-We use the soft-DTW function, which is differentiable and --- depending on its parameters --- not shift-invariant~\cite{cuturi2017soft,janati2020spatio,tavenard.blog.softdtw}.
-The loss function \DTWEnv~is the application of the soft-DTW function to the envelope of the two sounds being compared~\cite{lyons1997understanding}.
-The envelope is calculated by creating the STFT spectrogram of a sound (the same process used in Section~\ref{sec:fourier_specs}) and summing the values at each timestep.
-
-\subsection{The Synthesizers}
-\label{sec:programs}
-The synthesizer programs are meant to be simple examples that test the building blocks of digital sound synthesis.
-Subtractive, additive, and FM/AM synthesis are three of the most common techniques in sound design~\cite{smith1991viewpoints}.
-In subtractive synthesis, frequencies are removed from a sound via digital filters.
-In additive synthesis, complex sounds are created via the linear combination of simpler sounds~\cite{lyons1997understanding,smith2007introduction}.
-As discussed previously, FM/AM synthesis refers to the general technique of modulating the frequency or amplitude of a waveform (the carrier) by another waveform (the modulator).
-
-For each program, we provide the Faust code~\cite{orlarey2009faust}, which can be run in the online IDE~\footnote{\url{https://faustide.grame.fr/}}.
-Faust is a functional language for audio synthesis that can succinctly define signal processing chains.
-Following Braun's methodology~\cite{braun2024dac}, programs are first defined in Faust, then converted (or \textit{transpiled}) to differentiable Jax functions using the DawDreamer library\footnote{\url{https://github.com/DBraun/DawDreamer}}.
-
- Once the programs are transpiled to differentiable synthesizers, the iterative sound-matching procedure is:
- \begin{enumerate}
-    \item Initialize $\theta^*$ and $\hat{\theta}$, i.e., random generation of target and initial parameters uniformly over a predefined range
-    \item Generating the output of the synthesizer with $\hat{\theta}$ (the length of the output is set to 1 second with sample-rate of 44100 Hz) 
-    \item Calculating the loss between the target and output
-    \item Applying gradient updates to the synthesizer parameters
-    \item Repeating the second step with the updated parameters $\hat{\theta}$, until maximum number of iterations has been reached
- \end{enumerate}
-
-% A general overview of the programs and the predefined parameter ranges is given in Table~\ref{tab:programs}.
-
-In the following subsections, we describe the programs and their parameters in more detail, followed by the methodology in Section~\ref{sec:methodology} and results in Section~\ref{sec:results_posthoc}.
-
-% \begin{table*}[htbp]
-% \centering
-% \begin{adjustbox}{max width=\textwidth}
-% \begin{tabular}{|c|c|c|c|l|}
-% \hline
-% \textbf{ID} & \textbf{Parameters} & \makecell{\textbf{Param. 1}\\\textbf{Range}} & \makecell{\textbf{Param. 2}\\\textbf{Range}} & \textbf{Description} \\
-% \hline
-% \BPNoise & hp\_cut, lp\_cut & 50-1000 & 1-120 & Band-pass of Noise signal \\
-% \hline
-% \AddSineSaw & saw\_freq, sine\_freq & 20-1000 & 20-1000 & Addition of sine and saw waves \\
-% \hline
-% \AmpMod & amp, carrier & 1-20 & 20-1000 & LFO on amplitude of noise osc \\
-% \hline
-% \FMMod & amp, carrier & 0-5 & 0-4 & LFO on amplitude of saw osc \\
-% \hline
-% \end{tabular}
-% \end{adjustbox}
-% \caption{Overview of Program Parameters and the range of values each parameter can take.}
-% \label{tab:programs}
-% \end{table*}
-
-\begin{lstlisting}[caption={\BPNoise}, label={lst:program0}, language=Faust,
-                  float, floatplacement=!H, xleftmargin=1em, xrightmargin=0.5em, firstnumber=0, aboveskip=0em, belowskip=-1em]
-import("stdfaust.lib");
-lp_cut = hslider("lp_cut",900,50,1000,1);
-hp_cut = hslider("hp_cut",100,1,120,1);
-process = no.noise:fi.lowpass(3,lp_cut):fi.highpass(10,hp_cut);
-\end{lstlisting}
-
-\begin{lstlisting}[caption={\AddSineSaw}, label={lst:program1},language=Faust,float,floatplacement=!H,xleftmargin=1em,xrightmargin=0.5em,firstnumber=0,aboveskip=0em, belowskip=-1em]
-import("stdfaust.lib");
-saw_freq = hslider("saw_freq",800,20,1000,1);
-sine_freq = hslider("sine_freq",300,20,1000,1);
-sineOsc(f) = +(f/ma.SR) ~ ma.frac:*(2*ma.PI) : sin;
-sawOsc(f) = +(f/ma.SR) ~ ma.frac;
-process = sineOsc(sine_freq)+sawOsc(saw_freq);
-\end{lstlisting}
-
-\begin{lstlisting}[caption={\AmpMod}, label={lst:program2},language=Faust,float,floatplacement=!H,xleftmargin=1em,xrightmargin=0.5em,firstnumber=0,aboveskip=0em, belowskip=-1em]
-import("stdfaust.lib");
-amp = hslider("amp",0.5,0,5,0.01);
-modulator = hslider("modulator",0.5,0,4,0.01);
-sineOsc(f) = +(f/ma.SR) ~ ma.frac:*(2*ma.PI) : sin;
-process = no.noise*sineOsc(modulator)*amp;
-\end{lstlisting}
-
-\begin{lstlisting}[caption={\FMMod}, label={lst:program3},language=Faust,float,floatplacement=!H,xleftmargin=1em,xrightmargin=0.5em,firstnumber=0,aboveskip=0em, belowskip=-1em]
-import("stdfaust.lib");
-carrier = hslider("carrier",100,20,1000,1);
-amp = hslider("amp",6,1,20,1);
-sineOsc(f) = +(f/ma.SR) ~ ma.frac:*(2*ma.PI) : sin;
-sawOsc(f) = +(f/ma.SR) ~ ma.frac;
-process = sineOsc(amp)*sawOsc(carrier);
-\end{lstlisting}
-
-\subsubsection{\BPNoise}
-\label{sec:program0}
-\BPNoise{} is a bare bones example of subtractive synthesis using digital filters~\cite{smith2007introduction}.
-Despite the ubiquity of subtractive synthesis in practical sound design, it has rarely been tested in differentiable sound-matching.
-In this program, a noise signal is fed through a band-pass (BP) filter.
-This removes the frequencies outside the low-pass (LP) and high-pass (HP) cutoffs.
-The noise generator produces all frequencies with random variation over time.
-The LP filter removes frequencies over its threshold frequency, and the HP removes frequencies lower than its threshold~\cite{smith2007introduction}.
-The search parameters are the cutoff thresholds for the LP and HP filters.
-Listing~\ref{lst:program0} shows the Faust code for \BPNoise.
-
-\subsubsection{\AddSineSaw}
-\label{sec:program1}
-\AddSineSaw{} is an additive program that combines a saw and a sine wave function.
-Like subtractive synthesis, additive synthesis is a common approach to sound design that has not been extensively tested as a benchmark in sound-matching.
-The search parameters here are the frequency of the sine and saw oscillators.
-Listing~\ref{lst:program1} is the Faust code for \AddSineSaw.
-
-\subsubsection{\AmpMod}
-\label{sec:program2}
-\AmpMod{} involves amplitude modulation of a noise generator by an LFO with \textit{modulator} as its frequency and a global \textit{amp} value that does not change over time and applies to the entire signal.
-The search parameters for this program are the LFO frequency and the amp value.
-The amp value acts as a global volume control, and since the sounds are normalized before analysis, it is likely inconsequential to the performance.
-This program is meant to be a stepping stone to \FMMod, which utilizes proper AM synthesis.
-Listing~\ref{lst:program2} is the Faust code for \AmpMod.
-
-\subsubsection{\FMMod}
-\label{sec:program3}
-Relative to additive and subtractive synthesis, AM/FM synthesis are less common methods of sound design.
-Due to their frequent use in previous works, we also tested an AM synthesizer.
-The synthesizer program is the multiplication of a low frequency sine oscillator with frequency parameter \textit{amp}, and a saw oscillator with frequency parameter \textit{carrier}.
-Listing~\ref{lst:program3} is the Faust code for \FMMod.
-
-\subsection{Post hoc Analysis of Loss Performance}
-\label{sec:results_posthoc}
-As summarized in Figure~\ref{fig:posthoc_evaluation}, for each program and evaluation (or scoring) method, there are four sets of 1000 bootstrapped values, corresponding to the four loss functions.
-A post-hoc test is conducted in two stages in order to determine whether there is a difference in performance between loss functions for each program, and if so, which are the best performers.
-The first stage determines whether there is a difference in group means, with the null hypothesis being that all groups have similar mean ranks.
-The second stage ranks the loss functions from best to worst using the non-parametric Scott-Knott test (\gls{NPSK})~\cite{tantithamthavorn2017mvt,tantithamthavorn2018optimization}.
-The Kruskal-Wallis test is used for the first stage~\cite{kruskal1952use}; this test pools and ranks all evaluation measures for a program, then tests whether the differences in mean rank for all loss function groups is zero.
-The Kruskal-Wallis test calculates an H statistic that is compared to a chi-square distribution with k-1 degrees of freedom (k is the number of groups, or 4, and degrees of freedom is 3).
-Using the significance level of 0.05, if the H statistic is greater than the critical value of the chi-square distribution then the null hypothesis is rejected, meaning that at least one group has a mean rank significantly above others~\cite{kruskal1952use}.
-Table~\ref{tab:kruskal_auto} shows the Kruskal-Wallis results for every program and evaluation method, and shows that significant differences between loss function performance measures (i.e., the bootstrapped distributions of the evaluation for each loss function) exist in all cases except MSS for the \AmpMod{} program.
-
-\begin{table}[ht]
-\centering
-\caption{Do loss functions have different median performance? Kruskal-Wallis results by program and bootstrapped evaluation results.}
-\begin{tabular}{|c|c|c|c|c|}
-\hline
-\textbf{Program} & \textbf{Eval. Method} & \textbf{H-Stat.} & \textbf{P-value} & \textbf{Reject} \\
-\hline
-\BPNoise & MSS      & 81.19  & $1.70 \times 10^{-17}$ & Yes \\
-\BPNoise & P-Loss  & 163.60 & $3.07 \times 10^{-35}$ & Yes \\
-\BPNoise & Manual & 9.45  & $2.39 \times 10^{-2}$ & Yes \\
-\AddSineSaw & MSS      & 308.97 & $1.14 \times 10^{-66}$ & Yes \\
-\AddSineSaw & P-Loss  & 348.42 & $3.28 \times 10^{-75}$ & Yes \\
-\AddSineSaw & Manual & 9.45  & $2.39 \times 10^{-2}$ & Yes \\
-\AmpMod & MSS      & 1.35   & $0.7171$               & No \\
-\AmpMod & P-Loss  & 366.76 & $3.50 \times 10^{-79}$ & Yes \\
-\AmpMod & Manual & 32.71 & $3.70 \times 10^{-7}$ & Yes \\
-\FMMod & MSS      & 564.65 & $4.65 \times 10^{-122}$ & Yes \\
-\FMMod & P-Loss  & 229.19 & $2.07 \times 10^{-49}$ & Yes \\
-\FMMod & Manual &  207.58 & $9.69 \times 10^{-45}$ & Yes \\
-\hline
-\end{tabular}
-\label{tab:kruskal_auto}
-\end{table}
-
-% \begin{table}[ht]
-% \centering
-% \begin{tabular}{|c|c|c|c|}
-% \hline
-% \textbf{Program} & \textbf{H-Stat.} & \textbf{P-value} & \textbf{Reject} \\
-% \hline
-% \BPNoise & 9.45  & $2.39 \times 10^{-2}$ & Yes \\
-% \AddSineSaw & 9.45  & $2.39 \times 10^{-2}$ & Yes \\
-% \AmpMod & 32.71 & $3.70 \times 10^{-7}$ & Yes \\
-% \FMMod & 207.58 & $9.69 \times 10^{-45}$ & Yes \\
-% \hline
-% \end{tabular}
-% \caption{Manual ranking Kruskal-Wallis results. Significant differences are observed for all programs.}
-% \label{tab:kruskal_manual}
-% \end{table}
-
-\subsubsection{Best Performers For Each Program}
-Based on the bootstrapped distribution of scores, the NPSK algorithm ranks the loss functions from 1 (best) to a maximum of 4 (worst).
-In cases where the distributions are similar, multiple loss functions can be clustered into the same rank.
-In Figure~\ref{fig:npsk_all}, we use color-coded violin plots to visualize the best performers for each program.
-The corresponding colors for each rank are
-\colorbox{rank1}{\textcolor{black}{\textbf{1}}} \colorbox{rank2}{\textcolor{white}{\textbf{2}}} \colorbox{rank3}{\textcolor{white}{\textbf{3}}} \colorbox{rank4}{\textcolor{black}{\textcolor{white}{\textbf{4}}}}
-
-\input{extra_figures/ranking_tables}
-\input{extra_figures/npsk_figures}
-
-\subsubsection{\BPNoise}
-For this synthesizer program, the spectrogram-based models performed the best.
-This makes intuitive sense, as the visual effects of a band-pass filter on white noise are readily apparent in a spectrogram.
-As shown in Figure~\ref{fig:npsk_p0}, manual hearing test and MSS selected \SIMSESpec~as the best performer and \LoneSpec~as the second-best performer, while P-Loss gave the reverse order.
-
-\subsubsection{\AddSineSaw}
-As shown in Figure~\ref{fig:npsk_p1}, JTFS was the best performing loss function in all evaluation methods. Moreover, all evaluation methods produced identical results.
-
-\subsubsection{\AmpMod}
-IP-Loss and manual hearing tests yield identical rankings, and MSS results vary for for ranks 2 to 4.
-It is not surprising that MSS results differ from the hearing tests, as we saw in Table~\ref{tab:kruskal_auto}, MSS showed no significant differences between groups.
-As shown in Figure~\ref{fig:npsk_p2}, all models selected \DTWEnv~as the best performer.
-This may be due to \DTWEnv's focus on periodic changes in loudness.
-
-\subsubsection{\FMMod}
-\DTWEnv~is again the best performer here. As shown in Figure~\ref{fig:npsk_p3}, all methods of analysis gave identical rankings to all programs.
-
-\subsection{Consistency In Rankings}
-\label{sec:consistency_in_rankings}
-\subsubsection{Manual Ranks} Agreement between the manually assigned scores is measured using Spearman’s rank correlation, which provides both a correlation coefficient ($\rho$) ranging from –1 (perfect negative correlation) to 1 (perfect positive correlation), and a p-value testing the null hypothesis of no correlation~\cite{spearman1987proof,rebekic2015pearson}.
-Across all programs, the correlation was very strong ($\rho = 0.86$, $p < 10^{-180}$).
-Per-program correlations were also very strong: $\rho = 0.71$ for \BPNoise{} ($p < 10^{-25}$), $\rho = 0.64$ for \AddSineSaw{} ($p < 10^{-19}$), $\rho = 0.84$ for \AmpMod{} ($p < 10^{-43}$), and $\rho = 0.85$ for \FMMod{} ($p < 10^{-44}$).
-
-Table~\ref{tab:combined_ranks} shows the SNPK rankings of bootstrapped evaluation results for each program.
-Both MSS and P-Loss gave a different rank from the hearing results in 3 out of 16 cases, which shows consistency between automatic and manual hearing tests, at least for the simple programs used in this work.
-We also observe that top ranks were consistent across performance evaluation methodologies, with the exception of P-Loss in \BPNoise, which narrowly picks a different spectrogram-based loss function.
-% \input{extra_figures/ranking_tables}
-
-\subsection{Takeaways}
-An important takeaway is that the loss function that yields the best sound-matching outcomes varies depending on the synthesizer program.
-This is perhaps due to the interaction between how the parameters of the synthesizer influence the sound, and the core sonic features of the loss function.
-
-We also see that the \DTWEnv{} and \SIMSESpec{} (to our knowledge, not previously used in iterative sound-matching), lead to the best results in three of the four programs.
-This shows room for further creative approaches to differentiable sound-similarity measures, particularly since we observed that SOTA loss functions can be ineffective depending on the design of a synthesizer.
-
-Finally, we surprisingly see somewhat consistent results between the rankings assigned by manual hearing tests (which we take as the ground truth), and automatic measures of P-Loss and MSS.
-Due to the simplicity of our programs and the occasional deviations, we cannot conclude that these automatic measures are substitutes for human hearing tests.
-
-\section{Summary, Weaknesses, and Conclusion}
-\label{sec:summary_conclusion}
-\textbf{Summary:} Sound-matching is an umbrella term for the algorithmic programming of audio synthesizers, often with the goal of assisting sound designers.
-Here we provided a history of sound-matching, major issues in the field, and discussed the importance of ``differentiable iterative sound-matching'' as a natural extension of current literature.
-
-The main hypothesis tested here is whether the performance of differentiable loss functions is influenced by the synthesis techniques used for sound-matching.
-We created iterative sound-matching experiments by combining four different loss functions with four different sound synthesis programs.
-We ranked the performance of the iterative sound-matching pipelines for every loss function and program, and observed that the success of the pipeline (that is, how closely the output sound matches the target sound) is program dependent.
-In other words, different synthesizer programs work best with different loss functions.
-Notably, we see that our novel use of DTW and SIMSE based differentiable loss functions (\DTWEnv{} and \SIMSESpec) can outperform what are regarded as the SOTA loss functions in 3 of 4 cases, although this is highly synthesizer dependent.
-
-P-Loss and MSS have frequently been used as automatic performance measures, yet their ``preference'' has rarely been compared to human rankings.
-We observed that automatic performance measures and manual listening tests were often in agreement, despite this, manual verification of sound-matching results remains a necessity for future works.
-
-\textbf{Weaknesses:} 
-While we cannot prove that a universally best similarity measure does not exist, we can advocate for more creativity in the field based on our findings.
-Compared to previous work, we presented a more cohesive approach to iterative sound-matching which utilizes a variety of loss functions and synthesis methods.
-However, there are many other methods of synthesis and sound-similarity that can be combined in practically infinite ways.
-Due to this large search space, we set arbitrary parameters for the various signal processing functions.
-We used bare-bones versions of STFT and JTFS with fixed parameters.
-We did not test complex synthesizers using parallel and sequential DSP functions.
-Arbitrary hyperparameters such as learning rate and max number of iterations were selected for the DL pipeline, and only the RMSProp optimizer was tested.
-
- Like the majority of previous works, this work utilizes in-domain sounds; that is, the target sound is made by the synthesizer, and the parameters are already known.
-This simplifies the issues of measuring sound-similarity, but it is not a realistic scenario for practical sound-matching.
-This problem is left for future work, discussed in the next section.
-
-\label{sec:future}
-\textbf{Future Work: } The problem of periodic loss-landscapes, noted by many previous works, remains unaddressed~\cite{turian2020sorry,vahidi2023mesostructures,uzrad2024diffmoog,bruford2024synthesizer}.
-Perhaps this problem emerges due to the periodic nature of sound, which requires better loss landscape navigation methods.
-An optimizer that is more aware of fluctuations in the gradients would perhaps lead to better solutions than simple gradient descent.
-Viewing the sound-matching problem as ``the navigation of an agent from an arbitrary point in a gradient field to a target'' closely resembles many classical problems in the field of reinforcement learning (RL)~\cite{sutton2018reinforcement}.
-Naturally, the application of RL and other heuristic search techniques to the problem of iterative sound-matching would be an important contribution.
-
-Contemporary works often involve the application of domain specific and computationally expensive loss functions~\cite{han2023perceptual,uzrad2024diffmoog}, use of large neural networks~\cite{hershey2017cnn,cramer2019look}, or complex ensemble methods~\cite{turian2022hear}.
-Such models are useful but intractable; furthermore, training them requires the definition of simpler loss functions, which emphasizes the need for further development of differentiable and expressive loss function implementations.
-
-Like nearly all previous work in sound-matching, the main measure of success here was the accurate \textit{replication} of sounds (or synth parameters) rather than imitation of sounds outside of the synthesizer's domain.
-Replication of arbitrary features of sounds is an important component of sound-design, though it is much harder to define or measure.
-Future works could explore loss functions which only measure certain characteristics of sound (such as \DTWEnv), and whether they can pave the way for better imitation in sound-matching.
-
-\bibliographystyle{IEEEtran}
-\bibliography{references}
 \end{document}

### 2025-10-27 sections/background_related.tex
final files

@@
     \item $g(\theta)$: A parametric audio synthesizer $g$ with parameters $\theta$.
-    \item $x_\theta$: The synthesizer output, $x_\theta = g(\theta)$.”.
+    \item $x_\theta$: The synthesizer output, $x_\theta = g(\theta)$”.
     \item $\xzero$: The target sound to be replicated or imitated. 
@@
 
-In sound-matching, particularly in a Deep Learning (\gls{DL}) context~\cite{goodfellow2016deep}, the similarity function can also be called a \textit{loss} function, where the emphasis is on measurement and reduction of the distance between target and output.
+In sound-matching, particularly in a deep learning (\gls{DL}) context~\cite{goodfellow2016deep}, the similarity function can also be called a \textit{loss} function, where the emphasis is on measurement and reduction of the distance between target and output.
 It is important to note that there is a close relationship between the loss function $L$ and the sound representation function $\phi$.
@@
 \label{sec:in-domain}
-The choice of domain depends on whether we want to use the same synthesizer for the target and output sounds, the scenario that is called \textit{in-domain}, or have target sounds that came from sources other than the synthesizer, or \textit{out-of-domain}.
+The choice of domain depends on whether we want to use the same synthesizer for the target and output sounds, known as the \textit{in-domain} scenario, or have target sounds that came from sources other than the synthesizer, known as the \textit{out-of-domain} scenario.
 To paraphrase the description given by Masuda \textit{et al.}~\cite{masuda2021soundmatch}, if $g$, the synthesizer of choice, can accurately replicate the target sound $\xzero$, or put differently, if $\xzero$ itself is an output of $g$, then the sound-matching task is \textit{in-domain}.
@@
 In-domain tasks in general are simpler, and often there is a guarantee that there is a correct answer to the sound-matching problem, particularly if the goal is accurate replication of the sound.
-If the target sound is out-of-domain, replication is not guaranteed, and the goal becomes the \textit{imitation} of some aspect of sound.
+If the target sound is out-of-domain, replication is not guaranteed, and the goal becomes the \textit{imitation} of some aspects of sound.
 
@@
 
-The choice of \textit{heuristics} is yet another important attribute sound-matching.
+The choice of \textit{heuristics} is yet another important attribute of sound-matching.
 The goal of sound-matching is to find the optimal parameters $\theta^*$ that minimize the loss between the synthesizer output and the target sound.
@@
 \caption{Summary of select works in sound-matching.
-Gen.
-refers to whether the generations are done in 1-shot (generally, output by a neural network) or a search is conducted with the goal of iteratively getting closer to the target sound.}
+The \textbf{Gen.} column shows whether the generations are done in 1-shot (generally, output by a neural network) or a search is conducted with the goal of iteratively getting closer to the target sound.}
 \setlength{\tabcolsep}{3pt} % Reduce horizontal spacing
@@
 \hline
-Mitchel \textit{\textit{et al.}}~\cite{mitchell2007evolutionary} & Non-differentiable FM & Spectral Relative & Evolutionary & Replicate & In (Contrived) & 2007 & Iter \\
+Mitchell and Creasy~\cite{mitchell2007evolutionary} & Non-differentiable FM & Spectral Relative & Evolutionary & Replicate & In (Contrived) & 2007 & Iter \\
 \hline
@@
 This library provides different choices for DNNs, GAs, synthesizers, and feature extractors.
-Shier \textit{et al.} presented an experiment with a similar setup to Yee-King \textit{et al.}~\cite{yee2018automatic}, however, they found a genetic algorithm as the best performing.
+Shier \textit{et al.} presented an experiment with a similar setup to Yee-King \textit{et al.}~\cite{yee2018automatic}, however, they found a genetic algorithm as the best-performing.
 
@@
 
-\subsection{What is Lacking In the Field}
+\subsection{What Is Lacking In the Field}
 \label{sec:lacking}

### 2025-10-27 sections/introduction.tex
final files

@@
 A digital audio synthesizer is any software used for the creation and manipulation of audio.
-The possibilities afforded by these synthesizers are infinite, leading to their widespread adoption by artists and sound-designers~\cite{lyons1997understanding,russ1999sound,stranneby2004digital}.
+The possibilities afforded by these synthesizers are infinite, leading to their widespread adoption by artists and sound designers~\cite{lyons1997understanding,russ1999sound,stranneby2004digital}.
 A typical synthesizer has a number of parameterizable functions which affect the sound output in various ways, and manual sound-design often involves modifying the parameters until a conceptualized target sound is reached.

### 2025-10-27 sections/methodology_and_analysis.tex
final files

@@
 \label{sec:program0}
-\BPNoise{} is a bare bones example of subtractive synthesis using digital filters~\cite{smith2007introduction}.
+\BPNoise{} is a bare-bones example of subtractive synthesis using digital filters~\cite{smith2007introduction}.
 Despite the ubiquity of subtractive synthesis in practical sound design, it has rarely been tested in differentiable sound-matching.
@@
 This makes intuitive sense, as the visual effects of a band-pass filter on white noise are readily apparent in a spectrogram.
-As shown in Figure~\ref{fig:npsk_p0}, manual hearing test and MSS selected \SIMSESpec~as the best performer and \LoneSpec~as the second-best performer, while P-Loss gave the reverse order.
+As shown in Figure~\ref{fig:npsk_p0}, manual hearing tests and MSS selected \SIMSESpec~as the best performer and \LoneSpec~as the second-best performer, while P-Loss gave the reverse order.
 

### 2025-10-27 sections/summary_weakness_conclusion.tex
final files

@@
 \\We see somewhat consistent results between the rankings assigned by manual hearing tests (which we take as the ground truth), and automatic measures of P-Loss and MSS.
-With the exception of P-Loss in \BPNoise{} narrowly selecting a different spectrogram based loss, all measures of performance selected the same top performer.
+With the exception of P-Loss in \BPNoise{} narrowly selecting a different spectrogram-based loss, all measures of performance selected the same top performer.
 However, given the simplicity of our programs and the occasional deviations, we cannot conclude that these automatic measures are substitutes for human hearing tests.
@@
 We show the landscapes of the loss functions with simplified versions of \BPNoise{} (HP-Noise) and \AmpMod{} (S-Noise-AM), each reduced to a single parameter for one-dimensional visualization.
-HP-Noise uses only a high-pass filter with cut-off between 100–20,000 Hz, while S-Noise-AM varies only the modulation rate between 0.1–20 Hz.
-Loss values are normalized to 0–1 to allow comparison across functions.
-
+HP-Noise uses only a high-pass filter with cut-off between 100--20,000 Hz, while S-Noise-AM varies only the modulation rate between 0.1--20 Hz.
+To simplify comparisons, loss values and parameter ranges are respectively normalized to 0--1 and -1--1.
 Figure~\ref{fig:loss_landscape_noisebp} shows the HP-Noise results.
@@
 Replication of arbitrary features of sounds is an important component of sound-design, though it is much harder to define or measure.
-Future works could explore loss functions which only measure certain characteristics of sound (such as \DTWEnv), and whether they can pave the way for better imitation in sound-matching.
+Future work could explore loss functions which only measure certain characteristics of sound (such as \DTWEnv), and whether they can pave the way for better imitation in sound-matching.

### 2025-11-24 sections/background_related.tex
missed a hat for x in iterative loop figures
Updates from Overleaf

@@
     \item $g(\theta)$: A parametric audio synthesizer $g$ with parameters $\theta$.
-    \item $x_\theta$: The synthesizer output, $x_\theta = g(\theta)$”.
+    \item $x_\theta$: The synthesizer output, $x_\theta = g(\theta)$.
     \item $\xzero$: The target sound to be replicated or imitated. 
@@
 % Nodes
-\node (start) [text centered] {\( g(\hat{\theta}) = x_{\theta} \)};
+\node (start) [text centered] {\( g(\hat{\theta}) = x_{\hat{\theta}} \)};
 \node (L) [above of=start, right of=start, text centered] 
-    {\( L(\hat{\theta}, x_{0}) \propto \phi(x_{\theta}) - \phi(x_{0}) \)};
+    {\( L(\hat{\theta}, x_{0}) \propto \phi(x_{\hat{\theta}}) - \phi(x_{0}) \)};
 \node (optimize) [below of=L, right of=L, text centered] {Optimize $\hat{\theta}$};
@@
 \subsection{Historical Framing of Sound-Matching}
-Given the attributes discussed above, Table~\ref{tab:summary} summarizes relevant literature. Here we expand with a historical analysis of past sound-matching work.
+Table~\ref{tab:summary} summarizes relevant literature. Following the discussion of important sound-matching components, a historical analysis of past work can be given.
 
@@
 
-Uzrad \textit{et al.}~\cite{uzrad2024diffmoog} took another unique approach to sound-matching: using a differentiable \textit{synthesis chain} of DSP generators and effects and a loss function that combines P-Loss with a \textit{signal-chain loss}.
-The synthesizer is a customizable chain of effects, which feed one output as input to the next step of the chain; signal-chain loss compares the parameter and output difference at every output step in the chain~\cite{uzrad2024diffmoog}.
-Possible chain functionalities are FM/AM, Low-Frequency Oscillators (\gls{LFO}), filters, and envelopes.
+Uzrad \textit{et al.}~\cite{uzrad2024diffmoog} took another unique approach to sound-matching: using a differentiable \textit{synthesis chain} of DSP generators and effects along with a loss function that combines P-Loss with a \textit{signal-chain loss}.
+The synthesizer is a customizable chain of effects, which feeds one output as input to the next step of the chain; signal-chain loss compares the differences in parameter and output at every output step in the chain~\cite{uzrad2024diffmoog}.
+Possible chain functionalities are FM/AM, low-frequency oscillators (\gls{LFOs}), filters, and envelopes.
 Like the results shown by Masuda \textit{et al.}~\cite{masuda2021soundmatch}, better out-of-domain results were achieved when pre-trained on in-domain data and fine-tuned using out-of-domain NSynth data~\cite{engel2017neural}.
@@
 Audio embeddings can also be used as a similarity metric.
-In a recent work, Cherep \textit{et al.} used latent representations from the CLAP model~\cite{wu2023large} along with a differentiable synthesizer~\cite{synthhaxcherep2023} to create creative interpretations of sound effects.
+For example, Cherep \textit{et al.} used latent representations from the CLAP model~\cite{wu2023large} along with a differentiable synthesizer~\cite{synthhaxcherep2023} to create creative interpretations of sound effects.
 In their approach, a desired sound-effect is described in text and embedded using CLAP, followed by iterative updates from a gradient-free optimizer~\cite{evosax2022github} to the synthesizer's parameters to minimize embedding differences.
-Based on manual hearing tests, this approach did not produce the ``correct'' sounds more frequently than previous works~\cite{kreuk2022audiogen}.
+Based on manual hearing tests, this approach did not produce the ``correct'' sounds more often than prior methods~\cite{kreuk2022audiogen}.
 However, it did yield better scores for ``artistic-interpretation'' (what we referred to as ``imitation'').
-It should be noted that audio embedding models are often trained using simpler loss functions---particularly those which utilize Fourier transforms---therefore improving such simpler loss functions is likely to improve the embedding models.
+It should be noted that audio embedding models are often trained using simpler (often Fourier-based) loss functions, therefore improvements to the fundamental loss functions may also lead to better embeddings.
 
